### 2018-05-21 papers/1_dmatpol/paper-r0.tex
Dmatpol paper versions archived

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

% hyphenation
\usepackage{hyphenat}
%\usepackage{lipsum}

% tables
%\usepackage{dcolumn}
%\usepackage{multirow}
%\usepackage{hhline}
%\usepackage{booktabs}
%\usepackage{longtable}

% mathematics
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{upgreek}
%\allowdisplaybreaks

% symbols
%\usepackage{gensymb}
%\usepackage{textcomp}

%---------------------------------------------------
% happy integral
\newcommand{\rint}[1]{\mbox{\Large $ \int\limits_{\mbox{\tiny  $#1$}}$}}
% SHORTCUTS
%\newcolumntype{,}{D{.}{,}{2}}
\newcommand{\citee}[1]{\ensuremath{\scriptsize^{\citenum{#1}}}}
\newcommand{\HRule}{\rule{\linewidth}{0.2mm}}
% Quantum notation
\newcommand{\bra}[1]{\ensuremath{\bigl\langle {#1} \bigl\lvert}}
\newcommand{\ket}[1]{\ensuremath{\bigr\rvert {#1} \bigr\rangle}}
\newcommand{\braket}[2]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigr\rangle}}
\newcommand{\tbraket}[3]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigl\lvert {#3} \bigr\rangle}}
% Math
\newcommand{\pd}{\ensuremath{\partial}}
\newcommand{\DR}{\ensuremath{{\rm d} {\bf r}}}
%\newcommand{\BM}[1]{\ensuremath{\mbox{\boldmath${#1}$}}}
\newcommand{\BM}[1]{\bm{#1}}
% Chemistry (formulas)
\newcommand{\ch}[2]{\ensuremath{\mathrm{#1}_{#2}}}
% Math 
\newcommand{\VEC}[1]{\ensuremath{\mathrm{\mathbf{#1}}}}
% vector nabla
\newcommand{\Nabla}{\ensuremath{ \BM{\nabla}}}
% derivative
\newcommand{\FDer}[3]{\ensuremath{
\bigg(
\frac{\partial #1}{\partial #2}
\bigg)_{#3}}}
% diagonal second derivative
\newcommand{\SDer}[3]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2^2}
\biggr)_{#3}}}
% off-diagonal second derivative
\newcommand{\SSDer}[4]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2 \partial #3}
\biggr)_{#4}}}
% derivatives without bound
% derivative
\newcommand{\fderiv}[2]{\ensuremath{
\frac{\partial #1}{\partial #2}}}
% diagonal second derivative
\newcommand{\sderiv}[2]{\ensuremath{
\frac{\partial^2 #1}{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivd}[3]{\ensuremath{
\frac{\partial^2 #1}{\partial #2 \partial #3}
}}
% derivatives for tables
\newcommand{\fderivm}[2]{\ensuremath{
{\partial #1}/{\partial #2}}}
% diagonal second derivative
\newcommand{\sderivm}[2]{\ensuremath{
{\partial^2 #1}/{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivdm}[3]{\ensuremath{
{\partial^2 #1}/{\partial #2 \partial #3}
}}
% ERIs and OEIs
\newcommand{\OEIc}[3]{\ensuremath{\left(#1 \lvert #2 \rvert #3 \right)}}
\newcommand{\ERIc}[4]{\ensuremath{\left(#1 #2 \vert #3 #4 \right)}}

% Partial density and potential
\newcommand{\PartPot}[4]{\ensuremath{\frac{#1 #2}{\lvert #3-#4 \rvert }}}

% trace operator
\DeclareMathOperator{\Tr}{Tr}

\draft % marks overfull lines with a black rule on the right

\begin{document}

\title{Density Matrix Susceptibility} %Title of paper

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}
\thanks{Bulak}
\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, Wroc{\l}aw University of Technology, Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}
To be filled ...
\end{abstract}

\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

\section{\label{s:1}Introduction}

\section{\label{s:2}Theory}

The total electronic dipole moment of a closed-shell molecule can be expressed as
%
\begin{equation}\label{e:mu-total}
 {\BM{\upmu}}({\bf r}_0) = 
    2 \Tr{
        \left[
             {\bf D} {\mathbb{M}({\bf r}_0)}
        \right]} \;,
\end{equation}
%
where ${\bf D}$ is a one-particle density matrix and ${\mathbb{M}({\bf r}_0)}$ is certain representation
of the dipole moment operator defined with respect to origin at ${\bf r}_0$,
%
\begin{equation}\label{e:m}
 \left[ {\mathbb{M}({\bf r}_0)} \right]_{\alpha\beta} = \tbraket{\alpha}{{\bf r} - {\bf r}_0}{\beta} \;.
\end{equation}
%
In the above equations and throughout this Work, the basis set functions are labeled with Greek symbols.
%Note that ${\mathbb{M}}$ can be considered as a vector of $n\times n$ matrices with $n$ denoting the
%number of basis set functions.

In the Hartree\hyp{}Fock theory, ${\bf D}$ is approximated as
%
\begin{equation}
 {\bf D} = {\bf C}{\bf C}^\dagger \;,
\end{equation}
%
where ${\bf C}$ describes the occupied molecular orbitals as linear combinations
of atomic orbitals (AO's), referred here to as the LCAO-MO matrix.
When ${\bf D}$ is expressed in orthogonal basis, the density matrix is idempotent, i.e.,
%
\begin{equation}
 {\bf D}^2 = {\bf D} \;.
\end{equation}
%

%
\subsection{Molecule in External Electric Field}

Let us consider a case in which a molecule experiences external and generally non\hyp{}uniform electric field,
${\bf F}({\bf r})$. 
Then, if unchanging basis such as AO's is used, Eq.~\eqref{e:mu-total} can be recast as
%
\begin{equation}
 {\BM{\upmu}}({\bf r}_0) = 
     2 \Tr{ 
         \left[ 
              \left\{ {\bf D}^{(0)} + \delta {\bf D}[{\bf F}({\bf r})] \right\}  {\mathbb{M}({\bf r}_0)}
         \right] } \;,
\end{equation}
%
in which ${\bf D}^{(0)}$ is the unperturbed density matrix of the system whereas
$\delta {\bf D}$ is the associated change due to the external electric field distribution. 
In fact, $\delta {\bf D}$ is a complicated
but unknown function of ${\bf F}({\bf r})$ and basis set. 
%Moreover, it certainly depends on the choice of
%molecular orbitals because the unitary transformation . 
On the other hand it seems intuitive that $\delta {\bf D}$ should 
be directly related to the distortion of the electron density distribution, which is also described
by the polarization\hyp{}induced multipole moments. Then, the optimal $\delta {\bf D}$ minimizes
the total energy of the molecule under a constraint of reproducing the polarization\hyp{}induced 
distortion of the electron density distribution (or equivalently, all the multipole moments). 
Unfortunately, such a mathematical problem is very complicated due to its high\hyp{}order tensor nature
with respect to the unknown $\delta {\bf D}$ and no approximate analytical solutions have been proposed
up to date. Therefore, in this paper we adopt a somewhat different
approach that will require \emph{a priori} knowledge from training set of known cases.

\subsection{Polarization\hyp{}Induced Dipole Moment}

The first non\hyp{}vanishing induced moment is given by
%
\begin{equation} \label{e:dmu-exact}
 \delta {\BM{\upmu}}({\bf r}_0) = 
     2 \Tr{ 
         \left[ 
              \delta {\bf D} {\mathbb{M}({\bf r}_0)}
         \right] } \;.
\end{equation}
%
The change in the molecular orbitals can be parameterized as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
by defining the idempotency\hyp{}preserving variation
%
\begin{equation}
 \delta {\bf C} = {\BM\Delta} {\bf C} \;,
\end{equation}
%
where $\BM\Delta$ is a square non\hyp{}singular and non\hyp{}symmetric matrix.
McWeeny has shown that
%
\begin{equation} \label{e:dmat-change.exact}
 \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right]
                                   \left[ {\bf 1} + {\bf v}^\dagger{\bf v} \right]^{-1}
                                   \left[ {\bf D}^{(0)} + {\bf v}^\dagger \right] \;,
\end{equation}
%
where the auxiliary vector ${\bf v}$ is defined as
%
\begin{equation}
 {\bf v} \equiv \left[ {\bf 1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)}  \;.
\end{equation}
%
Expanding Eq.~\eqref{e:dmat-change.exact} in a Taylor series around ${\bf v}={\bf 0}$ and
truncating it at linear terms with respect to ${\BM\Delta}$ gives
%
\begin{subequations} 
 \begin{align}
 \delta {\bf D} &\cong \left[ \bf{1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)} + 
                        {\bf D}^{(0)} {\BM\Delta}^\dagger \left[ \bf{1} - {\bf D}^{(0)} \right]  
 \label{e:dD-1} \\  &= 
  \delta {\bf C}  {\bf C}^\dagger + {\bf C}\delta {\bf C}^\dagger
           - {\bf D}^{(0)} \delta {\bf C}  {\bf C}^\dagger - {\bf C}\delta {\bf C}^\dagger {\bf D}^{(0)} 
 \label{e:dD-2}
 \end{align}
\end{subequations}
%
The latter equality is true only if the basis set functions are orthogonal, i.e., ${{\bf C}^\dagger}{\bf C}={\bf 1}$,
which will be the case throughout the entire work, unless stated differently.
Substituting the result from Eq.~\eqref{e:dD-1}
into Eq.~\eqref{e:dmu-exact} leads to
%%
%\begin{equation} \label{e:delta-Taylor}
% \delta {\bf D} = \left[ {\bf v} + {\bf v}^\dagger \right]
%                + \left[ {\bf v}{\bf v}^\dagger - {\bf v}^\dagger{\bf v}\right] + \ldots 
%\end{equation}
%
\begin{equation} \label{e:dmu-4-exact.linear-approximation}
 \frac{1}{2} 
 \delta {\BM{\upmu}}
  \cong
   \Tr{ 
    \left[ 
         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
    \right] }
%
  +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
    \right] }
%
  -\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
    \right] }
%
  -\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
    \right] } \;,
\end{equation}
%
It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
induced dipole moment is independent on the origin. Therefore,
in the above result and also in later course of the Paper the dipole origin has been omitted. 
%where the dipole origin was omitted for notational clarity.
If the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
%
\begin{subequations}
 \begin{align}
  \Tr{ 
    \left[ 
         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
    \right] }
%
  +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
    \right] }
  &=
2 \Tr{ 
    \left[ 
         \widetilde{\mathbb{M}} \cdot \delta {\bf C}
   \right] }  \;,
%
\\
%
  \Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
    \right] }
%
 +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
    \right] }
  &=
2 \Tr{ 
    \left[ 
         \widetilde{\mathbb{K}} \cdot \delta {\bf C}
   \right] } \;.
%
 \end{align}
\end{subequations}
%
Here we introduced the auxiliary tensors
%
\begin{subequations}
 \begin{align}
   \widetilde{\mathbb{M}}  &\equiv {\bf C}^\dagger {\mathbb{M}}     \;,           \\
   \widetilde{\mathbb{K}}  &\equiv {\bf C}^\dagger {\mathbb{M}} {\bf D}^{(0)} \;, \\
   \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;.
 \end{align}
\end{subequations}
%
to simplify the notation. Hence, we have
%
\begin{equation} \label{e:dmu-l-vector}
  \frac{1}{4} 
 \delta {\BM{\upmu}} %({\bf r}_0) 
   =
   \Tr{ 
    \left[ 
         \widetilde{\mathbb{L}} \cdot \delta {\bf C}
    \right] }
   %
   \equiv \sum_{i}^{\rm occ} \sum_\alpha {\bf l}_{i\alpha} \delta C_{\alpha i} \;,
\end{equation}
%
where the Cartesian vector
%
\begin{equation}\label{e:l-vector}
 {\bf l}_{i\alpha} \equiv \left[ \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \right]_{i\alpha} 
      = \left[  {\bf C}^\dagger \mathbb{M} \left[ \bf{1} - \bf{D}^{(0)} \right] \right]_{i\alpha}
\end{equation}
%
and $i$ runs over occupied unperturbed MO's.
Note that the intermediate quantity $\widetilde{\mathbb{L}}$ contains the projector onto
the unoccupied MO space at the absence of the external electric field. It means that the
perturbed solution is constructed from the unperturbed virtual orbitals as well.
Although Eq.~\eqref{e:dmu-l-vector}
is exact up to first order in ${\BM\Delta}$, it cannot be inverted at this moment to obtain $\delta {\bf C}$
due to the presence of trace operation.

\subsection{Distributed Multipole Approach}

The above problem can be simplified if the trace operation in Eq.~\eqref{e:dmu-l-vector} is broken down into
two separate summations and then compared with the MO\hyp{}distributed partitioning 
of the polarization\hyp{}induced dipole moment. Formally,
%
\begin{equation} \label{e:mu-ind-distributed-general}
 \delta {\BM{\upmu}} = \sum_{i'}^{\rm occ} \delta {\BM{\upmu}}_{i'}  \;.
\end{equation}
%
The molecular orbitals in Eq.~\eqref{e:mu-ind-distributed-general}
are generally different from molecular orbitals in Eq.~\eqref{e:dmu-l-vector} because
any unitary transformation of MO's preserves the total dipole moment of the system. In other words,
Eq.~\eqref{e:dmu-l-vector} requires the following generalization
%
\begin{equation} \label{e:dmu-l-vector-X}
  \frac{1}{4} 
 \delta {\BM{\upmu}} %({\bf r}_0) 
   =
   \Tr{ 
    \left[ {\bf X}^\dagger
         \widetilde{\mathbb{L}}  \delta {\bf C}
          {\bf X}
    \right] } \;,
\end{equation}
%
in which ${\bf X}$ is, until now, an arbitrary unitary transformation of unperturbed molecular orbitals
%
\begin{equation}
 C_{\alpha i'} = \sum_i  C_{\alpha i} X_{ii'} \;,
\end{equation}
%
and is, in general, dependent on the external electric field.
Unfortunately, such an approach is greatly limited because now the target quantity $\delta {\bf C}$
is involved in summations that are second order in ${\bf X}$. However,
one can consider an opposite but much simpler problem when we look for such unitary transformation
that will result in the MO\hyp{}distributed dipole moments equivalent to 
$\sum_\alpha {\bf l}_{i\alpha} \delta C_{\alpha i}$ for $i$ running over \emph{known} set of unperturbed orbitals.
It means that we require that
%
\begin{equation} \label{e:dmu-l-vector-mo-transform}
 \frac{1}{4} \delta {\BM{\upmu}}_{i'}
   \cong
   \sum_\alpha {\bf l}_{i\alpha} \delta C_{\alpha i} \;.
\end{equation}
%
%which is to be understood that ${\BM{\upmu}}^{(i)}$ refers to the part of the total induced dipole moment
%that is compatible with the contribution to $\delta {\bf C}$ constructed from the known set of MO's.
%(which is different from ${\BM{\upmu}}_i$ which 
$\delta {\BM{\upmu}}_{i'}$ can be found from the coupled\hyp{}perturbed Hartree\hyp{}Fock theory,
according to which the molecular polarizability is given as 
%
\begin{equation} \label{e:pol-cphf}
 {\BM{\alpha}} = \Tr{{\mathbb{A}}} = \sum_i^{\rm occ} {\BM{\alpha}}_i  \;,
\end{equation}
%
where
%
\begin{equation} \label{e:pol-cphf-A}
 \left[{\mathbb{A}}\right]_{ij} = 2 \sum_r^{\rm vir} {\mathbb X}_{ir} \otimes {\mathbb{M}}_{rj}
\end{equation}
%
and ${\BM{\alpha}}_i$ is the distributed polarizability associated with the $i$th occupied MO.
In Eq.~\eqref{e:pol-cphf} the trace is performed along the occupied MO axes of ${\mathbb{A}}$
whereas the tensor outer product ($\otimes$) in Eq.~\eqref{e:pol-cphf-A} refers to the Cartesian components of
the perturbation operator $\hat{x}$ and dipole moment operator, projected onto the 
occupied and virtual MO subspaces. Note also, that ${\mathbb{A}}$ is a square matrix with
dimension $(N_{\rm occ} \times N_{\rm occ})$ of $(3 \times 3)$ matrices.

Now, the induced dipole moment associated with the $i$th occupied
MO is given by
%
\begin{equation} \label{e:dmu-dpol}
 \delta {\BM{\upmu}}_i = {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) \;.
\end{equation}
%
In Eq.~\eqref{e:dmu-dpol}, ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge
${\bf r}_i = \tbraket{i}{\hat{\bf r}}{i}$.
It becomes apparent that
%
\begin{equation} \label{e:dmu-dpol-expl}
 \delta {\BM{\upmu}}_{i'} = \left[ {\bf X}^\dagger {\mathbb A} {\bf X} \right]_{i'i'} \cdot {\bf F}({\bf r}_{i'}) \;.
\end{equation}
%
We can now insert the above result into Eq.~\eqref{e:dmu-l-vector-mo-transform} and, after dropping off the primed indices
and recasting the right hand side in a matrix form,
we get
%
\begin{equation} \label{e:dmu-l-vector-mo-transform-X}
 \frac{1}{4} \sum_{jk}^{\rm occ} X_{ji} \left[ {\bf A}_{jk} \cdot {\bf F}({\bf r}_i) \right] X_{ki}
   =
   \delta {\bf c}_i^T \cdot {\bf L}_i \;.
\end{equation}
%
Note that now ${\bf L}_i$ is a matrix of size $(n \times 3)$ and $\delta{\bf c}_i$ is just the $i$th column of the
change in the LCAO-MO matrix due to the polarization process.
%
%The above equation can be re-written in a matrix form as
%
%\begin{equation} 
% \left[ {\mathbb L} \right]_{\alpha \zeta \beta} = L_{\beta\alpha}^{(\zeta)}
%\end{equation}
%%
%of size $m \times 3 \times m$.
%Note that, in this notation, ${\BM{\upmu}}_i^T$ and $\overline{\BM\Delta}_i^T$ are row vectors
%whereas $\overline{\BM\Delta}_i^p$ is a column vector. 
The solution for $\delta {\bf C}$, given that ${\bf X}$ is known, is
%
\begin{equation} \label{e:mu-ind.matrix.linear.solution}
  \delta {\bf c}_i^T = \frac{1}{4}
                    \sum_{jk}^{\rm occ} X_{ji} X_{ki}
                    \left[ {\bf A}_{jk} \cdot {\bf F}({\bf r}_i) \right] \cdot
                    \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
\end{equation}
%
where $\left[ {\bf L}_i  \right]^{-1}_{\rm Left}$ is a left inverse
of ${\bf L}_i$ matrix 
%
\begin{equation} 
      \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
       \left[ {\bf L}_i^T {\bf L}_i \right]^{-1} {\bf L}_i^T 
\end{equation}
%
and has size $(3\times n)$. Note that the \emph{right inverse} of ${\bf L}_i$
does not exist because ${\bf L}_i {\bf L}_i^T$ is singular.
%Clearly, in the above approximation, ${\bf X}$ depends on the choice
%of unperturbed orbitals.

\subsection{Density Matrix Change}

Now we can obtain the final expression for the density matrix change 
due to external electric field that reproduces the 
induced dipole moment of the molecule, provided ${\bf X}$ is known. 
The change in the LCAO\hyp{}MO coefficient is parametereized as
%
\begin{equation} \label{e:dC-dmatpol}
 \delta C_{\alpha i} = {\bf b}_{\alpha i}[{\bf X}] \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the susceptibility tensor is defined as
%
\begin{equation} \label{e:susceptibility-b}
  b_{\alpha i;w}[{\bf X}] = \frac{1}{4} \sum_{jk}^{\rm occ} \sum_u^{x,y,z} X_{ji} X_{ki} A_{jk;uw} 
   \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  
\end{equation}
%
for $w=x,y,z$. The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
the expression for $\delta {\bf D}$ from Eq.~\eqref{e:dD-2} to finally give
%
\begin{equation}\label{e:final-model}
 \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the \emph{density matrix polarization susceptibility tensor} is
%
\begin{equation}  \label{e:susceptibility-B}
 {\bf B}^{(i)}_{\alpha\beta} = \left\{
%
                               C_{\alpha i} {\bf b}_{\beta i} + C_{\beta i} {\bf b}_{\alpha i}
                               - \sum_\gamma 
                                 \left( 
               D_{\alpha\gamma} C_{\beta i} + D_{\beta\gamma} C_{\alpha i}
                                 \right)
                                           {\bf b}_{\gamma i}
%
\right\}  \;.
\end{equation}
%
Hence, the 3D vector ${\bf B}^{(i)}_{\alpha\beta}$ describes the polarization affinity
of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.

It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i)}$ except of ${\bf X}$
are unique properties of electronically unperturbed molecule. 
However, Eq.~\eqref{e:final-model} is an approximation because (i) it is based on the approximate
relation in Eq.~\eqref{e:dmu-l-vector-mo-transform} and (ii) only total dipole moment (but not quadrupole
and higher moments) is exactly
reproduced by the current model. Due to the above reasons, the existence of ${\bf X}$ that
reproduces the exact perturbed density matrix with vanishing error is not guaranteed. 
Moreover, as it was pointed out before, ${\bf X}$ almost certainly depends on the external electric field. 
Next section adresses the problem of finding the optimal ${\bf X}$ that is treated as an adjustable
matrix.

%unperturbed molecule. This makes them perfect candidates for general effective fragment parameters.

\subsection{Determination of Optimal Unitary Transformation}

In this Work the major task is then brought to examining if it is possible to find such ${\bf X}$ that optimizes
the perturbed density matrix with the minimal error. If it would be the case, the next important question
would be how much the effective unitary transformation depends on the external electric field.

We start with recasting Eq.~\eqref{e:final-model} in a form that is simpler for further analysis, mainly
%
\begin{equation}
 \delta D_{\alpha\beta}^{(N)} \approx
  \sum_{ijk}^{\rm occ} X_{ji} X_{ki} S_{ijk}^{\alpha\beta(N)}  \;,
\end{equation}
%
where 
%
\begin{equation}
  S_{ijk}^{\alpha\beta(N)} \equiv 
   \frac{1}{4} \sum_{uw}^{x,y,z} A_{jk;uw} 
   \left\{
     C_{\alpha i} G_{i\beta ;u} + C_{\beta i} G_{i\alpha ;u} 
   - \sum_\gamma \left(
        D_{\alpha\gamma} C_{\beta i} + D_{\beta\gamma} C_{\alpha i} 
    \right) G_{i\gamma ;u}
   \right\}
   F_w^{(N)}({\bf r}_i)  \;,
\end{equation}
%
$N$ denotes the sample taken for analysis and
%
\begin{equation}
 G_{i\alpha ;u} \equiv  \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  \;.
\end{equation}
%
%for notational clarity. 
Now, let us assume that we have gathered $N$ samples with $N$ known exact estimates
of density matrix (each for different distribution of electric field) denoted as $\delta D_{\alpha\beta}^{(N),{\rm ref}}$. 
We define the objective, positive definite
error function
%
\begin{equation} \label{e:Z-func}
 Z[{\bf X}] = \sum_N \sum_{\alpha\beta} 
    \left( 
      \delta D_{\alpha\beta}^{(N)} - \delta D_{\alpha\beta}^{(N),{\rm ref}}
    \right)^2
\end{equation}
%
that is subject to minimization under unitary constraints for $\bf X$.
Writing $Z$ function explicitly we can see that it is fourth\hyp{}order with respect to
the unitary transformation
%
\begin{equation} \label{e:Z-func-explicit}
 Z[{\bf X}] = \sum_{ijklmn} X_{ji} X_{ki} X_{ml} X_{nl} R_{ijklmn}
    + \sum_{ijk} X_{ji} X_{ki} P_{ijk} 
    + Z_0
\end{equation}
%
where 
%
\begin{subequations}
 \begin{align}
  R_{ijklmn} &=  \sum_N \sum_{\alpha\beta} S_{ijk}^{\alpha\beta(N)} S_{lmn}^{\alpha\beta(N)} \\
  P_{ijk}    &=-2\sum_N \sum_{\alpha\beta} \delta D_{\alpha\beta}^{(N),{\rm ref}} S_{ijk}^{\alpha\beta(N)} \\
  Z_0        &=  \sum_N \sum_{\alpha\beta} \left[ \delta D_{\alpha\beta}^{(N),{\rm ref}} \right]^2
 \end{align}
\end{subequations}
%
We developed the iterative algorithm to find the global optimum of $Z$ and the Reader 
is referred to the Appendix~\ref{a:X-mat} for more details.

\section{\label{s:3}Calculation Details}

The theory that is necessary to validate the density matrix susceptibility model from Section~\ref{s:2}
was implemented in our in\hyp{}house plugin to Psi4 quantum chemistry program.
A few simple molecules were selected as test systems, that were subject to 
perturbations due to uniform electric fields as well as a collection of point charges
generating a non\hyp{}uniform electric fields. For finding a general unitary transformations,
100 random point charge distributions were selected for which an exact density matrices
were calculated.

Throughout the Work, the Hartree\hyp{}Fock level was assumed and the aug-cc-pVTZ basis set
was used. For all the calculations of the density matrix polarization susceptibility tensors, 
atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme.

\section{\label{s:4}Results and Discussion}

%The critical assessment of the proposed model is discussed in two stages: (i) first
%the capability of the model in capturing the polarization 
It is instructive to analyze first the capability of the model in capturing the polarization
process by studying the 3D electronic density distortions due to the electric field perturbation.
For this purpose we consider a special case of the model in which ${\bf X}={\bf 1}$, i.e., without
optimization of the particular density matrix elements with respect to benchmark. In such a case,
density matrix polarization susceptibility tensors are implicitly independent on the electric field
distribution and can be computed for an isolated molecule.

For this purpose we studied few molecules
in a uniform external electric field. Although the density matrix elements in a chosen AO basis
can be inconsistent with other quantities (such as Fock matrix and so on) the polarization

\section{\label{s:5}Summary and a few concluding remarks}
Test summary bla bla blaaaaa dfn fdj gf ads;jha g;ja gf; p.

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

%
\appendix

\section{\label{a:orig-dep} Origin-Dependence of Induced Dipole Moment}

We need to be aware that multipole integrals depend on the origin with respect to which they are evaluated.
However, the induced dipole moments defined in Eq.~\eqref{e:mu-ind-distributed-general} 
have to be origin independent. 
%Let us check if this is the case.

Let us start with noticing that
%
\begin{equation}
 \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
\end{equation}
%
Further we have
%
\begin{equation}
 \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
\end{equation}
%
But ${\bf C}^\dagger {\bf D}={\bf C}^\dagger {\bf C} {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
AO basis. This implies that
%
\begin{equation}
   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0)
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
 = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
\end{equation}
%
Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
defined in Eq.~\eqref{e:mu-ind-distributed-general} 
are origin independent.
Thus, one can compute dipole integrals with respect to any origin and resulting
susceptibilities ${\bf B}_{\alpha\beta}^{(i)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.

\section{\label{a:X-mat}Algorithm for Optimal Unitary Transformation}

The error function from Eq.~\eqref{e:Z-func-explicit} is optimized by using the generalized Jacobi iteration algorithm
which is an extension of the Edmiston\hyp{}Reudenberg optimization scheme to tackle high\hyp{}order
complicated tensor equations.

Here we discuss a general algorithm that minimizes or maximizes the following function
%
\begin{equation} \label{e:a.Z}
 Z[{\bf X}] = \sum_{ijklmn} X_{ki} X_{lj} X_{mi} X_{nj} R_{ijklmn}
             +\sum_{ijk}    X_{ji} X_{ki} P_{ijk}
\end{equation}
%
under unitary constraint for ${\bf X}$. ${\bf P}^{(3)}$ and ${\bf R}^{(6)}$ are the real, arbitrary three\hyp{} and
six\hyp{}rank tensors of size $N^3$ and $N^6$, respectively. Note that $Z_0$ term 
in Eq.~\eqref{e:Z-func-explicit} is constant
and therefore we omit this term in the present section.

Optimization of $ {\bf X} $ is factorized into a sequence of 2-dimensional Jacobi rotations with one real parameter
%
\begin{equation}
 {\bf X}^{\rm New} = {\bf X}^{\rm Old} \cdot {\bf U}^{IJ}(\gamma)
\end{equation}
%
where 
%
\begin{equation} \label{e:a.U}
 U_{ij}^{IJ}(\gamma) = \delta_{ij}
                       \left( 1-\delta_{iI} \right)
                       \left( 1-\delta_{jJ} \right) 
                       +
                       \cos(\gamma) \delta_{ij} \left( \delta_{iI} + \delta_{jJ} \right)
                       + 
                       \sin(\gamma) \left( 1-\delta_{ij} \right) 
                       \left( \delta_{iI} \delta_{jJ} - \delta_{iJ} \delta_{jI}\right)
\end{equation}
%
is the Jacobi transformation matrix constructed for the appropriately chosen $ I$th and $ J$th element ($I\neq J$)
from the entire $ N$-dimensional set. Each step is performed by finding the optimal pair
$(I,J)$ (out of all $N(N-1)/2$ pairs) for which optimal $\gamma*$ gives the largest decrease (increase) of $Z$ 
in the case of minimization (maximization). 
For the sake of algirithmic simplicity, 
every iteration after $ {\bf U}^{IJ}(\gamma) $ has been formed, $ {\bf X}^{\rm Old} $ 
is for a while assumed to be an identity matrix and the $ {\bf R} $ as well as $ {\bf P} $ tensors 
are transformed according to the following formulae
%
\begin{subequations}\label{e:a.rp-transformations}
 \begin{align}
 R_{ijklmn} &\rightarrow \sum_{k'l'm'n'} R_{ijk'l'm'n'} X_{k'k} X_{l'l} X_{m'm} X_{n'n} \;, \\
 P_{ijk}    &\rightarrow \sum_{j'k'}     P_{ijk} X_{j'j} X_{k'k}  \;.
 \end{align}
\end{subequations}
%
The full transformation matrix is accumulated in the memory buffer until convergence.

To find the optimal $\gamma^*$ for a chosen $(I,J)$ pair
all the roots of the first derivative of $Z$ with respect to $\gamma$
need to be found first, i.e.,
%
\begin{equation} \label{e:a.Z-grad}
 \frac{\partial Z[{\bf U}^{IJ}(\gamma)]}{\partial \gamma} = 0 \;.
\end{equation}
%
From expression for ${\bf U}^{IJ}$ in Eq.~\eqref{e:a.U} it is clear that Eq.~\eqref{e:a.Z-grad}
adopts the truncated Fourier series
%
\begin{equation} \label{e:a.Z-grad-Fourier}
 \frac{\partial Z[{\bf U}^{IJ}(\gamma)]}{\partial \gamma} = 
  a_0 + \sum_{q=1}^4 \left\{ a_q\cos(q\gamma) + b_q\sin(q\gamma) \right\} = 0\;.
\end{equation}
%
In general, if $Z$ function subject to optimize contains $Y$ as maximal power of ${\bf X}$
then the Fourier series will have $2Y+1$ coefficients and maximally $2Y$ roots. Therefore,
in the case of $Z$ in Eq.~\eqref{e:a.Z} there will be eight roots, from which only one
is to be chosen.

Finding all the Fourier coefficients is straightforward though tedious if attempted analytically. 
In the Supporting Information the analytical formulae for the Fourier coefficients are given. 
Once all the Fourier coefficients are found, Eq.~\eqref{e:a.Z-grad-Fourier} can be solved
by utilizing the Boyd's method.\cite{Boyd.JEngMath.2006} 
In our case, the roots are given by
%
\begin{equation}
 \gamma_n = \Re\left[-i\ln{\varepsilon_n}\right] \;,
\end{equation}
%
where $\varepsilon_n$ are the complex eigenvalues of the Frobenius\hyp{}Jacobi matrix
%
\begin{equation}
\begin{pmatrix}
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0\\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0\\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0\\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0\\
-\frac{a_4+ib_4}{a_4-ib_4} &
-\frac{a_3+ib_3}{a_4-ib_4} &
-\frac{a_2+ib_2}{a_4-ib_4} &
-\frac{a_1+ib_1}{a_4-ib_4} & 
-\frac{2a_0}{a_4-ib_4} &
-\frac{a_1-ib_1}{a_4-ib_4} &
-\frac{a_2-ib_2}{a_4-ib_4} &
-\frac{a_3-ib_3}{a_4-ib_4} &
\end{pmatrix}  \;.
\end{equation}
%
The roots can also be determined numerically. 
It is anticipated here that the above described method is not limited to Eq.~\eqref{e:a.Z} but can be directly applied 
to design unitary optimizations for more complicated tensor equations. In fact, the whole problem
reduces to finding the roots of the Fourier series.
It is also worth mentioning that the orbital localization method of Edmiston and Reudenberg
is the special case of the above algorighm. In their case the most of the Fourier coefficients
vanish leaving only those for $\cos(4\gamma)$. The according $Z$ function reaches 
the same maxima (minima) for angles that are constant through the iterative process,
in contrast to the general case when they need to be updated every iteration,
after the transformations similar to those in Eqs.~\eqref{e:a.rp-transformations}.

\bibliography{references}

\end{document}

### 2018-05-21 papers/1_dmatpol/paper-r1.tex
Dmatpol paper versions archived

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

% hyphenation
\usepackage{hyphenat}
%\usepackage{lipsum}

% tables
%\usepackage{dcolumn}
%\usepackage{multirow}
%\usepackage{hhline}
%\usepackage{booktabs}
%\usepackage{longtable}

% mathematics
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{upgreek}
%\allowdisplaybreaks

% symbols
%\usepackage{gensymb}
%\usepackage{textcomp}

%---------------------------------------------------
% happy integral
\newcommand{\rint}[1]{\mbox{\Large $ \int\limits_{\mbox{\tiny  $#1$}}$}}
% SHORTCUTS
%\newcolumntype{,}{D{.}{,}{2}}
\newcommand{\citee}[1]{\ensuremath{\scriptsize^{\citenum{#1}}}}
\newcommand{\HRule}{\rule{\linewidth}{0.2mm}}
% Quantum notation
\newcommand{\bra}[1]{\ensuremath{\bigl\langle {#1} \bigl\lvert}}
\newcommand{\ket}[1]{\ensuremath{\bigr\rvert {#1} \bigr\rangle}}
\newcommand{\braket}[2]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigr\rangle}}
\newcommand{\tbraket}[3]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigl\lvert {#3} \bigr\rangle}}
% Math
\newcommand{\pd}{\ensuremath{\partial}}
\newcommand{\DR}{\ensuremath{{\rm d} {\bf r}}}
%\newcommand{\BM}[1]{\ensuremath{\mbox{\boldmath${#1}$}}}
\newcommand{\BM}[1]{\bm{#1}}
% Chemistry (formulas)
\newcommand{\ch}[2]{\ensuremath{\mathrm{#1}_{#2}}}
% Math 
\newcommand{\VEC}[1]{\ensuremath{\mathrm{\mathbf{#1}}}}
% vector nabla
\newcommand{\Nabla}{\ensuremath{ \BM{\nabla}}}
% derivative
\newcommand{\FDer}[3]{\ensuremath{
\bigg(
\frac{\partial #1}{\partial #2}
\bigg)_{#3}}}
% diagonal second derivative
\newcommand{\SDer}[3]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2^2}
\biggr)_{#3}}}
% off-diagonal second derivative
\newcommand{\SSDer}[4]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2 \partial #3}
\biggr)_{#4}}}
% derivatives without bound
% derivative
\newcommand{\fderiv}[2]{\ensuremath{
\frac{\partial #1}{\partial #2}}}
% diagonal second derivative
\newcommand{\sderiv}[2]{\ensuremath{
\frac{\partial^2 #1}{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivd}[3]{\ensuremath{
\frac{\partial^2 #1}{\partial #2 \partial #3}
}}
% derivatives for tables
\newcommand{\fderivm}[2]{\ensuremath{
{\partial #1}/{\partial #2}}}
% diagonal second derivative
\newcommand{\sderivm}[2]{\ensuremath{
{\partial^2 #1}/{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivdm}[3]{\ensuremath{
{\partial^2 #1}/{\partial #2 \partial #3}
}}
% ERIs and OEIs
\newcommand{\OEIc}[3]{\ensuremath{\left(#1 \lvert #2 \rvert #3 \right)}}
\newcommand{\ERIc}[4]{\ensuremath{\left(#1 #2 \vert #3 #4 \right)}}

% Partial density and potential
\newcommand{\PartPot}[4]{\ensuremath{\frac{#1 #2}{\lvert #3-#4 \rvert }}}

% trace operator
\DeclareMathOperator{\Tr}{Tr}

\draft % marks overfull lines with a black rule on the right

\begin{document}

\title{Density Matrix Susceptibility} %Title of paper

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}
\thanks{Bulak}
\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, Wroc{\l}aw University of Technology, Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}
To be filled ...
\end{abstract}

\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

\section{\label{s:1}Introduction}

\section{\label{s:2}Theory}

The electronic energy of a closed\hyp{}shell molecule $A$ is given by the time\hyp{}independent Schr{\"o}dinger equation,
%
\begin{equation}
 E^{(0)} = \tbraket{\Psi^{(0)}}{\mathscr{H}^{(0)}}{\Psi^{(0)}} \;,
\end{equation}
%
where $E$ stands for the total electronic energy, $\Psi$ the electronic wavefunction and $\mathscr{H}$
the Hamiltonian operator. In the above, the superscript $(0)$
denotes the reference, or unperturbed, state that is assumed to be known.

Under the influence of an external electrostatic perturbation, the total electronic energy
becomes
%
\begin{equation}
 E[{\bf F}({\bf r})] = \tbraket{\Psi}{\mathscr{H}^{(0)} + \mathscr{V}({\bf F}({\bf r}))}{\Psi}
\end{equation}
%
in which ${\bf F}({\bf r})$ is a non\hyp{}uniform electric field distribution.
The electric field\hyp{}induced energy change is an unknown functional 
of the associated change in the one\hyp{}electron density,
%
\begin{equation}\label{e:rho}
 \rho({\bf r}) = 2\sum_{\alpha\beta} D_{\alpha\beta} \phi_\alpha({\bf r}) \phi^*_\beta({\bf r}) \;.
\end{equation}
%
In Eq.~\eqref{e:rho} $\phi_\alpha({\bf r})$ is a basis funtion, which can be for instance 
an atomic orbital (AO) or a molecular orbital (MO), and $D_{\alpha\beta}$ is the corresponding
one\hyp{}particle density matrix.
If perturbation\hyp{}independent basis is chosen, the electric field\hyp{}induced change in the one\hyp{}particle density
is given by
%
\begin{equation}\label{e:rho}
 \delta \rho({\bf r}) = 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \phi_\alpha({\bf r}) \phi^*_\beta({\bf r}) \;.
\end{equation}
%
Ideally, the electric field\hyp{}induced change in the ground state's one\hyp{}particle density 
follows its energy minimum of the perturbed state. In addition, all the perturbed 
excited state one\hyp{}electron densities must be excited state solutions of the Schr{\"o}dinger equation
with the exact energy\hyp{}minimized ground state.

Development of a first\hyp{}principles theory
that is capable of accurately capturing the electric field\hyp{}dependence of the one\hyp{}particle density
is extremely difficult. This is not only because of the very complicated (and generally unknown) solution
to the electronic correlation problem, but also due to the dimensionality of the problem: the quantitative
model has to accurately predict all the $n^2$ elements of the density matrix. Even relatively small errors
in a particular matrix element could accumulate tremendously when computing traces with other matrices,
and in turn, obstruct the evaluated properties with non\hyp{}negligible errors.
Therefore, in this Work we adopt the simplistic approach: we first develop a qualitative first\hyp{}principles model 
at the Hartree\hyp{}Fock (HF)
level of theory, and then generalize it to the arbitrary level using the functional form that was found 
by considering HF level,
but treating the model as a set of adjustable parameters further on.

\subsection{Ab Initio Density Matrix Susceptibilities: Hartree-Fock Level}

\subsubsection{Induced Dipole Moment}

It seems intuitive that $\delta {\bf D}$ should 
be directly related to the
polarization\hyp{}induced multipole moments because they are
a measure of the distortion of the electron density distribution due to the external electric
field. 
%Unfortunately, such a mathematical problem is very complicated due to its high\hyp{}order tensor nature
%with respect to the unknown $\delta {\bf D}$ and no approximate analytical solutions have been proposed
%up to date. Therefore, in this paper we adopt a somewhat different
%approach that will require \emph{a priori} knowledge from training set of known cases.
Therefore, 
we start our analysis by considering the induced electronic dipole moment 
of a closed-shell molecule, %which is the first non\hyp{}vanishing moment,
%
\begin{equation} \label{e:dmu-exact}
 \delta {\BM{\upmu}}({\bf r}_Q) = 
     2 \Tr{ 
         \left[ 
              \delta {\bf D} {\mathbb{M}({\bf r}_Q)}
         \right] } \;.
\end{equation}
%
where ${\mathbb{M}({\bf r}_Q)}$ is AO\hyp{}representation
of the dipole moment operator defined with respect to origin at ${\bf r}_Q$,
%
\begin{equation}\label{e:m}
 \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = \tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
\end{equation}
%
Note that if the total charge is conserved, the induced moments are independent on the origin.

In the Hartree\hyp{}Fock theory, ${\bf D}$ is approximated as
%
\begin{equation}
 {\bf D} = {\bf C}{\bf C}^\dagger \;,
\end{equation}
%
where ${\bf C}$ describes the occupied molecular orbitals as linear combinations
of atomic orbitals (AO's), referred here to as the LCAO-MO matrix.
When ${\bf D}$ is expressed in orthogonal basis, the density matrix is idempotent, i.e.,
%
\begin{equation}
 {\bf D}^2 = {\bf D} \;.
\end{equation}
%
The change in the molecular orbitals can be parameterized as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
by defining the idempotency\hyp{}preserving variation
%
\begin{equation}
 \delta {\bf C} = {\BM\Delta} \cdot {\bf C}^{(0)} \;,
\end{equation}
%
where $\BM\Delta$ is a square non\hyp{}singular and non\hyp{}symmetric matrix.
McWeeny has shown that
%
\begin{equation} \label{e:dmat-change.exact}
 \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right]
                                   \left[ {\bf 1} + {\bf v}^\dagger{\bf v} \right]^{-1}
                                   \left[ {\bf D}^{(0)} + {\bf v}^\dagger \right] \;,
\end{equation}
%
where the auxiliary vector ${\bf v}$ is defined as
%
\begin{equation}
 {\bf v} \equiv \left[ {\bf 1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)}  \;.
\end{equation}
%
Expanding Eq.~\eqref{e:dmat-change.exact} in a Taylor series around ${\bf v}={\bf 0}$ and
truncating it at linear terms with respect to ${\BM\Delta}$ gives
%
\begin{subequations} 
 \begin{align}
 \delta {\bf D} &\cong \left[ {\bf 1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)} + 
                        {\bf D}^{(0)} {\BM\Delta}^\dagger \left[ {\bf 1} - {\bf D}^{(0)} \right]  
 \label{e:dD-1} \\  &= 
  \delta {\bf C}  {\bf C}^{(0)\dagger} + {\bf C}^{(0)}\delta {\bf C}^\dagger
           - {\bf D}^{(0)} \delta {\bf C}  {\bf C}^{{(0)}\dagger} - {\bf C}^{(0)}\delta {\bf C}^\dagger {\bf D}^{(0)} 
 \label{e:dD-2}
 \end{align}
\end{subequations}
%
The latter equality is true only if the basis set functions are orthogonal, i.e., 
${\bf C}^{(0)\dagger}{\bf C}^{(0)}={\bf 1}$,
which will be the case throughout the entire work, unless stated differently.
Substituting the result from Eq.~\eqref{e:dD-1}
into Eq.~\eqref{e:dmu-exact} leads to
%%
%\begin{equation} \label{e:delta-Taylor}
% \delta {\bf D} = \left[ {\bf v} + {\bf v}^\dagger \right]
%                + \left[ {\bf v}{\bf v}^\dagger - {\bf v}^\dagger{\bf v}\right] + \ldots 
%\end{equation}
%
\begin{equation} \label{e:dmu-4-exact.linear-approximation}
 \frac{1}{2} 
 \delta {\BM{\upmu}}
  \cong
   \Tr{ 
    \left[ 
         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
    \right] }
%
  +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
    \right] }
%
  -\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
    \right] }
%
  -\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
    \right] } \;,
\end{equation}
%
It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
induced dipole moment is independent on the origin. Therefore,
in the above result and also in later course of the Paper the dipole origin has been omitted. 
%where the dipole origin was omitted for notational clarity.
If the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
%
\begin{subequations}
 \begin{align}
  \Tr{ 
    \left[ 
         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
    \right] }
%
  +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
    \right] }
  &=
2 \Tr{ 
    \left[ 
         \widetilde{\mathbb{M}} \cdot \delta {\bf C}
   \right] }  \;,
%
\\
%
  \Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
    \right] }
%
 +\Tr{ 
    \left[ 
         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
    \right] }
  &=
2 \Tr{ 
    \left[ 
         \widetilde{\mathbb{K}} \cdot \delta {\bf C}
   \right] } \;.
%
 \end{align}
\end{subequations}
%
Here we introduced the auxiliary tensors
%
\begin{subequations}
 \begin{align}
   \widetilde{\mathbb{M}}  &\equiv {\bf C}^\dagger {\mathbb{M}}     \;,           \\
   \widetilde{\mathbb{K}}  &\equiv {\bf C}^\dagger {\mathbb{M}} {\bf D}^{(0)} \;, \\
   \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;.
 \end{align}
\end{subequations}
%
to simplify the notation. Hence, we have
%
\begin{equation} \label{e:dmu-l-vector}
  \frac{1}{4} 
 \delta {\BM{\upmu}} %({\bf r}_0) 
   =
   \Tr{ 
    \left[ 
         \widetilde{\mathbb{L}} \cdot \delta {\bf C}
    \right] }
   %
   \equiv \sum_{i}^{\rm occ} \sum_\alpha {\bf l}_{i\alpha} \delta C_{\alpha i} \;,
\end{equation}
%
where the Cartesian vector
%
\begin{equation}\label{e:l-vector}
 {\bf l}_{i\alpha} \equiv \left[ \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \right]_{i\alpha} 
      = \left[  {\bf C}^\dagger \mathbb{M} \left[ \bf{1} - \bf{D}^{(0)} \right] \right]_{i\alpha}
\end{equation}
%
and $i$ runs over occupied unperturbed MO's.
Note that the intermediate quantity $\widetilde{\mathbb{L}}$ contains the projector onto
the unoccupied MO space at the absence of the external electric field. It means that the
perturbed solution is constructed from the unperturbed virtual orbitals as well.
Although Eq.~\eqref{e:dmu-l-vector}
is exact up to first order in ${\BM\Delta}$, it cannot be inverted at this moment to obtain $\delta {\bf C}$
due to the presence of trace operation, unless the electric field is uniform.
%in which ${\bf D}^{(0)}$ is the unperturbed density matrix of the system whereas
%$\delta {\bf D}$ is the associated change due to the external electric field distribution. 
%In fact, $\delta {\bf D}$ is a complicated
%but unknown function of ${\bf F}({\bf r})$ and basis set. 
%Moreover, it certainly depends on the choice of
%molecular orbitals because the unitary transformation . 
Let us partition the induced dipole moment into separate contributions
from the occupied MO's, i.e.,
%
\begin{equation} \label{e:dmu-part}
 \delta {\BM{\upmu}} = \sum_i^{\rm occ} \delta {\BM{\upmu}}_i
\end{equation}
%
where
%
\begin{equation} \label{e:dmu-dpol}
 \delta {\BM{\upmu}}_i = {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) \;.
\end{equation}
%
In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed polarizability tensor
associated with the $i$th MO whereas ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge
${\bf r}_i = \tbraket{i}{\hat{\bf r}}{i}$.
We can now insert the above result into Eq.~\eqref{e:dmu-l-vector} 
and after recasting the right hand side in a matrix form
we get
%
\begin{equation} \label{e:dmu-l-vector-mo-transform-X}
 \frac{1}{4} {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) 
   =
   \delta {\bf c}_i^T \cdot {\bf L}_i \;.
\end{equation}
%
Note that now ${\bf L}_i$ is a matrix of size $(n \times 3)$ and $\delta{\bf c}_i$ is just the $i$th column of the
change in the LCAO-MO matrix due to the polarization process.
%
%The above equation can be re-written in a matrix form as
%
%\begin{equation} 
% \left[ {\mathbb L} \right]_{\alpha \zeta \beta} = L_{\beta\alpha}^{(\zeta)}
%\end{equation}
%%
%of size $m \times 3 \times m$.
%Note that, in this notation, ${\BM{\upmu}}_i^T$ and $\overline{\BM\Delta}_i^T$ are row vectors
%whereas $\overline{\BM\Delta}_i^p$ is a column vector. 
The solution for $\delta {\bf C}$ is
%
\begin{equation} \label{e:mu-ind.matrix.linear.solution}
  \delta {\bf c}_i^T = \frac{1}{4}
            {\bf F}({\bf r}_i)  \cdot {\BM{\alpha}}_i \cdot 
                    \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
\end{equation}
%
where $\left[ {\bf L}_i  \right]^{-1}_{\rm Left}$ is a left inverse
of ${\bf L}_i$ matrix 
%
\begin{equation} 
      \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
       \left[ {\bf L}_i^T {\bf L}_i \right]^{-1} {\bf L}_i^T 
\end{equation}
%
and has size $(3\times n)$. Note that the \emph{right inverse} of ${\bf L}_i$
does not exist because ${\bf L}_i {\bf L}_i^T$ is singular.
%Clearly, in the above approximation, ${\bf X}$ depends on the choice
%of unperturbed orbitals.

\subsubsection{Density Matrix Change}

Now we can obtain the final expression for the density matrix change 
due to external electric field that reproduces the 
induced dipole moment of the molecule.
The change in the LCAO\hyp{}MO coefficient is parametereized as
%
\begin{equation} \label{e:dC-dmatpol}
 \delta C_{\alpha i} = {\bf b}_{\alpha i} \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the susceptibility tensor is defined as
%
\begin{equation} \label{e:susceptibility-b}
  b_{\alpha i;w} = \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
   \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  
\end{equation}
%
for $w=x,y,z$. The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
the expression for $\delta {\bf D}$ from Eq.~\eqref{e:dD-2} to finally give
%
\begin{equation}\label{e:final-model.HF}
 \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the \emph{density matrix polarization susceptibility tensor} is
%
\begin{equation}  \label{e:susceptibility-B}
 {\bf B}^{(i)}_{\alpha\beta} = \left\{
%
                               C_{\alpha i} {\bf b}_{\beta i} + C_{\beta i} {\bf b}_{\alpha i}
                               - \sum_\gamma 
                                 \left( 
               D_{\alpha\gamma} C_{\beta i} + D_{\beta\gamma} C_{\alpha i}
                                 \right)
                                           {\bf b}_{\gamma i}
%
\right\}  \;.
\end{equation}
%
Hence, the 3D vector ${\bf B}^{(i)}_{\alpha\beta}$ describes the polarization affinity
of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i)}$
are unique properties of electronically unperturbed molecule.

One very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
the density matrix increment dependent {\emph only} by susceptibility element associated with the same
pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
vector susceptibility, different for each density matrix element.

\subsection{Generalized Density Matrix Susceptibility Model}

\subsubsection{Functional form of the model}

Eq.~\eqref{e:final-model.HF}, which is the final result of the previous section, 
can be used as a guideline to construct a general model of density matrix polarization.
Quadratic terms with respect to the electric field as well as linear terms with respect to
electric field gradient can be added to the formulation, which leads to
%
\begin{equation}\label{e:final-model.General}
 \delta D_{\alpha\beta} = \sum_{i }^M \left\{
                                      {\bf B}^{(i ; 10)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  
%                       + \sum_{ij}^M {\bf B}^{(ij; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
                        +             {\bf B}^{(i ; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
                        +             {\bf B}^{(i ; 01)}_{\alpha\beta} : {\BM\nabla} \otimes {\bf F}({\bf r}_i) 
                        + \ldots \right\}
\end{equation}
%
%However, Eq.~\eqref{e:final-model} is an approximation because (i) it is based on the approximate
%relation in Eq.~\eqref{e:dmu-l-vector-mo-transform} and (ii) only total dipole moment (but not quadrupole
%and higher moments) is exactly
%reproduced by the current model. Due to the above reasons, the existence of ${\bf X}$ that
%reproduces the exact perturbed density matrix with vanishing error is not guaranteed. 
%Moreover, as it was pointed out before, ${\bf X}$ almost certainly depends on the external electric field. 
%Next section adresses the problem of finding the optimal ${\bf X}$ that is treated as an adjustable
%matrix.
In the above equation, the density matrix susceptibilities
${\bf B}^{(i; RP)}_{\alpha\beta}$, distributed over $M$ sites,
denote the $R$-th order response with respect to the electric field
and $P$-th order response with respect to the electric field gradient. Note that
${\bf B}^{(i; 10)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 20)}_{\alpha\beta}$
and ${\bf B}^{(i; 01)}_{\alpha\beta}$ are $(3\times 3)$ matrices. Also, the distributed expansion centres
(labeled as $i$) are generalized to any point in the molecule.

\subsubsection{Determining the generalized susceptibility tensors}

The exact mathematical form of the generalized density matrix susceptibilities is unknown. 
The first suscebility from Eq.~\eqref{e:final-model.HF} is valid only under the Hartree\hyp{}Fock
approximation, with a guarantee that only the overall charge and the dipole moment, 
but not higher multipole moments, are correctly described.
Our strategy is to treat the susceptibilities from Eq.~\eqref{e:final-model.General}
as adjustable parameters, rather than try to derive them from first principles theory.

Let  
$\left\{ {\bf F}^{(1)}({\bf r}), {\bf F}^{(2)}({\bf r}), \ldots, {\bf F}^{(N)}({\bf r}), \ldots \right\}$ 
be a set of $N_{\rm max}$ distinct and randomly sampled 
spatially non\hyp{}uniform distributions of the electric field. It is assumed that
the exact difference one\hyp{}particle density matrices (with respect to the unperturbed state)
defined as
%
\begin{equation}
 \delta \overline{\bf D}^{(N)} \equiv \overline{\bf D}^{(N)} - \overline{\bf D}^{(0)}
\end{equation}
%
are known for each sample (overline symbolizes the exact estimate).
%Let us assume that we a $\delta \overline{\bf D}^{(N)}$ 
%is the $N$th known (exact) difference one\hyp{}particle density matrix
%that was computed for the electric field distribution ${\bf F}^{(N)}({\bf r})$ chosen randomly.
Now, for each pair of the AO indices we construct the following 
parameterization of Eq.~\eqref{e:final-model.General}:
%
\begin{equation}\label{e:final-model.General.Parameters}
 \delta D^{(N)} = \sum_{i }^M \left\{ 
                              \sum_u^{x,y,z} S^{[1]}_{iu} F_{iu}^{(N)} 
%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{jw} \right) F_{iu}^{(N)} F_{ju}^{(N)}
                +             \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{iw} \right) F_{iu}^{(N)} F_{iu}^{(N)}
                +             \sum_{uw}^{x,y,z} \left( S^{[3]}_{iu} + S^{[3]}_{iw}\right) F_{iu,iw}^{(N)}
                        + \ldots \right\}
\end{equation}
%
(the Greek subscripts were omitted here for notational simplicity).
The multiple parameter blocks 
$S^{[1]}$, $S^{[2]}$ and $S^{[3]}$ that are associated with electric field and its gradients
appear in the first power, allowing for linear least\hyp{}squares regression.
Note that $B_u^{(i,(10))} = S^{[1]}_{iu}$, $B_{uw}^{(i,(20))} = S^{[2]}_{iu} + S^{[2]}_{iw}$
and $B_{uw}^{(i,(01))} = S^{[3]}_{iu} + S^{[3]}_{iw}$. The square bracket superscripts 
denote the block of the parameter space. We chose symmetric forms of ${\bf B}^{(i;20)}$
and ${\bf B}^{(i;01)}$ (antisymmetric form could be constructed as well by taking the difference
of the parameters from block 2 and 3, respectively). $F_{iu}^{(N)}$ and $F_{iu,iw}^{(N)}$ 
stand for the electric field $\left[ {\bf F}({\bf r}_i)\right]_u$ and the electric field
gradient $\left[ {\BM \nabla} \otimes {\bf F}({\bf r}_i) \right]_{uw}$, respectively.

To determine the optimim set 
$
 {\bf S} = 
\begin{pmatrix}
{\bf S}^{[1]} &
{\bf S}^{[2]} &
{\bf S}^{[3]}
\end{pmatrix}^T
$, we define the loss function $Z$ 
of the three blocks of adjustable parameters
that is subject to the least\hyp{}squares minimization:
%
\begin{equation}\label{e:Z}
 Z[{\bf S}] = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
\end{equation}
%
%where $N_{\rm max}$ is the total number of samples. 
%The linear form of Eq.~\eqref{e:final-model.General.Parameters} ensures that the Hessian 
%of $Z$
%is non\hyp{}singular and independent on the parameters. 
The Hessian of $Z$ computed with respect to the parameters is non\hyp{}singular 
and constant.
Therefore, the exact solution for the optimal parameters is given by a Newton equation
%
\begin{equation}\label{e:Newton}
 {\bf S} = -{\bf H}^{-1} \cdot {\bf g} \;,
\end{equation}
%
where ${\bf g}$ and ${\bf H}$ are the gradient vector and Hessian matrix, respectively.
%
%\begin{equation}\label{e:Newton.S}
% {\bf S} = 
%\begin{pmatrix}
%{\bf S}^{[1]} &
%{\bf S}^{[2]} &
%{\bf S}^{[3]}
%\end{pmatrix}^T
%\end{equation}
%
The explicit forms of the gradient and Hessian are given in the Appendix~\ref{a:blocks}.
Note that the dimensions of parameter space for the block 1, 2 and 3 are all
equal to $3M$.
%unperturbed molecule. This makes them perfect candidates for general effective fragment parameters.

\section{\label{s:3}Calculation Details}

The theory that is necessary to validate the density matrix susceptibility model from Section~\ref{s:2}
was implemented in our in\hyp{}house plugin to Psi4 quantum chemistry program.
A few simple molecules were selected as test systems, that were subject to 
perturbations due to uniform electric fields as well as collections of point charges
generating non\hyp{}uniform electric fields. 
%For finding a general unitary transformations,
%100 random point charge distributions were selected for which an exact density matrices
%were calculated.

In order to validate all the models discussed in this Contribution,
the electric field\hyp{}induced interaction energy, 
the dipole moment and the quadrupole moment were analysed in detail.
The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
was computed from the following exact formula:
%
\begin{multline}\label{e:dE}
 \delta E[{\bf F}] = 2\Tr{\left\{ {\bf D}^{(0)} \cdot {\bf V}[{\bf F}] \right\}}
                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
                                \left( \delta {\bf J} + \delta {\bf K}\right) \right\}} \\
                   + \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + 2 {\bf V}[{\bf F}] + {\bf f}^{(0)} \right) \right\}}
                   + E_{\rm nuc}({\bf F}) \;.
\end{multline}
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}[{\bf F}]$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
due to the nuclei.
The electric field\hyp{}induced dipole and quadrupole moments were evaluated from the following expressions:
%
\begin{subequations}\label{e:dmult}
  \begin{align}
   \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{uw}{\beta} \;. 
  \end{align}
\end{subequations}
%
To compute the interaction energy and the induced multipole moments, either exact or approximated forms
of the difference density matrix from Eq.~\eqref{e:final-model.General}
were used for a particular electric field distribution. The statistical evaluation of
errors with respect to the benchmark (full HF calculations) were based on 
the following statistical quality descriptors:
%
\begin{enumerate}
 \item Root Mean Square Energy Error (RMSE)
   \begin{equation}
     {\rm RMSE} = \sum_N \sqrt{ \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
   \end{equation}
 \item Root Mean Square Dipole Error (RMSD)
   \begin{equation}
     {\rm RMSD} = \sum_N \sqrt{ \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
   \end{equation}
 \item Root Mean Square Quadrupole Error (RMSQ)
   \begin{equation}
     {\rm RMSQ} = \sum_N \sqrt{ \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
   \end{equation}
\end{enumerate}
%
The first describes the ability of the model to reproduce the exact energetics of the system, 
wherease the latter two contain crucial information on how well the deformation of the one\hyp{}electron
density is described.

Throughout the Work, the Hartree\hyp{}Fock level was assumed and the aug-cc-pVTZ basis set
was used. For all the calculations of the density matrix polarization susceptibility tensors
from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme.

\section{\label{s:4}Results and Discussion}

\subsection{Ab Initio Model of Densty Matrix Polarization}

We first test the accuracy of the density matrix susceptibility tensors that were derived
in Eq.~\eqref{e:susceptibility-B} from first principles. 
For this purpose we have considered a water molecule 
under weak uniform electric fields (Figure~XXX). 
3D difference one\hyp{}electron density isosurface plots for sample electric fields
with respect to the unperturbed density (${\bf F} = {\bf 0}$)
are shown in Figure~XXX. The statistical evaluation of errors for a set of
100 random electric fields is shown in Figure~XXX. The shape of the basins 
of decreased/increased density are comparable with the exact result
obtained from HF/6-311++G** calculations. 
%It can be seen that the theory
%describes the polarization of the electron density distribution qualitatively well:
However, the accuracy in the density matrix elements is on overall very low which 
manifests with drammatic errors in the electronic total energy of the system.
It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
which neglects quadrupole and higher\hyp{}order induced multipole moments.

\subsection{General Model of Densty Matrix Polarization}

%However, Eq.~\eqref{e:final-model} is an approximation because (i) it is based on the approximate
%relation in Eq.~\eqref{e:dmu-l-vector-mo-transform} and (ii) only total dipole moment (but not quadrupole
%and higher moments) is exactly
%reproduced by the current model. Due to the above reasons, the existence of ${\bf X}$ that

%The critical assessment of the proposed model is discussed in two stages: (i) first
%the capability of the model in capturing the polarization 
%It is instructive to analyze first the capability of the model in capturing the polarization
%process by studying the 3D electronic density distortions due to the electric field perturbation.
%For this purpose we consider a , i.e., without
%optimization of the particular density matrix elements with respect to benchmark.

%For this purpose we studied few molecules
%in a uniform external electric field. Although the density matrix elements in a chosen AO basis
%can be inconsistent with other quantities (such as Fock matrix and so on) the polarization

\section{\label{s:5}Summary and a few concluding remarks}
Test summary bla bla blaaaaa dfn fdj gf ads;jha g;ja gf; p.

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

%
\appendix

\section{\label{a:orig-dep} Origin-Dependence of Induced Dipole Moment}

We need to be aware that multipole integrals depend on the origin with respect to which they are evaluated.
However, the induced dipole moments defined in Eq.~\eqref{e:mu-ind-distributed-general} 
have to be origin independent. 
%Let us check if this is the case.

Let us start with noticing that
%
\begin{equation}
 \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
\end{equation}
%
Further we have
%
\begin{equation}
 \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
\end{equation}
%
But ${\bf C}^\dagger {\bf D}={\bf C}^\dagger {\bf C} {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
AO basis. This implies that
%
\begin{equation}
   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0)
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
 = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
\end{equation}
%
Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
defined in Eq.~\eqref{e:mu-ind-distributed-general} 
are origin independent.
Thus, one can compute dipole integrals with respect to any origin and resulting
susceptibilities ${\bf B}_{\alpha\beta}^{(i)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.

\section{\label{a:blocks} Explicit Formulae for Gradient and Hessian Blocks}

The gradient vector ${\bf g}$ and Hessian matrix ${\bf H}$ 
are build from blocks associated with the type of parameters, i.e.,
%
\begin{equation}\label{e:Newton.gH}
 {\bf g} = 
\begin{pmatrix}
{\bf g}^{[1]} \\ 
{\bf g}^{[2]} \\ 
{\bf g}^{[3]}
\end{pmatrix} ,\quad
 {\bf H} = 
\begin{pmatrix}
{\bf H}^{[11]} & {\bf H}^{[12]} & {\bf H}^{[13]} \\ 
{\bf H}^{[21]} & {\bf H}^{[22]} & {\bf H}^{[23]} \\ 
{\bf H}^{[31]} & {\bf H}^{[32]} & {\bf H}^{[33]} 
\end{pmatrix} \;.
\end{equation}
%
The gradient element of the $r$th block and Hessian element of the $rs$th block read
%
\begin{subequations}
 \begin{align}
  g^{[r ]}_{kv}    &\equiv \frac{\partial   Z}{\partial S_{kv}^{[r]}} 
     =-2\sum_N \overline{\delta D}^{(N)}
               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}} \;,\\
  H^{[rs]}_{kv,lw} &\equiv \frac{\partial^2 Z}{\partial S_{kv}^{[r]} \partial S_{lw}^{[s]}}  
     = 2\sum_N 
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}}
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{lw}^{[s]}} \;.
 \end{align}
\end{subequations}
%
Note that the second derivatives of $\delta D^{(N)}$ vanish
due to the linearity with respect to the adjustable parameters.
Thus, the explicit formulae for the gradient are
%
\begin{subequations}
 \begin{align}
  g^{[1]}_{kv} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \;,\\
  g^{[2]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \widetilde{F}^{(N)}_k \;,\\
  g^{[3]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} \widetilde{F}^{(N)}_{kv} \;.
 \end{align}
\end{subequations}
%
The Hessian subsequently follows to be
%
\begin{subequations}
 \begin{align}
  H^{[11]}_{kv,lw} &= 2\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \;,\\
  H^{[22]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_l \;,\\
  H^{[33]}_{kv,lw} &= 8\sum_N \widetilde{F}^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
  H^{[12]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_l \;,\\
  H^{[13]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
  H^{[23]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_{lw} \;.
 \end{align}
\end{subequations}
%
In the above equations, the auxiliary quantities $\widetilde{F}^{(N)}_k$ 
and $\widetilde{F}^{(N)}_{kv}$ are defined as
%
\begin{subequations}
 \begin{align}
  \widetilde{F}^{(N)}_k    &\equiv \sum_w F^{(N)}_{kw} \;,\\
  \widetilde{F}^{(N)}_{kv} &\equiv \sum_w F^{(N)}_{kw,kv} \;.
 \end{align}
\end{subequations}
%
Note that due to the symmetry of the Hessian matrix, the blocks $21$, $31$ and $32$
are transposes of the blocks $12$, $13$ and $23$, respectively. The composite index
can be defined as $kv \equiv 3k+v$, for example.
\bibliography{references}

\end{document}

### 2018-05-21 papers/templates/aip/aipsamp.tex
Templates folder for papers

% ****** Start of file aipsamp.tex ******
%
%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
%   Version 4.1 of REVTeX, October 2009
%
%   Copyright (c) 2009 American Institute of Physics.
%
%   See the AIP README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex  aipsamp
%  2)  bibtex aipsamp
%  3)  latex  aipsamp
%  4)  latex  aipsamp
%
% Use this file as a source of example code for your aip document.
% Use the file aiptemplate.tex as a template for your document.
\documentclass[%
 aip,
%jmp,%
%bmf,%
 sd,%
%rsi,%
 amsmath,amssymb,
%preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

\begin{document}

\preprint{AIP/123-QED}

\title[Sample title]{Sample Title:\\with Forced Linebreak\footnote{Error!}}% Force line breaks with \\
\thanks{Footnote to title of article.}

\author{A. Author}
 \altaffiliation[Also at ]{Physics Department, XYZ University.}%Lines break automatically or can be forced with \\
\author{B. Author}%
 \email{Second.Author@institution.edu.}
\affiliation{ 
Authors' institution and/or address%\\This line break forced with \textbackslash\textbackslash
}%

\author{C. Author}
 \homepage{http://www.Second.institution.edu/~Charlie.Author.}
\affiliation{%
Second institution and/or address%\\This line break forced% with \\
}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
An article usually includes an abstract, a concise summary of the work
covered at length in the main body of the article. It is used for
secondary publications and for information retrieval purposes. 
%
Valid PACS numbers may be entered using the \verb+\pacs{#1}+ command.
\end{abstract}

\pacs{Valid PACS appear here}% PACS, the Physics and Astronomy
                             % Classification Scheme.
\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

\begin{quotation}
The ``lead paragraph'' is encapsulated with the \LaTeX\ 
\verb+quotation+ environment and is formatted as a single paragraph before the first section heading. 
(The \verb+quotation+ environment reverts to its usual meaning after the first sectioning command.) 
Note that numbered references are allowed in the lead paragraph.
%
The lead paragraph will only be found in an article being prepared for the journal \textit{Chaos}.
\end{quotation}

\section{\label{sec:level1}First-level heading:\protect\\ The line
break was forced \lowercase{via} \textbackslash\textbackslash}

This sample document demonstrates proper use of REV\TeX~4.1 (and
\LaTeXe) in manuscripts prepared for submission to AIP
journals. Further information can be found in the documentation included in the distribution or available at
\url{http://authors.aip.org} and in the documentation for 
REV\TeX~4.1 itself.

When commands are referred to in this example file, they are always
shown with their required arguments, using normal \TeX{} format. In
this format, \verb+#1+, \verb+#2+, etc. stand for required
author-supplied arguments to commands. For example, in
\verb+\section{#1}+ the \verb+#1+ stands for the title text of the
author's section heading, and in \verb+\title{#1}+ the \verb+#1+
stands for the title text of the paper.

Line breaks in section headings at all levels can be introduced using
\textbackslash\textbackslash. A blank input line tells \TeX\ that the
paragraph has ended.

\subsection{\label{sec:level2}Second-level heading: Formatting}

This file may be formatted in both the \texttt{preprint} (the default) and
\texttt{reprint} styles; the latter format may be used to 
mimic final journal output. Either format may be used for submission
purposes; however, for peer review and production, AIP will format the
article using the \texttt{preprint} class option. Hence, it is
essential that authors check that their manuscripts format acceptably
under \texttt{preprint}. Manuscripts submitted to AIP that do not
format correctly under the \texttt{preprint} option may be delayed in
both the editorial and production processes.

The \texttt{widetext} environment will make the text the width of the
full page, as on page~\pageref{eq:wideeq}. (Note the use the
\verb+\pageref{#1}+ to get the page number right automatically.) The
width-changing commands only take effect in \texttt{twocolumn}
formatting. It has no effect if \texttt{preprint} formatting is chosen
instead.

\subsubsection{\label{sec:level3}Third-level heading: Citations and Footnotes}

Citations in text refer to entries in the Bibliography;
they use the commands \verb+\cite{#1}+ or \verb+\onlinecite{#1}+. 
Because REV\TeX\ uses the \verb+natbib+ package of Patrick Daly, 
its entire repertoire of commands are available in your document;
see the \verb+natbib+ documentation for further details.
The argument of \verb+\cite+ is a comma-separated list of \emph{keys};
a key may consist of letters and numerals.

By default, citations are numerical; \cite{feyn54} author-year citations are an option. 
To give a textual citation, use \verb+\onlinecite{#1}+: (Refs.~\onlinecite{witten2001,epr,Bire82}). 
REV\TeX\ ``collapses'' lists of consecutive numerical citations when appropriate. 
REV\TeX\ provides the ability to properly punctuate textual citations in author-year style;
this facility works correctly with numerical citations only with \texttt{natbib}'s compress option turned off. 
To illustrate, we cite several together \cite{feyn54,witten2001,epr,Berman1983}, 
and once again (Refs.~\onlinecite{epr,feyn54,Bire82,Berman1983}). 
Note that, when numerical citations are used, the references were sorted into the same order they appear in the bibliography.

A reference within the bibliography is specified with a \verb+\bibitem{#1}+ command,
where the argument is the citation key mentioned above. 
\verb+\bibitem{#1}+ commands may be crafted by hand or, preferably,
generated by using Bib\TeX. 
The AIP styles for REV\TeX~4 include Bib\TeX\ style files
\verb+aipnum.bst+ and \verb+aipauth.bst+, appropriate for
numbered and author-year bibliographies,
respectively. 
REV\TeX~4 will automatically choose the style appropriate for 
the document's selected class options: the default is numerical, and
you obtain the author-year style by specifying a class option of \verb+author-year+.

This sample file demonstrates a simple use of Bib\TeX\ 
via a \verb+\bibliography+ command referencing the \verb+aipsamp.bib+ file.
Running Bib\TeX\ (in this case \texttt{bibtex
aipsamp}) after the first pass of \LaTeX\ produces the file
\verb+aipsamp.bbl+ which contains the automatically formatted
\verb+\bibitem+ commands (including extra markup information via
\verb+\bibinfo+ commands). If not using Bib\TeX, the
\verb+thebibiliography+ environment should be used instead.

\paragraph{Fourth-level heading is run in.}%
Footnotes are produced using the \verb+\footnote{#1}+ command. 
Numerical style citations put footnotes into the 
bibliography\footnote{Automatically placing footnotes into the bibliography requires using BibTeX to compile the bibliography.}.
Author-year and numerical author-year citation styles (each for its own reason) cannot use this method. 
Note: due to the method used to place footnotes in the bibliography, \emph{you
must re-run BibTeX every time you change any of your document's
footnotes}.

\section{Math and Equations}
Inline math may be typeset using the \verb+$+ delimiters. Bold math
symbols may be achieved using the \verb+bm+ package and the
\verb+\bm{#1}+ command it supplies. For instance, a bold $\alpha$ can
be typeset as \verb+$\bm{\alpha}$+ giving $\bm{\alpha}$. Fraktur and
Blackboard (or open face or double struck) characters should be
typeset using the \verb+\mathfrak{#1}+ and \verb+\mathbb{#1}+ commands
respectively. Both are supplied by the \texttt{amssymb} package. For
example, \verb+$\mathbb{R}$+ gives $\mathbb{R}$ and
\verb+$\mathfrak{G}$+ gives $\mathfrak{G}$

In \LaTeX\ there are many different ways to display equations, and a
few preferred ways are noted below. Displayed math will center by
default. Use the class option \verb+fleqn+ to flush equations left.

Below we have numbered single-line equations, the most common kind: 
\begin{eqnarray}
\chi_+(p)\alt{\bf [}2|{\bf p}|(|{\bf p}|+p_z){\bf ]}^{-1/2}
\left(
\begin{array}{c}
|{\bf p}|+p_z\\
px+ip_y
\end{array}\right)\;,
\\
\left\{%
 \openone234567890abc123\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}%
\right\}%
\label{eq:one}.
\end{eqnarray}
Note the open one in Eq.~(\ref{eq:one}).

Not all numbered equations will fit within a narrow column this
way. The equation number will move down automatically if it cannot fit
on the same line with a one-line equation:
\begin{equation}
\left\{
 ab12345678abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}%
\right\}.
\end{equation}

When the \verb+\label{#1}+ command is used [cf. input for
Eq.~(\ref{eq:one})], the equation can be referred to in text without
knowing the equation number that \TeX\ will assign to it. Just
use \verb+\ref{#1}+, where \verb+#1+ is the same name that used in
the \verb+\label{#1}+ command.

Unnumbered single-line equations can be typeset
using the \verb+\[+, \verb+\]+ format:
\[g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \]

\subsection{Multiline equations}

Multiline equations are obtained by using the \verb+eqnarray+
environment.  Use the \verb+\nonumber+ command at the end of each line
to avoid assigning a number:
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
\delta_{\sigma_1,-\sigma_2}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_jl_i\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1),
\end{eqnarray}
\begin{eqnarray}
\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
        (N^2-1)\nonumber \\
 & &\times \left( \sum_{i<j}\right)
  \sum_{\text{perm}}
 \frac{1}{S_{12}}
 \frac{1}{S_{12}}
 \sum_\tau c^f_\tau~.
\end{eqnarray}
\textbf{Note:} Do not use \verb+\label{#1}+ on a line of a multiline
equation if \verb+\nonumber+ is also used on that line. Incorrect
cross-referencing will result. Notice the use \verb+\text{#1}+ for
using a Roman font within a math environment.

To set a multiline equation without \emph{any} equation
numbers, use the \verb+\begin{eqnarray*}+,
\verb+\end{eqnarray*}+ format:
\begin{eqnarray*}
\sum \vert M^{\text{viol}}_g \vert ^2&=&g^{2n-4}_S(Q^2)~N^{n-2}
        (N^2-1)\\
 & &\times \left( \sum_{i<j}\right)
 \left(
  \sum_{\text{perm}}\frac{1}{S_{12}S_{23}S_{n1}}
 \right)
 \frac{1}{S_{12}}~.
\end{eqnarray*}
To obtain numbers not normally produced by the automatic numbering,
use the \verb+\tag{#1}+ command, where \verb+#1+ is the desired
equation number. For example, to get an equation number of
(\ref{eq:mynum}),
\begin{equation}
g^+g^+ \rightarrow g^+g^+g^+g^+ \dots ~,~~q^+q^+\rightarrow
q^+g^+g^+ \dots ~. \tag{2.6$'$}\label{eq:mynum}
\end{equation}

A few notes on \verb=\tag{#1}=. \verb+\tag{#1}+ requires
\texttt{amsmath}. The \verb+\tag{#1}+ must come before the
\verb+\label{#1}+, if any. The numbering set with \verb+\tag{#1}+ is
\textit{transparent} to the automatic numbering in REV\TeX{};
therefore, the number must be known ahead of time, and it must be
manually adjusted if other equations are added. \verb+\tag{#1}+ works
with both single-line and multiline equations. \verb+\tag{#1}+ should
only be used in exceptional case - do not use it to number all
equations in a paper.

Enclosing single-line and multiline equations in
\verb+\begin{subequations}+ and \verb+\end{subequations}+ will produce
a set of equations that are ``numbered'' with letters, as shown in
Eqs.~(\ref{subeq:1}) and (\ref{subeq:2}) below:
\begin{subequations}
\label{eq:whole}
\begin{equation}
\left\{
 abc123456abcdef\alpha\beta\gamma\delta1234556\alpha\beta
 \frac{1\sum^{a}_{b}}{A^2}
\right\},\label{subeq:1}
\end{equation}
\begin{eqnarray}
{\cal M}=&&ig_Z^2(4E_1E_2)^{1/2}(l_i^2)^{-1}
(g_{\sigma_2}^e)^2\chi_{-\sigma_2}(p_2)\nonumber\\
&&\times
[\epsilon_i]_{\sigma_1}\chi_{\sigma_1}(p_1).\label{subeq:2}
\end{eqnarray}
\end{subequations}
Putting a \verb+\label{#1}+ command right after the
\verb+\begin{subequations}+, allows one to
reference all the equations in a subequations environment. For
example, the equations in the preceding subequations environment were
Eqs.~(\ref{eq:whole}).

\subsubsection{Wide equations}
The equation that follows is set in a wide format, i.e., it spans
across the full page. The wide format is reserved for long equations
that cannot be easily broken into four lines or less:
\begin{widetext}
\begin{equation}
{\cal R}^{(\text{d})}=
 g_{\sigma_2}^e
 \left(
   \frac{[\Gamma^Z(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
  +\frac{[\Gamma^Z(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
 \right)
 + x_WQ_e
 \left(
   \frac{[\Gamma^\gamma(3,21)]_{\sigma_1}}{Q_{12}^2-M_W^2}
  +\frac{[\Gamma^\gamma(13,2)]_{\sigma_1}}{Q_{13}^2-M_W^2}
 \right)\;. \label{eq:wideeq}
\end{equation}
\end{widetext}
This is typed to show the output is in wide format.
(Since there is no input line between \verb+\equation+ and
this paragraph, there is no paragraph indent for this paragraph.)
\section{Cross-referencing}
REV\TeX{} will automatically number sections, equations, figure
captions, and tables. In order to reference them in text, use the
\verb+\label{#1}+ and \verb+\ref{#1}+ commands. To reference a
particular page, use the \verb+\pageref{#1}+ command.

The \verb+\label{#1}+ should appear in a section heading, within an
equation, or in a table or figure caption. The \verb+\ref{#1}+ command
is used in the text where the citation is to be displayed.  Some
examples: Section~\ref{sec:level1} on page~\pageref{sec:level1},
Table~\ref{tab:table1},%
\begin{table}
\caption{\label{tab:table1}This is a narrow table which fits into a
text column when using \texttt{twocolumn} formatting. Note that
REV\TeX~4 adjusts the intercolumn spacing so that the table fills the
entire width of the column. Table captions are numbered
automatically. This table illustrates left-aligned, centered, and
right-aligned columns.  }
\begin{ruledtabular}
\begin{tabular}{lcr}
Left\footnote{Note a.}&Centered\footnote{Note b.}&Right\\
\hline
1 & 2 & 3\\
10 & 20 & 30\\
100 & 200 & 300\\
\end{tabular}
\end{ruledtabular}
\end{table}
and Fig.~\ref{fig:epsart}.

\section{Figures and Tables}
Figures and tables are typically ``floats''; \LaTeX\ determines their
final position via placement rules. 
\LaTeX\ isn't always successful in automatically placing floats where you wish them.

Figures are marked up with the \texttt{figure} environment, the content of which
imports the image (\verb+\includegraphics+) followed by the figure caption (\verb+\caption+).
The argument of the latter command should itself contain a \verb+\label+ command if you
wish to refer to your figure with \verb+\ref+.

Import your image using either the \texttt{graphics} or
\texttt{graphix} packages. These packages both define the
\verb+\includegraphics{#1}+ command, but they differ in the optional
arguments for specifying the orientation, scaling, and translation of the figure.
Fig.~\ref{fig:epsart}%
\begin{figure}
\includegraphics{fig_1}% Here is how to import EPS art
\caption{\label{fig:epsart} A figure caption. The figure captions are
automatically numbered.}
\end{figure}
is small enough to fit in a single column, while
Fig.~\ref{fig:wide}%
\begin{figure*}
\includegraphics{fig_2}% Here is how to import EPS art
\caption{\label{fig:wide}Use the \texttt{figure*} environment to get a wide
figure, spanning the page in \texttt{twocolumn} formatting.}
\end{figure*}
is too wide for a single column,
so instead the \texttt{figure*} environment has been used.

The analog of the \texttt{figure} environment is \texttt{table}, which uses
the same \verb+\caption+ command.
However, you should type your caption command first within the \texttt{table}, 
instead of last as you did for \texttt{figure}.

The heart of any table is the \texttt{tabular} environment,
which represents the table content as a (vertical) sequence of table rows,
each containing a (horizontal) sequence of table cells. 
Cells are separated by the \verb+&+ character;
the row terminates with \verb+\\+. 
The required argument for the \texttt{tabular} environment
specifies how data are displayed in each of the columns. 
For instance, a column
may be centered (\verb+c+), left-justified (\verb+l+), right-justified (\verb+r+),
or aligned on a decimal point (\verb+d+). 
(Table~\ref{tab:table4}%
\begin{table}
\caption{\label{tab:table4}Numbers in columns Three--Five have been
aligned by using the ``d'' column specifier (requires the
\texttt{dcolumn} package). 
Non-numeric entries (those entries without
a ``.'') in a ``d'' column are aligned on the decimal point. 
Use the
``D'' specifier for more complex layouts. }
\begin{ruledtabular}
\begin{tabular}{ccddd}
One&Two&\mbox{Three}&\mbox{Four}&\mbox{Five}\\
\hline
one&two&\mbox{three}&\mbox{four}&\mbox{five}\\
He&2& 2.77234 & 45672. & 0.69 \\
C\footnote{Some tables require footnotes.}
  &C\footnote{Some tables need more than one footnote.}
  & 12537.64 & 37.66345 & 86.37 \\
\end{tabular}
\end{ruledtabular}
\end{table}
illustrates the use of decimal column alignment.)

Extra column-spacing may be be specified as well, although
REV\TeX~4 sets this spacing so that the columns fill the width of the
table.
Horizontal rules are typeset using the \verb+\hline+
command.
The doubled (or Scotch) rules that appear at the top and
bottom of a table can be achieved by enclosing the \texttt{tabular}
environment within a \texttt{ruledtabular} environment.
Rows whose columns span multiple columns can be typeset using \LaTeX's
\verb+\multicolumn{#1}{#2}{#3}+ command
(for example, see the first row of Table~\ref{tab:table3}).%
\begin{table*}
\caption{\label{tab:table3}This is a wide table that spans the page
width in \texttt{twocolumn} mode. It is formatted using the
\texttt{table*} environment. It also demonstrates the use of
\textbackslash\texttt{multicolumn} in rows with entries that span
more than one column.}
\begin{ruledtabular}
\begin{tabular}{ccccc}
 &\multicolumn{2}{c}{$D_{4h}^1$}&\multicolumn{2}{c}{$D_{4h}^5$}\\
 Ion&1st alternative&2nd alternative&lst alternative
&2nd alternative\\ \hline
 K&$(2e)+(2f)$&$(4i)$ &$(2c)+(2d)$&$(4f)$ \\
 Mn&$(2g)$\footnote{The $z$ parameter of these positions is $z\sim\frac{1}{4}$.}
 &$(a)+(b)+(c)+(d)$&$(4e)$&$(2a)+(2b)$\\
 Cl&$(a)+(b)+(c)+(d)$&$(2g)$\footnote{This is a footnote in a table that spans the full page
width in \texttt{twocolumn} mode. It is supposed to set on the full width of the page, just as the caption does. }
 &$(4e)^{\text{a}}$\\
 He&$(8r)^{\text{a}}$&$(4j)^{\text{a}}$&$(4g)^{\text{a}}$\\
 Ag& &$(4k)^{\text{a}}$& &$(4h)^{\text{a}}$\\
\end{tabular}
\end{ruledtabular}
\end{table*}

The tables in this document illustrate various effects.
Tables that fit in a narrow column are contained in a \texttt{table}
environment.
Table~\ref{tab:table3} is a wide table, therefore set with the
\texttt{table*} environment.
Lengthy tables may need to break across pages.
A simple way to allow this is to specify
the \verb+[H]+ float placement on the \texttt{table} or
\texttt{table*} environment.
Alternatively, using the standard \LaTeXe\ package \texttt{longtable} 
gives more control over how tables break and allows headers and footers 
to be specified for each page of the table.
An example of the use of \texttt{longtable} can be found
in the file \texttt{summary.tex} that is included with the REV\TeX~4
distribution.

There are two methods for setting footnotes within a table (these
footnotes will be displayed directly below the table rather than at
the bottom of the page or in the bibliography).
The easiest
and preferred method is just to use the \verb+\footnote{#1}+
command. This will automatically enumerate the footnotes with
lowercase roman letters.
However, it is sometimes necessary to have
multiple entries in the table share the same footnote.
In this case,
create the footnotes using
\verb+\footnotemark[#1]+ and \verb+\footnotetext[#1]{#2}+.
\texttt{\#1} is a numeric value.
Each time the same value for \texttt{\#1} is used, 
the same mark is produced in the table. 
The \verb+\footnotetext[#1]{#2}+ commands are placed after the \texttt{tabular}
environment. 
Examine the \LaTeX\ source and output for Tables~\ref{tab:table1} and 
\ref{tab:table2}%
\begin{table}
\caption{\label{tab:table2}A table with more columns still fits
properly in a column. Note that several entries share the same
footnote. Inspect the \LaTeX\ input for this table to see
exactly how it is done.}
\begin{ruledtabular}
\begin{tabular}{cccccccc}
 &$r_c$ (\AA)&$r_0$ (\AA)&$\kappa r_0$&
 &$r_c$ (\AA) &$r_0$ (\AA)&$\kappa r_0$\\
\hline
Cu& 0.800 & 14.10 & 2.550 &Sn\footnotemark[1]
& 0.680 & 1.870 & 3.700 \\
Ag& 0.990 & 15.90 & 2.710 &Pb\footnotemark[2]
& 0.450 & 1.930 & 3.760 \\
Au& 1.150 & 15.90 & 2.710 &Ca\footnotemark[3]
& 0.750 & 2.170 & 3.560 \\
Mg& 0.490 & 17.60 & 3.200 &Sr\footnotemark[4]
& 0.900 & 2.370 & 3.720 \\
Zn& 0.300 & 15.20 & 2.970 &Li\footnotemark[2]
& 0.380 & 1.730 & 2.830 \\
Cd& 0.530 & 17.10 & 3.160 &Na\footnotemark[5]
& 0.760 & 2.110 & 3.120 \\
Hg& 0.550 & 17.80 & 3.220 &K\footnotemark[5]
&  1.120 & 2.620 & 3.480 \\
Al& 0.230 & 15.80 & 3.240 &Rb\footnotemark[3]
& 1.330 & 2.800 & 3.590 \\
Ga& 0.310 & 16.70 & 3.330 &Cs\footnotemark[4]
& 1.420 & 3.030 & 3.740 \\
In& 0.460 & 18.40 & 3.500 &Ba\footnotemark[5]
& 0.960 & 2.460 & 3.780 \\
Tl& 0.480 & 18.90 & 3.550 & & & & \\
\end{tabular}
\end{ruledtabular}
\footnotetext[1]{Here's the first, from Ref.~\onlinecite{feyn54}.}
\footnotetext[2]{Here's the second.}
\footnotetext[3]{Here's the third.}
\footnotetext[4]{Here's the fourth.}
\footnotetext[5]{And etc.}
\end{table}
for an illustration.

All AIP journals require that the initial citation of
figures or tables be in numerical order.
\LaTeX's automatic numbering of floats is your friend here:
just put each \texttt{figure} environment immediately following 
its first reference (\verb+\ref+), as we have done in this example file.

\begin{acknowledgments}
We wish to acknowledge the support of the author community in using
REV\TeX{}, offering suggestions and encouragement, testing new versions,
\dots.
\end{acknowledgments}

\appendix

\section{Appendixes}

To start the appendixes, use the \verb+\appendix+ command.
This signals that all following section commands refer to appendixes
instead of regular sections. Therefore, the \verb+\appendix+ command
should be used only once---to set up the section commands to act as
appendixes. Thereafter normal section commands are used. The heading
for a section can be left empty. For example,
\begin{verbatim}
\appendix
\section{}
\end{verbatim}
will produce an appendix heading that says ``APPENDIX A'' and
\begin{verbatim}
\appendix
\section{Background}
\end{verbatim}
will produce an appendix heading that says ``APPENDIX A: BACKGROUND''
(note that the colon is set automatically).

If there is only one appendix, then the letter ``A'' should not
appear. This is suppressed by using the star version of the appendix
command (\verb+\appendix*+ in the place of \verb+\appendix+).

\section{A little more on appendixes}

Observe that this appendix was started by using
\begin{verbatim}
\section{A little more on appendixes}
\end{verbatim}

Note the equation number in an appendix:
\begin{equation}
E=mc^2.
\end{equation}

\subsection{\label{app:subsec}A subsection in an appendix}

You can use a subsection or subsubsection in an appendix. Note the
numbering: we are now in Appendix~\ref{app:subsec}.

\subsubsection{\label{app:subsubsec}A subsubsection in an appendix}
Note the equation numbers in this appendix, produced with the
subequations environment:
\begin{subequations}
\begin{eqnarray}
E&=&mc, \label{appa}
\\
E&=&mc^2, \label{appb}
\\
E&\agt& mc^3. \label{appc}
\end{eqnarray}
\end{subequations}
They turn out to be Eqs.~(\ref{appa}), (\ref{appb}), and (\ref{appc}).

\nocite{*}
\bibliography{aipsamp}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file aipsamp.tex ******

### 2018-05-21 oepdev/libgefp/gefp.h
Statistical data set structure introduced

@@ -12,6 +12,7 @@
 #include "psi4/libmints/vector3.h"
 #include "../liboep/oep.h"
 #include "../libutil/cphf.h"
+#include "../libutil/scf_perturb.h"
 
 namespace oepdev{
 
@@ -511,14 +512,14 @@ class PolarGEFactory : public GenEffParFactory //, public std::enable_shared_fro
    /// Randomly draw charge value
    double draw_charge();
 
-   /// Solve SCF equations to find perturbed one-particle density matrix due to uniform electric field
-   std::shared_ptr<psi::Matrix> perturbed_dmat(const std::shared_ptr<psi::Vector>& field);
+   /// Solve SCF equations to find perturbed state due to uniform electric field
+   std::shared_ptr<oepdev::RHFPerturbed> perturbed_state(const std::shared_ptr<psi::Vector>& field);
 
-   /// Solve SCF equations to find perturbed one-particle density matrix due to point charge
-   std::shared_ptr<psi::Matrix> perturbed_dmat(const std::shared_ptr<psi::Vector>& pos, const double& charge);
+   /// Solve SCF equations to find perturbed state due to point charge
+   std::shared_ptr<oepdev::RHFPerturbed> perturbed_state(const std::shared_ptr<psi::Vector>& pos, const double& charge);
 
-   /// Solve SCF equations to find perturbed one-particle density matrix due set of point charges
-   std::shared_ptr<psi::Matrix> perturbed_dmat(const std::shared_ptr<psi::Matrix>& charges);
+   /// Solve SCF equations to find perturbed state due set of point charges
+   std::shared_ptr<oepdev::RHFPerturbed> perturbed_state(const std::shared_ptr<psi::Matrix>& charges);
 
    /// Evaluate electric field at point (x,y,z) due to point charges 
    std::shared_ptr<psi::Vector> field_due_to_charges(const std::shared_ptr<psi::Matrix>& charges, 
@@ -640,12 +641,27 @@ class GeneralizedPolarGEFactory : public PolarGEFactory
    bool hasQuadrupolePolarizability_;
 
 
-   // --> Sets of statistical data <-- //
+   // --> Sets of statistical data <-- //  ----> TODO: Group them into objects (structures)
 
-   /// Reference density matrix set
-   std::vector<std::shared_ptr<psi::Matrix>> referenceDensityMatrixSet_;
-   /// Model density matrix set
[…]

### 2018-05-21 oepdev/libgefp/gefp_polar_base.cc

@@ -20,17 +20,25 @@ oepdev::GeneralizedPolarGEFactory::GeneralizedPolarGEFactory(std::shared_ptr<CPH
    mField_(options_.get_double("DMATPOL_SCALE_1")),
    Zinit_(-1.0),
    Z_(-1.0),
-   referenceDensityMatrixSet_({}),
-   modelDensityMatrixSet_({}),
+   referenceStatisticalSet_({{},{},{},{}}),
+   modelStatisticalSet_({{},{},{},{}}),
+   VMatrixSet_({}),
    electricFieldSet_({}),
    electricFieldGradientSet_({}),
    electricFieldSumSet_({}),
    electricFieldGradientSumSet_({})
 {
-   // Allocate memory for density matrix sets
+   // Allocate memory for matrix sets
    for (int n=0; n<nSamples_; ++n) {
-        referenceDensityMatrixSet_.push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
-        modelDensityMatrixSet_    .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
+        VMatrixSet_                                         .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
+        referenceStatisticalSet_.DensityMatrixSet           .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
+        referenceStatisticalSet_.InducedDipoleSet           .push_back(std::make_shared<psi::Vector>("", 3));
+        referenceStatisticalSet_.InducedQuadrupoleSet       .push_back(std::make_shared<psi::Matrix>("", 3, 3));
+        referenceStatisticalSet_.InducedInteractionEnergySet.push_back(0.0);
+        modelStatisticalSet_.DensityMatrixSet               .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
+        modelStatisticalSet_.InducedDipoleSet               .push_back(std::make_shared<psi::Vector>("", 3));
+        modelStatisticalSet_.InducedQuadrupoleSet           .push_back(std::make_shared<psi::Matrix>("", 3, 3));
+        modelStatisticalSet_.InducedInteractionEnergySet    .push_back(0.0);
    }
 }
 oepdev::GeneralizedPolarGEFactory::GeneralizedPolarGEFactory(std::shared_ptr<CPHF> cphf)
[…]

### 2018-05-21 oepdev/libgefp/gefp_polar_nonuniform_field_1.cc

@@ -28,7 +28,7 @@ void oepdev::LinearNonUniformEFieldPolarGEFactory::compute_gradient(int i, int j
         int sz = 3*s + z;
         double g = 0.0;
         for (int n=0; n<nSamples_; ++n) {                            
-             double dij = -referenceDensityMatrixSet_[n]->get(i, j);
+             double dij = -referenceStatisticalSet_.DensityMatrixSet[n]->get(i, j);
              g += dij * electricFieldSet_[n][s]->get(z);
         }
         Gradient_->set(sz, 0, g);

### 2018-05-21 oepdev/libgefp/gefp_polar_nonuniform_field_2.cc

@@ -32,7 +32,7 @@ void oepdev::QuadraticNonUniformEFieldPolarGEFactory::compute_gradient(int i, in
         double g1 = 0.0;
         double g2 = 0.0;
         for (int n=0; n<nSamples_; ++n) {                                                                  
-             double dij = -referenceDensityMatrixSet_[n]->get(i, j);
+             double dij = -referenceStatisticalSet_.DensityMatrixSet[n]->get(i, j);
              double fz = electricFieldSet_[n][s]->get(z);
              double fs = electricFieldSumSet_[n][s] * 2.0 * mField_;
              g1 += dij * fz;

### 2018-05-21 oepdev/libgefp/gefp_polar_uniform_field_1.cc

@@ -27,7 +27,7 @@ void oepdev::LinearUniformEFieldPolarGEFactory::compute_gradient(int i, int j)
    double g_y = 0.0;
    double g_z = 0.0;
    for (int n=0; n<nSamples_; ++n) {
-        double dij = -referenceDensityMatrixSet_[n]->get(i, j);
+        double dij = -referenceStatisticalSet_.DensityMatrixSet[n]->get(i, j);
         g_x += dij * electricFieldSet_[n][0]->get(0);
         g_y += dij * electricFieldSet_[n][0]->get(1);
         g_z += dij * electricFieldSet_[n][0]->get(2);

### 2018-05-22 papers/1_dmatpol/paper-r1.tex
RMSE indicator coded and working well

@@
 \begin{multline}\label{e:dE}
- \delta E[{\bf F}] = 2\Tr{\left\{ {\bf D}^{(0)} \cdot {\bf V}[{\bf F}] \right\}}
+ \delta E[{\bf F}] = 
+                     \Tr{\left\{ \delta {\bf D} \cdot
+                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}}
                    + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
-                                \left( \delta {\bf J} + \delta {\bf K}\right) \right\}} \\
-                   + \Tr{\left\{ \delta {\bf D} \cdot
-                                \left( {\bf H}^{(0)} + 2 {\bf V}[{\bf F}] + {\bf f}^{(0)} \right) \right\}}
+                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} 
                    + E_{\rm nuc}({\bf F}) \;.
@@
 the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
-${\bf V}[{\bf F}]$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
+${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
 and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
@@
    \begin{equation}
-     {\rm RMSE} = \sum_N \sqrt{ \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
+     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
    \end{equation}
@@
    \begin{equation}
-     {\rm RMSD} = \sum_N \sqrt{ \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
+     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
    \end{equation}
@@
    \begin{equation}
-     {\rm RMSQ} = \sum_N \sqrt{ \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
+     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
    \end{equation}

### 2018-05-22 main.cc

@@ -199,6 +199,9 @@ int read_options(std::string name, Options& options)
         options.add_double ("DMATPOL_TEST_FIELD_Y"  , 0.000                      );
         options.add_double ("DMATPOL_TEST_FIELD_Z"  , 0.008                      );
 
+        /*- Output file name for statistical evaluation for DMATPOL -*/
+        options.add_str    ("DMATPOL_OUT_STATS"     , "dmatpol.stats.dat"        );
+
      }
 
     return true;

### 2018-05-22 oepdev/libgefp/gefp.h

@@ -438,6 +438,7 @@ class GenEffParFactory //: public std::enable_shared_from_this<GenEffParFactory>
     *   - `DMATPOL_TEST_FIELD_X`  - test electric field in X direction. Default: `0.000` [au]
     *   - `DMATPOL_TEST_FIELD_Y`  - test electric field in Y direction. Default: `0.000` [au]
     *   - `DMATPOL_TEST_FIELD_Z`  - test electric field in Z direction. Default: `0.008` [au]
+    *   - `DMATPOL_OUT_STATS`     - output file name for statistical evaluation results. Default: `dmatpol.stats.dat`
     */
    static std::shared_ptr<GenEffParFactory> build(const std::string& type, 
                                                   std::shared_ptr<CPHF> cphf, psi::Options& opt);

### 2018-05-22 oepdev/libgefp/gefp_polar_base.cc

@@ -201,9 +201,8 @@ void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
    cout << " Statistical evaluation ...\n";
 
    // Grab some unperturbed quantities
-   std::shared_ptr<psi::Matrix> Hcore  = wfn_->H();
-   std::shared_ptr<psi::Matrix> Fock   = wfn_->Fa();
-   std::shared_ptr<psi::Matrix> D      = wfn_->Da();
+   std::shared_ptr<psi::Matrix> H = wfn_->Fa(); H->add(wfn_->H());
+   std::shared_ptr<psi::Matrix> D = wfn_->Da();
 
    // Compute model difference density matrices
    if (hasQuadrupolePolarizability_) {
@@ -220,62 +219,99 @@ void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
 
    // Compute changes in the Coulomb and exchange HF matrices
    std::vector<psi::SharedMatrix>& C_left = jk_->C_left();
+   std::vector<psi::SharedMatrix>& C_right= jk_->C_right();
    const std::vector<std::shared_ptr<psi::Matrix>>& J = jk_->J();
    const std::vector<std::shared_ptr<psi::Matrix>>& K = jk_->K();
 
-   C_left.clear();
+   C_left.clear(); C_right.clear();
    for (int n=0; n<nSamples_; ++n) {
-        std::shared_ptr<psi::Matrix> C = referenceStatisticalSet_.DensityMatrixSet[n]->partial_cholesky_factorize();
-        C_left.push_back(C);
+        std::shared_ptr<psi::Matrix> Cl = referenceStatisticalSet_.DensityMatrixSet[n]->clone();
+        std::shared_ptr<psi::Matrix> Cr= Cl->clone(); Cr->identity();
+        C_left.push_back(Cl);
+        C_right.push_back(Cr);
    }
    jk_->compute();
    for (int n=0; n<nSamples_; ++n) {
-        referenceStatisticalSet_.JKMatrixSet[n]->add(J[n]);
-        referenceStatisticalSet_.JKMatrixSet[n]->add(K[n]);
+        referenceStatisticalSet_.JKMatrixSet[n]->axpy(2.0,J[n]);
+        referenceStatisticalSet_.JKMatrixSet[n]->subtract(K[n]);
    }
-   C_left.clear();
+   C_left.clear(); C_right.clear();
    for (int n=0; n<nSamples_; ++n) {
[…]

### 2018-05-22 oepdev/libgefp/gefp_polar_nonuniform_base.cc

@@ -48,7 +48,13 @@ void oepdev::NonUniformEFieldPolarGEFactory::compute_samples(void)
         referenceStatisticalSet_.DensityMatrixSet[n]->subtract(wfn_->Da());
         VMatrixSet_[n]->copy(pert->Vpert());
 
+
         electricFieldSet_.push_back(fields_n);
+
+        referenceStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+            modelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+
+        cout << oepdev::string_sprintf(" Interaction Energy = %15.6f\n", pert->reference_energy() - wfn_->reference_energy());
    }
  
 }

### 2018-05-22 oepdev/libgefp/gefp_polar_uniform_base.cc

@@ -36,6 +36,11 @@ void oepdev::UniformEFieldPolarGEFactory::compute_samples(void)
         VMatrixSet_[n]->copy(pert->Vpert());
 
         electricFieldSet_.push_back(fields);
+
+        referenceStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+            modelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+
+        cout << oepdev::string_sprintf(" Interaction Energy = %15.6f\n", pert->reference_energy() - wfn_->reference_energy());
    }
 }
 // abstract methods

### 2018-05-22 oepdev/libutil/scf_perturb.cc

@@ -24,6 +24,7 @@ RHFPerturbed::RHFPerturbed(std::shared_ptr<psi::Wavefunction> ref_wfn, std::shar
 void RHFPerturbed::common_init()
 {
    Vpert_ = std::make_shared<psi::Matrix>("External Potential Matrix", basisset_->nbf(), basisset_->nbf());
+   nuclearInteractionEnergy_ = 0.0;
 }
 RHFPerturbed::~RHFPerturbed()
 {
@@ -88,7 +89,7 @@ void RHFPerturbed::perturb_Hcore()
   }
 
   // Save the perturbation potential one-electron matrix
-  V_->copy(Hadd);
+  Vpert_->copy(Hadd);
 
   // Add perturbation to Hcore matrix
   H_->add(Hadd);
@@ -99,16 +100,17 @@ void RHFPerturbed::perturb_Hcore()
   psi::timer_off("HF: Guess-Perturbed");
 
   // Add the contribution from nuclei
-  for (int z=0; z<3; ++z) nuclearrep_ -= perturbField_->get(z) * molecule_->nuclear_dipole().get(z);
+  for (int z=0; z<3; ++z) nuclearInteractionEnergy_ -= perturbField_->get(z) * molecule_->nuclear_dipole().get(z);
   for (int I=0; I<molecule_->natom(); ++I) {
        for (int n=0; n<perturbCharges_->charges.size(); ++n){
             double x = molecule_->x(I) - perturbCharges_->positions[n]->get(0);
             double y = molecule_->y(I) - perturbCharges_->positions[n]->get(1);
             double z = molecule_->z(I) - perturbCharges_->positions[n]->get(2);
             double r_In = sqrt(x*x + y*y + z*z);
-            nuclearrep_ += perturbCharges_->charges[n] * (double)(molecule_->Z(I)) / r_In;
+            nuclearInteractionEnergy_ += perturbCharges_->charges[n] * (double)(molecule_->Z(I)) / r_In;
        }
   }
+  nuclearrep_ += nuclearInteractionEnergy_;
 }
 
 } // EndNameSpace oepdev

### 2018-05-22 oepdev/libutil/scf_perturb.h

@@ -85,6 +85,9 @@ class RHFPerturbed : public psi::scf::RHF
     /// Get a copy of the perturbation potential one-electron matrix
     std::shared_ptr<psi::Matrix> Vpert() const {return std::make_shared<psi::Matrix>(Vpert_);}
 
+    /// Get the interaction energy of the nuclei with the perturbing potential
+    double nuclear_interaction_energy() const {return nuclearInteractionEnergy_;}
+
   protected:
 
     /// Perturbing electric field
@@ -94,6 +97,9 @@ class RHFPerturbed : public psi::scf::RHF
 
     /// Perturbation potential one-electron matrix
     std::shared_ptr<psi::Matrix> Vpert_;
+
+    /// Electrostatic interaction energy due to nuclei
+    double nuclearInteractionEnergy_;
     
     /// Add the electrostatic perturbation to the Hcore matrix
     virtual void perturb_Hcore();

### 2018-05-25 papers/1_dmatpol/paper-r1.tex
Bug in absolute quadrupole magnitude removed

@@
 due to the nuclei.
+In order to compare the performance of the distributed polarizability approach,
+the approximate interaction energy was also evaluated by using the expression:
+%
+\begin{equation}\label{e:dE-dpol}
+ \delta E[{\bf F}] \approx 2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} } 
+                   + \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
+                   + E_{\rm nuc}({\bf F}) \;,
+\end{equation}
+%
+in which ${\BM \alpha}_i$ is the distributed anisotropic polarizability tensor associated with the
+$i$th localized molecular orbital (LMO) centroid, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
+The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
+method. Orbital localization was performed by using the Boys method. Note that in Eq.~\eqref{e:dE-dpol}, 
+multipole approximation is introduced only for the part associated with the electric field\hyp{}induced
+change in the one\hyp{}particle density matrix, i.e., the polarization of the electron density distribution.
+
 The electric field\hyp{}induced dipole and quadrupole moments were evaluated from the following expressions:
@@
 %
+Quadrupole moments were subsequently transformed to the their traceless forms by
+%
+\begin{equation}
+  \delta \widetilde{\Theta}_{uw} = \delta \Theta_{uw} - \frac{1}{3} \delta_{uw} \Tr{\BM\Theta} \;.
+\end{equation}
+%
 To compute the interaction energy and the induced multipole moments, either exact or approximated forms
@@
  \item Root Mean Square Energy Error (RMSE)
-   \begin{equation}
+   \begin{equation}\label{e:rmse}
      {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
@@
  \item Root Mean Square Dipole Error (RMSD)
-   \begin{equation}
+   \begin{equation}\label{e:rmsd}
      {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
@@
  \item Root Mean Square Quadrupole Error (RMSQ)
-   \begin{equation}
-     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
+   \begin{equation}\label{e:rmsq}
+     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta \widetilde{\BM\Theta}^{(N)} - \overline{\delta \widetilde{\BM\Theta}^{(N)}}\rvert^2 }
    \end{equation}
@@
 wherease the latter two contain crucial information on how well the deformation of the one\hyp{}electron
-density is described.
-
-Throughout the Work, the Hartree\hyp{}Fock level was assumed and the aug-cc-pVTZ basis set
+density is described. For Eqs.~\eqref{e:rmsd} and \eqref{e:rmsq} the absolute magnitudes
+are defined as
+%
+\begin{subequations}\label{e:absmagn}
+  \begin{align}
+   \lvert {\BM\upmu} \rvert &= \sqrt{\mu_x^2+\mu_y^2+\mu_z^2} \;,\\
+   \lvert {\widetilde{\BM\Theta}}\rvert &= 
+          \sqrt{\widetilde{\Theta}_{zz}^2 + \frac{1}{3}\left(\widetilde{\Theta}_{xx}-\widetilde{\Theta}_{yy}\right)^2 
+          + \frac{4}{3}\left( \widetilde{\Theta}_{xy}^2 + \widetilde{\Theta}_{xz}^2 + \widetilde{\Theta}_{yz}^2 \right)} \;. 
+  \end{align}
+\end{subequations}
+%
+Throughout the Work, the Hartree\hyp{}Fock level was assumed and the 6-311++G** basis set
 was used. For all the calculations of the density matrix polarization susceptibility tensors

### 2018-05-25 oepdev/libgefp/gefp_polar_base.cc

@@ -393,11 +393,11 @@ void oepdev::GeneralizedPolarGEFactory::compute_ab_initio(void) {
 void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
    cout << " Statistical evaluation ...\n";
 
-   // Grab some unperturbed quantities
+   // ---> Grab some unperturbed quantities <--- //
    std::shared_ptr<psi::Matrix> H = wfn_->Fa(); H->add(wfn_->H());
    std::shared_ptr<psi::Matrix> D = wfn_->Da();
 
-   // Compute model difference density matrices
+   // ---> Compute model difference density matrices <--- //
    if (hasQuadrupolePolarizability_) {
       for (int n=0; n<nSamples_; ++n) {                                                                          
            modelStatisticalSet_.DensityMatrixSet[n]->copy(PolarizationSusceptibilities_->compute_density_matrix(
@@ -415,7 +415,7 @@ void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
       }
    }
 
-   // Compute changes in the Coulomb and exchange HF matrices
+   // ---> Compute changes in the Coulomb and exchange HF matrices <--- //
    std::vector<psi::SharedMatrix>& C_left = jk_->C_left();
    std::vector<psi::SharedMatrix>& C_right= jk_->C_right();
    const std::vector<std::shared_ptr<psi::Matrix>>& J = jk_->J();
@@ -516,10 +516,6 @@ void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
         double qr_yy= 2.0 * referenceStatisticalSet_.DensityMatrixSet[n]->vector_dot(qadInts[3]);
         double qr_yz= 2.0 * referenceStatisticalSet_.DensityMatrixSet[n]->vector_dot(qadInts[4]);
         double qr_zz= 2.0 * referenceStatisticalSet_.DensityMatrixSet[n]->vector_dot(qadInts[5]);
-               dr_av = sqrt(dr_x*dr_x + dr_y*dr_y + dr_z*dr_z);
-        double tr = (1.0 / 3.0) * (qr_xx + qr_yy + qr_zz);
[…]

### 2018-05-25 oepdev/libutil/util.cc

@@ -63,13 +63,12 @@ solve_scf(std::shared_ptr<Molecule> molecule,
 extern "C" double average_moment(std::shared_ptr<psi::Vector> moment)
 {
   const int l = moment->dim();
-  // Dipole
-  if (l == 3) {
+  if (l == 3) { // Dipole
      double da_x = moment->get(0);
      double da_y = moment->get(1);
      double da_z = moment->get(2);
      return sqrt(da_x*da_x + da_y*da_y + da_z*da_z);
-  } else if (l == 6) {
+  } else if (l == 6) { // Quadrupole
      double qa_xx= moment->get(0);
      double qa_xy= moment->get(1);
      double qa_xz= moment->get(2);
@@ -78,7 +77,7 @@ extern "C" double average_moment(std::shared_ptr<psi::Vector> moment)
      double qa_zz= moment->get(5);
      double ta = (1.0 / 3.0) * (qa_xx + qa_yy + qa_zz);
      return sqrt((qa_zz-ta)*(qa_zz-ta) + (4.0/3.0)*(qa_xy*qa_xy+qa_xz*qa_xz+qa_yz*qa_yz) 
-                                                  + (1.0/3.0)*((qa_xx-ta)*(qa_xx-ta) - (qa_yy-ta)*(qa_yy-ta)));
+                                                  + (1.0/3.0)*(pow((qa_xx-ta) - (qa_yy-ta), 2.0)));
   } else {throw PSIEXCEPTION("Wrong size of multipole moment vector!");}
 }
 

### 2018-05-25 papers/1_dmatpol/paper-r1.tex
Comparison with distributed polarizability model added

@@
 One very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
-the density matrix increment dependent {\emph only} by susceptibility element associated with the same
+the density matrix increment is dependent \emph{only} on the susceptibility element associated with the same
 pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
@@
  \delta E[{\bf F}] \approx 2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} } 
-                   + \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
+                   - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
                    + E_{\rm nuc}({\bf F}) \;,

### 2018-05-25 oepdev/libgefp/gefp.cc

@@ -186,7 +186,8 @@ oepdev::GenEffParFactory::GenEffParFactory(std::shared_ptr<psi::Wavefunction> wf
  options_(opt),
  nbf_(wfn->basisset()->nbf()),
  randomDistribution_(std::uniform_real_distribution<double>(-1.0, 1.0)),
- cphfSolver_(nullptr)
+ cphfSolver_(nullptr),
+ abInitioPolarizationSusceptibilitiesFactory_(nullptr)
 {
    // --> Future: encapsulate the ESP point space with vdW excluded spheres into an object and put to 3D points library <-- //
 

### 2018-05-25 oepdev/libgefp/gefp.h

@@ -512,6 +512,9 @@ class GenEffParFactory
 
    /// The CPHF object
    std::shared_ptr<CPHF> cphfSolver_;
+
+   /// Ab initio polarization susceptibility factory
+   std::shared_ptr<oepdev::GenEffParFactory> abInitioPolarizationSusceptibilitiesFactory_;
 };
 
 /** \brief Polarization GEFP Factory. Abstract Base.
@@ -705,6 +708,8 @@ class GeneralizedPolarGEFactory : public PolarGEFactory
 
    /// Reference statistical data
    StatisticalSet referenceStatisticalSet_;
+   /// Multipole reference statistical data
+   StatisticalSet referenceDpolStatisticalSet_;
    /// Model statistical data
    StatisticalSet modelStatisticalSet_;
    /// Ab Initio Model statistical data

### 2018-05-25 oepdev/libgefp/gefp_polar_base.cc

@@ -163,6 +163,7 @@ oepdev::GeneralizedPolarGEFactory::GeneralizedPolarGEFactory(std::shared_ptr<psi
    Zinit_(-1.0),
    Z_(-1.0),
    referenceStatisticalSet_({{},{},{},{},{}}),
+   referenceDpolStatisticalSet_({{},{},{},{},{}}),
    modelStatisticalSet_({{},{},{},{},{}}),
    abInitioModelStatisticalSet_({{},{},{},{},{}}),
    VMatrixSet_({}),
@@ -181,12 +182,19 @@ oepdev::GeneralizedPolarGEFactory::GeneralizedPolarGEFactory(std::shared_ptr<psi
         referenceStatisticalSet_.InducedDipoleSet           .push_back(std::make_shared<psi::Vector>("", 3));
         referenceStatisticalSet_.InducedQuadrupoleSet       .push_back(std::make_shared<psi::Vector>("", 6));
         referenceStatisticalSet_.InducedInteractionEnergySet.push_back(0.0);
+
         modelStatisticalSet_.DensityMatrixSet               .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
         modelStatisticalSet_.JKMatrixSet                    .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_));
         modelStatisticalSet_.InducedDipoleSet               .push_back(std::make_shared<psi::Vector>("", 3));
         modelStatisticalSet_.InducedQuadrupoleSet           .push_back(std::make_shared<psi::Vector>("", 6));
         modelStatisticalSet_.InducedInteractionEnergySet    .push_back(0.0);
+
         //if (hasAbInitioDipolePolarizability_) {
+
+           referenceDpolStatisticalSet_.InducedDipoleSet           .push_back(std::make_shared<psi::Vector>("", 3)); 
+           referenceDpolStatisticalSet_.InducedQuadrupoleSet       .push_back(std::make_shared<psi::Vector>("", 6));
+           referenceDpolStatisticalSet_.InducedInteractionEnergySet.push_back(0.0);
+
            abInitioModelStatisticalSet_.DensityMatrixSet           .push_back(std::make_shared<psi::Matrix>("", nbf_, nbf_)); 
[…]

### 2018-05-25 oepdev/libgefp/gefp_polar_nonuniform_base.cc

@@ -11,10 +11,6 @@ oepdev::NonUniformEFieldPolarGEFactory::NonUniformEFieldPolarGEFactory(std::shar
    // Atoms are assumed to be distributed centres
    nSites_ = wfn_->molecule()->natom();
 }
-//oepdev::NonUniformEFieldPolarGEFactory::NonUniformEFieldPolarGEFactory(std::shared_ptr<CPHF> cphf)
-// : oepdev::NonUniformEFieldPolarGEFactory(cphf, cphf->options())
-//{
-//}
 oepdev::NonUniformEFieldPolarGEFactory::~NonUniformEFieldPolarGEFactory()
 {
 
@@ -23,6 +19,8 @@ oepdev::NonUniformEFieldPolarGEFactory::~NonUniformEFieldPolarGEFactory()
 void oepdev::NonUniformEFieldPolarGEFactory::compute_samples(void)
 {
    int nq   = options_.get_double("DMATPOL_NTEST_CHARGE");
+
+   std::shared_ptr<psi::Matrix> D = wfn_->Da();
 
    for (int n=0; n<nSamples_; ++n) {
         std::shared_ptr<psi::Matrix> charges = std::make_shared<psi::Matrix>("Q", nq, 4);
@@ -69,6 +67,8 @@ void oepdev::NonUniformEFieldPolarGEFactory::compute_samples(void)
             }
             abInitioModelElectricFieldSet_.push_back(fields_abini_n);
             abInitioModelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+            referenceDpolStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy() 
+                         + 2.0*D->vector_dot(VMatrixSet_[n]);
         }
 
    }

### 2018-05-25 oepdev/libgefp/gefp_polar_uniform_base.cc

@@ -11,10 +11,6 @@ oepdev::UniformEFieldPolarGEFactory::UniformEFieldPolarGEFactory(std::shared_ptr
    // Set the number of sites to one since the electric field is uniform in space
    nSites_ = 1;
 }
-//oepdev::UniformEFieldPolarGEFactory::UniformEFieldPolarGEFactory(std::shared_ptr<CPHF> cphf)
-// : oepdev::UniformEFieldPolarGEFactory(cphf, cphf->options())
-//{
-//}
 oepdev::UniformEFieldPolarGEFactory::~UniformEFieldPolarGEFactory()
 {
 
@@ -22,6 +18,8 @@ oepdev::UniformEFieldPolarGEFactory::~UniformEFieldPolarGEFactory()
 // implementations of abstract methods from base
 void oepdev::UniformEFieldPolarGEFactory::compute_samples(void)
 {
+   std::shared_ptr<psi::Matrix> D = wfn_->Da();
+
    for (int n=0; n<nSamples_; ++n) {
         std::vector<std::shared_ptr<psi::Vector>> fields;
         std::shared_ptr<psi::Vector> field = draw_field();
@@ -38,8 +36,9 @@ void oepdev::UniformEFieldPolarGEFactory::compute_samples(void)
 
         electricFieldSet_.push_back(fields);
 
-        referenceStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
-            modelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+            referenceStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+                modelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
+
 
         cout << oepdev::string_sprintf(" Interaction Energy = %15.6f\n", pert->reference_energy() - wfn_->reference_energy());
 
@@ -47,6 +46,8 @@ void oepdev::UniformEFieldPolarGEFactory::compute_samples(void)
             for (int o=1; o<wfn_->doccpi()[0]; ++o) fields.push_back(field);
             abInitioModelElectricFieldSet_.push_back(fields);
             abInitioModelStatisticalSet_.InducedInteractionEnergySet[n] = pert->nuclear_interaction_energy();
[…]

### 2018-05-25 main.cc
Routine for DMATPOL calculations added in main

@@ -64,6 +64,7 @@
 #include "oepdev/libutil/solver.h"
 #include "oepdev/liboep/oep.h"
 #include "oepdev/libpsi/potential.h"
+#include "oepdev/libgefp/gefp.h"
 
 #include "psi4/libtrans/mospace.h"
 #include "psi4/libtrans/integraltransform.h"
@@ -275,6 +276,12 @@ SharedWavefunction oepdev(SharedWavefunction ref_wfn, Options& options)
         oep->compute();
 
     }
+    // ===> Density Matrix Susceptibility Tensor Model <=== //
+    else if (o_task == "DMATPOL") 
+    {
+        std::shared_ptr<oepdev::GenEffParFactory> factory = oepdev::GenEffParFactory::build("POLARIZATION", ref_wfn, options);
+        std::shared_ptr<oepdev::GenEffPar> parameters = factory->compute();
+    } 
     // ==> Create Wavefunction Union of two monomers <==
     else if (o_task == "SOLVER")
     {

### 2018-05-25 pymodule.py

@@ -72,6 +72,7 @@ def run_oepdev(name, **kwargs):
 
     # case when OEP build is requested
     if psi4.core.get_global_option("OEPDEV_TARGET").startswith("OEP") \
+    or psi4.core.get_global_option("OEPDEV_TARGET").startswith("DMAT") \
     or psi4.core.get_global_option("OEPDEV_TARGET") == "TEST":
            
        basis_df_oep = psi4.core.BasisSet.build(molecule  , "BASIS", psi4.core.get_global_option("DF_BASIS_OEP"),

### 2018-05-28 papers/1_dmatpol/paper-r1.tex
Paper updated with sketch of introduction

@@
 
+Molecular\hyp{}level exploration of condensed phase phenomena becomes increasingly unavoidable
+as technology and medicine continue to develop. Unfortunately, the computational cost of accurate
+electronic structure calculations grows extremely fast with the number of particles, greatly limiting
+the use of the state\hyp{}of\hyp{}the\hyp{}art first\hyp{}principles theories 
+when large and complex chemical systems such as
+solutions and biomolecules come into play. Although the advent in the computer technology 
+brought some hopes, it is still insufficient to follow up the expensive needs.
+Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques that link
+extremely accurate but localized pieces of information with the heuristic knowledge of an atomistic organisation 
+on a relatively large spatial scales, typically ranging beyond few nanometers.
+
+A class of molecular dynamics simulation techniques was developed over the years. That was possible
+due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
+semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
+refinings. The advent of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
+technique was a revolution in computational chemistry and allowed insightful theoretical studies
+of quantum mechanical effects in biological systems because the high\hyp{}level 
+and costly quantum mechanical calculations
+could be combined with low\hyp{}level and cheap classical dynamics. 
+The size of quantum level can be extended without compromising accuracy and maintaining good performance
+by using vast family of fragmentation methods, such as effective fragment potentials (EFP).
+The most difficult to tackle part of Hamiltonian can also be embedded within a widely understood effective environment
+leading to various embedding methodologies, realising similar goals.
+
+However, even with the modern advances in computational chemistry, 
+%accurate description of environmental effects
+accurate description of phenomena due to complex heterogeneous molecular environments 
+remains a challenge in the following areas: (i) describing the electronically excited states 
+(electronic solvatochromism), (ii) understanding
+the vibrational properties of molecules in ground and electronically excited states
+(vibrational or vibronic solvatochromisms). EFP methods are limited to the ground\hyp{}state
+chemistry at the Hartree\hyp{}Fock (HF) or density functional theory (DFT) levels.
+It is also quite difficult to generalize the embedding potential operator
+because of complicated nature of intermolecular interactions, that theoretically
+cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
+Therefore, Hamiltonian embedding is often realised by Coulombic embedding with polarization
+of QM level, and only rarely the Pauli exclusion principle is taken into account. Although
+such approach is very accurate in certain cases, it is not guaranteed that it is sufficient
+for all cases.
+
+In this Work, we explore the susceptibility of the molecule's electronic density distribution
+to polarize due to external electrostatic perturbation from first principles. 
+The deformation density is expressed in terms
+of the density matrix susceptibility (DMS) tensors that are properties of isolated molecules and can be
+extracted from quantum chemistry calculations. The focus on the density (matrix) instead of the energy
+(scalar) is beneficial since it provides the directional tensor description
+of the polarization that could be combined with other effects such as Pauli repulsion or dispersion,
+in contrast to the standard interaction energy approaches. We show that accurate polarization 
+of the one\hyp{}particle density
+matrix is achievable and can be realized as long as the one\hyp{}particle density matrices
+are available. It is believed that the proposed method
+could be a basis for a construction of an 
+alternative approach in fragment\hyp{}based
+modelling of extended molecular aggregates that is not limitted to the ground state chemistry.
+We introduce the necessary theory and derive the \emph{ab initio} density matrix susceptibility model
+at the HF approximation in Section II. This model serves as a basis for generalization to the 
+arbitrary one\hyp{}electron density with electronic correlation which is derived in Section II. 
+The implementation details in a computer code as well as the validation procedures are detailed
+in Section III. We then discuss the performance of the models in Section IV based on a few small molecules
+in constant and varying electric field distributions. Finally, we conclude the presentation wit a few 
+remarks for a future development in Section V.
+
+%
+%Understanding intricate mechanisms that govern
+%the biomachinery of life requires thorough
+
 \section{\label{s:2}Theory}
@@
 the Hamiltonian operator. In the above, the superscript $(0)$
-denotes the reference, or unperturbed, state that is assumed to be known.
-
+denotes the reference or unperturbed state that is assumed to be known.
 Under the influence of an external electrostatic perturbation, the total electronic energy
@@
 is extremely difficult. This is not only because of the very complicated (and generally unknown) solution
-to the electronic correlation problem, but also due to the dimensionality of the problem: the quantitative
-model has to accurately predict all the $n^2$ elements of the density matrix. Even relatively small errors
+to the electronic correlation problem, but also because
+a working model has to accurately predict all the $n^2$ elements of the density matrix
+where $n$ is the size of basis set. Even relatively small errors
 in a particular matrix element could accumulate tremendously when computing traces with other matrices,
 and in turn, obstruct the evaluated properties with non\hyp{}negligible errors.
-Therefore, in this Work we adopt the simplistic approach: we first develop a qualitative first\hyp{}principles model 
+Therefore, we adopt the following approach: we first develop a qualitative first\hyp{}principles model 
 at the Hartree\hyp{}Fock (HF)
@@
 by considering HF level,
-but treating the model as a set of adjustable parameters further on.
+but treating the density matrix susceptibility tensors as cetrain functions of adjustable parameters further on.
 
@@
 %
+(In this work we shall use the special open bold font for matrices of vectors, so that
+in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
 Note that if the total charge is conserved, the induced moments are independent on the origin.
@@
 of atomic orbitals (AO's), referred here to as the LCAO-MO matrix.
-When ${\bf D}$ is expressed in orthogonal basis, the density matrix is idempotent, i.e.,
+${\bf D}$ is an idempotent matrix and if expressed in orthogonal basis the following holds
 %
@@
 It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
-induced dipole moment is independent on the origin. Therefore,
-in the above result and also in later course of the Paper the dipole origin has been omitted. 
+induced dipole moment is independent on the origin. 
+This quarantees that the theory fulfills the total charge conservation requirement.
+Therefore,
+in the above result and also in later course of the Paper the dipole origin will be omitted. 
 %where the dipole origin was omitted for notational clarity.
-If the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
+
+Now, if the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
 %
@@
 and has size $(3\times n)$. Note that the \emph{right inverse} of ${\bf L}_i$
-does not exist because ${\bf L}_i {\bf L}_i^T$ is singular.
+does not exist because ${\bf L}_i {\bf L}_i^T$ is singular due to the fact that usually
+$n > 3$.
 %Clearly, in the above approximation, ${\bf X}$ depends on the choice
@@
 
-The exact mathematical form of the generalized density matrix susceptibilities is unknown. 
-The first suscebility from Eq.~\eqref{e:final-model.HF} is valid only under the Hartree\hyp{}Fock
-approximation, with a guarantee that only the overall charge and the dipole moment, 
+The exact mathematical form of the generalized density matrix susceptibilities 
+from Eq.~\eqref{e:final-model.General} is unknown. 
+The susceptibility from Eq.~\eqref{e:final-model.HF} is valid only under the Hartree\hyp{}Fock
+approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
+limited by the flavour of the distributed polarizability approximation), 
 but not higher multipole moments, are correctly described.
@@
 of the three blocks of adjustable parameters
-that is subject to the least\hyp{}squares minimization:
+that is subject to the least\hyp{}squares minimization,
 %
@@
 %
-Quadrupole moments were subsequently transformed to the their traceless forms by
+Quadrupole moments were subsequently transformed to their traceless forms by
 %

### 2018-05-31 papers/1_dmatpol/paper-r1.tex
Change format of the paper, added abstract and summary sketch, added figure

@@
 
-\documentclass[aip,graphicx]{revtex4-1}
+\documentclass[aip,amsmath,amssymb,reprint]{revtex4-1}
 %\documentclass[aip,reprint]{revtex4-1}
+
+\usepackage{graphicx}% Include figure files
+\usepackage{dcolumn}% Align table columns on decimal point
+\usepackage{bm}% bold math
+%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
+%\linenumbers\relax % Commence numbering lines
 
@@
 \usepackage{amsbsy}
-\usepackage{bm}
+%\usepackage{bm}
 \usepackage{mathrsfs}
@@
 
-\draft % marks overfull lines with a black rule on the right
+%\draft % marks overfull lines with a black rule on the right
 
@@
 
-\title{Density Matrix Susceptibility} %Title of paper
+\preprint{AIP/123-DMS}
+
+\title{Density Matrix Susceptibility Tensors} %Title of paper
 
@@
 \homepage[]{https://www.polonez.pwr.edu.pl}
-\thanks{Bulak}
+%\thanks{Bulak}
 \affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, Wroc{\l}aw University of Technology, Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}
@@
 \begin{abstract}
-To be filled ...
+
+%Considering the solvent effect on the correlated electronic wavefunctions, especially
+%those that describe the electronically excited state, is a great challenge of Quantum Chemistry.
+
+In this work, the electric field\hyp{}induced change in the one\hyp{}electron density
+has been expressed as a series of the one\hyp{}particle density matrix susceptibility tensors
+interacting with electric fields. The analytical approximate expressions were derived
+at the Hartree\hyp{}Fock theory, which served as a basis for construction of generalized
+model that can be used for an arbitrary form of wavefunction as long as the one\hyp{}particle
+density matrices are available. It is shown that it is possible to accurately predict the changes
+in the one\hyp{}electron density of water molecule in spatially non\hyp{}uniform electric field distribution
+generetated by point charges. The interaction\hyp{}induced energies
+dipole moments and quadrupole moments
+are quantitatively described by the present theory in weak and moderate electric fields.
+The possibility of extension of the proposed model with high\hyp{}order susceptibilities
+opens new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
 \end{abstract}
@@
 %
-\begin{equation}\label{e:rho}
+\begin{equation}\label{e:rho-1}
  \delta \rho({\bf r}) = 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \phi_\alpha({\bf r}) \phi^*_\beta({\bf r}) \;.
@@
 %
-\begin{equation} \label{e:dmu-4-exact.linear-approximation}
+\begin{multline} \label{e:dmu-4-exact.linear-approximation}
  \frac{1}{2} 
@@
          {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
-    \right] }
+    \right] } \\
 %
@@
     \right] } \;,
-\end{equation}
+\end{multline}
 %
@@
     \right] }
-  &=
+  & \\ \nonumber =
 2 \Tr{ 
@@
 %
-\begin{equation}  \label{e:susceptibility-B}
- {\bf B}^{(i)}_{\alpha\beta} = \left\{
-%
-                               C_{\alpha i} {\bf b}_{\beta i} + C_{\beta i} {\bf b}_{\alpha i}
+\begin{multline}  \label{e:susceptibility-B}
+ {\bf B}^{(i)}_{\alpha\beta} = \Big\{
+                               C_{\alpha i} {\bf b}_{\beta i} + C_{\beta i} {\bf b}_{\alpha i} \\
                                 - \sum_\gamma 
@@
 %
-\right\}  \;.
-\end{equation}
+\Big\}  \;.
+\end{multline}
 %
@@
 %
-\begin{equation}\label{e:final-model.General}
- \delta D_{\alpha\beta} = \sum_{i }^M \left\{
+\begin{multline}\label{e:final-model.General}
+ \delta D_{\alpha\beta} = \sum_{i }^M \Big\{
                                       {\bf B}^{(i ; 10)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  
 %                       + \sum_{ij}^M {\bf B}^{(ij; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
-                        +             {\bf B}^{(i ; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
+                        +             {\bf B}^{(i ; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \\
                         +             {\bf B}^{(i ; 01)}_{\alpha\beta} : {\BM\nabla} \otimes {\bf F}({\bf r}_i) 
-                        + \ldots \right\}
-\end{equation}
+                        + \ldots \Big\}
+\end{multline}
 %
@@
 %
-\begin{equation}\label{e:final-model.General.Parameters}
- \delta D^{(N)} = \sum_{i }^M \left\{ 
-                              \sum_u^{x,y,z} S^{[1]}_{iu} F_{iu}^{(N)} 
-%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{jw} \right) F_{iu}^{(N)} F_{ju}^{(N)}
-                +             \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{iw} \right) F_{iu}^{(N)} F_{iu}^{(N)}
+\begin{multline}\label{e:final-model.General.Parameters}
+ \delta D^{(N)} = \sum_{i }^M \Big\{ 
+                              \sum_u^{x,y,z} S^{[1]}_{iu} F_{iu}^{(N)} \\
+%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
+                +             \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{iw} \right) F_{iu}^{(N)} F_{iw}^{(N)} \\
                 +             \sum_{uw}^{x,y,z} \left( S^{[3]}_{iu} + S^{[3]}_{iw}\right) F_{iu,iw}^{(N)}
-                        + \ldots \right\}
-\end{equation}
+                        + \ldots \Big\}
+\end{multline}
 %
@@
 \begin{multline}\label{e:dE}
- \delta E[{\bf F}] = 
+ \delta E[{\bf F}] = E_{\rm nuc}({\bf F}) +
                      \Tr{\left\{ \delta {\bf D} \cdot
-                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}}
+                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}} \\
                    + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
-                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} 
-                   + E_{\rm nuc}({\bf F}) \;.
+                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} \;.
 \end{multline}
@@
 %
-\begin{equation}\label{e:dE-dpol}
- \delta E[{\bf F}] \approx 2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} } 
-                   - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
-                   + E_{\rm nuc}({\bf F}) \;,
-\end{equation}
+\begin{multline}\label{e:dE-dpol}
+ \delta E[{\bf F}] \approx E_{\rm nuc}({\bf F}) + 
+                     2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} }  \\
+                   - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \;,
+\end{multline}
 %
@@
 
-\subsection{Ab Initio Model of Densty Matrix Polarization}
-
-We first test the accuracy of the density matrix susceptibility tensors that were derived
-in Eq.~\eqref{e:susceptibility-B} from first principles. 
-For this purpose we have considered a water molecule 
-under weak uniform electric fields (Figure~XXX). 
-3D difference one\hyp{}electron density isosurface plots for sample electric fields
-with respect to the unperturbed density (${\bf F} = {\bf 0}$)
-are shown in Figure~XXX. The statistical evaluation of errors for a set of
-100 random electric fields is shown in Figure~XXX. The shape of the basins 
-of decreased/increased density are comparable with the exact result
-obtained from HF/6-311++G** calculations. 
+Performance descriptors of the \emph{ab initio} and generalized 
+models of the density matrix susceptibility are shown in Table~XXX. 
+We considered three different types of point charge environment that generate
+weak (0.001 -- 0.005 a.u.), 
+moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
+electric fields (on average) around the water molecule.
+%
+%for a set of varying point charge environments
+%with weak, moderate and strong electric fields are shown in Table~XXX.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
+electric fields.}
+Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
+quadrupole moments (G-I).}
+\end{figure*}
+
+%We first test the accuracy of the density matrix susceptibility tensors that were derived
+%in Eq.~\eqref{e:susceptibility-B} from first principles. 
+%For this purpose we have considered a water molecule 
+%under weak uniform electric fields (Figure~XXX). 
+%3D difference one\hyp{}electron density isosurface plots for sample electric fields
+%with respect to the unperturbed density (${\bf F} = {\bf 0}$)
+%are shown in Figure~XXX. The statistical evaluation of errors for a set of
+%100 random electric fields is shown in Figure~XXX. The shape of the basins 
+%of decreased/increased density are comparable with the exact result
+%obtained from HF/6-311++G** calculations. 
 %It can be seen that the theory
 %describes the polarization of the electron density distribution qualitatively well:
-However, the accuracy in the density matrix elements is on overall very low which 
-manifests with drammatic errors in the electronic total energy of the system.
-It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
-which neglects quadrupole and higher\hyp{}order induced multipole moments.
-
-\subsection{General Model of Densty Matrix Polarization}
+%However, the accuracy in the density matrix elements is on overall very low which 
+%manifests with drammatic errors in the electronic total energy of the system.
+%It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
+%which neglects quadrupole and higher\hyp{}order induced multipole moments.
+
+%\subsection{General Model of Densty Matrix Polarization}
 
@@
 \section{\label{s:5}Summary and a few concluding remarks}
-Test summary bla bla blaaaaa dfn fdj gf ads;jha g;ja gf; p.
+
+In this Work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
+at the HF level was derived for a closed shell molecule in a spatially non\hyp{}uniform
+electric field distribution. It was found that the change in the density matrix element
+is approximately proportional to the electric field evaluated at the distributed site associated with the 
+center of charge of the 
+localized molecular orbital. Subsequently, the \emph{ab initio} model served as a basis
+for the formulation of the generalized model that is, in principle, not limited to the HF theory,
+and takes into account higher\hyp{}order curvatures of the electric field distribution as well.
+The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibility tensors
+that are able to interact with the electric field, square of electric field and electric field gradients.
+The particular density matrix susceptibilities are determined through the least\hyp{}squares 
+linear regression with respect to the enseble of electrostatically petrurbed known density matrices.
+
+The above theoretically developed models were then used to calculate the electric field\hyp{}induced
+interaction energies, dipole moments and quadrupole moments of water molecule that
+interacts with various point charge environments generating electric fields of weak, moderate
+and strong magnitudes. It was found that, while the \emph{ab initio} model yelds qualitatively correct
+interaction energies and dipole moments, the generalized model can reach quantitative level of accuracy
+for prediction of interaction energies, dipole moments and even quadrupole moments. Although 
+errors in interaction energies produced by using the generalized model 
+was still at best one order of magnitude larger than the errors produced by using the
+usual approach based on the distributed polarizabilities, this work demonstrated that it is possible to
+accurately predict the change in the density matrix of electronic system, and the accuracy of the model
+can be improved by extending the parameter space of the susceptibility tensors.
+
+It is anticipated here that the above generalized models from Eq.~\eqref{e:final-model.General}
+should be studied in the future with an emphasis 
+on their further optimization, especially to reach RMSE below 1 kcal/mole
+in strong and very strong electric fields. Owing the capability of the proposed approach to correctly capture the 
+environment\hyp{}induced changes in the one\hyp{}particle density matrix, 
+the generalized density matrix susceptibility tensors 
+could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
+described at correlated levels of theory, particularly including the excited state chemistry.
 
@@
 
-\section{\label{a:orig-dep} Origin-Dependence of Induced Dipole Moment}
+\section{\label{a:orig-dep} Origin-Independence of the Induced Dipole Moment}
 
 We need to be aware that multipole integrals depend on the origin with respect to which they are evaluated.
-However, the induced dipole moments defined in Eq.~\eqref{e:mu-ind-distributed-general} 
+However, the induced dipole moments defined in Eq.~\eqref{e:dmu-exact} 
 have to be origin independent. 
@@
 Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
-defined in Eq.~\eqref{e:mu-ind-distributed-general} 
+defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
 are origin independent.

### 2018-06-01 papers/1_dmatpol/paper-r1.tex
Table in paper initiated

@@
 %\usepackage{dcolumn}
-%\usepackage{multirow}
+\usepackage{multirow}
 %\usepackage{hhline}
@@
 Performance descriptors of the \emph{ab initio} and generalized 
-models of the density matrix susceptibility are shown in Table~XXX. 
+models of the density matrix susceptibility are shown in Table~\ref{t:results}. 
+%
+% Please add the following required packages to your document preamble:
+\begin{table*}[]
+\centering
+\caption{My caption}
+\label{my-label}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+\textbf{}                                                                           &              & \textbf{An Initio} & \multicolumn{4}{c|}{\textbf{Generalized}}                     & \textbf{Multipole} \\ \hline
+\textbf{Rank of DMS}                                                                & \textbf{Set} & \textbf{1, 0}      & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
+\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W            & 0.067 (99.80)      & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8}
+                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
+                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W            & 3.36 (98.70)       & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8}
+                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
+                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W            & 6.22 (58.72)       & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8}
+                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
+                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+\end{tabular}
+\end{table*}
 We considered three different types of point charge environment that generate
@@
 moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
-electric fields (on average) around the water molecule.
+electric fields around the water molecule (see the histograms of the electric field distribution
+per atomic centre in Figure~XXX).
 %
@@
 that are able to interact with the electric field, square of electric field and electric field gradients.
-The particular density matrix susceptibilities are determined through the least\hyp{}squares 
-linear regression with respect to the enseble of electrostatically petrurbed known density matrices.
+The particular density matrix susceptibilities were determined through the least\hyp{}squares 
+linear regression with respect to the training enseble of electrostatically petrurbed states.
 
@@
 errors in interaction energies produced by using the generalized model 
-was still at best one order of magnitude larger than the errors produced by using the
+were still at best one order of magnitude larger than the errors produced by using the
 usual approach based on the distributed polarizabilities, this work demonstrated that it is possible to
@@
 
-It is anticipated here that the above generalized models from Eq.~\eqref{e:final-model.General}
+It is anticipated here that the generalized models from Eq.~\eqref{e:final-model.General}
 should be studied in the future with an emphasis 
 on their further optimization, especially to reach RMSE below 1 kcal/mole
-in strong and very strong electric fields. Owing the capability of the proposed approach to correctly capture the 
+in strong and very strong electric fields (0.05-0.07 a.u.). 
+Owing the capability of the proposed approach to correctly capture the 
 environment\hyp{}induced changes in the one\hyp{}particle density matrix, 

### 2018-06-01 papers/1_dmatpol/paper-r1.tex
Table updated

@@
 %\usepackage{hhline}
-%\usepackage{booktabs}
+\usepackage{booktabs}
 %\usepackage{longtable}
@@
 %
-% Please add the following required packages to your document preamble:
 \begin{table*}[]
@@
 \caption{My caption}
-\label{my-label}
+\label{t:results}
 \begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-\textbf{}                                                                           &              & \textbf{An Initio} & \multicolumn{4}{c|}{\textbf{Generalized}}                     & \textbf{Multipole} \\ \hline
-\textbf{Rank of DMS}                                                                & \textbf{Set} & \textbf{1, 0}      & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
+\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
+\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
 \multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8}
-                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
-                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8}
+                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
 \multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8}
-                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
-                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8}
+                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
 \multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8}
-                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8}
-                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
+                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8}
+                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
 \end{tabular}
 \end{table*}
-We considered three different types of point charge environment that generate
-weak (0.001 -- 0.005 a.u.), 
-moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
-electric fields around the water molecule (see the histograms of the electric field distribution
-per atomic centre in Figure~XXX).
-%
-%for a set of varying point charge environments
-%with weak, moderate and strong electric fields are shown in Table~XXX.
-
-\begin{figure*}
-\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
-\caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
-electric fields.}
-Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
-quadrupole moments (G-I).}
-\end{figure*}
 

### 2018-06-01 papers/1_dmatpol/paper-r1.tex
Revisions in the table and appendix A

@@
 \label{t:results}
-\begin{tabular}{|l|c|c|c|c|c|c|c|}
-\hline
-\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
-\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
-\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8}
-                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8}
-                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
-\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8}
-                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8}
-                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
-\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8}
-                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8}
-                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
-\end{tabular}
+\begin{ruledtabular}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{cddcddddddddcdd}
+%\hline\hline
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{\textbf{An Initio DMS}} &  & \multicolumn{8}{c}{\textbf{Generalized DMS}}                                                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
+\cline{2-3}
+\cline{5-12}
+%\cline{14-15}
+\textbf{Set}         & \multicolumn{2}{c}{\textbf{1, 0}}          &  & \multicolumn{2}{c}{\textbf{1, 0}} & \multicolumn{2}{c}{\textbf{2, 0}} & \multicolumn{2}{c}{\textbf{1, 1}} & \multicolumn{2}{c}{\textbf{2, 1}} &  & \multicolumn{2}{c}{\textbf{Polarizability}}         \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}                                                                                                                                                                                             \\ \hline
+W                    & 0.067               & (99.80)              &  & 0.076          & (99.74)          & 0.053          & (99.88)          & 0.074          & (99.76)          & 0.050          & (99.89)          &  & 0.004            & (100.00)            \\
+M                    & 1.07                & (96.96)              &  & 1.24           & (95.96)          & 0.86           & (98.06)          & 1.20           & (96.17)          & 0.81           & (98.27)          &  & 0.07             & (99.99)             \\
+S                    & 4.22                & (88.87)              &  & 5.15           & (83.44)          & 3.60           & (91.92)          & 5.02           & (84.25)          & 3.41           & (92.73)          &  & 0.33             & (99.93)             \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                                                                                                                                                                                                   \\ \hline
+W                    & 3.36                & (98.70)              &  & 1.03           & (99.88)          & 0.98           & (99.89)          & 0.52           & (99.97)          & 0.47           & (99.98)          &  & 3.36             & (98.70)             \\
+M                    & 14.72               & (98.50)              &  & 7.72           & (99.59)          & 6.86           & (99.67)          & 6.15           & (99.74)          & 5.38           & (99.80)          &  & 14.72            & (98.50)             \\
+S                    & 44.80               & (96.92)              &  & 33.06          & (98.32)          & 29.49          & (98.67)          & 28.74          & (98.73)          & 26.25          & (98.94)          &  & 44.80            & (96.92)             \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}}                                                                                                                                                                                               \\ \hline
+W                    & 6.22                & (58.72)              &  & 1.59           & (97.32)          & 1.49           & (97.62)          & 0.58           & (99.64)          & 0.48           & (99.75)          &  & 5.45             & (68.30)             \\
+M                    & 28.22               & (53.90)              &  & 10.45          & (93.68)          & 8.83           & (95.49)          & 8.25           & (96.06)          & 6.57           & (97.50)          &  & 23.94            & (66.83)             \\
+S                    & 78.07               & (40.53)              &  & 43.33          & (81.68)          & 35.73          & (87.55)          & 38.80          & (85.31)          & 31.59          & (90.26)          &  & 64.33            & (59.62) \\
+\bottomrule
+\end{tabular}%
+}
+\end{ruledtabular}
 \end{table*}
-
+%
+We considered three different types of point charge environment that generate
+weak (0.001 -- 0.005 a.u.), 
+moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
+electric fields around the water molecule (see the histograms of the electric field distribution
+per atomic centre in Figure~XXX).
+-%
+-%for a set of varying point charge environments
+-%with weak, moderate and strong electric fields are shown in Table~XXX.
+\begin{figure*}
+\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
+electric fields.}
+Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
+quadrupole moments (G-I).}
+\end{figure*}
+%
+%\begin{table*}[]
+%\centering
+%\caption{My caption}
+%\label{t:results}
+%\resizebox{\textwidth}{!}{%
+%\begin{tabular}{|l|c|c|c|c|c|c|c|}
+%\hline
+%\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
+%\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
+%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8}
+%                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8}
+%                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
+%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8}
+%                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8}
+%                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
+%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8}
+%                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8}
+%                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
+%\end{tabular}%
+%}
+%\end{table*}
 %We first test the accuracy of the density matrix susceptibility tensors that were derived
@@
 \begin{equation}
- \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) 
- = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
+ \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
 \end{equation}
@@
 \begin{equation}
- \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0) 
- = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
+ \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
 \end{equation}
@@
 %
-\begin{equation}
-   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_0) 
- = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0)
+\begin{multline}
+   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) \\
  = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
  = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
-\end{equation}
+\end{multline}
 %

### 2018-06-02 papers/1_dmatpol/paper-r1.tex
Table added, revised equations, added discussion

@@
 
-\title{Density Matrix Susceptibility Tensors} %Title of paper
+\title{One-Particle Density Matrix Polarization Susceptibility Tensors}
 
@@
 associated with the $i$th MO whereas ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge
-${\bf r}_i = \tbraket{i}{\hat{\bf r}}{i}$.
+${\bf r}_i = \tbraket{\phi_i}{\hat{\bf r}}{\phi_i}$.
 We can now insert the above result into Eq.~\eqref{e:dmu-l-vector} 
@@
 %
-Hence, the 3D vector ${\bf B}^{(i)}_{\alpha\beta}$ describes the polarization affinity
+Hence, the three dimensional vector ${\bf B}^{(i)}_{\alpha\beta}$ describes the polarization affinity
 of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
@@
 appear in the first power, allowing for linear least\hyp{}squares regression.
-Note that $B_u^{(i,(10))} = S^{[1]}_{iu}$, $B_{uw}^{(i,(20))} = S^{[2]}_{iu} + S^{[2]}_{iw}$
-and $B_{uw}^{(i,(01))} = S^{[3]}_{iu} + S^{[3]}_{iw}$. The square bracket superscripts 
+Note that $B_u^{(i;10)} = S^{[1]}_{iu}$, $B_{uw}^{(i;20)} = S^{[2]}_{iu} + S^{[2]}_{iw}$
+and $B_{uw}^{(i;01)} = S^{[3]}_{iu} + S^{[3]}_{iw}$. The square bracket superscripts 
 denote the block of the parameter space. We chose symmetric forms of ${\bf B}^{(i;20)}$
@@
 
-To determine the optimim set 
+To determine the optimum set 
 $
@@
 and constant.
-Therefore, the exact solution for the optimal parameters is given by a Newton equation
+Therefore, the exact solution for the optimal parameters is given by the Newton equation
 %
@@
 %
-where ${\bf g}$ and ${\bf H}$ are the gradient vector and Hessian matrix, respectively.
+where ${\bf g}$ and ${\bf H}$ are the gradient vector and the Hessian matrix, respectively.
 %
@@
 
-The theory that is necessary to validate the density matrix susceptibility model from Section~\ref{s:2}
-was implemented in our in\hyp{}house plugin to Psi4 quantum chemistry program.
-A few simple molecules were selected as test systems, that were subject to 
-perturbations due to uniform electric fields as well as collections of point charges
+The theory that is necessary to validate the density matrix susceptibility models from Section~\ref{s:2}
+was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.
+Water molecule, energy\hyp{}optimized by using the HF/6-311++G** method, 
+was selected as a test system which was subject to 
+electrostatic perturbations due to random collections of point charges
 generating non\hyp{}uniform electric fields. 
+Cartesian coordinates of perturbing charges were drawn from uniform distribution of a set of fourty
+three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
+and origin in the center of mass of water molecule.
+In selection of random points, space inside the union of spheres surrounding
+oxygen and hydrogen atoms with radii of 4.5 and 4.0 Bohrs, respectively, was
+kept free from any point charges due to potential problems with the convergence of the 
+self\hyp{}consistent field (SCF) calculations.
+Charge values were selected from the one\hyp{}dimensional uniform distribution
+of double precision numbers in the interval of $\vert q_{\rm max} \vert$, where $q_{\rm max}$
+was tuned to obtain a desired ensemble of weak, moderate and strong electric field magnitudes
+(with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly; cf. Table~\ref{t:fields}).
+
 %For finding a general unitary transformations,
@@
 %were calculated.
-
 In order to validate all the models discussed in this Contribution,
@@
 in which ${\BM \alpha}_i$ is the distributed anisotropic polarizability tensor associated with the
-$i$th localized molecular orbital (LMO) centroid, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
+$i$th localized molecular orbital (LMO) centroid (see Eq.~\eqref{e:dmu-dpol}). 
+%, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
 The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
@@
 change in the one\hyp{}particle density matrix, i.e., the polarization of the electron density distribution.
-
-The electric field\hyp{}induced dipole and quadrupole moments were evaluated from the following expressions:
+%This is because we focus in this Work on the pure polarization process. Therefore, comparison between
+%the results obtained by using Eq.~\eqref{e:dE-dpol} and Eq.~\eqref{e:dE} will reveal
+
+The electric field\hyp{}induced dipole and primitive quadrupole moments were evaluated from the following expressions:
 %
@@
    \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
-   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{uw}{\beta} \;. 
+   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{(u-u_{\rm CM})(w-w_{\rm CM})}{\beta} \;. 
   \end{align}
@@
 %
-Quadrupole moments were subsequently transformed to their traceless forms by
+where the subscript CM denotes the center of mass of water molecule.
+When the distributed polarizabilities were used instead of the polarization\hyp{}induced 
+difference density matrices, primitive induced quadrupole moments were computed from
+the induced dipole moments according to
+%
+\begin{equation}
+  \delta {\BM \Theta} = \sum_i^{\rm occ} 
+   \left\{ 
+       \delta {\BM \upmu}_i \otimes \left( {\bf r}_i - {\bf r}_{\rm CM} \right) 
+      + \left( {\bf r}_i - {\bf r}_{\rm CM} \right) \otimes \delta {\BM \upmu}_i
+   \right\} \;.
+\end{equation}
+%
+%where ${\bf r}_{\rm CM}$ is the center of mass of water molecule.
+
+Primitive quadrupole moments were subsequently transformed to their traceless forms by
 %
@@
 
+Three different types of point charge environment that generate
+weak (0.003 -- 0.007 a.u.), 
+moderate (0.01 -- 0.02 a.u.) and strong (0.02 -- 0.05 a.u.) 
+electric fields around the water molecule were considered (Table~\ref{t:fields}),
+because similar ranges of electric field are experienced by water molecules
+in liquid phase.\cite{Reischl.Kofinger.Dellago.MolPhys.2009,Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
+%
+\begin{table}%[]
+%\centering
+\caption{{\bf Average electric fields in statistical sets of electrostatically perturbed states
+of water molecule.} Each set was composed of 100 samples differing in the configuration
+of 40 charges generated from a uniform distribution.
+In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
+respectively.}
+\label{t:fields}
+\begin{ruledtabular}
+\begin{tabular}{ldcdcdcd}
+\multirow{2}{*}{\textbf{Set}} & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
+                              & \multicolumn{3}{c}{\textbf{Oxygen Atom}} & \textbf{} 
+                              & \multicolumn{3}{c}{\textbf{Hydrogen Atom}} \\
+\cline{2-4}
+\cline{6-8}
+\textbf{W}                    & 0.0046     & $\pm$     & 0.0019     &           & 0.0046     & $\pm$     & 0.0020     \\
+\textbf{M}                    & 0.0184     & $\pm$     & 0.0075     &           & 0.0186     & $\pm$     & 0.0078     \\
+\textbf{S}                    & 0.0368     & $\pm$     & 0.0150     &           & 0.0372     & $\pm$     & 0.0156    
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+%
 Performance descriptors of the \emph{ab initio} and generalized 
-models of the density matrix susceptibility are shown in Table~\ref{t:results}. 
-%
-\begin{table*}[]
-\centering
-\caption{My caption}
+models of the density matrix susceptibility, as well as the conventional distributed polarizability
+approach, are shown in Table~\ref{t:results}. 
+%
+\begin{table*}%[]
+%\centering
+\caption{{\bf Performance of the density matrix susceptibility tensors of water molecule 
+in weak, moderate and strong electric fields.}
+In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
+respectively. $R$-squared coefficients are shown in parentheses.
+Ranks of the density matrix susceptibility models are denoted as ($R_{\rm max}, P_{\rm max}$),
+where $R_{\rm max}$ and $P_{\rm max}$ are the maximum ranks of the electric field 
+and the electric field gradient in the expansion
+from Eq.~\eqref{e:final-model.General}, respectively.
+}
 \label{t:results}
@@
 %\cline{14-15}
-\textbf{Set}         & \multicolumn{2}{c}{\textbf{1, 0}}          &  & \multicolumn{2}{c}{\textbf{1, 0}} & \multicolumn{2}{c}{\textbf{2, 0}} & \multicolumn{2}{c}{\textbf{1, 1}} & \multicolumn{2}{c}{\textbf{2, 1}} &  & \multicolumn{2}{c}{\textbf{Polarizability}}         \\ \hline
+\textbf{Set}         
+& \multicolumn{2}{c}{\textbf{1, 0}}          
+&  
+& \multicolumn{2}{c}{\textbf{1, 0}} 
+& \multicolumn{2}{c}{\textbf{2, 0}} 
+& \multicolumn{2}{c}{\textbf{1, 1}} 
+& \multicolumn{2}{c}{\textbf{2, 1}} 
+&  
+& \multicolumn{2}{c}{\textbf{Polarizability}}         \\ \hline
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}                                                                                                                                                                                             \\ \hline
@@
 %
-We considered three different types of point charge environment that generate
-weak (0.001 -- 0.005 a.u.), 
-moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
-electric fields around the water molecule (see the histograms of the electric field distribution
-per atomic centre in Figure~XXX).
--%
--%for a set of varying point charge environments
--%with weak, moderate and strong electric fields are shown in Table~XXX.
-\begin{figure*}
+
+The \emph{ab initio} DMS model quantitatively reproduces the interaction energy and the dipole moment 
+in weak electric field (RMSE of 0.067~kcal/mol and RMSD of 3.4~mD). In stronger electric fields
+the predictions become less quantitative, with high RMSE of 4.2~kcal/mol and RMSD of 14.7~mD in strong electric fields.
+Induced quadrupole moment is described by the \emph{ab initio} DMS qualitatively well with $R^2$ not exceeding
+60\%. Generalized DMS model of the same rank ($R_{\rm max}$=1, $P_{\rm max}$=0) performs slightly worse
+for interaction energies but better for induced multipole moments, particularly decreasing the errors of induced quadrupole
+calculations almost by a factor of two in strong electric fields, and by a factor of three in weaker fields.
+Extending the parameter space improves the accuracy of the generalized DMS model, on overall. 
+Including DMS of type ${\bf B}^{(20)}$, i.e., that interact with the square of electric field,
+leads to improvement in prediction of all quantities evaluated in all electric field strengths considered in this Work. 
+Interestingly, interaction energies are somewhat insensitive to addition of the quadrupolar DMS 
+(${\bf B}^{(01)}$, interacting with the electric field gradient), in contrast to induced multipole moments
+which exhibit noticeable improvement of accuracy. The most accurate generalized DMS model presented here 
+($R_{\rm max}$=2, $P_{\rm max}$=1, which corresponds to including ${\bf B}^{(10)}$, ${\bf B}^{(20)}$ and ${\bf B}^{(01)}$)
+works quantitatively well in weak and medium electric fields with $R^2$ being systematically 97.5\% and higher,
+and with RMSE of 0.8~kcal/mol in moderate fields. 
+However, the errors in the interaction energy
+in strong fields are 3.4~kcal/mol, which is quite high as compared to the usually acceptable accuracy 
+of 1~kcal/mol in routine quantum chemistry calculations. 
+Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
+polarizable forcefield, that the average electric field experienced by water molecule in liquid phase
+in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
+This value corresponds to our moderate electric field set of samples (denoted as `M' in Tables~\ref{t:fields}
+and \ref{t:results}) for which the
+accuracy of the generalized DMS models is good 
+(see also error distributions shown in Figure~\ref{f:results}).
+Nevertheless, water molecules can experience
+stronger electric fields in certain geometries of hydrogen bonded clusters, 
+even up to 0.055 a.u. according to study by Reischl et~al.\cite{Reischl.Kofinger.Dellago.MolPhys.2009} 
+Therefore, in strong electric fields
+the present state of generalized DMS's must be further optimized in order to reach accuracy of at least
+1~kcal/mol.
+
+It is also instructive to compare the DMS models with the conventional approach based on the distributed
+dipole\hyp{}dipole
+polarizabilities. In terms of accuracy in interaction energy prediction, DMS models are on average one order 
+of magnitude less accurate than the distributed polarizability approach, which provides errors of 0.3~kcal/mol
+or less in all cases. On the other hand, DMS models better predict the induced dipole and quadrupole 
+moments even in weak fields. Certainly, taking into account the distributed quadrupole polarizabilities
+would greatly reduce errors of the standard multipole approach. However, it is believed that
+the DMS approach, although less (but still sufficiently) accurate in terms of interaction energy, 
+has an important advantage over the multipole\hyp{}based approach, i.e., it provides detailed
+tensor description of the spatial polarization of the electron density distribution.
+
+\begin{figure*}[h]
 \includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
@@
 We need to be aware that multipole integrals depend on the origin with respect to which they are evaluated.
-However, the induced dipole moments defined in Eq.~\eqref{e:dmu-exact} 
+However, the induced dipole moments defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
+(or equivalently in Eq.~\eqref{e:dmu-l-vector})
 have to be origin independent. 
@@
 \begin{equation}
- \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
+ \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
 \end{equation}
@@
 \begin{equation}
- \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
+ \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
 \end{equation}
@@
 \begin{multline}
-   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) \\
- = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
- = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
+   \left[ \widetilde{\mathbb{L}} \right]_{i\alpha} ({\bf r}_Q) 
+ = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
+ - \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) \\
+ = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0})   
+ - \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0})
+ = \left[ \widetilde{\mathbb{L}} \right]_{i\alpha} ({\bf 0}) \;.
 \end{multline}
@@
 %
-The gradient element of the $r$th block and Hessian element of the $rs$th block read
+The gradient element of the $r$th block and Hessian element of the $(rs)$th block read
 %

### 2018-06-03 papers/1_dmatpol/paper-r1.tex
Revision

@@
 
-\documentclass[aip,amsmath,amssymb,reprint]{revtex4-1}
+\documentclass[aip,amsmath,amssymb,reprint,floatfix]{revtex4-1}
 %\documentclass[aip,reprint]{revtex4-1}
@@
 solutions and biomolecules come into play. Although the advent in the computer technology 
-brought some hopes, it is still insufficient to follow up the expensive needs.
-Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques that link
-extremely accurate but localized pieces of information with the heuristic knowledge of an atomistic organisation 
+brought some hopes, it is still insufficient to follow up the expensive needs
+of computational chemistry.
+Due to this reason, it has been of particular importance nowadays 
+to develop bridging theoretical techniques that link
+extremely accurate but localized pieces of information 
+with the heuristic knowledge of an atomistic organisation and intermolecular interaction
 on a relatively large spatial scales, typically ranging beyond few nanometers.
 
-A class of molecular dynamics simulation techniques was developed over the years. That was possible
-due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
-semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
-refinings. The advent of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
+The advent of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
 technique was a revolution in computational chemistry and allowed insightful theoretical studies
@@
 could be combined with low\hyp{}level and cheap classical dynamics.
-The size of quantum level can be extended without compromising accuracy and maintaining good performance
-by using vast family of fragmentation methods, such as effective fragment potentials (EFP).
-The most difficult to tackle part of Hamiltonian can also be embedded within a widely understood effective environment
-leading to various embedding methodologies, realising similar goals.
+That was possible due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
+semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
+refinings.
+Understanding physics that rules intermolecular interactions helped to develop
+a wide family of force fields, and served a basis for embedding Hamiltonian
+within an effective environment. 
+%The size of quantum level can be extended without compromising accuracy and maintaining good performance
+%by using vast family of fragmentation methods, such as effective fragment potentials (EFP).
+%The most difficult to tackle part of Hamiltonian can also be embedded within a widely understood effective environment
+%leading to various embedding methodologies, realising similar goals.
 
@@
 the vibrational properties of molecules in ground and electronically excited states
-(vibrational or vibronic solvatochromisms). EFP methods are limited to the ground\hyp{}state
-chemistry at the Hartree\hyp{}Fock (HF) or density functional theory (DFT) levels.
-It is also quite difficult to generalize the embedding potential operator
+(vibrational or vibronic solvatochromisms). 
+Interface between QM and MM levels is usually realized by electrostatic embedding
+and only rarely includes the Pauli exclusion principle and dispersion. 
+Unfortunately, it is quite difficult to generalize the embedding potential operator
 because of complicated nature of intermolecular interactions, that theoretically
 cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
-Therefore, Hamiltonian embedding is often realised by Coulombic embedding with polarization
-of QM level, and only rarely the Pauli exclusion principle is taken into account. Although
-such approach is very accurate in certain cases, it is not guaranteed that it is sufficient
-for all cases.
-
+Perhaps the most frequently used efficient and accurate \emph{ab initio} fragmentation methods,
+the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
+chemistry at the Hartree\hyp{}Fock (HF) or density functional theory (DFT) levels,
+and, while they can approximately describe the environment composed of molecules in their ground states,
+they cannot be used for highly correlated wavefunctions such as electronically excited states.
+%Therefore, Hamiltonian embedding is often realised by Coulombic embedding with polarization
+%of QM level, and only rarely the Pauli exclusion principle is taken into account. Although
+%such approach is very accurate in certain cases, it is not guaranteed that it is sufficient
+%for all cases.
+
+Most of efforts to theoretically treat extended molecular aggregates are rooted in 
+energy\hyp{}based philosophy. However,
+in the past decade there has been considerable interest in density\hyp{}based approaches
+of a wide scope. In particular, the importance of
+one\hyp{}electron densities in chemistry is well known and acknowledged in the community,
+though still not solved exactly.\cite{Holas.March.PhysRevA.1991}
+%that seem to have an important advantage over the standard energetic approach.
+This Work challenges in establishing a connection bewteen the one\hyp{}particle
+density and the perturbation in the environment through introducing the
+density matrix susceptibility (DMS) tensors.
+The associated change in the one\hyp{}particle density matrix, $\delta {\bf D}$, is postulated to be
+deconvoluted into a product form of the perturbation susceptibility ${\bf B}_i$
+and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$.
+%and induced dipole moments $\delta {\BM\upmu}_i$ due to electric fields ${\bf E}_i$ 
+%at the centroids of the occupied molecular orbitals,
+%
+\begin{equation*} \label{e:initial-model}
+ \delta {\bf D} \approx \sum_i^{\text{Sites}} \text{[DMS of the $i$th site]} 
+  \cdot  \delta {\bf P}_i
+  %\rightarrow \sum_i^{\rm occ} {\bf B}^{(i)}_{\mu\nu} \cdot {\bf E}_i \;.
+\end{equation*}
+%
+While $\delta {\bf P}_i$ contain critical information 
+about the actual distortion of the electronic cloud,
+perturbation susceptibility is assumed to be a property of unperturbed 
+molecule and could be precomputed only once. 
+This could be adapted into an efficient effective 
+one\hyp{}electron fragment potential, in principle applicable to an arbitrary
+level of theory, as long as the one\hyp{}particle densities are available.
+%
+%a simple first\hyp{}principles model for prediction 
+%of the perturbation\hyp{}induced one\hyp{}electron density
+%change of structurally unrelaxed molecule is proposed here.
+
+The simplest, yet perhaps the most relevant for chemistry, type of perturbation 
+is electrostatic perturbation.
+It is worth 
+mentioning that, at that moment, there is no simple \emph{ab initio}
+method of computation of electron deformation densities without considering the entire molecular 
+aggregate.\cite{Horn.Yuezhi.Head-Gordon.JCP.2016,Horn.Head-Gordon.JCP.2015}
 In this Work, we explore the susceptibility of the molecule's electronic density distribution
@@
 The deformation density is expressed in terms
-of the density matrix susceptibility (DMS) tensors that are properties of isolated molecules and can be
+of the one\hyp{}particle density matrix polarization susceptibility tensors 
+that are properties of isolated molecules and can be
 extracted from quantum chemistry calculations. The focus on the density (matrix) instead of the energy
@@
 in Section III. We then discuss the performance of the models in Section IV based on a few small molecules
-in constant and varying electric field distributions. Finally, we conclude the presentation wit a few 
-remarks for a future development in Section V.
+in varying electric field distributions. Finally, in Section V we conclude the presentation with a few 
+remarks for future development.
 
@@
 where $\BM\Delta$ is a square non\hyp{}singular and non\hyp{}symmetric matrix.
-McWeeny has shown that
+McWeeny showed that
 %
@@
  {\bf l}_{i\alpha} \equiv \left[ \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \right]_{i\alpha} 
-      = \left[  {\bf C}^\dagger \mathbb{M} \left[ \bf{1} - \bf{D}^{(0)} \right] \right]_{i\alpha}
+      = \left[  {\bf C}^\dagger \mathbb{M} \left( \bf{1} - \bf{D}^{(0)} \right) \right]_{i\alpha}
 \end{equation}
@@
 The theory that is necessary to validate the density matrix susceptibility models from Section~\ref{s:2}
-was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.
-Water molecule, energy\hyp{}optimized by using the HF/6-311++G** method, 
+was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
+Water molecule, energy\hyp{}optimized by using the HF/6-311++G** 
+method\cite{Roothaan.RevModPhys.1951,Krishnan.Binkley.Seeger.Pople.JCP.1980}, 
 was selected as a test system which was subject to 
@@
 The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
-method. Orbital localization was performed by using the Boys method. Note that in Eq.~\eqref{e:dE-dpol}, 
+method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
+Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960} 
+Note that in Eq.~\eqref{e:dE-dpol}, 
 multipole approximation is introduced only for the part associated with the electric field\hyp{}induced
@@
 %where ${\bf r}_{\rm CM}$ is the center of mass of water molecule.
-
 Primitive quadrupole moments were subsequently transformed to their traceless forms by
@@
 %
+Methodology for the calculation of the electric field and electric field gradient due to point charges
+can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
+
 To compute the interaction energy and the induced multipole moments, either exact or approximated forms
@@
 %
-Throughout the Work, the Hartree\hyp{}Fock level was assumed and the 6-311++G** basis set
+Throughout the Work, the Hartree\hyp{}Fock level\cite{Roothaan.RevModPhys.1951} was assumed 
+and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
 was used. For all the calculations of the density matrix polarization susceptibility tensors
 from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
-atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme.
+atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme.\cite{Mayer.IJQC.2002}
 
@@
 has an important advantage over the multipole\hyp{}based approach, i.e., it provides detailed
-tensor description of the spatial polarization of the electron density distribution.
+tensor description of the spatial polarization of the electron density distribution, that could be
+combined with density\hyp{}based methods of interaction energy calculations.\cite{Mandado.Hermida-Ramon.JCTC.2011}
 

### 2018-07-19 papers/1_dmatpol/paper-r1.tex
Revision.

@@
 in the one\hyp{}electron density of water molecule in spatially non\hyp{}uniform electric field distribution
-generetated by point charges. The interaction\hyp{}induced energies
+generetated by point charges. The electric field\hyp{}induced interaction energies,
 dipole moments and quadrupole moments
-are quantitatively described by the present theory in weak and moderate electric fields.
-The possibility of extension of the proposed model with high\hyp{}order susceptibilities
-opens new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
+are quantitatively described by the present theory in weak and moderate electric fields (0.003 -- 0.02 a.u.).
+%It is believed that the possibility of extension of the proposed model with high\hyp{}order susceptibilities
+%opens new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
+It is believed that the proposed model, after extensive development on its optimization, could open
+new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
 \end{abstract}
@@
 when large and complex chemical systems such as
-solutions and biomolecules come into play. Although the advent in the computer technology 
-brought some hopes, it is still insufficient to follow up the expensive needs
-of computational chemistry.
+solutions and biomolecules come into play.\cite{Tomasi.Mennucci.Cammi.ChemRev.2005}
+%Although the advent in the computer technology 
+%brought some hopes, it is still insufficient to follow up the expensive needs
+%of computational chemistry.
 Due to this reason, it has been of particular importance nowadays 
@@
 extremely accurate but localized pieces of information 
-with the heuristic knowledge of an atomistic organisation and intermolecular interaction
+with certain knowledge of an atomistic organisation and intermolecular interaction
 on a relatively large spatial scales, typically ranging beyond few nanometers.
 
-The advent of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
+The discovery of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
 technique was a revolution in computational chemistry and allowed insightful theoretical studies
@@
 and costly quantum mechanical calculations
-could be combined with low\hyp{}level and cheap classical dynamics.
-That was possible due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
-semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
-refinings.
-Understanding physics that rules intermolecular interactions helped to develop
+could be combined with low\hyp{}level and cheap classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
+%That was possible due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
+%semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
+%refinings.
+Understanding physics that rules intermolecular interactions\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
+helped to develop
 a wide family of force fields, and served a basis for embedding Hamiltonian
-within an effective environment. 
+within an effective environment and fragmentation methods.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
 %The size of quantum level can be extended without compromising accuracy and maintaining good performance
@@
 remains a challenge in the following areas: (i) describing the electronically excited states 
-(electronic solvatochromism), (ii) understanding
-the vibrational properties of molecules in ground and electronically excited states
-(vibrational or vibronic solvatochromisms). 
+(electronic solvatochromism),\cite{Barbati.JACS.2014,
+Szabla.Sponer.Jiri.Gora.JPCL.2015,
+Bednarska.Zalesny.Tian.Murugan.Agren.Bartkowiak.Molecules.2017,
+Jedrzejewska.Grabarz.Bartkowiak.Osmialowski.SpectChimActA.2018} 
+(ii) understanding the vibrational properties of molecules in ground and electronically excited states
+(vibrational or vibronic solvatochromisms).\cite{Blasiak.Londergan.Webb.Cho.ACR.2017} 
 Interface between QM and MM levels is usually realized by electrostatic embedding
-and only rarely includes the Pauli exclusion principle and dispersion. 
-Unfortunately, it is quite difficult to generalize the embedding potential operator
+and only rarely includes the Pauli exclusion principle and dispersion.\cite{List.Olsen.Kongsted.PCCP.2016}
+Unfortunately, it is quite difficult to efficiently treat the environment\hyp{}induced potential operator
 because of complicated nature of intermolecular interactions, that theoretically
@@
 the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
-chemistry at the Hartree\hyp{}Fock (HF) or density functional theory (DFT) levels,
+chemistry at the Hartree\hyp{}Fock (HF)\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
+or density functional theory (DFT) levels\cite{Nguyen.Pachter.Day.JCP.2014},
 and, while they can approximately describe the environment composed of molecules in their ground states,
@@
 Most of efforts to theoretically treat extended molecular aggregates are rooted in 
-energy\hyp{}based philosophy. However,
+energy\hyp{} (or force field\hyp{})based philosophy. However,
 in the past decade there has been considerable interest in density\hyp{}based approaches
-of a wide scope. In particular, the importance of
+of a wide scope and aims.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
+Mandado.Hermida-Ramon.JCTC.2011,
+Sun.Chan.ACR.2016,
+Hedegard.Reiher.JCTC.2016} 
+In particular, the importance of
 one\hyp{}electron densities in chemistry is well known and acknowledged in the community,
@@
 %that seem to have an important advantage over the standard energetic approach.
-This Work challenges in establishing a connection bewteen the one\hyp{}particle
+With connection to the above, it would be incredibly insightful 
+to establish a first\hyp{}principles connection bewteen the one\hyp{}particle
 density and the perturbation in the environment through introducing the
-density matrix susceptibility (DMS) tensors.
-The associated change in the one\hyp{}particle density matrix, $\delta {\bf D}$, is postulated to be
-deconvoluted into a product form of the perturbation susceptibility ${\bf B}_i$
-and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$.
+\emph{density matrix susceptibility} (DMS) tensors. % because knowing the density distribution
+%is equivalent to knowing the state of the molecule.
+The main postulate of the present Work is that the associated change 
+in the one\hyp{}particle density matrix, $\delta {\bf D}$, can be
+deconvoluted into a product form of the generalized perturbation susceptibility ${\bf B}_i$
+and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$,
+evaluated at certain site $i$.
 %and induced dipole moments $\delta {\BM\upmu}_i$ due to electric fields ${\bf E}_i$ 
@@
 perturbation susceptibility is assumed to be a property of unperturbed 
-molecule and could be precomputed only once. 
-This could be adapted into an efficient effective 
+molecule and is to be precomputed only once. 
+Thus, such DMS could be adapted into an efficient effective 
 one\hyp{}electron fragment potential, in principle applicable to an arbitrary
 level of theory, as long as the one\hyp{}particle densities are available.
+It also means that such a technique would dramatically reduce costs of quantum chemistry calculations
+becuase the task of performing full computation on a molecular aggregate
+would reduce to much simpler matrix operations.
 %
@@
 The simplest, yet perhaps the most relevant for chemistry, type of perturbation 
-is electrostatic perturbation.
-It is worth 
-mentioning that, at that moment, there is no simple \emph{ab initio}
+is the electrostatic perturbation.
+It is worth mentioning that, at that moment, there is no simple \emph{ab initio}
 method of computation of electron deformation densities without considering the entire molecular 
-aggregate.\cite{Horn.Yuezhi.Head-Gordon.JCP.2016,Horn.Head-Gordon.JCP.2015}
+aggregate by using full quantum chemistry calculations.\cite{Horn.Yuezhi.Head-Gordon.JCP.2016,Horn.Head-Gordon.JCP.2015}
 In this Work, we explore the susceptibility of the molecule's electronic density distribution
@@
 (scalar) is beneficial since it provides the directional tensor description
-of the polarization that could be combined with other effects such as Pauli repulsion or dispersion,
+of the polarization that could be combined with other effects such as Pauli repulsion 
+or dispersion,\cite{Mandado.Hermida-Ramon.JCTC.2011}
 in contrast to the standard interaction energy approaches. We show that accurate polarization 
@@
 modelling of extended molecular aggregates that is not limitted to the ground state chemistry.
+
 We introduce the necessary theory and derive the \emph{ab initio} density matrix susceptibility model
-at the HF approximation in Section II. This model serves as a basis for generalization to the 
-arbitrary one\hyp{}electron density with electronic correlation which is derived in Section II. 
+at the HF approximation\cite{Roothaan.RevModPhys.1951} 
+in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalization to the 
+arbitrary one\hyp{}electron density with electronic correlation which is derived in Section~\ref{s:2}. 
 The implementation details in a computer code as well as the validation procedures are detailed
-in Section III. We then discuss the performance of the models in Section IV based on a few small molecules
-in varying electric field distributions. Finally, in Section V we conclude the presentation with a few 
+in Section~\ref{s:3}. We then discuss the performance of the models in Section~\ref{s:4} 
+%based on a few small molecules
+%in varying electric field distributions. 
+Finally, in Section~\ref{s:5} we conclude the presentation with a few 
 remarks for future development.
@@
 
-The electronic energy of a closed\hyp{}shell molecule $A$ is given by the time\hyp{}independent Schr{\"o}dinger equation,
+The electronic energy of a closed\hyp{}shell molecule is given by the time\hyp{}independent Schr{\"o}dinger equation,
 %
@@
 \begin{equation}\label{e:m}
- \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = \tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
-\end{equation}
-%
-(In this work we shall use the special open bold font for matrices of vectors, so that
+ \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = -e\tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
+\end{equation}
+%
+(In this Work we shall use the special open bold font for matrices of vectors, so that
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
@@
 
-In the Hartree\hyp{}Fock theory, ${\bf D}$ is approximated as
+In the Hartree\hyp{}Fock theory,\cite{Roothaan.RevModPhys.1951} ${\bf D}$ is approximated as
 %
@@
 induced dipole moment is independent on the origin. 
-This quarantees that the theory fulfills the total charge conservation requirement.
+This guarantees that the theory fulfills the total charge conservation requirement.
 Therefore,
@@
 and ${\bf B}^{(i; 01)}_{\alpha\beta}$ are $(3\times 3)$ matrices. Also, the distributed expansion centres
-(labeled as $i$) are generalized to any point in the molecule.
+(labeled as $i$) are generalized to any point in the molecule.\footnote{More general
+contribution due to square of electric field could be $\sum_{ij}^M {\bf B}^{(ij; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$.
+However, for simplicity we neglect the contributions from different sites $i$ and $j$
+in the present discussion and leave it for future development.
+}
 
@@
 (with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly; cf. Table~\ref{t:fields}).
+In this way, 100 random samples of 40 point charges were selected and exact solutions of HF\hyp{}SCF equations were 
+performed for each validation procedure.
 
@@
 The particular density matrix susceptibilities were determined through the least\hyp{}squares 
-linear regression with respect to the training enseble of electrostatically petrurbed states.
+linear regression with respect to the training enseble of electrostatically perturbed states.
 
@@
 in strong and very strong electric fields (0.05-0.07 a.u.). 
-Owing the capability of the proposed approach to correctly capture the 
+For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
+sites should be further investigated, preferably with post\hyp{}Hartree\hyp{}Fock methods.
+Nevertheless, owing the capability of the proposed approach to correctly capture the 
 environment\hyp{}induced changes in the one\hyp{}particle density matrix, 

### 2018-07-19 oepdev/libgefp/gefp_polar_base.cc

@@ -496,20 +496,36 @@ void oepdev::GeneralizedPolarGEFactory::compute_statistics(void) {
 
         // ---> Compute interaction energy <--- // nuclear part is already computed in void compute_samples(void)
 
-        double eint_refer = H->vector_dot(referenceStatisticalSet_.DensityMatrixSet[n]);
-        double eint_model = H->vector_dot(    modelStatisticalSet_.DensityMatrixSet[n]);
+        //double eint_refer = H->vector_dot(referenceStatisticalSet_.DensityMatrixSet[n]);
+        //double eint_model = H->vector_dot(    modelStatisticalSet_.DensityMatrixSet[n]);
 
-        std::shared_ptr<psi::Matrix> Dn1= D->clone(); Dn1->add(referenceStatisticalSet_.DensityMatrixSet[n]);
-        std::shared_ptr<psi::Matrix> Dn2= D->clone(); Dn2->add(    modelStatisticalSet_.DensityMatrixSet[n]);
-        std::shared_ptr<psi::Matrix> G = D->clone(); G->zero();
-        G->add(VMatrixSet_[n]); G->scale(2.0);
-        std::shared_ptr<psi::Matrix> G1 = G->clone(); G1->add(referenceStatisticalSet_.JKMatrixSet[n]);
-        std::shared_ptr<psi::Matrix> G2 = G->clone(); G2->add(    modelStatisticalSet_.JKMatrixSet[n]);
-        eint_refer+= G1->vector_dot(Dn1);
-        eint_model+= G2->vector_dot(Dn2);
+        //std::shared_ptr<psi::Matrix> Dn1= D->clone(); Dn1->add(referenceStatisticalSet_.DensityMatrixSet[n]);
+        //std::shared_ptr<psi::Matrix> Dn2= D->clone(); Dn2->add(    modelStatisticalSet_.DensityMatrixSet[n]);
+        //std::shared_ptr<psi::Matrix> G = D->clone(); G->zero();
+        //G->add(VMatrixSet_[n]); G->scale(2.0);
+        //double eint_coul = G->vector_dot(D);
+        //std::shared_ptr<psi::Matrix> G1 = G->clone(); G1->add(referenceStatisticalSet_.JKMatrixSet[n]);
+        //std::shared_ptr<psi::Matrix> G2 = G->clone(); G2->add(    modelStatisticalSet_.JKMatrixSet[n]);
+        //eint_refer+= G1->vector_dot(Dn1);
+        //eint_model+= G2->vector_dot(Dn2);
 
[…]

### 2018-07-20 oepdev/libgefp/gefp.cc
Second-order finite-field DMS implemented and works well.

@@ -100,6 +100,9 @@ std::shared_ptr<psi::Matrix> oepdev::GenEffPar::compute_density_matrix(std::vect
         //cout << densityMatrixDipolePolarizability_[n][0]->get(0,0) << endl;
         //cout << densityMatrixDipolePolarizability_[n][1]->get(0,0) << endl;
         //cout << densityMatrixDipolePolarizability_[n][2]->get(0,0) << endl;
+        //densityMatrixDipolePolarizability_[n][0]->print();
+        //densityMatrixDipolePolarizability_[n][1]->print();
+        //densityMatrixDipolePolarizability_[n][2]->print();
         if (hasDensityMatrixDipoleDipoleHyperpolarizability_) {
             //cout << densityMatrixDipoleDipoleHyperpolarizability_[n][0]->get(0,0) << endl;
             //cout << densityMatrixDipoleDipoleHyperpolarizability_[n][4]->get(0,0) << endl;
@@ -110,6 +113,15 @@ std::shared_ptr<psi::Matrix> oepdev::GenEffPar::compute_density_matrix(std::vect
             //cout << densityMatrixDipoleDipoleHyperpolarizability_[n][6]->get(0,0) << endl;
             //cout << densityMatrixDipoleDipoleHyperpolarizability_[n][5]->get(0,0) << endl;
             //cout << densityMatrixDipoleDipoleHyperpolarizability_[n][7]->get(0,0) << endl;
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][0]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][1]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][2]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][3]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][4]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][5]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][6]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][7]->print();
+            //densityMatrixDipoleDipoleHyperpolarizability_[n][8]->print();
             D->axpy(fx*fx, densityMatrixDipoleDipoleHyperpolarizability_[n][0]);
[…]

### 2018-07-20 oepdev/libgefp/gefp_polar_base.cc

@@ -361,7 +361,11 @@ void oepdev::GeneralizedPolarGEFactory::allocate(void)
   if (hasDipolePolarizability_           ) PolarizationSusceptibilities_->allocate(1, 0, nSites_, nbf_);
   if (hasDipoleDipoleHyperpolarizability_) PolarizationSusceptibilities_->allocate(2, 0, nSites_, nbf_);
   if (hasQuadrupolePolarizability_       ) PolarizationSusceptibilities_->allocate(0, 1, nSites_, nbf_);
-  if (hasAbInitioDipolePolarizability_   ) abInitioPolarizationSusceptibilities_->allocate(1, 0, nSitesAbInitio_, nbf_);
+  if (hasAbInitioDipolePolarizability_   ) {
+     abInitioPolarizationSusceptibilities_->allocate(1, 0, nSitesAbInitio_, nbf_);
+     if ((options_.get_int("DMATPOL_FIELD_RANK")==2) && (options_.get_bool("DMATPOL_FF_AB_INITIO")))
+     abInitioPolarizationSusceptibilities_->allocate(2, 0, nSitesAbInitio_, nbf_);
+  }
 
   // Ab Initio Model
   if (hasAbInitioDipolePolarizability_   ) compute_ab_initio();
@@ -402,6 +406,8 @@ void oepdev::GeneralizedPolarGEFactory::compute_ab_initio(void) {
   }
   std::shared_ptr<oepdev::GenEffPar> parameters = factory->compute();
   abInitioPolarizationSusceptibilities_->set_dipole_polarizability(parameters->dipole_polarizability());
+  if ((options_.get_int("DMATPOL_FIELD_RANK")==2) && options_.get_bool("DMATPOL_FF_AB_INITIO"))
+  abInitioPolarizationSusceptibilities_->set_dipole_dipole_hyperpolarizability(parameters->dipole_dipole_hyperpolarizability());
   abInitioPolarizationSusceptibilities_->set_centres(parameters->centres());
   abInitioPolarizationSusceptibilitiesFactory_ = factory;
 }

### 2018-07-20 oepdev/libgefp/gefp_polar_ffabinitio.cc

@@ -49,15 +49,15 @@ std::shared_ptr<oepdev::GenEffPar> oepdev::FFAbInitioPolarGEFactory::compute()
        Dmatset.push_back(perturbed_state(Fields[n])->Da());
   }
  
-  // Compute The Density Matrix Susceptibility Tensors 
+  // Compute The Density Matrix Dipole Susceptibility Tensors 
   const double h = 1.0/(2.0 * s);
 
-  std::vector<std::vector<std::shared_ptr<psi::Matrix>>> densityMatrixSusceptibility;
+  std::vector<std::vector<std::shared_ptr<psi::Matrix>>> densityMatrixDipolePolarizability;
   std::vector<std::shared_ptr<psi::Matrix>> B;
 
-  std::shared_ptr<psi::Matrix> Bx = matrixFactory->create_shared_matrix("B[x] Orthogonal AO Basis"); 
-  std::shared_ptr<psi::Matrix> By = matrixFactory->create_shared_matrix("B[y] Orthogonal AO Basis"); 
-  std::shared_ptr<psi::Matrix> Bz = matrixFactory->create_shared_matrix("B[z] Orthogonal AO Basis"); 
+  std::shared_ptr<psi::Matrix> Bx = matrixFactory->create_shared_matrix("B[x] Non-Orthogonal AO Basis"); 
+  std::shared_ptr<psi::Matrix> By = matrixFactory->create_shared_matrix("B[y] Non-Orthogonal AO Basis"); 
+  std::shared_ptr<psi::Matrix> Bz = matrixFactory->create_shared_matrix("B[z] Non-Orthogonal AO Basis"); 
 
   Bx->copy(Dmatset[0]); Bx->subtract(Dmatset[1]); Bx->scale(h);
   By->copy(Dmatset[2]); By->subtract(Dmatset[3]); By->scale(h);
@@ -67,12 +67,84 @@ std::shared_ptr<oepdev::GenEffPar> oepdev::FFAbInitioPolarGEFactory::compute()
   B.push_back(By);
   B.push_back(Bz);
 
-  densityMatrixSusceptibility.push_back(B);
-  cout << " Density Matrix Susceptibility computed!\n";
+  densityMatrixDipolePolarizability.push_back(B);
+  cout << " Density Matrix Dipole Polarizability computed!\n";
+
+  // Compute The Density Matrix Dipole-Dipole Hyperpolarizability Tensors 
[…]

### 2018-07-20 papers/1_dmatpol/do-r2.sh
Second iteration of paper draft initiated.

@@ -1,0 +1,6 @@
+#!/bin/bash
+base=paper-r2
+
+latex  ${base}.tex
+bibtex ${base}.aux
+dvipdf ${base}.dvi

### 2018-07-22 papers/1_dmatpol/paper-r2.tex
Second iteration of paper draft initiated.
New figure added to the paper.

--- 2018-07-19 papers/1_dmatpol/paper-r1.tex
@@
 In this work, the electric field\hyp{}induced change in the one\hyp{}electron density
-has been expressed as a series of the one\hyp{}particle density matrix susceptibility tensors
+has been expressed as a series of the one\hyp{}particle density matrix susceptibilities
 interacting with electric fields. The analytical approximate expressions were derived
@@
 
-The discovery of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
+The development of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
 technique was a revolution in computational chemistry and allowed insightful theoretical studies
@@
 helped to develop
-a wide family of force fields, and served a basis for embedding Hamiltonian
+a wide family of force fields, and served as a basis for embedding Hamiltonian
 within an effective environment and fragmentation methods.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
@@
 cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
-Perhaps the most frequently used efficient and accurate \emph{ab initio} fragmentation methods,
+Perhaps the most frequently used efficient and reasonably accurate \emph{ab initio} fragmentation methods,
 the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
 chemistry at the Hartree\hyp{}Fock (HF)\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
-or density functional theory (DFT) levels\cite{Nguyen.Pachter.Day.JCP.2014},
-and, while they can approximately describe the environment composed of molecules in their ground states,
+or density functional theory (DFT) levels\cite{Nguyen.Pachter.Day.JCP.2014}.
+Unfortunately, while they can approximately describe the environment composed of molecules in their ground states,
 they cannot be used for highly correlated wavefunctions such as electronically excited states.
@@
 %is equivalent to knowing the state of the molecule.
-The main postulate of the present Work is that the associated change 
+The main postulate of the present work is that the associated change 
 in the one\hyp{}particle density matrix, $\delta {\bf D}$, can be
-deconvoluted into a product form of the generalized perturbation susceptibility ${\bf B}_i$
+represented in a product form of the generalized perturbation susceptibility ${\bf B}_i$
 and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$,
@@
 %
-While $\delta {\bf P}_i$ contain critical information 
+While $\delta {\bf P}_i$ contains critical information 
 about the actual distortion of the electronic cloud,
@@
 aggregate by using full quantum chemistry calculations.\cite{Horn.Yuezhi.Head-Gordon.JCP.2016,Horn.Head-Gordon.JCP.2015}
-In this Work, we explore the susceptibility of the molecule's electronic density distribution
+In this work, we explore the susceptibility of the molecule's electronic density distribution
 to polarize due to external electrostatic perturbation from first principles. 
@@
 %
-Ideally, the electric field\hyp{}induced change in the ground state's one\hyp{}particle density 
+The electric field\hyp{}induced change in the ground state's one\hyp{}particle density 
 follows its energy minimum of the perturbed state. In addition, all the perturbed 
@@
 with the exact energy\hyp{}minimized ground state. 
-
-Development of a first\hyp{}principles theory
+In the limit of a weak and spatially uniform electric field,
+the polarization\hyp{}induced density matrix change is given by the follwoing Taylor expansion
+%
+\begin{equation}\label{e:dD-Taylor}
+ \delta D_{\alpha\beta} = \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}}  \cdot {\bf F} 
+   + \frac{1}{2} 
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} : {\bf F} \otimes {\bf F}
+   + \ldots 
+\end{equation}
+%
+In the above expression, the density matrix susceptibilities are simply equal to the derivatives of the density matrix
+with respect to electric field (up to a multiplicative constant).
+
+In a general case, i.e., when electric field is non\hyp{}uniform, the development of a first\hyp{}principles theory
 that is capable of accurately capturing the electric field\hyp{}dependence of the one\hyp{}particle density
-is extremely difficult. This is not only because of the very complicated (and generally unknown) solution
+becomes extremely difficult. This is not only because of the very complicated (and generally unknown) solution
 to the electronic correlation problem, but also because
@@
 where $n$ is the size of basis set. Even relatively small errors
-in a particular matrix element could accumulate tremendously when computing traces with other matrices,
+in a particular matrix element could accumulate tremendously when performing two\hyp{}
+and four\hyp{}index contractions,
 and in turn, obstruct the evaluated properties with non\hyp{}negligible errors.
-Therefore, we adopt the following approach: we first develop a qualitative first\hyp{}principles model 
-at the Hartree\hyp{}Fock (HF)
+Single\hyp{}centered expansions, like the one from Eq.~\eqref{e:dD-Taylor}, although valid for 
+an arbitrary level of theory,
+are probably not sufficient enough to accurately describe the local polarization
+of electron density distribution when the electric fields are highly non\hyp{}uniform in space.
+Due to all of these reasons, we adopt the following approach: 
+we first develop a qualitative first\hyp{}principles distributed site model at the Hartree\hyp{}Fock (HF)
 level of theory, and then generalize it to the arbitrary level using the functional form that was found 
@@
 ${\bf C}^{(0)\dagger}{\bf C}^{(0)}={\bf 1}$,
-which will be the case throughout the entire work, unless stated differently.
+which will be the case throughout the entire work, unless stated otherwise.
 Substituting the result from Eq.~\eqref{e:dD-1}
@@
 Therefore,
-in the above result and also in later course of the Paper the dipole origin will be omitted. 
+in the above result and also in later course of the paper the dipole origin will be omitted. 
 %where the dipole origin was omitted for notational clarity.
@@
 was selected as a test system which was subject to 
-electrostatic perturbations due to random collections of point charges
+electrostatic perturbations due to randomly selected uniform electric fields and random collections of point charges
 generating non\hyp{}uniform electric fields. 
-Cartesian coordinates of perturbing charges were drawn from uniform distribution of a set of fourty
+Components of spatially uniform electric fields were determined 
+from the uniform distribution in the interval of $\vert 0.02 \vert$ a.u.
+generating average field of 0.035 a.u. in an ensemble of 100 samples.
+Cartesian coordinates of perturbing charges were drawn from the uniform distribution of a set of fourty
 three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
@@
 self\hyp{}consistent field (SCF) calculations.
-Charge values were selected from the one\hyp{}dimensional uniform distribution
+Electric charge values were selected from the one\hyp{}dimensional uniform distribution
 of double precision numbers in the interval of $\vert q_{\rm max} \vert$, where $q_{\rm max}$
@@
 (with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly; cf. Table~\ref{t:fields}).
-In this way, 100 random samples of 40 point charges were selected and exact solutions of HF\hyp{}SCF equations were 
+In this way, independent sets of 100 random samples of uniform electric fields as well as 
+40 point charges were selected and exact solutions of HF\hyp{}SCF equations were 
 performed for each validation procedure.
@@
 %were calculated.
-In order to validate all the models discussed in this Contribution,
-the electric field\hyp{}induced interaction energy, 
+In order to validate all the models discussed in this contribution,
+the electric field\hyp{}induced polarization energy, 
 the dipole moment and the quadrupole moment were analysed in detail.
@@
 \begin{multline}\label{e:dE}
- \delta E[{\bf F}] = E_{\rm nuc}({\bf F}) +
-                     \Tr{\left\{ \delta {\bf D} \cdot
-                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}} \\
-                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
-                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} \;.
+ \delta E^{\rm Pol}[{\bf F}] = 
+                     \Tr{\left\{ {\bf D}^{(0)} \cdot
+                                \left( 2\delta {\bf J} - \delta {\bf K}  \right) \right\}} \\
+                   + \Tr{\left\{ \delta {\bf D} \cdot
+                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} + 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} \;.
 \end{multline}
@@
 and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
-the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
-due to the nuclei.
+the polarization of the electron density distribution, respectively. 
+%$E_{\rm nuc}({\bf F})$ denotes the interaction energy due to the nuclei.
 In order to compare the performance of the distributed polarizability approach,
@@
 %
-\begin{multline}\label{e:dE-dpol}
- \delta E[{\bf F}] \approx E_{\rm nuc}({\bf F}) + 
-                     2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} }  \\
+\begin{equation}\label{e:dE-dpol}
+ \delta E^{\rm Pol}[{\bf F}] \approx %E_{\rm nuc}({\bf F}) + 
+                     %2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} }  \\
                    - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \;,
-\end{multline}
+\end{equation}
 %
@@
 Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
-Note that in Eq.~\eqref{e:dE-dpol}, 
-multipole approximation is introduced only for the part associated with the electric field\hyp{}induced
-change in the one\hyp{}particle density matrix, i.e., the polarization of the electron density distribution.
-%This is because we focus in this Work on the pure polarization process. Therefore, comparison between
-%the results obtained by using Eq.~\eqref{e:dE-dpol} and Eq.~\eqref{e:dE} will reveal
 
@@
 %
-where the subscript CM denotes the center of mass of water molecule.
+where the subscript `CM' denotes the center of mass of water molecule.
 When the distributed polarizabilities were used instead of the polarization\hyp{}induced 
@@
 
+\subsection{\label{ss:41}Water molecule in spatially uniform electric field}
+
+It is instructive to start first from the discussion of the polarization under the uniform electric field
+condition because, in this case, it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
+directly by numerical differentiation. Here we consider DMS up to second order, i.e.,
+%
+\begin{subequations}\label{e:exact-DMS}
+ \begin{align}
+  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}}   \\
+  {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} 
+ \end{align}
+\end{subequations}
+%
+and the associated error analysis is shown in Figure~\ref{f:fig-1}.
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\caption{\label{f:fig-1} A figure caption. The figure captions are
+automatically numbered.}
+\end{figure}
+
+\subsection{\label{ss:42}Water molecule in spatially non\hyp{}uniform electric field}
+
 Three different types of point charge environment that generate
@@
 \caption{{\bf Average electric fields in statistical sets of electrostatically perturbed states
-of water molecule.} Each set was composed of 100 samples differing in the configuration
+of water molecule surrounded by point charges.} Each set was composed of 100 samples differing in the configuration
 of 40 charges generated from a uniform distribution.
@@
 calculations almost by a factor of two in strong electric fields, and by a factor of three in weaker fields.
-Extending the parameter space improves the accuracy of the generalized DMS model, on overall. 
+By and large, extending the parameter space improves the accuracy of the generalized DMS model. 
 Including DMS of type ${\bf B}^{(20)}$, i.e., that interact with the square of electric field,
@@
 Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
-polarizable forcefield, that the average electric field experienced by water molecule in liquid phase
+polarizable force field, that the average electric field experienced by water molecule in liquid phase
 in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
@@
 dipole\hyp{}dipole
-polarizabilities. In terms of accuracy in interaction energy prediction, DMS models are on average one order 
+polarizabilities. In terms of accuracy in interaction energy prediction, DMS models are, on average, one order 
 of magnitude less accurate than the distributed polarizability approach, which provides errors of 0.3~kcal/mol
@@
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-2.eps}
 \caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
-electric fields.}
+non\hyp{}uniform electric fields.}
 Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
-quadrupole moments (G-I).}
+quadrupole moments (G-I). Comparison with the distributed polarizability model is also shown in this Figure
+in panels C, F and I.}
 \end{figure*}

### 2018-07-22 papers/1_dmatpol/paper-r2.tex
Revision in second iteration: Notation regularized and added subsection in the discussion

@@
 The implementation details in a computer code as well as the validation procedures are detailed
-in Section~\ref{s:3}. We then discuss the performance of the models in Section~\ref{s:4} 
-%based on a few small molecules
-%in varying electric field distributions. 
+in Section~\ref{s:3}. We then discuss the performance of the models in Section~\ref{s:4}.
 Finally, in Section~\ref{s:5} we conclude the presentation with a few 
 remarks for future development.
-
-%
-%Understanding intricate mechanisms that govern
-%the biomachinery of life requires thorough
 
@@
 with the exact energy\hyp{}minimized ground state.
+
 In the limit of a weak and spatially uniform electric field,
-the polarization\hyp{}induced density matrix change is given by the follwoing Taylor expansion
+the polarization\hyp{}induced density matrix change is given by the follwoing Taylor expansion:
 %
@@
 with respect to electric field (up to a multiplicative constant).
-
 In a general case, i.e., when electric field is non\hyp{}uniform, the development of a first\hyp{}principles theory
@@
 are probably not sufficient enough to accurately describe the local polarization
-of electron density distribution when the electric fields are highly non\hyp{}uniform in space.
+of electron density distribution when electric fields are highly non\hyp{}uniform in space.
 Due to all of these reasons, we adopt the following approach: 
-we first develop a qualitative first\hyp{}principles distributed site model at the Hartree\hyp{}Fock (HF)
+we first develop a qualitative first\hyp{}principles distributed\hyp{}site model at the Hartree\hyp{}Fock (HF)
 level of theory, and then generalize it to the arbitrary level using the functional form that was found 
@@
          \left[ 
-              \delta {\bf D} {\mathbb{M}({\bf r}_Q)}
+              \delta {\bf D} \cdot {\mathbb{M}({\bf r}_Q)}
          \right] } \;.
@@
 \begin{equation}
- {\bf D} = {\bf C}{\bf C}^\dagger \;,
+ {\bf D} = {\bf C} \cdot {\bf C}^\dagger \;,
 \end{equation}
@@
 \begin{equation} \label{e:dmat-change.exact}
- \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right]
-                                   \left[ {\bf 1} + {\bf v}^\dagger{\bf v} \right]^{-1}
+ \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
+                                   \left[ {\bf 1} + {\bf v}^\dagger \cdot {\bf v} \right]^{-1} \cdot
                                    \left[ {\bf D}^{(0)} + {\bf v}^\dagger \right] \;,
@@
 \begin{equation}
- {\bf v} \equiv \left[ {\bf 1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)}  \;.
+ {\bf v} \equiv \left[ {\bf 1} - {\bf D}^{(0)} \right] \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  \;.
 \end{equation}
@@
  \begin{align}
- \delta {\bf D} &\cong \left[ {\bf 1} - {\bf D}^{(0)} \right] {\BM\Delta} {\bf D}^{(0)} + 
-                        {\bf D}^{(0)} {\BM\Delta}^\dagger \left[ {\bf 1} - {\bf D}^{(0)} \right]  
+ \delta {\bf D} &\cong \left[ {\bf 1} - {\bf D}^{(0)} \right] \cdot {\BM\Delta} \cdot {\bf D}^{(0)} + 
+                        {\bf D}^{(0)} \cdot {\BM\Delta}^\dagger \cdot \left[ {\bf 1} - {\bf D}^{(0)} \right]  
  \label{e:dD-1} \\  &= 
-  \delta {\bf C}  {\bf C}^{(0)\dagger} + {\bf C}^{(0)}\delta {\bf C}^\dagger
-           - {\bf D}^{(0)} \delta {\bf C}  {\bf C}^{{(0)}\dagger} - {\bf C}^{(0)}\delta {\bf C}^\dagger {\bf D}^{(0)} 
+  \delta {\bf C} \cdot {\bf C}^{(0)\dagger} + {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \nonumber \\
+           &\qquad - {\bf D}^{(0)} \cdot \delta {\bf C} \cdot {\bf C}^{{(0)}\dagger} 
+                   - {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \cdot {\bf D}^{(0)} 
  \label{e:dD-2}
@@
 The latter equality is true only if the basis set functions are orthogonal, i.e., 
-${\bf C}^{(0)\dagger}{\bf C}^{(0)}={\bf 1}$,
+${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$,
 which will be the case throughout the entire work, unless stated otherwise.
@@
     \left[ 
-         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
+         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
     \right] }
@@
     \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
+         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
     \right] } \\
@@
     \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
+         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
     \right] }
@@
     \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
+         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
     \right] } \;,
@@
 Now, if the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
-%
-\begin{subequations}
- \begin{align}
-  \Tr{ 
-    \left[ 
-         {\mathbb{M}} {\BM\Delta} {\bf D}^{(0)}  
-    \right] }
-%
-  +\Tr{ 
-    \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger}
-    \right] }
-  & \\ \nonumber =
+%%
+%\begin{subequations}
+% \begin{align}
+%  \Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
+%    \right] }  &
+%%
+%  \nonumber\\ 
+%  +\Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
+%    \right] }
+% &=
+%2 \Tr{ 
+%    \left[ 
+%         \widetilde{\mathbb{M}} \cdot \delta {\bf C}
+%   \right] }  \;, 
+%%
+%\\
+%%
+%  \Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
+%    \right] }  &
+%%
+%   \nonumber\\
+% +\Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
+%    \right] }
+% &= 
+%2 \Tr{ 
+%    \left[ 
+%         \widetilde{\mathbb{K}} \cdot \delta {\bf C}
+%   \right] } \;.
+%%
+% \end{align}
+%\end{subequations}
+%%
+%
+\begin{equation}
+ \frac{1}{2} 
+ \delta {\BM{\upmu}}
+  \cong
 2 \Tr{ 
@@
          \widetilde{\mathbb{M}} \cdot \delta {\bf C}
-   \right] }  \;,
-%
-\\
-%
-  \Tr{ 
-    \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta} {\bf D}^{(0)}
    \right] }  
-%
- +\Tr{ 
-    \left[ 
-         {\mathbb{M}} {\bf D}^{(0)} {\BM\Delta}^{\dagger} {\bf D}^{(0)}
-    \right] }
-  & \\ \nonumber =
+-
 2 \Tr{ 
@@
    \right] } \;.
-%
- \end{align}
-\end{subequations}
+\end{equation}
 %
@@
  \begin{align}
-   \widetilde{\mathbb{M}}  &\equiv {\bf C}^\dagger {\mathbb{M}}     \;,           \\
-   \widetilde{\mathbb{K}}  &\equiv {\bf C}^\dagger {\mathbb{M}} {\bf D}^{(0)} \;, \\
+   \widetilde{\mathbb{M}}  &\equiv {\bf C}^{(0)\dagger} \cdot {\mathbb{M}}     \;,           \\
+   \widetilde{\mathbb{K}}  &\equiv {\bf C}^{(0)\dagger} \cdot {\mathbb{M}} \cdot {\bf D}^{(0)} \;, \\
    \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;.
@@
  {\bf l}_{i\alpha} \equiv \left[ \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \right]_{i\alpha} 
-      = \left[  {\bf C}^\dagger \mathbb{M} \left( \bf{1} - \bf{D}^{(0)} \right) \right]_{i\alpha}
+      = \left[  {\bf C}^\dagger \cdot \mathbb{M} \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \right]_{i\alpha}
 \end{equation}
@@
       \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
-       \left[ {\bf L}_i^T {\bf L}_i \right]^{-1} {\bf L}_i^T 
+       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T 
 \end{equation}
@@
 and has size $(3\times n)$. Note that the \emph{right inverse} of ${\bf L}_i$
-does not exist because ${\bf L}_i {\bf L}_i^T$ is singular due to the fact that usually
+does not exist because ${\bf L}_i \cdot {\bf L}_i^T$ is singular due to the fact that usually
 $n > 3$.
@@
 induced dipole moment of the molecule.
-The change in the LCAO\hyp{}MO coefficient is parametereized as
+The change in the LCAO\hyp{}MO coefficient is parameterized as
 %
 \begin{equation} \label{e:dC-dmatpol}
- \delta C_{\alpha i} = {\bf b}_{\alpha i} \cdot {\bf F}({\bf r}_i)  \;,
+ \delta C_{\alpha i} = {\bf b}_{\alpha i}^{(10)} \cdot {\bf F}({\bf r}_i)  \;,
 \end{equation}
@@
 \begin{equation} \label{e:susceptibility-b}
-  b_{\alpha i;w} = \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
+  b_{\alpha i;w}^{(10)}= \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
    \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  
@@
 %
-for $w=x,y,z$. The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
+for $w=x,y,z$. 
+The superscript `$(10)$' denotes the fact that the susceptibility introduced here
+is first\hyp{}order with respect to the electric field and zeroth\hyp{}order with respect to
+the electric field gradient.
+The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
 the expression for $\delta {\bf D}$ from Eq.~\eqref{e:dD-2} to finally give
@@
 \begin{equation}\label{e:final-model.HF}
- \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
+ \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i;10)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
 \end{equation}
@@
 \begin{multline}  \label{e:susceptibility-B}
- {\bf B}^{(i)}_{\alpha\beta} = \Big\{
-                               C_{\alpha i} {\bf b}_{\beta i} + C_{\beta i} {\bf b}_{\alpha i} \\
+ {\bf B}^{(i;10)}_{\alpha\beta} = \Big\{
+                               C_{\alpha i} {\bf b}_{\beta i}^{(10)} + C_{\beta i} {\bf b}_{\alpha i}^{(10)} \\
                                 - \sum_\gamma 
@@
                                  \right)
-                                           {\bf b}_{\gamma i}
+                                           {\bf b}_{\gamma i}^{(10)}
 %
@@
 %
-Hence, the three dimensional vector ${\bf B}^{(i)}_{\alpha\beta}$ describes the polarization affinity
+Hence, the three dimensional vector ${\bf B}^{(i;10)}_{\alpha\beta}$ describes the polarization affinity
 of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
-It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i)}$
+It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i;10)}$
 are unique properties of electronically unperturbed molecule.
@@
 pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
-vector susceptibility, different for each density matrix element.
+vector susceptibility, different for each density matrix element. This feature will be used to
+generalize the model to an arbitrary level of theory and electric field curvature in the next section.
 
@@
 limited by the flavour of the distributed polarizability approximation), 
-but not higher multipole moments, are correctly described.
+but not higher multipole moments and polarization energy, are correctly described.
 Our strategy is to treat the susceptibilities from Eq.~\eqref{e:final-model.General}
@@
 was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-Water molecule, energy\hyp{}optimized by using the HF/6-311++G** 
-method\cite{Roothaan.RevModPhys.1951,Krishnan.Binkley.Seeger.Pople.JCP.1980}, 
+Water molecule
+%, energy\hyp{}optimized by using the HF/6-311++G** 
+%method\cite{Roothaan.RevModPhys.1951,Krishnan.Binkley.Seeger.Pople.JCP.1980}, 
 was selected as a test system which was subject to 
-electrostatic perturbations due to randomly selected uniform electric fields and random collections of point charges
+electrostatic perturbations due to random uniform electric fields and random collections of point charges
 generating non\hyp{}uniform electric fields. 
@@
 the dipole moment and the quadrupole moment were analysed in detail.
-The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
+The electric field\hyp{}induced polarization energy at the HF level, $\delta E^{\rm Pol}$,
 was computed from the following exact formula:
@@
 To compute the interaction energy and the induced multipole moments, either exact or approximated forms
-of the difference density matrix from Eq.~\eqref{e:final-model.General}
+of the difference density matrix from Eq.~\eqref{e:dD-Taylor}, Eq.~\eqref{e:final-model.HF} or Eq.~\eqref{e:final-model.General},
 were used for a particular electric field distribution. The statistical evaluation of
@@
    \begin{equation}\label{e:rmse}
-     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
+     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 } \;;
    \end{equation}
@@
    \begin{equation}\label{e:rmsd}
-     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
+     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 } \;;
    \end{equation}
@@
    \begin{equation}\label{e:rmsq}
-     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta \widetilde{\BM\Theta}^{(N)} - \overline{\delta \widetilde{\BM\Theta}^{(N)}}\rvert^2 }
+     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta \widetilde{\BM\Theta}^{(N)} - \overline{\delta \widetilde{\BM\Theta}^{(N)}}\rvert^2 } \;.
    \end{equation}
@@
 The first describes the ability of the model to reproduce the exact energetics of the system, 
-wherease the latter two contain crucial information on how well the deformation of the one\hyp{}electron
+whereas the latter two contain crucial information on how well the deformation of the one\hyp{}electron
 density is described. For Eqs.~\eqref{e:rmsd} and \eqref{e:rmsq} the absolute magnitudes
@@
 %
-Throughout the Work, the Hartree\hyp{}Fock level\cite{Roothaan.RevModPhys.1951} was assumed 
+Throughout the work, the Hartree\hyp{}Fock level\cite{Roothaan.RevModPhys.1951} was assumed 
 and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
-was used. For all the calculations of the density matrix polarization susceptibility tensors
+was used. Before all the calculations, water molecule was energy\hyp{}optimized.
+For all the calculations of the density matrix polarization susceptibility tensors
 from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
-atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme.\cite{Mayer.IJQC.2002}
+atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme,\cite{Mayer.IJQC.2002}
+and further back\hyp{}transformed to the non\hyp{}orthogonal AO basis.
+To compute numerical first and second derivatives of the one\hyp{}particle density matrix
+with respect to the electric field from Eq.~\eqref{e:dD-Taylor}, 
+the following finite\hyp{}difference formulae were utilized:
+%
+\begin{subequations}\label{e:ff}
+  \begin{align}
+    f_u    &= \frac{f(h_u) - f(-h_u)}{2h_u} \;,\\
+    f_{uu} &= \frac{f(h_u) + f(-h_u) - 2f(0)}{h_u^2} \;,\\
+    f_{uw} &= \frac{f(h_u,h_w) + f(-h_u,-h_w) + 2f(0)}{2h_uh_w} \nonumber \\ 
+           & \; -\frac{f(h_u,0) - f(-h_u,0) - f(0,h_w) - f(0,-h_w)}{2h_uh_w} \;,
+  \end{align}
+\end{subequations}
+%
+where 
+the subscripts denote the Cartesian component of the electric field and $h$ is the differentiation step
+assumed to be 0.0005~a.u.
 
@@
  \begin{align}
-  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}}   \\
+  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \\
   {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
-     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} 
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;,
  \end{align}
@@
 \includegraphics[width=0.5\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
-\caption{\label{f:fig-1} A figure caption. The figure captions are
-automatically numbered.}
+\caption{\label{f:fig-1} {\bf Performance of the density matrix polarization susceptibility tensors of water molecule
+in moderate uniform electric field.} The notation $(R_{\max}0)$ denotes the maximum rank of the electric field
+and zero rank of electric field gradient in the expansion from Eq.~\eqref{e:final-model.General}.
+Electric field magnitudes in a range from 0 to 0.035 a.u. were tested in 100 random samples with random direction of electric field.
+} 
 \end{figure}
@@
 accuracy of the generalized DMS models is good 
-(see also error distributions shown in Figure~\ref{f:results}).
+(see also error distributions shown in Figure~\ref{f:fig-2}).
 Nevertheless, water molecules can experience
@@
 \includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-2.eps}
-\caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
+\caption{\label{f:fig-2} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
 non\hyp{}uniform electric fields.}
@@
 
+\subsection{\label{ss:43}Limitations of the current generalized DMS model}
+
+The need of including at least quadratic terms with respect to the electric field becomes clear
+already when considering the simplest case with uniform electric field (Figure~\ref{f:fig-1}).
+However, using the linear regression model from Eq.~\eqref{e:final-model.General.Parameters} 
+to extract such quadratic dependence
+yields wrong polarization energies, despite the one\hyp{}electron properties are accurately reproduced
+(Figure~\ref{f:fig-2} and Table~\ref{t:results}).
+This is probably because the current generalized DMS model is not fully utilizing the variational flexibility
+of the loss function in Eq.~\eqref{e:Z} due to the linear parameterization of a non\hyp{}linear susceptibilities 
+(Eq.~\eqref{e:final-model.General}). 
+For example, the second\hyp{}order susceptibility, ${\bf B}^{(20)}$, is constrained to
+form
+%
+\begin{equation}
+ B_{uw}^{(20)} \approx S_u^{[2]} + S_w^{[2]} \;,
+\end{equation}
+%
+%where ${\bf S}$ is the parameter block. 
+leading to symmetric matrices.\footnote{Extending the model by adding additional independent parameter block, i.e.,
+$B_{uw}^{(20)} \approx S_u^{[2]} + S_w^{[3]}$, leads to singularity in the Hessian matrix. Therefore,
+if the linear regression model is considered, the parameter blocks need to be the same.} 
+Note however, that not all symmetric matrices
+can be factored out in a way shown above which puts significant limits to a linear regression method
+currently used.
+It is also natural to expect that $B_{uw}^{(20)}$ does not need to be symmetric if it refers to the distributed site(s).
+Although such linear parameterization leads to very simple, one\hyp{}step optimization problem 
+from Eq.~\eqref{e:Newton} (Hessian is parameter\hyp{}independent), it
+is unfortunately insufficient for an accurate description of two\hyp{}electron effects.
+On the other hand, constructing more sophisticated models of $B_{uw}^{(20)}$ and $B_{uw}^{(01)}$ 
+requires introducing non\hyp{}linearity to the model, which is out of scope of the present contribution.
+This shall be studied in detail elsewhere.
+
 \section{\label{s:5}Summary and a few concluding remarks}
@@
 
-\section{\label{a:blocks} Explicit Formulae for Gradient and Hessian Blocks}
+\section{\label{a:blocks} Explicit Formulae for Gradient and Hessian Blocks in Linear Regression DMS Model}
 

### 2018-07-23 papers/1_dmatpol/paper-r2.tex
Revision in notation

@@
 \begin{multline}  \label{e:susceptibility-B}
- {\bf B}^{(i;10)}_{\alpha\beta} = \Big\{
+ {\bf B}^{(i;10)}_{\alpha\beta} = %\Big\{
                                C_{\alpha i} {\bf b}_{\beta i}^{(10)} + C_{\beta i} {\bf b}_{\alpha i}^{(10)} \\
@@
 %
-\Big\}  \;.
+%\Big\} 
+ \;.
 \end{multline}
@@
 Hence, the three dimensional vector ${\bf B}^{(i;10)}_{\alpha\beta}$ describes the polarization affinity
-of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
+of the density matrix element associated with the $\alpha$th and $\beta$th orthogonal AO's.
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
@@
 
-It is instructive to start first from the discussion of the polarization under the uniform electric field
+It is instructive to start first from the discussion of polarization under the uniform electric field
 condition because, in this case, it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
@@
  \begin{align}
-  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \\
+  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B10} \\
   {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
-     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;,
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B20}
  \end{align}
@@
 \end{figure}
+%
+All the models that are linear with respect to the electric field incorrectly predict the polarization energy.
+Interestingly, the predicted values are in perfect correlation with the exact estimates with a slope of $\sim$2.
+Incorporation of the second\hyp{}order term from Eq.~\eqref{e:B20} 
+obtained by numerical differentiation dramatically reduces the errors
+leading to quantitative accuracy (with RMSE of 0.029 kcal/mol and $R^2=99.58\%$;
+see also red filled dots in Figure~\ref{f:fig-1}).
+What is particularly
 

### 2018-07-23 papers/1_dmatpol/paper-r2.tex
Discussion on uniform fields added.

@@
 \begin{equation*} \label{e:initial-model}
- \delta {\bf D} \approx \sum_i^{\text{Sites}} \text{[DMS of the $i$th site]} 
-  \cdot  \delta {\bf P}_i
+% \delta {\bf D} \approx \sum_i^{\text{Sites}} \text{[DMS of the $i$th site]} 
+  \delta {\bf D} \approx \sum_i^{\text{Sites}} {\bf B}_i
+  \cdot  \delta {\bf P}_i \;.
   %\rightarrow \sum_i^{\rm occ} {\bf B}^{(i)}_{\mu\nu} \cdot {\bf E}_i \;.
@@
 %
-(In this Work we shall use the special open bold font for matrices of vectors, so that
+(In this work we shall use the special open bold font for matrices of vectors, so that
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
@@
 for $w=x,y,z$. 
-The superscript `$(10)$' denotes the fact that the susceptibility introduced here
+The superscript `$(10)$' underlines the fact that the susceptibility introduced here
 is first\hyp{}order with respect to the electric field and zeroth\hyp{}order with respect to
@@
 The Hessian of $Z$ computed with respect to the parameters is non\hyp{}singular
-and constant.
+and parameter\hyp{}independent (constant).
 Therefore, the exact solution for the optimal parameters is given by the Newton equation
@@
   {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
-     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B20}
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;. \label{e:B20}
  \end{align}
@@
 %
-and the associated error analysis is shown in Figure~\ref{f:fig-1}.
+The associated error distributions in prediction of polarization energies 
+are shown in scatter plot in Figure~\ref{f:fig-1}, whereas
+statistical performance descriptors of polarization energies and induced multipole moments
+are gathered in Table~XXX.
 \begin{figure}[h]
@@
 and zero rank of electric field gradient in the expansion from Eq.~\eqref{e:final-model.General}.
-Electric field magnitudes in a range from 0 to 0.035 a.u. were tested in 100 random samples with random direction of electric field.
+Electric field magnitudes in a range from 0 to 0.035 a.u. were tested in 100 samples with randomly selected 
+directions of electric field.
 } 
@@
 %
-All the models that are linear with respect to the electric field incorrectly predict the polarization energy.
+All the tested models that are linear with respect to the electric field 
+incorrectly reproduce the polarization energy.
 Interestingly, the predicted values are in perfect correlation with the exact estimates with a slope of $\sim$2.
+%especially the \emph{ab initio} models from Eq.~\eqref{e:final-model.HF} and Eq.~\eqref{e:dD-Taylor}
 Incorporation of the second\hyp{}order term from Eq.~\eqref{e:B20} 
@@
 leading to quantitative accuracy (with RMSE of 0.029 kcal/mol and $R^2=99.58\%$;
-see also red filled dots in Figure~\ref{f:fig-1}).
-What is particularly
+see also red filled circles in Figure~\ref{f:fig-1}). On the other hand,
+induced dipole and quadrupole moments are reproduced quantitatively well already by
+using first\hyp{}order DMS, ${\bf B}^{(10)}$, with $R^2$ coefficients being systematically
+greater than 95\%. This observation can be understood by the fact that the induced dipole
+and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
+energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
+quadratic terms in the expansion from Eq.~\eqref{e:dD-Taylor}.
+
+Performance of the generalized DMS models based on the linear regression 
+is inferior with respect to the \emph{ab initio} approach in terms of calculations of polarization energy.
+Although the linear model (blue open circles in Figure~\ref{f:fig-1})
+still predicts the correct slope of $\sim$2 at this level of approximation with RMSE of 1.19 kcal/mol,
+yet the quadratic model (red open circles)
+improves the performance only slightly, yelding RMSE as high as 0.89 kcal/mol
+and produces significant spread of points around the correlation line. This striking difference
+between the second\hyp{}order \emph{ab initio} and generalized models originates from the limitations
+in the current fitting procedure, that is used to approximate non\hyp{}linear susceptibilities (see further discussion).
 
@@
 
+To study the capability of the presented distributed models to capture 
+density matrix polarization in inhomogeneous electric fields, we analyse water molecule
+surrounded by point charges.
 Three different types of point charge environment that generate
@@
 Including DMS of type ${\bf B}^{(20)}$, i.e., that interact with the square of electric field,
-leads to improvement in prediction of all quantities evaluated in all electric field strengths considered in this Work. 
+leads to improvement in prediction of all quantities evaluated in all electric field strengths considered in this work. 
 Interestingly, interaction energies are somewhat insensitive to addition of the quadrupolar DMS 
@@
 It is also natural to expect that $B_{uw}^{(20)}$ does not need to be symmetric if it refers to the distributed site(s).
-Although such linear parameterization leads to very simple, one\hyp{}step optimization problem 
-from Eq.~\eqref{e:Newton} (Hessian is parameter\hyp{}independent), it
-is unfortunately insufficient for an accurate description of two\hyp{}electron effects.
+Although such linear parameterization enables a relatively simple, one\hyp{}step optimization
+from Eq.~\eqref{e:Newton}, 
+it is unfortunately insufficient for an accurate description of two\hyp{}electron effects which manifest
+here in energy by the Fermi electron correlation at the HF level.
 On the other hand, constructing more sophisticated models of $B_{uw}^{(20)}$ and $B_{uw}^{(01)}$ 
@@
 
-In this Work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
+In this work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
 at the HF level was derived for a closed shell molecule in a spatially non\hyp{}uniform

### 2018-07-27 papers/1_dmatpol/paper-r3.tex
Field Gradient removed fromthe paper

--- 2018-07-22 papers/1_dmatpol/paper-r2.tex
@@
 dipole moments and quadrupole moments
-are quantitatively described by the present theory in weak and moderate electric fields (0.003 -- 0.02 a.u.).
-%It is believed that the possibility of extension of the proposed model with high\hyp{}order susceptibilities
-%opens new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
-It is believed that the proposed model, after extensive development on its optimization, could open
-new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
+are quantitatively described by the present theory in weak, moderate and even very strong electric fields (0.003 -- 0.07 a.u.).
+It is believed that the proposed model could open
+new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems
+at arbitrary level of theory.
+%at post\hyp{}Hartree\hyp{}Fock levels of theory.
 \end{abstract}
@@
 solutions and biomolecules come into play.\cite{Tomasi.Mennucci.Cammi.ChemRev.2005}
-%Although the advent in the computer technology 
-%brought some hopes, it is still insufficient to follow up the expensive needs
-%of computational chemistry.
 Due to this reason, it has been of particular importance nowadays 
@@
 could be combined with low\hyp{}level and cheap classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
-%That was possible due to the fruitful incorporation of accurate \emph{ab initio} methods and a 
-%semi\hyp{}classical notion of molecular motion combined with extensive experimental and semi\hyp{}empirical
-%refinings.
 Understanding physics that rules intermolecular interactions\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
@@
 within an effective environment and fragmentation methods.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
-%The size of quantum level can be extended without compromising accuracy and maintaining good performance
-%by using vast family of fragmentation methods, such as effective fragment potentials (EFP).
-%The most difficult to tackle part of Hamiltonian can also be embedded within a widely understood effective environment
-%leading to various embedding methodologies, realising similar goals.
 
 However, even with the modern advances in computational chemistry, 
-%accurate description of environmental effects
 accurate description of phenomena due to complex heterogeneous molecular environments 
@@
 they cannot be used for highly correlated wavefunctions such as electronically excited states.
-%Therefore, Hamiltonian embedding is often realised by Coulombic embedding with polarization
-%of QM level, and only rarely the Pauli exclusion principle is taken into account. Although
-%such approach is very accurate in certain cases, it is not guaranteed that it is sufficient
-%for all cases.
 
@@
 though still not solved exactly.\cite{Holas.March.PhysRevA.1991}
-%that seem to have an important advantage over the standard energetic approach.
 With connection to the above, it would be incredibly insightful 
@@
 density and the perturbation in the environment through introducing the
-\emph{density matrix susceptibility} (DMS) tensors. % because knowing the density distribution
-%is equivalent to knowing the state of the molecule.
+\emph{density matrix susceptibility} (DMS) tensors. 
 The main postulate of the present work is that the associated change 
@@
 evaluated at certain site $i$.
-%and induced dipole moments $\delta {\BM\upmu}_i$ due to electric fields ${\bf E}_i$ 
-%at the centroids of the occupied molecular orbitals,
 %
 \begin{equation*} \label{e:initial-model}
- \delta {\bf D} \approx \sum_i^{\text{Sites}} \text{[DMS of the $i$th site]} 
-  \cdot  \delta {\bf P}_i
-  %\rightarrow \sum_i^{\rm occ} {\bf B}^{(i)}_{\mu\nu} \cdot {\bf E}_i \;.
+  \delta {\bf D} \approx \sum_i^{\text{Sites}} {\bf B}_i
+  \cdot  \delta {\bf P}_i \;.
 \end{equation*}
@@
 would reduce to much simpler matrix operations.
-%
-%a simple first\hyp{}principles model for prediction 
-%of the perturbation\hyp{}induced one\hyp{}electron density
-%change of structurally unrelaxed molecule is proposed here.
 
@@
 field. 
-%Unfortunately, such a mathematical problem is very complicated due to its high\hyp{}order tensor nature
-%with respect to the unknown $\delta {\bf D}$ and no approximate analytical solutions have been proposed
-%up to date. Therefore, in this paper we adopt a somewhat different
-%approach that will require \emph{a priori} knowledge from training set of known cases.
 Therefore, 
 we start our analysis by considering the induced electronic dipole moment 
-of a closed-shell molecule, %which is the first non\hyp{}vanishing moment,
+of a closed-shell molecule, 
 %
@@
 into Eq.~\eqref{e:dmu-exact} leads to
-%%
-%\begin{equation} \label{e:delta-Taylor}
-% \delta {\bf D} = \left[ {\bf v} + {\bf v}^\dagger \right]
-%                + \left[ {\bf v}{\bf v}^\dagger - {\bf v}^\dagger{\bf v}\right] + \ldots 
-%\end{equation}
 %
@@
 Now, if the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
-%%
-%\begin{subequations}
-% \begin{align}
-%  \Tr{ 
-%    \left[ 
-%         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
-%    \right] }  &
-%%
-%  \nonumber\\ 
-%  +\Tr{ 
-%    \left[ 
-%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
-%    \right] }
-% &=
-%2 \Tr{ 
-%    \left[ 
-%         \widetilde{\mathbb{M}} \cdot \delta {\bf C}
-%   \right] }  \;, 
-%%
-%\\
-%%
-%  \Tr{ 
-%    \left[ 
-%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
-%    \right] }  &
-%%
-%   \nonumber\\
-% +\Tr{ 
-%    \left[ 
-%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
-%    \right] }
-% &= 
-%2 \Tr{ 
-%    \left[ 
-%         \widetilde{\mathbb{K}} \cdot \delta {\bf C}
-%   \right] } \;.
-%%
-% \end{align}
-%\end{subequations}
-%%
 %
@@
 due to the presence of trace operation, unless the electric field is uniform.
-%in which ${\bf D}^{(0)}$ is the unperturbed density matrix of the system whereas
-%$\delta {\bf D}$ is the associated change due to the external electric field distribution. 
-%In fact, $\delta {\bf D}$ is a complicated
-%but unknown function of ${\bf F}({\bf r})$ and basis set. 
-%Moreover, it certainly depends on the choice of
-%molecular orbitals because the unitary transformation . 
 Let us partition the induced dipole moment into separate contributions
@@
 change in the LCAO-MO matrix due to the polarization process.
-%
-%The above equation can be re-written in a matrix form as
-%
-%\begin{equation} 
-% \left[ {\mathbb L} \right]_{\alpha \zeta \beta} = L_{\beta\alpha}^{(\zeta)}
-%\end{equation}
-%%
-%of size $m \times 3 \times m$.
-%Note that, in this notation, ${\BM{\upmu}}_i^T$ and $\overline{\BM\Delta}_i^T$ are row vectors
-%whereas $\overline{\BM\Delta}_i^p$ is a column vector. 
 The solution for $\delta {\bf C}$ is
@@
 $n > 3$.
-%Clearly, in the above approximation, ${\bf X}$ depends on the choice
-%of unperturbed orbitals.
 
@@
 \begin{equation} \label{e:dC-dmatpol}
- \delta C_{\alpha i} = {\bf b}_{\alpha i}^{(10)} \cdot {\bf F}({\bf r}_i)  \;,
+ \delta C_{\alpha i} = {\bf b}_{\alpha}^{(i;1)} \cdot {\bf F}({\bf r}_i)  \;,
 \end{equation}
@@
 \begin{equation} \label{e:susceptibility-b}
-  b_{\alpha i;w}^{(10)}= \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
+  b_{\alpha;w}^{(i;1)}= \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
    \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  
@@
 for $w=x,y,z$. 
-The superscript `$(10)$' denotes the fact that the susceptibility introduced here
-is first\hyp{}order with respect to the electric field and zeroth\hyp{}order with respect to
-the electric field gradient.
+The superscript `$(1)$' underlines the fact that the susceptibility introduced here
+is first\hyp{}order with respect to the electric field.
 The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
@@
 \begin{equation}\label{e:final-model.HF}
- \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i;10)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
+ \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i;1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
 \end{equation}
@@
 \begin{multline}  \label{e:susceptibility-B}
- {\bf B}^{(i;10)}_{\alpha\beta} = \Big\{
-                               C_{\alpha i} {\bf b}_{\beta i}^{(10)} + C_{\beta i} {\bf b}_{\alpha i}^{(10)} \\
+ {\bf B}^{(i;1)}_{\alpha\beta} = 
+                               C_{\alpha i} {\bf b}_{\beta}^{(i;1)} + C_{\beta i} {\bf b}_\alpha^{(i;1)} \\
                                 - \sum_\gamma 
@@
                                  \right)
-                                           {\bf b}_{\gamma i}^{(10)}
-%
-\Big\}  \;.
+                                           {\bf b}_{\gamma}^{(i;1)}
+ \;.
 \end{multline}
 %
-Hence, the three dimensional vector ${\bf B}^{(i;10)}_{\alpha\beta}$ describes the polarization affinity
-of the density matrix element associated with $\alpha$th and $\beta$th orthogonal AO's.
+Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ describes the polarization affinity
+of the density matrix element associated with the $\alpha$th and $\beta$th orthogonal AO's.
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
-It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i;10)}$
+It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
 are unique properties of electronically unperturbed molecule.
@@
  \delta D_{\alpha\beta} = \sum_{i }^M \Big\{
-                                      {\bf B}^{(i ; 10)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  
-%                       + \sum_{ij}^M {\bf B}^{(ij; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
-                        +             {\bf B}^{(i ; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \\
-                        +             {\bf B}^{(i ; 01)}_{\alpha\beta} : {\BM\nabla} \otimes {\bf F}({\bf r}_i) 
+                                      {\bf B}^{(i ; 1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \\
+%                       + \sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
+                        +             {\bf B}^{(i ; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) 
                         + \ldots \Big\}
@@
 %
-%However, Eq.~\eqref{e:final-model} is an approximation because (i) it is based on the approximate
-%relation in Eq.~\eqref{e:dmu-l-vector-mo-transform} and (ii) only total dipole moment (but not quadrupole
-%and higher moments) is exactly
-%reproduced by the current model. Due to the above reasons, the existence of ${\bf X}$ that
-%reproduces the exact perturbed density matrix with vanishing error is not guaranteed. 
-%Moreover, as it was pointed out before, ${\bf X}$ almost certainly depends on the external electric field. 
-%Next section adresses the problem of finding the optimal ${\bf X}$ that is treated as an adjustable
-%matrix.
 In the above equation, the density matrix susceptibilities
-${\bf B}^{(i; RP)}_{\alpha\beta}$, distributed over $M$ sites,
-denote the $R$-th order response with respect to the electric field
-and $P$-th order response with respect to the electric field gradient. Note that
-${\bf B}^{(i; 10)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 20)}_{\alpha\beta}$
-and ${\bf B}^{(i; 01)}_{\alpha\beta}$ are $(3\times 3)$ matrices. Also, the distributed expansion centres
+${\bf B}^{(i; R)}_{\alpha\beta}$, distributed over $M$ sites,
+denote the $R$-th order response with respect to the electric field.
+Note that
+${\bf B}^{(i; 1)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 2)}_{\alpha\beta}$
+is $(3\times 3)$ matrix. Also, the distributed expansion centres
 (labeled as $i$) are generalized to any point in the molecule.\footnote{More general
-contribution due to square of electric field could be $\sum_{ij}^M {\bf B}^{(ij; 20)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$.
+contribution due to square of electric field could be $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$.
 However, for simplicity we neglect the contributions from different sites $i$ and $j$
@@
 are known for each sample (overline symbolizes the exact estimate).
-%Let us assume that we a $\delta \overline{\bf D}^{(N)}$ 
-%is the $N$th known (exact) difference one\hyp{}particle density matrix
-%that was computed for the electric field distribution ${\bf F}^{(N)}({\bf r})$ chosen randomly.
 Now, for each pair of the AO indices we construct the following 
@@
 %               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
-                +             \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{iw} \right) F_{iu}^{(N)} F_{iw}^{(N)} \\
-                +             \sum_{uw}^{x,y,z} \left( S^{[3]}_{iu} + S^{[3]}_{iw}\right) F_{iu,iw}^{(N)}
+                +             \sum_{u}^{x,y,z} \sum_{w<u} r_{uw} S^{[2]}_{iuw} F_{iu}^{(N)} F_{iw}^{(N)} 
                         + \ldots \Big\}
@@
 (the Greek subscripts were omitted here for notational simplicity).
-The multiple parameter blocks 
-$S^{[1]}$, $S^{[2]}$ and $S^{[3]}$ that are associated with electric field and its gradients
+In the above equation, $B_u^{(i;1)} = S^{[1]}_{iu}$ and $B_{uw}^{(i;2)} = r_{uw} S^{[2]}_{iuw}$,
+where $r_{uw}$ is the symmetry factor equal to 1 for diagonal elements and 2 for off\hyp{}diagonal
+elements of $B_{uw}^{(i;2)}$.
+Note that the multiple parameter blocks 
+$S^{[1]}$ and $S^{[2]}$
 appear in the first power, allowing for linear least\hyp{}squares regression.
-Note that $B_u^{(i;10)} = S^{[1]}_{iu}$, $B_{uw}^{(i;20)} = S^{[2]}_{iu} + S^{[2]}_{iw}$
-and $B_{uw}^{(i;01)} = S^{[3]}_{iu} + S^{[3]}_{iw}$. The square bracket superscripts 
-denote the block of the parameter space. We chose symmetric forms of ${\bf B}^{(i;20)}$
-and ${\bf B}^{(i;01)}$ (antisymmetric form could be constructed as well by taking the difference
-of the parameters from block 2 and 3, respectively). $F_{iu}^{(N)}$ and $F_{iu,iw}^{(N)}$ 
-stand for the electric field $\left[ {\bf F}({\bf r}_i)\right]_u$ and the electric field
-gradient $\left[ {\BM \nabla} \otimes {\bf F}({\bf r}_i) \right]_{uw}$, respectively.
+The square bracket superscripts denote the block of the parameter space.
 
@@
 {\bf S}^{[1]} &
-{\bf S}^{[2]} &
-{\bf S}^{[3]}
+{\bf S}^{[2]} 
 \end{pmatrix}^T
@@
 %
-%where $N_{\rm max}$ is the total number of samples. 
-%The linear form of Eq.~\eqref{e:final-model.General.Parameters} ensures that the Hessian 
-%of $Z$
-%is non\hyp{}singular and independent on the parameters. 
 The Hessian of $Z$ computed with respect to the parameters is non\hyp{}singular
-and constant.
+and parameter\hyp{}independent (constant).
 Therefore, the exact solution for the optimal parameters is given by the Newton equation
@@
 where ${\bf g}$ and ${\bf H}$ are the gradient vector and the Hessian matrix, respectively.
-%
-%\begin{equation}\label{e:Newton.S}
-% {\bf S} = 
-%\begin{pmatrix}
-%{\bf S}^{[1]} &
-%{\bf S}^{[2]} &
-%{\bf S}^{[3]}
-%\end{pmatrix}^T
-%\end{equation}
-%
 The explicit forms of the gradient and Hessian are given in the Appendix~\ref{a:blocks}.
-Note that the dimensions of parameter space for the block 1, 2 and 3 are all
-equal to $3M$.
-%unperturbed molecule. This makes them perfect candidates for general effective fragment parameters.
+Note that the dimensions of parameter space for the block 1 and 2 are
+equal to $3M$ and $6M$, respectively.
 
@@
 Water molecule
-%, energy\hyp{}optimized by using the HF/6-311++G** 
-%method\cite{Roothaan.RevModPhys.1951,Krishnan.Binkley.Seeger.Pople.JCP.1980}, 
 was selected as a test system which was subject to 
@@
 
-%For finding a general unitary transformations,
-%100 random point charge distributions were selected for which an exact density matrices
-%were calculated.
 In order to validate all the models discussed in this contribution,
@@
 the polarization of the electron density distribution, respectively. 
-%$E_{\rm nuc}({\bf F})$ denotes the interaction energy due to the nuclei.
 In order to compare the performance of the distributed polarizability approach,
@@
 \begin{equation}\label{e:dE-dpol}
- \delta E^{\rm Pol}[{\bf F}] \approx %E_{\rm nuc}({\bf F}) + 
-                     %2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} }  \\
+ \delta E^{\rm Pol}[{\bf F}] \approx 
                    - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \;,
@@
 $i$th localized molecular orbital (LMO) centroid (see Eq.~\eqref{e:dmu-dpol}).
-%, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
 The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
@@
 %
-%where ${\bf r}_{\rm CM}$ is the center of mass of water molecule.
 Primitive quadrupole moments were subsequently transformed to their traceless forms by
@@
 %
-Methodology for the calculation of the electric field and electric field gradient due to point charges
+Methodology for the calculation of the electric field due to point charges
 can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
@@
  \begin{align}
-  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \\
+  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B10} \\
   {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
-     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;,
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;. \label{e:B20}
  \end{align}
@@
 %
-and the associated error analysis is shown in Figure~\ref{f:fig-1}.
+and consider electric field magnitudes in a range from very weak (~0.003 a.u.) to strong (0.07 a.u.).
+The associated error distributions in prediction of polarization energies 
+are shown in scatter plot in Figure~\ref{f:fig-1}, whereas the
+statistical performance descriptors of polarization energies and induced multipole moments
+are gathered in Table~\ref{t:results}.
+%
+\begin{table*}%[]
+\caption{{\bf Performance of the density matrix susceptibility tensors of water molecule 
+in weak, moderate and strong electric fields.}
+In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
+respectively. $R$-squared coefficients are shown in parentheses.
+Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$
+where $R_{\rm max}$ is the maximum power in the expansion
+from Eq.~\eqref{e:final-model.General}.
+}
+\label{t:results}
+\begin{ruledtabular}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{cddddddcddddcdd}
+%\hline\hline
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalized DMS}}                                                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
+\cline{2-8}
+\cline{9-12}
+%\cline{14-15}
+\textbf{}         
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1^*}$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2^*}$}
+&
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$} 
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$} 
+&  
+& \multicolumn{2}{c}{\textbf{Polarizability}}        \\ \hline
+\multicolumn{15}{c}{} \\ 
+\multicolumn{15}{c}{\textbf{Uniform Electric Field}} \\ 
+\multicolumn{15}{c}{} \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}              \\ %\hline
+   & 4.041 & (-) & 4.823 & (-) & 0.231 & (98.40) && 5.0121 & (-) & 0.173 & (99.10) && 0.112 & (99.63) \\ 
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                    \\ %\hline
+   & 30.46 & (97.91) & 30.46 & (97.91) & 18.02 & (99.27) && 24.83 & (98.61) & 7.13 & (99.89) && 30.46 & (97.91) \\ 
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}} \\ %\hline
+   & 47.89 & (69.17) & 39.24 & (79.30) &  8.56 & (99.02) && 37.63 & (80.96) & 3.94 & (99.79) && 50.66 & (65.49) \\ \hline
+\multicolumn{15}{c}{} \\ 
+\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}} \\ 
+\multicolumn{15}{c}{} \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}   \\ 
+W & 0.067 & (-)     & - & (-) & - & (-) && 0.076 & (-)     & 0.003 & (99.45) && 0.004 & (98.64)\\ 
+M & 1.072 & (-)     & - & (-) & - & (-) && 1.236 & (-)     & 0.057 & (99.08) && 0.070 & (98.58)\\ 
+S & 4.220 & (-)     & - & (-) & - & (-) && 5.147 & (-)     & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}  \\ 
+W &  3.36 & (98.71) & - & (-) & -  & (-) &&  1.03 & (99.88) &  0.94 & (99.90) &&  3.36 & (98.71) \\ 
+M & 14.72 & (98.50) & - & (-) & -  & (-) &&  7.72 & (99.59) &  4.55 & (99.86) && 14.72 & (98.50)\\ 
+S & 44.80 & (96.92) & - & (-) & -  & (-) && 33.06 & (98.33) & 15.96 & (99.61) && 44.80 & (96.92) \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}} \\ 
+W &  6.22 & (58.72) & - & (-) & - & (-) &&  1.59 & (97.32) &  1.42 & (97.86) &&  5.45 & (68.29) \\ 
+M & 28.22 & (53.90) & - & (-) & - & (-) && 10.45 & (93.68) &  5.68 & (98.13) && 23.94 & (66.83) \\ 
+S & 78.07 & (40.53) & - & (-) & - & (-) && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ \hline
+\bottomrule
+\end{tabular}%
+}
+\end{ruledtabular}
+\end{table*}
+%
 \begin{figure}[h]
@@
 \caption{\label{f:fig-1} {\bf Performance of the density matrix polarization susceptibility tensors of water molecule
-in moderate uniform electric field.} The notation $(R_{\max}0)$ denotes the maximum rank of the electric field
-and zero rank of electric field gradient in the expansion from Eq.~\eqref{e:final-model.General}.
-Electric field magnitudes in a range from 0 to 0.035 a.u. were tested in 100 random samples with random direction of electric field.
+in uniform electric field.} The notation $(R=R_{\max})$ denotes the maximum rank of the electric field
+in the expansion from Eq.~\eqref{e:final-model.General}.
+Electric field magnitudes in a range from 0.0 to 0.07 a.u. were tested in 100 samples with randomly selected 
+directions of electric field.
 } 
 \end{figure}
+%
+All the tested models that are linear with respect to the electric field,
+i.e., Eq.~\eqref{e:final-model.HF} as well as Eqs.~\eqref{e:dD-Taylor} and \eqref{e:final-model.General} 
+truncated at linear susceptibilities, ${\bf B}^{(1)}$,
+incorrectly reproduce the polarization energy.
+Interestingly, the predicted values are in noticeable correlation with the exact estimates with a slope of $\sim$2.
+Incorporation of the second\hyp{}order term
+obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
+or by using the generalized DMS model (red open circles)
+dramatically reduces the errors
+leading to quantitative accuracy in all tested field magnitudes 
+with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalized DMS model.
+On the other hand, the
+induced dipole and quadrupole moments are reproduced reasonably well already by
+using the first\hyp{}order DMS, ${\bf B}^{(10)}$, with $R^2$ coefficients being 
+greater than 97\% in case of dipole moments and grater than 80\% in case of quadrupole moments.
+This observation can be understood by the fact that the induced dipole
+and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
+energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
+quadratic terms in the expansion from Eq.~\eqref{e:dD-Taylor}.
+
+In comparison to the dipole\hyp{}dipole polarizability approach, 
+Quadratic DMS models perform only slightly worse 
+(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalized quadratic DMS
+model and the polarizability model are used, respectively).
+However, quadratic DMS models better predict he induced multipole moments with
+%work better
+%when predicting the induced multipole moments with
+$R^2$ systematically greater than 99\%. 
+Certainly, taking into account the distributed quadrupole polarizabilities
+would greatly reduce errors of the standard dipole\hyp{}dipole polarizability approach.
+Nevertheless, it is clear that the second\hyp{}order
+DMS models are quantitative even at very strong uniform electric fields up to 0.07 a.u.
+%On the other hand, DMS models better predict the induced dipole and quadrupole 
+%moments even in weak fields.
 
@@
 electric fields around the water molecule were considered (Table~\ref{t:fields}),
-because similar ranges of electric field are experienced by water molecules
-in liquid phase.\cite{Reischl.Kofinger.Dellago.MolPhys.2009,Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
-%
-\begin{table}%[]
-%\centering
+%
+\begin{table}[b]
 \caption{{\bf Average electric fields in statistical sets of electrostatically perturbed states
@@
 %
+because similar ranges of electric field are experienced by water molecules
+in liquid phase.\cite{Reischl.Kofinger.Dellago.MolPhys.2009,Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
 Performance descriptors of the \emph{ab initio} and generalized 
@@
 approach, are shown in Table~\ref{t:results}.
-%
-\begin{table*}%[]
-%\centering
-\caption{{\bf Performance of the density matrix susceptibility tensors of water molecule 
-in weak, moderate and strong electric fields.}
-In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
-respectively. $R$-squared coefficients are shown in parentheses.
-Ranks of the density matrix susceptibility models are denoted as ($R_{\rm max}, P_{\rm max}$),
-where $R_{\rm max}$ and $P_{\rm max}$ are the maximum ranks of the electric field 
-and the electric field gradient in the expansion
-from Eq.~\eqref{e:final-model.General}, respectively.
-}
-\label{t:results}
-\begin{ruledtabular}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{cddcddddddddcdd}
-%\hline\hline
-\toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{\textbf{An Initio DMS}} &  & \multicolumn{8}{c}{\textbf{Generalized DMS}}                                                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
-\cline{2-3}
-\cline{5-12}
-%\cline{14-15}
-\textbf{Set}         
-& \multicolumn{2}{c}{\textbf{1, 0}}          
-&  
-& \multicolumn{2}{c}{\textbf{1, 0}} 
-& \multicolumn{2}{c}{\textbf{2, 0}} 
-& \multicolumn{2}{c}{\textbf{1, 1}} 
-& \multicolumn{2}{c}{\textbf{2, 1}} 
-&  
-& \multicolumn{2}{c}{\textbf{Polarizability}}         \\ \hline
-\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}                                                                                                                                                                                             \\ \hline
-W                    & 0.067               & (99.80)              &  & 0.076          & (99.74)          & 0.053          & (99.88)          & 0.074          & (99.76)          & 0.050          & (99.89)          &  & 0.004            & (100.00)            \\
-M                    & 1.07                & (96.96)              &  & 1.24           & (95.96)          & 0.86           & (98.06)          & 1.20           & (96.17)          & 0.81           & (98.27)          &  & 0.07             & (99.99)             \\
-S                    & 4.22                & (88.87)              &  & 5.15           & (83.44)          & 3.60           & (91.92)          & 5.02           & (84.25)          & 3.41           & (92.73)          &  & 0.33             & (99.93)             \\ \hline
-\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                                                                                                                                                                                                   \\ \hline
-W                    & 3.36                & (98.70)              &  & 1.03           & (99.88)          & 0.98           & (99.89)          & 0.52           & (99.97)          & 0.47           & (99.98)          &  & 3.36             & (98.70)             \\
-M                    & 14.72               & (98.50)              &  & 7.72           & (99.59)          & 6.86           & (99.67)          & 6.15           & (99.74)          & 5.38           & (99.80)          &  & 14.72            & (98.50)             \\
-S                    & 44.80               & (96.92)              &  & 33.06          & (98.32)          & 29.49          & (98.67)          & 28.74          & (98.73)          & 26.25          & (98.94)          &  & 44.80            & (96.92)             \\ \hline
-\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}}                                                                                                                                                                                               \\ \hline
-W                    & 6.22                & (58.72)              &  & 1.59           & (97.32)          & 1.49           & (97.62)          & 0.58           & (99.64)          & 0.48           & (99.75)          &  & 5.45             & (68.30)             \\
-M                    & 28.22               & (53.90)              &  & 10.45          & (93.68)          & 8.83           & (95.49)          & 8.25           & (96.06)          & 6.57           & (97.50)          &  & 23.94            & (66.83)             \\
-S                    & 78.07               & (40.53)              &  & 43.33          & (81.68)          & 35.73          & (87.55)          & 38.80          & (85.31)          & 31.59          & (90.26)          &  & 64.33            & (59.62) \\
-\bottomrule
-\end{tabular}%
-}
-\end{ruledtabular}
-\end{table*}
-%
-
-The \emph{ab initio} DMS model quantitatively reproduces the interaction energy and the dipole moment 
-in weak electric field (RMSE of 0.067~kcal/mol and RMSD of 3.4~mD). In stronger electric fields
-the predictions become less quantitative, with high RMSE of 4.2~kcal/mol and RMSD of 14.7~mD in strong electric fields.
-Induced quadrupole moment is described by the \emph{ab initio} DMS qualitatively well with $R^2$ not exceeding
-60\%. Generalized DMS model of the same rank ($R_{\rm max}$=1, $P_{\rm max}$=0) performs slightly worse
-for interaction energies but better for induced multipole moments, particularly decreasing the errors of induced quadrupole
-calculations almost by a factor of two in strong electric fields, and by a factor of three in weaker fields.
-By and large, extending the parameter space improves the accuracy of the generalized DMS model. 
-Including DMS of type ${\bf B}^{(20)}$, i.e., that interact with the square of electric field,
-leads to improvement in prediction of all quantities evaluated in all electric field strengths considered in this Work. 
-Interestingly, interaction energies are somewhat insensitive to addition of the quadrupolar DMS 
-(${\bf B}^{(01)}$, interacting with the electric field gradient), in contrast to induced multipole moments
-which exhibit noticeable improvement of accuracy. The most accurate generalized DMS model presented here 
-($R_{\rm max}$=2, $P_{\rm max}$=1, which corresponds to including ${\bf B}^{(10)}$, ${\bf B}^{(20)}$ and ${\bf B}^{(01)}$)
-works quantitatively well in weak and medium electric fields with $R^2$ being systematically 97.5\% and higher,
-and with RMSE of 0.8~kcal/mol in moderate fields. 
-However, the errors in the interaction energy
-in strong fields are 3.4~kcal/mol, which is quite high as compared to the usually acceptable accuracy 
-of 1~kcal/mol in routine quantum chemistry calculations. 
+
+Similarly as in the case of uniform electric field, the linear DMS models, i.e.,
+the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} as well as
+the linear DMS model, are insufficient to correctly describe the polarization energy,
+although the correlation with benchmark is still very clear with a slope of $\sim2$
+(Figure~\ref{f:fig-2}A). Again, induced dipole moments are reproduced very well even
+in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$).
+Induced quadrupole moments are well described only by the generalized DMS model.
+For example, in medium electric fields RMSQ is 10.4$\times 10^{-3}$D\AA{ }with $R^2=93.7\%$
+whereas the \emph{ab initio} DMS model yealds RMSQ of 28.2$\times 10^{-3}$D\AA{ }($R^2=53.9\%$).
+
+The most sophisticated model studied in this work, the generalized quadratic DMS model, yields excellent predictions 
+of the polarization energies and induced multipole moments in all the electric field strengths tested.
+The errors in predicting polarization energy are reduced by an order of magnitude when compared with
+the linear DMS models. Improvement of accuracy can also be observed in the case of induced multipole moment predictions
+but the effect is less significant (improvement roughly by a factor of 2).
+What is particularly interesting is that in weak and moderate electric fields
+the generalized quadratic DMS model performs better than the distributed dipole\hyp{}dipole
+polarizability model, whereas in strong electric fields the situation is reversed
+(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.2\%$ for the DMS and polarizability model, respectively).
+Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% in strong electric fields for polarization
+energy, and even higher $R^2$ for induced multipole moments.
 Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
@@
 and \ref{t:results}) for which the
-accuracy of the generalized DMS models is good 
-(see also error distributions shown in Figure~\ref{f:fig-2}).
-Nevertheless, water molecules can experience
+accuracy of the generalized quadratic DMS model is excellent
+Water molecules can experience
 stronger electric fields in certain geometries of hydrogen bonded clusters, 
 even up to 0.055 a.u. according to study by Reischl et~al.\cite{Reischl.Kofinger.Dellago.MolPhys.2009} 
-Therefore, in strong electric fields
-the present state of generalized DMS's must be further optimized in order to reach accuracy of at least
-1~kcal/mol.
-
-It is also instructive to compare the DMS models with the conventional approach based on the distributed
-dipole\hyp{}dipole
-polarizabilities. In terms of accuracy in interaction energy prediction, DMS models are, on average, one order 
-of magnitude less accurate than the distributed polarizability approach, which provides errors of 0.3~kcal/mol
-or less in all cases. On the other hand, DMS models better predict the induced dipole and quadrupole 
-moments even in weak fields. Certainly, taking into account the distributed quadrupole polarizabilities
-would greatly reduce errors of the standard multipole approach. However, it is believed that
-the DMS approach, although less (but still sufficiently) accurate in terms of interaction energy, 
+This electric field magnitude falls into our strong electric field regime in which the generalized quadratic
+DMS model still provides accurate predictions (see also error distributions 
+shown in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H). 
+It is emphasized here that the current model can be easily extended by taking into account
+third\hyp{}order terms. 
+%This however falls out of scope of this contribution and it will be studied
+%elsewhere. 
+Therefore, it is believed that
+the generalized DMS approach 
 has an important advantage over the multipole\hyp{}based approach, i.e., it provides detailed
@@
 \end{figure*}
-%
-%\begin{table*}[]
-%\centering
-%\caption{My caption}
-%\label{t:results}
-%\resizebox{\textwidth}{!}{%
-%\begin{tabular}{|l|c|c|c|c|c|c|c|}
-%\hline
-%\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
-%\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
-%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8}
-%                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8}
-%                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
-%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8}
-%                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8}
-%                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
-%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8}
-%                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8}
-%                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
-%\end{tabular}%
-%}
-%\end{table*}
-%We first test the accuracy of the density matrix susceptibility tensors that were derived
-%in Eq.~\eqref{e:susceptibility-B} from first principles. 
-%For this purpose we have considered a water molecule 
-%under weak uniform electric fields (Figure~XXX). 
-%3D difference one\hyp{}electron density isosurface plots for sample electric fields
-%with respect to the unperturbed density (${\bf F} = {\bf 0}$)
-%are shown in Figure~XXX. The statistical evaluation of errors for a set of
-%100 random electric fields is shown in Figure~XXX. The shape of the basins 
-%of decreased/increased density are comparable with the exact result
-%obtained from HF/6-311++G** calculations. 
-%It can be seen that the theory
-%describes the polarization of the electron density distribution qualitatively well:
-%However, the accuracy in the density matrix elements is on overall very low which 
-%manifests with drammatic errors in the electronic total energy of the system.
-%It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
-%which neglects quadrupole and higher\hyp{}order induced multipole moments.
-
-%\subsection{General Model of Densty Matrix Polarization}
-
-%However, Eq.~\eqref{e:final-model} is an approximation because (i) it is based on the approximate
-%relation in Eq.~\eqref{e:dmu-l-vector-mo-transform} and (ii) only total dipole moment (but not quadrupole
-%and higher moments) is exactly
-%reproduced by the current model. Due to the above reasons, the existence of ${\bf X}$ that
-
-%The critical assessment of the proposed model is discussed in two stages: (i) first
-%the capability of the model in capturing the polarization 
-%It is instructive to analyze first the capability of the model in capturing the polarization
-%process by studying the 3D electronic density distortions due to the electric field perturbation.
-%For this purpose we consider a , i.e., without
-%optimization of the particular density matrix elements with respect to benchmark.
-
-%For this purpose we studied few molecules
-%in a uniform external electric field. Although the density matrix elements in a chosen AO basis
-%can be inconsistent with other quantities (such as Fock matrix and so on) the polarization
-
-\subsection{\label{ss:43}Limitations of the current generalized DMS model}
-
-The need of including at least quadratic terms with respect to the electric field becomes clear
-already when considering the simplest case with uniform electric field (Figure~\ref{f:fig-1}).
-However, using the linear regression model from Eq.~\eqref{e:final-model.General.Parameters} 
-to extract such quadratic dependence
-yields wrong polarization energies, despite the one\hyp{}electron properties are accurately reproduced
-(Figure~\ref{f:fig-2} and Table~\ref{t:results}).
-This is probably because the current generalized DMS model is not fully utilizing the variational flexibility
-of the loss function in Eq.~\eqref{e:Z} due to the linear parameterization of a non\hyp{}linear susceptibilities 
-(Eq.~\eqref{e:final-model.General}). 
-For example, the second\hyp{}order susceptibility, ${\bf B}^{(20)}$, is constrained to
-form
-%
-\begin{equation}
- B_{uw}^{(20)} \approx S_u^{[2]} + S_w^{[2]} \;,
-\end{equation}
-%
-%where ${\bf S}$ is the parameter block. 
-leading to symmetric matrices.\footnote{Extending the model by adding additional independent parameter block, i.e.,
-$B_{uw}^{(20)} \approx S_u^{[2]} + S_w^{[3]}$, leads to singularity in the Hessian matrix. Therefore,
-if the linear regression model is considered, the parameter blocks need to be the same.} 
-Note however, that not all symmetric matrices
-can be factored out in a way shown above which puts significant limits to a linear regression method
-currently used.
-It is also natural to expect that $B_{uw}^{(20)}$ does not need to be symmetric if it refers to the distributed site(s).
-Although such linear parameterization leads to very simple, one\hyp{}step optimization problem 
-from Eq.~\eqref{e:Newton} (Hessian is parameter\hyp{}independent), it
-is unfortunately insufficient for an accurate description of two\hyp{}electron effects.
-On the other hand, constructing more sophisticated models of $B_{uw}^{(20)}$ and $B_{uw}^{(01)}$ 
-requires introducing non\hyp{}linearity to the model, which is out of scope of the present contribution.
-This shall be studied in detail elsewhere.
 
@@
 and takes into account higher\hyp{}order curvatures of the electric field distribution as well.
-The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibility tensors
-that are able to interact with the electric field, square of electric field and electric field gradients.
+The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibilities
+that are able to interact with the electric field and square of electric field.
 The particular density matrix susceptibilities were determined through the least\hyp{}squares 
@@
 interaction energies, dipole moments and quadrupole moments of water molecule that
-interacts with various point charge environments generating electric fields of weak, moderate
-and strong magnitudes. It was found that, while the \emph{ab initio} model yelds qualitatively correct
-interaction energies and dipole moments, the generalized model can reach quantitative level of accuracy
-for prediction of interaction energies, dipole moments and even quadrupole moments. Although 
-errors in interaction energies produced by using the generalized model 
-were still at best one order of magnitude larger than the errors produced by using the
-usual approach based on the distributed polarizabilities, this work demonstrated that it is possible to
+interacts with uniform electric fields as well as 
+various point charge environments generating electric fields of weak, moderate
+and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
+interaction energies, the generalized model can reach quantitative level of accuracy
+for prediction of interaction energies, dipole moments and even quadrupole moments
+once at least quadratic terms with respect to the electric field are taken into account. 
+This work demonstrated that it is possible to
 accurately predict the change in the density matrix of electronic system, and the accuracy of the model
@@
 should be studied in the future with an emphasis 
-on their further optimization, especially to reach RMSE below 1 kcal/mole
-in strong and very strong electric fields (0.05-0.07 a.u.). 
+on their further optimization.
+%, especially to reach RMSE below 1 kcal/mole
+%in strong and very strong electric fields (0.05-0.07 a.u.). 
 For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
@@
 {\bf g}^{[1]} \\ 
-{\bf g}^{[2]} \\ 
-{\bf g}^{[3]}
+{\bf g}^{[2]} 
 \end{pmatrix} ,\quad
@@
 \begin{pmatrix}
-{\bf H}^{[11]} & {\bf H}^{[12]} & {\bf H}^{[13]} \\ 
-{\bf H}^{[21]} & {\bf H}^{[22]} & {\bf H}^{[23]} \\ 
-{\bf H}^{[31]} & {\bf H}^{[32]} & {\bf H}^{[33]} 
+{\bf H}^{[11]} & {\bf H}^{[12]}  \\ 
+{\bf H}^{[21]} & {\bf H}^{[22]}  
 \end{pmatrix} \;.

### 2018-07-27 oepdev/libgefp/gefp_polar_base.cc
Field Gradients result in singular Hessians. The DMATPOL models utilizing gradients are not working.

@@ -362,9 +362,9 @@ void oepdev::GeneralizedPolarGEFactory::save(int i, int j)
            PolarizationSusceptibilities_->quadrupole_polarizability(n, 7)->set(i, j, val_yz);
            PolarizationSusceptibilities_->quadrupole_polarizability(n, 8)->set(i, j, val_zz);
       }
-  } else {
-    throw psi::PSIEXCEPTION("The model you refer to is not implemented.");
-  }
+  } //else {
+  //  throw psi::PSIEXCEPTION("The model you refer to is not implemented.");
+  //}
 
   //if (!hasDipoleDipoleHyperpolarizability_ and !hasQuadrupolePolarizability_) {
   //    for (int n=0; n<nSites_; ++n) {

### 2018-07-27 oepdev/libgefp/gefp_polar_nonuniform_field_1_grad_1.cc

@@ -27,53 +27,170 @@ void oepdev::LinearGradientNonUniformEFieldPolarGEFactory::compute_gradient(int 
    const int d = nParametersBlock_[0];
 
    for (int s=0; s<nSites_; ++s) {
-   for (int z=0; z<3; ++z) {
-        int sz = 3*s + z;
-        double g1 = 0.0;
-        double g2 = 0.0;
-        for (int n=0; n<nSamples_; ++n) {                                                                  
-             double dij = -referenceStatisticalSet_.DensityMatrixSet[n]->get(i, j);
-             double fz = electricFieldSet_[n][s]->get(z);
-             double gs = electricFieldGradientSumSet_[n][s]->get(z) * 2.0;
-             g1 += dij * fz;
-             g2 += dij * gs;
+        double g1_x = 0.0;                                                        
+        double g1_y = 0.0;
+        double g1_z = 0.0;
+        double g2_xx= 0.0;
+        double g2_yy= 0.0;
+        double g2_zz= 0.0;
+        double g2_xy= 0.0;
+        double g2_xz= 0.0;
+        double g2_yz= 0.0;
+                                                                                  
+        for (int n=0; n<nSamples_; ++n) {
+             double fx = electricFieldSet_[n][s]->get(0);
+             double fy = electricFieldSet_[n][s]->get(1);
+             double fz = electricFieldSet_[n][s]->get(2);
+             double fxx= electricFieldGradientSet_[n][s]->get(0, 0);
+             double fxy= electricFieldGradientSet_[n][s]->get(0, 1);
+             double fxz= electricFieldGradientSet_[n][s]->get(0, 2);
+             double fyy= electricFieldGradientSet_[n][s]->get(1, 1);
+             double fyz= electricFieldGradientSet_[n][s]->get(1, 2);
+             double fzz= electricFieldGradientSet_[n][s]->get(2, 2);
+             double dij=-referenceStatisticalSet_.DensityMatrixSet[n]->get(i, j);
+             g1_x += dij * fx;
+             g1_y += dij * fy;
+             g1_z += dij * fz;
+             g2_xx += dij * fxx;
[…]

### 2018-07-27 papers/1_dmatpol/paper-r3.tex
Equation in the Appendix B updated

@@
 The gradient vector ${\bf g}$ and Hessian matrix ${\bf H}$ 
-are build from blocks associated with the type of parameters, i.e.,
+are build from blocks associated with a particular type of parameters, i.e.,
 %
@@
 %
-The gradient element of the $r$th block and Hessian element of the $(rs)$th block read
+where the block indices 1 and 2 correspond to the first\hyp{} and second\hyp{}order susceptibilities, respectively.
+Note that the second derivatives of $\delta D^{(N)}$ 
+with respect to the adjustable parameters vanish
+due to the linear functional form of Eq.~\eqref{e:final-model.General.Parameters}.
+Thus, the gradient element of the $r$th block and Hessian element of the $(rs)$th block read
 %
@@
  \begin{align}
-  g^{[r ]}_{kv}    &\equiv \frac{\partial   Z}{\partial S_{kv}^{[r]}} 
+  g^{[r ]}    &\equiv \frac{\partial   Z}{\partial S^{[r]}} 
      =-2\sum_N \overline{\delta D}^{(N)}
-               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}} \;,\\
-  H^{[rs]}_{kv,lw} &\equiv \frac{\partial^2 Z}{\partial S_{kv}^{[r]} \partial S_{lw}^{[s]}}  
+               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}} \;,\\
+  H^{[rs]} &\equiv \frac{\partial^2 Z}{\partial S^{[r]} \partial S^{[s]}}  
      = 2\sum_N 
-        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}}
-        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{lw}^{[s]}} \;.
+        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}}
+        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[s]}} \;.
  \end{align}
@@
 %
-Note that the second derivatives of $\delta D^{(N)}$ vanish
-due to the linearity with respect to the adjustable parameters.
-Thus, the explicit formulae for the gradient are
+The explicit formulae for the gradient are
 %
@@
   g^{[1]}_{kv} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \;,\\
-  g^{[2]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \widetilde{F}^{(N)}_k \;,\\
-  g^{[3]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} \widetilde{F}^{(N)}_{kv} \;.
+  g^{[2]}_{kvw} &=-2r_{vw} \sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} F^{(N)}_{kw} \;.
  \end{align}
@@
   H^{[11]}_{kv,lw} &= 2\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \;,\\
-  H^{[22]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_l \;,\\
-  H^{[33]}_{kv,lw} &= 8\sum_N \widetilde{F}^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
-  H^{[12]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_l \;,\\
-  H^{[13]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
-  H^{[23]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_{lw} \;.
+  H^{[12]}_{kv,lv'w'} &= 2r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{lv'} F^{(N)}_{lw'}  \;,\\
+  H^{[22]}_{kvw,lv'w'} &= 2r_{vw} r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{kw} F^{(N)}_{lv'} F^{(N)}_{lw'} \;.
  \end{align}
@@
 %
-In the above equations, the auxiliary quantities $\widetilde{F}^{(N)}_k$ 
-and $\widetilde{F}^{(N)}_{kv}$ are defined as
-%
-\begin{subequations}
- \begin{align}
-  \widetilde{F}^{(N)}_k    &\equiv \sum_w F^{(N)}_{kw} \;,\\
-  \widetilde{F}^{(N)}_{kv} &\equiv \sum_w F^{(N)}_{kw,kv} \;.
- \end{align}
-\end{subequations}
-%
-Note that due to the symmetry of the Hessian matrix, the blocks $21$, $31$ and $32$
-are transposes of the blocks $12$, $13$ and $23$, respectively. The composite index
-can be defined as $kv \equiv 3k+v$, for example.
+Note that due to the symmetry of the Hessian matrix, the block $21$
+is a transpose of the block $12$. 
+The composite indices $kv$ and $kvw$ are constructed from the distributed site index $k$
+and the appropriate symmetry\hyp{}corrected Cartesian component 
+of a particular DMS tensor ($v$ for first\hyp{}order,
+and $vw$, $w<v$ for second\hyp{}order susceptibility tensor, respectively).
 \bibliography{references}

### 2018-07-29 papers/1_dmatpol/paper-r3.tex
Revision

@@
 
-In this work, the electric field\hyp{}induced change in the one\hyp{}electron density
-has been expressed as a series of the one\hyp{}particle density matrix susceptibilities
-interacting with electric fields. The analytical approximate expressions were derived
+The electric field\hyp{}induced change in the one\hyp{}electron density
+has been expressed as a series of the one\hyp{}particle density matrix susceptibilities.
+The analytical approximate expressions were derived
 at the Hartree\hyp{}Fock theory, which served as a basis for construction of generalized
@@
 density matrices are available. It is shown that it is possible to accurately predict the changes
-in the one\hyp{}electron density of water molecule in spatially non\hyp{}uniform electric field distribution
-generetated by point charges. The electric field\hyp{}induced interaction energies,
-dipole moments and quadrupole moments
-are quantitatively described by the present theory in weak, moderate and even very strong electric fields (0.003 -- 0.07 a.u.).
+in the one\hyp{}electron density of water molecule in spatially uniform electric field, as well as
+in spatially non\hyp{}uniform electric field distribution
+generetated by point charges. The electric field\hyp{}induced polarization energies,
+dipole moments and quadrupole moments are quantitatively described by the present theory 
+in weak, moderate and even very strong electric fields (0.003 -- 0.07 a.u.)
+when quadratic terms with respect to the electric field are accounted for.
 It is believed that the proposed model could open
-new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems
-at arbitrary level of theory.
-%at post\hyp{}Hartree\hyp{}Fock levels of theory.
+new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
 \end{abstract}
@@
 %
-and consider electric field magnitudes in a range from very weak (~0.003 a.u.) to strong (0.07 a.u.).
+and consider electric field magnitudes in a range from very weak (~0.003 a.u.) to very strong (0.07 a.u.).
 The associated error distributions in prediction of polarization energies 
@@
 \caption{{\bf Performance of the density matrix susceptibility tensors of water molecule
-in weak, moderate and strong electric fields.}
-In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
-respectively. $R$-squared coefficients are shown in parentheses.
-Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$
-where $R_{\rm max}$ is the maximum power in the expansion
-from Eq.~\eqref{e:final-model.General}.
+in uniform and non\hyp{}uniform electric fields.$^{a,b}$}
 }
@@
 \multicolumn{15}{c}{} \\ 
-\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}} \\ 
+\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}$^{c}$} \\ 
 \multicolumn{15}{c}{} \\ \hline
@@
 M & 28.22 & (53.90) & - & (-) & - & (-) && 10.45 & (93.68) &  5.68 & (98.13) && 23.94 & (66.83) \\ 
-S & 78.07 & (40.53) & - & (-) & - & (-) && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ \hline
+S & 78.07 & (40.53) & - & (-) & - & (-) && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ 
 \bottomrule
@@
 \end{ruledtabular}
+%
+\footnotesize{\flushleft{
+$^a$ $R$-squared coefficients are shown in parentheses. `n.d.' stands for `not determined'. \\
+$^b$ Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.\\
+$^c$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields 
+that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).\\
+}}
+
 \end{table*}
@@
 in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
-This value corresponds to our moderate electric field set of samples (denoted as `M' in Tables~\ref{t:fields}
-and \ref{t:results}) for which the
-accuracy of the generalized quadratic DMS model is excellent
+This value corresponds to our moderate electric field set of samples (denoted as `M' in Tables~\ref{t:results}
+and \ref{t:fields}) for which the
+accuracy of the generalized quadratic DMS model is excellent.
 Water molecules can experience

### 2018-07-31 papers/1_dmatpol/paper-r3.tex
Figures and Tables revised.

@@
 The electric field\hyp{}induced change in the one\hyp{}electron density
-has been expressed as a series of the one\hyp{}particle density matrix susceptibilities.
+has been expressed as a series of the one\hyp{}particle density matrix susceptibilities (DMS)
+interacting with electric field.
 The analytical approximate expressions were derived
@@
 It is believed that the proposed model could open
-new routes in quantum chemistry of fast and efficient calculations of strongly perturbed systems.
+new routes in Quantum Chemistry of fast and efficient calculations of strongly perturbed systems.
 \end{abstract}
@@
 Unfortunately, it is quite difficult to efficiently treat the environment\hyp{}induced potential operator
-because of complicated nature of intermolecular interactions, that theoretically
+because of complicated nature of intermolecular interactions that theoretically
 cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
-Perhaps the most frequently used efficient and reasonably accurate \emph{ab initio} fragmentation methods,
+Perhaps one of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods,
 the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
-chemistry at the Hartree\hyp{}Fock (HF)\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
+chemistry at the Hartree\hyp{}Fock (HF)\cite{Roothaan.RevModPhys.1951,Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
 or density functional theory (DFT) levels\cite{Nguyen.Pachter.Day.JCP.2014}.
@@
 Due to all of these reasons, we adopt the following approach: 
-we first develop a qualitative first\hyp{}principles distributed\hyp{}site model at the Hartree\hyp{}Fock (HF)
+we first develop a qualitative first\hyp{}principles distributed\hyp{}site model at the Hartree\hyp{}Fock
 level of theory, and then generalize it to the arbitrary level using the functional form that was found 
@@
 %
-where ${\mathbb{M}({\bf r}_Q)}$ is AO\hyp{}representation
+where ${\mathbb{M}({\bf r}_Q)}$ is atomic orbital (AO)\hyp{}representation
 of the dipole moment operator defined with respect to origin at ${\bf r}_Q$,
@@
 %
-where ${\bf C}$ describes the occupied molecular orbitals as linear combinations
-of atomic orbitals (AO's), referred here to as the LCAO-MO matrix.
+where ${\bf C}$ describes the occupied molecular orbitals (MO's) as linear combinations
+of AO's, referred here to as the LCAO-MO matrix.
 ${\bf D}$ is an idempotent matrix and if expressed in orthogonal basis the following holds
@@
 %
-In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed polarizability tensor
-associated with the $i$th MO whereas ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge
+In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarizability tensor
+associated with the $i$th MO whereas ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge,
 ${\bf r}_i = \tbraket{\phi_i}{\hat{\bf r}}{\phi_i}$.
@@
 and after recasting the right hand side in a matrix form
+and assuming real quantities
 we get
@@
 where $\left[ {\bf L}_i  \right]^{-1}_{\rm Left}$ is a left inverse
-of ${\bf L}_i$ matrix 
+of ${\bf L}_i$ matrix,
 %
@@
       \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
-       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T 
+       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T   \;,
 \end{equation}
@@
 can be used as a guideline to construct a general model of density matrix polarization.
-Quadratic terms with respect to the electric field as well as linear terms with respect to
-electric field gradient can be added to the formulation, which leads to
+Quadratic (and higher\hyp{}order) terms with respect to the electric field
+can be added to the formulation, which leads to
 %
@@
 ${\bf B}^{(i; 1)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 2)}_{\alpha\beta}$
-is $(3\times 3)$ matrix. Also, the distributed expansion centres
+is a $(3\times 3)$ matrix. Also, the distributed expansion centres
 (labeled as $i$) are generalized to any point in the molecule.\footnote{More general
-contribution due to square of electric field could be $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$.
-However, for simplicity we neglect the contributions from different sites $i$ and $j$
+contribution due to square of electric field could 
+be $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$. 
+However, for simplicity, we neglect the contributions from different sites $i$ and $j$
 in the present discussion and leave it for future development.
@@
 be a set of $N_{\rm max}$ distinct and randomly sampled 
-spatially non\hyp{}uniform distributions of the electric field. It is assumed that
+spatial distributions of electric field. It is assumed that
 the exact difference one\hyp{}particle density matrices (with respect to the unperturbed state)
@@
 \end{pmatrix}^T
-$, we define the loss function $Z$ 
+$, we define a loss function $Z$ 
 of the three blocks of adjustable parameters
@@
 Components of spatially uniform electric fields were determined 
-from the uniform distribution in the interval of $\vert 0.02 \vert$ a.u.
+from the uniform distribution in the interval of $\vert 0.0404 \vert$ a.u.
 generating average field of 0.035 a.u. in an ensemble of 100 samples.
-Cartesian coordinates of perturbing charges were drawn from the uniform distribution of a set of fourty
-three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
+Cartesian coordinates of perturbing charges were drawn from the uniform distribution 
+of a set of fourty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
 and origin in the center of mass of water molecule.
@@
 was tuned to obtain a desired ensemble of weak, moderate and strong electric field magnitudes
-(with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly; cf. Table~\ref{t:fields}).
+(with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly).
 In this way, independent sets of 100 random samples of uniform electric fields as well as 
@@
 the electric field\hyp{}induced polarization energy, 
-the dipole moment and the quadrupole moment were analysed in detail.
+dipole moment and quadrupole moment were analysed in detail.
 The electric field\hyp{}induced polarization energy at the HF level, $\delta E^{\rm Pol}$,
@@
 the polarization of the electron density distribution, respectively. 
-In order to compare the performance of the distributed polarizability approach,
+In order to compare the performance of the usual 
+distributed dipole\hyp{}dipole polarizability approach,
 the approximate interaction energy was also evaluated by using the expression:
@@
 %
-in which ${\BM \alpha}_i$ is the distributed anisotropic polarizability tensor associated with the
+in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarizability tensor associated with the
 $i$th localized molecular orbital (LMO) centroid (see Eq.~\eqref{e:dmu-dpol}).
@@
    \begin{equation}\label{e:rmsd}
-     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 } \;;
+     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \left( \lvert \delta {\BM\upmu}^{(N)} \rvert 
+                                             - \lvert \overline{\delta {\BM\upmu}^{(N)} \rvert }\right)^2 } \;;
    \end{equation}
@@
    \begin{equation}\label{e:rmsq}
-     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta \widetilde{\BM\Theta}^{(N)} - \overline{\delta \widetilde{\BM\Theta}^{(N)}}\rvert^2 } \;.
+     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \left( \lvert \delta \widetilde{\BM\Theta}^{(N)} \rvert 
+                                             - \lvert \overline{\delta \widetilde{\BM\Theta}^{(N)} \rvert }\right)^2 } \;.
    \end{equation}
@@
   \begin{align}
-    f_u    &= \frac{f(h_u) - f(-h_u)}{2h_u} \;,\\
-    f_{uu} &= \frac{f(h_u) + f(-h_u) - 2f(0)}{h_u^2} \;,\\
-    f_{uw} &= \frac{f(h_u,h_w) + f(-h_u,-h_w) + 2f(0)}{2h_uh_w} \nonumber \\ 
-           & \; -\frac{f(h_u,0) - f(-h_u,0) - f(0,h_w) - f(0,-h_w)}{2h_uh_w} \;,
+    D_u    &= \frac{D(F_u) - D(-F_u)}{2F_u} \;,\\
+    D_{uu} &= \frac{D(F_u) + D(-F_u) - 2D^{(0)}+}{F_u^2} \;,\\
+    D_{uw} &= \frac{D(F_u,F_w) + D(-F_u,-F_w) + 2D^{(0)}}{2F_uF_w} \nonumber \\ 
+           & \; -\frac{D(F_u) + D(-F_u) + D(F_w) + D(-F_w)}{2F_uF_w} \;,
   \end{align}
@@
 where 
-the subscripts denote the Cartesian component of the electric field and $h$ is the differentiation step
-assumed to be 0.0005~a.u.
+the subscripts $u,w=x,y,z$ denote the derivatives with respect 
+to the Cartesian component of the electric field, and the finite difference
+steps $F_{u(w)}$ were assumed to be 0.0005~a.u for all components.
 
@@
 %
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\caption{\label{f:fig-1} {\bf Performance of the DMS tensors of water molecule
+in uniform electric field.} 
+$R_{\rm max}$ denotes the maximum rank of the electric field
+in the expansion from Eq.~\eqref{e:final-model.General}.
+Electric field magnitudes in a range from 0.0 to 0.07 a.u. were tested in 100 samples with randomly selected 
+directions of electric field.
+} 
+\end{figure}
+%
+
+All the tested models that are linear with respect to the electric field,
+i.e., from Eq.~\eqref{e:final-model.HF} (black solid squares in Figure~\ref{f:fig-1}) 
+as well as truncated at linear susceptibilities ${\bf B}^{(1)}$ 
+from Eqs.~\eqref{e:dD-Taylor} (blue solid circles) and \eqref{e:final-model.General} (blue open circles),
+incorrectly reproduce the polarization energy.
+Interestingly, the predicted values are in noticeable 
+and same correlation with the exact estimates with a slope of $\sim$2.
+Incorporation of the second\hyp{}order term
+obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
+or by using the generalized DMS model (red open circles)
+dramatically reduces the errors
+leading to quantitative accuracy in all tested field magnitudes 
+with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalized DMS model.
+On the other hand, the
+induced dipole and quadrupole moments are reproduced reasonably well already by
+using the first\hyp{}order DMS with $R^2$ coefficients being 
+greater than 97\% in case of dipole moments and from 69\% up to 81\% in case of quadrupole moments.
+This observation can be understood by the fact that the induced dipole
+and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
+energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
+quadratic terms in the expansion from Eq.~\eqref{e:dD-Taylor} or Eq.~\eqref{e:final-model.General}.
+
+In comparison to the dipole\hyp{}dipole polarizability approach, 
+quadratic DMS models perform only slightly worse 
+(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalized quadratic DMS
+model and the polarizability model are used, respectively).
+However, quadratic DMS models better predict the induced multipole moments with
+%work better
+%when predicting the induced multipole moments with
+$R^2$ systematically greater than 99\%. 
+Certainly, taking into account the distributed quadrupole polarizabilities
+would greatly reduce errors of the standard distributed polarizability approach.
+Nevertheless, it is clear that the second\hyp{}order
+DMS models are quantitative even at very strong uniform electric fields up to 0.07 a.u.
+%On the other hand, DMS models better predict the induced dipole and quadrupole 
+%moments even in weak fields.
+
+%
 \begin{table*}%[]
-\caption{{\bf Performance of the density matrix susceptibility tensors of water molecule
+\caption{{\bf Performance of the DMS tensors of water molecule
 in uniform and non\hyp{}uniform electric fields.$^{a,b}$}
@@
 \toprule
-\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalized DMS}}                                                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
+\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalized DMS}}
+                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
 \cline{2-8}
@@
 \textbf{}         
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$}
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1^*}$}
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=2^*}$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^c$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^d$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$$^d$}
 &
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$} 
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$} 
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^e$} 
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$$^e$} 
 &  
@@
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}              \\ %\hline
-   & 4.041 & (-) & 4.823 & (-) & 0.231 & (98.40) && 5.0121 & (-) & 0.173 & (99.10) && 0.112 & (99.63) \\ 
+   & 4.041 & (n.d) & 4.823 & (n.d.) & 0.231 & (98.40) && 5.0121 & (n.d.) & 0.173 & (99.10) && 0.112 & (99.63) \\ 
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                    \\ %\hline
@@
 \multicolumn{15}{c}{} \\ 
-\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}$^{c}$} \\ 
+\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}$^{f}$} \\ 
 \multicolumn{15}{c}{} \\ \hline
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}   \\ 
-W & 0.067 & (-)     & - & (-) & - & (-) && 0.076 & (-)     & 0.003 & (99.45) && 0.004 & (98.64)\\ 
-M & 1.072 & (-)     & - & (-) & - & (-) && 1.236 & (-)     & 0.057 & (99.08) && 0.070 & (98.58)\\ 
-S & 4.220 & (-)     & - & (-) & - & (-) && 5.147 & (-)     & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
+W & 0.067 & (n.d.)     & - & (-) & - & (-) && 0.076 & (-)     & 0.003 & (99.45) && 0.004 & (98.64)\\ 
+M & 1.072 & (n.d.)     & - & (-) & - & (-) && 1.236 & (-)     & 0.057 & (99.08) && 0.070 & (98.58)\\ 
+S & 4.220 & (n.d.)     & - & (-) & - & (-) && 5.147 & (-)     & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}  \\ 
@@
 \footnotesize{\flushleft{
-$^a$ $R$-squared coefficients are shown in parentheses. `n.d.' stands for `not determined'. \\
-$^b$ Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.\\
-$^c$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields 
+$^a$ $R$-squared coefficients are shown in parentheses. `$n.d.$' stands for `not determined'. 
+$^b$ Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ 
+where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.
+$^c$ \emph{Ab initio} HF model, Eq.~\eqref{e:final-model.HF}.
+$^d$ \emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.
+$^e$ Linear\hyp{}regression generalized model, Eq.~\eqref{e:final-model.General}.
+$^f$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields 
 that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).\\
 }}
-
 \end{table*}
-%
-\begin{figure}[h]
-\includegraphics[width=0.5\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
-\caption{\label{f:fig-1} {\bf Performance of the density matrix polarization susceptibility tensors of water molecule
-in uniform electric field.} The notation $(R=R_{\max})$ denotes the maximum rank of the electric field
-in the expansion from Eq.~\eqref{e:final-model.General}.
-Electric field magnitudes in a range from 0.0 to 0.07 a.u. were tested in 100 samples with randomly selected 
-directions of electric field.
-} 
-\end{figure}
-%
-All the tested models that are linear with respect to the electric field,
-i.e., Eq.~\eqref{e:final-model.HF} as well as Eqs.~\eqref{e:dD-Taylor} and \eqref{e:final-model.General} 
-truncated at linear susceptibilities, ${\bf B}^{(1)}$,
-incorrectly reproduce the polarization energy.
-Interestingly, the predicted values are in noticeable correlation with the exact estimates with a slope of $\sim$2.
-Incorporation of the second\hyp{}order term
-obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
-or by using the generalized DMS model (red open circles)
-dramatically reduces the errors
-leading to quantitative accuracy in all tested field magnitudes 
-with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalized DMS model.
-On the other hand, the
-induced dipole and quadrupole moments are reproduced reasonably well already by
-using the first\hyp{}order DMS, ${\bf B}^{(10)}$, with $R^2$ coefficients being 
-greater than 97\% in case of dipole moments and grater than 80\% in case of quadrupole moments.
-This observation can be understood by the fact that the induced dipole
-and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
-energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
-quadratic terms in the expansion from Eq.~\eqref{e:dD-Taylor}.
-
-In comparison to the dipole\hyp{}dipole polarizability approach, 
-Quadratic DMS models perform only slightly worse 
-(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalized quadratic DMS
-model and the polarizability model are used, respectively).
-However, quadratic DMS models better predict he induced multipole moments with
-%work better
-%when predicting the induced multipole moments with
-$R^2$ systematically greater than 99\%. 
-Certainly, taking into account the distributed quadrupole polarizabilities
-would greatly reduce errors of the standard dipole\hyp{}dipole polarizability approach.
-Nevertheless, it is clear that the second\hyp{}order
-DMS models are quantitative even at very strong uniform electric fields up to 0.07 a.u.
-%On the other hand, DMS models better predict the induced dipole and quadrupole 
-%moments even in weak fields.
 
@@
 
-To study the capability of the presented distributed models to capture 
-density matrix polarization in inhomogeneous electric fields, we analyse water molecule
+To study the capability of the distributed DMS models to capture 
+density matrix polarization in inhomogeneous electric fields, we analyse a water molecule
 surrounded by point charges.
@@
 Performance descriptors of the \emph{ab initio} and generalized 
-models of the density matrix susceptibility, as well as the conventional distributed polarizability
+models of the density matrix susceptibility, as well as 
+the conventional dipole\hyp{}dipole distributed polarizability
 approach, are shown in Table~\ref{t:results}.
 
-Similarly as in the case of uniform electric field, the linear DMS models, i.e.,
-the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} as well as
-the linear DMS model, are insufficient to correctly describe the polarization energy,
-although the correlation with benchmark is still very clear with a slope of $\sim2$
-(Figure~\ref{f:fig-2}A). Again, induced dipole moments are reproduced very well even
+Similarly as in uniform electric field, 
+the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} 
+as well as
+the linear generalized DMS model, are insufficient to correctly describe the polarization energy 
+in non\hyp{}uniform electric field,
+although the correlation with benchmark is still very clear with a slope of $\sim$2
+(Figure~\ref{f:fig-2}A). 
+Again, induced dipole moments are reproduced very well even
 in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$).
@@
 \includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-2.eps}
-\caption{\label{f:fig-2} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
+\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of water molecule in moderate
 non\hyp{}uniform electric fields.}
@@
 The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibilities
-that are able to interact with the electric field and square of electric field.
+that are able to interact with the electric field and square of the electric field.
 The particular density matrix susceptibilities were determined through the least\hyp{}squares 
@@
 For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
-sites should be further investigated, preferably with post\hyp{}Hartree\hyp{}Fock methods.
-Nevertheless, owing the capability of the proposed approach to correctly capture the 
+sites should be further investigated, preferably with post\hyp{}Hartree\hyp{}Fock methods
+and electronically excited states.
+Nevertheless, owing to the capability of the proposed approach to correctly capture the 
 environment\hyp{}induced changes in the one\hyp{}particle density matrix, 
@@
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
+Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 
+for reading and commenting on the manuscript.
 \end{acknowledgments}
@@
 
-We need to be aware that multipole integrals depend on the origin with respect to which they are evaluated.
+Multipole integrals depend on the origin with respect to which they are evaluated.
 However, the induced dipole moments defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
@@
 have to be origin independent. 
-%Let us check if this is the case.
-
 Let us start with noticing that
@@
  \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
-\end{equation}
-%
-But ${\bf C}^\dagger {\bf D}={\bf C}^\dagger {\bf C} {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
+ = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \cdot {\bf D} \right]_{i\alpha} \;.
+\end{equation}
+%
+But ${\bf C}^\dagger \cdot {\bf D}={\bf C}^\dagger \cdot {\bf C} \cdot {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
 AO basis. This implies that
@@
 Thus, one can compute dipole integrals with respect to any origin and resulting
-susceptibilities ${\bf B}_{\alpha\beta}^{(i)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
+susceptibilities ${\bf B}_{\alpha\beta}^{(i;1)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
 
@@
  \begin{align}
-  g^{[1]}_{kv} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \;,\\
-  g^{[2]}_{kvw} &=-2r_{vw} \sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} F^{(N)}_{kw} \;.
+  g^{[1]}_{ku} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{ku} \;,\\
+  g^{[2]}_{kuw} &=-2r_{uw} \sum_N \overline{\delta D}^{(N)} F^{(N)}_{ku} F^{(N)}_{kw} \;.
  \end{align}
@@
  \begin{align}
-  H^{[11]}_{kv,lw} &= 2\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \;,\\
-  H^{[12]}_{kv,lv'w'} &= 2r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{lv'} F^{(N)}_{lw'}  \;,\\
-  H^{[22]}_{kvw,lv'w'} &= 2r_{vw} r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{kw} F^{(N)}_{lv'} F^{(N)}_{lw'} \;.
+  H^{[11]}_{ku,lw} &= 2\sum_N F^{(N)}_{ku} F^{(N)}_{lw} \;,\\
+  H^{[12]}_{ku,lu'w'} &= 2r_{u'w'} \sum_N F^{(N)}_{ku} F^{(N)}_{lu'} F^{(N)}_{lw'}  \;,\\
+  H^{[22]}_{kuw,lu'w'} &= 2r_{uw} r_{u'w'} \sum_N F^{(N)}_{ku} F^{(N)}_{kw} F^{(N)}_{lu'} F^{(N)}_{lw'} \;.
  \end{align}
@@
 is a transpose of the block $12$. 
-The composite indices $kv$ and $kvw$ are constructed from the distributed site index $k$
-and the appropriate symmetry\hyp{}corrected Cartesian component 
-of a particular DMS tensor ($v$ for first\hyp{}order,
-and $vw$, $w<v$ for second\hyp{}order susceptibility tensor, respectively).
+The composite indices $ku$ and $kuw$ are constructed from the distributed site index $k$
+and the appropriate symmetry\hyp{}adapted ($w<u$) Cartesian component 
+of a particular DMS tensor: $u$ for the first\hyp{}order,
+and $uw$ for the second\hyp{}order susceptibility tensor, respectively.
+
+% -----------------------
 \bibliography{references}
+% -----------------------
 

### 2018-08-01 papers/1_dmatpol/paper-r3.tex
Revision

@@
-%% ****** Start of file aiptemplate.tex ****** %
 %%
-%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
-%%   Version 4.1 of 9 October 2009.
+%%   Version 3.1 of 31 July 2018.
 %%
-%
-% This is a template for producing documents for use with 
-% the REVTEX 4.1 document class and the AIP substyles.
-% 
-% Copy this file to another name and then work on that file.
-% That way, you always have this original template file to use.
-
 \documentclass[aip,amsmath,amssymb,reprint,floatfix]{revtex4-1}
@@
 \usepackage{hyphenat}
-%\usepackage{lipsum}
 
@@
 %\thanks{Bulak}
-\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, Wroc{\l}aw University of Technology, Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}
+\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
+Wroc{\l}aw University of Science and Technology, 
+Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}
 
@@
 has been expressed as a series of the one\hyp{}particle density matrix susceptibilities (DMS)
-interacting with electric field.
+interacting with the spatial distribution of electric field.
 The analytical approximate expressions were derived
@@
 in spatially non\hyp{}uniform electric field distribution
-generetated by point charges. The electric field\hyp{}induced polarization energies,
+generetated by point charges. When both linear and quadratic terms with respect to the electric field 
+are accounted for,
+the electric field\hyp{}induced polarization energies,
 dipole moments and quadrupole moments are quantitatively described by the present theory 
-in weak, moderate and even very strong electric fields (0.003 -- 0.07 a.u.)
-when quadratic terms with respect to the electric field are accounted for.
+in electric fields from weak to very strong in magnitude (0.003 -- 0.07 a.u.).
 It is believed that the proposed model could open
@@
 The development of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
-technique was a revolution in computational chemistry and allowed insightful theoretical studies
-of quantum mechanical effects in biological systems because the high\hyp{}level
-and costly quantum mechanical calculations
-could be combined with low\hyp{}level and cheap classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
+technique was a revolution in computational chemistry which allows insightful theoretical studies
+of quantum mechanical effects in biological systems because the costly quantum mechanical calculations
+could be combined with lower\hyp{}level and much less expensive
+classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
 Understanding physics that rules intermolecular interactions\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
@@
 Interface between QM and MM levels is usually realized by electrostatic embedding
-and only rarely includes the Pauli exclusion principle and dispersion.\cite{List.Olsen.Kongsted.PCCP.2016}
+and only rarely the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
 Unfortunately, it is quite difficult to efficiently treat the environment\hyp{}induced potential operator
-because of complicated nature of intermolecular interactions that theoretically
+due to the complicated nature of intermolecular interactions that theoretically
 cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
-Perhaps one of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods,
+One of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods,
 the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
@@
 In particular, the importance of
-one\hyp{}electron densities in chemistry is well known and acknowledged in the community,
-though still not solved exactly.\cite{Holas.March.PhysRevA.1991}
+one\hyp{}electron densities in chemistry is well known and acknowledged in the community,\cite{Holas.March.PhysRevA.1991}
+though obtaining the exact density has been an undersolved issue for most of chemical systems.
 With connection to the above, it would be incredibly insightful 
@@
 method of computation of electron deformation densities without considering the entire molecular 
-aggregate by using full quantum chemistry calculations.\cite{Horn.Yuezhi.Head-Gordon.JCP.2016,Horn.Head-Gordon.JCP.2015}
+aggregate by using full quantum chemistry calculations.\cite{Horn.Head-Gordon.JCP.2015,
+Mandado.Hermida-Ramon.JCTC.2011}
 In this work, we explore the susceptibility of the molecule's electronic density distribution
@@
 where $n$ is the size of basis set. Even relatively small errors
-in a particular matrix element could accumulate tremendously when performing two\hyp{}
+in a particular matrix element will accumulate when performing two\hyp{}
 and four\hyp{}index contractions,
@@
 an arbitrary level of theory,
-are probably not sufficient enough to accurately describe the local polarization
+are not sufficient enough to accurately describe the local polarization
 of electron density distribution when electric fields are highly non\hyp{}uniform in space.
@@
 
-It seems intuitive that $\delta {\bf D}$ should 
+$\delta {\bf D}$ should 
 be directly related to the
@@
               \delta {\bf D} \cdot {\mathbb{M}({\bf r}_Q)}
-         \right] } \;.
+         \right] } \;,
 \end{equation}
@@
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
-Note that if the total charge is conserved, the induced moments are independent on the origin.
+Note that if the total charge is conserved, the induced moments are independent of the origin.
 
@@
            &\qquad - {\bf D}^{(0)} \cdot \delta {\bf C} \cdot {\bf C}^{{(0)}\dagger} 
-                   - {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \cdot {\bf D}^{(0)} 
+                   - {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \cdot {\bf D}^{(0)}  \;.
  \label{e:dD-2}
@@
          {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
-    \right] } \;,
+    \right] } \;.
 \end{multline}
@@
 It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
-induced dipole moment is independent on the origin. 
+induced dipole moment is independent of the origin. 
 This guarantees that the theory fulfills the total charge conservation requirement.
@@
    \widetilde{\mathbb{K}}  &\equiv {\bf C}^{(0)\dagger} \cdot {\mathbb{M}} \cdot {\bf D}^{(0)} \;, \\
-   \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;.
+   \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;
  \end{align}
@@
 due to the presence of trace operation, unless the electric field is uniform.
-Let us partition the induced dipole moment into separate contributions
+Let the induced dipole moment be partitioned into separate contributions
 from the occupied MO's, i.e.,
@@
 \begin{equation} \label{e:dmu-part}
- \delta {\BM{\upmu}} = \sum_i^{\rm occ} \delta {\BM{\upmu}}_i
+ \delta {\BM{\upmu}} = \sum_i^{\rm occ} \delta {\BM{\upmu}}_i \;,
 \end{equation}
@@
                         +             {\bf B}^{(i ; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) 
-                        + \ldots \Big\}
+                        + \ldots \Big\} \;.
 \end{multline}
@@
 Note that the multiple parameter blocks 
-$S^{[1]}$ and $S^{[2]}$
+${\bf S}^{[1]}$ and ${\bf S}^{[2]}$
 appear in the first power, allowing for linear least\hyp{}squares regression.
@@
 $, we define a loss function $Z$ 
-of the three blocks of adjustable parameters
+of the two blocks of adjustable parameters
 that is subject to the least\hyp{}squares minimization,
@@
 %
-The Hessian of $Z$ computed with respect to the parameters is non\hyp{}singular
-and parameter\hyp{}independent (constant).
+The Hessian of $Z$ computed with respect to the parameters 
+is parameter\hyp{}independent (constant) and generally non\hyp{}singular
+as long as the electric fields on all distributed sites are different.
 Therefore, the exact solution for the optimal parameters is given by the Newton equation
@@
 difference density matrices, primitive induced quadrupole moments were computed from
-the induced dipole moments according to
+the induced dipole moments (given in Eq.~\eqref{e:dmu-dpol}) according to
 %
@@
     D_u    &= \frac{D(F_u) - D(-F_u)}{2F_u} \;,\\
-    D_{uu} &= \frac{D(F_u) + D(-F_u) - 2D^{(0)}+}{F_u^2} \;,\\
+    D_{uu} &= \frac{D(F_u) + D(-F_u) - 2D^{(0)}}{F_u^2} \;,\\
     D_{uw} &= \frac{D(F_u,F_w) + D(-F_u,-F_w) + 2D^{(0)}}{2F_uF_w} \nonumber \\ 
@@
   {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
-     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;. \label{e:B20}
+     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B20}
  \end{align}
@@
 %
-and consider electric field magnitudes in a range from very weak (~0.003 a.u.) to very strong (0.07 a.u.).
+and electric field magnitudes in a range from very weak (~0.003 a.u.) to very strong (0.07 a.u.).
 The associated error distributions in prediction of polarization energies 
@@
 \caption{{\bf Performance of the DMS tensors of water molecule
-in uniform and non\hyp{}uniform electric fields.$^{a,b}$}
+in uniform and non\hyp{}uniform electric field$^{a,b}$}
 }
@@
 \caption{{\bf Average electric fields in statistical sets of electrostatically perturbed states
-of water molecule surrounded by point charges.} Each set was composed of 100 samples differing in the configuration
-of 40 charges generated from a uniform distribution.
-In this Table, `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields,
-respectively.}
+of water molecule surrounded by point charges$^a$}
+}
 \label{t:fields}
@@
 \begin{tabular}{ldcdcdcd}
-\multirow{2}{*}{\textbf{Set}} & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
+\multirow{2}{*}{\textbf{Set$^b$}} & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
                                   & \multicolumn{3}{c}{\textbf{Oxygen Atom}} & \textbf{} 
@@
 \end{ruledtabular}
+%
+\footnotesize{\flushleft{
+$^a$ Each set was composed of 100 samples differing in the configuration of 40 charges generated from a uniform distribution.
+$^b$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields, respectively.\\
+}}
+
 \end{table}
@@
 The most sophisticated model studied in this work, the generalized quadratic DMS model, yields excellent predictions 
-of the polarization energies and induced multipole moments in all the electric field strengths tested.
+of the polarization energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
 The errors in predicting polarization energy are reduced by an order of magnitude when compared with
@@
 the generalized quadratic DMS model performs better than the distributed dipole\hyp{}dipole
-polarizability model, whereas in strong electric fields the situation is reversed
+polarizability model, whereas only in strong electric fields the situation is reversed
 (RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.2\%$ for the DMS and polarizability model, respectively).
-Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% in strong electric fields for polarization
-energy, and even higher $R^2$ for induced multipole moments.
+Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% even in strong electric fields for polarization
+energy, and also very high $R^2$ for induced multipole moments.
+
 Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
@@
 even up to 0.055 a.u. according to study by Reischl et~al.\cite{Reischl.Kofinger.Dellago.MolPhys.2009} 
-This electric field magnitude falls into our strong electric field regime in which the generalized quadratic
+This electric field magnitude falls into our strong electric field regime (denoted as `S')
+in which the generalized quadratic
 DMS model still provides accurate predictions (see also error distributions 
 shown in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H). 
-It is emphasized here that the current model can be easily extended by taking into account
-third\hyp{}order terms. 
+Even at the current state of development, the generalized DMS model
+is of comparable energetic accuracy as the distributed polarizability model. 
+It is emphasized here that the model can be extended by taking into account
+third\hyp{}order terms which should reduce the errors even further.
 %This however falls out of scope of this contribution and it will be studied
@@
 non\hyp{}uniform electric fields.}
-Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
-quadrupole moments (G-I). Comparison with the distributed polarizability model is also shown in this Figure
-in panels C, F and I.}
+In this Figure, `Ab Initio DMS' (panels A, D and G) 
+refers to the model from Eq.~\eqref{e:final-model.HF} and
+`Generalized DMS' (panels B, E and H) 
+refers to the generalized model with $R_{\rm max}=2$ 
+where $R_{\rm max}$ is the maximum power
+in the expansion from Eq.~\eqref{e:final-model.General}.
+%Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
+%quadrupole moments (G-I). 
+Comparison with the distributed dipole\hyp{}dipole polarizability model 
+is also shown in this Figure in panels C, F and I.}
 \end{figure*}
@@
  \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
+ = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \right]_{i\alpha} {\bf r}_Q  \;.
 \end{equation}
@@
  \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \cdot {\bf D} \right]_{i\alpha} \;.
+ = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \cdot {\bf D} \right]_{i\alpha} {\bf r}_Q \;.
 \end{equation}
@@
 {\bf H}^{[21]} & {\bf H}^{[22]}  
-\end{pmatrix} \;.
+\end{pmatrix} \;,
 \end{equation}

### 2018-08-02 papers/1_dmatpol/paper-r3.tex
Ikhusia's comments added

@@
 interacting with the spatial distribution of electric field.
-The analytical approximate expressions were derived
-at the Hartree\hyp{}Fock theory, which served as a basis for construction of generalized
+The analytical approximate expressions are derived
+at the Hartree\hyp{}Fock theory, which serves as a basis for construction of generalized
 model that can be used for an arbitrary form of wavefunction as long as the one\hyp{}particle
 density matrices are available. It is shown that it is possible to accurately predict the changes
-in the one\hyp{}electron density of water molecule in spatially uniform electric field, as well as
+in the one\hyp{}electron ground\hyp{}state density of water molecule in spatially uniform electric field, as well as
 in spatially non\hyp{}uniform electric field distribution
-generetated by point charges. When both linear and quadratic terms with respect to the electric field 
+generated by point charges. When both linear and quadratic terms with respect to the electric field 
 are accounted for,
@@
 It is believed that the proposed model could open
-new routes in Quantum Chemistry of fast and efficient calculations of strongly perturbed systems.
+new routes in Quantum Chemistry for fast and efficient calculations of molecular properties
+in condensed phases.
 \end{abstract}
@@
 
-Most of efforts to theoretically treat extended molecular aggregates are rooted in 
+Most of the efforts to theoretically treat extended molecular aggregates are rooted in 
 energy\hyp{} (or force field\hyp{})based philosophy. However,
 in the past decade there has been considerable interest in density\hyp{}based approaches
-of a wide scope and aims.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
+of a wider scope and aims.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
 Mandado.Hermida-Ramon.JCTC.2011,
@@
 though obtaining the exact density has been an undersolved issue for most of chemical systems.
-With connection to the above, it would be incredibly insightful 
-to establish a first\hyp{}principles connection bewteen the one\hyp{}particle
-density and the perturbation in the environment through introducing the
+Given that fact, it would be incredibly insightful 
+to establish a first\hyp{}principles connection between the one\hyp{}particle
+density and the perturbation in the environment through introduction of
 \emph{density matrix susceptibility} (DMS) tensors. 
@@
 Thus, such DMS could be adapted into an efficient effective 
-one\hyp{}electron fragment potential, in principle applicable to an arbitrary
+one\hyp{}electron fragment potential which in principle, is applicable to an arbitrary
 level of theory, as long as the one\hyp{}particle densities are available.
 It also means that such a technique would dramatically reduce costs of quantum chemistry calculations
-becuase the task of performing full computation on a molecular aggregate
+because the task of performing full computation on a molecular aggregate
 would reduce to much simpler matrix operations.
 
-The simplest, yet perhaps the most relevant for chemistry, type of perturbation 
+The simplest type of perturbation, yet the most relevant for chemistry, 
 is the electrostatic perturbation.
-It is worth mentioning that, at that moment, there is no simple \emph{ab initio}
-method of computation of electron deformation densities without considering the entire molecular 
+It is worth mentioning that, at the moment, there is no simple \emph{ab initio}
+method to compute electron deformation densities without considering the entire molecular 
 aggregate by using full quantum chemistry calculations.\cite{Horn.Head-Gordon.JCP.2015,
@@
 and four\hyp{}index contractions,
-and in turn, obstruct the evaluated properties with non\hyp{}negligible errors.
+and in turn, affect the evaluated properties with non\hyp{}negligible errors.
 Single\hyp{}centered expansions, like the one from Eq.~\eqref{e:dD-Taylor}, although valid for 
@@
 ${\bf r}_i = \tbraket{\phi_i}{\hat{\bf r}}{\phi_i}$.
-We can now insert the above result into Eq.~\eqref{e:dmu-l-vector} 
-and after recasting the right hand side in a matrix form
-and assuming real quantities
-we get
+By inserting the above result into Eq.~\eqref{e:dmu-l-vector},
+recasting the right hand side in a matrix form
+and assuming real quantities,
+one can separate the contributions coming from occupied MO's as
 %
@@
 %
-Note that now ${\bf L}_i$ is a matrix of size $(n \times 3)$ and $\delta{\bf c}_i$ is just the $i$th column of the
+Note that now ${\bf L}_i$ is a matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
 change in the LCAO-MO matrix due to the polarization process.
@@
       \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
-       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T   \;,
-\end{equation}
-%
-and has size $(3\times n)$. Note that the \emph{right inverse} of ${\bf L}_i$
+       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T   
+\end{equation}
+%
+with the dimension of $3\times n$. Note that the right inverse of ${\bf L}_i$
 does not exist because ${\bf L}_i \cdot {\bf L}_i^T$ is singular due to the fact that usually
@@
 ${\bf B}^{(i; 1)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 2)}_{\alpha\beta}$
-is a $(3\times 3)$ matrix. Also, the distributed expansion centres
+is a $3\times 3$ matrix. Also, the distributed expansion centres
 (labeled as $i$) are generalized to any point in the molecule.\footnote{More general
@@
 Our strategy is to treat the susceptibilities from Eq.~\eqref{e:final-model.General}
-as adjustable parameters, rather than try to derive them from first principles theory.
+as adjustable parameters, rather than try to derive them from the first principles theory.
 
@@
 Water molecule
-was selected as a test system which was subject to 
+was selected as a test system which was subjected to 
 electrostatic perturbations due to random uniform electric fields and random collections of point charges
@@
 generating average field of 0.035 a.u. in an ensemble of 100 samples.
-Cartesian coordinates of perturbing charges were drawn from the uniform distribution 
+Cartesian coordinates of point charges were drawn from the uniform distribution 
 of a set of fourty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
-and origin in the center of mass of water molecule.
+with an origin that is located in the center of mass of water molecule.
 In selection of random points, space inside the union of spheres surrounding
@@
    \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
-   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{(u-u_{\rm CM})(w-w_{\rm CM})}{\beta} \;. 
+   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{(u-u_{\rm CM})(w-w_{\rm CM})}{\beta} \;, 
   \end{align}
@@
 The first describes the ability of the model to reproduce the exact energetics of the system, 
-whereas the latter two contain crucial information on how well the deformation of the one\hyp{}electron
+whereas the following two contain crucial information on how well the deformation of the one\hyp{}electron
 density is described. For Eqs.~\eqref{e:rmsd} and \eqref{e:rmsq} the absolute magnitudes
@@
 and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
-was used. Before all the calculations, water molecule was energy\hyp{}optimized.
+was used. Before all the calculations, water molecule was energy\hyp{}optimized
+by using tight convergence criterion for energy.
 For all the calculations of the density matrix polarization susceptibility tensors
@@
 
-It is instructive to start first from the discussion of polarization under the uniform electric field
-condition because, in this case, it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
+Polarization under the uniform electric field is the simplest type of electrostatic perturbation
+whereby it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
 directly by numerical differentiation. Here we consider DMS up to second order, i.e.,
@@
 and electric field magnitudes in a range from very weak (~0.003 a.u.) to very strong (0.07 a.u.).
+The \emph{ab initio} model from Eq.~\eqref{e:final-model.HF} and
+the generalized model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
 The associated error distributions in prediction of polarization energies 
@@
 incorrectly reproduce the polarization energy.
-Interestingly, the predicted values are in noticeable 
-and same correlation with the exact estimates with a slope of $\sim$2.
+Interestingly, the predicted values are in strong
+correlation with the exact estimates and exhibit a similar slope of $\sim$2.
 Incorporation of the second\hyp{}order term
@@
 On the other hand, the
-induced dipole and quadrupole moments are reproduced reasonably well already by
+induced dipole and quadrupole moments are reproduced reasonably well by
 using the first\hyp{}order DMS with $R^2$ coefficients being 
@@
 In this work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
-at the HF level was derived for a closed shell molecule in a spatially non\hyp{}uniform
+at the HF level was derived for a closed shell molecule in a spatial
 electric field distribution. It was found that the change in the density matrix element
-is approximately proportional to the electric field evaluated at the distributed site associated with the 
+is approximately linearly proportional to the electric field evaluated at the distributed site associated with the 
 center of charge of the 
@@
 for the formulation of the generalized model that is, in principle, not limited to the HF theory,
-and takes into account higher\hyp{}order curvatures of the electric field distribution as well.
-The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibilities
-that are able to interact with the electric field and square of the electric field.
-The particular density matrix susceptibilities were determined through the least\hyp{}squares 
-linear regression with respect to the training enseble of electrostatically perturbed states.
+and takes into account also quadratic and higher\hyp{}order terms with respect to the electric field.
+The generalized model, which consisted of a set of adjustable parameters,
+definined the first\hyp{} and second\hyp{}order density matrix susceptibilities
+that were determined through the least\hyp{}squares 
+linear regression with respect to the training ensemble of electrostatically perturbed states.
 
@@
 for prediction of interaction energies, dipole moments and even quadrupole moments
-once at least quadratic terms with respect to the electric field are taken into account. 
+once, at least quadratic terms with respect to the electric field, are taken into account. 
 This work demonstrated that it is possible to

### 2018-08-02 papers/1_dmatpol/paper-r3.tex
Revision

@@
 The electric field\hyp{}induced change in the one\hyp{}electron density
-has been expressed as a series of the one\hyp{}particle density matrix susceptibilities (DMS)
-interacting with the spatial distribution of electric field.
-The analytical approximate expressions are derived
-at the Hartree\hyp{}Fock theory, which serves as a basis for construction of generalized
-model that can be used for an arbitrary form of wavefunction as long as the one\hyp{}particle
-density matrices are available. It is shown that it is possible to accurately predict the changes
-in the one\hyp{}electron ground\hyp{}state density of water molecule in spatially uniform electric field, as well as
-in spatially non\hyp{}uniform electric field distribution
-generated by point charges. When both linear and quadratic terms with respect to the electric field 
-are accounted for,
-the electric field\hyp{}induced polarization energies,
-dipole moments and quadrupole moments are quantitatively described by the present theory 
+has been expressed as a series of the one\hyp{}particle density matrix 
+susceptibilities (DMS) interacting with the spatial distribution of electric field.
+The analytical approximate expressions are derived at the Hartree\hyp{}Fock theory, 
+which serves as a basis for construction of generalized model 
+that can be used for an arbitrary form of wavefunction as long as the one\hyp{}particle
+density matrices are available. It is shown that it is possible to accurately predict 
+the changes in the one\hyp{}electron ground\hyp{}state density of water molecule 
+in spatially uniform electric field, as well as in spatially non\hyp{}uniform 
+electric field distribution generated by point charges. When both linear 
+and quadratic terms with respect to the electric field are accounted for,
+the electric field\hyp{}induced polarization energies, dipole moments 
+and quadrupole moments are quantitatively described by the present theory 
 in electric fields from weak to very strong in magnitude (0.003 -- 0.07 a.u.).
-It is believed that the proposed model could open
-new routes in Quantum Chemistry for fast and efficient calculations of molecular properties
-in condensed phases.
+It is believed that the proposed model could open new routes in Quantum Chemistry 
+for fast and efficient calculations of molecular properties in condensed phases.
 \end{abstract}
@@
 In the limit of a weak and spatially uniform electric field,
-the polarization\hyp{}induced density matrix change is given by the follwoing Taylor expansion:
+the polarization\hyp{}induced density matrix change is given by the following Taylor expansion:
 %
@@
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
-Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 
-for reading and commenting on the manuscript.
+%Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 
+%for reading and commenting on the manuscript.
 \end{acknowledgments}
@@
 (or equivalently in Eq.~\eqref{e:dmu-l-vector})
-have to be origin independent. 
-Let us start with noticing that
+have to be origin independent. Note that
 %
@@
  \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \right]_{i\alpha} {\bf r}_Q  \;.
-\end{equation}
-%
-Further we have
+ = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \right]_{i\alpha} {\bf r}_Q  
+\end{equation}
+%
+and
 %
@@
 But ${\bf C}^\dagger \cdot {\bf D}={\bf C}^\dagger \cdot {\bf C} \cdot {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
-AO basis. This implies that
+AO basis which implies that
 %
@@
 Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
-defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
-are origin independent.
+defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} are origin independent.
 Thus, one can compute dipole integrals with respect to any origin and resulting

### 2018-08-03 papers/1_dmatpol/paper-submit.tex
Submition version added

--- 2018-08-02 papers/1_dmatpol/paper-r3.tex
@@
 which serves as a basis for construction of generalized model 
-that can be used for an arbitrary form of wavefunction as long as the one\hyp{}particle
-density matrices are available. It is shown that it is possible to accurately predict 
+that can be used for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix
+as long as the latter are available by means of quantum chemistry methods. 
+It is shown that it is possible to accurately predict 
 the changes in the one\hyp{}electron ground\hyp{}state density of water molecule 
@@
 and quadrupole moments are quantitatively described by the present theory 
-in electric fields from weak to very strong in magnitude (0.003 -- 0.07 a.u.).
-It is believed that the proposed model could open new routes in Quantum Chemistry 
+in electric fields ranging from weak to very strong (0.001 -- 0.07 a.u.).
+It is believed that the proposed model could open new routes in quantum chemistry 
 for fast and efficient calculations of molecular properties in condensed phases.
@@
 of quantum mechanical effects in biological systems because the costly quantum mechanical calculations
-could be combined with lower\hyp{}level and much less expensive
+can be combined with lower\hyp{}level and much less expensive
 classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
@@
 Most of the efforts to theoretically treat extended molecular aggregates are rooted in 
-energy\hyp{} (or force field\hyp{})based philosophy. However,
+energy\hyp{}based philosophy. However,
 in the past decade there has been considerable interest in density\hyp{}based approaches
@@
 one\hyp{}electron densities in chemistry is well known and acknowledged in the community,\cite{Holas.March.PhysRevA.1991}
-though obtaining the exact density has been an undersolved issue for most of chemical systems.
+though obtaining the accurate electron density distribution has been an undersolved issue for most of chemical systems
+in condensed phases.
 Given that fact, it would be incredibly insightful 
@@
 It is worth mentioning that, at the moment, there is no simple \emph{ab initio}
-method to compute electron deformation densities without considering the entire molecular 
+method to compute polarization\hyp{}induced electron deformation densities 
+without considering the entire molecular 
 aggregate by using full quantum chemistry calculations.\cite{Horn.Head-Gordon.JCP.2015,
@@
 In the limit of a weak and spatially uniform electric field,
-the polarization\hyp{}induced density matrix change is given by the following Taylor expansion:
+the polarization\hyp{}induced density matrix change of any electronic state
+or electronic transition is given by the following Taylor expansion:
 %
@@
 %
-In the above expression, the density matrix susceptibilities are simply equal to the derivatives of the density matrix
+In the above expression, the density matrix susceptibilities 
+are simply equal to the derivatives of the density matrix
 with respect to electric field (up to a multiplicative constant).
-In a general case, i.e., when electric field is non\hyp{}uniform, the development of a first\hyp{}principles theory
-that is capable of accurately capturing the electric field\hyp{}dependence of the one\hyp{}particle density
-becomes extremely difficult. This is not only because of the very complicated (and generally unknown) solution
-to the electronic correlation problem, but also because
-a working model has to accurately predict all the $n^2$ elements of the density matrix
+In a general case, i.e., when electric field is relatively strong 
+and non\hyp{}uniform, single\hyp{}centered expansions, like the one 
+from Eq.~\eqref{e:dD-Taylor},
+%although valid for an arbitrary level of theory in uniform fields,
+are not sufficient enough to accurately describe the local polarization
+of electron density distribution.
+
+The development of a general first\hyp{}principles theory
+that is capable of accurately capturing the electric field\hyp{}dependence 
+of the one\hyp{}particle density becomes extremely difficult. This is not 
+only because of the very complicated (and generally unknown) solution
+to the electronic correlation problem, but also because a working model 
+has to accurately predict all the $n^2$ elements of the density matrix
 where $n$ is the size of basis set. Even relatively small errors
 in a particular matrix element will accumulate when performing two\hyp{}
-and four\hyp{}index contractions,
-and in turn, affect the evaluated properties with non\hyp{}negligible errors.
-Single\hyp{}centered expansions, like the one from Eq.~\eqref{e:dD-Taylor}, although valid for 
-an arbitrary level of theory,
-are not sufficient enough to accurately describe the local polarization
-of electron density distribution when electric fields are highly non\hyp{}uniform in space.
+and four\hyp{}index contractions, and in turn, affect the evaluated properties 
+with non\hyp{}negligible errors.
+
 Due to all of these reasons, we adopt the following approach: 
-we first develop a qualitative first\hyp{}principles distributed\hyp{}site model at the Hartree\hyp{}Fock
-level of theory, and then generalize it to the arbitrary level using the functional form that was found 
-by considering HF level,
-but treating the density matrix susceptibility tensors as cetrain functions of adjustable parameters further on.
+we first develop a qualitative \emph{ab initio} distributed\hyp{}site 
+model at the Hartree\hyp{}Fock level of theory. Further on, we generalize it 
+to an arbitrary level and type of density matrix
+using the functional form that was found by considering HF level as a guideline,
+but treating the density matrix susceptibility tensors as cetrain functions 
+of adjustable parameters.
 
@@
 
-$\delta {\bf D}$ should 
+$\delta {\bf D}$ from Eq.~\eqref{e:rho-1} should 
 be directly related to the
@@
 field. 
-Therefore, 
-we start our analysis by considering the induced electronic dipole moment 
-of a closed-shell molecule, 
+Therefore, we start our analysis by considering the induced 
+ground\hyp{}state electronic dipole moment of a closed\hyp{}shell molecule, 
 %
@@
 \begin{equation}\label{e:m}
- \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = -e\tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
-\end{equation}
-%
-(In this work we shall use the special open bold font for matrices of vectors, so that
+ \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = -\tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
+\end{equation}
+%
+(In this work we shall use the double struck font for matrices of vectors, so that
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
@@
 Let the induced dipole moment be partitioned into separate contributions
-from the occupied MO's, i.e.,
+from the occupied localized MO's (LMO's), i.e.,
 %
@@
 In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarizability tensor
-associated with the $i$th MO whereas ${\bf F}({\bf r}_i)$ is the electric field evaluated at the $i$th MO centroid of charge,
+associated with the $i$th LMO whereas ${\bf F}({\bf r}_i)$ 
+is the electric field evaluated at the $i$th LMO charge centroid,
 ${\bf r}_i = \tbraket{\phi_i}{\hat{\bf r}}{\phi_i}$.
@@
 %
-Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ describes the polarization affinity
-of the density matrix element associated with the $\alpha$th and $\beta$th orthogonal AO's.
+Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ 
+describes the local linear polarization affinity 
+of the density matrix element 
+associated with the $i$th LMO centroid as well as the $\alpha$th and $\beta$th orthogonal AO's.
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
-It is emphasized here that all elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
+It is emphasized here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
 are unique properties of electronically unperturbed molecule. 
-
-One very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
+However, they are valid only under the Hartree\hyp{}Fock
+approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
+limited by the flavour of the distributed polarizability approximation), 
+but not higher multipole moments and polarization energy, are correctly described.
+
+Nevertheless, one very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
 the density matrix increment is dependent \emph{only} on the susceptibility element associated with the same
@@
 vector susceptibility, different for each density matrix element. This feature will be used to
-generalize the model to an arbitrary level of theory and electric field curvature in the next section.
+generalize the model to an arbitrary level of theory and order of electric field response in the next section.
 
@@
 in the present discussion and leave it for future development.
-}
-
-\subsubsection{Determining the generalized susceptibility tensors}
-
-The exact mathematical form of the generalized density matrix susceptibilities 
+} Note that Eq.~\eqref{e:final-model.General} can refer to any type of one\hyp{}particle density matrix
+such as ground\hyp{}state, excited\hyp{}state or transition density matrix.
+However, the exact mathematical form of the generalized density matrix susceptibilities 
 from Eq.~\eqref{e:final-model.General} is unknown.
-The susceptibility from Eq.~\eqref{e:final-model.HF} is valid only under the Hartree\hyp{}Fock
-approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
-limited by the flavour of the distributed polarizability approximation), 
-but not higher multipole moments and polarization energy, are correctly described.
-Our strategy is to treat the susceptibilities from Eq.~\eqref{e:final-model.General}
-as adjustable parameters, rather than try to derive them from the first principles theory.
-
-Let  
-$\left\{ {\bf F}^{(1)}({\bf r}), {\bf F}^{(2)}({\bf r}), \ldots, {\bf F}^{(N)}({\bf r}), \ldots \right\}$ 
+
+\subsubsection{Determining the generalized susceptibility tensors by linear regression}
+
+Let $\left\{ {\bf F}^{(1)}({\bf r}), {\bf F}^{(2)}({\bf r}), \ldots, {\bf F}^{(N)}({\bf r}), \ldots \right\}$ 
 be a set of $N_{\rm max}$ distinct and randomly sampled 
@@
 elements of $B_{uw}^{(i;2)}$.
-Note that the multiple parameter blocks 
-${\bf S}^{[1]}$ and ${\bf S}^{[2]}$
+The multiple parameter blocks
+(${\bf S}^{[1]}$, ${\bf S}^{[2]}$ and so on)
 appear in the first power, allowing for linear least\hyp{}squares regression.
@@
 
-To determine the optimum set 
+To determine the optimum set,
 $
@@
 {\bf S}^{[1]} &
-{\bf S}^{[2]} 
+{\bf S}^{[2]} & \cdots
 \end{pmatrix}^T
 $, we define a loss function $Z$ 
-of the two blocks of adjustable parameters
 that is subject to the least\hyp{}squares minimization,
@@
 \begin{equation}\label{e:Z}
- Z[{\bf S}] = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
+ Z({\bf S}) = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
 \end{equation}
@@
 where ${\bf g}$ and ${\bf H}$ are the gradient vector and the Hessian matrix, respectively.
-The explicit forms of the gradient and Hessian are given in the Appendix~\ref{a:blocks}.
-Note that the dimensions of parameter space for the block 1 and 2 are
+The explicit forms of the gradient and Hessian up to second\hyp{}order 
+are given in the Appendix~\ref{a:blocks}.
+Note that in this case the dimensions of parameter space for the block 1 and 2 are
 equal to $3M$ and $6M$, respectively.
@@
 was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-Water molecule
-was selected as a test system which was subjected to 
+Water molecule was selected as a test system which was subjected to 
 electrostatic perturbations due to random uniform electric fields and random collections of point charges
@@
 Components of spatially uniform electric fields were determined 
-from the uniform distribution in the interval of $\vert 0.0404 \vert$ a.u.
-generating average field of 0.035 a.u. in an ensemble of 100 samples.
+from the uniform distribution in the interval 
+of $\vert 0.0404 \vert$ a.u. generating average field of 0.035 a.u. in an ensemble of 100 samples.
 Cartesian coordinates of point charges were drawn from the uniform distribution 
@@
 Electric charge values were selected from the one\hyp{}dimensional uniform distribution
-of double precision numbers in the interval of $\vert q_{\rm max} \vert$, where $q_{\rm max}$
+in the interval of $\vert q_{\rm max} \vert$, where $q_{\rm max}$
 was tuned to obtain a desired ensemble of weak, moderate and strong electric field magnitudes
 (with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly).
-In this way, independent sets of 100 random samples of uniform electric fields as well as 
-40 point charges were selected and exact solutions of HF\hyp{}SCF equations were 
+In this way, independent sets of 100 random samples of either uniform electric field or
+point charge distributions were selected and exact solutions of HF\hyp{}SCF equations were 
 performed for each validation procedure.
+In the case of non\hyp{}uniform electric fields, the distribution centres 
+of the \emph{ab initio} DMS model (Eq.~\eqref{e:final-model.HF}) were assumed 
+to be LMO centroids, while atoms were used for the generalized DMS models (Eq.~\eqref{e:final-model.General}).
+This is because various LMO centroids are often located at the same point which would
+introduce linear dependencies in the Hessian matrix from Eq.~\eqref{e:Newton.gH}.
 
@@
 In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
-the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
+the one\hyp{}electron core Hamiltonian matrix of an unperturbed HF ground state,
 ${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
 and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
-the polarization of the electron density distribution, respectively. 
+the polarization of the electron density, respectively. 
 In order to compare the performance of the usual 
@@
 in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarizability tensor associated with the
-$i$th localized molecular orbital (LMO) centroid (see Eq.~\eqref{e:dmu-dpol}).
+$i$th LMO centroid (see Eq.~\eqref{e:dmu-dpol}).
 The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
@@
 The first describes the ability of the model to reproduce the exact energetics of the system, 
-whereas the following two contain crucial information on how well the deformation of the one\hyp{}electron
+whereas the following two contain additional information on how well the deformation of the one\hyp{}electron
 density is described. For Eqs.~\eqref{e:rmsd} and \eqref{e:rmsq} the absolute magnitudes
@@
 was used. Before all the calculations, water molecule was energy\hyp{}optimized
-by using tight convergence criterion for energy.
+with a maximum energy, force and displacement thresholds of 10$^{-6}$, 3$\times$$10^{-4}$ 
+and 1.2$\times$$10^{-3}$ a.u., respectively.
 For all the calculations of the density matrix polarization susceptibility tensors
@@
 where 
-the subscripts $u,w=x,y,z$ denote the derivatives with respect 
+the subscripts $u,w=x,y,z$ on the right\hyp{}hand sides denote the derivatives with respect 
 to the Cartesian component of the electric field, and the finite difference
@@
 %
-and electric field magnitudes in a range from very weak (~0.003 a.u.) to very strong (0.07 a.u.).
+and electric field magnitudes in a range from very weak ($<$0.001 a.u.) to very strong ($\sim$0.07 a.u.).
 The \emph{ab initio} model from Eq.~\eqref{e:final-model.HF} and
@@
 Interestingly, the predicted values are in strong
-correlation with the exact estimates and exhibit a similar slope of $\sim$2.
+correlation with the exact estimates and exhibit similar slope of $\sim$2.
 Incorporation of the second\hyp{}order term
@@
 greater than 97\% in case of dipole moments and from 69\% up to 81\% in case of quadrupole moments.
-This observation can be understood by the fact that the induced dipole
+This observation can be understood given the fact that the induced dipole
 and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
@@
 Nevertheless, it is clear that the second\hyp{}order
-DMS models are quantitative even at very strong uniform electric fields up to 0.07 a.u.
+DMS models are quantitative even at very strong uniform electric fields.
 %On the other hand, DMS models better predict the induced dipole and quadrupole 
@@
 To study the capability of the distributed DMS models to capture 
-density matrix polarization in inhomogeneous electric fields, we analyse a water molecule
-surrounded by point charges.
+density matrix polarization in inhomogeneous electric fields, a water molecule
+surrounded by point charges is analysed.
 Three different types of point charge environment that generate
@@
 (Figure~\ref{f:fig-2}A). 
-Again, induced dipole moments are reproduced very well even
+Again, induced dipole moments are reproduced very well also
 in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$).
@@
 (RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.2\%$ for the DMS and polarizability model, respectively).
-Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% even in strong electric fields for polarization
+Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% 
+even in strong electric fields for polarization
 energy, and also very high $R^2$ for induced multipole moments.
@@
 shown in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H).
-Even at the current state of development, the generalized DMS model
+
+Even at the current state of development, the generalized quadratic DMS model
 is of comparable energetic accuracy as the distributed polarizability model. 
-It is emphasized here that the model can be extended by taking into account
-third\hyp{}order terms which should reduce the errors even further.
+It is emphasized here that the model can be easily extended by taking into account
+third\hyp{}order terms as well if fine accuracy is required.
 %This however falls out of scope of this contribution and it will be studied
 %elsewhere. 
+In addition, any type of one\hyp{}particle density matrix can be treated by using
+generalized DMS models, which is much more difficult to realize when a conventional multipole
+model is used.
 Therefore, it is believed that
@@
 interaction energies, dipole moments and quadrupole moments of water molecule that
-interacts with uniform electric fields as well as 
-various point charge environments generating electric fields of weak, moderate
+is subjected to perturbation due to uniform electric field, as well as 
+various point charge environments generating non\hyp{}uniform electric field of weak, moderate
 and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
-interaction energies, the generalized model can reach quantitative level of accuracy
+interaction energies, the atom\hyp{}distributed generalized model can reach quantitative level of accuracy
 for prediction of interaction energies, dipole moments and even quadrupole moments
@@
 should be studied in the future with an emphasis 
-on their further optimization.
-%, especially to reach RMSE below 1 kcal/mole
-%in strong and very strong electric fields (0.05-0.07 a.u.). 
+on their further optimization and application for molecular aggregates.
 For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed

### 2018-08-03 papers/1_dmatpol/do-submit.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+base=paper-submit
+
+latex  ${base}.tex
+bibtex ${base}.aux
+dvipdf ${base}.dvi

### 2018-08-06 papers/1_dmatpol/paper-submit.tex
Revisiom papers added

@@
 which serves as a basis for construction of generalized model 
-that can be used for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix
-as long as the latter are available by means of quantum chemistry methods. 
+that is designed for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix.
 It is shown that it is possible to accurately predict 
@@
 electronic structure calculations grows extremely fast with the number of particles, greatly limiting
-the use of the state\hyp{}of\hyp{}the\hyp{}art first\hyp{}principles theories 
-when large and complex chemical systems such as
-solutions and biomolecules come into play.\cite{Tomasi.Mennucci.Cammi.ChemRev.2005}
-Due to this reason, it has been of particular importance nowadays 
-to develop bridging theoretical techniques that link
-extremely accurate but localized pieces of information 
-with certain knowledge of an atomistic organisation and intermolecular interaction
-on a relatively large spatial scales, typically ranging beyond few nanometers.
-
-The development of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM)
-technique was a revolution in computational chemistry which allows insightful theoretical studies
-of quantum mechanical effects in biological systems because the costly quantum mechanical calculations
-can be combined with lower\hyp{}level and much less expensive
-classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
-Understanding physics that rules intermolecular interactions\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
-helped to develop
-a wide family of force fields, and served as a basis for embedding Hamiltonian
-within an effective environment and fragmentation methods.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
-
-However, even with the modern advances in computational chemistry, 
-accurate description of phenomena due to complex heterogeneous molecular environments 
-remains a challenge in the following areas: (i) describing the electronically excited states 
-(electronic solvatochromism),\cite{Barbati.JACS.2014,
-Szabla.Sponer.Jiri.Gora.JPCL.2015,
-Bednarska.Zalesny.Tian.Murugan.Agren.Bartkowiak.Molecules.2017,
+the use of the state\hyp{}of\hyp{}the\hyp{}art first\hyp{}principles theories when large and complex 
+chemical systems such as solutions and biomolecules come into play.\cite{
+Xu.Guidez.Bertoni.Gordon.JCP.2018,Tomasi.Mennucci.Cammi.ChemRev.2005}
+Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques 
+that link extremely accurate but localized pieces of information with certain knowledge of an atomistic 
+organisation and intermolecular interaction on a relatively large spatial scales, typically ranging beyond 
+few nanometers.
+
+The development of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM) technique 
+was a revolution in computational chemistry which allows insightful theoretical studies of quantum mechanical
+effects in biological systems because costly quantum mechanical calculations can be combined with lower\hyp{}level 
+and much less expensive classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
+Fruitful incorporation of accurate \emph{ab initio} methods with extensive experimental and semi\hyp{}empirical
+refinings led to development of a wide variety of force fields,\cite{Demerdash.Yap.T-Head-Gordon.AnnuRevPhysChem.2014}
+whereas sophisticated partitioning of Hamiltonian or electron density enabled numerous schemes for efficient fragmentation
+within a QM level.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
+%Gao.Truhlar.Wang.Mazack.Loffler.Provorse.Rehak.ACR.2014}
+
+However, even with the modern advances in computational chemistry, accurate description of phenomena due 
+to complex heterogeneous molecular environments remains a challenge in the following areas: (i) describing 
+the electronically excited states (electronic solvatochromism),\cite{Barbati.JACS.2014,
+Szabla.Sponer.Jiri.Gora.JPCL.2015,Bednarska.Zalesny.Tian.Murugan.Agren.Bartkowiak.Molecules.2017,
 Jedrzejewska.Grabarz.Bartkowiak.Osmialowski.SpectChimActA.2018} 
 (ii) understanding the vibrational properties of molecules in ground and electronically excited states
-(vibrational or vibronic solvatochromisms).\cite{Blasiak.Londergan.Webb.Cho.ACR.2017} 
-Interface between QM and MM levels is usually realized by electrostatic embedding
-and only rarely the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
+(vibrational or vibronic solvatochromisms).\cite{Blasiak.Londergan.Webb.Cho.ACR.2017,
+Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018} 
+Interface between QM and MM levels is usually realized by electrostatic embedding and only rarely 
+the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
 Unfortunately, it is quite difficult to efficiently treat the environment\hyp{}induced potential operator
-due to the complicated nature of intermolecular interactions that theoretically
-cannot be solely envisioned as a sum of electrostatic, dispersion and Pauli effects.
-One of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods,
-the effective fragment potential (EFP) methods, are limited to the ground\hyp{}state
-chemistry at the Hartree\hyp{}Fock (HF)\cite{Roothaan.RevModPhys.1951,Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
-or density functional theory (DFT) levels\cite{Nguyen.Pachter.Day.JCP.2014}.
+due to complicated nature of intermolecular interactions that theoretically cannot be solely envisioned 
+as a sum of separate contributions coming from electrostatic, dispersion and Pauli 
+effects.\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
+For example,
+one of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods, 
+the effective fragment potential (EFP) methods, are limited to the Hartree\hyp{}Fock 
+(HF)\cite{Roothaan.RevModPhys.1951,Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
+or density functional theory (DFT) levels\cite{Hohenberg.Kohn.PhysRev.1964,Kohn.Sham.PhysRev.1965,Nguyen.Pachter.Day.JCP.2014}.
 Unfortunately, while they can approximately describe the environment composed of molecules in their ground states,
@@
 
-Most of the efforts to theoretically treat extended molecular aggregates are rooted in 
-energy\hyp{}based philosophy. However,
-in the past decade there has been considerable interest in density\hyp{}based approaches
-of a wider scope and aims.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
-Mandado.Hermida-Ramon.JCTC.2011,
-Sun.Chan.ACR.2016,
-Hedegard.Reiher.JCTC.2016} 
-In particular, the importance of
-one\hyp{}electron densities in chemistry is well known and acknowledged in the community,\cite{Holas.March.PhysRevA.1991}
-though obtaining the accurate electron density distribution has been an undersolved issue for most of chemical systems
-in condensed phases.
-Given that fact, it would be incredibly insightful 
-to establish a first\hyp{}principles connection between the one\hyp{}particle
-density and the perturbation in the environment through introduction of
-\emph{density matrix susceptibility} (DMS) tensors. 
-The main postulate of the present work is that the associated change 
-in the one\hyp{}particle density matrix, $\delta {\bf D}$, can be
-represented in a product form of the generalized perturbation susceptibility ${\bf B}_i$
-and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$,
-evaluated at certain site $i$.
+Most of the efforts to theoretically treat extended molecular aggregates are rooted in energy\hyp{}based 
+philosophy. However, in the past two decade there has been considerable interest in density\hyp{}based approaches
+of a wider scope and aims.\cite{Yang.Lee.JCP.1995,
+Mukamel.Berman.JCP.2003,
+Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
+Elliott.Burke.Cohen.Wasserman.PhysRevA.2010,
+Mandado.Hermida-Ramon.JCTC.2011,Sun.Chan.ACR.2016,Hedegard.Reiher.JCTC.2016} 
+In particular, the importance of one\hyp{}electron densities in chemistry is well known and acknowledged 
+in the community,\cite{Holas.March.PhysRevA.1991,Yang.PhysRevLett.1991,Yang.Lee.JCP.1995} 
+though obtaining the accurate electron density distribution 
+has been an undersolved issue for most of chemical systems in condensed phases. Given that fact, it would be 
+incredibly insightful to establish a first\hyp{}principles connection between the one\hyp{}particle density 
+and the perturbation in the environment through introduction of \emph{density matrix susceptibility} (DMS) tensors. 
+The main postulate of the present work is that the associated change in the one\hyp{}particle density matrix, 
+$\delta {\bf D}$, can be represented in a product form of the generalized perturbation susceptibility ${\bf B}_i$
+and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$, evaluated at certain site $i$,
 %
@@
 %
-While $\delta {\bf P}_i$ contains critical information 
-about the actual distortion of the electronic cloud,
-perturbation susceptibility is assumed to be a property of unperturbed 
-molecule and is to be precomputed only once. 
-Thus, such DMS could be adapted into an efficient effective 
-one\hyp{}electron fragment potential which in principle, is applicable to an arbitrary
-level of theory, as long as the one\hyp{}particle densities are available.
-It also means that such a technique would dramatically reduce costs of quantum chemistry calculations
-because the task of performing full computation on a molecular aggregate
-would reduce to much simpler matrix operations.
-
-The simplest type of perturbation, yet the most relevant for chemistry, 
-is the electrostatic perturbation.
-It is worth mentioning that, at the moment, there is no simple \emph{ab initio}
-method to compute polarization\hyp{}induced electron deformation densities 
-without considering the entire molecular 
-aggregate by using full quantum chemistry calculations.\cite{Horn.Head-Gordon.JCP.2015,
+While $\delta {\bf P}_i$ contains critical information about the actual distortion of the electronic cloud,
+perturbation susceptibility is assumed to be a property of unperturbed molecule and is to be precomputed only once. 
+Thus, such DMS could be adapted into an efficient effective one\hyp{}electron fragment potential which in principle, 
+is applicable to an arbitrary level of theory, as long as the one\hyp{}particle densities are available by means of
+quantum chemistry methods. It also means that such a technique would dramatically reduce costs of quantum chemistry 
+calculations because the task of performing full computation on a molecular aggregate would reduce to much simpler 
+matrix operations.
+
+The simplest type of perturbation, yet the most relevant for chemistry, is the electrostatic perturbation.
+It is worth mentioning that, at the moment, there is no simple \emph{ab initio} method to compute 
+polarization\hyp{}induced electron deformation densities without considering the entire molecular 
+aggregate by using full quantum chemistry calculations.\cite{
+Ma.Politzer.JCP.2004,
+Horn.Head-Gordon.JCP.2015,
 Mandado.Hermida-Ramon.JCTC.2011}
-In this work, we explore the susceptibility of the molecule's electronic density distribution
+In this work, we explore the susceptibility of the molecule's electronic density
 to polarize due to external electrostatic perturbation from first principles. 
@@
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}   \\ 
-W & 0.067 & (n.d.)     & - & (-) & - & (-) && 0.076 & (-)     & 0.003 & (99.45) && 0.004 & (98.64)\\ 
-M & 1.072 & (n.d.)     & - & (-) & - & (-) && 1.236 & (-)     & 0.057 & (99.08) && 0.070 & (98.58)\\ 
-S & 4.220 & (n.d.)     & - & (-) & - & (-) && 5.147 & (-)     & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
+W & 0.067 & (n.d.)  & - &  -  & - &  -  && 0.076 & (n.d.)  & 0.003 & (99.45) && 0.004 & (98.64) \\ 
+M & 1.072 & (n.d.)  & - &  -  & - &  -  && 1.236 & (n.d.)  & 0.057 & (99.08) && 0.070 & (98.58) \\ 
+S & 4.220 & (n.d.)  & - &  -  & - &  -  && 5.147 & (n.d.)  & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}  \\ 
-W &  3.36 & (98.71) & - & (-) & -  & (-) &&  1.03 & (99.88) &  0.94 & (99.90) &&  3.36 & (98.71) \\ 
-M & 14.72 & (98.50) & - & (-) & -  & (-) &&  7.72 & (99.59) &  4.55 & (99.86) && 14.72 & (98.50)\\ 
-S & 44.80 & (96.92) & - & (-) & -  & (-) && 33.06 & (98.33) & 15.96 & (99.61) && 44.80 & (96.92) \\ \hline
+W &  3.36 & (98.71) & - &  -  & - &  -  &&  1.03 & (99.88) &  0.94 & (99.90) &&  3.36 & (98.71) \\ 
+M & 14.72 & (98.50) & - &  -  & - &  -  &&  7.72 & (99.59) &  4.55 & (99.86) && 14.72 & (98.50) \\ 
+S & 44.80 & (96.92) & - &  -  & - &  -  && 33.06 & (98.33) & 15.96 & (99.61) && 44.80 & (96.92) \\ \hline
 \multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}} \\ 
-W &  6.22 & (58.72) & - & (-) & - & (-) &&  1.59 & (97.32) &  1.42 & (97.86) &&  5.45 & (68.29) \\ 
-M & 28.22 & (53.90) & - & (-) & - & (-) && 10.45 & (93.68) &  5.68 & (98.13) && 23.94 & (66.83) \\ 
-S & 78.07 & (40.53) & - & (-) & - & (-) && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ 
+W &  6.22 & (58.72) & - &  -  & - &  -  &&  1.59 & (97.32) &  1.42 & (97.86) &&  5.45 & (68.29) \\ 
+M & 28.22 & (53.90) & - &  -  & - &  -  && 10.45 & (93.68) &  5.68 & (98.13) && 23.94 & (66.83) \\ 
+S & 78.07 & (40.53) & - &  -  & - &  -  && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ 
 \bottomrule
@@
 as well as
-the linear generalized DMS model, are insufficient to correctly describe the polarization energy 
+the generalized linear DMS model, are insufficient to correctly describe the polarization energy 
 in non\hyp{}uniform electric field,
@@
 This electric field magnitude falls into our strong electric field regime (denoted as `S')
-in which the generalized quadratic
-DMS model still provides accurate predictions (see also error distributions 
+in which the generalized quadratic DMS model still provides accurate predictions (see also error distributions 
 shown in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H).

### 2018-08-07 papers/1_dmatpol/submit/paper-submit.tex
Submission level introduced
Reports and figures added to submission platform

--- 2018-08-06 papers/1_dmatpol/paper-submit.tex
@@
 
-\title{One-Particle Density Matrix Polarization Susceptibility Tensors}
+\title{One-Particle Density Matrix Polarisation Susceptibility Tensors}
 
@@
 susceptibilities (DMS) interacting with the spatial distribution of electric field.
-The analytical approximate expressions are derived at the Hartree\hyp{}Fock theory, 
-which serves as a basis for construction of generalized model 
+The analytic approximate expressions are derived at the Hartree\hyp{}Fock theory, 
+which serves as a basis for construction of generalised model 
 that is designed for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix.
@@
 and quadratic terms with respect to the electric field are accounted for,
-the electric field\hyp{}induced polarization energies, dipole moments 
+the electric field\hyp{}induced polarisation energies, dipole moments 
 and quadrupole moments are quantitatively described by the present theory 
@@
 Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques 
-that link extremely accurate but localized pieces of information with certain knowledge of an atomistic 
+that link extremely accurate but localised pieces of information with certain knowledge of an atomistic 
 organisation and intermolecular interaction on a relatively large spatial scales, typically ranging beyond 
@@
 Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018} 
-Interface between QM and MM levels is usually realized by electrostatic embedding and only rarely 
+Interface between QM and MM levels is usually realised by electrostatic embedding and only rarely 
 the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
@@
 Most of the efforts to theoretically treat extended molecular aggregates are rooted in energy\hyp{}based 
-philosophy. However, in the past two decade there has been considerable interest in density\hyp{}based approaches
+philosophy. However, in the past two decades there has been considerable interest in density\hyp{}based approaches
 of a wider scope and aims.\cite{Yang.Lee.JCP.1995,
@@
 though obtaining the accurate electron density distribution 
-has been an undersolved issue for most of chemical systems in condensed phases. Given that fact, it would be 
+has been an unresolved issue for most of chemical systems in condensed phases. Given that fact, it would be 
 incredibly insightful to establish a first\hyp{}principles connection between the one\hyp{}particle density 
@@
 The main postulate of the present work is that the associated change in the one\hyp{}particle density matrix, 
-$\delta {\bf D}$, can be represented in a product form of the generalized perturbation susceptibility ${\bf B}_i$
+$\delta {\bf D}$, can be represented in a product form of the generalised perturbation susceptibility ${\bf B}_i$
 and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$, evaluated at certain site $i$,
@@
 It is worth mentioning that, at the moment, there is no simple \emph{ab initio} method to compute 
-polarization\hyp{}induced electron deformation densities without considering the entire molecular 
+polarisation\hyp{}induced electron deformation densities without considering the entire molecular 
 aggregate by using full quantum chemistry calculations.\cite{
@@
 In this work, we explore the susceptibility of the molecule's electronic density
-to polarize due to external electrostatic perturbation from first principles. 
+to polarise due to external electrostatic perturbation from first principles. 
 The deformation density is expressed in terms
-of the one\hyp{}particle density matrix polarization susceptibility tensors 
+of the one\hyp{}particle density matrix polarisation susceptibility tensors 
 that are properties of isolated molecules and can be
@@
 (scalar) is beneficial since it provides the directional tensor description
-of the polarization that could be combined with other effects such as Pauli repulsion 
+of the polarisation that could be combined with other effects such as Pauli repulsion 
 or dispersion,\cite{Mandado.Hermida-Ramon.JCTC.2011}
-in contrast to the standard interaction energy approaches. We show that accurate polarization 
+in contrast to the standard interaction energy approaches. We show that accurate polarisation 
 of the one\hyp{}particle density
-matrix is achievable and can be realized as long as the one\hyp{}particle density matrices
+matrix is achievable and can be realised as long as the one\hyp{}particle density matrices
 are available. It is believed that the proposed method
@@
 alternative approach in fragment\hyp{}based
-modelling of extended molecular aggregates that is not limitted to the ground state chemistry.
+modelling of extended molecular aggregates that is not limited to the ground state chemistry.
 
@@
 at the HF approximation\cite{Roothaan.RevModPhys.1951} 
-in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalization to the 
+in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalisation to the 
 arbitrary one\hyp{}electron density with electronic correlation which is derived in Section~\ref{s:2}. 
@@
 %
-In Eq.~\eqref{e:rho} $\phi_\alpha({\bf r})$ is a basis funtion, which can be for instance 
+In Eq.~\eqref{e:rho} $\phi_\alpha({\bf r})$ is a basis function, which can be for instance 
 an atomic orbital (AO) or a molecular orbital (MO), and $D_{\alpha\beta}$ is the corresponding
@@
 excited state one\hyp{}electron densities must be excited state solutions of the Schr{\"o}dinger equation
-with the exact energy\hyp{}minimized ground state.
+with the exact energy\hyp{}minimised ground state.
 
 In the limit of a weak and spatially uniform electric field,
-the polarization\hyp{}induced density matrix change of any electronic state
+the polarisation\hyp{}induced density matrix change of any electronic state
 or electronic transition is given by the following Taylor expansion:
@@
 In a general case, i.e., when electric field is relatively strong 
-and non\hyp{}uniform, single\hyp{}centered expansions, like the one 
+and non\hyp{}uniform, single\hyp{}centred expansions, like the one 
 from Eq.~\eqref{e:dD-Taylor},
 %although valid for an arbitrary level of theory in uniform fields,
-are not sufficient enough to accurately describe the local polarization
+are not sufficient enough to accurately describe the local polarisation
 of electron density distribution.
@@
 we first develop a qualitative \emph{ab initio} distributed\hyp{}site 
-model at the Hartree\hyp{}Fock level of theory. Further on, we generalize it 
+model at the Hartree\hyp{}Fock level of theory. Further on, we generalise it 
 to an arbitrary level and type of density matrix
 using the functional form that was found by considering HF level as a guideline,
-but treating the density matrix susceptibility tensors as cetrain functions 
+but treating the density matrix susceptibility tensors as certain functions 
 of adjustable parameters.
@@
 be directly related to the
-polarization\hyp{}induced multipole moments because they are
+polarisation\hyp{}induced multipole moments because they are
 a measure of the distortion of the electron density distribution due to the external electric
@@
 %
-The change in the molecular orbitals can be parameterized as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
+The change in the molecular orbitals can be parameterised as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
 by defining the idempotency\hyp{}preserving variation
@@
 induced dipole moment is independent of the origin. 
-This guarantees that the theory fulfills the total charge conservation requirement.
+This guarantees that the theory fulfils the total charge conservation requirement.
 Therefore,
@@
 Let the induced dipole moment be partitioned into separate contributions
-from the occupied localized MO's (LMO's), i.e.,
+from the occupied localised MO's (LMO's), i.e.,
 %
@@
 %
-In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarizability tensor
+In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarisability tensor
 associated with the $i$th LMO whereas ${\bf F}({\bf r}_i)$ 
@@
 Note that now ${\bf L}_i$ is a matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
-change in the LCAO-MO matrix due to the polarization process.
+change in the LCAO-MO matrix due to the polarisation process.
 The solution for $\delta {\bf C}$ is
@@
 induced dipole moment of the molecule.
-The change in the LCAO\hyp{}MO coefficient is parameterized as
+The change in the LCAO\hyp{}MO coefficient is parameterised as
 %
@@
 %
-where the \emph{density matrix polarization susceptibility tensor} is
+where the \emph{density matrix polarisation susceptibility tensor} is
 %
@@
 Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ 
-describes the local linear polarization affinity 
+describes the local linear polarisation affinity 
 of the density matrix element 
@@
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
-It is emphasized here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
+It is emphasised here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
 are unique properties of electronically unperturbed molecule. 
@@
 approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
-limited by the flavour of the distributed polarizability approximation), 
-but not higher multipole moments and polarization energy, are correctly described.
+limited by the flavour of the distributed polarisability approximation), 
+but not higher multipole moments and polarisation energy, are correctly described.
 
 Nevertheless, one very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
-the density matrix increment is dependent \emph{only} on the susceptibility element associated with the same
+the density matrix increment is dependent only on the susceptibility element associated with the same
 pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
 vector susceptibility, different for each density matrix element. This feature will be used to
-generalize the model to an arbitrary level of theory and order of electric field response in the next section.
-
-\subsection{Generalized Density Matrix Susceptibility Model}
+generalise the model to an arbitrary level of theory and order of electric field response in the next section.
+
+\subsection{Generalised Density Matrix Susceptibility Model}
 
@@
 Eq.~\eqref{e:final-model.HF}, which is the final result of the previous section, 
-can be used as a guideline to construct a general model of density matrix polarization.
+can be used as a guideline to construct a general model of density matrix polarisation.
 Quadratic (and higher\hyp{}order) terms with respect to the electric field
@@
 is a $3\times 3$ matrix. Also, the distributed expansion centres
-(labeled as $i$) are generalized to any point in the molecule.\footnote{More general
+(labelled as $i$) are generalised to any point in the molecule.\footnote{More general
 contribution due to square of electric field could 
@@
 such as ground\hyp{}state, excited\hyp{}state or transition density matrix.
-However, the exact mathematical form of the generalized density matrix susceptibilities 
+However, the exact mathematical form of the generalised density matrix susceptibilities 
 from Eq.~\eqref{e:final-model.General} is unknown.
 
-\subsubsection{Determining the generalized susceptibility tensors by linear regression}
+\subsubsection{Determining the generalised susceptibility tensors by linear regression}
 
@@
 %
-are known for each sample (overline symbolizes the exact estimate).
+are known for each sample (overline symbolises the exact estimate).
 Now, for each pair of the AO indices we construct the following 
-parameterization of Eq.~\eqref{e:final-model.General}:
+parameterisation of Eq.~\eqref{e:final-model.General}:
 %
@@
 $, we define a loss function $Z$ 
-that is subject to the least\hyp{}squares minimization,
+that is subject to the least\hyp{}squares minimisation,
 %
@@
 Cartesian coordinates of point charges were drawn from the uniform distribution 
-of a set of fourty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
-with an origin that is located in the center of mass of water molecule.
+of a set of forty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
+with an origin that is located in the centre of mass of water molecule.
 In selection of random points, space inside the union of spheres surrounding
@@
 of the \emph{ab initio} DMS model (Eq.~\eqref{e:final-model.HF}) were assumed 
-to be LMO centroids, while atoms were used for the generalized DMS models (Eq.~\eqref{e:final-model.General}).
+to be LMO centroids, while atoms were used for the generalised DMS models (Eq.~\eqref{e:final-model.General}).
 This is because various LMO centroids are often located at the same point which would
@@
 In order to validate all the models discussed in this contribution,
-the electric field\hyp{}induced polarization energy, 
+the electric field\hyp{}induced polarisation energy, 
 dipole moment and quadrupole moment were analysed in detail.
-The electric field\hyp{}induced polarization energy at the HF level, $\delta E^{\rm Pol}$,
+The electric field\hyp{}induced polarisation energy at the HF level, $\delta E^{\rm Pol}$,
 was computed from the following exact formula:
@@
 and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
-the polarization of the electron density, respectively. 
+the polarisation of the electron density, respectively. 
 In order to compare the performance of the usual 
-distributed dipole\hyp{}dipole polarizability approach,
+distributed dipole\hyp{}dipole polarisability approach,
 the approximate interaction energy was also evaluated by using the expression:
@@
 %
-in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarizability tensor associated with the
+in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarisability tensor associated with the
 $i$th LMO centroid (see Eq.~\eqref{e:dmu-dpol}).
-The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
+The distributed polarisabilities were computed by utilising the coupled\hyp{}perturbed Hartree\hyp{}Fock
 method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
-Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
+Orbital localisation was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
 
@@
 %
-where the subscript `CM' denotes the center of mass of water molecule.
-When the distributed polarizabilities were used instead of the polarization\hyp{}induced 
+where the subscript `CM' denotes the centre of mass of water molecule.
+When the distributed polarisabilities were used instead of the polarisation\hyp{}induced 
 difference density matrices, primitive induced quadrupole moments were computed from
@@
 and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
-was used. Before all the calculations, water molecule was energy\hyp{}optimized
+was used. Before all the calculations, water molecule was energy\hyp{}optimised
 with a maximum energy, force and displacement thresholds of 10$^{-6}$, 3$\times$$10^{-4}$ 
 and 1.2$\times$$10^{-3}$ a.u., respectively.
-For all the calculations of the density matrix polarization susceptibility tensors
+For all the calculations of the density matrix polarisation susceptibility tensors
 from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
-atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme,\cite{Mayer.IJQC.2002}
+atomic basis was orthogonalised by using the L{\"o}wdin symmetric orthogonalisation scheme,\cite{Mayer.IJQC.2002}
 and further back\hyp{}transformed to the non\hyp{}orthogonal AO basis.
@@
 with respect to the electric field from Eq.~\eqref{e:dD-Taylor}, 
-the following finite\hyp{}difference formulae were utilized:
+the following finite\hyp{}difference formulae were utilised:
 %
@@
 
-Polarization under the uniform electric field is the simplest type of electrostatic perturbation
+Polarisation under the uniform electric field is the simplest type of electrostatic perturbation
 whereby it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
@@
 The \emph{ab initio} model from Eq.~\eqref{e:final-model.HF} and
-the generalized model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
-The associated error distributions in prediction of polarization energies 
+the generalised model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
+The associated error distributions in prediction of polarisation energies 
 are shown in scatter plot in Figure~\ref{f:fig-1}, whereas the
-statistical performance descriptors of polarization energies and induced multipole moments
+statistical performance descriptors of polarisation energies and induced multipole moments
 are gathered in Table~\ref{t:results}.
 %
-\begin{figure}[h]
-\includegraphics[width=0.5\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
+\begin{figure}[t]
+\includegraphics[width=0.5\textwidth]{fig-1.eps}
 \caption{\label{f:fig-1} {\bf Performance of the DMS tensors of water molecule
@@
 from Eqs.~\eqref{e:dD-Taylor} (blue solid circles) and \eqref{e:final-model.General} (blue open circles),
-incorrectly reproduce the polarization energy.
+incorrectly reproduce the polarisation energy.
 Interestingly, the predicted values are in strong
@@
 obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
-or by using the generalized DMS model (red open circles)
+or by using the generalised DMS model (red open circles)
 dramatically reduces the errors
 leading to quantitative accuracy in all tested field magnitudes 
-with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalized DMS model.
+with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalised DMS model.
 On the other hand, the
@@
 This observation can be understood given the fact that the induced dipole
-and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
+and quadrupole moment is approximately proportional to the external electric field, whereas the polarisation
 energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
@@
 
-In comparison to the dipole\hyp{}dipole polarizability approach, 
+In comparison to the dipole\hyp{}dipole polarisability approach, 
 quadratic DMS models perform only slightly worse 
-(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalized quadratic DMS
-model and the polarizability model are used, respectively).
+(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalised quadratic DMS
+model and the polarisability model are used, respectively).
 However, quadratic DMS models better predict the induced multipole moments with
@@
 $R^2$ systematically greater than 99\%. 
-Certainly, taking into account the distributed quadrupole polarizabilities
-would greatly reduce errors of the standard distributed polarizability approach.
+Certainly, taking into account the distributed quadrupole polarisabilities
+would greatly reduce errors of the standard distributed polarisability approach.
 Nevertheless, it is clear that the second\hyp{}order
@@
 \toprule
-\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalized DMS}}
+\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalised DMS}}
                                                                   &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
@@
 &  
-& \multicolumn{2}{c}{\textbf{Polarizability}}        \\ \hline
+& \multicolumn{2}{c}{\textbf{Polarisability}}        \\ \hline
 \multicolumn{15}{c}{} \\ 
@@
 $^d$ \emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.
-$^e$ Linear\hyp{}regression generalized model, Eq.~\eqref{e:final-model.General}.
+$^e$ Linear\hyp{}regression generalised model, Eq.~\eqref{e:final-model.General}.
 $^f$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields 
@@
 To study the capability of the distributed DMS models to capture 
-density matrix polarization in inhomogeneous electric fields, a water molecule
+density matrix polarisation in inhomogeneous electric fields, a water molecule
 surrounded by point charges is analysed.
@@
 in liquid phase.\cite{Reischl.Kofinger.Dellago.MolPhys.2009,Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
-Performance descriptors of the \emph{ab initio} and generalized 
+Performance descriptors of the \emph{ab initio} and generalised 
 models of the density matrix susceptibility, as well as 
-the conventional dipole\hyp{}dipole distributed polarizability
-approach, are shown in Table~\ref{t:results}.
+the conventional dipole\hyp{}dipole distributed polarisability
+approach, are shown in Table~\ref{t:results}, whereas the distribution of errors
+in strong electric fields are additionally plotted in Figure~\ref{f:fig-2}A--I.
 
@@
 the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} 
-as well as
-the generalized linear DMS model, are insufficient to correctly describe the polarization energy 
-in non\hyp{}uniform electric field,
-although the correlation with benchmark is still very clear with a slope of $\sim$2
-(Figure~\ref{f:fig-2}A). 
-Again, induced dipole moments are reproduced very well also
-in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$).
-Induced quadrupole moments are well described only by the generalized DMS model.
-For example, in medium electric fields RMSQ is 10.4$\times 10^{-3}$D\AA{ }with $R^2=93.7\%$
-whereas the \emph{ab initio} DMS model yealds RMSQ of 28.2$\times 10^{-3}$D\AA{ }($R^2=53.9\%$).
-
-The most sophisticated model studied in this work, the generalized quadratic DMS model, yields excellent predictions 
-of the polarization energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
-The errors in predicting polarization energy are reduced by an order of magnitude when compared with
+as well as the generalised linear DMS model, are insufficient to correctly describe the polarisation energy 
+in non\hyp{}uniform electric field, although the correlation with benchmark is still very clear with a slope of $\sim$2
+(Figure~\ref{f:fig-2}A). Again, induced dipole moments are reproduced very well also
+in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$; Figure~\ref{f:fig-2}D and \ref{f:fig-2}E).
+Induced quadrupole moments are well described only by the generalised DMS model.
+For example, in moderate electric fields RMSQ is 10.4$\times 10^{-3}$D\AA{ }with $R^2=93.7\%$
+whereas the \emph{ab initio} DMS model yields RMSQ of 28.2$\times 10^{-3}$D\AA{ }($R^2=53.9\%$).
+
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{fig-2.eps}
+\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of water molecule in strong
+non\hyp{}uniform electric fields.}
+In this Figure, `Ab Initio DMS' (panels A, D and G) 
+refers to the model from Eq.~\eqref{e:final-model.HF} and
+`Generalised DMS' (panels B, E and H) 
+refers to the generalised model with $R_{\rm max}=2$ 
+where $R_{\rm max}$ is the maximum power
+in the expansion from Eq.~\eqref{e:final-model.General}.
+%Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
+%quadrupole moments (G-I). 
+Comparison with the distributed dipole\hyp{}dipole polarisability model 
+is also shown in this Figure in panels C, F and I. 
+The ensemble of electric fields corresponds to the `S' set
+from Tables~\ref{t:results} and \ref{t:fields}.}
+\end{figure*}
+
+The most sophisticated model studied in this work, the generalised quadratic DMS model, yields excellent predictions 
+of the polarisation energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
+The errors in predicting polarisation energy are reduced by an order of magnitude when compared with
 the linear DMS models. Improvement of accuracy can also be observed in the case of induced multipole moment predictions
@@
 What is particularly interesting is that in weak and moderate electric fields
-the generalized quadratic DMS model performs better than the distributed dipole\hyp{}dipole
-polarizability model, whereas only in strong electric fields the situation is reversed
-(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.2\%$ for the DMS and polarizability model, respectively).
+the generalised quadratic DMS model performs better than the distributed dipole\hyp{}dipole
+polarisability model, whereas only in strong electric fields the situation is reversed
+(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.1\%$ for the DMS and polarisability model, respectively; 
+see also Figure~\ref{f:fig-2}B and \ref{f:fig-2}C).
 Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% 
-even in strong electric fields for polarization
+even in strong electric fields for polarisation
 energy, and also very high $R^2$ for induced multipole moments.
@@
 Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
-polarizable force field, that the average electric field experienced by water molecule in liquid phase
+polarisable force field, that the average electric field experienced by water molecule in liquid phase
 in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
@@
 and \ref{t:fields}) for which the
-accuracy of the generalized quadratic DMS model is excellent.
+accuracy of the generalised quadratic DMS model is excellent.
 Water molecules can experience
@@
 This electric field magnitude falls into our strong electric field regime (denoted as `S')
-in which the generalized quadratic DMS model still provides accurate predictions (see also error distributions 
-shown in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H).
-
-Even at the current state of development, the generalized quadratic DMS model
-is of comparable energetic accuracy as the distributed polarizability model. 
-It is emphasized here that the model can be easily extended by taking into account
-third\hyp{}order terms as well if fine accuracy is required.
+in which the generalised quadratic DMS model still provides accurate predictions (see also error distributions 
+in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H).
+
+\section{\label{s:5}Summary and a few concluding remarks}
+
+%In this work, the \emph{ab initio} model of the 
+%one\hyp{}particle density matrix susceptibility tensors
+%at the HF level was derived for a closed shell molecule 
+%in a spatial electric field distribution. It was found 
+%that the change in the density matrix element is approximately 
+%linearly proportional to the electric field evaluated at the 
+%distributed site associated with the centre of charge of the 
+%localised molecular orbital. Subsequently, the \emph{ab initio} 
+%model served as a basis for the formulation of the generalised 
+%model that is, in principle, not limited to the HF theory,
+%and takes into account also quadratic and higher\hyp{}order terms 
+%with respect to the electric field. The generalised model, which 
+%consisted of a set of adjustable parameters, defined the first\hyp{} 
+%and second\hyp{}order density matrix susceptibilities that were 
+%determined through the least\hyp{}squares linear regression 
+%with respect to the training ensemble of electrostatically 
+%perturbed states.
+%
+%The above theoretically developed models were then used 
+%to calculate the electric field\hyp{}induced interaction 
+%energies, dipole moments and quadrupole moments of water 
+%molecule that is subjected to perturbation due to uniform 
+%electric field, as well as various point charge environments 
+%generating non\hyp{}uniform electric field of weak, moderate
+%and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
+%interaction energies, the atom\hyp{}distributed generalised model can reach quantitative level of accuracy
+%for prediction of interaction energies, dipole moments and even quadrupole moments
+%once, at least quadratic terms with respect to the electric field, are taken into account. 
+In this work, it was demonstrated that the generalized DMS models from Eq.~\eqref{e:final-model.General}
+can accurately describe the polarization\hyp{}induced deformation of the one\hyp{}particle ground\hyp{}state electronic density
+due to the uniform and non\hyp{}uniform distribution of electric field.
+It was also shown that the accuracy of the model can be improved by extending the parameter space of the susceptibility tensors.
+Even at the current state of development, the generalised quadratic DMS model
+is of comparable energetic accuracy as the distributed polarisability model (RMSE < 0.4~kcal/mol).
+%It is emphasised here that the model can be easily extended by taking into account
+%third\hyp{}order terms as well if fine accuracy is required.
+%Moreover, the generalized DMS model can be used to any type of one\hyp{}particle density matrices at arbitrary level of theory.
 %This however falls out of scope of this contribution and it will be studied
 %elsewhere. 
-In addition, any type of one\hyp{}particle density matrix can be treated by using
-generalized DMS models, which is much more difficult to realize when a conventional multipole
+Moreover, any type of one\hyp{}particle density matrix could be treated by using
+generalised DMS models, which is much more difficult to realise when a conventional multipole
 model is used.
 Therefore, it is believed that
-the generalized DMS approach
+the generalised DMS approach
 has an important advantage over the multipole\hyp{}based approach, i.e., it provides detailed
-tensor description of the spatial polarization of the electron density distribution, that could be
-combined with density\hyp{}based methods of interaction energy calculations.\cite{Mandado.Hermida-Ramon.JCTC.2011}
-
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-2.eps}
-\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of water molecule in moderate
-non\hyp{}uniform electric fields.}
-In this Figure, `Ab Initio DMS' (panels A, D and G) 
-refers to the model from Eq.~\eqref{e:final-model.HF} and
-`Generalized DMS' (panels B, E and H) 
-refers to the generalized model with $R_{\rm max}=2$ 
-where $R_{\rm max}$ is the maximum power
-in the expansion from Eq.~\eqref{e:final-model.General}.
-%Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
-%quadrupole moments (G-I). 
-Comparison with the distributed dipole\hyp{}dipole polarizability model 
-is also shown in this Figure in panels C, F and I.}
-\end{figure*}
-
-\section{\label{s:5}Summary and a few concluding remarks}
-
-In this work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
-at the HF level was derived for a closed shell molecule in a spatial
-electric field distribution. It was found that the change in the density matrix element
-is approximately linearly proportional to the electric field evaluated at the distributed site associated with the 
-center of charge of the 
-localized molecular orbital. Subsequently, the \emph{ab initio} model served as a basis
-for the formulation of the generalized model that is, in principle, not limited to the HF theory,
-and takes into account also quadratic and higher\hyp{}order terms with respect to the electric field.
-The generalized model, which consisted of a set of adjustable parameters,
-definined the first\hyp{} and second\hyp{}order density matrix susceptibilities
-that were determined through the least\hyp{}squares 
-linear regression with respect to the training ensemble of electrostatically perturbed states.
-
-The above theoretically developed models were then used to calculate the electric field\hyp{}induced
-interaction energies, dipole moments and quadrupole moments of water molecule that
-is subjected to perturbation due to uniform electric field, as well as 
-various point charge environments generating non\hyp{}uniform electric field of weak, moderate
-and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
-interaction energies, the atom\hyp{}distributed generalized model can reach quantitative level of accuracy
-for prediction of interaction energies, dipole moments and even quadrupole moments
-once, at least quadratic terms with respect to the electric field, are taken into account. 
-This work demonstrated that it is possible to
-accurately predict the change in the density matrix of electronic system, and the accuracy of the model
-can be improved by extending the parameter space of the susceptibility tensors.
-
-It is anticipated here that the generalized models from Eq.~\eqref{e:final-model.General}
+tensor description of the spatial polarisation of the electron density distribution, that could be
+combined for example with density\hyp{}based methods of interaction energy calculations.\cite{Mandado.Hermida-Ramon.JCTC.2011}
+
+It is anticipated here that the generalised models from Eq.~\eqref{e:final-model.General}
 should be studied in the future with an emphasis 
-on their further optimization and application for molecular aggregates.
+on their further optimisation and application for molecular aggregates.
 For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
-sites should be further investigated, preferably with post\hyp{}Hartree\hyp{}Fock methods
+sites should be further investigated, preferably considering post\hyp{}Hartree\hyp{}Fock methods
 and electronically excited states.
 Nevertheless, owing to the capability of the proposed approach to correctly capture the 
-environment\hyp{}induced changes in the one\hyp{}particle density matrix, 
-the generalized density matrix susceptibility tensors 
+polarization of ground state's electron density,
+the generalised density matrix susceptibility tensors 
 could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
@@
 %
-Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
+Therefore, it is proved that the polarisation\hyp{}induced distributed dipole moments 
 defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} are origin independent.

### 2018-08-07 papers/1_dmatpol/submit/make.fig-1.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+figure_dir="./"
+
+gnuplot plot.fig-1.gpl
+
+for plot in fig-1 ; do
+    epstopdf ${plot}-inc.eps
+    pdflatex ${plot}.tex
+    pdftops -eps ${plot}.pdf
+    #mv -v ${plot}.pdf ${plot}.eps ${figure_dir}
+    rm -v ${plot}.pdf
+    rm -v ${plot}.aux ${plot}.tex ${plot}.log ${plot}-inc*
+done

### 2018-08-07 papers/1_dmatpol/submit/make.fig-2.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+figure_dir="./"
+
+gnuplot plot.fig-2.gpl
+
+for plot in fig-2 ; do
+    epstopdf ${plot}-inc.eps
+    pdflatex ${plot}.tex
+    pdftops -eps ${plot}.pdf
+    #mv -v ${plot}.pdf ${plot}.eps ${figure_dir}
+    rm -v ${plot}.pdf
+    rm -v ${plot}.aux ${plot}.tex ${plot}.log ${plot}-inc*
+done

### 2018-08-09 papers/1_dmatpol/submit/paper-submit.tex
Final review I.

@@
 \usepackage{bm}% bold math
+%\usepackage{txfonts}
 %\usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 
+% EU logo
+\newcommand*{\InlineImage}[1]{%
+    \raisebox{-.02\baselineskip}{%
+        \includegraphics[
+        height=\baselineskip,
+        width=\baselineskip,
+        keepaspectratio,
+        ]{#1}%
+    }%
+}
+
 %\draft % marks overfull lines with a black rule on the right
@@
 that link extremely accurate but localised pieces of information with certain knowledge of an atomistic 
-organisation and intermolecular interaction on a relatively large spatial scales, typically ranging beyond 
+organisation and intermolecular interaction on relatively large spatial scales, typically ranging beyond 
 few nanometers.
@@
 and much less expensive classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
-Fruitful incorporation of accurate \emph{ab initio} methods with extensive experimental and semi\hyp{}empirical
+Combination of accurate \emph{ab initio} methods with extensive experimental and semi\hyp{}empirical
 refinings led to development of a wide variety of force fields,\cite{Demerdash.Yap.T-Head-Gordon.AnnuRevPhysChem.2014}
@@
 %
-to simplify the notation. Hence, we have
+to simplify the notation. Hence,
 %
@@
 Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
-No.~665778. This project is funded by National Science Centre, Poland 
+No.~665778. \InlineImage{eu-logo.eps} This project is funded by National Science Centre, Poland 
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 

### 2018-08-10 papers/1_dmatpol/submit/paper-submit.tex
Reference and main text check-up final

@@
 of the \emph{ab initio} DMS model (Eq.~\eqref{e:final-model.HF}) were assumed 
-to be LMO centroids, while atoms were used for the generalised DMS models (Eq.~\eqref{e:final-model.General}).
-This is because various LMO centroids are often located at the same point which would
-introduce linear dependencies in the Hessian matrix from Eq.~\eqref{e:Newton.gH}.
+to be the LMO centroids (resulting in the five\hyp{}centre model), 
+while the atoms were used for the generalised DMS models (Eq.~\eqref{e:final-model.General}, 
+resulting in the three\hyp{}centre models).
+This is because certain LMO centroids are often located at the same point which would
+introduce linear dependencies in the Hessian matrix from Eq.~\eqref{e:Newton.gH}
+when using the generalised DMS scheme.
 
@@
 In order to compare the performance of the usual 
-distributed dipole\hyp{}dipole polarisability approach,
+distributed dipole\hyp{}dipole polarisability approach,\cite{Li.Netzloff.Gordon.JCP.2006}
 the approximate interaction energy was also evaluated by using the expression:
@@
    \begin{equation}\label{e:rmse}
-     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 } \;;
+     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N^{N_{\rm max}} \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 } \;;
    \end{equation}
@@
    \begin{equation}\label{e:rmsd}
-     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \left( \lvert \delta {\BM\upmu}^{(N)} \rvert 
+     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N^{N_{\rm max}} \left( \lvert \delta {\BM\upmu}^{(N)} \rvert 
                                              - \lvert \overline{\delta {\BM\upmu}^{(N)} \rvert }\right)^2 } \;;
@@
    \begin{equation}\label{e:rmsq}
-     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \left( \lvert \delta \widetilde{\BM\Theta}^{(N)} \rvert 
+     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N^{N_{\rm max}} \left( \lvert \delta \widetilde{\BM\Theta}^{(N)} \rvert 
                                              - \lvert \overline{\delta \widetilde{\BM\Theta}^{(N)} \rvert }\right)^2 } \;.
@@
 to the Cartesian component of the electric field, and the finite difference
-steps $F_{u(w)}$ were assumed to be 0.0005~a.u for all components.
+steps $F_{u(w)}$ were assumed to be 0.0005~a.u. for all components.
 
@@
 \includegraphics[width=0.5\textwidth]{fig-1.eps}
-\caption{\label{f:fig-1} {\bf Performance of the DMS tensors of water molecule
+\caption{\label{f:fig-1} {\bf Performance of the DMS tensors of the ground\hyp{}state water molecule
 in uniform electric field.} 
@@
 \begin{table*}%[]
-\caption{{\bf Performance of the DMS tensors of water molecule
+\caption{{\bf Performance of the DMS tensors of ground\hyp{}state water molecule
 in uniform and non\hyp{}uniform electric field$^{a,b}$}
@@
 \begin{table}[b]
-\caption{{\bf Average electric fields in statistical sets of electrostatically perturbed states
+\caption{{\bf Average electric fields in statistical sets of electrostatically perturbed ground states
 of water molecule surrounded by point charges$^a$}
@@
 because similar ranges of electric field are experienced by water molecules
-in liquid phase.\cite{Reischl.Kofinger.Dellago.MolPhys.2009,Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
+in liquid phase.\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013,Reischl.Kofinger.Dellago.MolPhys.2009}
 Performance descriptors of the \emph{ab initio} and generalised 
@@
 the conventional dipole\hyp{}dipole distributed polarisability
-approach, are shown in Table~\ref{t:results}, whereas the distribution of errors
+approach, are shown in Table~\ref{t:results}, whereas the error distributions
 in strong electric fields are additionally plotted in Figure~\ref{f:fig-2}A--I.
@@
 \includegraphics[width=\textwidth]{fig-2.eps}
-\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of water molecule in strong
-non\hyp{}uniform electric fields.}
+\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of ground\hyp{}state water molecule in strong
+non\hyp{}uniform electric field.}
 In this Figure, `Ab Initio DMS' (panels A, D and G) 
@@
 of the polarisation energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
-The errors in predicting polarisation energy are reduced by an order of magnitude when compared with
+The errors in predicting polarisation energy are reduced by over an order of magnitude when compared with
 the linear DMS models. Improvement of accuracy can also be observed in the case of induced multipole moment predictions
-but the effect is less significant (improvement roughly by a factor of 2).
+but the effect is less significant.
+%(improvement roughly by a factor of 2).
 What is particularly interesting is that in weak and moderate electric fields
@@
 see also Figure~\ref{f:fig-2}B and \ref{f:fig-2}C).
-Nevertheless, the quadratic DMS model yields still very high $R^2$ of 97.2\% 
-even in strong electric fields for polarisation
-energy, and also very high $R^2$ for induced multipole moments.
+Nevertheless, the quadratic DMS model yields still very accurate predictions 
+of polarization energy and induced multipole moments, even in strong electric field.
 
@@
 in which the generalised quadratic DMS model still provides accurate predictions (see also error distributions 
-in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H).
+in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H). Therefore, it is believed that
+the generalised quadratic DMS model is capable of accurately describing 
+the polarization of ground\hyp{}state water molecule in inhomogeneous 
+electric fields that are comparable in strength to condensed phase conditions.
 
@@
 %once, at least quadratic terms with respect to the electric field, are taken into account. 
-In this work, it was demonstrated that the generalized DMS models from Eq.~\eqref{e:final-model.General}
+In this work, it was demonstrated that the generalised DMS models from Eq.~\eqref{e:final-model.General}
 can accurately describe the polarization\hyp{}induced deformation of the one\hyp{}particle ground\hyp{}state electronic density
@@
 Even at the current state of development, the generalised quadratic DMS model
-is of comparable energetic accuracy as the distributed polarisability model (RMSE < 0.4~kcal/mol).
+is of comparable energetic accuracy as the distributed polarisability model (RMSE of 0.4~kcal/mol in strong 
+electric fields and less than 0.06~kcal/mol in moderate and weak fields).
 %It is emphasised here that the model can be easily extended by taking into account
@@
 the generalised DMS approach
-has an important advantage over the multipole\hyp{}based approach, i.e., it provides detailed
+has an important advantage over the multipole\hyp{}based approach, i.e., it provides more detailed
 tensor description of the spatial polarisation of the electron density distribution, that could be
@@
 
-Multipole integrals depend on the origin with respect to which they are evaluated.
-However, the induced dipole moments defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
+Multipole one\hyp{}electron integrals depend on the origin with respect to which they are evaluated.
+However, the induced dipole moment defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
 (or equivalently in Eq.~\eqref{e:dmu-l-vector})
-have to be origin independent. Note that
+has to be origin\hyp{}independent. Note that
 %
@@
 Therefore, it is proved that the polarisation\hyp{}induced distributed dipole moments 
-defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} are origin independent.
-Thus, one can compute dipole integrals with respect to any origin and resulting
-susceptibilities ${\bf B}_{\alpha\beta}^{(i;1)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
+defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} are origin\hyp{}independent.
+Thus, one can compute one\hyp{}electron dipole integrals with respect to any origin 
+and resulting the susceptibilities ${\bf B}_{\alpha\beta}^{(i;1)}$ 
+from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
 
@@
 
-The gradient vector ${\bf g}$ and Hessian matrix ${\bf H}$ 
-are build from blocks associated with a particular type of parameters, i.e.,
+The gradient vector ${\bf g}$ and the Hessian matrix ${\bf H}$ 
+are built from blocks associated with a particular type of parameters, i.e.,
 %

### 2018-08-10 papers/1_dmatpol/submit/paper-submit.tex
Rederivation. Complex numbers checkup and revision.

@@
   \delta {\bf D} \approx \sum_i^{\text{Sites}} {\bf B}_i
-  \cdot  \delta {\bf P}_i \;.
+  \cdot  \delta {\bf P}_i \;,
 \end{equation*}
 %
+where the `$\cdot$' symbol denotes the inner tensor (dot) product.
 While $\delta {\bf P}_i$ contains critical information about the actual distortion of the electronic cloud,
@@
 %
+where the `$\otimes$' symbol denotes the outer tensor product.
 In the above expression, the density matrix susceptibilities 
@@
 The latter equality is true only if the basis set functions are orthogonal, i.e., 
-${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$,
-which will be the case throughout the entire work, unless stated otherwise.
+${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$.
+Such a condition is necessary in order to meet the charge conservation
+requirement (see Appendix~\ref{a:orig-dep} for a detailed discussion).
+Therefore orthogonality of AO's is assumed from that point, unless stated otherwise.
+%Also, from this point onwards we assume that the molecular orbitals 
+%and the transformation matrix $\BM \Delta$ are real.
 Substituting the result from Eq.~\eqref{e:dD-1}
@@
 %
-\begin{multline} \label{e:dmu-4-exact.linear-approximation}
- \frac{1}{2} 
- \delta {\BM{\upmu}}
+%\begin{multline} \label{e:dmu-4-exact.linear-approximation}
+% \frac{1}{2} 
+% \delta {\BM{\upmu}}({\bf r}_Q)
+%  \cong
+%   \Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
+%    \right] }
+%%
+%  +\Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
+%    \right] } \\
+%%
+%  -\Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
+%    \right] }
+%%
+%  -\Tr{ 
+%    \left[ 
+%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
+%    \right] } \;.
+%\end{multline}
+%
+%This can be recast to
+%
+\begin{equation}\label{e:dmu-l-vector}
+ \frac{1}{4} 
+ \delta {\BM{\upmu}}({\bf r}_Q)
   \cong
+  \Re \left\{ 
  \Tr{ 
     \left[ 
-         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
+         \delta {\bf C} \cdot \mathbb{L}({\bf r}_Q)
    \right] }  
-%
-  +\Tr{ 
-    \left[ 
-         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
-    \right] } \\
-%
-  -\Tr{ 
-    \left[ 
-         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
-    \right] }
-%
-  -\Tr{ 
-    \left[ 
-         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
-    \right] } \;.
-\end{multline}
-%
+  \right\} \;,
+\end{equation}
+%
+where the `$\Re$' symbol denotes the real part,
+ and the intermediate quantity
+%
+\begin{equation} \label{e:L-projector}
+ \mathbb{L}({\bf r}_Q) \equiv {\bf C}^{\dagger} \cdot \mathbb{M}({\bf r}_Q) \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \;.
+\end{equation}
+%
+Note that $\mathbb{L}({\bf r}_Q)$ contains the projector onto
+the unoccupied MO space at the absence of the external electric field. It means that the
+perturbed solution is constructed from the unperturbed virtual orbitals as well.
+Note also the similarity between Eq.~\eqref{e:L-projector} and Eq.~\eqref{e:dmu-exact}.
 It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
@@
 This guarantees that the theory fulfils the total charge conservation requirement.
-Therefore,
-in the above result and also in later course of the paper the dipole origin will be omitted. 
-%where the dipole origin was omitted for notational clarity.
-
-Now, if the AO basis is orthogonal, Eq.~\eqref{e:dmu-4-exact.linear-approximation} can be simplified by realising that
-%
-\begin{equation}
- \frac{1}{2} 
- \delta {\BM{\upmu}}
-  \cong
-2 \Tr{ 
-    \left[ 
-         \widetilde{\mathbb{M}} \cdot \delta {\bf C}
-   \right] }  
--
-2 \Tr{ 
-    \left[ 
-         \widetilde{\mathbb{K}} \cdot \delta {\bf C}
-   \right] } \;.
-\end{equation}
-%
-Here we introduced the auxiliary tensors
-%
-\begin{subequations}
- \begin{align}
-   \widetilde{\mathbb{M}}  &\equiv {\bf C}^{(0)\dagger} \cdot {\mathbb{M}}     \;,           \\
-   \widetilde{\mathbb{K}}  &\equiv {\bf C}^{(0)\dagger} \cdot {\mathbb{M}} \cdot {\bf D}^{(0)} \;, \\
-   \widetilde{\mathbb{L}}  &\equiv \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \;
- \end{align}
-\end{subequations}
-%
-to simplify the notation. Hence,
-%
-\begin{equation} \label{e:dmu-l-vector}
-  \frac{1}{4} 
- \delta {\BM{\upmu}} %({\bf r}_0) 
-   =
-   \Tr{ 
-    \left[ 
-         \widetilde{\mathbb{L}} \cdot \delta {\bf C}
-    \right] }
-   %
-   \equiv \sum_{i}^{\rm occ} \sum_\alpha {\bf l}_{i\alpha} \delta C_{\alpha i} \;,
-\end{equation}
-%
-where the Cartesian vector
-%
-\begin{equation}\label{e:l-vector}
- {\bf l}_{i\alpha} \equiv \left[ \widetilde{\mathbb{M}} - \widetilde{\mathbb{K}} \right]_{i\alpha} 
-      = \left[  {\bf C}^\dagger \cdot \mathbb{M} \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \right]_{i\alpha}
-\end{equation}
-%
-and $i$ runs over occupied unperturbed MO's.
-Note that the intermediate quantity $\widetilde{\mathbb{L}}$ contains the projector onto
-the unoccupied MO space at the absence of the external electric field. It means that the
-perturbed solution is constructed from the unperturbed virtual orbitals as well.
+Therefore, in later course of the paper the dipole origin will be omitted.
+
 Although Eq.~\eqref{e:dmu-l-vector}
-is exact up to first order in ${\BM\Delta}$, it cannot be inverted at this moment to obtain $\delta {\bf C}$
-due to the presence of trace operation, unless the electric field is uniform.
+is exact up to first order in ${\BM\Delta}$, it cannot be inverted at this moment to obtain $\delta {\bf C}$.
+If the orbitals are real, the analytic solution exists
+provided that the electric field is uniform. In a more general case, i.e., when the electric field
+is non\hyp{}uniform, the analytic solution cannot be found without further approximations.
 Let the induced dipole moment be partitioned into separate contributions
@@
 %
-where
+where the induced dipole moment associated with the $i$th LMO
+is assumed to be linearly dependent on the local electric field, i.e.,
 %
 \begin{equation} \label{e:dmu-dpol}
- \delta {\BM{\upmu}}_i = {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) \;.
+ \delta {\BM{\upmu}}_i \approx {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) \;.
 \end{equation}
@@
 recasting the right hand side in a matrix form
-and assuming real quantities,
-one can separate the contributions coming from occupied MO's as
+and assuming real orbitals,
+one can separate the contributions coming from the occupied MO's as
 %
@@
    =
-   \delta {\bf c}_i^T \cdot {\bf L}_i \;.
-\end{equation}
-%
-Note that now ${\bf L}_i$ is a matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
+   \delta {\bf c}_i^{\rm T} \cdot {\bf L}_i \;.
+\end{equation}
+%
+Note that now ${\bf L}_i \equiv \left[ \mathbb{L} \right]_i$ 
+is a real rectangular matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
 change in the LCAO-MO matrix due to the polarisation process.
@@
 \begin{equation} \label{e:mu-ind.matrix.linear.solution}
-  \delta {\bf c}_i^T = \frac{1}{4}
+  \delta {\bf c}_i^{\rm T} = \frac{1}{4}
             {\bf F}({\bf r}_i)  \cdot {\BM{\alpha}}_i \cdot 
@@
       \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
-       \left[ {\bf L}_i^T \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^T   
+       \left[ {\bf L}_i^{\rm T} \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^{\rm T}   
 \end{equation}
@@
 with the dimension of $3\times n$. Note that the right inverse of ${\bf L}_i$
-does not exist because ${\bf L}_i \cdot {\bf L}_i^T$ is singular due to the fact that usually
+does not exist because ${\bf L}_i \cdot {\bf L}_i^{\rm T}$ is singular due to the fact that usually
 $n > 3$.

### 2018-08-12 papers/1_dmatpol/submit/paper-submit.tex
Major revision of the induced dipole moment origin-independence proof

@@
 %
-The latter equality is true only if the basis set functions are orthogonal, i.e., 
+The latter equality is true in a special case when the basis set functions are orthogonal, i.e., 
 ${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$.
-Such a condition is necessary in order to meet the charge conservation
-requirement (see Appendix~\ref{a:orig-dep} for a detailed discussion).
-Therefore orthogonality of AO's is assumed from that point, unless stated otherwise.
+%Such a condition is necessary in order to meet the charge conservation
+%requirement (see Appendix~\ref{a:orig-dep} for a detailed discussion).
+%Therefore orthogonality of AO's is assumed from that point, unless stated otherwise.
 %Also, from this point onwards we assume that the molecular orbitals 
 %and the transformation matrix $\BM \Delta$ are real.
-Substituting the result from Eq.~\eqref{e:dD-1}
-into Eq.~\eqref{e:dmu-exact} leads to
+By substituting the McWeeny's result from Eq.~\eqref{e:dD-1}
+into Eq.~\eqref{e:dmu-exact} one can show that
 %
@@
 %         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
-%    \right] } \;.
+%    \right] } \;,
 %\end{multline}
-%
-%This can be recast to
+%%
+%which can be recast to
 %
@@
    \right] }  
-  \right\} \;,
-\end{equation}
-%
-where the `$\Re$' symbol denotes the real part,
+  \right\} \;.
+\end{equation}
+%
+In the above expression, the `$\Re$' symbol denotes the real part,
 and the intermediate quantity
@@
 \begin{equation} \label{e:L-projector}
- \mathbb{L}({\bf r}_Q) \equiv {\bf C}^{\dagger} \cdot \mathbb{M}({\bf r}_Q) \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \;.
+ \mathbb{L}({\bf r}_Q) \equiv {\bf C}^{(0)\dagger} \cdot \mathbb{M}({\bf r}_Q) \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \;.
 \end{equation}
@@
 perturbed solution is constructed from the unperturbed virtual orbitals as well.
-Note also the similarity between Eq.~\eqref{e:L-projector} and Eq.~\eqref{e:dmu-exact}.
-It can be proven (see Appendix~\ref{a:orig-dep}) that such obtained
-induced dipole moment is independent of the origin. 
-This guarantees that the theory fulfils the total charge conservation requirement.
-Therefore, in later course of the paper the dipole origin will be omitted.
+%Note also the similarity between Eq.~\eqref{e:L-projector} and Eq.~\eqref{e:dmu-exact}.
+
+For the sake of simplicity, 
+the orthogonality of AO's is assumed in this work, unless stated otherwise.
+It is proven in the Appendix~\ref{a:orig-dep} that under such circumstances the
+induced dipole moment obtained in Eq.~\eqref{e:L-projector} is independent of the origin. 
+This guarantees that the theory fulfils the total charge conservation requirement and,
+therefore, in later course of the paper, the dipole origin will be omitted.
 
@@
  \delta D^{(N)} = \sum_{i }^M \Big\{ 
-                              \sum_u^{x,y,z} S^{[1]}_{iu} F_{iu}^{(N)} \\
-%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( S^{[2]}_{iu} + S^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
-                +             \sum_{u}^{x,y,z} \sum_{w<u} r_{uw} S^{[2]}_{iuw} F_{iu}^{(N)} F_{iw}^{(N)} 
+                              \sum_u^{x,y,z} s^{[1]}_{iu} F_{iu}^{(N)} \\
+%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( s^{[2]}_{iu} + s^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
+                +             \sum_{u}^{x,y,z} \sum_{w<u} r_{uw} s^{[2]}_{iuw} F_{iu}^{(N)} F_{iw}^{(N)} 
                         + \ldots \Big\}
@@
 (the Greek subscripts were omitted here for notational simplicity).
-In the above equation, $B_u^{(i;1)} = S^{[1]}_{iu}$ and $B_{uw}^{(i;2)} = r_{uw} S^{[2]}_{iuw}$,
+In the above equation, $B_u^{(i;1)} = s^{[1]}_{iu}$ and $B_{uw}^{(i;2)} = r_{uw} s^{[2]}_{iuw}$,
 where $r_{uw}$ is the symmetry factor equal to 1 for diagonal elements and 2 for off\hyp{}diagonal
@@
 The multiple parameter blocks
-(${\bf S}^{[1]}$, ${\bf S}^{[2]}$ and so on)
+(${\bf s}^{[1]}$, ${\bf s}^{[2]}$ and so on)
 appear in the first power, allowing for linear least\hyp{}squares regression.
@@
 $
- {\bf S} = 
+ {\bf s} = 
 \begin{pmatrix}
-{\bf S}^{[1]} &
-{\bf S}^{[2]} & \cdots
-\end{pmatrix}^T
+{\bf s}^{[1]} &
+{\bf s}^{[2]} & \cdots
+\end{pmatrix}^{\rm T}
 $, we define a loss function $Z$ 
@@
 \begin{equation}\label{e:Z}
- Z({\bf S}) = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
+ Z({\bf s}) = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
 \end{equation}
@@
 \begin{equation}\label{e:Newton}
- {\bf S} = -{\bf H}^{-1} \cdot {\bf g} \;,
+ {\bf s} = -{\bf H}^{-1} \cdot {\bf g} \;,
 \end{equation}
@@
 %
-\begin{equation}
+\begin{equation}\label{e:dTheta-dpol}
   \delta {\BM \Theta} = \sum_i^{\rm occ} 
@@
 in the expansion from Eq.~\eqref{e:final-model.General}.
-%Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
-%quadrupole moments (G-I). 
-Comparison with the distributed dipole\hyp{}dipole polarisability model 
+Comparison with the distributed dipole\hyp{}dipole polarisability model %(Eq.~\eqref{e:dE-dpol})
 is also shown in this Figure in panels C, F and I. 
@@
 Even at the current state of development, the generalised quadratic DMS model
-is of comparable energetic accuracy as the distributed polarisability model (RMSE of 0.4~kcal/mol in strong 
+is of comparable and quantitative energetic accuracy as the distributed polarisability model (RMSE of 0.4~kcal/mol in strong 
 electric fields and less than 0.06~kcal/mol in moderate and weak fields).
@@
 Multipole one\hyp{}electron integrals depend on the origin with respect to which they are evaluated.
-However, the induced dipole moment defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
-(or equivalently in Eq.~\eqref{e:dmu-l-vector})
-has to be origin\hyp{}independent. Note that
+However, the induced dipole moment defined in Eq.~\eqref{e:L-projector} 
+has to be origin\hyp{}independent. Denoting the overlap AO integrals 
+by $S_{\alpha\beta} = \braket{\alpha}{\beta}$,
+it follows that $\mathbb{M}_u({\bf r}_Q) = \mathbb{M}_u({\bf 0}) - u_Q {\bf S}$
+for $u=x,y,z$. Thus,
 %
 \begin{equation}
- \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \right]_{i\alpha} {\bf r}_Q  
-\end{equation}
-%
-and
+ \mathbb{L}_u({\bf r}_Q) = \mathbb{L}_u({\bf 0})
+                          - u_Q {\bf C}^{(0)\dagger} \cdot {\bf S} \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) \;,
+\end{equation}
+%
+where
 %
 \begin{equation}
- \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0}) - \left[ {\bf C}^\dagger \cdot {\bf D} \right]_{i\alpha} {\bf r}_Q \;.
-\end{equation}
-%
-But ${\bf C}^\dagger \cdot {\bf D}={\bf C}^\dagger \cdot {\bf C} \cdot {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
-AO basis which implies that
-%
-\begin{multline}
-   \left[ \widetilde{\mathbb{L}} \right]_{i\alpha} ({\bf r}_Q) 
- = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf r}_Q) 
- - \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf r}_Q) \\
- = \left[ \widetilde{\mathbb{M}} \right]_{i\alpha} ({\bf 0})   
- - \left[ \widetilde{\mathbb{K}} \right]_{i\alpha} ({\bf 0})
- = \left[ \widetilde{\mathbb{L}} \right]_{i\alpha} ({\bf 0}) \;.
-\end{multline}
-%
-Therefore, it is proved that the polarisation\hyp{}induced distributed dipole moments 
-defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} are origin\hyp{}independent.
-Thus, one can compute one\hyp{}electron dipole integrals with respect to any origin 
-and resulting the susceptibilities ${\bf B}_{\alpha\beta}^{(i;1)}$ 
-from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
+ \mathbb{L}_u({\bf 0}) = {\bf C}^{(0)\dagger} \cdot \mathbb{M}_u({\bf 0}) \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) \;.
+\end{equation}
+%
+But
+%
+\begin{equation}
+  {\bf C}^{(0)\dagger} \cdot {\bf S} \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) 
+= {\bf C}^{(0)\dagger} \cdot \left( {\bf S} - {\bf 1} \right)
+\end{equation}
+%
+since ${\bf C}^{(0)\dagger} \cdot {\bf S} \cdot {\bf C}^{(0)} = {\bf 1}$. 
+In the simplest case of ${\bf S} = {\bf 1}$ (i.e., the AO basis functions are orthogonal) 
+$\mathbb{L}_u({\bf r}_Q) = \mathbb{L}_u({\bf 0})$ which guarantees charge conservation.
+When a non\hyp{}orthogonal AO basis is used, 
+the following general condition must be satisfied:
+%
+\begin{equation}
+ \Re \left\{
+  \Tr{
+  \left[
+  \delta {\bf C} \cdot {\bf C}^{(0)\dagger} \cdot \left( {\bf S} - {\bf 1} \right)  
+ \right]}
+ \right\} = 0
+\end{equation}
+%
+Due to the fact that ${\BM \Delta}$ considered here is accurate only up to first\hyp{}order, 
+the above expression is not valid. 
+Therefore, the polarisation\hyp{}induced distributed dipole moments 
+defined in Eq.~\eqref{e:L-projector} are exactly origin\hyp{}independent
+as long as the AO basis is orthogonalised.
+In such a case, one can compute one\hyp{}electron dipole integrals with respect to any origin 
+and the resulting susceptibilities $\mathbb{B}^{(i;1)}$ 
+from Eq.~\eqref{e:susceptibility-B} will be uniquely defined. They can be back\hyp{}transformed
+to the original (non\hyp{}orthogonal) AO basis, 
+${\bf X} \cdot \mathbb{B}^{(i;1)} \cdot {\bf X}$, where $\bf X$ is an orthogonaliser.
+The numerical tests confirmed that $\Tr\left[\delta{\bf D}\cdot{\bf S}\right] \sim 10^{-16}$
+with $\delta{\bf D}$ given by Eq.~\eqref{e:final-model.HF} and susceptibilities transformed
+to the non\hyp{}orthogonal AO basis.
 
@@
  \begin{align}
-  g^{[r ]}    &\equiv \frac{\partial   Z}{\partial S^{[r]}} 
+  g^{[r ]}    &\equiv \frac{\partial   Z}{\partial s^{[r]}} 
      =-2\sum_N \overline{\delta D}^{(N)}
-               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}} \;,\\
-  H^{[rs]} &\equiv \frac{\partial^2 Z}{\partial S^{[r]} \partial S^{[s]}}  
+               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[r]}} \;,\\
+  H^{[rs]} &\equiv \frac{\partial^2 Z}{\partial s^{[r]} \partial s^{[s]}}  
      = 2\sum_N 
-        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}}
-        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[s]}} \;.
+        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[r]}}
+        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[s]}} \;.
  \end{align}
@@
 Note that due to the symmetry of the Hessian matrix, the block $21$
-is a transpose of the block $12$. 
-The composite indices $ku$ and $kuw$ are constructed from the distributed site index $k$
-and the appropriate symmetry\hyp{}adapted ($w<u$) Cartesian component 
-of a particular DMS tensor: $u$ for the first\hyp{}order,
-and $uw$ for the second\hyp{}order susceptibility tensor, respectively.
+is a transpose of the block $12$. The composite indices $ku$ and $kuw$ 
+are constructed from the distributed site index $k$ and the appropriate 
+symmetry\hyp{}adapted ($w<u$) Cartesian component of a particular DMS tensor: 
+$u$ for the first\hyp{}order, and $uw$ for the second\hyp{}order susceptibility 
+tensor, respectively. The method described above can be easily extended 
+to third and higher orders.
 

### 2018-08-13 papers/1_dmatpol/submit/paper-submit.tex
Citation added and switch to reprint mode

@@
 %%
-\documentclass[aip,amsmath,amssymb,reprint,floatfix]{revtex4-1}
+\documentclass[aip,jcp,preprint,amsmath,amssymb,floatfix]{revtex4-1}
 %\documentclass[aip,reprint]{revtex4-1}
@@
 
-\preprint{AIP/123-DMS}
+%\preprint{AIP/123-DMS}
 
@@
 (vibrational or vibronic solvatochromisms).\cite{Blasiak.Londergan.Webb.Cho.ACR.2017,
-Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018} 
+Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018,Lewis.Gruenke.Oliver.Ballottari.Bassi.Fleming.JPCL.2016} 
 Interface between QM and MM levels is usually realised by electrostatic embedding and only rarely 
@@
   \delta {\bf c}_i^{\rm T} = \frac{1}{4}
-            {\bf F}({\bf r}_i)  \cdot {\BM{\alpha}}_i \cdot 
+                    {\bf F}^{\rm T}({\bf r}_i) \cdot {\BM{\alpha}}_i^{\rm T} \cdot
                     \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
+%                   {\BM{\alpha}}_i : {\bf F}({\bf r}_i)  \otimes
+%                   \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
 \end{equation}
@@
 %
-\begin{multline}  \label{e:susceptibility-B}
+%\begin{multline}  \label{e:susceptibility-B} % PREPRINT
+\begin{equation}  \label{e:susceptibility-B}
  {\bf B}^{(i;1)}_{\alpha\beta} = 
-                               C_{\alpha i} {\bf b}_{\beta}^{(i;1)} + C_{\beta i} {\bf b}_\alpha^{(i;1)} \\
+                               C_{\alpha i} {\bf b}_{\beta}^{(i;1)} + C_{\beta i} {\bf b}_\alpha^{(i;1)} %\\ PREPRINT
                                 - \sum_\gamma 
@@
  \;.
-\end{multline}
+\end{equation}
+%\end{multline} % PREPRINT
 %
@@
 %
-\begin{multline}\label{e:final-model.General}
+%\begin{multline}\label{e:final-model.General} % PREPRINT
+\begin{equation}\label{e:final-model.General}
  \delta D_{\alpha\beta} = \sum_{i }^M \Big\{
-                                      {\bf B}^{(i ; 1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \\
+                                      {\bf B}^{(i ; 1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  %\\ PREPRINT
 %                       + \sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
@@
                         + \ldots \Big\} \;.
-\end{multline}
+\end{equation}
+%\end{multline} % PREPRINT
 %
@@
 %
-\begin{multline}\label{e:final-model.General.Parameters}
+%\begin{multline}\label{e:final-model.General.Parameters} % PREPRINT
+\begin{equation}\label{e:final-model.General.Parameters}
  \delta D^{(N)} = \sum_{i }^M \Big\{ 
-                              \sum_u^{x,y,z} s^{[1]}_{iu} F_{iu}^{(N)} \\
+                              \sum_u^{x,y,z} s^{[1]}_{iu} F_{iu}^{(N)} %\\ % PREPRINT
 %               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( s^{[2]}_{iu} + s^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
@@
                         + \ldots \Big\}
-\end{multline}
+\end{equation}
+%\end{multline} % PREPRINT
 %
@@
 %
-\begin{multline}\label{e:dE}
+%\begin{multline}\label{e:dE} % PREPRINT
+\begin{equation}\label{e:dE}
  \delta E^{\rm Pol}[{\bf F}] = 
                      \Tr{\left\{ {\bf D}^{(0)} \cdot
-                                \left( 2\delta {\bf J} - \delta {\bf K}  \right) \right\}} \\
+                                \left( 2\delta {\bf J} - \delta {\bf K}  \right) \right\}} %\\ % PREPRINT
                    + \Tr{\left\{ \delta {\bf D} \cdot
                                 \left( {\bf H}^{(0)} + {\bf f}^{(0)} + 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} \;.
-\end{multline}
+\end{equation}
+%\end{multline} % PREPRINT
 %
@@
  \begin{align}
-  {\bf B}^{(10)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B10} \\
-  {\bf B}^{(20)}_{\alpha\beta} &= \frac{1}{2} 
+  {\bf B}^{(1)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B10} \\
+  {\bf B}^{(2)}_{\alpha\beta} &= \frac{1}{2} 
      \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B20}
@@
 i.e., from Eq.~\eqref{e:final-model.HF} (black solid squares in Figure~\ref{f:fig-1}) 
-as well as truncated at linear susceptibilities ${\bf B}^{(1)}$ 
+as well as truncated at linear susceptibilities $\mathbb{B}^{(1)}$ 
 from Eqs.~\eqref{e:dD-Taylor} (blue solid circles) and \eqref{e:final-model.General} (blue open circles),
@@
 \begin{table*}%[]
-\caption{{\bf Performance of the DMS tensors of ground\hyp{}state water molecule
-in uniform and non\hyp{}uniform electric field$^{a,b}$}
+\caption[Performance of the DMS tensors of ground\hyp{}state water molecule
+in uniform and non\hyp{}uniform electric field]{\label{t:results}
+{\bf Performance of the DMS tensors of ground\hyp{}state water molecule
+in uniform and non\hyp{}uniform electric field}  
+\footnotemark[1]\footnotemark[2]
 }
-\label{t:results}
-\begin{ruledtabular}
+%\begin{ruledtabular}
 \resizebox{\textwidth}{!}{%
 \begin{tabular}{cddddddcddddcdd}
-%\hline\hline
-\toprule
+\hline\hline
+%\toprule
 \multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalised DMS}}
@@
 \textbf{}         
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^c$}
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^d$}
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$$^d$}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[3]}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[4]}
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$\footnotemark[4]}
 &
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$$^e$} 
-& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$$^e$} 
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[5]} 
+& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$\footnotemark[5]} 
 &  
@@
 \multicolumn{15}{c}{} \\ 
-\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}$^{f}$} \\ 
+\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}\footnotemark[6]} \\ 
 \multicolumn{15}{c}{} \\ \hline
@@
 S & 78.07 & (40.53) & - &  -  & - &  -  && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ 
-\bottomrule
+%\bottomrule
+\hline\hline
 \end{tabular}%
 }
-\end{ruledtabular}
-%
-\footnotesize{\flushleft{
-$^a$ $R$-squared coefficients are shown in parentheses. `$n.d.$' stands for `not determined'. 
-$^b$ Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ 
-where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.
-$^c$ \emph{Ab initio} HF model, Eq.~\eqref{e:final-model.HF}.
-$^d$ \emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.
-$^e$ Linear\hyp{}regression generalised model, Eq.~\eqref{e:final-model.General}.
-$^f$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields 
-that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).\\
-}}
+%\end{ruledtabular}
+\footnotetext[1]{$R$-squared coefficients are shown in parentheses. `$n.d.$' stands for `not determined'.}
+\footnotetext[2]{Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.}
+\footnotetext[3]{\emph{Ab initio} HF model, Eq.~\eqref{e:final-model.HF}.}
+\footnotetext[4]{\emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.}
+\footnotetext[5]{Linear\hyp{}regression generalised model, Eq.~\eqref{e:final-model.General}.}
+\footnotetext[6]{`W', `M' and `S' denote `weak', `moderate' and `strong' electric fields that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).}
 \end{table*}
@@
 \begin{table}[b]
-\caption{{\bf Average electric fields in statistical sets of electrostatically perturbed ground states
-of water molecule surrounded by point charges$^a$}
+\caption[Average electric fields in statistical sets of electrostatically perturbed ground states
+of water molecule surrounded by point charges]
+{{\bf Average electric fields in statistical sets of electrostatically perturbed ground states
+of water molecule surrounded by point charges\footnotemark[1]}
 }
@@
 \begin{tabular}{ldcdcdcd}
-\multirow{2}{*}{\textbf{Set$^b$}} & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
+\multirow{2}{*}{\textbf{Set\footnotemark[2]}} 
+                                  & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
                                   & \multicolumn{3}{c}{\textbf{Oxygen Atom}} & \textbf{} 
@@
 %
-\footnotesize{\flushleft{
-$^a$ Each set was composed of 100 samples differing in the configuration of 40 charges generated from a uniform distribution.
-$^b$ `W', `M' and `S' denote `weak', `moderate' and `strong' electric fields, respectively.\\
-}}
-
+\footnotetext[1]{Each set was composed of 100 samples differing in the configuration of 40 charges generated from a uniform distribution.}
+\footnotetext[2]{`W', `M' and `S' denote `weak', `moderate' and `strong' electric fields, respectively.}
 \end{table}
@@
 No.~665778. \InlineImage{eu-logo.eps} This project is funded by National Science Centre, Poland 
-(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 
+(registration No.~2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 
 %Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 

### 2018-08-13 papers/1_dmatpol/submit/do-submit.sh

@@ -3,4 +3,6 @@ base=paper-submit
 
 latex  ${base}.tex
 bibtex ${base}.aux
+latex  ${base}.tex
+latex  ${base}.tex
 dvipdf ${base}.dvi

### 2018-08-22 papers/1_dmatpol/submit/paper-submit.tex
Revision. Correct reference format - removing footnote.

@@
 \usepackage{bm}% bold math
-%\usepackage{txfonts}
-%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
-%\linenumbers\relax % Commence numbering lines
 
@@
 % tables
-%\usepackage{dcolumn}
 \usepackage{multirow}
-%\usepackage{hhline}
 \usepackage{booktabs}
-%\usepackage{longtable}
 
@@
 \usepackage{amsbsy}
-%\usepackage{bm}
 \usepackage{mathrsfs}
 \usepackage{upgreek}
-%\allowdisplaybreaks
-
-% symbols
-%\usepackage{gensymb}
-%\usepackage{textcomp}
 
@@
 
-\title{One-Particle Density Matrix Polarisation Susceptibility Tensors}
+\title{One-Particle Density Matrix Polarization Susceptibility Tensors}
 
@@
 \email[]{blasiak.bartosz@gmail.com}
-\homepage[]{https://www.polonez.pwr.edu.pl}
-%\thanks{Bulak}
+\homepage[]{http://www.polonez.pwr.edu.pl}
 \affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
@@
 \begin{abstract}
-
-%Considering the solvent effect on the correlated electronic wavefunctions, especially
-%those that describe the electronically excited state, is a great challenge of Quantum Chemistry.
 
@@
 The analytic approximate expressions are derived at the Hartree\hyp{}Fock theory, 
-which serves as a basis for construction of generalised model 
+which serves as a basis for construction of generalized model 
 that is designed for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix.
@@
 and quadratic terms with respect to the electric field are accounted for,
-the electric field\hyp{}induced polarisation energies, dipole moments 
+the electric field\hyp{}induced polarization energies, dipole moments 
 and quadrupole moments are quantitatively described by the present theory 
@@
 Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques 
-that link extremely accurate but localised pieces of information with certain knowledge of an atomistic 
-organisation and intermolecular interaction on relatively large spatial scales, typically ranging beyond 
+that link extremely accurate but localized pieces of information with certain knowledge of an atomistic 
+organization and intermolecular interaction on relatively large spatial scales, typically ranging beyond 
 few nanometers.
@@
 within a QM level.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
-%Gao.Truhlar.Wang.Mazack.Loffler.Provorse.Rehak.ACR.2014}
 
@@
 Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018,Lewis.Gruenke.Oliver.Ballottari.Bassi.Fleming.JPCL.2016} 
-Interface between QM and MM levels is usually realised by electrostatic embedding and only rarely 
+Interface between QM and MM levels is usually realized by electrostatic embedding and only rarely 
 the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
@@
 The main postulate of the present work is that the associated change in the one\hyp{}particle density matrix, 
-$\delta {\bf D}$, can be represented in a product form of the generalised perturbation susceptibility ${\bf B}_i$
+$\delta {\bf D}$, can be represented in a product form of the generalized perturbation susceptibility ${\bf B}_i$
 and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$, evaluated at certain site $i$,
@@
 It is worth mentioning that, at the moment, there is no simple \emph{ab initio} method to compute 
-polarisation\hyp{}induced electron deformation densities without considering the entire molecular 
+polarization\hyp{}induced electron deformation densities without considering the entire molecular 
 aggregate by using full quantum chemistry calculations.\cite{
@@
 In this work, we explore the susceptibility of the molecule's electronic density
-to polarise due to external electrostatic perturbation from first principles. 
+to polarize due to external electrostatic perturbation from first principles. 
 The deformation density is expressed in terms
-of the one\hyp{}particle density matrix polarisation susceptibility tensors 
+of the one\hyp{}particle density matrix polarization susceptibility tensors 
 that are properties of isolated molecules and can be
@@
 (scalar) is beneficial since it provides the directional tensor description
-of the polarisation that could be combined with other effects such as Pauli repulsion 
+of the polarization that could be combined with other effects such as Pauli repulsion 
 or dispersion,\cite{Mandado.Hermida-Ramon.JCTC.2011}
-in contrast to the standard interaction energy approaches. We show that accurate polarisation 
+in contrast to the standard interaction energy approaches. We show that accurate polarization 
 of the one\hyp{}particle density
-matrix is achievable and can be realised as long as the one\hyp{}particle density matrices
+matrix is achievable and can be realized as long as the one\hyp{}particle density matrices
 are available. It is believed that the proposed method
@@
 alternative approach in fragment\hyp{}based
-modelling of extended molecular aggregates that is not limited to the ground state chemistry.
-
-We introduce the necessary theory and derive the \emph{ab initio} density matrix susceptibility model
-at the HF approximation\cite{Roothaan.RevModPhys.1951} 
-in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalisation to the 
+modeling of extended molecular aggregates that is not limited to the ground state chemistry.
+
+The necessary theory and the derivation of the \emph{ab initio} density matrix susceptibility model
+at the HF approximation\cite{Roothaan.RevModPhys.1951} are elaborated
+in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalization to the 
 arbitrary one\hyp{}electron density with electronic correlation which is derived in Section~\ref{s:2}. 
 The implementation details in a computer code as well as the validation procedures are detailed
-in Section~\ref{s:3}. We then discuss the performance of the models in Section~\ref{s:4}.
-Finally, in Section~\ref{s:5} we conclude the presentation with a few 
+in Section~\ref{s:3}. The performance of the models is then discussed in Section~\ref{s:4}.
+Finally, in Section~\ref{s:5} the presentation is concluded with a few 
 remarks for future development.
@@
 excited state one\hyp{}electron densities must be excited state solutions of the Schr{\"o}dinger equation
-with the exact energy\hyp{}minimised ground state.
+with the exact energy\hyp{}minimized ground state.
 
 In the limit of a weak and spatially uniform electric field,
-the polarisation\hyp{}induced density matrix change of any electronic state
+the polarization\hyp{}induced density matrix change of any electronic state
 or electronic transition is given by the following Taylor expansion:
@@
 In a general case, i.e., when electric field is relatively strong 
-and non\hyp{}uniform, single\hyp{}centred expansions, like the one 
+and non\hyp{}uniform, single\hyp{}centered expansions, like the one 
 from Eq.~\eqref{e:dD-Taylor},
-%although valid for an arbitrary level of theory in uniform fields,
-are not sufficient enough to accurately describe the local polarisation
+are not sufficient enough to accurately describe the local polarization
 of electron density distribution.
@@
 
-Due to all of these reasons, we adopt the following approach: 
-we first develop a qualitative \emph{ab initio} distributed\hyp{}site 
-model at the Hartree\hyp{}Fock level of theory. Further on, we generalise it 
+Due to all of these reasons, the following approach is adopted here: 
+a qualitative \emph{ab initio} distributed\hyp{}site 
+model at the Hartree\hyp{}Fock level of theory is developed first. 
+Further on, we generalize it 
 to an arbitrary level and type of density matrix
@@
 be directly related to the
-polarisation\hyp{}induced multipole moments because they are
+polarization\hyp{}induced multipole moments because they are
 a measure of the distortion of the electron density distribution due to the external electric
-field. 
-Therefore, we start our analysis by considering the induced 
-ground\hyp{}state electronic dipole moment of a closed\hyp{}shell molecule, 
+field. Therefore, as a starting point, the induced 
+ground\hyp{}state electronic dipole moment of a closed\hyp{}shell molecule
+will be considered,
 %
@@
 %
-(In this work we shall use the double struck font for matrices of vectors, so that
+(In this work the double struck font for matrices of vectors or higher\hyp{}order tensors is used, so that
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
@@
 %
-The change in the molecular orbitals can be parameterised as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
+The change in the molecular orbitals can be parameterized as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
 by defining the idempotency\hyp{}preserving variation
@@
 ${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$.
-%Such a condition is necessary in order to meet the charge conservation
-%requirement (see Appendix~\ref{a:orig-dep} for a detailed discussion).
-%Therefore orthogonality of AO's is assumed from that point, unless stated otherwise.
-%Also, from this point onwards we assume that the molecular orbitals 
-%and the transformation matrix $\BM \Delta$ are real.
 By substituting the McWeeny's result from Eq.~\eqref{e:dD-1}
@@
 induced dipole moment obtained in Eq.~\eqref{e:L-projector} is independent of the origin. 
-This guarantees that the theory fulfils the total charge conservation requirement and,
+This guarantees that the theory fulfills the total charge conservation requirement and,
 therefore, in later course of the paper, the dipole origin will be omitted.
@@
 Let the induced dipole moment be partitioned into separate contributions
-from the occupied localised MO's (LMO's), i.e.,
+from the occupied localized MO's (LMO's), i.e.,
 %
@@
 %
-In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarisability tensor
+In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarizability tensor
 associated with the $i$th LMO whereas ${\bf F}({\bf r}_i)$ 
@@
 is a real rectangular matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
-change in the LCAO-MO matrix due to the polarisation process.
+change in the LCAO-MO matrix due to the polarization process.
 The solution for $\delta {\bf C}$ is
@@
 
-Now we can obtain the final expression for the density matrix change 
-due to external electric field that reproduces the 
-induced dipole moment of the molecule.
-The change in the LCAO\hyp{}MO coefficient is parameterised as
+Once the approximate relationship between the orbital change and the external electric field
+has been established in Eq.~\eqref{e:mu-ind.matrix.linear.solution},
+one can obtain the final expression for the density matrix change 
+that reproduces the induced dipole moment of the molecule.
+The change in the LCAO\hyp{}MO coefficient is parameterized as
 %
@@
 %
-where the \emph{density matrix polarisation susceptibility tensor} is
+where the \emph{density matrix polarization susceptibility tensor} is
 %
@@
 Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ 
-describes the local linear polarisation affinity 
+describes the local linear polarization affinity 
 of the density matrix element 
@@
 Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
-It is emphasised here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
+It is emphasized here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
 are unique properties of electronically unperturbed molecule. 
@@
 approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
-limited by the flavour of the distributed polarisability approximation), 
-but not higher multipole moments and polarisation energy, are correctly described.
+limited by the flavor of the distributed polarizability approximation), 
+but not higher multipole moments and polarization energy, are correctly described.
 
@@
 vector susceptibility, different for each density matrix element. This feature will be used to
-generalise the model to an arbitrary level of theory and order of electric field response in the next section.
-
-\subsection{Generalised Density Matrix Susceptibility Model}
+generalize the model to an arbitrary level of theory and order of electric field response in the next section.
+
+\subsection{Generalized Density Matrix Susceptibility Model}
 
@@
 Eq.~\eqref{e:final-model.HF}, which is the final result of the previous section, 
-can be used as a guideline to construct a general model of density matrix polarisation.
+can be used as a guideline to construct a general model of density matrix polarization.
 Quadratic (and higher\hyp{}order) terms with respect to the electric field
@@
 In the above equation, the density matrix susceptibilities
-${\bf B}^{(i; R)}_{\alpha\beta}$, distributed over $M$ sites,
+$\mathbb{B}^{(i; R)}$, distributed over $M$ sites,
 denote the $R$-th order response with respect to the electric field.
@@
 ${\bf B}^{(i; 1)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 2)}_{\alpha\beta}$
-is a $3\times 3$ matrix. Also, the distributed expansion centres
-(labelled as $i$) are generalised to any point in the molecule.\footnote{More general
+is a $3\times 3$ matrix. Also, the distributed expansion centers
+(labeled as $i$) are generalized to any point in the molecule.
+In principle, a slightly more general
 contribution due to square of electric field could 
-be $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$. 
-However, for simplicity, we neglect the contributions from different sites $i$ and $j$
-in the present discussion and leave it for future development.
-} Note that Eq.~\eqref{e:final-model.General} can refer to any type of one\hyp{}particle density matrix
+be formulated here as $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$,
+defining the second\hyp{}order susceptibilities due to pairs of different sites $i$ and $j$. 
+However, in the present work such contributions are neglected and only `diagonal' susceptibilities
+(for simplicity denoted as $\mathbb{B}^{(i; 2)}$ instead of $\mathbb{B}^{(ii; 2)}$) are considered.
+
+Eq.~\eqref{e:final-model.General} can refer to any type of one\hyp{}particle density matrix
 such as ground\hyp{}state, excited\hyp{}state or transition density matrix.
-However, the exact mathematical form of the generalised density matrix susceptibilities 
+However, the exact mathematical form of the generalized density matrix susceptibilities 
 from Eq.~\eqref{e:final-model.General} is unknown.
 
-\subsubsection{Determining the generalised susceptibility tensors by linear regression}
+\subsubsection{Determining the generalized susceptibility tensors by linear regression}
 
@@
 %
-are known for each sample (overline symbolises the exact estimate).
-Now, for each pair of the AO indices we construct the following 
-parameterisation of Eq.~\eqref{e:final-model.General}:
+are known for each sample (overline symbolizes the exact estimate).
+Now, for each pair of the AO indices the following 
+parameterization of Eq.~\eqref{e:final-model.General} is constructed:
 %
@@
 \end{pmatrix}^{\rm T}
-$, we define a loss function $Z$ 
-that is subject to the least\hyp{}squares minimisation,
+$, a loss function $Z$ 
+that is subject to the least\hyp{}squares minimization, is defined as
 %
@@
 from the uniform distribution in the interval 
-of $\vert 0.0404 \vert$ a.u. generating average field of 0.035 a.u. in an ensemble of 100 samples.
+of $\vert 0.0404 \vert$ a.u. generating average field of 0.038 a.u. in an ensemble of 100 samples.
 Cartesian coordinates of point charges were drawn from the uniform distribution 
 of a set of forty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
-with an origin that is located in the centre of mass of water molecule.
+with an origin that is located in the center of mass of water molecule.
 In selection of random points, space inside the union of spheres surrounding
@@
 performed for each validation procedure.
-In the case of non\hyp{}uniform electric fields, the distribution centres 
+In the case of non\hyp{}uniform electric fields, the distribution centers 
 of the \emph{ab initio} DMS model (Eq.~\eqref{e:final-model.HF}) were assumed 
-to be the LMO centroids (resulting in the five\hyp{}centre model), 
-while the atoms were used for the generalised DMS models (Eq.~\eqref{e:final-model.General}, 
-resulting in the three\hyp{}centre models).
+to be the LMO centroids (resulting in the five\hyp{}center model), 
+while the atoms were used for the generalized DMS models (Eq.~\eqref{e:final-model.General}, 
+resulting in the three\hyp{}center models).
 This is because certain LMO centroids are often located at the same point which would
 introduce linear dependencies in the Hessian matrix from Eq.~\eqref{e:Newton.gH}
-when using the generalised DMS scheme.
+when using the generalized DMS scheme.
 
 In order to validate all the models discussed in this contribution,
-the electric field\hyp{}induced polarisation energy, 
-dipole moment and quadrupole moment were analysed in detail.
-The electric field\hyp{}induced polarisation energy at the HF level, $\delta E^{\rm Pol}$,
+the electric field\hyp{}induced polarization energy, 
+dipole moment and quadrupole moment were analyzed in detail.
+The electric field\hyp{}induced polarization energy at the HF level, $\delta E^{\rm Pol}$,
 was computed from the following exact formula:
@@
 and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
-the polarisation of the electron density, respectively. 
+the polarization of the electron density, respectively. 
 In order to compare the performance of the usual 
-distributed dipole\hyp{}dipole polarisability approach,\cite{Li.Netzloff.Gordon.JCP.2006}
+distributed dipole\hyp{}dipole polarizability approach,\cite{Li.Netzloff.Gordon.JCP.2006}
 the approximate interaction energy was also evaluated by using the expression:
@@
 %
-in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarisability tensor associated with the
+in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarizability tensor associated with the
 $i$th LMO centroid (see Eq.~\eqref{e:dmu-dpol}).
-The distributed polarisabilities were computed by utilising the coupled\hyp{}perturbed Hartree\hyp{}Fock
+The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
 method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
-Orbital localisation was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
+Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
 
@@
 %
-where the subscript `CM' denotes the centre of mass of water molecule.
-When the distributed polarisabilities were used instead of the polarisation\hyp{}induced 
+where the subscript `CM' denotes the center of mass of water molecule.
+When the distributed polarizabilities were used instead of the polarization\hyp{}induced 
 difference density matrices, primitive induced quadrupole moments were computed from
@@
 and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
-was used. Before all the calculations, water molecule was energy\hyp{}optimised
+was used. Before all the calculations, water molecule was energy\hyp{}optimized
 with a maximum energy, force and displacement thresholds of 10$^{-6}$, 3$\times$$10^{-4}$ 
 and 1.2$\times$$10^{-3}$ a.u., respectively.
-For all the calculations of the density matrix polarisation susceptibility tensors
+For all the calculations of the density matrix polarization susceptibility tensors
 from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
-atomic basis was orthogonalised by using the L{\"o}wdin symmetric orthogonalisation scheme,\cite{Mayer.IJQC.2002}
+atomic basis was orthogonalized by using the L{\"o}wdin symmetric orthogonalization scheme,\cite{Mayer.IJQC.2002}
 and further back\hyp{}transformed to the non\hyp{}orthogonal AO basis.
@@
 with respect to the electric field from Eq.~\eqref{e:dD-Taylor}, 
-the following finite\hyp{}difference formulae were utilised:
+the following finite\hyp{}difference formulae were utilized:
 %
@@
 
-Polarisation under the uniform electric field is the simplest type of electrostatic perturbation
+Polarization under the uniform electric field is the simplest type of electrostatic perturbation
 whereby it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
-directly by numerical differentiation. Here we consider DMS up to second order, i.e.,
+directly by numerical differentiation. Here, the DMS's up to second order are considered, i.e.,
 %
@@
 The \emph{ab initio} model from Eq.~\eqref{e:final-model.HF} and
-the generalised model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
-The associated error distributions in prediction of polarisation energies 
+the generalized model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
+The associated error distributions in prediction of polarization energies 
 are shown in scatter plot in Figure~\ref{f:fig-1}, whereas the
-statistical performance descriptors of polarisation energies and induced multipole moments
+statistical performance descriptors of polarization energies and induced multipole moments
 are gathered in Table~\ref{t:results}.
@@
 from Eqs.~\eqref{e:dD-Taylor} (blue solid circles) and \eqref{e:final-model.General} (blue open circles),
-incorrectly reproduce the polarisation energy.
+incorrectly reproduce the polarization energy.
 Interestingly, the predicted values are in strong
@@
 obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
-or by using the generalised DMS model (red open circles)
+or by using the generalized DMS model (red open circles)
 dramatically reduces the errors
 leading to quantitative accuracy in all tested field magnitudes 
-with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalised DMS model.
+with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalized DMS model.
 On the other hand, the
@@
 This observation can be understood given the fact that the induced dipole
-and quadrupole moment is approximately proportional to the external electric field, whereas the polarisation
+and quadrupole moment is approximately proportional to the external electric field, whereas the polarization
 energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
@@
 
-In comparison to the dipole\hyp{}dipole polarisability approach, 
+In comparison to the dipole\hyp{}dipole polarizability approach, 
 quadratic DMS models perform only slightly worse 
-(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalised quadratic DMS
-model and the polarisability model are used, respectively).
+(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalized quadratic DMS
+model and the polarizability model are used, respectively).
 However, quadratic DMS models better predict the induced multipole moments with
-%work better
-%when predicting the induced multipole moments with
 $R^2$ systematically greater than 99\%. 
-Certainly, taking into account the distributed quadrupole polarisabilities
-would greatly reduce errors of the standard distributed polarisability approach.
+Certainly, taking into account the distributed quadrupole polarizabilities
+would greatly reduce errors of the standard distributed polarizability approach.
 Nevertheless, it is clear that the second\hyp{}order
 DMS models are quantitative even at very strong uniform electric fields.
-%On the other hand, DMS models better predict the induced dipole and quadrupole 
-%moments even in weak fields.
 
@@
 %\toprule
-\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalised DMS}}
+\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalized DMS}}
                                                                   &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
@@
 &  
-& \multicolumn{2}{c}{\textbf{Polarisability}}        \\ \hline
+& \multicolumn{2}{c}{\textbf{Polarizability}}        \\ \hline
 \multicolumn{15}{c}{} \\ 
@@
 \footnotetext[4]{\emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.}
-\footnotetext[5]{Linear\hyp{}regression generalised model, Eq.~\eqref{e:final-model.General}.}
+\footnotetext[5]{Linear\hyp{}regression generalized model, Eq.~\eqref{e:final-model.General}.}
 \footnotetext[6]{`W', `M' and `S' denote `weak', `moderate' and `strong' electric fields that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).}
@@
 To study the capability of the distributed DMS models to capture 
-density matrix polarisation in inhomogeneous electric fields, a water molecule
-surrounded by point charges is analysed.
+density matrix polarization in inhomogeneous electric fields, a water molecule
+surrounded by point charges is analyzed.
 Three different types of point charge environment that generate
@@
 in liquid phase.\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013,Reischl.Kofinger.Dellago.MolPhys.2009}
-Performance descriptors of the \emph{ab initio} and generalised 
+Performance descriptors of the \emph{ab initio} and generalized 
 models of the density matrix susceptibility, as well as 
-the conventional dipole\hyp{}dipole distributed polarisability
+the conventional dipole\hyp{}dipole distributed polarizability
 approach, are shown in Table~\ref{t:results}, whereas the error distributions
@@
 the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} 
-as well as the generalised linear DMS model, are insufficient to correctly describe the polarisation energy 
+as well as the generalized linear DMS model, are insufficient to correctly describe the polarization energy 
 in non\hyp{}uniform electric field, although the correlation with benchmark is still very clear with a slope of $\sim$2
@@
 in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$; Figure~\ref{f:fig-2}D and \ref{f:fig-2}E).
-Induced quadrupole moments are well described only by the generalised DMS model.
+Induced quadrupole moments are well described only by the generalized DMS model.
 For example, in moderate electric fields RMSQ is 10.4$\times 10^{-3}$D\AA{ }with $R^2=93.7\%$
@@
 refers to the model from Eq.~\eqref{e:final-model.HF} and
-`Generalised DMS' (panels B, E and H) 
-refers to the generalised model with $R_{\rm max}=2$ 
+`Generalized DMS' (panels B, E and H) 
+refers to the generalized model with $R_{\rm max}=2$ 
 where $R_{\rm max}$ is the maximum power
 in the expansion from Eq.~\eqref{e:final-model.General}.
-Comparison with the distributed dipole\hyp{}dipole polarisability model %(Eq.~\eqref{e:dE-dpol})
+Comparison with the distributed dipole\hyp{}dipole polarizability model %(Eq.~\eqref{e:dE-dpol})
 is also shown in this Figure in panels C, F and I. 
@@
 
-The most sophisticated model studied in this work, the generalised quadratic DMS model, yields excellent predictions 
-of the polarisation energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
-The errors in predicting polarisation energy are reduced by over an order of magnitude when compared with
+The most sophisticated model studied in this work, the generalized quadratic DMS model, yields excellent predictions 
+of the polarization energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
+The errors in predicting polarization energy are reduced by over an order of magnitude when compared with
 the linear DMS models. Improvement of accuracy can also be observed in the case of induced multipole moment predictions
@@
 What is particularly interesting is that in weak and moderate electric fields
-the generalised quadratic DMS model performs better than the distributed dipole\hyp{}dipole
-polarisability model, whereas only in strong electric fields the situation is reversed
-(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.1\%$ for the DMS and polarisability model, respectively; 
+the generalized quadratic DMS model performs better than the distributed dipole\hyp{}dipole
+polarizability model, whereas only in strong electric fields the situation is reversed
+(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.1\%$ for the DMS and polarizability model, respectively; 
 see also Figure~\ref{f:fig-2}B and \ref{f:fig-2}C).
@@
 Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
-polarisable force field, that the average electric field experienced by water molecule in liquid phase
+polarizable force field, that the average electric field experienced by water molecule in liquid phase
 in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
@@
 and \ref{t:fields}) for which the
-accuracy of the generalised quadratic DMS model is excellent.
+accuracy of the generalized quadratic DMS model is excellent.
 Water molecules can experience
@@
 This electric field magnitude falls into our strong electric field regime (denoted as `S')
-in which the generalised quadratic DMS model still provides accurate predictions (see also error distributions 
+in which the generalized quadratic DMS model still provides accurate predictions (see also error distributions 
 in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H). Therefore, it is believed that
-the generalised quadratic DMS model is capable of accurately describing 
+the generalized quadratic DMS model is capable of accurately describing 
 the polarization of ground\hyp{}state water molecule in inhomogeneous 
@@
 
-%In this work, the \emph{ab initio} model of the 
-%one\hyp{}particle density matrix susceptibility tensors
-%at the HF level was derived for a closed shell molecule 
-%in a spatial electric field distribution. It was found 
-%that the change in the density matrix element is approximately 
-%linearly proportional to the electric field evaluated at the 
-%distributed site associated with the centre of charge of the 
-%localised molecular orbital. Subsequently, the \emph{ab initio} 
-%model served as a basis for the formulation of the generalised 
-%model that is, in principle, not limited to the HF theory,
-%and takes into account also quadratic and higher\hyp{}order terms 
-%with respect to the electric field. The generalised model, which 
-%consisted of a set of adjustable parameters, defined the first\hyp{} 
-%and second\hyp{}order density matrix susceptibilities that were 
-%determined through the least\hyp{}squares linear regression 
-%with respect to the training ensemble of electrostatically 
-%perturbed states.
-%
-%The above theoretically developed models were then used 
-%to calculate the electric field\hyp{}induced interaction 
-%energies, dipole moments and quadrupole moments of water 
-%molecule that is subjected to perturbation due to uniform 
-%electric field, as well as various point charge environments 
-%generating non\hyp{}uniform electric field of weak, moderate
-%and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
-%interaction energies, the atom\hyp{}distributed generalised model can reach quantitative level of accuracy
-%for prediction of interaction energies, dipole moments and even quadrupole moments
-%once, at least quadratic terms with respect to the electric field, are taken into account. 
-In this work, it was demonstrated that the generalised DMS models from Eq.~\eqref{e:final-model.General}
-can accurately describe the polarization\hyp{}induced deformation of the one\hyp{}particle ground\hyp{}state electronic density
+In this work, it was demonstrated that the generalized DMS models from Eq.~\eqref{e:final-model.General}
+can accurately describe the polarization\hyp{}induced deformation 
+of the one\hyp{}particle ground\hyp{}state electronic density
 due to the uniform and non\hyp{}uniform distribution of electric field.
-It was also shown that the accuracy of the model can be improved by extending the parameter space of the susceptibility tensors.
-Even at the current state of development, the generalised quadratic DMS model
-is of comparable and quantitative energetic accuracy as the distributed polarisability model (RMSE of 0.4~kcal/mol in strong 
+It was also shown that the accuracy of the model can be improved 
+by extending the parameter space of the susceptibility tensors.
+Even at the current state of development, the generalized quadratic DMS model
+is of comparable and quantitative energetic accuracy as the distributed 
+polarizability model (RMSE of 0.4~kcal/mol in strong 
 electric fields and less than 0.06~kcal/mol in moderate and weak fields).
-%It is emphasised here that the model can be easily extended by taking into account
-%third\hyp{}order terms as well if fine accuracy is required.
-%Moreover, the generalized DMS model can be used to any type of one\hyp{}particle density matrices at arbitrary level of theory.
-%This however falls out of scope of this contribution and it will be studied
-%elsewhere. 
 Moreover, any type of one\hyp{}particle density matrix could be treated by using
-generalised DMS models, which is much more difficult to realise when a conventional multipole
-model is used.
-Therefore, it is believed that
-the generalised DMS approach
+generalized DMS models, which is much more difficult to realize when a conventional multipole
+model is used. Therefore, it is believed that the generalized DMS approach
 has an important advantage over the multipole\hyp{}based approach, i.e., it provides more detailed
-tensor description of the spatial polarisation of the electron density distribution, that could be
+tensor description of the spatial polarization of the electron density distribution, that could be
 combined for example with density\hyp{}based methods of interaction energy calculations.\cite{Mandado.Hermida-Ramon.JCTC.2011}
 
-It is anticipated here that the generalised models from Eq.~\eqref{e:final-model.General}
+It is anticipated here that the generalized models from Eq.~\eqref{e:final-model.General}
 should be studied in the future with an emphasis 
-on their further optimisation and application for molecular aggregates.
-For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
+on their further optimization and application for molecular aggregates.
+For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, 
+as well as the choice of distributed
 sites should be further investigated, preferably considering post\hyp{}Hartree\hyp{}Fock methods
@@
 polarization of ground state's electron density,
-the generalised density matrix susceptibility tensors 
+the generalized density matrix susceptibility tensors 
 could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
@@
 (registration No.~2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 
-%Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 
-%for reading and commenting on the manuscript.
 \end{acknowledgments}
@@
  \right]}
- \right\} = 0
+ \right\} = 0 \;.
 \end{equation}
@@
 the above expression is not valid. 
-Therefore, the polarisation\hyp{}induced distributed dipole moments 
+Therefore, the polarization\hyp{}induced distributed dipole moments 
 defined in Eq.~\eqref{e:L-projector} are exactly origin\hyp{}independent
-as long as the AO basis is orthogonalised.
+as long as the AO basis is orthogonalized.
 In such a case, one can compute one\hyp{}electron dipole integrals with respect to any origin 
@@
 to the original (non\hyp{}orthogonal) AO basis, 
-${\bf X} \cdot \mathbb{B}^{(i;1)} \cdot {\bf X}$, where $\bf X$ is an orthogonaliser.
+${\bf X} \cdot \mathbb{B}^{(i;1)} \cdot {\bf X}$, where $\bf X$ is an orthogonalizer.
 The numerical tests confirmed that $\Tr\left[\delta{\bf D}\cdot{\bf S}\right] \sim 10^{-16}$
@@
 % -----------------------
+%\bibliographystyle{aipnum4-1}
 \bibliography{references}

### 2018-10-01 papers/1_dmatpol/revision/paper-revision-1.tex
Revision files for DMS paper added

--- 2018-08-22 papers/1_dmatpol/submit/paper-submit.tex
@@
 %%
-\documentclass[aip,jcp,preprint,amsmath,amssymb,floatfix]{revtex4-1}
+\documentclass[aip,jcp,reprint,amsmath,amssymb,floatfix]{revtex4-1}
 %\documentclass[aip,reprint]{revtex4-1}
@@
 or density functional theory (DFT) levels\cite{Hohenberg.Kohn.PhysRev.1964,Kohn.Sham.PhysRev.1965,Nguyen.Pachter.Day.JCP.2014}.
-Unfortunately, while they can approximately describe the environment composed of molecules in their ground states,
+While they can approximately describe the environment composed of molecules in their ground states,
 they cannot be used for highly correlated wavefunctions such as electronically excited states. 
+In such cases, the full QM method needs to be utilized on the excited state wavefunction, that is 
+embedded within the EFP environment through the electrostatic or polarizable embedding.\cite{Ghosh.JPCA.2017}
 
@@
 whereby it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
-directly by numerical differentiation. Here, the DMS's up to second order are considered, i.e.,
+by using either the coupled\hyp{}perturbed Hartree\hyp{}Fock method\cite{Horn.Head-Gordon.JCP.2015}
+or numerical differentiation. Here, the DMS's up to second order are considered, i.e.,
 %
@@
 
+\subsection{\label{s:computational-cost}Computational cost and implementation of the DMS models}
+
+Although the cost of computation of the perturbed one\hyp{}electron density matrices
+from the DMS tensors and the distribution of the electric field is negligible, 
+obtaining the generalized DMS matrix elements for a particular molecule
+involves substantial amount of full QM calculations necessary to fit the DMS tensor elements.
+This is, however, a relatively expensive operation that has to be performed only once.
+%To get statistically relevant set of perturbed states, it is recommended to perform at least 100 
+%full QM calculations. 
+Since each of such preparatory calculations is an independent computational task, 
+its automatization and parallelization is straightforward. Moreover, the implementation
+of the method does not depend on the level of theory because only the perturbed density matrices
+are needed. The computation of the gradient 
+and Hessian matrix is inexpensive and its efficiency is limited only by the speed of Hessian matrix inversion.
+Therefore, the method can be applied to all molecular fragments at any level of theory, provided
+the one\hyp{}particle density matrices are available. In the illustrations shown in this work,
+a set of 100 independent HF/6-311++G** calculations was sufficient to get statistically relevant set of perturbed 
+ground state density matrices. This task was concluded within approximately 5 minutes per each DMS fit
+on a single\hyp{}CPU computer.
+
 \section{\label{s:5}Summary and a few concluding remarks}

### 2018-10-01 papers/1_dmatpol/revision/reply-letter-1.tex
Revision files for DMS paper added

\documentclass[12pt]{article}
%\usepackage{a4wide}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{xcolor}
%\usepackage[a4paper, margin=2.7cm]{geometry}
\usepackage{amsmath, amsthm, amsfonts, amssymb}
\usepackage{mathrsfs}           % \mathscr font.

\title{Review of Manuscript ID: A18.08.0126}
\author{Bartosz B{\l}asiak}
% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
 \section*{Reviewer \thereviewer}}

\newenvironment{point}
 {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
 {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point} \bigskip \noindent 
 {\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
 {\medskip \noindent \begin{sf} \color{blue} \textbf{Reply}:\ }
 {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\ #2
 \ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
 \medskip \end{sf}}

\begin{document}
\maketitle
\section*{Response to the reviewers}

I would like to thank the Editor and two Reviewers for the time and insights given to my manuscript 
and for supporting the work to be published in \emph{J. Chem. Phys.}
I appreciate the positive and constructive feedbacks given, and agree with the reviewers’ 
suggestions for revisions. The reviewers’ comments have been fully addressed as outlined below. 
In addition, I have also introduced some minor adjustments in the notation.
Now, I hope that the revised manuscript is suitable for publication in \emph{J. Chem. Phys.}

\clearpage
\reviewersection

Recommendation: Publish as is \\
New Potential Energy Surface: No \\
Reviewer  (Comments to the Author): \\

I think this manuscript is very interesting. The model developed provides a very good reference 
to be used in connection with testing of less sophisticated schemes, 
e.g. models where a quantum chemical description is coupled to a classical polarizable model 
for the environment. Also the suggested use of the model within the field of aggregation 
is very interesting. As such I recommend publication of this manuscript.

\begin{point}
One very minor issue relates to the statement in the introduction saying 
that the EFP model is limited to either HF or DFT. Indeed the EFP model has been 
formulated also within correlated wave function approaches like CC. 
 \label{pt:rev-1}
\end{point}

\begin{reply}
Thanks for pointing this out. 
On the one hand, the \emph{ab initio} parameters derived from the EFP methods
can be obtained by using only HF or DFT wavefunctions, 
because no correlated wavefunction formulation of EFP method
was developed so far. %However, I agree with the Reviewer that the EFP method is not limited to only HF or DFT
On the other hand, I agree with the Reviewer that the EFP method is not limited to only HF or DFT
in a sense that it is possible to use it together with correlated wavefunction methods like CC
in a hybrid fashion (such as CC-EOM/EFP method). 
%In such circumstances, the QM method
%describes electronically excited states, while the EFP method is used for ground-state solvent molecules.
Therefore, to clarify this issue, the following sentence (shown in bold font) was added on Page 3: 
\begin{quote}
While they [the EFP methods] can approximately describe the environment composed of molecules in their ground states,
they cannot be used for highly correlated wavefunctions such as electronically excited states. 
\textbf{\noindent
In such cases, the full QM method needs to be utilized on the excited state wavefunction, that is 
embedded within the EFP environment through the electrostatic 
or polarizable embedding. [D. Ghosh, \emph{J. Chem. Phys. A} {\bf 121}, 741 (2017).] }
\end{quote}
\end{reply}

\clearpage
\reviewersection
Recommendation: Revision \\
New Potential Energy Surface: No \\
Comments to the Author: \\

The paper entitled `One-Particle Density Matrix Polarization Susceptibility Tensors' 
presents a new way to describe electric field induced changes in a molecule by the mean of 
one-particle density matrix susceptibilities (DMS) interacting with a spacial distribution 
of electric field. The author applied this method to the water molecule considering uniform 
and non-uniform electric-field distributions and showed that the generalized DMS model 
is able to reproduce electric-field induced changes in the polarization energy, dipole moment 
and quadrupole moment characterized at the Hartree-Fock level of theory. This study 
is the first step prior to provide an efficient effective one-electron fragment potential 
to reduce the cost of quantum chemistry calculations in condensed phase. This is 
an interesting article that is suitable to be published in the Journal of Chemical Physics. 
Though, I have some minor comments/questions that should be answered prior to the publication.

\begin{point}
When I finished reading this MS, I was left hungering for a bit more. Since this DMS scheme 
could be a more precise alternative to multipole-based approaches to describe aggregates, 
I think an application on some small water aggregates (or dimer of water) and comparing 
to full HF calculations could be nice addition. But I understand that this method is still 
in its early stage and that this is probably not yet possible. In any case is a bit more 
testing on other systems may be also of mixed polarity would improve the quality of the MS. 
 \label{pt:rev-2.1}
\end{point}

\begin{reply}
I definitely agree with the Reviewer that an additional illustration on some molecular
dimer would much increase the quality of the present manuscript. Unfortunately, as the Reviewer also
understands, this requires a separate article since other effects such as Pauli exchange-repulsion
need to be included and combined with polarization in order to compare to the full HF result. 
Due to this reason, I decided to write a separate manuscript on that issue in the future.
\end{reply}

\begin{point}
The author should discuss what is the computational cost and applicability (automatization) 
of this method since quite a lot of calculations for all the spatial distributions are needed for fitting the DMS. 
 \label{pt:rev-2.2}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
On page 14 it is written in section IV.A.: 
`Polarization under the uniform electric field is the simplest type 
of electrostatic perturbation whereby it is possible to compute 
the exact DMS according to Eq. (5) directly by numerical differentiation.' 
Note that this is also possible analytically with the CPHF method. 
 \label{pt:rev-2.3}
\end{point}

\begin{reply}
I agree with the remark, thanks. I changed the sentence to incorporate this:
\begin{quote}
Polarization under the uniform electric field is the simplest type of electrostatic perturbation
whereby it is possible to compute the exact DMS according to Eq.~(5)
\textbf{\noindent
by using either the coupled-perturbed Hartree-Fock method [P. R. Horn and M. Head-Gordon, \emph{J. Chem. Phys.} {\bf 143}, 114111 (2015).]
or numerical differentiation.}
\end{quote}
\end{reply}

\end{document}

### 2018-10-08 papers/1_dmatpol/revision/paper-revision-1.tex
Revision of the paper submitted

@@
 %%
-\documentclass[aip,jcp,reprint,amsmath,amssymb,floatfix]{revtex4-1}
+\documentclass[aip,jcp,preprint,amsmath,amssymb,floatfix]{revtex4-1}
 %\documentclass[aip,reprint]{revtex4-1}
@@
 in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
-Note that if the total charge is conserved, the induced moments are independent of the origin.
+Note that if the total charge is conserved, the induced dipole moment is independent of the origin.
 
@@
 \begin{equation} \label{e:dmat-change.exact}
- \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
+ \delta {\bf D} =  {\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
                                    \left[ {\bf 1} + {\bf v}^\dagger \cdot {\bf v} \right]^{-1} \cdot
@@
  {\bf B}^{(i;1)}_{\alpha\beta} = 
-                               C_{\alpha i} {\bf b}_{\beta}^{(i;1)} + C_{\beta i} {\bf b}_\alpha^{(i;1)} %\\ PREPRINT
+                               C_{\alpha i}^{(0)} {\bf b}_{\beta}^{(i;1)} + C_{\beta i}^{(0)} {\bf b}_\alpha^{(i;1)} %\\ PREPRINT
                                 - \sum_\gamma 
                                  \left( 
-               D_{\alpha\gamma} C_{\beta i} + D_{\beta\gamma} C_{\alpha i}
+               D_{\alpha\gamma} C_{\beta i}^{(0)} + D_{\beta\gamma} C_{\alpha i}^{(0)}
                                  \right)
@@
 
-\subsection{\label{s:computational-cost}Computational cost and implementation of the DMS models}
+\subsection{\label{s:computational-cost}Computational cost and automatization of the generalized DMS model}
 
 Although the cost of computation of the perturbed one\hyp{}electron density matrices
-from the DMS tensors and the distribution of the electric field is negligible, 
-obtaining the generalized DMS matrix elements for a particular molecule
-involves substantial amount of full QM calculations necessary to fit the DMS tensor elements.
-This is, however, a relatively expensive operation that has to be performed only once.
+from the generalized DMS tensors and the distribution of the electric field is negligible, 
+obtaining the generalized DMS tensors for a particular molecule
+involves substantial amount of full QM calculations. %necessary to fit the DMS tensor elements.
+This is, however, a relatively expensive operation that has to be performed only once
+when the EFP\hyp{}like strategy is adopted. 
+%Note that the DMS parameters 
+%are environment\hyp{}independent and, due to this reason, could be used
+%as effective fragment parameters.
 %To get statistically relevant set of perturbed states, it is recommended to perform at least 100 
 %full QM calculations. 
-Since each of such preparatory calculations is an independent computational task, 
-its automatization and parallelization is straightforward. Moreover, the implementation
+Note also that each of such preparatory calculations is an independent computational task, 
+and hence, its automatization and parallelization is straightforward. Moreover, the implementation
 of the method does not depend on the level of theory because only the perturbed density matrices
-are needed. The computation of the gradient 
-and Hessian matrix is inexpensive and its efficiency is limited only by the speed of Hessian matrix inversion.
-Therefore, the method can be applied to all molecular fragments at any level of theory, provided
+are needed. The least\hyp{}squares fit, i.e., the computation of the gradient
+and Hessian matrix including its inversion, is inexpensive.
+This step can be implemented in an independent script or program taking as input the
+perturbed density matrices obtained by using any available quantum chemistry package. 
+%and its efficiency is limited only by the speed of Hessian matrix inversion.
+Therefore, the linear regression method to obtain the DMS tensors could be applied 
+to any molecular fragment at a chosen level of theory
+in a fairly black box manner, provided that
 the one\hyp{}particle density matrices are available. In the illustrations shown in this work,
-a set of 100 independent HF/6-311++G** calculations was sufficient to get statistically relevant set of perturbed 
-ground state density matrices. This task was concluded within approximately 5 minutes per each DMS fit
-on a single\hyp{}CPU computer.
+a set of 100 independent full QM calculations was sufficient to get statistically relevant set of perturbed 
+ground state density matrices. 
+%This task was concluded within approximately 5 minutes per each DMS fit
+%on a single\hyp{}CPU computer.
 

### 2018-10-08 papers/1_dmatpol/revision/reply-letter-1.tex
Revision of the paper submitted

@@
 and for supporting the work to be published in \emph{J. Chem. Phys.}
-I appreciate the positive and constructive feedbacks given, and agree with the reviewers’ 
-suggestions for revisions. The reviewers’ comments have been fully addressed as outlined below. 
-In addition, I have also introduced some minor adjustments in the notation.
+I appreciate the positive and constructive feedback given, and agree with the Reviewers’ 
+suggestions for revisions. The Reviewers’ comments have been fully addressed as outlined below. 
+%In addition, I have also introduced some minor adjustments in the notation.
 Now, I hope that the revised manuscript is suitable for publication in \emph{J. Chem. Phys.}
@@
 I definitely agree with the Reviewer that an additional illustration on some molecular
-dimer would much increase the quality of the present manuscript. Unfortunately, as the Reviewer also
-understands, this requires a separate article since other effects such as Pauli exchange-repulsion
+dimer would much increase the quality of the present manuscript. On the other hand, as the Reviewer also
+understands, this requires a more elaborate theoretical section since other effects such as Pauli exchange-repulsion
 need to be included and combined with polarization in order to compare to the full HF result. 
-Due to this reason, I decided to write a separate manuscript on that issue in the future.
+After an attempt to extend the scope of this manucript by the above illustration,
+I decided to write a separate manuscript on that issue in the future, because the additional
+material would distract the reader from the key focus of this article and introduce more aspects
+to be separately discussed (effects of exchange-repulsion and intermolecular charge transfer, 
+the choice of definition of induction energy
+within a dimer and so on). 
 \end{reply}
@@
 \begin{reply}
-
+To discuss these aspects in more detail, I have added one subsection entitled
+\emph{Computational cost and automatization of the generalized DMS model}
+within the Results and Discussion section (IV.C) on Page 20.
 \end{reply}
@@
 \begin{reply}
-I agree with the remark, thanks. I changed the sentence to incorporate this:
+Thanks. I changed the sentence on Page 14 as follows:
 \begin{quote}

### 2018-11-27 include/doxygen/oepdev_manual.h
Documentation revision

@@ -465,7 +465,7 @@ __Table 2.__ Target models vs benchmarks and competitor models.
 
 //--------------------------------------------------------------------------------------------
 
-/*! \page pprogramming Contributing to oep-dev
+/*! \page pprogramming Contributing to OEP-Dev
     \tableofcontents
 
 OepDev is a plugin to Psi4. Therefore it should follow the programming etiquette of Psi4. Also,

### 2018-11-27 oepdev/lib3d/dmtp.h

@@ -25,29 +25,99 @@ class DMTPole;
 /** \brief Multipole Convergence. 
  *
  * Handles the convergence of the distributed multipole expansions up to hexadecapole.
+ * Takes shared pointers to existing DMTPole objects and computes the generalized
+ * property:
+ *  - energy
+ *  - potential
+ *  from the DMTP sets. The results are stored in vector of length equal to the number
+ *  of DMTP's in a set decribed by DMTPole objects given. 
+ *
+ * \note
+ *  The number of DMTP's in each object has to be the same.
  */
 class MultipoleConvergence
 {
   public:
+    /** 
+     * Convergence level of the multipole expansion:
+     *
+     * @param R1 - qq term
+     * @param R2 - qd and sum of the above
+     * @param R3 - qQ, dd and sum of the above
+     * @param R4 - qO, dQ and sum of the above
+     * @param R5 - qH, dO, QQ and sum of the above
+     *
+     */
     enum ConvergenceLevel {R1, R2, R3, R4, R5};
+
+    /** 
+     * Property to be evaluated from interacting DMTP's:
+     *
+     * @param Energy    - generalized energy
+     * @param Potential - generalized potential
+     *
+     */
     enum Property {Energy, Potential};
+
+    /** \brief Construct from two shared DMTPole objects.
+     *
+     * @param dmtp1      - first DMTPole object
+     * @param dmtp2      - second DMTPole object
+     * @param max_clevel - maximul allowed convergence level
+     */
     MultipoleConvergence(std::shared_ptr<DMTPole> dmtp1, std::shared_ptr<DMTPole> dmtp2, 
                          ConvergenceLevel max_clevel = R5);
+
+    /// Destructor
     virtual ~MultipoleConvergence();
+
+    /** Compute the generalized property
+     * 
+     *  @param property - generalized Property
+     */
     void compute(Property property = Energy);
+
+    /** Grab the generalized property at specified level of convergence
+     * 
+     * @param clevel - ConvergenceLevel
+     * @return vector of results (each element corresponds to each DMTP pair in a set)
+     */
     std::shared_ptr<psi::Vector> level(ConvergenceLevel clevel = R5);
+
+
   protected:
+
+    /// Maximum allowed convergence level
     ConvergenceLevel max_clevel_;
+    /// First DMTP set
     std::shared_ptr<DMTPole> dmtp_1_;
[…]

### 2018-11-27 oepdev/lib3d/dmtp.h
Recentering of DMTP documented but needs revision

@@ -117,7 +117,6 @@ class MultipoleConvergence
  *  - superimposition
  *  - recentering the origins
  *  - computing the generalized property from another DMTP set
- * /TODO
  */
 class DMTPole : public std::enable_shared_from_this<DMTPole>
 {
@@ -212,10 +211,75 @@ class DMTPole : public std::enable_shared_from_this<DMTPole>
     /// Set the distributed hexadecapoles for the \f$ i \f$th distribution
     void set_hexadecapoles(psi::SharedMatrix M, int i) {hexadecapoles_[i] = std::make_shared<psi::Matrix>(M);}
 
-    /// Change origins of the distributed multipole moments of ith set
-    virtual void recenter(psi::SharedMatrix new_origins, int i);
-    /// Change origins of the distributed multipole moments of all sets
-    virtual void recenter(psi::SharedMatrix new_origins);
+    /** 
+     *  Change origins of the distributed multipole moments of all sets
+     *
+     *  @param new_origins - matrix with coordinates of the new origins \f$ \{ {\bf r}_{\rm new} \}\f$.
+     *  \note The number of origins has to be equal to the number of distributed centres.
+     *
+     *  Recentering of the multipoles affects the distributed dipoles and higher moments.
+     *  The moments are given as
+     *  \f{align*}{
+     *   q_{\rm new}  &= q_{\rm old} \\
+     *   {\boldsymbol{\upmu}}_{\rm new}   &= {\boldsymbol{\upmu}}_{\rm old} - q_{\rm old} {\boldsymbol \Delta}^{(1)} \\
+     *   {\boldsymbol{\Theta}}_{\rm new}  &= {\boldsymbol{\Theta}}_{\rm old} + q_{\rm old} {\boldsymbol \Delta}^{(2)}
+     *       - \sum_{\mathscr{P}_2} \mathscr{P}_2 \left[
+     *          \left(
+     *          {\boldsymbol{\upmu}}_{\rm old} + q_{\rm old} {\bf r}_{\rm old} 
+     *          \right)
+     *          \otimes {\boldsymbol \Delta}^{(1)} 
+     *          \right] \\
+     *   {\boldsymbol{\Omega}}_{\rm new}  &= {\boldsymbol{\Omega}}_{\rm old} - q_{\rm old} {\boldsymbol \Delta}^{(3)}
[…]

### 2018-11-27 oepdev/export/export_dmtp.cc
Recentering of one DMTP set moved to protected interface

@@ -27,7 +27,6 @@ void export_dmtp(py::module &m) {
     typedef std::vector<psi::SharedMatrix> (oepdev::DMTPole::*get_vecmat)(void) const;
     typedef void (oepdev::DMTPole::*set_matrix_int)(psi::SharedMatrix, int);
     typedef void (oepdev::DMTPole::*set_vecmat)(std::vector<psi::SharedMatrix>);
-    typedef void (oepdev::DMTPole::*recenter_set)(psi::SharedMatrix, int);
     typedef void (oepdev::DMTPole::*recenter_all)(psi::SharedMatrix);
 
     /* Function get_camm */
@@ -90,7 +89,6 @@ void export_dmtp(py::module &m) {
 
   	.def       ("n_sites"          ,                 &oepdev::DMTPole::n_sites                    , "Number of DMTP sites", py::return_value_policy::copy)
   	.def       ("n_dmtp"           ,                 &oepdev::DMTPole::n_dmtp                     , "Number of DMTP sets", py::return_value_policy::copy)
-  	.def       ("recenter"         , recenter_set(   &oepdev::DMTPole::recenter)                  , "Change origins of the distributed multipole moments of ith set")
   	.def       ("recenter"         , recenter_all(   &oepdev::DMTPole::recenter)                  , "Change origins of the distributed multipole moments of all sets")
 	.def       ("translate"        ,                 &oepdev::DMTPole::translate                  , "Translate the DMTP sets")
 	.def       ("rotate"           ,                 &oepdev::DMTPole::rotate                     , "Rotate the DMTP sets")

### 2018-11-27 oepdev/lib3d/dmtp.h

@@ -279,7 +279,7 @@ class DMTPole : public std::enable_shared_from_this<DMTPole>
      *  each distributed site of multipoles is independent of the others.
      *  TODO - Finish for octupoles and hexadecapoles! -> define the permutation operators!
      */
-    virtual void recenter(psi::SharedMatrix new_origins)
+    virtual void recenter(psi::SharedMatrix new_origins);
 
     /// Translate the DMTP sets
     void translate(psi::SharedVector transl);

### 2019-02-04 papers/1_dmatpol/revision/paper-revision-1.tex
Test of gamma scaling of XC density

@@
 Wroc{\l}aw University of Science and Technology, 
-Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}
+Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}
 
@@
 \begin{equation} \label{e:dmat-change.exact}
- \delta {\bf D} =  {\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
+ \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
                                    \left[ {\bf 1} + {\bf v}^\dagger \cdot {\bf v} \right]^{-1} \cdot
@@
                                  \left( 
-               D_{\alpha\gamma} C_{\beta i}^{(0)} + D_{\beta\gamma} C_{\alpha i}^{(0)}
+               D_{\alpha\gamma}^{(0)} C_{\beta i}^{(0)} + D_{\beta\gamma}^{(0)} C_{\alpha i}^{(0)}
                                  \right)
@@
 Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
-No.~665778. \InlineImage{eu-logo.eps} This project is funded by National Science Centre, Poland 
+No.~665778. This project is funded by National Science Centre, Poland 
 (registration No.~2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 

### 2019-02-04 gefp/gefp/density/partitioning.py

@@ -602,6 +602,7 @@ class DensityDecomposition:
 
         for i in range(self.aggregate.nfragments()):
             w_i = numpy.sqrt(self.matrix["n"])
+            print(" Test: Molecule %d: %16.3f" % (i+1, _f_test(w_i)))
             c_i = self.matrix["c"]
             scale = math.sqrt(self.xc_scale)
             w_i*= scale
@@ -977,6 +978,7 @@ class DensityDecomposition:
         ns = numpy.sqrt(n); N = c.shape[0]
         D  = numpy.zeros((N, N), numpy.float64)
         #scale = self.xc_scale ** (1.0 - ns)
+        print(" Test: %16.3f" % (_f_test(n)))
         scale = math.sqrt(self.xc_scale)
         ns *= scale
         for i in range(n.size):
@@ -1040,4 +1042,12 @@ class DensityDecomposition:
         v2 = value * c2
         line = "%18.8f   %18.8f   %18.8f" % (value, v1, v2)
         return line
+    def _f_test(self, n):
+        N = n.sum()
+        _f = 0.0
+        for i in range(len(n)):
+            for j in range(len(n)):
+                _f += math.sqrt(n[i] * n[j])
+        _f = N*N / _f
+        return _f
 

### 2019-05-17 papers/2_oep_technique/paper-r0.tex
Paper 2 draft was initialized

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

\title{Ab Initio Effective One-Electron Potentials 
for Repulsive and Charge-Transfer Interaction Energies at Hartree-Fock Level}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Marta Cho{\l}uj} 
\author{Joanna Bednarska}
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}
The concept of the effective one-electron potentials has been useful for many decades
in efficient description of electronic structure of chemical systems, especially extended
molecular aggregates such as interacting molecules in condensed phases. 
In this Work, \emph{ab initio} effective potentials are studied for a number of theories
of the intemolecular interaction energy.
\end{abstract}

\pacs{}

\maketitle

\tableofcontents

\section{\label{s:1}Introduction}
\section{\label{s:2}Theory}

Solution of the Schr{\"o}dinger equation provides the wavefunction
of the molecular aggregate $AB$,
%
\begin{equation}
 E^{AB} = \tBraKet{\Psi^{AB}}{\mathscr{H}}{\Psi^{AB}} \;,
\end{equation}
%
from which any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation}
 P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \;.
\end{equation}
%

\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Observable}

Within a supermolecular approch, the part of a property that arises due to the
intermolecular interaction between $A$ and $B$ is given by
%
\begin{equation} \label{e:delta-p}
	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} - 
	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} -
	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}\;.
\end{equation}
%
In a certain basis, expectation values of any operator $\mathscr{O}$ evaluated for
any system $X$ can be formally
expanded in a following series
%
\begin{equation} \label{e:exp-val-series}
	\tBraKet{\Psi^{X}}{\mathscr{O}}{\Psi^{X}} =
	\sum_{ij} C^{X}_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} C^{X}_{ijkl} \BraKet{ij}{kl}
\end{equation}
%
where $\hat{h}$ is the core Hamiltonian one\hyp{}electron operator
and the tensors
${\bf C}^{(2)}$ and ${\bf C}^{(4)}$ are characteristic for the property of interest.
Consider an approximation in which the basis functions $\phi_i$ 
are localized on either molecule. 
For this case, application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} yields the partitioning of the interaction\hyp{}induced property,
for which the one\hyp{}electron part reads
%
\begin{equation} \label{e:delta-p-part-1el}
	\sum_{ij} \Big\{ 
	C^{AA}_{ij}  \BraKet{A}{A} + %\tBraKet{i^A}{\hat{h}}{j^A} + 
	C^{BB}_{ij}  \BraKet{B}{B} + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	C^{AB}_{ij}  \BraKet{A}{B}   %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{equation}
%
and the two\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-2el}
	\sum_{ijkl} \Big\{  
	C^{AAAA}_{ijkl} \BraKet{AA}{AA} +    %\BraKet{i^Aj^A}{k^Al^A} +
	C^{BBBB}_{ijkl} \BraKet{BB}{BB} + \\ %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] + 
	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + \\
	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] +
	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\}
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\BraKet{A}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.

Therefore, the interaction\hyp{}induced property can be recast in a following way

\subsection{\label{ss:2.2}Effective One-Electron Potential}

The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
can be defined as
%
\begin{equation}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. The density can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation}
	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written by
%
\begin{equation}
	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
	\ket{\alpha} 
	\left[
	\int
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
	\rho^{\rm eff}({\bf r})
	d{\bf r} 
	\right]
	\bra{\beta}
\end{equation}
%
with the Coulomb operator defined by
%
\begin{equation}
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
	f({\bf r}) 
	\int
	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%
for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tbraket{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \sum_{\gamma\delta} \braket{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}
\end{equation}
%
In the above equation and throughout the work, the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\braket{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2
\end{equation}
%

\subsection{\label{ss:2.3}}

%n brief, the main principle
%is to rewrite the arbirtary function $f$ as\footnote{
%I use Coulomb notation for ERIs, i.e.,
%$\ERIc{\phi_i}{\phi_j}{\phi_k}{\phi_l} \equiv \iint d{\bf r}_1 d{\bf r}_2
%\phi_i^{*}({\bf r}_1)\phi_j({\bf r}_1) \frac{1}{\lvert {\bf r}_1 - {\bf r}_2 \rvert} \phi_k^{*}({\bf r}_2)\phi_l({\bf r}_2)
%$
%}
%
\begin{subequations}
 \begin{align}
 f\left[ 
   \ERIc{\phi_i^A}{\phi_j^A}{\phi_k^B}{\phi_l^B}
 \right] &= \OEIc{\phi_i^A}{v_{kl}^B}{\phi_j^A} \rightarrow \text{ multipole expansion or density fitting}\\
%
 f\left[ 
   \ERIc{\phi_i^A}{\phi_j^B}{\phi_k^B}{\phi_l^B}
 \right] &= \OEIc{\phi_i^A}{v_{kl}^B}{\phi_j^B} \rightarrow \text{ density fitting}
 \end{align}
\end{subequations}
%
where $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
or basis function.
$v_{kl}^B$ is the effective one-electron potential derived from the partial
effective densities $\rho_{kl}^B({\bf r}) = \phi_k^B({\bf r})\phi_l^B({\bf r})$.
The summations over $k$ and $l$ can be incorporated into the generalized one-electron potential
$v_{\text{eff}}^B$
to produce
%
\begin{equation}
 \sum_{ij}\sum_{kl\in B} f\left[ 
   \ERIc{\phi_i}{\phi_j}{\phi_k^B}{\phi_l^B}
 \right] = \sum_{ij} \OEIc{\phi_i}{v_{\text{eff}}^B}{\phi_j} 
\end{equation}
%
Thus, the total computational effort is extremely reduced from the fourth-fold
sum involving evaluation of ERIs to the two-fold sums of cheaper one-electron integrals.
It is also possible to generalize the above expression even further by
summing over all possible functions $f_t$
%
\begin{equation} \label{e:ft-reduction}
 \sum_t \sum_{ij}\sum_{kl\in B} f_t\left[ 
   \ERIc{\phi_i}{\phi_j}{\phi_k^B}{\phi_l^B}
 \right] = \sum_{ij} \OEIc{\phi_i}{v_{\text{eff}}^B}{\phi_j} 
\end{equation}
%
The above design has the advantage that the effective potentials are fully first-principles
and no extensive case-dependent fitting procedures are necessary.
Only the effective potentials of \emph{independent} fragment $B$ need to be determined once and for all
and stored in a file. Note also, that, in principle, there is no approximation 
made here at that moment.

\section{\label{s:3}Calculation Details}
\section{\label{s:4}Results and Discussion}
\section{\label{s:5}Summary and a few concluding remarks}

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

% -----------------------
\bibliography{references}
% -----------------------

\end{document}

### 2019-05-20 papers/2_oep_technique/paper-r0.tex
Mistakes in documentation corrected. Paper II adds

@@
 \section{\label{s:1}Introduction}
-\section{\label{s:2}Theory}
+\section{\label{s:2}Effective One-Electron Potentials: Configuration Interaction}
 
-Solution of the Schr{\"o}dinger equation provides the wavefunction
-of the molecular aggregate $AB$,
+The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
+electronic state
+of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
+%
+\begin{equation} \label{e:schrodinger}
+ E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
+\end{equation}
+%
+for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
+with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
+and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
+From the above solution, any property $P$, understood as an expectation value of the
+associated operator $\mathscr{P}$, can be computed according to
+%
+\begin{equation} \label{e:prop}
+ P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
+\end{equation}
+%
+The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
+the configuration interaction (CI) expansion around the reference 
+(approximate) wavefunction, $\Psi^{(0)}$,
+typically the solution to the Hartree\hyp{}Fock (HF) problem,
+%
+\begin{equation} \label{e:ci}
+ \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
+	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
+\end{equation}
+%
+In the above equation,
+$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
+with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
+one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
+virtual (unoccupied) molecular orbital, obtained from the HF solution.
+By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
+and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
+operator matrix element between
+the CI configurations, it can be shown that the following holds
+%
+\begin{equation} \label{e:exp-val-series}
+	P =
+	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
+	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
+\end{equation}
+%
+where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $,
+$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
 %
 \begin{equation}
- E^{AB} = \tBraKet{\Psi^{AB}}{\mathscr{H}}{\Psi^{AB}} \;,
+	\BraKet{ij}{kl} \equiv
+	\iint 
+	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
+	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
+	d{\bf r}_1 d{\bf r}_2
 \end{equation}
 %
-from which any property $P$, understood as an expectation value of the
-associated operator $\mathscr{P}$, can be computed according to
-%
-\begin{equation}
- P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \;.
-\end{equation}
-%
+whereas the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
+and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
+are well defined and characteristic for the property of interest.
+%In this work we are interested in cases when
 
-\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Observable}
+%$\Psi^{AB}$. In this Section we consider theories to calculate the effect of the interaction
+%of molecules $A$ and $B$ on the property $P$ knowing only the unperturbed wavefunctions,
+%$\Psi^{A}$ and $\Psi^{B}$.
+
+\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property: Localized Basis Set Approximation}
 
@@
 %
-In a certain basis, expectation values of any operator $\mathscr{O}$ evaluated for
-any system $X$ can be formally
-expanded in a following series
-%
-\begin{equation} \label{e:exp-val-series}
-	\tBraKet{\Psi^{X}}{\mathscr{O}}{\Psi^{X}} =
-	\sum_{ij} C^{X}_{ij} \tBraKet{i}{\hat{h}}{j}
-	+ \sum_{ijkl} C^{X}_{ijkl} \BraKet{ij}{kl}
-\end{equation}
-%
-where $\hat{h}$ is the core Hamiltonian one\hyp{}electron operator
-and the tensors
-${\bf C}^{(2)}$ and ${\bf C}^{(4)}$ are characteristic for the property of interest.
+where $\Psi^{AB}$ and $\Psi^{A(B)}$ are the wavefunctions of the interacting molecular aggregate
+and the unperturbed (non\hyp{}interacting) molecules, respectively.
+
 Consider an approximation in which the basis functions $\phi_i$ 

### 2019-05-20 oepdev/libsolver/solver.h

@@ -606,7 +606,7 @@ class RepulsionEnergySolver : public OEPDevSolver
  *            \left\{
  *               V^A_{nj} + 2\sum_{k\in A}^{\rm Occ_A} (1-\delta_{ik})(nj \vert kk)
  *            \right\}
- *            -(nj \vert ij)
+ *            +(nj \vert ij)
  *            \right]
  *            + \sum_{k\in A}^{\rm Occ_A}\sum_{j\in B}^{\rm Occ_B}
  *            S_{kj}(1-\delta_{ik}) (ik \vert nj)
@@ -623,10 +623,10 @@ class RepulsionEnergySolver : public OEPDevSolver
  * \f]
  * where 
  * \f[
- *   V^2_{in} = \frac{V^B_{in} - \sum_{m\in A}^{\rm All_A} V_{im}S_{mn}^B}
+ *   V^2_{in} = \frac{V^{EF,B}_{in} - \sum_{m\in A}^{\rm All_A} V^{EF,B}_{im}S_{mn}^B}
  *                   {1-\sum_{m\in A}^{\rm All_A} S_{mn}^2}
  *          \left\{
- *             V^B_{in} - \sum_{m\in A}^{\rm All_A} V_{im}^BS_{mn}
+ *             V^{EF,B}_{in} - \sum_{m\in A}^{\rm All_A} V_{im}^{EF,B}S_{mn}
  *            + \sum_{j\in B}^{\rm Occ_B} S_{ij} \left( T_{nj} - \sum_{m\in A}^{\rm All_A} S_{nm} T_{mj} \right)
  *          \right\}
  * \f]

### 2019-05-20 gefp/gefp/density/opdm.py
Control over occupancy deviation added by n_eps option. Documentation added.

@@ -13,16 +13,68 @@ from abc import ABC, abstractmethod
 __all__ = ["Density"]
 
 class Density:
-    """\
+    """
+ --------------------------------------------------------------------------------------------------------------------------
+                                                Electron Density
+
  Handles the Electron Density Distribution.
 
- Class methods:
-  Density.natural_orbitals
-  Density.generalized_density
-  Density.orthogonalize_OPDM
-  Density.orthogonalizer
-  Density.deorthogonalizer
+ --------------------------------------------------------------------------------------------------------------------------
 
+ Usage as container class: 
+ 
+  1) Initialize container object:
+     
+     density = Density(D = None, jk = None)
+
+     where:
+       o D  - the density matrix in AO or MO basis
+       o jk - the psi4::JK object for AO basis JK calculations
+
+  2) Grab the density matrix
+
+     D = density.matrix()
+
+  3) Computations in AO basis:
+
+     o compute 1-electron energy (does not require JK object to be set)
+
+       e_1 = density.compute_1el_energy(D, V1)
+
+     The below require jk to be set:
+
+     o compute 2-electron energy (J-type expression)
+
+       e_2j = density.compute_2el_energy(D_left, D_right, type='j')
+
+     o compute 2-electron energy (K-type expression)
+
+       e_2k = density.compute_2el_energy(D_left, D_right, type='k')
+
+     o compute J matrix (or K matrix if type=='k'):
+
+       J = density.generalized_JK(D, type='j')
+
+ --------------------------------------------------------------------------------------------------------------------------
+
+ Usage as method class. 
+
+ Using 'Density' as a class of methods do not require object initialization.
+ The list of class methods is given below:
+
+   o Density.natural_orbitals     - compute natural orbitals
+   o Density.generalized_density  - compute generalized OPDM
+   o Density.orthogonalize_OPDM   - compute orthogonalized OPDM
+   o Density.orthogonalizer       - compute orthogonalizer matrix
+   o Density.deorthogonalizer     - compute deorthogonalizer rmatrix
+
+ Usage: 
+   result = Density.'class method name'
+
+ See respective documentation for each of them for further details.
+
+ --------------------------------------------------------------------------------------------------------------------------
+                                                                                  Last Revision: Gundelfingen, 20 May 2019
 """
[…]

### 2019-05-20 papers/2_oep_technique/paper-r0.tex
Elaborated on the basics of OEP technique

@@
 \section{\label{s:1}Introduction}
+
+\section{\label{s:2}Effective One-Electron Potentials: General Form}
+
+The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
+produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
+is 
+%
+\begin{equation}
+	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
+\end{equation}
+%
+where ${\bf r}$ is a spatial coordinate. Formally, the effective density can be expanded in terms of an effective
+one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
+%
+\begin{equation}
+	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
+	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
+\end{equation}
+%
+Based on that, the operator form of the effective potential 
+can be written as
+%
+\begin{equation}
+	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
+	\ket{\alpha} 
+	\left[
+	\int
+	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
+	\rho^{\rm eff}({\bf r})
+	d{\bf r} 
+	\right]
+	\bra{\beta}  
+\end{equation}
+%
+with the Coulomb operator defined by
+%
+\begin{equation}
+	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
+	f({\bf r}) 
+	\int
+	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+\end{equation}
+%
+for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
+is therefore given by
+%
+\begin{equation}
+	\tbraket{\alpha}{ \hat{v}^{\rm eff} }{\beta}
+	= \sum_{\gamma\delta} \braket{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
+\end{equation}
+%
+In the above equation and throughout the work, the electron repulsion integral (ERI)
+is defined according to
+%
+\begin{equation}
+	\braket{\alpha\beta}{\gamma\delta} \equiv
+	\iint 
+	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
+	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
+	d{\bf r}_1 d{\bf r}_2  \;.
+\end{equation}
+%
+
 \section{\label{s:2}Effective One-Electron Potentials: Configuration Interaction}
@@
 	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
-	d{\bf r}_1 d{\bf r}_2
+	d{\bf r}_1 d{\bf r}_2  \;,
 \end{equation}
@@
 
-\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property: Localized Basis Set Approximation}
+\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property}
 
-Within a supermolecular approch, the part of a property that arises due to the
-intermolecular interaction between $A$ and $B$ is given by
+Consider a molecular aggregate consisting of two interacting 
+molecules $A$ and $B$. The part of property that arises due to the
+intermolecular interaction between $A$ and $B$ can be calculated
+based on the supermolecular approach,
 %
-\begin{equation} \label{e:delta-p}
-	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} - 
-	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} -
-	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}\;.
-\end{equation}
+\begin{multline} \label{e:delta-p}
+	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \\- 
+	\left(
+	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} +
+	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}
+	\right)
+	\;,
+\end{multline}
 %
@@
 and the unperturbed (non\hyp{}interacting) molecules, respectively.
+The notation $\Delta^{AB}$ emphasizes on the two\hyp{}body character of the interaction\hyp{}induced
+property. In this work we focus on re\hyp{}expressing $\Delta^{AB}P$ in terms of separate
+fragments associated with either molecule. In general case when the basis set
+is complete, such an operation is not possible. However,
+assuming an approximation
+in which the basis functions $\phi_i$ 
+are localized on either molecule it allows one to partition the basis set space
+and the resulting 2\hyp{} and 4\hyp{}index tensor elements
+into subsets `belonging' either to molecule $A$ or to molecule $B$.
+Such an approximation is essentially a foundation of any fragment\hyp{}based strategy
+and, in particular, designing \emph{ab initio} force fields such as EFP2
+of SolEFP.
 
-Consider an approximation in which the basis functions $\phi_i$ 
-are localized on either molecule. 
-For this case, application of Eq.~\eqref{e:exp-val-series}
-to Eq.~\eqref{e:delta-p} yields the partitioning of the interaction\hyp{}induced property,
+Application of Eq.~\eqref{e:exp-val-series}
+to Eq.~\eqref{e:delta-p} and adopting the localized basis set approximation
+yields the partitioning of the interaction\hyp{}induced property,
+%
+\begin{equation}
+ \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
+\end{equation}
+%
 for which the one\hyp{}electron part reads
 %
-\begin{equation} \label{e:delta-p-part-1el}
+\begin{multline} \label{e:delta-p-part-1el}
+ \Delta^{AB} P(1) \approx
 	\sum_{ij} \Big\{ 
 	C^{AA}_{ij}  \BraKet{A}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
-	C^{BB}_{ij}  \BraKet{B}{B} + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
+	C^{BB}_{ij}  \BraKet{B}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
 2	C^{AB}_{ij}  \BraKet{A}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
 	\Big\}
-\end{equation}
+\end{multline}
 %
-and the two\hyp{}electron part reads
+and the two\hyp{}electron part is given by
 %
 \begin{multline} \label{e:delta-p-part-2el}
+ \Delta^{AB} P(2) \approx
 	\sum_{ijkl} \Big\{ 
-	C^{AAAA}_{ijkl} \BraKet{AA}{AA} +    %\BraKet{i^Aj^A}{k^Al^A} +
-	C^{BBBB}_{ijkl} \BraKet{BB}{BB} + \\ %\BraKet{i^Bj^B}{k^Bl^B} +
-	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] + 
-	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + \\
-	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] +
+	C^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
+	C^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
+	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
+	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
+	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
 	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
-	\Big\}
+	\Big\} \;.
 \end{multline}
@@
 given their partitioning scheme between molecules $A$ and $B$.
+All the summation components from Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el} 
+can be grouped into the three
+sub\hyp{}categories: (i) Coulomb\hyp{}like terms
+with integrals of the type $\BraKet{AA}{BB}$, (ii) overlap\hyp{}like terms with $\BraKet{AA}{AB}$
+and $\BraKet{AB}{BB}$ and (iii) exchange\hyp{}like terms with $\BraKet{AB}{AB}$.
 
-Therefore, the interaction\hyp{}induced property can be recast in a following way
+%Therefore, the interaction\hyp{}induced property can be recast in a following way
 
-\subsection{\label{ss:2.2}Effective One-Electron Potential}
+\subsection{\label{ss:2.3}Defining Effective Potentials}
 
-The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
-produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
-can be defined as
+Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
+ERI's. 
 %
  \begin{equation}
-	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+ \mathcal{F}
+ \left[ 
+   \ERIc{i}{j}{k^B}{l^B}
+	 \right] = \OEIc{i}{v_{kl}^B}{j}  \;,
  \end{equation}
 %
-where ${\bf r}$ is a spatial coordinate. The density can be expanded in terms of an effective
-one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
-%
-\begin{equation}
-	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
-	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})
-\end{equation}
-%
-Based on that, the operator form of the effective potential 
-can be written by
-%
-\begin{equation}
-	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
-	\ket{\alpha} 
-	\left[
-	\int
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
-	\rho^{\rm eff}({\bf r})
-	d{\bf r} 
-	\right]
-	\bra{\beta}
-\end{equation}
-%
-with the Coulomb operator defined by
-%
-\begin{equation}
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
-	f({\bf r}) 
-	\int
-	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-\end{equation}
-%
-for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
-is therefore given by
-%
-\begin{equation}
-	\tbraket{\alpha}{ \hat{v}^{\rm eff} }{\beta}
-	= \sum_{\gamma\delta} \braket{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}
-\end{equation}
-%
-In the above equation and throughout the work, the electron repulsion integral (ERI)
-is defined according to
-%
-\begin{equation}
-	\braket{\alpha\beta}{\gamma\delta} \equiv
-	\iint 
-	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
-	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
-	d{\bf r}_1 d{\bf r}_2
-\end{equation}
-%
-
-\subsection{\label{ss:2.3}}
-
-%n brief, the main principle
-%is to rewrite the arbirtary function $f$ as\footnote{
-%I use Coulomb notation for ERIs, i.e.,
-%$\ERIc{\phi_i}{\phi_j}{\phi_k}{\phi_l} \equiv \iint d{\bf r}_1 d{\bf r}_2
-%\phi_i^{*}({\bf r}_1)\phi_j({\bf r}_1) \frac{1}{\lvert {\bf r}_1 - {\bf r}_2 \rvert} \phi_k^{*}({\bf r}_2)\phi_l({\bf r}_2)
-%$
-%}
-%
-\begin{subequations}
- \begin{align}
- f\left[ 
-   \ERIc{\phi_i^A}{\phi_j^A}{\phi_k^B}{\phi_l^B}
- \right] &= \OEIc{\phi_i^A}{v_{kl}^B}{\phi_j^A} \rightarrow \text{ multipole expansion or density fitting}\\
-%
- f\left[ 
-   \ERIc{\phi_i^A}{\phi_j^B}{\phi_k^B}{\phi_l^B}
- \right] &= \OEIc{\phi_i^A}{v_{kl}^B}{\phi_j^B} \rightarrow \text{ density fitting}
- \end{align}
-\end{subequations}
-%
-where $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
-or basis function.
+where 
+% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
+%or basis function.
 $v_{kl}^B$ is the effective one-electron potential derived from the partial
@@
 \begin{equation}
- \sum_{ij}\sum_{kl\in B} f\left[ 
-   \ERIc{\phi_i}{\phi_j}{\phi_k^B}{\phi_l^B}
- \right] = \sum_{ij} \OEIc{\phi_i}{v_{\text{eff}}^B}{\phi_j} 
+	\sum_{ij}\sum_{kl\in B} \mathcal{F}\left[ 
+   \ERIc{i}{j}{k^B}{l^B}
+ \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j}  \;.
 \end{equation}
 %
-Thus, the total computational effort is extremely reduced from the fourth-fold
-sum involving evaluation of ERIs to the two-fold sums of cheaper one-electron integrals.
+Thus, the total computational effort is, in principle, reduced from the fourth-fold
+sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
 It is also possible to generalize the above expression even further by
-summing over all possible functions $f_t$
+summing over all possible functionals ${\mathcal{F}}_t$
 %
 \begin{equation} \label{e:ft-reduction}
- \sum_t \sum_{ij}\sum_{kl\in B} f_t\left[ 
-   \ERIc{\phi_i}{\phi_j}{\phi_k^B}{\phi_l^B}
- \right] = \sum_{ij} \OEIc{\phi_i}{v_{\text{eff}}^B}{\phi_j} 
+	\sum_t \sum_{ij}\sum_{kl\in B} {\mathcal{F}}_t\left[ 
+   \ERIc{i}{j}{k^B}{l^B}
+ \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j} \;.
 \end{equation}
@@
 
+\subsection{\label{ss:2.4}Interaction-Induced Property: Effective Potentials}
+
+The above technique can be now applied to the Coulomb\hyp{}like and overlap\hyp{}like 
+terms from Eq.~\eqref{e:delta-p-part-2el}. For example, the Coulomb\hyp{}like contributions
+can be rewritten without making any approximation as
+%
+\begin{equation}
+ \sum_{ijkl} 
+	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
+ \equiv
+	\sum_{kl\in B} 
+	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
+\end{equation}
+%
+with
+%
+\begin{equation}
+ \hat{v}^{{\rm eff},A} 
+	\equiv \sum_{ij\in A} {s} \left[ C^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
+\end{equation}
+%
+Similarly for the overlap\hyp{}like term we have
+%
+\begin{equation}
+ \sum_{ijkl} 
+	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
+ \equiv
+	\sum_{k\in A} \sum_{l\in B} 
+	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
+\end{equation}
+%
+and the associated effective potential reads
+%
+\begin{equation}
+ \hat{v}^{{\rm eff},A} 
+	\equiv \sum_{ij\in A} {s} \left[ C^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
+\end{equation}
+%
+Note also that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
+Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
+%
+\begin{multline}
+ \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
+	+ \sum_{i\in A} \sum_{j\in B} 
+	\left\{
+	  \tBraKet{i}{\hat{v}^{{\rm eff},A}_1 }{j}
+	+ \tBraKet{i}{\hat{v}^{{\rm eff},B}_1 }{j}
+	\right\} \\ +
+	\frac{1}{2}
+	\sum_{i\in A} \sum_{k\in A}
+	  \tBraKet{i}{\hat{v}^{{\rm eff},B}_2 }{k}
+	 +
+	 \frac{1}{2}
+        \sum_{l\in B} \sum_{l\in B}
+          \tBraKet{l}{\hat{v}^{{\rm eff},A}_2 }{l}
+\end{multline}
+%
+where the exchange\hyp{}like component is
+%
+\begin{equation}
+	\Delta_{\rm ex}^{AB} P = 
+	\sum_{ik\in A} \sum_{jl\in B}
+	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] \;.
+\end{equation}
+%
+whereas
+the overlap\hyp{}like and the Coulomb\hyp{}like 
+OEP's are defined as follows
+%
+\begin{align}
+	\hat{v}^{{\rm eff},A}_1 &\equiv C_{ij}^{AB} 
+	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
+	%
+	\hat{v}^{{\rm eff},A}_2 &\equiv 
+	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
+\end{align}
+%
+and accordingly for the twin terms (molecule $B$).
+
+\section{\label{s.334}Practical Calculations with OEP's}
+
+There are two general cases of basis function partitioning scheme
+in which it is possible to define an effective one\hyp{}electron
+potential.
+The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
+The second type is an overlap\hyp{}like contribution, can be approximated with the density fitting. Mainly,
+%
+\begin{subequations}
+ \begin{align}
+ \mathcal{F}
+ \left[ 
+   \ERIc{i^A}{j^A}{k^B}{l^B}
+	 \right] = \OEIc{i^A}{v_{kl}^B}{j^A} \\ \nonumber \rightarrow \text{ multipole expansion or density fitting}\\ 
+%
+ \mathcal{F}
+ \left[ 
+   \ERIc{i^A}{j^B}{k^B}{l^B}
+	 \right] = \OEIc{i^A}{v_{kl}^B}{j^B} \\ \rightarrow \text{ density fitting} \nonumber
+ \end{align}
+\end{subequations}
+%
+
+\subsection{\label{s.333}Distributed Multipole Expansion of OEP's}
+
+\subsection{\label{s.334}Density Fitting of OEP's}
+
+To get the \emph{ab initio} representation of a overlap\hyp{}like OEP, 
+one can use a procedure similar to
+the typical density fitting or resolution of identity, both of which are nowadays widely used 
+to compute electron-repulsion integrals (ERI's) more efficiently.
+
+\subsubsection{Density Fitting in Nearly-Complete Space}
+
+An arbitrary one-electron potential of molecule $A$ acting on any state vector 
+associated with molecule $A$ can be expanded in an *auxiliary space* centered 
+on $A$ as
+\begin{equation}
+   v\vert i) = \sum_{\xi\eta} v\vert \xi) [{\bf S}^{-1}]_{\xi\eta} ( \eta \vert i)
+\end{equation}
+under the necessary assumption that the auxiliary basis set is nearly complete.
+In a special case when the basis set is orthogonal (e.g., molecular orbitals)
+the above relation simplifies to
+\begin{equation}
+   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
+\end{equation}
+It can be easily shown that the above general and exact expansion can be obtained by 
+performing a density fitting in the complete space. We expand the LHS of the first equation 
+on this page in a series of the auxiliary basis functions scaled by the undetermined expansion coefficients:
+\begin{equation}
+  v\vert i) = \sum_{\xi} {G_{i\xi}} \vert \xi)
+\end{equation}
+which we shall refer here as to the matrix form of the OEP operator.
+By constructing the least-squares objective function 
+\begin{equation}
+ Z[\{G^{(i)}_\xi\}] = \int d{\bf r}_1
+                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi G^{(i)}_\xi \varphi_\xi({\bf r}_1) \right]^2
+\end{equation}
+and requiring that
+\begin{equation}
+ \frac{\partial Z[\{G^{(i)}_\xi\}]}{\partial G^{(i)}_\mu} = 0 \text{ for all $\mu$}
+\end{equation}
+we find the coefficients $ G^{(i)}_\xi $ to be
+\begin{equation}
+ {\bf G}^{(i)} = {\bf v}^{(i)} \cdot {\bf S}^{-1}
+\end{equation}
+where
+\begin{align}
+ v^{(i)}_\eta &= (\eta \vert vi) \\
+ S_{\eta\xi}  &= (\eta \vert \xi)
+\end{align}
+or explitictly
+\begin{equation}
+  {G_{i\xi}} = \sum_\eta [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
+\end{equation}
+identical to what we obtained from application of the resolution of identity 
+in space spanned by non-orthogonal complete set of basis vectors.
+
+Since matrix elements of an OEP operator in auxiliary space can be computed 
+in the same way as the matrix elements with any other basis function, 
+one can formally write the following identity
+\begin{equation}
+ (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
+\end{equation}
+where $ X $ is an arbitrary orbital.
+When the other orbital
+does not belong to molecule $A$ but to the (changing) environment, it is 
+straightforward to compute the resulting matrix element, which is simply given as  
+\begin{equation}
+   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
+\end{equation}
+where $j$ denotes the other (environmental) basis function.
+
+\subsubsection{Density Fitting in Incomplete Space}
+
+Density fitting scheme from previous section has practical disadvantage of a nearly-complete basis set
+being usually very large (spanned by large amount of basis set vectors). Any non-complete basis set
+won't work in the previous example. Since most of basis sets used in quantum chemistry do not form a nealy complete
+set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
+matrix elements of the OEP operator in a incomplete auxiliary space. This can be achieved by minimizing 
+the following objective function
+\begin{multline}
+	Z[\{G^{(i)}_\xi\}] = \iint d{\bf r}_1 d{\bf r}_2 \Bigg\{
+                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi G^{(i)}_\xi \varphi_\xi({\bf r}_1) \right]
+                  \\ \times \frac{1}{\vert {\bf r}_1 - {\bf r}_2 \vert}  
+                     \left[    v({\bf r}_2) \phi_i({\bf r}_2) - \sum_\xi G^{(i)}_\eta\varphi_\eta({\bf r}_1) \right]
+		    \Bigg\}
+\end{multline}
+Thus requesting that 
+\begin{equation}
+ \frac{\partial Z[\{G^{(i)}_\xi\}]}{\partial G^{(i)}_\mu} = 0 \text{ for all $\mu$}
+\end{equation}
+we find the coefficients $ G^{(i)}_\xi $ to be
+\begin{equation}
+ {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
+\end{equation}
+where 
+\begin{align}
+ b^{(i)}_\eta &= (\eta \vert\vert vi) \\
+ A_{\eta\xi}  &= (\eta \vert\vert \xi)
+\end{align}
+The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
+and $ {\bf r}_2 $. Thus, it is clear that in order to use this generalized density fitting scheme
+one must to compute two\hyp{}centre electron repulsion integrals
+as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
+The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.
+
 \section{\label{s:3}Calculation Details}
 \section{\label{s:4}Results and Discussion}
+\subsection{\label{ss.4.1}Pauli-Repulsion Interaction Energy}
+\subsection{\label{ss.4.2}Charge-Transfer Interaction Energy}
 \section{\label{s:5}Summary and a few concluding remarks}
@@
 
+%
+\appendix
+
+\section{\label{a:mcmurchie-davidson} McMurchie-Davidson Method for Asymmetric Electron Repulsion Integrals}
+
+Evaluation of integrals $(ijk\vert l)$,
+that are necessary for the generalized density fitting of OEP's,
+can be easily carried out by re\hyp{}expressing the
+unnormalized product of three primitive Gaussian\hyp{}type (GTO) functions,
+%
+\begin{equation} \label{e:1}
+[ijk] \equiv \phi_i({\bf r}) \phi_j({\bf r}) \phi_k({\bf r}) \;,
+\end{equation}
+%
+in terms of Hermite functions. It can be shown that
+%
+\begin{multline} \label{e:2}
+   [ijk] = E_{ijk} \sum_{N=0}^{n_1+n_2+n_3} \sum_{L=0}^{l_1+l_2+l+3} \sum_{M=0}^{m_1+m_2+m_3} 
+	\\
+          d_N^{n_1n_2n_3} d_L^{l_1l_2l_3} d_M^{m_1m_2m_3}
+          \Lambda_N(x_R)\Lambda_L(y_R)\Lambda_M(z_R)e^{-\alpha_Rr_R^2}
+\end{multline}
+%
+in which the McMurchie-Davidson $d3$ coefficients are given by the following recurrence relationships
+%
+ \begin{align} \label{e:3}
+  d_N^{n_1+1,n_2,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+                             + \vert {\bf R} - {\bf A}\vert_x d_N^{n_1n_2n_3} \nonumber \\ 
+			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
+  d_N^{n_1,n_2+1,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+                             + \vert {\bf R} - {\bf B}\vert_x d_N^{n_1n_2n_3} \nonumber \\
+			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
+  d_N^{n_1,n_2,n_3+1} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+                             + \vert {\bf R} - {\bf C}\vert_x d_N^{n_1n_2n_3} \nonumber \\
+			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} 
+\end{align}
+%
+with $d_0^{000} = 1$.
+In Eq.~\eqref{e:1},
+$\phi_i({\bf r})$ is given by
+%
+\begin{equation}
+\phi_i({\bf r}) \equiv x_A^{n_1} y_A^{l_1} z_A^{m_1} e^{-\alpha_1r_A^2}
+\end{equation}
+%
+where 
+${\bf r}_A \equiv {\bf r} - {\bf A}$,
+${\bf A}$ is the centre of the GTO, $\alpha_1$ its exponent, whereas $n_1,l_1,m_1$
+the Cartesian angular momenta, with the total angular momentum $\theta_1 = n_1+l_1+m_1$.
+
+It can be easily shown that the multiplicative constant $E_{ijk}$ is given by
+%
+\begin{multline}
+  E_{ijk}(\alpha_1,\alpha_2,\alpha_3)  = \exp{\left[-\frac{\alpha_1\alpha_2}
+                                        {\alpha_1+\alpha_2}\vert {\bf A}-{\bf B}\vert^2\right]} \\ \times
+                                         \exp{\left[-\frac{(\alpha_1+\alpha_2)\alpha_3}
+                                        {\alpha_1+\alpha_2+\alpha_3}
+                                         \vert {\bf P}-{\bf C}\vert^2\right]} 
+\end{multline}
+%
+
 % -----------------------

### 2019-05-21 papers/2_oep_technique/paper-r0.tex
Paper II CT-OEP with added discussion on OEP basicts from CI expansion

@@
 
-\title{Ab Initio Effective One-Electron Potentials 
-for Repulsive and Charge-Transfer Interaction Energies at Hartree-Fock Level}
+\title{Ab Initio Effective One-Electron Potentials Revisited:
+General Theory and Example Applications for Charge-Transfer Energy Calculations}
 
@@
 molecular aggregates such as interacting molecules in condensed phases. 
-In this Work, \emph{ab initio} effective potentials are studied for a number of theories
-of the intemolecular interaction energy.
+%In this Work, \emph{ab initio} effective potentials are studied for a number of theories
+%of the intemolecular interaction energy.
 \end{abstract}
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-eff}
 	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
@@
 %
-\begin{equation}
+\begin{equation} \label{e:d-spectral}
 	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:oep-operator}
 	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
-	\ket{\alpha} 
+	\Ket{\alpha} 
 	\left[
@@
 	\right]
-	\bra{\beta}  
+	\Bra{\beta}  
 \end{equation}
@@
 \begin{equation}
-	\tbraket{\alpha}{ \hat{v}^{\rm eff} }{\beta}
-	= \sum_{\gamma\delta} \braket{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
+	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
+	= \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
 \end{equation}
 %
-In the above equation and throughout the work, the electron repulsion integral (ERI)
+In the above equation and throughout the work, 
+%
+\begin{equation}
+\tBraKet{\alpha}{\mathcal{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathcal{O}(1) \phi_\beta({\bf r}) d{\bf r} 
+\end{equation}
+%
+for any one\hyp{}electron operator $\mathcal{O}(1)$ 
+and the electron repulsion integral (ERI)
 is defined according to
@@
 \begin{equation}
-	\braket{\alpha\beta}{\gamma\delta} \equiv
+	\BraKet{\alpha\beta}{\gamma\delta} \equiv
 	\iint 
@@
 %
-where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $,
-$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
+%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
+%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
+%%
+%\begin{equation}
+%	\BraKet{ij}{kl} \equiv
+%	\iint 
+%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
+%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
+%	d{\bf r}_1 d{\bf r}_2  \;,
+%\end{equation}
+%%
+where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
+and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
+are well defined and characteristic for the property of interest.
+In fact, they are certain functions of the CI coefficients, i.e.,
 %
 \begin{equation}
-	\BraKet{ij}{kl} \equiv
-	\iint 
-	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
-	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
-	d{\bf r}_1 d{\bf r}_2  \;,
+{\bf P} = {\bf P}
+\left(
+ \left\{
+  C_{r}^{a}, C_{rs}^{ab}, \ldots 
+ \right\} 
+\right)
 \end{equation}
 %
-whereas the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
-and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
-are well defined and characteristic for the property of interest.
+and, in principle, are known as long as the solution to the CI problem 
+or its approximation is available.
+
 %In this work we are interested in cases when
@@
 %
-\begin{equation}
+\begin{equation} \label{e:delta-p-part}
  \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
@@
 	\sum_{ij} \Big\{ 
-	C^{AA}_{ij}  \BraKet{A}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
-	C^{BB}_{ij}  \BraKet{B}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
-2	C^{AB}_{ij}  \BraKet{A}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
+	P^{AA}_{ij}  \tBraKet{A}{\hat{h}}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
+	P^{BB}_{ij}  \tBraKet{B}{\hat{h}}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
+2	P^{AB}_{ij}  \tBraKet{A}{\hat{h}}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
 	\Big\}
@@
 	\sum_{ijkl} \Big\{ 
-	C^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
-	C^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
-	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
-	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
-	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
-	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
+	P^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
+	P^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
+	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
+	{s} \left[ P^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
+	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
+	{s} \left[ P^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
 	\Big\} \;.
@@
 in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
-the shorthand notation $\BraKet{A}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
+the shorthand notation $\tBraKet{A}{\hat{h}}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
 $\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
@@
 Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
-ERI's. 
+ERI's. In this work, OEP's are defined by the following transformation
 %
@@
  \left[ 
-   \ERIc{i}{j}{k^B}{l^B}
-	 \right] = \OEIc{i}{v_{kl}^B}{j}  \;,
+   \BraKet{ij}{k^Al^A}
+	 \right] = \tBraKet{i}{\hat{v}_{kl}^A}{j}  \;,
  \end{equation}
@@
 %or basis function.
-$v_{kl}^B$ is the effective one-electron potential derived from the partial
-effective densities $\rho_{kl}^B({\bf r}) = \phi_k^B({\bf r})\phi_l^B({\bf r})$.
-The summations over $k$ and $l$ can be incorporated into the generalized one-electron potential
-$v_{\text{eff}}^B$
+$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
+with the
+effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
+The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
+$\hat{v}_{\text{eff}}^A$
 to produce
@@
 \begin{equation}
-	\sum_{ij}\sum_{kl\in B} \mathcal{F}\left[ 
-   \ERIc{i}{j}{k^B}{l^B}
- \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j}  \;.
+	\sum_{ij}\sum_{kl\in A} \mathcal{F}\left[ 
+   \BraKet{ij}{k^Al^A}
+ \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j}  \;.
 \end{equation}
@@
 \begin{equation} \label{e:ft-reduction}
-	\sum_t \sum_{ij}\sum_{kl\in B} {\mathcal{F}}_t\left[ 
-   \ERIc{i}{j}{k^B}{l^B}
- \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j} \;.
+	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
+   \BraKet{ij}{k^Al^A}
+ \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
 \end{equation}
 %
-The above design has the advantage that the effective potentials are fully first-principles
-and no extensive case-dependent fitting procedures are necessary.
-Only the effective potentials of \emph{independent} fragment $B$ need to be determined once and for all
-and stored in a file. Note also, that, in principle, there is no approximation 
-made here at that moment.
+The above design has the advantage that it opens the possibility to define first\hyp{}principles
+effective fragments as long as the $P_{ij}$ and $P_{ijkl}$ 
+from Eq.~\eqref{e:exp-val-series} are computable and can be approximately
+partitioned in between the interacting fragments.
+%the resulting effective potentials are fully first\hyp{}principles
+%and no extensive case\hyp{}dependent fitting procedures are necessary as long as the $P_{ij}$ and $P_{ijkl}$
+%are computable.
+%Only the effective potentials of \emph{independent} fragment $A$ need to be determined once and for all
+%and stored in a file. 
+%Note also, that, in principle, there is no approximation 
+%made here at that moment.
 
@@
 
-The above technique can be now applied to the Coulomb\hyp{}like and overlap\hyp{}like 
-terms from Eq.~\eqref{e:delta-p-part-2el}. For example, the Coulomb\hyp{}like contributions
-can be rewritten without making any approximation as
+The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
+under the localized basis set approximation 
+from Eq.~\eqref{e:delta-p-part}.
+%Coulomb\hyp{}like and overlap\hyp{}like 
+%terms from Eq.~\eqref{e:delta-p-part-2el}. 
+For example, the Coulomb\hyp{}like contributions
+can be rewritten without making any further approximation as
 %
@@
  \sum_{ijkl} 
-	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
+	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
  \equiv
@@
  \hat{v}^{{\rm eff},A} 
-	\equiv \sum_{ij\in A} {s} \left[ C^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
+	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
 \end{equation}
@@
  \sum_{ijkl} 
-	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
+	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
  \equiv
@@
  \hat{v}^{{\rm eff},A} 
-	\equiv \sum_{ij\in A} {s} \left[ C^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
+	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
 \end{equation}
 %
-Note also that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
+Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
 Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
@@
 	+ \sum_{i\in A} \sum_{j\in B} 
-	\left\{
-	  \tBraKet{i}{\hat{v}^{{\rm eff},A}_1 }{j}
-	+ \tBraKet{i}{\hat{v}^{{\rm eff},B}_1 }{j}
-	\right\} \\ +
-	\frac{1}{2}
+	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
+	 \\ +
 	\sum_{i\in A} \sum_{k\in A}
-	  \tBraKet{i}{\hat{v}^{{\rm eff},B}_2 }{k}
+	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
 	 +
-	 \frac{1}{2}
-        \sum_{l\in B} \sum_{l\in B}
-          \tBraKet{l}{\hat{v}^{{\rm eff},A}_2 }{l}
+        \sum_{j\in B} \sum_{l\in B}
+          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
 \end{multline}
@@
 	\Delta_{\rm ex}^{AB} P = 
-	\sum_{ik\in A} \sum_{jl\in B}
-	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] \;.
+	s \sum_{ik\in A} \sum_{jl\in B}
+	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
 \end{equation}
 %
-whereas
+and
 the overlap\hyp{}like and the Coulomb\hyp{}like 
@@
 \begin{align}
-	\hat{v}^{{\rm eff},A}_1 &\equiv C_{ij}^{AB} 
-	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
+	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
+	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \\
 	%
-	\hat{v}^{{\rm eff},A}_2 &\equiv 
-	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
+	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
+        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
+        \\
+	&+ \frac{1}{2}
+        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} 
 \end{align}
 %
-and accordingly for the twin terms (molecule $B$).
+and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.
 
-\section{\label{s.334}Practical Calculations with OEP's}
+\section{\label{s.334}Practical OEP-Based Calculations}
 
@@
 in which it is possible to define an effective one\hyp{}electron
-potential.
+potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
 The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
@@
 %
-\begin{subequations}
- \begin{align}
- \mathcal{F}
- \left[ 
-   \ERIc{i^A}{j^A}{k^B}{l^B}
-	 \right] = \OEIc{i^A}{v_{kl}^B}{j^A} \\ \nonumber \rightarrow \text{ multipole expansion or density fitting}\\ 
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[b]
+\caption[Types of matrix elements with OEP operators]
+{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
+}
+\label{t:oep-matrix-element-types}
+\begin{ruledtabular}
+\begin{tabular}{lcccc}
+Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
+                    && $\tBraKet{i}{\hat{v}^{{\rm eff},A}}{j} $ && $\tBraKet{j}{\hat{v}^{{\rm eff},A}}{l}$ \\ 
+	\cline{1-5}
+Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
+Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
+DF\footnotemark[2]/RI\footnotemark[3] Form    
+&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
+&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
+DMTP\footnotemark[4] Form                     
+&& --  &  &   \\
+\end{tabular}
+\end{ruledtabular}
 %
- \mathcal{F}
- \left[ 
-   \ERIc{i^A}{j^B}{k^B}{l^B}
-	 \right] = \OEIc{i^A}{v_{kl}^B}{j^B} \\ \rightarrow \text{ density fitting} \nonumber
- \end{align}
-\end{subequations}
+\footnotetext[1]{Test footnotemark.}
+\footnotetext[2]{Density Fitting}
+\footnotetext[3]{Resolution of Identity}
+\footnotetext[4]{Distributed Multipole Expansion}
+%
+\end{table}
+}
 %
@@
 An arbitrary one-electron potential of molecule $A$ acting on any state vector 
-associated with molecule $A$ can be expanded in an *auxiliary space* centered 
+associated with molecule $A$ can be expanded in an auxiliary space centered 
 on $A$ as
 \begin{equation}
-   v\vert i) = \sum_{\xi\eta} v\vert \xi) [{\bf S}^{-1}]_{\xi\eta} ( \eta \vert i)
+   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
 \end{equation}
-under the necessary assumption that the auxiliary basis set is nearly complete.
-In a special case when the basis set is orthogonal (e.g., molecular orbitals)
-the above relation simplifies to
+under the necessary assumption that the auxiliary basis set is nearly complete,
+i.e., 
+$\sum_{\xi\eta} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
+%In a special case when the basis set is orthogonal (e.g., molecular orbitals)
+%the above relation simplifies to
+%\begin{equation}
+%   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
+%\end{equation}
+The above general and exact expansion can be obtained by 
+utilizing the resolution of identity or, equivalently, density fitting
+in the complete space,
+%
 \begin{equation}
-   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
+ \hat{v}^{A}\Ket{i} = \sum_{\xi} V^A_{i\xi} \Ket{\xi} \;.
 \end{equation}
-It can be easily shown that the above general and exact expansion can be obtained by 
-performing a density fitting in the complete space. We expand the LHS of the first equation 
-on this page in a series of the auxiliary basis functions scaled by the undetermined expansion coefficients:
+%
+In the above equation,
+the matrix ${\bf V}^A$
+is the representation of the OEP operator $\hat{v}^{A}$
+in the complete basis $\{ \varphi_\xi \}$.
+By constructing the least\hyp{}squares objective function 
+%
 \begin{equation}
-  v\vert i) = \sum_{\xi} {G_{i\xi}} \vert \xi)
+ Z[{\bf V}^A] = \int 
+                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]^2
+d{\bf r}_1
 \end{equation}
-which we shall refer here as to the matrix form of the OEP operator.
-By constructing the least-squares objective function 
+%
+and requiring that
+%
 \begin{equation}
- Z[\{G^{(i)}_\xi\}] = \int d{\bf r}_1
-                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi G^{(i)}_\xi \varphi_\xi({\bf r}_1) \right]^2
+ \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
 \end{equation}
-and requiring that
+%
+one finds the coefficients of the $i$th row of ${\bf V}^A$ to be
+%
 \begin{equation}
- \frac{\partial Z[\{G^{(i)}_\xi\}]}{\partial G^{(i)}_\mu} = 0 \text{ for all $\mu$}
+% {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
+  V^A_{i\xi} = \sum_\eta [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
 \end{equation}
-we find the coefficients $ G^{(i)}_\xi $ to be
+%
+where
+%
+\begin{align}
+ a^{(i)}_\eta &= \int \varphi^*_\eta({\bf r}) \hat{v}^A \phi_i({\bf r}) d{\bf r} \;,\\  
+ S_{\eta\xi}  &= \int \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) d{\bf r} \;.
+\end{align}
+%
+%Since matrix elements of an OEP operator in auxiliary space can be computed 
+%in the same way as the matrix elements with any other basis function, 
+%one can formally write the following identity
+%\begin{equation}
+% (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
+%\end{equation}
+%where $ X $ is an arbitrary orbital.
+%When the other orbital
+%does not belong to molecule $A$ but to the (changing) environment, it is 
+%straightforward to compute the resulting matrix element, which is simply given as  
+%\begin{equation}
+%   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
+%\end{equation}
+%where $j$ denotes the other (environmental) basis function.
+%
+The working form of $a^{(i)}_\eta$ can be found by applying 
+potential for from Eq.~\eqref{e:v-eff}
+along with the spectral representation of the effective density from Eq.~\eqref{e:d-spectral} 
+which gives
+%
 \begin{equation}
- {\bf G}^{(i)} = {\bf v}^{(i)} \cdot {\bf S}^{-1}
+ a^{(i)}_\eta = \sum_{x\in A} W_{\eta i}^{(x)} + 
+ \sum_{\alpha\beta} D^A_{\alpha\beta} 
+  \BraKet{\alpha\beta}{\eta i} \;.
+%  \iint \frac{ \phi_\alpha^*({\bf r}_1)  \phi_\beta({\bf r}_1) \varphi^*_\eta({\bf r}_2) \phi_i({\bf r}_2)}
+% {\vert {\bf r}_1 - {\bf r}_2 \vert }
+% d{\bf r}_1 d{\bf r}_2
 \end{equation}
-where
-\begin{align}
- v^{(i)}_\eta &= (\eta \vert vi) \\
- S_{\eta\xi}  &= (\eta \vert \xi)
-\end{align}
-or explitictly
-\begin{equation}
-  {G_{i\xi}} = \sum_\eta [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
-\end{equation}
-identical to what we obtained from application of the resolution of identity 
-in space spanned by non-orthogonal complete set of basis vectors.
-
-Since matrix elements of an OEP operator in auxiliary space can be computed 
-in the same way as the matrix elements with any other basis function, 
-one can formally write the following identity
-\begin{equation}
- (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
-\end{equation}
-where $ X $ is an arbitrary orbital.
-When the other orbital
-does not belong to molecule $A$ but to the (changing) environment, it is 
-straightforward to compute the resulting matrix element, which is simply given as  
-\begin{equation}
-   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
-\end{equation}
-where $j$ denotes the other (environmental) basis function.
+%
 
@@
 \begin{multline}
-	Z[\{G^{(i)}_\xi\}] = \iint d{\bf r}_1 d{\bf r}_2 \Bigg\{
-                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi G^{(i)}_\xi \varphi_\xi({\bf r}_1) \right]
+	Z[{\bf V}^A] = \iint 
+        \Bigg\{
+                     \left[    v^A({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]
                   \\ \times \frac{1}{\vert {\bf r}_1 - {\bf r}_2 \vert}  
-                     \left[    v({\bf r}_2) \phi_i({\bf r}_2) - \sum_\xi G^{(i)}_\eta\varphi_\eta({\bf r}_1) \right]
+                     \left[    v^A({\bf r}_2) \phi_i({\bf r}_2) - \sum_\xi V^A_{i\xi} \varphi_\eta({\bf r}_1) \right]
 		    \Bigg\}
+         d{\bf r}_1 d{\bf r}_2  \;.
 \end{multline}
-Thus requesting that 
+Requesting that 
 \begin{equation}
- \frac{\partial Z[\{G^{(i)}_\xi\}]}{\partial G^{(i)}_\mu} = 0 \text{ for all $\mu$}
+ \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
 \end{equation}
-we find the coefficients $ G^{(i)}_\xi $ to be
+one obtains
 \begin{equation}
- {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
+% {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
+  V^A_{i\xi} = \sum_\eta [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
 \end{equation}
@@
 \begin{align}
- b^{(i)}_\eta &= (\eta \vert\vert vi) \\
- A_{\eta\xi}  &= (\eta \vert\vert \xi)
+ b^{(i)}_\eta &= \iint 
+                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 
+                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
+                 d{\bf r}_1 d{\bf r}_2 \;, \\
+ A_{\eta\xi}  &= \iint 
+                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
+                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
+                 d{\bf r}_1 d{\bf r}_2 \;.
 \end{align}
-The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
-and $ {\bf r}_2 $. Thus, it is clear that in order to use this generalized density fitting scheme
-one must to compute two\hyp{}centre electron repulsion integrals
-as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
-The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.
+%
+%The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
+%and $ {\bf r}_2 $. 
+Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI,
+$b^{(i)}_\eta$ matrix elements are not at all trivial to evaluate
+because the OEP operator, which contains integration over an electron coordinate,
+is present inside the double integral. Therefore, the following triple integral
+has to be computed
+%
+\begin{equation}
+ b^{(i)}_\eta = \iiint 
+           \frac{ \varphi^*_\eta({\bf r}_1) \phi_i({\bf r}_2)  \rho^{\rm eff}({\bf r}_3) }
+            {\vert {\bf r}_1 - {\bf r}_2 \vert \vert {\bf r}_3 - {\bf r}_2 \vert}
+           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3
+\end{equation}
+%
+Computation of all the necessery integrals of this kind is very costly 
+and impractical even for medium sized molecules. 
+However, one can introduce the effective potential in order to eliminate one integration. 
+This can be achieved by performing additional density fitting in a nearly complete intermediate basis 
+%
+\begin{equation}
+ \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} \Ket{\varepsilon}
+\end{equation}
+%
+where the symbol `Int' denotes the intermediate basis from a chosen resolution of identity.
+
+%Therefore, in order to use this generalized density fitting scheme
+%one must to compute two\hyp{}centre electron repulsion integrals
+%as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
+%The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.
 
@@
 
+%%
+%\appendix
 %
-\appendix
-
-\section{\label{a:mcmurchie-davidson} McMurchie-Davidson Method for Asymmetric Electron Repulsion Integrals}
-
-Evaluation of integrals $(ijk\vert l)$,
-that are necessary for the generalized density fitting of OEP's,
-can be easily carried out by re\hyp{}expressing the
-unnormalized product of three primitive Gaussian\hyp{}type (GTO) functions,
+%\section{\label{a:mcmurchie-davidson} McMurchie-Davidson Method for Asymmetric Electron Repulsion Integrals}
 %
-\begin{equation} \label{e:1}
-[ijk] \equiv \phi_i({\bf r}) \phi_j({\bf r}) \phi_k({\bf r}) \;,
-\end{equation}
+%Evaluation of integrals $(ijk\vert l)$,
+%that are necessary for the generalized density fitting of OEP's,
+%can be easily carried out by re\hyp{}expressing the
+%unnormalized product of three primitive Gaussian\hyp{}type (GTO) functions,
+%%
+%\begin{equation} \label{e:1}
+%[ijk] \equiv \phi_i({\bf r}) \phi_j({\bf r}) \phi_k({\bf r}) \;,
+%\end{equation}
+%%
+%in terms of Hermite functions. It can be shown that
+%%
+%\begin{multline} \label{e:2}
+%   [ijk] = E_{ijk} \sum_{N=0}^{n_1+n_2+n_3} \sum_{L=0}^{l_1+l_2+l+3} \sum_{M=0}^{m_1+m_2+m_3} 
+%	\\
+%          d_N^{n_1n_2n_3} d_L^{l_1l_2l_3} d_M^{m_1m_2m_3}
+%          \Lambda_N(x_R)\Lambda_L(y_R)\Lambda_M(z_R)e^{-\alpha_Rr_R^2}
+%\end{multline}
+%%
+%in which the McMurchie-Davidson $d3$ coefficients are given by the following recurrence relationships
+%%
+% \begin{align} \label{e:3}
+%  d_N^{n_1+1,n_2,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+%                             + \vert {\bf R} - {\bf A}\vert_x d_N^{n_1n_2n_3} \nonumber \\ 
+%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
+%  d_N^{n_1,n_2+1,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+%                             + \vert {\bf R} - {\bf B}\vert_x d_N^{n_1n_2n_3} \nonumber \\
+%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
+%  d_N^{n_1,n_2,n_3+1} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
+%                             + \vert {\bf R} - {\bf C}\vert_x d_N^{n_1n_2n_3} \nonumber \\
+%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} 
+%\end{align}
+%%
+%with $d_0^{000} = 1$.
+%In Eq.~\eqref{e:1},
+%$\phi_i({\bf r})$ is given by
+%%
+%\begin{equation}
+%\phi_i({\bf r}) \equiv x_A^{n_1} y_A^{l_1} z_A^{m_1} e^{-\alpha_1r_A^2}
+%\end{equation}
+%%
+%where 
+%${\bf r}_A \equiv {\bf r} - {\bf A}$,
+%${\bf A}$ is the centre of the GTO, $\alpha_1$ its exponent, whereas $n_1,l_1,m_1$
+%the Cartesian angular momenta, with the total angular momentum $\theta_1 = n_1+l_1+m_1$.
 %
-in terms of Hermite functions. It can be shown that
-%
-\begin{multline} \label{e:2}
-   [ijk] = E_{ijk} \sum_{N=0}^{n_1+n_2+n_3} \sum_{L=0}^{l_1+l_2+l+3} \sum_{M=0}^{m_1+m_2+m_3} 
-	\\
-          d_N^{n_1n_2n_3} d_L^{l_1l_2l_3} d_M^{m_1m_2m_3}
-          \Lambda_N(x_R)\Lambda_L(y_R)\Lambda_M(z_R)e^{-\alpha_Rr_R^2}
-\end{multline}
-%
-in which the McMurchie-Davidson $d3$ coefficients are given by the following recurrence relationships
-%
- \begin{align} \label{e:3}
-  d_N^{n_1+1,n_2,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-                             + \vert {\bf R} - {\bf A}\vert_x d_N^{n_1n_2n_3} \nonumber \\ 
-			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
-  d_N^{n_1,n_2+1,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-                             + \vert {\bf R} - {\bf B}\vert_x d_N^{n_1n_2n_3} \nonumber \\
-			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
-  d_N^{n_1,n_2,n_3+1} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-                             + \vert {\bf R} - {\bf C}\vert_x d_N^{n_1n_2n_3} \nonumber \\
-			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} 
-\end{align}
-%
-with $d_0^{000} = 1$.
-In Eq.~\eqref{e:1},
-$\phi_i({\bf r})$ is given by
-%
-\begin{equation}
-\phi_i({\bf r}) \equiv x_A^{n_1} y_A^{l_1} z_A^{m_1} e^{-\alpha_1r_A^2}
-\end{equation}
-%
-where 
-${\bf r}_A \equiv {\bf r} - {\bf A}$,
-${\bf A}$ is the centre of the GTO, $\alpha_1$ its exponent, whereas $n_1,l_1,m_1$
-the Cartesian angular momenta, with the total angular momentum $\theta_1 = n_1+l_1+m_1$.
-
-It can be easily shown that the multiplicative constant $E_{ijk}$ is given by
-%
-\begin{multline}
-  E_{ijk}(\alpha_1,\alpha_2,\alpha_3)  = \exp{\left[-\frac{\alpha_1\alpha_2}
-                                        {\alpha_1+\alpha_2}\vert {\bf A}-{\bf B}\vert^2\right]} \\ \times
-                                         \exp{\left[-\frac{(\alpha_1+\alpha_2)\alpha_3}
-                                        {\alpha_1+\alpha_2+\alpha_3}
-                                         \vert {\bf P}-{\bf C}\vert^2\right]} 
-\end{multline}
-%
+%It can be easily shown that the multiplicative constant $E_{ijk}$ is given by
+%%
+%\begin{multline}
+%  E_{ijk}(\alpha_1,\alpha_2,\alpha_3)  = \exp{\left[-\frac{\alpha_1\alpha_2}
+%                                        {\alpha_1+\alpha_2}\vert {\bf A}-{\bf B}\vert^2\right]} \\ \times
+%                                         \exp{\left[-\frac{(\alpha_1+\alpha_2)\alpha_3}
+%                                        {\alpha_1+\alpha_2+\alpha_3}
+%                                         \vert {\bf P}-{\bf C}\vert^2\right]} 
+%\end{multline}
+%%
 

### 2019-05-21 papers/2_oep_technique/paper-r0.tex
Generalized density fitting in incomplete basis sketched

@@
 %
+\begin{subequations}
 \begin{align}
 	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
-	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \\
+	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \;, \\
 	%
@@
 	&+ \frac{1}{2}
-        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} 
+        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
 \end{align}
+\end{subequations}
 %
@@
 %
+\begin{subequations}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 %
@@
 %
-The working form of $a^{(i)}_\eta$ can be found by applying 
+The working formula for $a^{(i)}_\eta$ can be found by applying 
 potential for from Eq.~\eqref{e:v-eff}
@@
 where 
+\begin{subequations}
 \begin{align}
@@
 \begin{equation}
- \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} \Ket{\varepsilon}
+ \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} \Ket{\varepsilon} \;,
 \end{equation}
@@
 where the symbol `Int' denotes the intermediate basis from a chosen resolution of identity.
-
+The working equation is therefore
+%
+\begin{equation}
+ b^{(i)}_\eta = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} R_{\varepsilon\eta} 
+\end{equation}
+%
+with
+%
+\begin{equation}
+ R_{\varepsilon\eta} \equiv \iint 
+                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
+                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
+                 d{\bf r}_1 d{\bf r}_2 \;,
+\end{equation}
+%
+which can be easily evaluated and the matrix elements of the OEP operator found.
 %Therefore, in order to use this generalized density fitting scheme

### 2019-05-22 papers/2_oep_technique/paper-r0.tex
More about OEP definition

@@
 %
-\begin{multline}
+\begin{multline} \label{e:ci-oep-master}
  \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
@@
 \begin{subequations}
-\begin{align}
+\begin{align} \label{e:oep-exch-like}
 	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
@@
         + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
-        \\
-	&+ \frac{1}{2}
+        \\ \label{e:oep-coul-like}
+	& \qquad\qquad+ \frac{1}{2}
         s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
@@
 and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.
+
+\subsection{\label{s.242}Separability of P-Coefficients}
+
+Until that point, the effective potentials from Eqs.~\eqref{e:oep-exch-like}
+and \eqref{e:oep-coul-like} are not separable in terms of the interacting molecules
+because the $\bf P$ tensors are derived from the CI wavefunction of
+the entire interacting system $AB$. However, if there exist an approximate
+separation as, for example,
+%
+\begin{align}
+ P^{AAAB}_{mkij} &\approx X^A_{mk} Y^{AB}_{ij} \;, \\
+ P^{AABB}_{ikjl} &\approx Z^A_{ik} Y^{AB}_{} Z^{ B}_{jl} \;,
+\end{align}
+%
+and if the matrices ${\bf X}^{A(B)}$, 
+${\bf Z}^{A(B)}$ and ${\bf P}^{AA(BB)}$ as well as the tensors
+${\bf P}^{AAAA(BBBB)}$ 
+are properties of the unperturbed (isolated) molecules,
+further rearrangements in Eq.~\eqref{e:ci-oep-master} are possible
+leading to computationally more efficient expressions.
+
+Note also that the matrices ${\bf Y}^{AB}$ and ${\bf P}^{AB}$ depend on their instantaneous 
+relative positions.
+
+%then the molecule\hyp{}specific OEP's can be defined.
 
@@
 The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
-The second type is an overlap\hyp{}like contribution, can be approximated with the density fitting. Mainly,
+The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. Mainly,
 %

### 2019-05-24 papers/2_oep_technique/paper-r0.tex
Working formulae derived for DF-temr in CT energy of Murrell et al

@@
 
-\section{\label{s:2}Effective One-Electron Potentials: Configuration Interaction}
+\section{\label{s:3}Effective One-Electron Potentials: Configuration Interaction}
 
@@
 (approximate) wavefunction, $\Psi^{(0)}$,
-typically the solution to the Hartree\hyp{}Fock (HF) problem,
+typically the solution to the Hartree\hyp{}Fock (HF) problem\cite{Roothaan.RevModPhys.1951},
 %
@@
 
-\subsection{\label{ss:2.3}Defining Effective Potentials}
+\subsection{\label{ss:2.3}Incorporating Electron Repulsion Integrals into Effective Potentials}
 
@@
 
-\subsection{\label{ss:2.4}Interaction-Induced Property: Effective Potentials}
+\subsection{\label{ss:2.4}Definition of Interaction-Induced Property Effective Potentials}
 
@@
 
-\subsection{\label{s.334}Density Fitting of OEP's}
+\subsection{\label{s.3344}Density Fitting of OEP's}
 
@@
 on $A$ as
+%
 \begin{equation}
-   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
+   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta}^{\rm RI} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
 \end{equation}
+%
 under the necessary assumption that the auxiliary basis set is nearly complete,
 i.e., 
-$\sum_{\xi\eta} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
+$\sum_{\xi\eta}^{\rm RI} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
+In the equations above, 
+the `RI' symbol denotes a certain auxiliary basis set that fulfills the resolution of identity. 
 %In a special case when the basis set is orthogonal (e.g., molecular orbitals)
@@
 %\end{equation}
-The above general and exact expansion can be obtained by 
-utilizing the resolution of identity or, equivalently, density fitting
-in the complete space,
+Such a general expansion can be obtained by 
+utilizing the density fitting
+in the nearly\hyp{}complete space,
 %
 \begin{equation}
- \hat{v}^{A}\Ket{i} = \sum_{\xi} V^A_{i\xi} \Ket{\xi} \;.
+ \hat{v}^{A}\Ket{i} = \sum_{\xi}^{\rm RI} V^A_{i\xi} \Ket{\xi} \;.
 \end{equation}
@@
 the matrix ${\bf V}^A$
-is the representation of the OEP operator $\hat{v}^{A}$
-in the complete basis $\{ \varphi_\xi \}$.
-By constructing the least\hyp{}squares objective function 
+is the projection of the state vector $\hat{v}^{A} \Ket{i}$
+onto the complete basis $\{ \varphi_\xi \}$.
+Let $Z_i[{\bf V}^A]$ be the least\hyp{}squares objective function 
+%
+\begin{equation} \label{e:z-compl}
+ Z_i[{\bf V}^A] = \int \vert \Xi_i({\bf r}) \vert^2 d{\bf r}
+\end{equation}
+%
+with the error density defined by
 %
 \begin{equation}
- Z[{\bf V}^A] = \int 
-                     \left[    v({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]^2
-d{\bf r}_1
+ \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm RI} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
 \end{equation}
 %
-and requiring that
+By requiring that
 %
-\begin{equation}
- \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
+\begin{equation} \label{e:z-necessary-requirement}
+ \frac{\partial Z_i[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
 \end{equation}
@@
 % {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
-  V^A_{i\xi} = \sum_\eta [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
+  V^A_{i\xi} = \sum_\eta^{\rm RI} [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
 \end{equation}
 %
-where
+where the auxiliary matrices are given by
 %
@@
 The working formula for $a^{(i)}_\eta$ can be found by applying 
-potential for from Eq.~\eqref{e:v-eff}
+potential form from Eq.~\eqref{e:v-eff}
 along with the spectral representation of the effective density from Eq.~\eqref{e:d-spectral} 
-which gives
+which finally gives
 %
 \begin{equation}
- a^{(i)}_\eta = \sum_{x\in A} W_{\eta i}^{(x)} + 
+ a^{(i)}_\eta = \sum_{x} W_{\eta i}^{(x)} + 
  \sum_{\alpha\beta} D^A_{\alpha\beta} 
@@
 
-Density fitting scheme from previous section has practical disadvantage of a nearly-complete basis set
-being usually very large (spanned by large amount of basis set vectors). Any non-complete basis set
-won't work in the previous example. Since most of basis sets used in quantum chemistry do not form a nealy complete
+Density fitting scheme from previous section has practical disadvantage of a nearly\hyp{}complete basis set
+being usually very large (spanned by large amount of basis set vectors). 
+Since most of basis sets used in quantum chemistry do not form a nealy complete
 set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
@@
 the following objective function
-\begin{multline}
-	Z[{\bf V}^A] = \iint 
-        \Bigg\{
-                     \left[    v^A({\bf r}_1) \phi_i({\bf r}_1) - \sum_\xi V^A_{i\xi} \varphi_\xi({\bf r}_1) \right]
-                  \\ \times \frac{1}{\vert {\bf r}_1 - {\bf r}_2 \vert}  
-                     \left[    v^A({\bf r}_2) \phi_i({\bf r}_2) - \sum_\xi V^A_{i\xi} \varphi_\eta({\bf r}_1) \right]
-		    \Bigg\}
+%
+\begin{equation} \label{e:z-incompl}
+	Z_i[{\bf V}^A] = \iint 
+        \frac{ \Xi_i^*({\bf r}_1) \Xi_i({\bf r}_2) }{\vert {\bf r}_1 - {\bf r}_2 \vert}  
          d{\bf r}_1 d{\bf r}_2  \;.
-\end{multline}
-Requesting that 
+\end{equation}
+%
+with the error density defined by
+%
 \begin{equation}
- \frac{\partial Z[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
+ \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm DF} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
 \end{equation}
+%
+where the symbol `DF' denotes the generally incomplete auxiliary basis set.
+From the requirement given in Eq.~\eqref{e:z-necessary-requirement}
 one obtains
-\begin{equation}
+%
+\begin{equation} \label{e:v-noncompl}
 % {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
-  V^A_{i\xi} = \sum_\eta [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
+  V^A_{i\xi} = \sum_\eta^{\rm DF} [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
 \end{equation}
+%
 where 
+%
 \begin{subequations}
@@
 %and $ {\bf r}_2 $. 
-Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI,
-$b^{(i)}_\eta$ matrix elements are not at all trivial to evaluate
+Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI 
+that can be evaluated by standard means,
+$b^{(i)}_\eta$ matrix elements are not at all trivial to calculate
 because the OEP operator, which contains integration over an electron coordinate,
-is present inside the double integral. Therefore, the following triple integral
-has to be computed
+is present inside the double integral. Therefore, the following triple integral,
 %
-\begin{equation}
+\begin{equation} \label{e:triple-integral}
  b^{(i)}_\eta = \iiint 
@@
             {\vert {\bf r}_1 - {\bf r}_2 \vert \vert {\bf r}_3 - {\bf r}_2 \vert}
-           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3
+           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3 \;,
 \end{equation}
 %
-Computation of all the necessery integrals of this kind is very costly 
+has to be computed.
+Obtaining all the necessery integrals of this kind directly 
+by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 
 and impractical even for medium sized molecules. 
@@
 \begin{equation}
- \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} \Ket{\varepsilon} \;,
+ \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} \Ket{\varepsilon} \;,
 \end{equation}
 %
-where the symbol `Int' denotes the intermediate basis from a chosen resolution of identity.
-The working equation is therefore
+%where the symbol `RI' denotes the intermediate basis from a chosen resolution of identity.
+The working equation is therefore given by
 %
@@
 %
-which can be easily evaluated and the matrix elements of the OEP operator found.
+which can be easily evaluated and the matrix elements of the OEP operator found
+from Eq.~\eqref{e:v-noncompl}.
 %Therefore, in order to use this generalized density fitting scheme
@@
 
-\section{\label{s:3}Calculation Details}
-\section{\label{s:4}Results and Discussion}
-\subsection{\label{ss.4.1}Pauli-Repulsion Interaction Energy}
-\subsection{\label{ss.4.2}Charge-Transfer Interaction Energy}
-\section{\label{s:5}Summary and a few concluding remarks}
+It is emphasized here that, as long as the state vector $\Ket{i}$, OEP operator $\hat{v}^A$ 
+and the auxiliary and intermediate basis sets depend solely on one unperturbed molecule $A$, the matrix elements
+$V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.
+
+\section{\label{s:4}Calculation Details}
+
+All the models that were used to test the theory presented in this work
+were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
+
+\section{\label{s:5}Results and Discussion}
+\subsection{\label{ss.5.1}Pauli-Repulsion Interaction Energy}
+\subsection{\label{ss.5.2}Charge-Transfer Interaction Energy}
+
+Charge transfer (CT) between molecules occurs when the net electronic populations 
+of interacting molecules change which leads to an additional stabilization 
+of a molecular aggregate. Since the effect of CT on the interaction energy 
+is known to be often non\hyp{}negligible, it is quite important to model 
+CT energies with reasonable accuracy and efficiency when performing energy 
+and molecular dynamics calculations with \emph{ab initio} force fields, 
+especially in donor\hyp{}acceptor systems such as H\hyp{}bonded species. 
+However, estimating the CT stabilization energy is usually much more expensive 
+than estimating the contributions due to other effects such as electrostatics, 
+exchange\hyp{}repulsion and dispersion.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
+
+As an illustration of application of the OEP technique for efficient calculation 
+of CT energy we consider the second\hyp{}order perturbation theory of Murrell et al.,
+in which the energy associated with the charge transfer from molecule $A$ to $B$ is given by
+%
+\begin{equation} \label{e:ct-murell-etal.e}
+ E^{A\rightarrow B} = 2 \sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir} 
+  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n}
+\end{equation}
+%
+%coupling constants between occupied and virtual molecular orbitals 
+%are re-expressed by the effective one-electron potentials. The resulting formula provides new possibilities 
+%of optimization and applications in condensed phase simulations. [3]
+%
+where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
+of the occupied (denoted by `Occ') and virtual (denoted by `Vir') 
+HF molecular orbitals, respectively,
+whereas the coupling constant is given by
+%
+\begin{multline} \label{e:ct-murell-etal.vin}
+ V^{A\rightarrow B}_{in} = 
+        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
+      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
+      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
+        S_{kj} \left( 1 - \delta_{ik} \right) 
+        \BraKet{nj}{ik} \;.
+\end{multline}
+%
+In the above expression, the following effective potentials,
+%
+\begin{subequations} 
+\begin{align} \label{e:ct-murell-etal.vtot-vi}
+ \hat{v}^B_{\rm tot} &= \hat{v}^B_{\rm nuc} + 2\sum_{j\in B}^{\rm Occ} \hat{v}^B_{jj} \quad\text{ and }\\ 
+ \hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
+\end{align}
+\end{subequations}
+%
+were introduced without making any approximation to the original equation
+from Ref. One can immediately notice that the five summation terms
+from Eq.~\eqref{e:ct-murell-etal.vin} can be classified into three groups
+regarding the type of ERI's that are required:
+(i) $\BraKet{AB}{BB}$ -- the first two terms;
+(ii) $\BraKet{AA}{BB}$ -- the third term and
+(iii) $\BraKet{BB}{AA}$ -- the two last terms. 
+%Note that groups (ii) and (iii) correspond to the Coulomb interactions between
+%The group (i) corresponds to the overlap\hyp{}like integrals, whereas the groups (ii) and (iii)
+%correspond to the Coulomb\hyp{}like integrals
+
+\paragraph{Group (i).}
+Group (i) can be rewritten by noticing that
+%
+\begin{equation} \label{e:ct-murell-etal.group-i.notice}
+ \sum_{j\in B}^{\rm Occ} 
+ \BraKet{nj}{ij} =
+ \Bra{i} \left[ \hat{v}^B_{nj} \Ket{j} \right]  \;,
+\end{equation}
+%
+which, by combining with the first term from Eq.~\eqref{e:ct-murell-etal.vin}, 
+allows to apply the density fitting in an
+incomplete basis set as follows
+%
+\begin{equation}
+\Bra{i} \left[ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
+\cong \Bra{i} \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta}
+\end{equation}
+%
+with the error density defined by
+%
+\begin{multline}
+ \Xi_n({\bf r}) = 
+  \left\{
+      v_{nuc}^B({\bf r}) + 2 \sum_{j\in B}^{\rm Occ}
+   \int \frac{ \vert \phi_j({\bf r}') \vert^2 }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+  \right\}
+  \phi_n({{\bf r}})  \\
+ - \sum_{j\in B}^{\rm Occ} \phi_j({{\bf r}})
+   \int \frac{ \phi^*_n({\bf r}') \phi_j({{\bf r}'}) }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+ - \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \varphi_\eta({\bf r}) \;.
+\end{multline}
+%
+Substituting the above equation into Eq.~\eqref{e:z-incompl}
+%and then
+%using resolution of identity to simplify the 
+leads to
+%
+\begin{equation}
+ V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf A}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
+\end{equation}
+%
+where
+%
+\begin{multline}
+ b^{(n)}_\eta = \iint 
+           d{\bf r}_1 d{\bf r}_2  
+           \frac{ \varphi^*_\eta({\bf r}_1) }
+            {\vert {\bf r}_1 - {\bf r}_2 \vert } \\ \times
+          \left[ 
+           \hat{v}^B_{\rm tot}  \phi_n({\bf r}_2) 
+         - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \phi_j({{\bf r}_2})
+           \right]  \;.
+\end{multline}
+%
+The above result is given by a sum of triple integrals from Eq.~\eqref{e:triple-integral}.
+However, application of the resolution of identity as follows
+%
+\begin{equation}
+ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
+ \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon}
+\end{equation}
+%
+leads to a much simpler formula
+that requires only the one- and two\hyp{}electron integrals. 
+Therefore, by combining
+equations %which equations?
+one arrives to
+%
+\begin{equation}
+ V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} 
+          \sum_{\varepsilon\in B}^{\rm RI}
+         \left[ {\bf A}^{-1} \right]_{\xi\eta}
+         H_{n\varepsilon}^B R_{\varepsilon \eta} \;,
+\end{equation}
+%
+where
+%
+\begin{equation}
+ H_{n\varepsilon}^B = \sum_{\eta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\eta}
+   a_\eta^{(n)}
+\end{equation}
+%
+and
+%%
+\begin{align}
+ a^{(n)}_\eta &= \tBraKet{\eta}{\hat{v}^B_{\rm tot}}{n}
+      - \sum_{j\in B}^{\rm Occ} \tBraKet{\eta}{\hat{v}^B_{nj}}{j} \nonumber \\
+ &= \sum_{y\in B} W^{(y)}_{\eta n} 
+  + \sum_{j\in B}^{\rm Occ} 
+  \left\{
+   2\BraKet{\eta n}{jj} - \BraKet{\eta j}{nj} 
+  \right\} \;.
+\end{align}
+%
+Note that all the calculations that are required to obtain $V^B_{n\xi}$ are performed
+solely on the densities and basis sets associated with the unperturbed molecule $B$.
+Therefore, $V^B_{n\xi}$ can be considered as effective fragment parameters
+used to compute the first two terms of Eq.~\eqref{e:ct-murell-etal.vin} by
+%
+\begin{equation}
+        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
+      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
+       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;,
+\end{equation}
+%
+which is a great simplification over the original form of group (i)
+because only the overlap integrals between the $i$th MO on molecule $A$
+and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated on the fly.
+
+\paragraph{Group (ii).}
+\paragraph{Group (iii).}
+\paragraph{Final OEP-based form of the coupling constant.}
+Gathering the results from Eqs. the coupling constant
+can be rewritten as follows
+%
+\begin{equation} \label{e:ct-murell-etal.vin.oep}
+ V_{in}^{A\rightarrow B}
+       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;.
+\end{equation}
+%
+Note that the only approximations made on it so fare was the application of density fitting
+and resolution of identity. If the associated basis sets are sufficiently large, the errors
+due to this approximation can be minimal and negligible.
+
+%he above formula does not explicitly define an OEP due to the summation over different
+%state vectors. 
+%However, it can be achieved
+%by transforming from MO to AO basis, i.e.,
+%%
+%\begin{equation}
+% \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
+% = \sum_{\beta\in B}^{\rm AO} 
+% \left\{ 
+%  C_{\beta n} \hat{v}^B_{\rm tot} - \sum_{j\in B}^{\rm Occ} C_{\beta j} \hat{v}^B_{nj} 
+% \right\} \Ket{\beta} \;,
+%\end{equation}
+%%
+%which leads to the OEP in spatial representation
+%%
+%\begin{multline}
+% v_{\beta n}({\bf r}) \equiv C_{\beta n} v_{nuc}^B({\bf r}) \\ + 
+%  \sum_{j\in B}^{\rm Occ} \Bigg\{ 
+%   2 C_{\beta n} \int \frac{ \vert \phi_j({\bf r}') \vert^2 }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+% %\\
+% - C_{\beta j} \int \frac{ \phi^*_n({\bf r}') \phi_j({\bf r}') }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+%  \Bigg\}  
+%%
+%\end{multline}
+%%
+
+\section{\label{s:6}Summary and a few concluding remarks}
 

### 2019-05-24 papers/2_oep_technique/paper-r0.tex
Group II of terms for CT energy described

@@
 \paragraph{Group (ii).}
+The term belonging to this group can be considered as a sum of interaction
+energies between the total charge density distribution of molecule $B$
+and the partial density $\rho_{ik}({\bf r})$ of molecule $A$,
+weighted by the overlap integrals $S_{nk}$. Using the distributed multipole 
+expansion this can be written as
+%
+\begin{equation} \label{e:ct.group-2}
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
+\end{equation}
+%
+Consider now a situation in which the localized occupied MO's are used.
+If the multipole series expansions as truncated
+at the distributed charges $q_i$ and dipole moments ${\BM{\upmu}_i}$, that are
+located at the centroids ${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$, 
+then the interaction energy can be written approximately as
+%
+\begin{multline}
+ \rho_{ki}^A \odot \rho^B_{\rm tot} 
+ \approx 
+ q_{ki} 
+ \left[
+  %\sum_{i\in A}^{\rm LC} 
+ \sum_{y\in B}^{\rm AT}
+  \frac{Z_y}{\vert {\bf r}_y - {\bf r}_i \vert } 
+ +
+ \sum_{j\in B}^{\rm LC}
+  \frac{q_j}{\vert {\bf r}_j - {\bf r}_i \vert } 
+ \right] \\
+ + \text{ dipole terms } \;.
+\end{multline}
+%
+Note however that $q_{ik} = -\delta_{ik}$ and $q_j = -1$. Moreover, 
+if the localized orbitals are perfectly spherical distributed dipole moments vanish.
+This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows
+%
+\begin{equation}
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \approx S_{ni}  \left[
+  %\sum_{i\in A}^{\rm LC} 
+ \sum_{y\in B}^{\rm AT}
+  \frac{Z_y}{ r_{yi} } 
+ -
+ \sum_{j\in B}^{\rm LC}
+  \frac{1}{r_{ji}} 
+ \right] \;.
+\end{equation}
+%
+Therefore, only overlap integrals are needed and relative distances between
+atomic and LC's which leads to a great reduction of the evaluation cost
+as compared to the original equation or the full multipole expansion.
+We shall refer to this approximation as to
+the spherical overlap approximation (SOA). Note that in order for this approximation to be valid,
+occupied molecular orbitals need to be localized.
+
+%If localized occupied MO's are used and only distributed charges ar, then $\rho_{ki} \approx \delta_{ki}$
+
 \paragraph{Group (iii).}

### 2019-05-27 papers/2_oep_technique/paper-r0.tex
All 3 groups for CT-OEP energy added. Final working formula to be implemented added

@@
  V^{A\rightarrow B}_{in} = 
-        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
+        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
@@
 The above result is given by a sum of triple integrals from Eq.~\eqref{e:triple-integral}.
-However, application of the resolution of identity as follows
+However, the following application of the resolution of identity,
 %
@@
  \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
- \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon}
+ \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon} \;,
 \end{equation}
@@
 \begin{equation}
-        \tBraKet{i}{\hat{v}^B_{\rm ch} }{n} 
+        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
@@
 and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated on the fly.
+Note that the only approximation made so far was the application of density fitting
+and resolution of identity. If the auxiliary and intermediate
+basis sets are sufficiently large, the errors
+due to this approximation can be minimal and negligible.
 
@@
 weighted by the overlap integrals $S_{nk}$. Using the distributed multipole 
-expansion this can be written as
+expansion this group can be written as
 %
@@
 Consider now a situation in which the localized occupied MO's are used.
-If the multipole series expansions as truncated
-at the distributed charges $q_i$ and dipole moments ${\BM{\upmu}_i}$, that are
+If the multipole series expansions are truncated
+at the distributed charges $q_i$ and dipole moments ${\BM{\upmu}_i}$, 
 located at the centroids ${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$, 
-then the interaction energy can be written approximately as
+then the interaction energy becomes
 %
@@
   %\sum_{i\in A}^{\rm LC} 
- \sum_{y\in B}^{\rm AT}
+ \sum_{y\in B}^{\rm At}
   \frac{Z_y}{\vert {\bf r}_y - {\bf r}_i \vert } 
  +
- \sum_{j\in B}^{\rm LC}
+ 2\sum_{j\in B}^{\rm LC}
   \frac{q_j}{\vert {\bf r}_j - {\bf r}_i \vert } 
@@
 if the localized orbitals are perfectly spherical distributed dipole moments vanish.
-This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows
+This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows:
 %
@@
   %\sum_{i\in A}^{\rm LC} 
- \sum_{y\in B}^{\rm AT}
+ \sum_{y\in B}^{\rm At}
   \frac{Z_y}{ r_{yi} } 
@@
  \sum_{j\in B}^{\rm LC}
-  \frac{1}{r_{ji}} 
+  \frac{2}{r_{ji}} 
  \right] \;.
@@
 %
-Therefore, only overlap integrals are needed and relative distances between
-atomic and LC's which leads to a great reduction of the evaluation cost
-as compared to the original equation or the full multipole expansion.
+Therefore, only overlap integrals and relative distances between
+atomic and LC's are needed, which leads to a great reduction of the
+calculation cost,
+as compared either to the original expression or to the multipole expansion (left\hyp{} and right\hyp{}hand sides
+of Eq.~\eqref{e:ct.group-2}, respectively).
 We shall refer to this approximation as to
@@
 
-%If localized occupied MO's are used and only distributed charges ar, then $\rho_{ki} \approx \delta_{ki}$
-
 \paragraph{Group (iii).}
-\paragraph{Final OEP-based form of the coupling constant.}
+The terms with the overlap integrals involving the occupied MO on $A$
+can be combined into a single summation term, i.e.,
+\begin{multline}
+       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
+        S_{kj} \left( 1 - \delta_{ik} \right) 
+        \BraKet{nj}{ik} \\ = -
+ \sum_{k\in A}^{\rm Occ} 
+ \sum_{j\in B}^{\rm Occ}
+ S_{kj} 
+ \tBraKet{j}{
+ \underbrace{
+ \left\{ 
+  \delta_{ik} \left( \hat{v}^A_{k} + \hat{v}^A_{ik} \right)
+  - \hat{v}^A_{ik}
+ \right\} 
+  }_{\hat{v}^{A,{\rm eff}}_{ik}}
+ }{n}  \\ \cong -
+%
+  \sum_{k\in A}^{\rm Occ} 
+ \sum_{j\in B}^{\rm Occ}
+ S_{kj} 
+ \rho_{nj}^B \odot \rho^{A,{\rm eff}}_{ik} \;,
+\end{multline}
+%
+where the effective potential ${v}^{A,{\rm eff}}_{ik}$ (with the associated 
+effective density $\rho^{A,{\rm eff}}_{ik}$) 
+is defined by
+%
+\begin{equation}
+ {v}^{A,{\rm eff}}_{ik}({\bf r}) \equiv
+ \delta_{ik} 
+ \left[
+  v^A_{\rm tot} ({\bf r}) - 2 v^A_{k} ({\bf r}) + v^A_{ik} ({\bf r})
+ \right] 
+  - v^A_{ik} ({\bf r}) \;.
+\end{equation}
+%
+Now, it is first assumed here that $\rho^{B}_{nj} ({\bf r}) \approx 0$ if occupied MO's on $B$ are localized.
+This means that
+%
+\begin{multline}
+        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
+        S_{kj} \left( 1 - \delta_{ik} \right) 
+        \BraKet{nj}{ik}  \\
+ \approx 0 \;.
+\end{multline}
+%
+The above result is a crude approximation, since virtual MO's can overlap with occupied MO's to some extent.
+In order to maintain $\rho^{B}_{nj}$ density, it is approximately represented here by a set of effective 
+cumulative
+atomic charges $\{ q^{B,(nj)}_{y} \}$ associated with the effective one\hyp{}particle density matrix
+%
+\begin{equation}
+ D^{B,(nj)}_{\beta\delta} = C_{\beta n} C_{\delta j} \;.
+\end{equation}
+%
+In this work, the effective charges were computed via the CAMM or ESP methods as
+discussed in Section~XXX. Applying the SOA leads to 
+%
+\begin{multline}
+        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
+        S_{kj} \left( 1 - \delta_{ik} \right) 
+        \BraKet{nj}{ik}  \\
+ \approx  -
+ \sum_{j\in B}^{\rm Occ} S_{ij}
+ \sum_{y\in B}^{\rm At} 
+ q^{B,(nj)}_{y} 
+ \left[ 
+   \sum_{x\in A}^{\rm At}
+   \frac{Z_x}{r_{xy}}
+  + \frac{2}{r_{iy}}
+  - \sum_{k\in A}^{\rm Occ}
+    \frac{2}{r_{ky}} 
+ \right]
+ \;.
+\end{multline}
+%
+
+\paragraph{Final OEP-based forms of the coupling constant.}
 Gathering the results from Eqs. the coupling constant
-can be rewritten as follows
-%
-\begin{equation} \label{e:ct-murell-etal.vin.oep}
- V_{in}^{A\rightarrow B}
-       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;.
-\end{equation}
-%
-Note that the only approximations made on it so fare was the application of density fitting
-and resolution of identity. If the associated basis sets are sufficiently large, the errors
-due to this approximation can be minimal and negligible.
-
+can be given as follows
+%
+\begin{subequations}
+\begin{align} \label{e:ct-murell-etal.vin.oep}
+\Approx{ V_{in}^{A\rightarrow B} }{1}
+       &= \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} 
+ +
+S_{ni} u_i^{BA} \;, \\
+%
+\Approx{ V_{in}^{A\rightarrow B} }{2} &=
+\Approx{ V_{in}^{A\rightarrow B} }{1} -
+  \sum_{j   \in B}^{\rm Occ} S_{ij}
+  \sum_{y   \in B}^{\rm At} q_y^{B,(nj)} w_{yi}^{BA}
+\;,
+\end{align}
+\end{subequations}
+%
+where the auxiliary variables are
+%
+\begin{subequations}
+\begin{align} \label{e:ct-murell-etal.vin.oep.aux-var}
+ u_i^{BA} &\equiv    
+ \sum_{y\in B}^{\rm At}
+  \frac{Z_y}{ r_{yi} } 
+ -
+ \sum_{j\in B}^{\rm LC}
+  \frac{2}{r_{ji}} 
+                \;, \\
+ w_{yi}^{BA} &\equiv 
+   \sum_{x\in A}^{\rm At}
+   \frac{Z_x}{r_{xy}}
+  + \frac{2}{r_{iy}}
+  - \sum_{k\in A}^{\rm Occ}
+    \frac{2}{r_{ky}} 
+                \;.
+\end{align}
+\end{subequations}
+%
+To compute the interaction energy due to CT from $A$ to $B$ by using the above SOA\hyp{}based
+approximations, coupling elements need to be transformed from localized to canonical MO basis.
+The final working formula reads
+%
+\begin{equation} \label{e:ect-oep.final}
+ E^{A\rightarrow B} \approx
+ 2 
+ \sum_{i\in A}^{\rm CMO}
+ \sum_{n\in B}^{\rm Vir}
+ \frac{
+ \lvert
+   \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
+   \Approx{V_{i'n}^{A\rightarrow B}}{} 
+ \rvert^2 }{\varepsilon_i - \varepsilon_n} \;,
+\end{equation}
+%
+where ${\bf L}^A$ is the CMO\hyp{}LMO transformation matrix for occupied orbitals of $A$. 
+The total CT energy is given by the sum of the above contribution and the twin contribution
+due to CT from $B$ to $A$.
+
+%
 %he above formula does not explicitly define an OEP due to the summation over different
@@
 \section{\label{s:6}Summary and a few concluding remarks}
+
+Bla.
 

### 2019-05-27 oepdev/liboep/oep.h
DF-based OEP type implemented for CT energy calculations

@@ -294,9 +294,9 @@ class ChargeTransferEnergyOEPotential : public OEPotential
     void common_init();
 
     /// Auxiliary computers
-    void compute_otto_ladik_v1();
-    void compute_otto_ladik_v2();
-    void compute_otto_ladik_v3();
+    void compute_murrell_etal_v1();
+    void compute_murrell_etal_v2();
+    void compute_murrell_etal_v3();
 };
 
 

### 2019-05-27 oepdev/liboep/oep_energy_ct.cc

@@ -1,12 +1,14 @@
 #include "oep.h"
 #include "oep_gdf.h"
 #include "../lib3d/esp.h"
+#include "../lib3d/dmtp.h"
 #include "../libutil/integrals_iter.h"
 #include "psi4/libqt/qt.h"
 
 using namespace oepdev;
 
 using SharedField3D = std::shared_ptr<oepdev::Field3D>;
+using SharedDMTPole = std::shared_ptr<oepdev::DMTPole>;
 
 // <============== CT Energy ===============> //
 
@@ -28,6 +30,8 @@ ChargeTransferEnergyOEPotential::ChargeTransferEnergyOEPotential(SharedWavefunct
 ChargeTransferEnergyOEPotential::~ChargeTransferEnergyOEPotential() {}
 void ChargeTransferEnergyOEPotential::common_init() 
 {
+    name_ = "HF Charge-Transfer Energy";
+
     int n1 = primary_->nbf();
     int n2 = auxiliary_->nbf();
     int n3 = wfn_->molecule()->natom();
@@ -36,19 +40,89 @@ void ChargeTransferEnergyOEPotential::common_init()
     SharedMatrix mat_2 = std::make_shared<psi::Matrix>("Q1", n3, n1);
     SharedMatrix mat_3 = std::make_shared<psi::Matrix>("Q2", n3, 1);
 
-    OEPType type_1 = {"Otto-Ladik.V1", true , n1, mat_1};
-    OEPType type_2 = {"Otto-Ladik.V2", false, n1, mat_2};
-    OEPType type_3 = {"Otto-Ladik.V3", false,  1, mat_3};
+    OEPType type_1 = {"Murrell-etal.V1", true , n1, mat_1};
+    OEPType type_2 = {"Murrell-etal.V2", false, n1, mat_2};
+    OEPType type_3 = {"Murrell-etal.V3", false,  1, mat_3};
 
     oepTypes_[type_1.name] = type_1;
     oepTypes_[type_2.name] = type_2;
     oepTypes_[type_3.name] = type_3;
 }
 
-void ChargeTransferEnergyOEPotential::compute(const std::string& oepType) {}
+void ChargeTransferEnergyOEPotential::compute(const std::string& oepType) 
+{
+  if (oepType == "Murrell-etal.V1"  ) compute_murrell_etal_v1() ;
+  else throw psi::PSIEXCEPTION("OEPDEV: Error. Incorrect OEP type specified!\n");
+}
[…]

### 2019-05-29 papers/2_oep_technique/paper-r0.tex
Notation of 1-1 ERI unifie
General theory chapter elaborated. Notation revised.

@@
 
-\section{\label{s:2}Effective One-Electron Potentials: General Form}
+\section{\label{s:2}Effective One-Electron Potentials: General Theory}
 
@@
 %
-where ${\bf r}$ is a spatial coordinate. Formally, the effective density can be expanded in terms of an effective
+where ${\bf r}$ is a spatial coordinate. 
+For convenience, the total effective density can be split into nuclear and electronic contributions,
+%
+\begin{equation} \label{e:rho-eff}
+ \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
+\end{equation}
+%
+where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
+Formally, the electronic part can be expanded in terms of an effective
 one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
@@
 \begin{equation} \label{e:d-spectral}
-	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
+	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
 	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
@@
 \begin{equation} \label{e:oep-operator}
-	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
+	\hat{v}^{\rm eff} = 
+        \lambda \hat{v}_{\rm nuc} +
+        \sum_{\alpha\beta} 
 	\Ket{\alpha} 
@@
 %
-with the Coulomb operator defined by
-%
-\begin{equation}
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
+with the nuclear
+and Coulomb operators defined by
+%
+\begin{subequations}
+\begin{align}
+        \hat{v}_{\rm nuc} &\equiv \\
+	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
 	f({\bf r}) 
@@
 	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-\end{equation}
+\end{align}
+\end{subequations}
 %
@@
 	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
-	= \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
-\end{equation}
-%
+	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
+        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;,
+\end{equation}
+%
+where 
+%
+\begin{equation}
+ W_{\alpha\beta}^{(x)} = 
+ Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
+\end{equation}
+%
+$Z_x$ is the atomic number of the $x$th atom,
+and the symbol `At' denotes all atoms that contribute to the effective potential.
 In the above equation and throughout the work, 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:gdf-coml.v}
 % {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
@@
 \begin{equation}
- a^{(i)}_\eta = \sum_{x} W_{\eta i}^{(x)} + 
+ a^{(i)}_\eta = \lambda \sum_{x} W_{\eta i}^{(x)} + 
  \sum_{\alpha\beta} D^A_{\alpha\beta} 
@@
 % {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
-  V^A_{i\xi} = \sum_\eta^{\rm DF} [{\bf A}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
+  V^A_{i\xi} = \sum_\eta^{\rm DF} [{\bf R}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
 \end{equation}
@@
                  d{\bf r}_1 d{\bf r}_2 \;, \\
- A_{\eta\xi}  &= \iint 
+ R_{\eta\xi}  &= \iint 
                        \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
@@
 %and $ {\bf r}_2 $. 
-Note that, while $A_{\eta\xi}$ is a typical 2\hyp{}center ERI 
+Note that, while $R_{\eta\xi}$ is a typical 2\hyp{}center ERI 
 that can be evaluated by standard means,
@@
 \begin{equation}
- b^{(i)}_\eta = \sum_\varepsilon^{\rm Int} H_{i\varepsilon} R_{\varepsilon\eta} 
-\end{equation}
-%
-with
-%
-\begin{equation}
- R_{\varepsilon\eta} \equiv \iint 
-                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
-                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
-                 d{\bf r}_1 d{\bf r}_2 \;,
-\end{equation}
-%
-which can be easily evaluated and the matrix elements of the OEP operator found
-from Eq.~\eqref{e:v-noncompl}.
+ b^{(i)}_\eta = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} R_{\varepsilon\eta} \;,
+\end{equation}
+%
+%with
+%%
+%\begin{equation}
+% R_{\varepsilon\eta} \equiv \iint 
+%                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
+%                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
+%                 d{\bf r}_1 d{\bf r}_2 \;,
+%\end{equation}
+%%
+which can be easily evaluated by noting that
+$H_{i\varepsilon}$ are given by Eq.~\eqref{e:gdf-coml.v}.
+Thus, the matrix elements of the OEP operator can be found
+by using Eq.~\eqref{e:v-noncompl}.
 %Therefore, in order to use this generalized density fitting scheme
@@
 (iii) $\BraKet{BB}{AA}$ -- the two last terms. 
+Note also that there are no exchange\hyp{}like terms needed in this case.
+Therefore, all the contributions can be re\hyp{}cast in terms of the OEP's.
 %Note that groups (ii) and (iii) correspond to the Coulomb interactions between
@@
 \begin{equation}
- V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf A}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
+ V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf R}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
 \end{equation}
@@
           \sum_{\varepsilon\in B}^{\rm RI}
-         \left[ {\bf A}^{-1} \right]_{\xi\eta}
-         H_{n\varepsilon}^B R_{\varepsilon \eta} \;,
+         \left[ {\bf R}^{-1} \right]_{\xi\eta}
+         R_{\eta \varepsilon} 
+         H_{n\varepsilon}^B \;,
 \end{equation}
@@
 \begin{equation}
- H_{n\varepsilon}^B = \sum_{\eta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\eta}
-   a_\eta^{(n)}
+ H_{n\varepsilon}^B = \sum_{\zeta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\zeta}
+   a_\zeta^{(n)}
 \end{equation}
@@
 \begin{align}
- a^{(n)}_\eta &= \tBraKet{\eta}{\hat{v}^B_{\rm tot}}{n}
-      - \sum_{j\in B}^{\rm Occ} \tBraKet{\eta}{\hat{v}^B_{nj}}{j} \nonumber \\
- &= \sum_{y\in B} W^{(y)}_{\eta n} 
+ a^{(n)}_\zeta &= \tBraKet{\zeta}{\hat{v}^B_{\rm tot}}{n}
+      - \sum_{j\in B}^{\rm Occ} \tBraKet{\zeta}{\hat{v}^B_{nj}}{j} \nonumber \\
+ &= \sum_{y\in B} W^{(y)}_{\zeta n} 
   + \sum_{j\in B}^{\rm Occ} 
   \left\{
-   2\BraKet{\eta n}{jj} - \BraKet{\eta j}{nj} 
+   2\BraKet{\zeta n}{jj} - \BraKet{\zeta j}{nj} 
   \right\} \;.
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ct.group-i.final}
         \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ct.group-ii.final}
       - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ct.group-iii.oep}
  {v}^{A,{\rm eff}}_{ik}({\bf r}) \equiv
@@
 %
-\begin{multline}
+\begin{multline} \label{e:ct.group-iii.final.a}
         - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
@@
 %
-The above result is a crude approximation, since virtual MO's can overlap with occupied MO's to some extent.
-In order to maintain $\rho^{B}_{nj}$ density, it is approximately represented here by a set of effective 
+The above result is a crude approximation, since the channels $j\rightarrow n$
+are quite important contributions in most of perturbation theory treatments.
+%virtual MO's can penetrate the occupied MO's to some extent.
+In order to include the $\rho^{B}_{nj}$ density, it is approximately represented here by a set of effective 
 cumulative
@@
 In this work, the effective charges were computed via the CAMM or ESP methods as
-discussed in Section~XXX. Applying the SOA leads to 
-%
-\begin{multline}
+discussed in Section~XXX. Applying the SOA 
+and the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
+leads to 
+%
+\begin{multline} \label{e:ct.group-iii.final.b}
         - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
@@
 To compute the interaction energy due to CT from $A$ to $B$ by using the above SOA\hyp{}based
-approximations, coupling elements need to be transformed from localized to canonical MO basis.
-The final working formula reads
-%
-\begin{equation} \label{e:ect-oep.final}
+approximations, coupling elements need to be transformed from localized to canonical MO basis, i.e.,
+%
+\begin{equation} \label{e:ect-oep.prefinal}
  E^{A\rightarrow B} \approx
@@
 where ${\bf L}^A$ is the CMO\hyp{}LMO transformation matrix for occupied orbitals of $A$. 
+Note however, that the effective potential parameters from the density fitted term of group (i) 
+do not involve any occupied orbitals.
+Therefore, to save computational costs, only groups (ii) and (iii) need to be transformed
+whereas for group (i) the overlap integrals from Eq.~\eqref{e:ct.group-i.final} 
+can be already evaluated in CMO basis.
+Thus, the final working formula for the interaction energy due to CT from $A$ to $B$ reads
+%
+\begin{multline} \label{e:ect-oep.final}
+ E^{A\rightarrow B} \approx
+ 2 
+ \sum_{i\in A}^{\rm CMO}
+ \sum_{n\in B}^{\rm Vir}
+ \frac{1}{\varepsilon_i - \varepsilon_n} \times 
+ \Bigg(
+   \Approx{V_{in}^{A\rightarrow B}}{\rm G1} \\
+  +
+   \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
+   \left\{
+   \Approx{V_{i'n}^{A\rightarrow B}}{\rm G2} 
+  +\Approx{V_{i'n}^{A\rightarrow B}}{\rm G3}
+  \right\}
+ \Bigg)^2 \;,
+\end{multline}
+%
+where the subscripts G$n$ ($n$ = 1, 2, 3) denote a particular group of 
+terms from Eqs.~\eqref{e:ct.group-i.final}, \eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.a}
+or \eqref{e:ct.group-iii.final.b}. 
 The total CT energy is given by the sum of the above contribution and the twin contribution
-due to CT from $B$ to $A$.
+due to CT from molecule $B$ to $A$.
 

### 2019-05-29 oepdev/liboep/oep.h
Localization of orbitals moved from OEPDevSolver to OEPotential objects to maintain orbital consistency

@@ -13,6 +13,7 @@
 #include "psi4/libmints/basisset.h"
 #include "psi4/libmints/matrix.h"
 #include "psi4/libmints/vector.h"
+#include "psi4/libmints/local.h"
 #include "../libpsi/integral.h"
 #include "../libpsi/potential.h"
 #include "../lib3d/space3d.h"
@@ -27,6 +28,7 @@ using SharedBasisSet     = std::shared_ptr<BasisSet>;
 using SharedMatrix       = std::shared_ptr<Matrix>;
 using SharedVector       = std::shared_ptr<Vector>;
 using SharedDMTPole      = std::shared_ptr<DMTPole>;
+using SharedLocalizer    = std::shared_ptr<Localizer>;
 /** \addtogroup OEPDEV_OEPS
  * @{
  */
@@ -72,6 +74,8 @@ class OEPotential : public std::enable_shared_from_this<OEPotential>
     SharedBasisSet auxiliary_;
     /// Intermediate Basis set
     SharedBasisSet intermediate_;
+    /// Molecular Orbital Localizer
+    SharedLocalizer localizer_;
 
     /// Name of this OEP;
     std::string name_;
@@ -86,8 +90,10 @@ class OEPotential : public std::enable_shared_from_this<OEPotential>
     std::shared_ptr<psi::OneBodyAOInt> OEInt_;
     /// One-electron potential shared pointer
     std::shared_ptr<oepdev::PotentialInt> potInt_;
-    /// Occupied orbitals
+    /// Occupied orbitals: Canonical (CMO)
     std::shared_ptr<psi::Matrix> cOcc_;
+    /// Occupied orbitals: Localized (LMO)
+    std::shared_ptr<psi::Matrix> lOcc_;
     /// Virtual orbitals
     std::shared_ptr<psi::Matrix> cVir_;
 
@@ -159,6 +165,9 @@ class OEPotential : public std::enable_shared_from_this<OEPotential>
     /// Write potential to a cube file
     virtual void write_cube(const std::string& oepType, const std::string& fileName);
 
+    /// Localize Occupied MO's
+    virtual void localize(void);
+
     /// Rotate 
     virtual void rotate(const Matrix& rotmat);
     /// Translate
@@ -185,6 +194,9 @@ class OEPotential : public std::enable_shared_from_this<OEPotential>
 
     /// Retrieve wavefunction object
     SharedWavefunction wfn() const {return wfn_;}
+
[…]

### 2019-05-29 oepdev/liboep/oep_base.cc

@@ -28,6 +28,15 @@ void OEPotential::common_init(void)
    potInt_      = std::make_shared<oepdev::PotentialInt>(intsFactory_->spherical_transform(), primary_, primary_, 0);
    cOcc_        = wfn_->Ca_subset("AO","OCC");
    cVir_        = wfn_->Ca_subset("AO","VIR");
+   lOcc_        = nullptr; //std::make_shared<psi::Matrix>();
+   localizer_   = nullptr; //psi::Localizer::build(options_.get_str("SOLVER_CT_LOCALIZER"), primary_, cOcc_, options_);
+}
+void OEPotential::localize(void) 
+{
+   std::string o_loc = options_.get_str("SOLVER_CT_LOCALIZER");
+   localizer_ = psi::Localizer::build(o_loc, primary_, cOcc_, options_);
+   localizer_->localize();
+   lOcc_ = localizer_->L();
 }
 std::shared_ptr<OEPotential> OEPotential::build(const std::string& category, SharedWavefunction wfn, Options& options)
 {

### 2019-05-29 oepdev/libsolver/solver_energy_ct.cc

@@ -2,13 +2,10 @@
 //#include "psi4/libdpd/dpd.h"
 
 #include "solver.h"
-#include "psi4/libmints/local.h"
 
 using namespace std;
 using namespace psi;
 using namespace oepdev;
-
-using SharedLocalizer       = std::shared_ptr<psi::Localizer>;
 
 // CT Solver//
 ChargeTransferEnergySolver::ChargeTransferEnergySolver(SharedWavefunctionUnion wfn_union)
@@ -633,7 +630,7 @@ double ChargeTransferEnergySolver::compute_benchmark_otto_ladik(){
 double ChargeTransferEnergySolver::compute_benchmark_efp2(){}
 double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
 {
-  // ===> Compute OEP Objects <=== //
+  // ===> Set-up OEP Objects <=== //
   SharedOEPotential oep_1 = oepdev::OEPotential::build("CHARGE TRANSFER ENERGY",
                                                        wfn_union_->l_wfn(0), 
                                                        wfn_union_->l_auxiliary(0), 
@@ -646,6 +643,10 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
                                                        wfn_union_->options());
   oep_1->compute("Murrell-etal.V1.GDF");
   oep_2->compute("Murrell-etal.V1.GDF");
+  oep_1->localize();
+  oep_2->localize();
+  oep_1->compute("Murrell-etal.V3.CAMM-nj");
+  oep_2->compute("Murrell-etal.V3.CAMM-nj");
 
   // ===> Compute Overlap Matrices <=== //
   int nbf_p1 = wfn_union_->l_nbf(0);
@@ -695,13 +696,8 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
   std::shared_ptr<psi::Matrix> S2 = psi::Matrix::doublet(Ca_occ_2, Sao_1a2p, true, true ); // OCC(B) x AUX(A)
 
   // ---> Localize occupied orbitals <--- //
-  std::string o_loc = wfn_union_->options().get_str("SOLVER_CT_LOCALIZER"); 
[…]

### 2019-05-29 oepdev/liboep/oep_energy_ct.cc
Revision of the group III code

@@ -135,7 +135,7 @@ void ChargeTransferEnergyOEPotential::compute_murrell_etal_v3_camm_nj()
 
   for (int j=0; j<nocc; ++j) {
        for (int n=0; n<nvir; ++n) {
-            int idx = nocc*j + n;
+          //int idx = nvir*j + n;
 
             /* Exclude nuclear part */
             trans.push_back(true);
@@ -146,6 +146,7 @@ void ChargeTransferEnergyOEPotential::compute_murrell_etal_v3_camm_nj()
             double** oed_p = oed->pointer();
             for (int a=0; a<nbf; ++a) {
             for (int b=0; b<nbf; ++b) {
+               //oed_p[a][b] =(lOcc_->get(a,j) * cVir_->get(b,n) + lOcc_->get(b,j) * cVir_->get(a,n)) / 2.0;
                  oed_p[a][b] = lOcc_->get(a,j) * cVir_->get(b,n);
             }}
             oeds.push_back(oed);

### 2019-05-29 oepdev/libsolver/solver_energy_ct.cc

@@ -746,7 +746,7 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
             double v = 0;
             for (int j=0; j<nocc_2; ++j) {              
                  for (int y=0; y<mol_2->natom(); ++y) {
-                      v += S12->get(i,j) * w_1->get(i, y) * q_2[nocc_2*j+n]->get(y, 0);
+                      v += S12->get(i,j) * w_1->get(i, y) * q_2[nvir_2*j+n]->get(y, 0);
                  }
             }
             v_ab_v3->set(i, n, v);
@@ -758,7 +758,7 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
             double v = 0;
             for (int j=0; j<nocc_1; ++j) {              
                  for (int y=0; y<mol_1->natom(); ++y) {
-                      v += S12->get(j,i) * w_2->get(i, y) * q_1[nocc_1*j+n]->get(y, 0);
+                      v += S12->get(j,i) * w_2->get(i, y) * q_1[nvir_1*j+n]->get(y, 0);
                  }
             }
             v_ba_v3->set(i, n, v);
@@ -922,13 +922,11 @@ std::shared_ptr<psi::Matrix> ChargeTransferEnergySolver::compute_w_matrix(
        }
        //
        for (int i=0; i<nocc_1; ++i) {
-            double v = vy;
             //
             double riy = sqrt(pow(rmo_1[0]->get(i) - mol_2->x(y), 2.0) +
                               pow(rmo_1[1]->get(i) - mol_2->y(y), 2.0) +
                               pow(rmo_1[2]->get(i) - mol_2->z(y), 2.0) );
-            v += 2.0 / riy;
-            w->set(i, y, v);
+            w->set(i, y, vy + 2.0/riy );
        }
   }
   w->scale(-1.0);

### 2019-05-29 papers/2_oep_technique/paper-r0.tex
Section about computational costs drafted

@@
  \left[
-  v^A_{\rm tot} ({\bf r}) - 2 v^A_{k} ({\bf r}) + v^A_{ik} ({\bf r})
+  v^A_{\rm tot} ({\bf r}) - 2 v^A_{kk} ({\bf r}) + v^A_{ik} ({\bf r})
  \right] 
@@
 
+\subsection{\label{s:413s6}Reduction of computational costs}
+
+Until this moment in this work, two models have been tested with application of the OEP technique.
+In this section, computational cost is estimated from the formulas.
+In
+Table~\ref{t:oep-costs} 
+%
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[b]
+\caption[Computational cost of the benchmark and OEP-based methods: calculation of coupling constant]
+{{\bf Computational cost of the benchmark and OEP-based methods\footnotemark[1]}
+}
+\label{t:oep-costs}
+\begin{ruledtabular}
+\begin{tabular}{lcccccc}
+Method             && Murrell et al. &&    EFP2        &&      OEP     \\ 
+	\cline{1-7}
+Group (i)         &&  &&        &&   $obQ$     \\
+Group (ii)        &&  &&        &&   $ob$      \\
+Group (iii)       &&  &&        &&   $ob(n+o)$     \\
+Total Cost        &&  && $10ob^2 + 2b^3 + o^2b^2$       &&   $ob(Q+n+o)$     \\
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
+$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
+%
+\end{table}
+}
+%
+It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
+and many orders of magnitude lesser than the original theory of Murrell et al.
+
 \section{\label{s:6}Summary and a few concluding remarks}

### 2019-06-28 papers/conference/1.MQM_2019/poster-80x120cm.tex
Poster draft for MQM2019

\RequirePackage[]{paralist}
\documentclass[final]{beamer} % beamer 3.10: do NOT use option hyperref={pdfpagelabels=false} !
%
% IBS 2016 conference - POSTER
%
%                 Gundelfingen, 24.06.2019, Germany
%
%\documentclass[final,hyperref={pdfpagelabels=false}]{beamer} % beamer 3.07: get rid of beamer warnings
\mode<presentation> {  %% check http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php for examples
  \usetheme{Berlin}    %% you should define your own theme e.g. for big headlines using your own logos 
}
\mode<presentation>{\usetheme{I6pd2}}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{amsmath, amsthm, amssymb, latexsym, upgreek, amsbsy, mathrsfs}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{booktabs}
\usepackage{caption}

\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usefonttheme[onlymath]{serif}
\definecolor{orange}{RGB}{255,127,0}
\definecolor{yellowish}{RGB}{255,255,153}

\newcommand{\BM}[1]{\ensuremath{\mbox{\boldmath${#1}$}}}
\newcommand{\exedout}{%
  \rule{0.8\textwidth}{0.5\textwidth}%
}
\newcommand{\BraKet}[2]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigr\rangle}}
\newcommand{\tBraKet}[3]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigl\lvert {#3} \bigr\rangle}}

% ====================================================================================
\graphicspath{{figures/}}

\usepackage[orientation=portrait,scale=1.4,debug]{beamerposter}  
%
\title{\huge Fast Evaluation of Intermolecular Charge Transfer Energy: 
New Approach Based on One-Electron Effective Potentials
}
\author{\underline{Bartosz B{\l}asiak}, Joanna D. Bednarska, Marta Cho{\l}uj, Wojciech Bartkowiak}
\institute{
Department of Physical and Quantum Chemistry, Wroc{\l}aw University of Technology, Wroc{\l}aw 50-370, Poland}
\date[July. 4th, 2019]{July. 4th, 2019}

\setlength{\paperwidth}{84.0cm}
\setlength{\paperheight}{118.0cm}

% ==========================================================================================================================
\begin{document}

\begin{frame}{OEP-Based Charge Transfer (CT) Interaction Energy}

\begin{textblock*}{440mm}(10mm,135mm)
\begin{block}{Ab Initio Charge Transfer Energy in Fragment-Based Modelling}
 \begin{itemize}
   \item CT energy is an {\color{green} important} stabilizing contribution to the total interaction energy
   \item It is {\color{red} very costly} to compute and generally not directly included in most force fields
   \item {\color{green} EFP2 model} implements efficient CT energy estomator
   \item However, EFP2 CT energy is {\color{red} 30-40 times slower} then evaluating other components
   \item In effect, this CT term is usually {\color{red} not included} in condensed matter modelling
 \end{itemize}
\end{block}
\end{textblock*}

\begin{textblock*}{250mm}(10mm,253mm)
\begin{block}{Our Approach}
 \begin{itemize}
   \item Consider Murrell et al.'s perturbation theory
   \item Apply OEP method to eliminate ERI's
   \item Adapt the CT term to the EFP2 method
 \end{itemize}
\end{block}
\end{textblock*}

\begin{textblock*}{180mm}(270mm,253mm)
\begin{block}{CT energy by Murrell et al.}
  \vskip-2ex
  \begin{equation*}
    \Delta E_{\rm CT}^{A\rightarrow B} = 2\sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir}
     \frac{\vert V_{in}^{A\rightarrow B} \vert^2 }{\varepsilon_i - \varepsilon_n}
  \end{equation*}
  \vskip0.4ex
\end{block}
\end{textblock*}

\begin{textblock*}{367mm}(463mm,137mm)
\begin{exampleblock}{One-Electron Effective Potential (OEP) Operators}
OEP's can be introduced to eliminate electron repulsion integrals from parent theories:
\begin{equation*}
 \sum_t \sum_{ij} \sum_{kl\in A} \mathscr{F}_t 
 \left[ 
  \left< ij \vert k^A l^A \right>
 \right] 
  = \left<  
      i \vert \hat{v}^A_{\rm eff} \vert j
    \right>
\end{equation*}
\end{exampleblock}
\end{textblock*}

\begin{textblock*}{367mm}(463mm,233mm)
\begin{exampleblock}{Types of matrix elements of OEP operators}
\vskip0.1ex
\begin{table}[b]
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}_{\rm eff}^{A}}{j} $ && $\tBraKet{j}{\hat{v}_{\rm eff}^{A}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
Handling            && Density Fitting                          && Distributed Multipoles                  \\
\end{tabular}
%
\end{table}
\vskip-0.1ex
\end{exampleblock}
\end{textblock*}

% - Graph 1
\begin{textblock*}{340mm}(10mm,802mm)
\begin{figure}
\includegraphics[width=1.0\textwidth]{fig-1-s.eps}
%\caption{\bf OEP-based model reproduces CT energies of the parent Murrell et al.'s [3] model.}
\end{figure}
\end{textblock*}

% - Arrow
\begin{textblock*}{60mm}(192mm,902mm)
%\begin{figure}
\includegraphics[scale=0.4]{arrow.eps}
%\end{figure}
\end{textblock*}

% - Scheme 
\begin{textblock*}{820mm}(10mm,341mm)
\begin{exampleblock}{Coupling Constant $\boldsymbol{ V_{in}^{A\rightarrow B} } $}
\begin{figure}
\includegraphics[width=1.0\textwidth]{scheme.eps}
\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al. EFP parameters
are shown in blue.}
\end{figure}
\end{exampleblock}
\end{textblock*}

% ----- RULE DIVIDING THEORY AND RESULTS
\begin{textblock*}{820mm}(6mm,820mm)
\noindent\rule{83cm}{5.4pt}
\end{textblock*}

% ----- RULE DIVIDING RESULTS AND ACKNOWLEDGEMENTS
\begin{textblock*}{820mm}(6mm,1070mm)
\noindent\rule{83cm}{5.4pt}
\end{textblock*}

%
%
%% ----------------------- SUMMARY
%%
%
% --------------------- ACKNOWLEDGEMENTS
\begin{textblock*}{820mm}(10mm,1070mm)
\begin{block}{Acknowledgements}
 This project is carried out under POLONEZ programme which has received funding from the European Union's
 Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
 No.~665778. This project is funded by National Science Centre, Poland 
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
 We thank
 Wroc{\l}aw Centre for Networking and Supercomputing (WCSS) for
 providing high quality computational resources.
\end{block}
\end{textblock*}
%
%%
%
%% --------------------- REFERENCES
%\begin{textblock*}{440mm}(10mm,1106mm)
%\begin{exampleblock}{\color{blue} References}
%    References\\
%    \begin{small}
%[1] Gordon, M. S.; Fedorov. D. G.; Pruitt, S. R.; Slipchenko L. V.; Chem. Rev., 2012, 112, 632
%[2] Murrell, J. N.; Randic, M.; Williams O. R.; Proc. Roy. Soc. A, 1965, 284, 566
%[3] Li, H.; Gordon, M. S.; J. Chem. Phys., 2006, 124, 214108
%              [2] B. B{\l}asiak, M. Maj, R. W. G{\'o}ra, M. Cho, \newline\emph{J. Chem. Theory Comput.}, 11, 3259 {\bf{\small 2015}};
%    \end{small}         
%%\end{exampleblock}
%\end{textblock*}
%
%
%
\end{frame}
\end{document}

### 2019-06-28 papers/conference/1.MQM_2019/data/water_water.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+pymol -p <<EOF
+load water_water.xyz
+set_view (\
+     0.077019170,    0.975163639,   -0.207657859,\
+    -0.356887996,    0.221440658,    0.907520950,\
+     0.930967569,    0.004214230,    0.365079701,\
+     0.000000000,    0.000000000,  -14.178204536,\
+    -0.027711868,   -0.007276356,   -0.205429554,\
+    11.178204536,   17.178203583,  -20.000000000 )
+set stick_radius, 0.06
+set sphere_scale, 0.15
+show sticks
+show spheres
+set transparency, 0.9
+set opaque_background, off
+set ray_opaque_background, off
+ray 900, 900
+png water_water, dpi=600
+EOF
+

### 2019-06-28 papers/conference/1.MQM_2019/do.sh

@@ -1,0 +1,2 @@
+latex  poster-80x120cm.tex
+dvipdf poster-80x120cm.dvi

### 2019-06-28 oepdev/libsolver/solver_energy_ct.cc
Poster uptdated

@@ -1094,7 +1094,7 @@ double ChargeTransferEnergySolver::compute_benchmark_efp2()
   // ---> Timer-off <--- //
   psi::timer_off("Solver E(CT) EFP2 MO-Expanded");
   t_time += clock(); // Clock END
-  cout << " o TIME EFP2: " << t_time << endl;
+  cout << " o TIME EFP2: " << ((double)t_time/CLOCKS_PER_SEC) << endl;
     
   // ---> Print results <--- //
   if (wfn_union_->options().get_int("PRINT") > 0) 
@@ -1290,7 +1290,7 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
   double e_tot = e_ab + e_ba;
   psi::Process::environment.globals["ECT OEP KCAL"] = e_tot*627.509;
 
-  cout << " o TIME OEP : " << t_time << endl;
+  cout << " o TIME OEP : " << ((double)t_time/CLOCKS_PER_SEC) << endl;
   //cout << " Time OEP : " << t_end - t_begin - (t2 - t1) << endl;
 
   // ---> Print <--- //

### 2019-06-28 papers/conference/1.MQM_2019/poster-84x118cm.tex
Poster revision

--- 2019-06-28 papers/conference/1.MQM_2019/poster-80x120cm.tex
@@
  \begin{itemize}
-   \item CT energy is an {\color{green} important} stabilizing contribution to the total interaction energy
+   \item CT energy is an {\color{green} important} stabilizing contribution to the total interaction energy [1]
    \item It is {\color{red} very costly} to compute and generally not directly included in most force fields
-   \item {\color{green} EFP2 model} implements efficient CT energy estomator
-   \item However, EFP2 CT energy is {\color{red} 30-40 times slower} then evaluating other components
-   \item In effect, this CT term is usually {\color{red} not included} in condensed matter modelling
+   \item {\color{green} EFP2 model} implements efficient CT energy estomator [2,3]
+   \item However, EFP2 CT energy is {\color{red} 30-40 times slower} then evaluating other components [3]
+   \item In effect, this CT term is usually {\color{red} not included} in condensed matter modelling [4]
  \end{itemize}
@@
  \begin{itemize}
-   \item Consider Murrell et al.'s perturbation theory
+   \item Consider Murrell et al.'s perturbation theory [5]
    \item Apply OEP method to eliminate ERI's
-   \item Adapt the CT term to the EFP2 method
+   \item Adapt the CT term to the EFP2 method [2,3]
  \end{itemize}
@@
 \begin{textblock*}{180mm}(270mm,253mm)
-\begin{block}{CT energy by Murrell et al.}
+\begin{block}{CT energy by Murrell et al. [5]}
   \vskip-2ex
@@
 % - Graph 1
-\begin{textblock*}{340mm}(10mm,802mm)
+\begin{textblock*}{340mm}(10mm,807mm)
 \begin{figure}
@@
 % - Arrow
-\begin{textblock*}{60mm}(192mm,902mm)
-%\begin{figure}
-\includegraphics[scale=0.4]{arrow.eps}
-%\end{figure}
+\begin{textblock*}{60mm}(220mm,903mm)
+\includegraphics[scale=0.37]{arrow.eps}
+\end{textblock*}
+
+% - Graph 2
+\begin{textblock*}{340mm}(290mm,820mm)
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{fig-2.eps}
+%\caption{\bf OEP-based model reproduces CT energies of the parent Murrell et al.'s [3] model.}
+\end{figure}
 \end{textblock*}
@@
 \includegraphics[width=1.0\textwidth]{scheme.eps}
-\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al. EFP parameters
+\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al.[5] EFP parameters
 are shown in blue.}
@@
 \begin{textblock*}{820mm}(6mm,820mm)
+\noindent\rule{83cm}{5.4pt}
+\end{textblock*}
+
+\begin{textblock*}{257mm}(573mm,824mm)
+\begin{block}{Summary and Outlook}
+ \begin{itemize}
+   \item OEP-based model reproduces CT energies of the parent Murrell et al.'s [5] model
+   \item OEP-based model is 4-7 times faster than EFP2
+   \item Further optimization is possible: using QUAMBO orbitals [4]
+   \item OEP's can be applied to other problems as well
+ \end{itemize}
+\end{block}
+\end{textblock*}
+
+\begin{textblock*}{257mm}(573mm,970mm)
+%\begin{exampleblock}{\color{blue} References}
+    References: %\\
+    \begin{small}
+%[1] Gordon, M. S.; Fedorov. D. G.; Pruitt, S. R.; Slipchenko L. V.; \emph{Chem. Rev.} {\bf 2012}, 112, 632
+[1] Gordon, M. S. et al.; \emph{Chem. Rev.} {\bf 2012}, 112, 632 
+[2] Li, H.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf 2006}, 124, 214108
+[3] Xu, P.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf 2013}, 139, 194104
+[4] Kaliman, I. A.; Slipchenko, L. V.; \emph{J. Comput. Chem.} {\bf 2015}, 36, 129
+[5] Murrell, J. N.; Randic, M.; Williams O. R.; \emph{Proc. Roy. Soc. A} {\bf 1965}, 284, 566
+    \end{small}         
+%\end{exampleblock}
+\end{textblock*}
+\begin{textblock*}{257mm}(573mm,964mm)
 \noindent\rule{83cm}{5.4pt}
@@
 \end{textblock*}
-%
-%%
-%
-%% --------------------- REFERENCES
-%\begin{textblock*}{440mm}(10mm,1106mm)
-%\begin{exampleblock}{\color{blue} References}
-%    References\\
-%    \begin{small}
-%[1] Gordon, M. S.; Fedorov. D. G.; Pruitt, S. R.; Slipchenko L. V.; Chem. Rev., 2012, 112, 632
-%[2] Murrell, J. N.; Randic, M.; Williams O. R.; Proc. Roy. Soc. A, 1965, 284, 566
-%[3] Li, H.; Gordon, M. S.; J. Chem. Phys., 2006, 124, 214108
-%              [2] B. B{\l}asiak, M. Maj, R. W. G{\'o}ra, M. Cho, \newline\emph{J. Chem. Theory Comput.}, 11, 3259 {\bf{\small 2015}};
-%    \end{small}         
-%%\end{exampleblock}
-%\end{textblock*}
-%
-%
-%
+
 \end{frame}

### 2019-06-28 papers/conference/1.MQM_2019/do.sh

@@ -1,2 +1,2 @@
-latex  poster-80x120cm.tex
-dvipdf poster-80x120cm.dvi
+latex  poster-84x118cm.tex
+dvipdf poster-84x118cm.dvi

### 2019-06-28 papers/conference/1.MQM_2019/poster-84x118cm.tex
Revision of Poster

@@
 
+% --- Motivation
+
 \begin{textblock*}{440mm}(10mm,135mm)
@@
    \item It is {\color{red} very costly} to compute and generally not directly included in most force fields
-   \item {\color{green} EFP2 model} implements efficient CT energy estomator [2,3]
+   \item {\color{green} EFP2 model} implements efficient CT energy estimator [2,3]
    \item However, EFP2 CT energy is {\color{red} 30-40 times slower} then evaluating other components [3]
@@
  \end{itemize}
+ \vskip0.4ex
 \end{block}
@@
 
-\begin{textblock*}{250mm}(10mm,253mm)
+% --- Our method
+
+\begin{textblock*}{250mm}(10mm,261mm)
 \begin{block}{Our Approach}
@@
 
-\begin{textblock*}{180mm}(270mm,253mm)
+\begin{textblock*}{180mm}(270mm,261mm)
 \begin{block}{CT energy by Murrell et al. [5]}
@@
 \begin{table}[b]
+\def\arraystretch{1.2}
 \begin{tabular}{lcccc}
@@
 % - Scheme 
-\begin{textblock*}{820mm}(10mm,341mm)
+\begin{textblock*}{820mm}(10mm,355mm)
 \begin{exampleblock}{Coupling Constant $\boldsymbol{ V_{in}^{A\rightarrow B} } $}
 \begin{figure}
-\includegraphics[width=1.0\textwidth]{scheme.eps}
-\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al.[5] EFP parameters
-are shown in blue.}
+\includegraphics[width=1.0\textwidth]{scheme-r1.eps}
+\vskip0.9ex
+\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al. [5] 
+             EFP parameters are shown in blue.}
 \end{figure}
@@
 \end{textblock*}
+
+% --- Summary and References
 
@@
 \begin{textblock*}{257mm}(573mm,970mm)
-%\begin{exampleblock}{\color{blue} References}
     References: %\\
@@
 %[1] Gordon, M. S.; Fedorov. D. G.; Pruitt, S. R.; Slipchenko L. V.; \emph{Chem. Rev.} {\bf 2012}, 112, 632
-[1] Gordon, M. S. et al.; \emph{Chem. Rev.} {\bf 2012}, 112, 632 
-[2] Li, H.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf 2006}, 124, 214108
-[3] Xu, P.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf 2013}, 139, 194104
-[4] Kaliman, I. A.; Slipchenko, L. V.; \emph{J. Comput. Chem.} {\bf 2015}, 36, 129
-[5] Murrell, J. N.; Randic, M.; Williams O. R.; \emph{Proc. Roy. Soc. A} {\bf 1965}, 284, 566
+[1] Gordon, M. S. et al.; \emph{Chem. Rev.} {\bf{\small 2012}}, 112, 632 
+[2] Li, H.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf{\small 2006}}, 124, 214108
+[3] Xu, P.; Gordon, M. S.; \emph{J. Chem. Phys.} {\bf{\small 2013}}, 139, 194104
+[4] Kaliman, I. A.; Slipchenko, L. V.; \emph{J. Comput. Chem.} {\bf{\small 2015}}, 36, 129
+[5] Murrell, J. N.; Randic, M.; Williams O. R.; \emph{Proc. Roy. Soc. A} {\bf{\small 1965}}, 284, 566
     \end{small}         
-%\end{exampleblock}
 \end{textblock*}
+% ----- RULE DIVIDING RESULTS AND REFERENCES
 \begin{textblock*}{257mm}(573mm,964mm)
@@
 
-%
-%
-%% ----------------------- SUMMARY
-%%
-%
 % --------------------- ACKNOWLEDGEMENTS

### 2019-06-29 papers/conference/1.MQM_2019/poster-84x118cm.tex
Some revision

@@
 
-\begin{textblock*}{250mm}(10mm,261mm)
+\begin{textblock*}{249mm}(10mm,261mm)
 \begin{block}{Our Approach}
@@
 \begin{textblock*}{180mm}(270mm,261mm)
-\begin{block}{CT energy by Murrell et al. [5]}
+\begin{block}{CT Energy by Murrell et al. [5]}
   \vskip-2ex
@@
 \begin{textblock*}{367mm}(463mm,233mm)
-\begin{exampleblock}{Types of matrix elements of OEP operators}
+\begin{exampleblock}{Types of Matrix Elements of OEP Operators}
 \vskip0.1ex
@@
 \vskip0.9ex
-\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al. [5] 
-             EFP parameters are shown in blue.}
+\caption{\bf Application of the OEP technique to significantly simplify original theory of Murrell et al. [5] (left). 
+             Fragment parameters are shown in blue.}
 \end{figure}
@@
  \begin{itemize}
-   \item OEP-based model reproduces CT energies of the parent Murrell et al.'s [5] model
-   \item OEP-based model is 4-7 times faster than EFP2
-   \item Further optimization is possible: using QUAMBO orbitals [4]
-   \item OEP's can be applied to other problems as well
+   \item OEP-based model {\color{blue} reproduces} CT energies of the parent Murrell et al.'s [5] model
+   \item OEP-based model is 4-7 times {\color{blue} faster} than EFP2
+   \item Further optimization is {\color{blue} possible}: using QUAMBO orbitals [4]
+   \item OEP's can be applied to {\color{blue} other problems} as well
  \end{itemize}

### 2019-07-08 papers/2_oep_technique/paper-r0.tex
Paper revision. Added section about multipole distribution

@@
 The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
-The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. Mainly,
+The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
 %
@@
 Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
-                    && $\tBraKet{i}{\hat{v}^{{\rm eff},A}}{j} $ && $\tBraKet{j}{\hat{v}^{{\rm eff},A}}{l}$ \\ 
+                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
 	\cline{1-5}
@@
 DMTP\footnotemark[4] Form                     
-&& --  &  &   \\
+&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
 \end{tabular}
@@
 
+This scheme is most applicable in the case of matrix elements of the type
+$
+ \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
+$
+because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
+and $\rho^A_{\rm eff}$. 
+In general, given certain two effective one\hyp{}electron density distributions,
+the associated effective Coulombic interaction energy can be estimated from the classical formula
+according to
+%
+\begin{equation}
+ \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
+ {\vert {\bf r}_1 - {\bf r}_2 \vert} 
+d{\bf r}_1 d{\bf r}_2 \;.
+\end{equation}
+%
+The above double integral can be approximated by the
+distributed multipole (DMTP)
+expansions of the respective densities as follows
+%
+\begin{multline}
+  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
+ \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
+ \sum_{u\in A} \sum_{w\in B} \Big\{ 
+ T^{(0)}_{uw}
+ q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
+ - {\bf T}^{(1)}_{uw} \cdot 
+   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
+ - {\bf T}^{(2)}_{uw} : 
+  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
+ \ldots
+ \Big\}
+\end{multline}
+%
+In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
+are the effective distributed monopole (charge)
+and dipole moment, respectively, centered on the $u$th site, 
+whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
+that can be found elsewhere.
+The symbol `$\odot$' denotes the sum of all the tensor contractions
+performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
+The choice of the distribution centres as well as the truncation order of the multipole expansion
+is crucial in compromising the accuracy and computational cost of the resulting expressions.
+There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
+of Stone and Anderton, %cite
+the cumulative atomic multipole moments of Sokalski and Poirier, %cite
+the localised distributed multipole expansion (LMTP) of Etchtebest et al. %cite
+or schemes based on fitting to electrostatic potential.
+
 \subsection{\label{s.3344}Density Fitting of OEP's}
 
-To get the \emph{ab initio} representation of a overlap\hyp{}like OEP, 
+This scheme is most useful in the case of matrix elements of the type
+$
+ \tBraKet{i^A}{\hat{v}_{\rm eff}^A}{j^B}
+$.
+To get the \emph{ab initio} representation of such an overlap\hyp{}like matrix element,
 one can use a procedure similar to
 the typical density fitting or resolution of identity, both of which are nowadays widely used 
-to compute electron-repulsion integrals (ERI's) more efficiently.
+to compute electron\hyp{}repulsion integrals (ERI's) more efficiently.
 
@@
 
-An arbitrary one-electron potential of molecule $A$ acting on any state vector 
+An arbitrary one\hyp{}electron potential of molecule $A$ acting on any state vector 
 associated with molecule $A$ can be expanded in an auxiliary space centered 
@@
 \subsection{\label{ss.5.1}Pauli-Repulsion Interaction Energy}
+
+At short intermolecular interactions, wavefunction overlap increases
+which causes the electrons from either molecules to move apart to minimize the repulsive interactions.
+The resulting interaction energy can be estimated based on perturbation theory approaches.
+In this section, we demonstrate the use of OEP operators to significanlty reduce the computational
+cost of such calculations, that can be used in fragment\hyp{}based modelling of extended molecular
+aggregates in terms of force field methods.
+
+As a starting point we consider perturbation theory model of Murrell et al.,
+in which repulsive energy is originally given by
+%
+\begin{equation}
+ \Delta^{AB} E^{\rm Rep} = 
+ \Delta^{AB} E^{\rm Rep}(\mathcal{O}(S)) + 
+ \Delta^{AB} E^{\rm Rep}(\mathcal{O}(S^2))
+\end{equation}
+%
+where the first\hyp{}order term with respect to the intermolecular overlap is
+%
+\begin{multline}
+  \Delta^{AB}  E^{\rm Rep}(\mathcal{O}(S)) = -2\sum_{a\in A} \sum_{b\in B}
+               S_{ab} \Big\{
+           V^A_{ab} 
+         + V^B_{ab} \\
+ + \sum_{c\in A} \left[ 2(ab \vert cc) - (ac \vert bc) \right] 
+ + \sum_{d\in B} \left[ 2(ab \vert dd) - (ad \vert bd) \right]
+                \Big\}
+\end{multline}
+%
+and the second\hyp{}order term is given in Eq.(XXX) in Ref. %cite
+In this work, we focus to simplify only the former term because this is usually
+the bottleneck of evaluating the Pauli repulsive contributions to
+forces in molecular dynamics when using the effective potential method, whereas
+very inexpensive approximation to the latter term has already been worked out previously. %cite
+
+As another benchmark model for accuracy, we compare our results with 
+the exchange and repulsion energy term 
+from the density\hyp{}based scheme of partitioning the interaction energy
+(here abbreviated as DDS),
+that was developed by Mandado and Hermida-Ramon. 
+%In the above equations, the Pauli deformation density 
+%%
+%\begin{equation}
+%   \Delta {\bf D}^{\rm Pauli} \equiv {\bf D}^{oo} - {\bf D}
+%\end{equation}
+%%
+%where \f$ {\bf D}^{oo}\f$ and \f$ {\bf D}\f$ are the density matrix formed from
+%mutually orthogonal sets of molecular orbitals within the entire aggregate (formed
+%by symmetric orthogonalization of MO's) and the density matrix of the unperturbed
+%system (that can be understood as a Hadamard sum \f$ {\bf D} \equiv {\bf D}^A \oplus {\bf D}^B\f$).
+% 
+%At HF level, the Pauli deformation density matrix is given by
+%\f[
+%  \Delta {\bf D}^{\rm Pauli} = {\bf C} \left[ {\bf S}^{-1} - {\bf 1} \right] {\bf C}^\dagger
+%\f]
+%whereas the density matrix constructed from mutually orthogonal orbitals is
+%\f[
+%   {\bf D}^{oo} = {\bf C} {\bf S}^{-1} {\bf C}^\dagger
+%\f]
+%In the above equations, \f$ {\bf S} \f$ is the overlap matrix between doubly occupied molecular orbitals
+% of the entire aggregate.
+It is also mentioned here that the DDS model of exchange-repulsion at HF level of theory
+is equivalent to the perturbative method of Hayes and Stone and the two methods
+yield the same results, although the exchange and repulsive interaction energies
+are slightly different.
+
+\subsubsection{\label{ssdff.sd}Defining OEP's for Repulsive Potential}
+
 \subsection{\label{ss.5.2}Charge-Transfer Interaction Energy}

### 2019-07-10 papers/2_oep_technique/paper-r0.tex
Paper revision and basis set module minirevision

@@
 \author{Marta Cho{\l}uj} 
-\author{Joanna Bednarska}
+\author{Joanna D. Bednarska}
 \author{Wojciech Bartkowiak}
@@
         \lambda \hat{v}_{\rm nuc} +
-        \sum_{\alpha\beta} 
-	\Ket{\alpha} 
+        %\sum_{\alpha\beta} 
+	%\Ket{\alpha} 
+        \int d{\bf r} \Ket{{\bf r}} 
 	\left[
 	\int
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
-	\rho^{\rm eff}({\bf r})
-	d{\bf r} 
+	\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
+	\rho^{\rm eff}({\bf r}')
+	d{\bf r}' 
 	\right]
-	\Bra{\beta}  
+	%\Bra{\beta}  
+        \Bra{{\bf r}}
 \end{equation}
@@
  \Delta^{AB} E^{\rm Rep} = 
- \Delta^{AB} E^{\rm Rep}(\mathcal{O}(S)) + 
- \Delta^{AB} E^{\rm Rep}(\mathcal{O}(S^2))
+ \Delta^{AB} E^{\rm Rep}(S^{-1}) + 
+ \Delta^{AB} E^{\rm Rep}(S^{-2}) \;,
 \end{equation}
@@
 %
-\begin{multline}
-  \Delta^{AB}  E^{\rm Rep}(\mathcal{O}(S)) = -2\sum_{a\in A} \sum_{b\in B}
-               S_{ab} \Big\{
-           V^A_{ab} 
-         + V^B_{ab} \\
- + \sum_{c\in A} \left[ 2(ab \vert cc) - (ac \vert bc) \right] 
- + \sum_{d\in B} \left[ 2(ab \vert dd) - (ad \vert bd) \right]
+\begin{multline} \label{e:rep.murrell-etal.S1}
+  \Delta^{AB}  E^{\rm Rep}(S^{-1}) = -2\sum_{i\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}
+               S_{ij} \Big\{
+           V^A_{ij} 
+         + V^B_{ji} \\
+ + \sum_{k\in A}^{\rm Occ} \left[ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right] 
+ + \sum_{l\in B}^{\rm Occ} \left[ 2\BraKet{ij}{ll} - \BraKet{il}{jl} \right]
                 \Big\}
@@
 very inexpensive approximation to the latter term has already been worked out previously. %cite
+Note that the Coulomb and exchange integrals can be re\hyp{}cast as follows:
+%
+\begin{subequations}
+ \begin{align}
+ \BraKet{ij}{kk} &\equiv \sum_{\mu\in A} 
+     C_{\mu i}^A \tBraKet{\mu}{\hat{v}_{kk}}{j} \;, \\
+ \BraKet{ik}{jk} &\equiv \sum_{\mu\in A} 
+     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
+ \end{align}
+\end{subequations}
+%
+%Similarly, the exchange integrals reduce to
+%
+%\begin{equation}
+% \BraKet{ik}{jk} \equiv \sum_{\mu\in A} 
+%     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
+%\end{equation}
+%
+In the above equations, the auxiliary potential operators are given by
+%
+\begin{equation}
+  \hat{v}_{ik} \equiv \int d{\bf r} \Ket{{\bf r}} 
+        \left[
+        \int d{\bf r}' \frac{\phi_i^{*}({\bf r}') \phi_k({\bf r}')}{\vert {\bf r}' - {\bf r}\vert}
+        \right] \Bra{{\bf r}} \;.
+\end{equation}
+%
+Now, using the prescription in Eq.~\eqref{e:ft-reduction} one can define a joint OEP
+operator from nuclear, Coulomb and exchange parts as
+%
+\begin{equation}
+ V^A_{ij} + 
+ \sum_{k\in A}^{\rm Occ} 
+  \left\{ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right\}
+\equiv \sum_{\mu\in A} \tBraKet{\mu}{ 
+\hat{v}^{A[\mu i]}_{\rm eff}
+ }{j}
+\end{equation}
+%
+with
+%
+\begin{equation}
+ \hat{v}^{A[\mu i]}_{\rm eff} \equiv C_{\mu i} \hat{v}^A_{\rm nuc} + 
+ \sum_{k\in A}^{\rm Occ} \left[
+ 2C_{\mu i}^A \hat{v}^A_{kk} - C_{\mu k}^A \hat{v}^A_{ik}
+ \right] \;.
+\end{equation}
+%
+%The nuclear part can also be incorporated within the effective potential
+%resulting in the following reduction
+%
+%
+On the other hand, it immediately follows that
+%
+\begin{equation}
+ \hat{v}^{A[\mu i]}_{\rm eff} \Ket{\mu} = 
+  \sum_{k\in A}^{\rm Occ} \left\{
+     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
+  \right\} \;.
+\end{equation}
+%
+For practical calculations, the right hand side of the above equation can be expanded 
+in the auxiliary basis,
+%
+\begin{equation}
+  \sum_{k\in A}^{\rm Occ} \left\{
+     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
+  \right\} \cong
+  \sum_{\xi\in A}^{\rm DF} 
+  V_{\xi i}^A \Ket{\xi} \;,
+\end{equation}
+%
+where the matrix ${\bf V}^A$ can be considered as effective fragment parameters.
+Doing the same operations on the twin operators associated with the molecule $B$
+original theory of Murrell et. al from Eq.~\eqref{e:rep.murrell-etal.S1}
+reduces to
+%
+\begin{equation}
+  \Delta^{AB}  E^{\rm Rep}(S^{-1}) \cong 
+ -2\sum_{i\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}
+               S_{ij} \Big\{
+           \sum_{\xi \in A}^{\rm DF} V_{\xi i}^A S_{\xi j}
+         + \sum_{\eta\in B}^{\rm DF} V_{\eta j}^B S_{\eta i}
+                \Big\}
+\end{equation}
+%
+which is a fully OEP\hyp{}based result. Note, that now only 
+overlap integrals, which are relatively computationally very inexpensive,
+need to be evaluated.
 

### 2019-07-10 gefp/gefp/basis/optimize.py

@@ -15,18 +15,21 @@ __all__ = ["DFBasis", "DFBasisOptimizer", "OEP"]
 
 def make_bastempl(templ, param): return templ % tuple(param)
 
-def bas(mol, role='ORBITAL'):
-    "Requires setting 'bas.templ' and 'bas.param' inside bas prior to executing bas"
+def oepfitbasis(mol, role='ORBITAL'):
+    "Requires setting 'oepfitbasis.templ' and 'oepfitbasis.param' inside bas prior to executing bas"
     basstrings = {}
     mol.set_basis_all_atoms("oepfit", role=role)
-    basstrings['oepfit'] = make_bastempl(bas.templ, bas.param)
+    basstrings['oepfit'] = make_bastempl(oepfitbasis.templ, oepfitbasis.param)
     return basstrings
+
+# -------------------------------------------------------------------------------------------------- #
 
 class DFBasis:
   """
  Basis set object to be optimized
 """
   def __init__(self, mol, templ_file='templ.dat', param_file='param.dat'):
+      "Initialize"
       # bound molecule to this basis set
       self.mol= mol
       # initial parameters
@@ -39,9 +42,9 @@ class DFBasis:
   def basisset(self, param=None):
       "Construct basis set from optimization parameters"
       if param is None: param = self.param_0
-      bas.param = param
-      bas.templ = self.templ
-      basis = psi4.core.BasisSet.build(self.mol, 'BASIS', bas)
+      oepfitbasis.param = param
+      oepfitbasis.templ = self.templ
+      basis = psi4.core.BasisSet.build(self.mol, 'BASIS', oepfitbasis)
       self.basis = basis
       return basis
 
@@ -52,6 +55,7 @@ class OEP(ABC):
  OEP object that defines the V matrix necessary for GDF
 """
   def __init__(self, wfn, dfbasis):
+      "Initialize"
       # wavefunction
       self.wfn = wfn
       # DF basis to optimize
@@ -90,10 +94,12 @@ class OEP_Pauli(OEP):
  OEP for S1 term in Murrell et al.'s theory of Pauli repulsion
  """
   def __init__(self, wfn, dfbasis):
+      "Initialize"
       super(OEP_Pauli, self).__init__(wfn, dfbasis)
 
   # - implementation - #
   def _compute_V(self):
+      "Compute OEP V for Pauli S1 term in Murrell et al.'s theory"
       Ca= self.wfn.Ca_subset("AO", "OCC").to_array(dense=True)
[…]

### 2019-07-11 papers/2_oep_technique/paper-r0.tex
Paper update: added more discussion about basis fitting

@@
 
-The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
+The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
 produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
@@
         \int d{\bf r} \Ket{{\bf r}} 
-	\left[
-	\int
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
-	\rho^{\rm eff}({\bf r}')
-	d{\bf r}' 
-	\right]
+        v^{\rm eff}({\bf r})
+	%\left[
+	%\int
+	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
+	%\rho^{\rm eff}({\bf r}')
+	%d{\bf r}' 
+	%\right]
 	%\Bra{\beta}  
@@
 with the nuclear
-and Coulomb operators defined by
-%
-\begin{subequations}
-\begin{align}
-        \hat{v}_{\rm nuc} &\equiv \\
-	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
-	f({\bf r}) 
-	\int
-	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-\end{align}
-\end{subequations}
-%
-for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
+%and Coulomb operators 
+operator
+defined by
+%
+%\begin{subequations}
+%\begin{align}
+\begin{equation}
+        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
+                     \int d{\bf r} \Ket{{\bf r}} 
+                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
+                     \Bra{{\bf r}} \;.
+	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
+	%f({\bf r}) 
+	%\int
+	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
+\end{equation}
+%\end{align}
+%\end{subequations}
+%
+%for any one\hyp{}electron function $f({\bf r})$. 
+and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
+The matrix element of the effective potential operator
 is therefore given by
@@
 \begin{equation}
-\tBraKet{\alpha}{\mathcal{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathcal{O}(1) \phi_\beta({\bf r}) d{\bf r} 
-\end{equation}
-%
-for any one\hyp{}electron operator $\mathcal{O}(1)$ 
+\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
+\end{equation}
+%
+for any one\hyp{}electron operator $\mathscr{O}(1)$ 
 and the electron repulsion integral (ERI)
@@
 for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
-with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
+with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator,
+$\hat{h}(1) = \hat{v}_{\rm nuc} -\frac{1}{2}\sum_e^{\rm el} \nabla_e^2$,
 and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-oep.rep}
   \sum_{k\in A}^{\rm Occ} \left\{
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-oep.ct}
 \Bra{i} \left[ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
@@
 %%
-%\appendix
+\appendix
+
+\section{Optimized Auxiliary Basis Sets for OEP Applications\label{a:auxiliary-basis}}
+
+To fit the auxiliary DF basis for the treatment of the overlap\hyp{}like
+OEP matrix elements with the operator $\hat{v}_{\rm eff}$, 
+the following objective function is minimized
+%
+\begin{equation} \label{e:a1-obj}
+ Z[\{\xi\}] = \sum_{\alpha i} \left[ 
+     \tBraKet{\alpha}{\hat{v}_{\rm eff}}{i} - 
+     \sum_{\xi}^{\rm DF} V_{\xi i} S_{\alpha \xi} 
+    \right]^2 \;,
+\end{equation}
+%
+where $\{\xi\}$ is the auxiliary basis set to optimize, whereas $\{\alpha\}$
+is the `test' basis set used to probe the accuracy of the density fitting.
+The forms of the Pauli and charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
+can be directly derived from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
+by expanding the MO's in terms of atomic basis functions. The working formulae
+are given below:
+%
+\begin{subequations} \label{e:a1-v.oep}
+ \begin{align}
+   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm Rep}}{i}
+     &= \sum_{x}^{\rm At} W_{\alpha i}^{(x)} \nonumber   \\ 
+        + \sum_{\beta\gamma\delta} &
+           \left\{ 
+             2 C_{\beta i} D_{\gamma\delta} - C_{\gamma i} D_{\beta \delta}
+           \right\}
+           \BraKet{\alpha\beta}{\gamma\delta} \;,\\
+   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{i} \;.
+     &=
+ \end{align}
+\end{subequations}
+%
+In the above equations, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
+is the one\hyp{}particle density matrix in AO basis.
+
 %

### 2019-07-11 include/oepdev_options.h
Interaction energy method choice options added.

@@ -68,6 +68,22 @@ int read_options(std::string name, Options& options)
         /*- Exclude octupoles in CT energy calculations? (relevant if above is "DMTP") -*/
         options.add_bool   ("EFP2_CT_NO_OCTUPOLES"  , true                       );
 
+        /*- Which methods of interaction energy are to be used? -*/
+        options.add_bool   ("OEPDEV_SOLVER_EINT_COUL_AO"  , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_COUL_MO"  , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_COUL_ESP" , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_COUL_CAMM", true                 ); 
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_HS"   , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_DDS"  , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_MRW"  , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_OL"   , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_OEP1" , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_OEP2" , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_REP_EFP2" , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_CT_OL"    , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_CT_OEP"   , true                 );
+        options.add_bool   ("OEPDEV_SOLVER_EINT_CT_EFP2"  , true                 );
+
 
         // ---> CPHF Global Options  <--- //
 

### 2019-07-11 main.cc

@@ -162,30 +162,30 @@ SharedWavefunction oepdev(SharedWavefunction ref_wfn, Options& options)
         {
             std::shared_ptr<oepdev::OEPDevSolver> solver = oepdev::OEPDevSolver::build("ELECTROSTATIC ENERGY", wfn_union); 
                                                                                                                             
-            double el1 = solver->compute_benchmark("AO_EXPANDED"    );
-            double el2 = solver->compute_benchmark("MO_EXPANDED"    );
-            double el3 = solver->compute_oep_based("ESP_SYMMETRIZED");
-	    double el4 = solver->compute_oep_based("CAMM"           );
+            if (options.get_bool("OEPDEV_SOLVER_EINT_COUL_AO"  )) solver->compute_benchmark("AO_EXPANDED"    );
+            if (options.get_bool("OEPDEV_SOLVER_EINT_COUL_MO"  )) solver->compute_benchmark("MO_EXPANDED"    );
+            if (options.get_bool("OEPDEV_SOLVER_EINT_COUL_ESP" )) solver->compute_oep_based("ESP_SYMMETRIZED");
+	    if (options.get_bool("OEPDEV_SOLVER_EINT_COUL_CAMM")) solver->compute_oep_based("CAMM"           );
         }
         else if (o_solver_type == "REPULSION_ENERGY") 
         {
             std::shared_ptr<oepdev::OEPDevSolver> solver = oepdev::OEPDevSolver::build("REPULSION ENERGY", wfn_union);
                                                                                                                             
-            double e_stone = solver->compute_benchmark("HAYES_STONE"  );
-            double e_dens  = solver->compute_benchmark("DENSITY_BASED");
-            double e_murr  = solver->compute_benchmark("MURRELL_ETAL" );
-            double e_otla  = solver->compute_benchmark("OTTO_LADIK"   );
-            double e_oep1  = solver->compute_oep_based("MURRELL_ETAL_GDF_ESP");
-	    double e_oep2  = solver->compute_oep_based("MURRELL_ETAL_GDF_CAMM");
-            double e_efp2  = solver->compute_benchmark("EFP2"         );        
+            if (options.get_bool("OEPDEV_SOLVER_EINT_REP_HS"  )) solver->compute_benchmark("HAYES_STONE"          );
[…]

### 2019-07-11 oepdev/libsolver/solver.h

@@ -43,6 +43,23 @@ using SharedOEPotential        = std::shared_ptr<OEPotential>;
  *  - `CHARGE TRANSFER ENERGY`
  *
  * # Options
+ * 
+ * ## Interaction Property Method      
+ * - `OEPDEV_SOLVER_EINT_COUL_AO`    - Coulombic energy: AO expanded
+ * - `OEPDEV_SOLVER_EINT_COUL_MO`    - Coulombic energy: MO expanded
+ * - `OEPDEV_SOLVER_EINT_COUL_ESP`   - Coulombic energy: ESP
+ * - `OEPDEV_SOLVER_EINT_COUL_CAMM`  - Coulombic energy: CAMM
+ * - `OEPDEV_SOLVER_EINT_REP_HS`     - Exchange-repulsion energy: Hayes-Stone
+ * - `OEPDEV_SOLVER_EINT_REP_DDS`    - Exchange-repulsion energy: DDS        
+ * - `OEPDEV_SOLVER_EINT_REP_MRW`    - Exchange-repulsion energy: Murrell et al.
+ * - `OEPDEV_SOLVER_EINT_REP_OL`     - Exchange-repulsion energy: Otto-Ladik 
+ * - `OEPDEV_SOLVER_EINT_REP_OEP1`   - Exchange-repulsion energy: OEP (S1: GDF, S2: ESP)
+ * - `OEPDEV_SOLVER_EINT_REP_OEP2`   - Exchange-repulsion energy: OEP (S1: GDF, S2: CAMM)
+ * - `OEPDEV_SOLVER_EINT_REP_EFP2`   - Exchange-repulsion energy: EFP2       
+ * - `OEPDEV_SOLVER_EINT_CT_OL`      - Charge-transfer energy: Otto-Ladik 
+ * - `OEPDEV_SOLVER_EINT_CT_OEP`     - Charge-transfer energy: OEP
+ * - `OEPDEV_SOLVER_EINT_CT_EFP2`    - Charge-transfer energy: EFP2        
+ *
  * ## Generalized density fitting (GDF) options:
  *  - `OEPDEV_DF_TYPE` - type of the GDF. Default: `DOUBLE`. Other: `SINGLE`.
  *  - `DF_BASIS_OEP` - auxiliary basis set. Default: `sto-3g`.
@@ -303,7 +320,7 @@ class ElectrostaticEnergySolver : public OEPDevSolver
  * <tr><th> Keyword  <th>Method Description  
[…]

### 2019-07-11 oepdev/libsolver/solver_energy_pauli.cc

@@ -41,7 +41,8 @@ double RepulsionEnergySolver::compute_benchmark(const std::string& method)
   double e = 0.0;
   if      (method == "DEFAULT" ||
            method == "HAYES_STONE"  ) e = compute_benchmark_hayes_stone();
-  else if (method == "DENSITY_BASED") e = compute_benchmark_density_based();
+  else if (method == "DDS"          ) e = compute_benchmark_dds();
+  else if (method == "DENSITY_BASED") e = compute_benchmark_dds();
   else if (method == "MURRELL_ETAL" ) e = compute_benchmark_murrell_etal();
   else if (method == "OTTO_LADIK"   ) e = compute_benchmark_otto_ladik();
   else if (method == "EFP2"         ) e = compute_benchmark_efp2();
@@ -51,7 +52,7 @@ double RepulsionEnergySolver::compute_benchmark(const std::string& method)
   }
   return e;
 }
-double RepulsionEnergySolver::compute_benchmark_density_based() {
+double RepulsionEnergySolver::compute_benchmark_dds() {
   double e             = 0.0;
   double e_Pauli_nuc   = 0.0;
   double e_Pauli_Pauli = 0.0;

### 2019-07-11 gefp/gefp/basis/optimize.py
CT OEP added to auxiliary basis set optimization

@@ -109,7 +109,8 @@ class OEP(ABC):
   @classmethod
   def create(cls, name, wfn, dfbasis):
       "Create OEP instance"
-      if name.lower() == 'pauli': oep = OEP_Pauli(wfn, dfbasis)
+      if   name.lower() == 'pauli': oep = OEP_Pauli(wfn, dfbasis)
+      elif name.lower() == 'ct'   : oep = OEP_CT   (wfn, dfbasis)
       else: raise ValueError("Not existent OEP chosen")
       return oep
   
@@ -118,7 +119,31 @@ class OEP(ABC):
       "Compute matrix representation of OEP"
       pass
 
-class OEP_Pauli(OEP):
+class OEP_FockLike(OEP):
+  """
+ OEP for S1 term in Murrell et al.'s theory of Pauli repulsion
+ """
+  def __init__(self, wfn, dfbasis):
+      "Initialize"
+      super(OEP_FockLike, self).__init__(wfn, dfbasis)
+
+  @abstractmethod
+  def _compute_Ca(self): pass
+
+  # - implementation - #
+  def _compute_V(self):
+      "Compute Fock-Like OEP Matrix"
+      Ca= self._compute_Ca()
+      Da= self.wfn.Da().to_array(dense=True)
+      Vao= self.mints.ao_potential(self.basis_test, self.basis_prim).to_array(dense=True)
+      V  = numpy.dot(Vao, Ca)
+      #
+      V += 2.0 * numpy.einsum("bi,mn,mnba->ai", Ca, Da, self.eri_pppt)
+      V -=       numpy.einsum("mi,nb,mnba->ai", Ca, Da, self.eri_pppt)
+      return V
+
+
+class OEP_Pauli(OEP_FockLike):
   """
  OEP for S1 term in Murrell et al.'s theory of Pauli repulsion
  """
@@ -127,18 +152,20 @@ class OEP_Pauli(OEP):
       super(OEP_Pauli, self).__init__(wfn, dfbasis)
 
   # - implementation - #
-  def _compute_V(self):
-      "Compute OEP V for Pauli S1 term in Murrell et al.'s theory"
-      Ca= self.wfn.Ca_subset("AO", "OCC").to_array(dense=True)
-      Da= self.wfn.Da().to_array(dense=True)
-      Vao= self.mints.ao_potential(self.basis_test, self.basis_prim).to_array(dense=True)
-      V  = numpy.dot(Vao, Ca)
-      #
[…]

### 2019-08-26 papers/3_eet.I.Theory/paper-r0.tex
EET Paper draft initialized

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

\title{Excitonic Energy Transfer Couplings at Short Distances: Efficient Transfer-Integral Fragmentation Approach
Based on Effective One-Electron Operators}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}
\author{Robert W. G{\'o}ra}
\author{Marta Cho{\l}uj} 
\author{Joanna D. Bednarska}
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}
Bla.
\end{abstract}

\pacs{}

\maketitle

\tableofcontents

\section{\label{s:1}Introduction}

\section{\label{s:2}Theory}

In the simplest version of the TI/CIS approach, the Hamiltonian of the 
complex of molecule $A$ and $B$ is constructed assuming the configuration interaction
singles (CIS) approximation involving the following four configurations
%
\begin{align}
 \Big| \Phi_1 \Big> &= \Big| \Psi_A^{(e)} \otimes \Psi_B^{(g)} \Big> \\
 \Big| \Phi_2 \Big> &= \Big| \Psi_A^{(g)} \otimes \Psi_B^{(e)} \Big> \\
 \Big| \Phi_3 \Big> &= \Big| \Psi_A^{(+)} \otimes \Psi_B^{(-)} \Big> \\
 \Big| \Phi_4 \Big> &= \Big| \Psi_A^{(-)} \otimes \Psi_B^{(+)} \Big> 
\end{align}
%
where $g$ and $e$ superscripts denote the ground and excited state of a molecule,
`+' and `-' label the cationic and anionic state, respectively, 
whereas $ \Big| \Psi_X \otimes \Psi_Y \Big> $ denotes the antisymmetrized Hartree product
of the monomer wavefunctions.
The associated diagonal Hamiltonian matrix elements can be defined as
%
\begin{align}
 \Big< \Phi_1 \Big| \mathscr{H} -E_{0} \Big| \Phi_1 \Big> &\equiv E_1 = E^A_{e\rightarrow g} 
     + \sum_{\mu\nu\in A} \left( P_{\nu\mu}^{A(e)} - P_{\nu\mu}^{A(g)}\right) \times
     \left\{ V_{\mu\nu}^{B({\rm nuc})} + \sum_{\lambda\sigma\in B} P_{\lambda\sigma}^{B(g)} 
     \left[ (\mu\nu | \sigma\lambda) - \frac{1}{2} (\mu\lambda | \sigma\nu) \right] \right\} \\
 \Big< \Phi_2 \Big| \mathscr{H} -E_{0} \Big| \Phi_2 \Big> &\equiv E_2 = E^B_{e\rightarrow g} 
     + \sum_{\mu\nu\in B} \left( P_{\nu\mu}^{B(e)} - P_{\nu\mu}^{B(g)}\right) \times
     \left\{ V_{\mu\nu}^{A({\rm nuc})} + \sum_{\lambda\sigma\in A} P_{\lambda\sigma}^{A(g)} 
     \left[ (\mu\nu | \sigma\lambda) - \frac{1}{2} (\mu\lambda | \sigma\nu) \right] \right\} \\
 \Big< \Phi_3 \Big| \mathscr{H} -E_{0} \Big| \Phi_3 \Big> &\equiv E_3 = 
 -\varepsilon_H^A + \varepsilon_L^B - \big( H^A H^A \big| L^B L^B \big)  \\
 \Big< \Phi_4 \Big| \mathscr{H} -E_{0} \Big| \Phi_4 \Big> &\equiv E_4 = 
  \varepsilon_L^A - \varepsilon_H^B - \big( L^A L^A \big| H^B H^B \big)  
\end{align}
%
The associated off\hyp{}diagonal Hamiltonian matrix elements are
%
\begin{align}
 \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_2 \Big> &\equiv V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl}\\
 \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm ET1} \\
 \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm ET2} \\
 \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm HT1} \\
 \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm HT2} \\
 \Big< \Phi_3 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm CT } 
\end{align}
%
where the Forster-type Coulombic (Coul), Dexter-type exchange (Exch), remaining overlap correction (Ovrl),
as well
as the electron, hole and charge (ET, HT, CT) transfer contributions are defined.

For a closed-shell system, the EET coupling constant for two electronic transitions
can then be given approximately by
%
\begin{equation}
  V \approx V^{\rm Direct} + V^{\rm Inirect}
\end{equation}
%
where the overlap\hyp{}corrected direct and indirect coupling constants are 
%
\begin{align}
  V^{\rm Direct  } &= V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl} \\
  V^{\rm Indirect} &= V^{\rm TI-2} + V^{\rm TI-3}
\end{align}
%
with
%
\begin{align}
 V^{\rm TI-2} &=-\frac{V^{\rm ET1} V^{\rm HT2}}{E_3-E_1} -\frac{V^{\rm ET2} V^{\rm HT1}}{E_4-E_1} \\
 V^{\rm TI-3} &= \frac{V^{\rm CT} \left( V^{\rm ET1} V^{\rm ET2} + V^{\rm HT1} V^{\rm HT2}\right) }{(E_3-E_1)(E_4-E_1)}
\end{align}
%
The exchange-Coulomb coupling takes the form
%
\begin{align}
 V^{\rm Coul} &= \frac{V^{{\rm Coul},(0)}}{1 - S_{12}^2} \\
 V^{\rm Exch} &= \frac{V^{{\rm Exch},(0)}}{1 - S_{12}^2} \\
 V^{\rm Ovrl} &=-\frac{(E_1+E_2)S_{12}}{2(1-S_{12}^2)}
\end{align}
%
The overlap-corrected ET, HT and CT matrix elements read
%
\begin{align}
 V^{\rm ET1} &= \left[ 1 - S_{13}^2 \right]^{-1} \left\{ V^{{\rm ET1},(0)} - \frac{1}{2} (E_1+E_2) S_{13} \right\} \\
 V^{\rm ET2} &= \left[ 1 - S_{24}^2 \right]^{-1} \left\{ V^{{\rm ET2},(0)} - \frac{1}{2} (E_1+E_2) S_{24} \right\} \\
 V^{\rm HT1} &= \left[ 1 - S_{14}^2 \right]^{-1} \left\{ V^{{\rm HT1},(0)} - \frac{1}{2} (E_1+E_2) S_{14} \right\} \\
 V^{\rm HT2} &= \left[ 1 - S_{23}^2 \right]^{-1} \left\{ V^{{\rm HT2},(0)} - \frac{1}{2} (E_1+E_2) S_{23} \right\} \\
 V^{\rm CT } &= \left[ 1 - S_{34}^2 \right]^{-1} \left\{ V^{{\rm CT },(0)} - \frac{1}{2} (E_1+E_2) S_{34} \right\} 
\end{align}
%
In the above equatons, the superscript (0) denotes that the matrix elements are not affected by the
overlap between molecular wavefunctions, and are given by
%
\begin{align}
 V^{{\rm Coul},(0)} &= \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
  (\mu\nu | \sigma\lambda) \\
 V^{{\rm Exch},(0)} &=-\frac{1}{2} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
  (\mu\lambda | \sigma\nu) \\
 V^{{\rm ET1},(0)} &= t_{H\rightarrow L}^A \left\{ \big( L^A \big| \mathscr{F} \big| L^B \big) 
   + 2 \big( L^A H^A \big| H^A L^B \big) - \big( L^A L^B \big| H^A H^A \big)  \right\} \\
 V^{{\rm ET2},(0)} &= t_{H\rightarrow L}^B \left\{ \big( L^A \big| \mathscr{F} \big| L^B \big) 
   + 2 \big( L^A H^B \big| H^B L^B \big) - \big( L^A L^B \big| H^B H^B \big)  \right\} \\
 V^{{\rm HT1},(0)} &= t_{H\rightarrow L}^A \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
   + 2 \big( H^A L^A \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^A \big)  \right\} \\
 V^{{\rm HT2},(0)} &= t_{H\rightarrow L}^B \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
   + 2 \big( H^A L^B \big| L^B H^B \big) - \big( H^A H^B \big| L^B L^B \big)  \right\} \\
 V^{{\rm CT },(0)} &= 
     2 \big( H^A L^B \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^B \Big) 
\end{align}
%
In the above, $\mathscr{F}$ is the Fock operator whereas $H$ and $L$ denote the 
HOMO and LUMO orbitals, respectively. It was found that the overlap corrections
are quite negligible.

\subsection{\label{s:2.1}Indirect coupling via electron and hole-transfer}

\subsection{\label{s:2.2}Indirect coupling via charge transfer}

\subsection{\label{s:2.3}Approximate pure exchange coupling}

\subsection{\label{s:2.4}Complete model of EET coupling}

\section{\label{s:3}Calculation Details}

All the models that were used to test the theory presented in this work
were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}

\section{\label{s:4}Results and Discussion}

\section{\label{s:5}Summary and a few concluding remarks}

Bla.

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

% -----------------------
\bibliography{references}
% -----------------------

\end{document}

### 2019-08-26 papers/3_eet.I.Theory/do-r0.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+base=paper-r0
+
+latex  ${base}.tex
+bibtex ${base}.aux
+dvipdf ${base}.dvi

### 2019-08-27 papers/3_eet.I.Theory/paper-r0.tex
EET Paper: Fujimoto's TI/CIS model described. OEPs for ET and HT terms derived and described.

@@
 \begin{abstract}
-Bla.
+What is EET? Short distances. Write about cost of current models. In this work, we developed a new model 
+that is
+based in the original TI/CIS model by Fujimoto. Write about cost reduction by introducing OEP's.
 \end{abstract}
@@
 
-\section{\label{s:2}Theory}
+Here why we want to care about EET at short distances. Why we want to devlop new model?
+Provide literature review
+of methods, then mention particularly the TDFI and TI/CIS model of Fujimoto, which
+is the starting point.
+
+\subsection{\label{s:2.0}Fujimoto's model of EET Coupling: Review}
 
@@
 complex of molecule $A$ and $B$ is constructed assuming the configuration interaction
-singles (CIS) approximation involving the following four configurations
+approximation involving the following four configurations
 %
+\begin{subequations}
 \begin{align}
- \Big| \Phi_1 \Big> &= \Big| \Psi_A^{(e)} \otimes \Psi_B^{(g)} \Big> \\
- \Big| \Phi_2 \Big> &= \Big| \Psi_A^{(g)} \otimes \Psi_B^{(e)} \Big> \\
- \Big| \Phi_3 \Big> &= \Big| \Psi_A^{(+)} \otimes \Psi_B^{(-)} \Big> \\
- \Big| \Phi_4 \Big> &= \Big| \Psi_A^{(-)} \otimes \Psi_B^{(+)} \Big> 
+ \big| \Phi_1 \big> &= \big| \Psi_A^{(e)} \otimes \Psi_B^{(g)} \big> \;, \\
+ \big| \Phi_2 \big> &= \big| \Psi_A^{(g)} \otimes \Psi_B^{(e)} \big> \;, \\
+ \big| \Phi_3 \big> &= \big| \Psi_A^{(+)} \otimes \Psi_B^{(-)} \big> \;, \\
+ \big| \Phi_4 \big> &= \big| \Psi_A^{(-)} \otimes \Psi_B^{(+)} \big> \;,
 \end{align}
+\end{subequations}
 %
@@
 `+' and `-' label the cationic and anionic state, respectively, 
-whereas $ \Big| \Psi_X \otimes \Psi_Y \Big> $ denotes the antisymmetrized Hartree product
-of the monomer wavefunctions.
-The associated diagonal Hamiltonian matrix elements can be defined as
-%
-\begin{align}
- \Big< \Phi_1 \Big| \mathscr{H} -E_{0} \Big| \Phi_1 \Big> &\equiv E_1 = E^A_{e\rightarrow g} 
-     + \sum_{\mu\nu\in A} \left( P_{\nu\mu}^{A(e)} - P_{\nu\mu}^{A(g)}\right) \times
-     \left\{ V_{\mu\nu}^{B({\rm nuc})} + \sum_{\lambda\sigma\in B} P_{\lambda\sigma}^{B(g)} 
-     \left[ (\mu\nu | \sigma\lambda) - \frac{1}{2} (\mu\lambda | \sigma\nu) \right] \right\} \\
- \Big< \Phi_2 \Big| \mathscr{H} -E_{0} \Big| \Phi_2 \Big> &\equiv E_2 = E^B_{e\rightarrow g} 
-     + \sum_{\mu\nu\in B} \left( P_{\nu\mu}^{B(e)} - P_{\nu\mu}^{B(g)}\right) \times
-     \left\{ V_{\mu\nu}^{A({\rm nuc})} + \sum_{\lambda\sigma\in A} P_{\lambda\sigma}^{A(g)} 
-     \left[ (\mu\nu | \sigma\lambda) - \frac{1}{2} (\mu\lambda | \sigma\nu) \right] \right\} \\
- \Big< \Phi_3 \Big| \mathscr{H} -E_{0} \Big| \Phi_3 \Big> &\equiv E_3 = 
- -\varepsilon_H^A + \varepsilon_L^B - \big( H^A H^A \big| L^B L^B \big)  \\
- \Big< \Phi_4 \Big| \mathscr{H} -E_{0} \Big| \Phi_4 \Big> &\equiv E_4 = 
-  \varepsilon_L^A - \varepsilon_H^B - \big( L^A L^A \big| H^B H^B \big)  
-\end{align}
-%
-The associated off\hyp{}diagonal Hamiltonian matrix elements are
-%
-\begin{align}
- \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_2 \Big> &\equiv V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl}\\
- \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm ET1} \\
- \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm ET2} \\
- \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm HT1} \\
- \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm HT2} \\
- \Big< \Phi_3 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm CT } 
-\end{align}
-%
-where the Forster-type Coulombic (Coul), Dexter-type exchange (Exch), remaining overlap correction (Ovrl),
-as well
-as the electron, hole and charge (ET, HT, CT) transfer contributions are defined.
-
-For a closed-shell system, the EET coupling constant for two electronic transitions
-can then be given approximately by
+whereas
 %
 \begin{equation}
-  V \approx V^{\rm Direct} + V^{\rm Inirect}
+\big| \Psi_X \otimes \Psi_Y \big> \equiv \mathscr{A} \Big\{ \big| \Psi_X \big> \otimes \big| \Psi_Y \big> \Big\}
 \end{equation}
 %
-where the overlap\hyp{}corrected direct and indirect coupling constants are 
+denotes the antisymmetrized Hartree product
+of the monomer wavefunctions with $\mathscr{A}$ being the standard antisymmetrizer operator. 
+Instead of diagonalizing the Hamiltonian in such basis,
+Fujimoto used the perturbation theory to obtain the approximate expression for the EET coupling
+constant,
 %
+\begin{multline}
+  V \approx \left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right> 
+   - \sum_{n=3,4} \frac{\left< \Phi_1 \vert \mathscr{H} \vert \Phi_n \right> 
+                        \left< \Phi_n \vert \mathscr{H} \vert \Phi_2 \right>}{E_n - E_1} \\
+   + \sum_{m,n=3,4}^{m\ne n}
+     \frac{\left< \Phi_1 \vert \mathscr{H} \vert \Phi_m \right>
+           \left< \Phi_m \vert \mathscr{H} \vert \Phi_n \right>
+           \left< \Phi_n \vert \mathscr{H} \vert \Phi_2 \right>}{(E_m-E_1)(E_n-E_1)} \;,
+\end{multline}
+%
+where $E_n \equiv \left< \Phi_n \vert \mathscr{H} -E_{0} \vert \Phi_n \right>$
+and $E_0$ is the ground state energy of the aggregate.
+The first two site energies are given by
+%
+\begin{subequations}
 \begin{align}
-  V^{\rm Direct  } &= V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl} \\
-  V^{\rm Indirect} &= V^{\rm TI-2} + V^{\rm TI-3}
+ E_1 \equiv \big< \Phi_1 \big| \mathscr{H} -E_{0} \big| \Phi_1 \big> &= E^A_{e\rightarrow g} 
+  + \Delta E^A_{e\rightarrow g}(B) \;, \\
+ E_2 \equiv \big< \Phi_2 \big| \mathscr{H} -E_{0} \big| \Phi_2 \big> &= E^B_{e\rightarrow g} 
+  + \Delta E^B_{e\rightarrow g}(A) \;,
 \end{align}
+\end{subequations}
 %
-with
+where $E^X_{e\rightarrow g}$ is the excitation energy of isolated molecule $X$ whereas 
+$\Delta E^X_{e\rightarrow g}(Y)$ accounts for the environmental effect due to other molecule,
 %
+\begin{equation}
+ \Delta E^X_{e\rightarrow g}(Y) \cong \int \left[ \rho^{X}_{{\rm el},(e)}({\bf r}) - \rho^{X}_{{\rm el},(g)}({\bf r}) \right]
+         v^{{\rm eff},Y}({\bf r}) \; d{\bf r}
+\end{equation}
+%
+with the effective one\hyp{}electron potential given by
+%
+\begin{multline}
+ v^{{\rm eff},Y}({\bf r}_1) \equiv v^{Y}_{\rm nuc}({\bf r})
+  + \sum_{ij}^{\rm Occ} \int d{\bf r}_2 \times \\
+ \Big\{
+ \frac{\phi_i^*({\bf r}_1) \phi_i({\bf r}_1)
+                    \phi_j^*({\bf r}_2) \phi_j({\bf r}_2)}{\vert {\bf r}_1 - {\bf r}_2 \vert}
+  - \frac{1}{2}
+   \frac{\phi_i^*({\bf r}_1 \phi_j({\bf r}_1 
+                 \phi_j^*({\bf r}_2 \phi_i({\bf r}_2)}{\vert {\bf r}_1 - {\bf r}_2 \vert}
+  \Big\} \;.
+\end{multline}
+%
+If the basis functions 3 and 4 are approximated as HOMO and LUMO orbitals of monomers,
+the remaining two site energies are given by
+%
+\begin{subequations}
 \begin{align}
- V^{\rm TI-2} &=-\frac{V^{\rm ET1} V^{\rm HT2}}{E_3-E_1} -\frac{V^{\rm ET2} V^{\rm HT1}}{E_4-E_1} \\
- V^{\rm TI-3} &= \frac{V^{\rm CT} \left( V^{\rm ET1} V^{\rm ET2} + V^{\rm HT1} V^{\rm HT2}\right) }{(E_3-E_1)(E_4-E_1)}
+ E_3 &\equiv \big< \Phi_3 \big| \mathscr{H} -E_{0} \big| \Phi_3 \big> \approx \nonumber \\ 
+ &\qquad\qquad -\varepsilon_H^A + \varepsilon_L^B - \big( \phi_H^A \phi_H^A \big| \phi_L^B \phi_L^B \big)  \;, \\
+ E_4 &\equiv \big< \Phi_4 \big| \mathscr{H} -E_{0} \big| \Phi_4 \big> \approx \nonumber \\
+ &\qquad\qquad \varepsilon_L^A - \varepsilon_H^B - \big( \phi_L^A \phi_L^A \big| \phi_H^B \phi_H^B \big)  \;.
 \end{align}
+\end{subequations}
 %
-The exchange-Coulomb coupling takes the form
+In the above, $\varepsilon_j^X$ is the energy of the $j$th Hartree-Fock orbital $\phi_j^X$ associated with molecule $X$,
+and the two\hyp{}electron integral is defined by
 %
-\begin{align}
- V^{\rm Coul} &= \frac{V^{{\rm Coul},(0)}}{1 - S_{12}^2} \\
- V^{\rm Exch} &= \frac{V^{{\rm Exch},(0)}}{1 - S_{12}^2} \\
- V^{\rm Ovrl} &=-\frac{(E_1+E_2)S_{12}}{2(1-S_{12}^2)}
-\end{align}
+\begin{equation}
+	\braket{\alpha\beta}{\gamma\delta} \equiv
+	\iint 
+	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
+	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
+	d{\bf r}_1 d{\bf r}_2  \;.
+\end{equation}
 %
-The overlap-corrected ET, HT and CT matrix elements read
-%
-\begin{align}
- V^{\rm ET1} &= \left[ 1 - S_{13}^2 \right]^{-1} \left\{ V^{{\rm ET1},(0)} - \frac{1}{2} (E_1+E_2) S_{13} \right\} \\
- V^{\rm ET2} &= \left[ 1 - S_{24}^2 \right]^{-1} \left\{ V^{{\rm ET2},(0)} - \frac{1}{2} (E_1+E_2) S_{24} \right\} \\
- V^{\rm HT1} &= \left[ 1 - S_{14}^2 \right]^{-1} \left\{ V^{{\rm HT1},(0)} - \frac{1}{2} (E_1+E_2) S_{14} \right\} \\
- V^{\rm HT2} &= \left[ 1 - S_{23}^2 \right]^{-1} \left\{ V^{{\rm HT2},(0)} - \frac{1}{2} (E_1+E_2) S_{23} \right\} \\
- V^{\rm CT } &= \left[ 1 - S_{34}^2 \right]^{-1} \left\{ V^{{\rm CT },(0)} - \frac{1}{2} (E_1+E_2) S_{34} \right\} 
-\end{align}
-%
-In the above equatons, the superscript (0) denotes that the matrix elements are not affected by the
-overlap between molecular wavefunctions, and are given by
-%
-\begin{align}
- V^{{\rm Coul},(0)} &= \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
-  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
-  (\mu\nu | \sigma\lambda) \\
- V^{{\rm Exch},(0)} &=-\frac{1}{2} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
-  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
-  (\mu\lambda | \sigma\nu) \\
- V^{{\rm ET1},(0)} &= t_{H\rightarrow L}^A \left\{ \big( L^A \big| \mathscr{F} \big| L^B \big) 
-   + 2 \big( L^A H^A \big| H^A L^B \big) - \big( L^A L^B \big| H^A H^A \big)  \right\} \\
- V^{{\rm ET2},(0)} &= t_{H\rightarrow L}^B \left\{ \big( L^A \big| \mathscr{F} \big| L^B \big) 
-   + 2 \big( L^A H^B \big| H^B L^B \big) - \big( L^A L^B \big| H^B H^B \big)  \right\} \\
- V^{{\rm HT1},(0)} &= t_{H\rightarrow L}^A \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
-   + 2 \big( H^A L^A \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^A \big)  \right\} \\
- V^{{\rm HT2},(0)} &= t_{H\rightarrow L}^B \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
-   + 2 \big( H^A L^B \big| L^B H^B \big) - \big( H^A H^B \big| L^B L^B \big)  \right\} \\
- V^{{\rm CT },(0)} &= 
-     2 \big( H^A L^B \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^B \Big) 
-\end{align}
-%
-In the above, $\mathscr{F}$ is the Fock operator whereas $H$ and $L$ denote the 
-HOMO and LUMO orbitals, respectively. It was found that the overlap corrections
-are quite negligible.
+
+%First, in Section II we briefly review the Fujimoto's TI/CIS model of EET coupling
+%which is treated in this work as a reference.
+%Then, further in this Section, we derive the OEP model based on the TI/CIS model (here referred to as the
+%which is validated in Section III against the TI/CIS method for a few model complexes. 
+First, in Section II we derive the OEP model based on the TI/CIS model (here referred to as the
+OEP-TI/CIS model). Next, we validate it against the TI/CIS method for a few model complexes
+in Section III. 
+Finally, we conclude our work in Section IV.
+
+\section{\label{s:2}OEP-Based Model}
+
+%The associated off\hyp{}diagonal Hamiltonian matrix elements are
+%%
+%\begin{align}
+% \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_2 \Big> &\equiv V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl}\\
+% \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm ET1} \\
+% \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm ET2} \\
+% \Big< \Phi_1 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm HT1} \\
+% \Big< \Phi_2 \Big| \mathscr{H} \Big| \Phi_3 \Big> &\equiv V^{\rm HT2} \\
+% \Big< \Phi_3 \Big| \mathscr{H} \Big| \Phi_4 \Big> &\equiv V^{\rm CT } 
+%\end{align}
+%%
+%where the Forster-type Coulombic (Coul), Dexter-type exchange (Exch), remaining overlap correction (Ovrl),
+%as well
+%as the electron, hole and charge (ET, HT, CT) transfer contributions are defined.
+
+%In the above equatons, the superscript (0) denotes that the matrix elements are not affected by the
+%overlap between molecular wavefunctions, and are given by
+%%
+%\begin{align}
+% V^{{\rm Coul},(0)} &= \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
+%  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
+%  (\mu\nu | \sigma\lambda) \\
+% V^{{\rm Exch},(0)} &=-\frac{1}{2} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
+%  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
+%  (\mu\lambda | \sigma\nu) \\
+% V^{{\rm HT1},(0)} &= t_{H\rightarrow L}^A \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
+%   + 2 \big( H^A L^A \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^A \big)  \right\} \\
+% V^{{\rm HT2},(0)} &= t_{H\rightarrow L}^B \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
+%   + 2 \big( H^A L^B \big| L^B H^B \big) - \big( H^A H^B \big| L^B L^B \big)  \right\} \\
+% V^{{\rm CT },(0)} &= 
+%     2 \big( H^A L^B \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^B \Big) 
+%\end{align}
+%%
+%In the above, $\mathscr{F}$ is the Fock operator whereas $H$ and $L$ denote the 
+%HOMO and LUMO orbitals, respectively. It was found that the overlap corrections
+%are quite negligible.
+
+To account for the EET coupling via charge transfer states
 
 \subsection{\label{s:2.1}Indirect coupling via electron and hole-transfer}
+
+The matrix elements that are needed for the electron transfer (ET) phenomena
+is
+%
+\begin{multline}\label{e:v0-et1}
+ V^{{\rm ET1},(0)} \equiv \left< \Phi_1 \vert \mathscr{H} \vert \Phi_3 \right> \approx 
+ t_{H\rightarrow L}^A \times \Big\{ 
+ \big( \phi_L^A \big| \mathscr{F} \big| \phi_L^B \big) \\
+   + 2 \big( \phi_L^A \phi_H^A \big| \phi_H^A \phi_L^B \big) - \big( \phi_L^A \phi_L^B \big| \phi_H^A \phi_H^A \big) 
+ \Big\} \;,
+\end{multline}
+%
+along with its twin term, $V^{{\rm ET2},(0)} \equiv \left< \Phi_2 \vert \mathscr{H} \vert \Phi_4 \right>$,
+which can be found in the original work of Fujimoto.
+In the above, $\mathscr{F}$ is the Fock operator of the entire molecular aggregate.
+Here, we approximate the Fock operator by constructing it from the unperturbed
+Fock operators of isolated monomers,
+%
+\begin{equation}
+ \mathscr{F} \approx \mathscr{F}^{A}_0 + \mathscr{F}^{B}_0 \;,
+\end{equation}
+%
+where
+%
+\begin{equation}
+ \mathscr{F}^{X}_0 = \hat{T}^{X} + \hat{V}^{X}_{\rm nuc} + \hat{J}({\bf P}^{X}_g) - \frac{1}{2} \hat{K}({\bf P}^{X}_g) \;.
+\end{equation}
+%
+Note that, without loss of generality, one can rewrite Eq.~\eqref{e:v0-et1}
+as
+%
+\begin{multline}
+ V^{{\rm ET1},(0)} = t_{H\rightarrow L}^A \times \Big\{ 
+ \tbraket{\phi_L^A}{\mathscr{F}}{\phi_L^B} \\
+  + 2 \tbraket{\phi_H^A}{\hat{v}_{LH}^A}{\phi_L^B}
+  -   \tbraket{\phi_L^A}{\hat{v}_{HH}^A}{\phi_L^B}
+ \Big\} \;,
+\end{multline}
+%
+in which the effective potential operator\cite{Blasiak.Bednarska.Choluj.Bartkowiak.XXXX} is defined by
+%
+\begin{equation}
+ \hat{v}_{ij}^X = \int d{\bf r} \Ket{{\bf r}} v_{ij}^X({\bf r})
+ \Bra{{\bf r}}
+\end{equation}
+%
+with its spatial form given as
+%
+\begin{equation}
+ v_{ij}^X({\bf r}) \equiv \int d{\bf r}' \frac{\phi_i^*({\bf r}') \phi_j({\bf r}') }{\vert {\bf r} - {\bf r}'\vert} \;.
+\end{equation}
+%
+Note here that $\phi_j({\bf r}) \equiv \BraKet{{\bf r}}{\phi_j}$.
+Now, 
+by using the OEP technique\cite{Blasiak.Bednarska.Choluj.Bartkowiak.XXXX}
+for overlap\hyp{}like effective potential operator matrix elements of the type AB
+one can gather operators associated to a particular molecule and group them
+into one effective one\hyp{}electron operator
+as
+%
+\begin{equation}
+ \bra{\phi_L^A} \left\{ \mathscr{F}_0^A - \hat{v}^A_{HH} \right\}
+ + \bra{\phi_H^A} 2 \hat{v}^A_{LH} 
+ \cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm ET1}}_{\xi,HL}
+\end{equation}
+%
+and
+%
+\begin{equation}
+ \mathscr{F}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm ET1}}_{\eta,L} \ket{\eta} 
+\end{equation}
+%
+In the above transformations, that involve the generalized density fitting
+in an auxiliary basis set space,
+$V^{A,{\rm ET1}}_{\xi,HL}$ and $V^{B,{\rm ET1}}_{\eta,L}$
+are the OEP matrix elements associated with the above
+effective potentials, separately defined for molecule $A$ and $B$. This allows one to
+further recast $V^{{\rm ET1},(0)}$ into
+%
+\begin{multline}\label{e:v0-et1-oep}
+ V^{{\rm ET1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
+ \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A,{\rm ET1}}_{\xi,HL} +
+ \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B,{\rm ET1}}_{\eta,L}
+ \Big\} \;,
+\end{multline}
+%
+which has a particularly simple form as compared to the original expression
+in Eq.~\eqref{e:v0-et1} that involves four\hyp{}center electron repulsion integrals.
+Note that if 
+the OEP matrices are considered
+as effective fragment parameters, %pre\hyp{}computed only once.
+only overlap integrals between auxiliary basis functions and LUMO orbitals of molecule
+$A$ and $B$ are necessary to compute, i.e.,
+%
+\begin{equation}
+ S^{XY}_{\xi U} = \sum_{\beta\in Y} S_{\xi\beta}^{XY} C_{\beta U}^Y \text{ for $\xi\in X$} \;,
+\end{equation}
+%
+in which 
+%
+\begin{equation}
+ S_{\xi\beta}^{XY} = \int d{\bf r} \varphi_\xi^*({\bf r}) \varphi_\beta({\bf r})
+\end{equation}
+%
+and $C_{\beta U}^Y$ is the SCF LCAO-MO matrix for isolated molecule $Y$.
+It is straightforward to show that the twin term ET2 is
+%
+\begin{multline}\label{e:v0-et2-oep}
+ V^{{\rm ET2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
+ \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B,{\rm ET1}}_{\eta,HL} +
+ \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A,{\rm ET1}}_{\xi,L}
+ \Big\} \;.
+\end{multline}
+%
+
+Similar considerations apply to the hole transfer (HT) matrix elements.
+The expression
+%
+\begin{multline}\label{e:v0-ht1}
+ V^{{\rm HT1},(0)} \equiv \left< \Phi_1 \vert \mathscr{H} \vert \Phi_4 \right> \approx 
+ t_{H\rightarrow L}^A \times \Big\{ 
+-\big(\phi_H^A \big| \mathscr{F} \big| \phi_H^B \big) \\
+   + 2 \big( \phi_H^A \phi_L^A \big| \phi_L^A \phi_H^B \big) - \big( \phi_H^A \phi_H^B \big| \phi_L^A \phi_L^A \big) 
+ \Big\} 
+\end{multline}
+%
+can be recast as
+%
+\begin{multline}\label{e:v0-ht1-oep}
+ V^{{\rm HT1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
+ \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A,{\rm HT1}}_{\xi,HL} +
+ \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B,{\rm HT1}}_{\eta,H}
+ \Big\} \;,
+\end{multline}
+%
+with the effective potential matrices given by
+%
+%
+\begin{subequations}
+\begin{align}
+ \bra{\phi_H^A} \left\{ -\mathscr{F}_0^A - \hat{v}^A_{LL} \right\}
+ + \bra{\phi_L^A} 2 \hat{v}^A_{HL} 
+ &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT1}}_{\xi,HL} \\
+ -\mathscr{F}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm ET1}}_{\eta,H} \ket{\eta} 
+\end{align}
+\end{subequations}
+%
+The twin term HT2 is accordingly
+%
+\begin{multline}\label{e:v0-ht2-oep}
+ V^{{\rm HT2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
+ \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B,{\rm HT2}}_{\eta,HL} +
+ \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A,{\rm HT2}}_{\xi,H}
+ \Big\} \;.
+\end{multline}
+%
+As can be seen, to compute Hamiltonian matrix elements
+associated with the electron and hole transfer, eight in total
+different OEP matrices need to be pre\hyp{}computed and stored in a file.
+Note that each of these matrices is actually just a vector of length
+equal to the size of the auxiliary basis set chosen. Therefore,
+it can be predicted that the computational cost of evaluation of the OEP\hyp{}based Hamiltonian
+matrix elements is negligible as compared to the original TI/CIS method.
+In Appendix~\ref{a:gdf-formulae}, explicit working formulae
+for the OEP matrix elements are given in terms of the AO's.
 
@@
 
-\section{\label{s:3}Calculation Details}
+To summarize, the overall form of the OEP-TI/CIS model is still given as in the original TI/CIS model
+by
+%
+\begin{equation}
+  V \approx V^{\rm Direct} + V^{\rm Inirect} \;,
+\end{equation}
+%
+where the so called `direct' and `indirect' coupling constants are respectively
+%
+\begin{subequations}
+\begin{align}
+  V^{\rm Direct  } &\equiv V^{\rm Coul} + V^{\rm Exch} + V^{\rm Ovrl} \;,\\
+  V^{\rm Indirect} &\equiv V^{\rm TI-2} + V^{\rm TI-3} \;,
+\end{align}
+\end{subequations}
+%
+with
+%
+\begin{subequations}
+\begin{align}
+ V^{\rm TI-2} &\equiv-\frac{V^{\rm ET1} V^{\rm HT2}}{E_3-E_1} -\frac{V^{\rm ET2} V^{\rm HT1}}{E_4-E_1}  \;,\\
+ V^{\rm TI-3} &\equiv \frac{V^{\rm CT} \left( V^{\rm ET1} V^{\rm ET2} + V^{\rm HT1} V^{\rm HT2}\right) }{(E_3-E_1)(E_4-E_1)} \;.
+\end{align}
+\end{subequations}
+%
+However, 
+the differences between the original Fujimoto's model and the newly derived OEP\hyp{}based model 
+are: 
+%
+\begin{enumerate}
+ \item The Hamiltonian matrix elements are evaluated from the fragment effective 
+parameters, for which only the one\hyp{}electron integrals are needed. Therefore, the computational
+cost is significantly reduced as compared to the parent TI/CIS model. In particular, 
+$V^{\rm Coul}$ is evaluated by using the TrCAMM method developed previously,
+$V^{\rm Exch}$ and $V^{\rm CT}$ are computed from the Mulliken approximated
+formulae in Eqs. XXX, and XXX, as well as $V^{\rm ET1}$, $V^{\rm ET2}$,
+$V^{\rm HT1}$ and $V^{\rm HT2}$ are treated by the generalized density fitting
+with OEP effective matrices as effective parameters spanned in auxiliary AO basis set 
+(Eqs.XXX); 
+ \item The relaxation effects due to the intermolecular interactions are not included
+in the current OEP\hyp{}based model.
+Note that in the original TI/CIS model the basis states and all the one\hyp{}particle density matrices are built from the
+self\hyp{}consistently adjusted fragment densities according to the DFI method.
+\end{enumerate}
+%
+\paragraph*{Overlap corrections to the Hamiltonian.}
+Since in the TI/CIS method the basis states are constructed from the monomer wavefunctions,
+the overlap effects should be properly taken into account. It was reported previously
+that overlap correction to the $\left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right>$ 
+matrix element brings negligible contributions to the 
+total EET coupling (c.f. the `overlap' contribution in Ref.). On the other hand,
+the importance of the overlap effects on the other off\hyp{}diagonal Hamiltonian matrix elements were not directly studied. 
+After the overlap correction, the constituents of the direct coupling take the form
+%
+\begin{subequations}\label{e:ovrl-direct}
+\begin{align}
+ V^{\rm Coul} &= \frac{V^{{\rm Coul},(0)}}{1 - S_{12}^2} \;, \\
+ V^{\rm Exch} &= \frac{V^{{\rm Exch},(0)}}{1 - S_{12}^2} \;, \\
+ V^{\rm Ovrl} &=-\frac{(E_1+E_2)S_{12}}{2(1-S_{12}^2)}   \;.
+\end{align}
+\end{subequations}
+%
+The overlap\hyp{}corrected ET, HT and CT matrix elements read
+%
+\begin{equation}\label{e:ovrl-indirect}
+ V^{\rm t} = \left[ 1 - S_{\rm t}^2 \right]^{-1} \left\{ V^{{\rm t},(0)} - \frac{1}{2} (E_1+E_2) S_{\rm t} \right\} \;,
+\end{equation}
+%
+where `t' denotes one of the matrix element types (ET1, ET2, HT1, HT2, CT) 
+and $S_{\rm t}$ is the appropriate overlap integral between basis states.
+Since evaluation
+of Eqs.~\eqref{e:ovrl-direct} and \eqref{e:ovrl-indirect} 
+is computationally inexpensive, we shall consider the overlap\hyp{}corrected
+Hamiltonian in the present work.
+
+\section{\label{s:4}Results and Discussion}
 
@@
 were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-
-\section{\label{s:4}Results and Discussion}
 
@@
 
+\appendix
+
+\section{Explicit formulae for ET and HT OEP matrix elements\label{a:gdf-formulae}}
+
+As developed in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.XXXX}, 
+the matrix elements of the OEP operators
+are computed from the generalized density fitting scheme, according to which
+%
+\begin{equation}
+ V^X_{\xi, M} = \sum_{\xi'\in X}^{\rm DF} 
+                \sum_{\varepsilon\in X}^{\rm RI}
+                [{\bf R}^{-1}]_{\xi\xi'} R_{\xi'\varepsilon} H^X_{\varepsilon, M} \;,
+\end{equation}
+%
+where
+%
+\begin{equation}
+ H^X_{\varepsilon, M} = \sum_{\varepsilon'\in X}^{\rm RI}
+                        [{\bf S}^{-1}]_{\varepsilon\varepsilon'} a_{\varepsilon'}^{X; M} \;.
+\end{equation}
+%
+In the above equations, the auxiliary matrices are defined by
+%
+\begin{equation}
+ S_{\eta\xi}  = \int 
+                       \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) 
+                 d{\bf r} 
+\end{equation}
+%
+and
+%
+\begin{equation}
+ R_{\eta\xi}  = \iint 
+                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
+                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
+                 d{\bf r}_1 d{\bf r}_2 \;.
+\end{equation}
+%
+The elements of vector ${\bf a}^X$ for $M=$ ET1, ET2, HT1 and HT2 and $X=A, B$ are given by
+%
+\begin{subequations}
+\begin{align}
+a_{\alpha}^{A;{\rm ET1}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \nonumber \\
+ &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+ \left\{ 2C^A_{\beta  H} C^A_{\gamma L} - C^A_{\beta  L} C^A_{\gamma H}\right\} C^A_{\delta H} \;,\\
+%
+a_{\alpha}^{B;{\rm ET1}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
+%
+a_{\alpha}^{B;{\rm ET2}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \nonumber \\
+ &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+ \left\{ 2C^B_{\beta  H} C^B_{\gamma L} - C^B_{\beta  L} C^B_{\gamma H}\right\} C^B_{\delta H} \;,\\
+%
+a_{\alpha}^{A;{\rm ET2}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;,\\
+%
+a_{\alpha}^{A;{\rm HT1}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \nonumber \\
+ &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+ \left\{ 2C^A_{\beta  L} C^A_{\gamma H} - C^A_{\beta  H} C^A_{\gamma L}\right\} C^A_{\delta L} \;,\\
+%
+a_{\alpha}^{B;{\rm HT1}} &=-\sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
+%
+a_{\alpha}^{B;{\rm HT2}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \nonumber \\
+ &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+ \left\{ 2C^B_{\beta  L} C^B_{\gamma H} - C^B_{\beta  H} C^B_{\gamma L}\right\} C^B_{\delta L} \;,\\
+%
+a_{\alpha}^{A;{\rm HT2}} &=-\sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;.
+\end{align}
+\end{subequations}
+%
 % -----------------------

### 2019-08-27 papers/3_eet.I.Theory/paper-r0.tex
Entire OEP-based EET model described

@@
 %
-\begin{multline}
+\begin{multline}\label{e:v0-fujimoto}
   V \approx \left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right> 
@@
 
+The first term in Eq.~\eqref{e:v0-fujimoto} is the, so called, `direct' coupling, which is 
+composed of two contributions:
+%
+\begin{equation}
+ \left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right> = V^{{\rm Coul},(0)} + V^{{\rm Exch},(0)}
+\end{equation}
+%
+where the F{\"o}rster (or Coulomb) coupling is given by
+%
+\begin{equation}
+ V^{{\rm Coul},(0)} = \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
+  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
+  \braket{\mu\nu}{\sigma\lambda}  \;,
+\end{equation}
+%
+whereas the Dexter (or exchange) coupling is
+%
+\begin{equation}
+ V^{{\rm Exch},(0)} =-\frac{1}{2} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
+  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
+  \braket{\mu\nu}{\sigma\lambda}  \;.
+\end{equation}
+%
+In the above equations, ${\bf P}^{g\rightarrow e(X)}$ is the transition total one\hyp{}
+particle density matrix of molecule $X$ in AO representation. $V^{{\rm Coul},(0)}$,
+which is usually a leading term for singlet-singlet EET (but vanishes for triplet-triplet EET)
+can be efficiently evaluated by various multipole expansion models published to date.
+$V^{{\rm Exch},(0)}$ is known to be negligible, even for close intermolecular contacts.
+However, unlike the direct coupling, the second and third term in Eq.~\eqref{e:v0-fujimoto}, which is collectively
+referred to by Fujimoto as the `indirect' coupling, are not negligible at short intermolecular distances.
+Unfortunately, evaluation of the indirect coupling is quite expensive and requires evaluation
+of the electron repulsion integrals (ERI's).
+
 %First, in Section II we briefly review the Fujimoto's TI/CIS model of EET coupling
@@
 %which is validated in Section III against the TI/CIS method for a few model complexes. 
+Since the TI/CIS model was found very accurate to estimate EET coupling constants for many organic
+molecules, we consider this model as a starting point to develop computationally more effcient
+scheme that does not involve electron repulsion integrals at all.
 First, in Section II we derive the OEP model based on the TI/CIS model (here referred to as the
@@
 %%
-%\begin{align}
-% V^{{\rm Coul},(0)} &= \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
-%  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
-%  (\mu\nu | \sigma\lambda) \\
-% V^{{\rm Exch},(0)} &=-\frac{1}{2} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
-%  P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
-%  (\mu\lambda | \sigma\nu) \\
-% V^{{\rm HT1},(0)} &= t_{H\rightarrow L}^A \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
-%   + 2 \big( H^A L^A \big| L^A H^B \big) - \big( H^A H^B \big| L^A L^A \big)  \right\} \\
-% V^{{\rm HT2},(0)} &= t_{H\rightarrow L}^B \left\{-\big( H^A \big| \mathscr{F} \big| H^B \big) 
-%   + 2 \big( H^A L^B \big| L^B H^B \big) - \big( H^A H^B \big| L^B L^B \big)  \right\} \\
 % V^{{\rm CT },(0)} &= 
@@
 %%
-%In the above, $\mathscr{F}$ is the Fock operator whereas $H$ and $L$ denote the 
-%HOMO and LUMO orbitals, respectively. It was found that the overlap corrections
-%are quite negligible.
 
-To account for the EET coupling via charge transfer states
+To account for the EET coupling via charge transfer states ...
 
@@
 
+The Hamiltonian matrix element that is associated with the charge transfer process
+is
+%
+\begin{equation}\label{e:v0-ct}
+ V^{{\rm CT },(0)} = 
+     2 \braket{\phi_H^A \phi_L^B}{\phi_L^A \phi_H^B}
+      -\braket{\phi_H^A \phi_H^B}{\phi_L^A \phi_L^B}
+\end{equation}
+%
+Since it involves ERI's of type $\braket{AB}{AB}$ OEP's cannot be directly defined here.
+However, Mulliken approximation allows one to express ERI's in terms of the
+overlap and Coulomb integrals, i.e.,
+%
+\begin{equation}
+ \braket{ij}{kl} \approx \frac{1}{4} S_{ij}S_{kl}
+ \left[ \braket{ii}{kk} + \braket{ii}{ll} + \braket{jj}{kk} + \braket{jj}{ll}\right] \;.
+\end{equation}
+%
+Application of the above approximation to Eq.~\eqref{e:v0-ct} yields
+%
+\begin{multline}\label{e:v0-ct.mulliken}
+ V^{{\rm CT },(0)} \approx \frac{1}{2} S_{HL}^{AB} S_{LH}^{AB} \times \\
+  \left[ r_{HL}^A + r_{HL}^B + \braket{\phi_H^A \phi_H^A}{\phi_H^B \phi_H^B} 
+                             + \braket{\phi_L^A \phi_L^A}{\phi_L^B \phi_L^B}\right] \\
+ -\frac{1}{4} S_{HH}^{AB} S_{LL}^{AB} \times \\
+  \left[ r_{HL}^A + r_{HL}^B + \braket{\phi_H^A \phi_H^A}{\phi_L^B \phi_L^B} 
+                             + \braket{\phi_L^A \phi_L^A}{\phi_H^B \phi_H^B}\right]  \;,
+\end{multline}
+%
+where 
+%
+\begin{equation} 
+ r_{UW}^X \equiv \braket{\phi_U^X \phi_U^X}{\phi_W^X \phi_W^X} \;.
+\end{equation}
+%
+First of all, note that $r_{HL}^A$ and $r_{HL}^B$ are just constant numbers
+associated with the isolated monomers and 
+independent on the molecular aggregate. Also, the remaining four ERI's
+can be considered as Coulombic interactions between HOMO and LUMO orbitals
+of unperturbed molecular wavefunctions.
+Then, applying distributed multipole expansion and truncating it on the monopole
+one finds that
+%
+\begin{equation} 
+ \braket{\phi_H^A \phi_H^A}{\phi_H^B \phi_H^B} \approx \frac{1}{\vert {\bf r}_H^A - {\bf r}_H^B \vert} \;,
+\end{equation}
+%
+where the charge centroids of HOMO orbitals are given by
+%
+\begin{equation} 
+ {\bf r}_H^X = \tbraket{\phi_H^X}{\hat{\bf r}}{\phi_H^X} \;.
+\end{equation}
+%
+Note that, once occupied molecular orbitals are localized, the above approximation
+should yield correct description. LUMO orbitals are usually not much localized.
+Here, we use the distributed cumulative atomic charges to represent them, %cite CAMM paper
+%
+\begin{equation} 
+ q^X_{x,L} =-\sum_{\alpha\in x} \sum_\beta S_{\alpha\beta} C_{\alpha L}^X C_{\beta L}^X
+\end{equation}
+%
+and one can find that assuming this one obtains
+%
+\begin{equation} 
+ \braket{\phi_L^A \phi_L^A}{\phi_L^B \phi_L^B} 
+ \approx \sum_{x\in A}^{\rm At} \sum_{y\in B}^{\rm At}
+ \frac{q_{x;L}^A q_{y;L}^B}{\vert {\bf r}_x - {\bf r}_y \vert} \;.
+\end{equation}
+%
+The two remaining ERI's can be then approximated as follows:
+%
+\begin{equation} 
+ \braket{\phi_H^A \phi_H^A}{\phi_L^B \phi_L^B} 
+ \approx -\sum_{y\in B}^{\rm At}
+ \frac{q_{y;L}^B}{\vert {\bf r}_H^A - {\bf r}_y \vert} \;,
+\end{equation}
+%
+and similarly for the twin ERI. To summarize the derivation, the CT matrix element
+is approximately given by
+%
+\begin{multline}\label{e:v0-ct.mulliken.working}
+ V^{{\rm CT },(0)} \approx \frac{1}{2} S_{HL}^{AB} S_{LH}^{AB} \times 
+  \Big[ r_{HL}^A + r_{HL}^B \\ 
+ + \frac{1}{\vert {\bf r}_H^A - {\bf r}_H^B \vert} 
+ + \sum_{x\in A}^{\rm At} \sum_{y\in B}^{\rm At} 
+ \frac{q_{x;L}^A q_{y;L}^B}{\vert {\bf r}_x - {\bf r}_y \vert} \Big] \\
+ -\frac{1}{4} S_{HH}^{AB} S_{LL}^{AB} \times 
+  \Big[ r_{HL}^A + r_{HL}^B \\
+ -\sum_{y\in B}^{\rm At}
+ \frac{q_{y;L}^B}{\vert {\bf r}_H^A - {\bf r}_y \vert} 
+ -\sum_{x\in A}^{\rm At}
+ \frac{q_{x;L}^A}{\vert {\bf r}_H^B - {\bf r}_x \vert} \Big]\;.
+\end{multline}
+%
+
 \subsection{\label{s:2.3}Approximate pure exchange coupling}
+
+Mulliken approximation can also be used to treat the pure exchange coupling (Eq.).
+It is straightforward to show that
+%
+\begin{multline}\label{e:v0-exch-mulliken}
+V^{{\rm Exch},(0)} \approx -\frac{1}{8} \sum_{\mu\nu\in A} \sum_{\lambda\sigma\in B} 
+ P_{\nu\mu}^{g\rightarrow e(A)} P_{\lambda\sigma}^{g\rightarrow e(B)} 
+ S_{\mu\lambda}  S_{\sigma\nu} \times \\
+ \left[ \braket{\mu\mu}{\sigma\sigma} + \braket{\lambda\lambda}{\nu\nu} 
+      + \braket{\mu\mu}{\nu\nu} + \braket{\lambda\lambda}{\sigma\sigma}\right] \;.
+\end{multline}
+%
+However, ERI's that involve AO's in two different molecules can be safely neglected.
+That leads to the following simplified expression:
+%
+\begin{multline}\label{e:v0-exch.mulliken.working}
+V^{{\rm Exch},(0)} \approx 
+  -\frac{1}{8} \Big\{ \sum_{\mu\nu\in A} P_{\nu\mu}^{g\rightarrow e(A)} G_{\mu\nu}^A
+    [{\bf s}^{AB} {\bf P}^{g\rightarrow e(B)} {\bf s}^{BA} ]_{\mu\nu} \\
+  + \sum_{\sigma\lambda\in B} P_{\lambda\sigma}^{g\rightarrow e(B)} G_{\lambda\sigma}^B
+    [{\bf s}^{BA} {\bf P}^{g\rightarrow e(A)} {\bf s}^{AB} ]_{\sigma\lambda} 
+  \Big\}
+\end{multline}
+%
+with $G_{\mu\nu} \equiv \braket{\mu\mu}{\nu\nu}$. Note that $G_{\mu\nu}^A$ and $G_{\lambda\sigma}^B$
+are just constant matrices that can be considered as effective fragment potentials.
 
@@
 $V^{\rm Exch}$ and $V^{\rm CT}$ are computed from the Mulliken approximated
-formulae in Eqs. XXX, and XXX, as well as $V^{\rm ET1}$, $V^{\rm ET2}$,
+formulae in Eqs.~\eqref{e:v0-exch.mulliken.working}, and \eqref{e:v0-ct.mulliken.working}, respectively,
+as well as $V^{\rm ET1}$, $V^{\rm ET2}$,
 $V^{\rm HT1}$ and $V^{\rm HT2}$ are treated by the generalized density fitting
 with OEP effective matrices as effective parameters spanned in auxiliary AO basis set 
-(Eqs.XXX); 
+(Eqs.~\eqref{e:v0-et1-oep}, \eqref{e:v0-et2-oep}, \eqref{e:v0-ht1-oep}, \eqref{e:v0-ht2-oep}
+and Eqs.~\eqref{e:a:voep}-\eqref{e:a:aoep}); 
  \item The relaxation effects due to the intermolecular interactions are not included
@@
 %
-\begin{equation}
+\begin{equation}\label{e:a:voep}
  V^X_{\xi, M} = \sum_{\xi'\in X}^{\rm DF} 
@@
 %
-The elements of vector ${\bf a}^X$ for $M=$ ET1, ET2, HT1 and HT2 and $X=A, B$ are given by
+The elements of vector ${\bf a}^{X;M}$ for $M=$ ET1, ET2, HT1 and HT2 and $X=A, B$ are given by
 %
-\begin{subequations}
+\begin{subequations}\label{e:a:aoep}
 \begin{align}

### 2019-08-28 papers/3_eet.I.Theory/paper-r0.tex
Equations of OEP matrices updated

@@
 where $E^X_{e\rightarrow g}$ is the excitation energy of isolated molecule $X$ whereas 
-$\Delta E^X_{e\rightarrow g}(Y)$ accounts for the environmental effect due to other molecule,
+$\Delta E^X_{e\rightarrow g}(Y)$ accounts for the environmental effect due to the other molecule,
 %
@@
   - \frac{1}{2}
-   \frac{\phi_i^*({\bf r}_1 \phi_j({\bf r}_1 
-                 \phi_j^*({\bf r}_2 \phi_i({\bf r}_2)}{\vert {\bf r}_1 - {\bf r}_2 \vert}
+   \frac{\phi_i^*({\bf r}_1) \phi_j({\bf r}_1)
+                 \phi_j^*({\bf r}_2) \phi_i({\bf r}_2)}{\vert {\bf r}_1 - {\bf r}_2 \vert}
   \Big\} \;.
@@
  + \bra{\phi_H^A} 2 \hat{v}^A_{LH} 
- \cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm ET1}}_{\xi,HL}
+ \cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A;{\rm ET}}_{\xi;HL}
 \end{equation}
@@
 \begin{equation}
- \mathscr{F}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm ET1}}_{\eta,L} \ket{\eta} 
+ \mathscr{F}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B;{\rm ET}}_{\eta;L} \ket{\eta} 
 \end{equation}
@@
 in an auxiliary basis set space,
-$V^{A,{\rm ET1}}_{\xi,HL}$ and $V^{B,{\rm ET1}}_{\eta,L}$
+$V^{A;{\rm ET}}_{\xi;HL}$ and $V^{B;{\rm ET}}_{\eta;L}$
 are the OEP matrix elements associated with the above
@@
  V^{{\rm ET1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
- \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A,{\rm ET1}}_{\xi,HL} +
- \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B,{\rm ET1}}_{\eta,L}
+ \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A;{\rm ET}}_{\xi;HL} +
+ \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B;{\rm ET}}_{\eta;L}
  \Big\} \;,
@@
 only overlap integrals between auxiliary basis functions and LUMO orbitals of molecule
-$A$ and $B$ are necessary to compute, i.e.,
+$A$ and $B$ are necessary to be computed, i.e.,
 %
@@
 %
-and $C_{\beta U}^Y$ is the SCF LCAO-MO matrix for isolated molecule $Y$.
+and $C_{\beta U}^X$ is the SCF LCAO-MO matrix for isolated molecule $X$.
 It is straightforward to show that the twin term ET2 is
@@
  V^{{\rm ET2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
- \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B,{\rm ET1}}_{\eta,HL} +
- \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A,{\rm ET1}}_{\xi,L}
+ \sum_{\eta\in B}^{\rm DF} S_{\eta L}^{BA} V^{B;{\rm ET}}_{\eta;HL} +
+ \sum_{\xi\in A}^{\rm DF} S_{\xi L}^{AB} V^{A;{\rm ET}}_{\xi;L}
  \Big\} \;.
@@
 %
+Note that OEP matrices for ET1 and ET2 terms share similar structure.
+Therefore a joint superscript `ET' is used here to label the OEP matrices
+and additional subscripts
+$L$ and $HL$ denote the dependence on only LUMO and on both LUMO and HOMO orbitals, respectively.
 
@@
  V^{{\rm HT1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
- \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A,{\rm HT1}}_{\xi,HL} +
- \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B,{\rm HT1}}_{\eta,H}
+ \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;HL} +
+ \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;H}
  \Big\} \;,
@@
  + \bra{\phi_L^A} 2 \hat{v}^A_{HL} 
- &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT1}}_{\xi,HL} \\
- -\mathscr{F}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm ET1}}_{\eta,H} \ket{\eta} 
+ &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT}}_{\xi,HL} \\
+ -\mathscr{F}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm HT}}_{\eta,H} \ket{\eta} 
 \end{align}
@@
  V^{{\rm HT2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
- \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B,{\rm HT2}}_{\eta,HL} +
- \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A,{\rm HT2}}_{\xi,H}
+ \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;HL} +
+ \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;H}
  \Big\} \;.
@@
      2 \braket{\phi_H^A \phi_L^B}{\phi_L^A \phi_H^B}
-      -\braket{\phi_H^A \phi_H^B}{\phi_L^A \phi_L^B}
+      -\braket{\phi_H^A \phi_H^B}{\phi_L^A \phi_L^B} \;.
 \end{equation}
@@
 \begin{equation} 
- r_{UW}^X \equiv \braket{\phi_U^X \phi_U^X}{\phi_W^X \phi_W^X} \;.
+ r_{HL}^X \equiv \braket{\phi_H^X \phi_H^X}{\phi_L^X \phi_L^X} \;.
 \end{equation}
@@
 %
+Thus, the computational cost of the above expression is negligible as
+compared to the original formula in Eq.~\eqref{e:v0-ct}.
 
@@
 with $G_{\mu\nu} \equiv \braket{\mu\mu}{\nu\nu}$. Note that $G_{\mu\nu}^A$ and $G_{\lambda\sigma}^B$
-are just constant matrices that can be considered as effective fragment potentials.
+are just constant matrices that can be considered as effective fragment parameters.
 
@@
 %
-where the so called `direct' and `indirect' coupling constants are respectively
+where the direct and indirect coupling constants are respectively
 %
@@
 %
-However, 
-the differences between the original Fujimoto's model and the newly derived OEP\hyp{}based model 
-are: 
-%
-\begin{enumerate}
- \item The Hamiltonian matrix elements are evaluated from the fragment effective 
-parameters, for which only the one\hyp{}electron integrals are needed. Therefore, the computational
-cost is significantly reduced as compared to the parent TI/CIS model. In particular, 
-$V^{\rm Coul}$ is evaluated by using the TrCAMM method developed previously,
-$V^{\rm Exch}$ and $V^{\rm CT}$ are computed from the Mulliken approximated
-formulae in Eqs.~\eqref{e:v0-exch.mulliken.working}, and \eqref{e:v0-ct.mulliken.working}, respectively,
-as well as $V^{\rm ET1}$, $V^{\rm ET2}$,
-$V^{\rm HT1}$ and $V^{\rm HT2}$ are treated by the generalized density fitting
-with OEP effective matrices as effective parameters spanned in auxiliary AO basis set 
-(Eqs.~\eqref{e:v0-et1-oep}, \eqref{e:v0-et2-oep}, \eqref{e:v0-ht1-oep}, \eqref{e:v0-ht2-oep}
-and Eqs.~\eqref{e:a:voep}-\eqref{e:a:aoep}); 
- \item The relaxation effects due to the intermolecular interactions are not included
-in the current OEP\hyp{}based model.
-Note that in the original TI/CIS model the basis states and all the one\hyp{}particle density matrices are built from the
-self\hyp{}consistently adjusted fragment densities according to the DFI method.
-\end{enumerate}
-%
-\paragraph*{Overlap corrections to the Hamiltonian.}
-Since in the TI/CIS method the basis states are constructed from the monomer wavefunctions,
-the overlap effects should be properly taken into account. It was reported previously
-that overlap correction to the $\left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right>$ 
-matrix element brings negligible contributions to the 
-total EET coupling (c.f. the `overlap' contribution in Ref.). On the other hand,
-the importance of the overlap effects on the other off\hyp{}diagonal Hamiltonian matrix elements were not directly studied. 
-After the overlap correction, the constituents of the direct coupling take the form
+In the above, the direct EET coupling contributions are
 %
@@
 %
-The overlap\hyp{}corrected ET, HT and CT matrix elements read
+whereas the indirect EET coupling is evaluated from
+the overlap\hyp{}corrected ET, HT and CT matrix elements that read
 %
@@
 and $S_{\rm t}$ is the appropriate overlap integral between basis states.
-Since evaluation
+
+Despite the same general form,
+the differences between the original Fujimoto's model and the newly derived OEP\hyp{}based model 
+are significant:
+%
+\begin{enumerate}
+ \item The Hamiltonian matrix elements are evaluated from the fragment effective 
+parameters, for which only the one\hyp{}electron integrals are needed. Therefore, the computational
+cost is significantly reduced as compared to the parent TI/CIS model. In particular, 
+$V^{{\rm Coul},(0)}$ is evaluated by using the TrCAMM method developed previously,
+$V^{{\rm Exch},(0)}$ and $V^{{\rm CT},(0)}$ are computed from the Mulliken approximated
+formulae in Eqs.~\eqref{e:v0-exch.mulliken.working} and \eqref{e:v0-ct.mulliken.working}, respectively,
+as well as $V^{\rm ET1}$, $V^{\rm ET2}$,
+$V^{{\rm HT1},(0)}$ and $V^{{\rm HT2},(0)}$ are treated by the generalized density fitting
+with OEP effective matrices as effective parameters spanned in auxiliary AO basis set 
+(Eqs.~\eqref{e:v0-et1-oep}, \eqref{e:v0-et2-oep}, \eqref{e:v0-ht1-oep}, \eqref{e:v0-ht2-oep}
+and Eqs.~\eqref{e:a:voep}-\eqref{e:a:aoep}). Note that evaluation
 of Eqs.~\eqref{e:ovrl-direct} and \eqref{e:ovrl-indirect} 
-is computationally inexpensive, we shall consider the overlap\hyp{}corrected
-Hamiltonian in the present work.
+is computationally inexpensive because it involves computing only approximate overlap
+integrals between basis states;
+ \item The relaxation effects due to the intermolecular interactions are not included
+in the current OEP\hyp{}based model.
+Note that in the original TI/CIS model the basis states and all the one\hyp{}particle density matrices are built from the
+self\hyp{}consistently adjusted fragment densities according to the DFI method.
+\end{enumerate}
+%
+%\paragraph*{Overlap corrections to the Hamiltonian.}
+%Since in the TI/CIS method the basis states are constructed from the monomer wavefunctions,
+%the overlap effects should be properly taken into account. It was reported previously
+%that overlap correction to the $\left< \Phi_1 \vert \mathscr{H} \vert \Phi_2 \right>$ 
+%matrix element brings negligible contributions to the 
+%total EET coupling (c.f. the `overlap' contribution in Ref.). On the other hand,
+%the importance of the overlap effects on the other off\hyp{}diagonal Hamiltonian matrix elements were not directly studied. 
+%After the overlap correction, the constituents of the direct coupling take the form
+%
+%Since evaluation
+%of Eqs.~\eqref{e:ovrl-direct} and \eqref{e:ovrl-indirect} 
+%is computationally inexpensive, we shall consider the overlap\hyp{}corrected
+%Hamiltonian in the present work.
 
@@
                 \sum_{\varepsilon\in X}^{\rm RI}
-                [{\bf R}^{-1}]_{\xi\xi'} R_{\xi'\varepsilon} H^X_{\varepsilon, M} \;,
+                [{\bf R}^{-1}]_{\xi\xi'} R_{\xi'\varepsilon} H^{X;M}_{\varepsilon; N} \;,
 \end{equation}
@@
 \begin{equation}
- H^X_{\varepsilon, M} = \sum_{\varepsilon'\in X}^{\rm RI}
-                        [{\bf S}^{-1}]_{\varepsilon\varepsilon'} a_{\varepsilon'}^{X; M} \;.
+ H^{X;M}_{\varepsilon; N} = \sum_{\varepsilon'\in X}^{\rm RI}
+                        [{\bf S}^{-1}]_{\varepsilon\varepsilon'} a_{\varepsilon';N}^{X; M} \;.
 \end{equation}
@@
 %
-The elements of vector ${\bf a}^{X;M}$ for $M=$ ET1, ET2, HT1 and HT2 and $X=A, B$ are given by
+The elements of vector ${\bf a}^{X;M}_N$ for $M=$ ET1, ET2, HT1 and HT2 and $N=$ H, L and HL are given by
 %
@@
 \begin{align}
-a_{\alpha}^{A;{\rm ET1}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \nonumber \\
+%
+a_{\alpha;L}^{X;{\rm ET}} &= \sum_\beta C_{\beta L}^X F_{\alpha\beta}^X \;,\\
+%
+a_{\alpha;HL}^{X;{\rm ET}} &= a_{\alpha;L}^{X;{\rm ET}} \nonumber \\
  &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
- \left\{ 2C^A_{\beta  H} C^A_{\gamma L} - C^A_{\beta  L} C^A_{\gamma H}\right\} C^A_{\delta H} \;,\\
+ \left\{ 2C^X_{\beta  H} C^X_{\gamma L} - C^X_{\beta  L} C^X_{\gamma H}\right\} C^X_{\delta H} \;,\\
 %
-a_{\alpha}^{B;{\rm ET1}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
+a_{\alpha;H}^{X;{\rm HT}} &=-\sum_\beta C_{\beta H}^X F_{\alpha\beta}^X \;,\\
 %
-a_{\alpha}^{B;{\rm ET2}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \nonumber \\
+a_{\alpha;HL}^{X;{\rm HT}} &=a_{\alpha;H}^{X;{\rm HT}} \nonumber \\
  &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
- \left\{ 2C^B_{\beta  H} C^B_{\gamma L} - C^B_{\beta  L} C^B_{\gamma H}\right\} C^B_{\delta H} \;,\\
+ \left\{ 2C^X_{\beta  L} C^X_{\gamma H} - C^X_{\beta  H} C^X_{\gamma L}\right\} C^X_{\delta L} \;.
 %
-a_{\alpha}^{A;{\rm ET2}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;,\\
-%
-a_{\alpha}^{A;{\rm HT1}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \nonumber \\
- &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
- \left\{ 2C^A_{\beta  L} C^A_{\gamma H} - C^A_{\beta  H} C^A_{\gamma L}\right\} C^A_{\delta L} \;,\\
-%
-a_{\alpha}^{B;{\rm HT1}} &=-\sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
-%
-a_{\alpha}^{B;{\rm HT2}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \nonumber \\
- &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
- \left\{ 2C^B_{\beta  L} C^B_{\gamma H} - C^B_{\beta  H} C^B_{\gamma L}\right\} C^B_{\delta L} \;,\\
-%
-a_{\alpha}^{A;{\rm HT2}} &=-\sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;.
+%a_{\alpha}^{A;{\rm ET1}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \nonumber \\
+% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+% \left\{ 2C^A_{\beta  H} C^A_{\gamma L} - C^A_{\beta  L} C^A_{\gamma H}\right\} C^A_{\delta H} \;,\\
+%%
+%a_{\alpha}^{B;{\rm ET1}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
+%%
+%a_{\alpha}^{B;{\rm ET2}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \nonumber \\
+% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+% \left\{ 2C^B_{\beta  H} C^B_{\gamma L} - C^B_{\beta  L} C^B_{\gamma H}\right\} C^B_{\delta H} \;,\\
+%%
+%a_{\alpha}^{A;{\rm ET2}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;,\\
+%%
+%a_{\alpha}^{A;{\rm HT1}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \nonumber \\
+% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+% \left\{ 2C^A_{\beta  L} C^A_{\gamma H} - C^A_{\beta  H} C^A_{\gamma L}\right\} C^A_{\delta L} \;,\\
+%%
+%a_{\alpha}^{B;{\rm HT1}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \;,\\
+%%
+%a_{\alpha}^{B;{\rm HT2}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \nonumber \\
+% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+% \left\{ 2C^B_{\beta  L} C^B_{\gamma H} - C^B_{\beta  H} C^B_{\gamma L}\right\} C^B_{\delta L} \;,\\
+%%
+%a_{\alpha}^{A;{\rm HT2}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \;.
 \end{align}
@@
 %
+Here, the Fock matrix in space spanned by the intermediate and primary AO basis sets
+needs to be computed. The respective contributions to ${\bf a}$ vectors can be
+written in a convenient to implement form as
+%
+\begin{multline}
+ \sum_\beta C_{\beta U}^X F_{\alpha\beta}^X = 
+ \sum_\beta C_{\beta U}^X h_{\alpha\beta}^X \\
+   + \sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
+     \left[ P_{\delta\gamma}^{X(g)} C_{\beta U}^X - \frac{1}{2} P_{\beta\gamma}^{X(g)} C_{\delta U}^X \right] \;,
+\end{multline}
+%
+where ${\bf h}^X$ is the one\hyp{}electron core Hamiltonian of molecule $X$.
+
 % -----------------------

### 2019-08-28 oepdev/liboep/oep.h
OEP types for ET and HT implemented

@@ -329,14 +329,7 @@ class ChargeTransferEnergyOEPotential : public OEPotential
 /**\brief Generalized One-Electron Potential for EET coupling calculations.
  * 
  *  Contains the following OEP types:
- *    - `Fujimoto.A.ET1` 
- *    - `Fujimoto.B.ET1` 
- *    - `Fujimoto.A.ET2` 
- *    - `Fujimoto.B.ET2` 
- *    - `Fujimoto.A.HT1` 
- *    - `Fujimoto.B.HT1` 
- *    - `Fujimoto.A.HT2`
- *    - `Fujimoto.B.HT2`
+ *    - `Fujimoto.GDF` - Joint OEP type for ET(L), ET(HL), HT(H) and HT(HL)
  */
 class EETCouplingOEPotential : public OEPotential 
 {
@@ -356,6 +349,8 @@ class EETCouplingOEPotential : public OEPotential
     void common_init();
 
     /// Auxiliary computers
+    void compute_fujimoto_gdf();
+
 
 };
 

### 2019-08-28 oepdev/liboep/oep_coupling_eet.cc

@@ -31,30 +31,118 @@ void EETCouplingOEPotential::common_init()
     int n2 = auxiliary_->nbf();
     int n3 = wfn_->molecule()->natom();
 
-    // Implement these matrices (here provide correct dimensions)
-    SharedMatrix mat_1 = std::make_shared<psi::Matrix>("ET1-NotImplemented", 1, 1);  
-    SharedMatrix mat_2 = std::make_shared<psi::Matrix>("ET2-NotImplemented", 1, 1);  
-    SharedMatrix mat_3 = std::make_shared<psi::Matrix>("ET1-NotImplemented", 1, 1);  
-    SharedMatrix mat_4 = std::make_shared<psi::Matrix>("ET2-NotImplemented", 1, 1);  
-    SharedMatrix mat_5 = std::make_shared<psi::Matrix>("ET1-NotImplemented", 1, 1);  
-    SharedMatrix mat_6 = std::make_shared<psi::Matrix>("ET2-NotImplemented", 1, 1);  
+    // OEP Matrices
+    SharedMatrix mat_gdf = std::make_shared<psi::Matrix>("ET and HT", n2, 4);  
 
     // Provide correct classes (DF- or ESP-based) and number of OEP's involved in each type
-    OEPType type_1 = {"Fujimoto.ET1", false,  0, mat_1};
-    OEPType type_2 = {"Fujimoto.ET2", false,  0, mat_2};
-    OEPType type_3 = {"Fujimoto.HT1", false,  0, mat_3};
-    OEPType type_4 = {"Fujimoto.HT2", false,  0, mat_4};
-    OEPType type_5 = {"Fujimoto.CT1", false,  0, mat_5};
-    OEPType type_6 = {"Fujimoto.CT2", false,  0, mat_6};
+    OEPType type_1 = {"Fujimoto.GDF", true, 4, mat_gdf};
 
+    // Register OEPType's
     oepTypes_[type_1.name] = type_1;
-    oepTypes_[type_2.name] = type_2;
-    oepTypes_[type_3.name] = type_3;
-    oepTypes_[type_4.name] = type_4;
-    oepTypes_[type_5.name] = type_5;
-    oepTypes_[type_6.name] = type_6;
 }
 
-void EETCouplingOEPotential::compute(const std::string& oepType) {}
+void EETCouplingOEPotential::compute(const std::string& oepType) 
+{
[…]

### 2019-11-05 papers/2_oep_technique/paper-r1.tex
Paper II revision

--- 2019-07-11 papers/2_oep_technique/paper-r0.tex
@@
 
-\section{\label{s:1}Introduction}
-
-\section{\label{s:2}Effective One-Electron Potentials: General Theory}
-
+%\section{\label{s:1}Introduction}
+
+\section{\label{ss.5.2}Introduction}
+
+Charge transfer (CT) between molecules occurs when the net electronic populations 
+of interacting molecules change which leads to an additional stabilization 
+of a molecular aggregate. The effect of CT on the interaction energy 
+is known to be often non\hyp{}negligible, especially
+in donor\hyp{}acceptor systems such as H\hyp{}bonded species
+and charged complexes.
+\emph{Ab initio} fragmentation methods, such as
+the second generation of the effective fragment potential method (EFP2),
+are often used to describe the homogeneous extended molecular systems 
+such as solutions
+and recently even biomolecules. EFP2 relies on the intermolecular 
+perturbation theory, and the total interaction energy is approximated as
+a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
+charge densities of the monomers, exchange\hyp{}repulsion energy,
+induction energy due to mutual polarization, intermolecular dispersion energy,
+as well as the CT energy,
+%
+\begin{equation}
+ U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
+\end{equation}
+%
+However, calculating the CT stabilization energy is usually much more expensive 
+than evaluating the contributions due to other effects.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
+%However, the computational cost of the CT term is usually at least few times larger
+%than the next costly contribution $U^{\rm Ex-Rep}$, which is still too costly
+%to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$. 
+In effect,
+this part of the EFP2 energy is usually ignored in most of the applications. In the 
+recent LIBEFP library, CT term were even not implemented due to those reasons.
+
+In this work, we propose an alternative formulation of the CT energy compatible with the
+EFP2 method. We start from the perturbation theory of Murrell et al.
+and
+introduce the effective one\hyp{}electron potential (OEP) operators
+to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
+Later we validate the model against the EFP2 method in Section XXX. Finally,
+we conclude this work and provide few remarks on future development in Section XXX.
+
+\section{\label{s:2}Effective One-Electron Potential Operators}
+
+To establish the notation within the present work, 
+the formalism of one\hyp{}electron potential (OEP) operators is
+given first.
 The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
@@
 
-\section{\label{s:3}Effective One-Electron Potentials: Configuration Interaction}
-
-The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
-electronic state
-of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
-%
-\begin{equation} \label{e:schrodinger}
- E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
-\end{equation}
-%
-for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
-with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator,
-$\hat{h}(1) = \hat{v}_{\rm nuc} -\frac{1}{2}\sum_e^{\rm el} \nabla_e^2$,
-and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
-From the above solution, any property $P$, understood as an expectation value of the
-associated operator $\mathscr{P}$, can be computed according to
-%
-\begin{equation} \label{e:prop}
- P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
-\end{equation}
-%
-The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
-the configuration interaction (CI) expansion around the reference 
-(approximate) wavefunction, $\Psi^{(0)}$,
-typically the solution to the Hartree\hyp{}Fock (HF) problem\cite{Roothaan.RevModPhys.1951},
-%
-\begin{equation} \label{e:ci}
- \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
-	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
-\end{equation}
-%
-In the above equation,
-$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
-with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
-one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
-virtual (unoccupied) molecular orbital, obtained from the HF solution.
-By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
-and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
-operator matrix element between
-the CI configurations, it can be shown that the following holds
-%
-\begin{equation} \label{e:exp-val-series}
-	P =
-	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
-	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
-\end{equation}
-%
-%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
-%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
-%%
-%\begin{equation}
-%	\BraKet{ij}{kl} \equiv
-%	\iint 
-%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
-%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
-%	d{\bf r}_1 d{\bf r}_2  \;,
-%\end{equation}
-%%
-where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
-and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
-are well defined and characteristic for the property of interest.
-In fact, they are certain functions of the CI coefficients, i.e.,
-%
-\begin{equation}
-{\bf P} = {\bf P}
-\left(
- \left\{
-  C_{r}^{a}, C_{rs}^{ab}, \ldots 
- \right\} 
-\right)
-\end{equation}
-%
-and, in principle, are known as long as the solution to the CI problem 
-or its approximation is available.
-
-%In this work we are interested in cases when
-
-%$\Psi^{AB}$. In this Section we consider theories to calculate the effect of the interaction
-%of molecules $A$ and $B$ on the property $P$ knowing only the unperturbed wavefunctions,
-%$\Psi^{A}$ and $\Psi^{B}$.
-
-\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property}
-
-Consider a molecular aggregate consisting of two interacting 
-molecules $A$ and $B$. The part of property that arises due to the
-intermolecular interaction between $A$ and $B$ can be calculated
-based on the supermolecular approach,
-%
-\begin{multline} \label{e:delta-p}
-	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \\- 
-	\left(
-	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} +
-	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}
-	\right)
-	\;,
-\end{multline}
-%
-where $\Psi^{AB}$ and $\Psi^{A(B)}$ are the wavefunctions of the interacting molecular aggregate
-and the unperturbed (non\hyp{}interacting) molecules, respectively.
-The notation $\Delta^{AB}$ emphasizes on the two\hyp{}body character of the interaction\hyp{}induced
-property. In this work we focus on re\hyp{}expressing $\Delta^{AB}P$ in terms of separate
-fragments associated with either molecule. In general case when the basis set
-is complete, such an operation is not possible. However,
-assuming an approximation
-in which the basis functions $\phi_i$ 
-are localized on either molecule it allows one to partition the basis set space
-and the resulting 2\hyp{} and 4\hyp{}index tensor elements
-into subsets `belonging' either to molecule $A$ or to molecule $B$.
-Such an approximation is essentially a foundation of any fragment\hyp{}based strategy
-and, in particular, designing \emph{ab initio} force fields such as EFP2
-of SolEFP.
-
-Application of Eq.~\eqref{e:exp-val-series}
-to Eq.~\eqref{e:delta-p} and adopting the localized basis set approximation
-yields the partitioning of the interaction\hyp{}induced property,
-%
-\begin{equation} \label{e:delta-p-part}
- \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
-\end{equation}
-%
-for which the one\hyp{}electron part reads
-%
-\begin{multline} \label{e:delta-p-part-1el}
- \Delta^{AB} P(1) \approx
-	\sum_{ij} \Big\{ 
-	P^{AA}_{ij}  \tBraKet{A}{\hat{h}}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
-	P^{BB}_{ij}  \tBraKet{B}{\hat{h}}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
-2	P^{AB}_{ij}  \tBraKet{A}{\hat{h}}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
-	\Big\}
-\end{multline}
-%
-and the two\hyp{}electron part is given by
-%
-\begin{multline} \label{e:delta-p-part-2el}
- \Delta^{AB} P(2) \approx
-	\sum_{ijkl} \Big\{ 
-	P^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
-	P^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
-	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
-	{s} \left[ P^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
-	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
-	{s} \left[ P^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
-	\Big\} \;.
-\end{multline}
-%
-To simplify the resulting expressions
-in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
-the shorthand notation $\tBraKet{A}{\hat{h}}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
-$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
-was introduced, and
-the real orbitals 
-were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
-The operator $s$ permutes and sums over all the combinations of the basis function indices
-given their partitioning scheme between molecules $A$ and $B$.
-All the summation components from Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el} 
-can be grouped into the three
-sub\hyp{}categories: (i) Coulomb\hyp{}like terms
-with integrals of the type $\BraKet{AA}{BB}$, (ii) overlap\hyp{}like terms with $\BraKet{AA}{AB}$
-and $\BraKet{AB}{BB}$ and (iii) exchange\hyp{}like terms with $\BraKet{AB}{AB}$.
-
-%Therefore, the interaction\hyp{}induced property can be recast in a following way
+%\subsection{\label{s.334}Practical OEP-Based Calculations}
 
@@
 
-\subsection{\label{ss:2.4}Definition of Interaction-Induced Property Effective Potentials}
-
-The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
-under the localized basis set approximation 
-from Eq.~\eqref{e:delta-p-part}.
-%Coulomb\hyp{}like and overlap\hyp{}like 
-%terms from Eq.~\eqref{e:delta-p-part-2el}. 
-For example, the Coulomb\hyp{}like contributions
-can be rewritten without making any further approximation as
-%
-\begin{equation}
- \sum_{ijkl} 
-	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
- \equiv
-	\sum_{kl\in B} 
-	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
-\end{equation}
-%
-with
-%
-\begin{equation}
- \hat{v}^{{\rm eff},A} 
-	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
-\end{equation}
-%
-Similarly for the overlap\hyp{}like term we have
-%
-\begin{equation}
- \sum_{ijkl} 
-	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
- \equiv
-	\sum_{k\in A} \sum_{l\in B} 
-	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
-\end{equation}
-%
-and the associated effective potential reads
-%
-\begin{equation}
- \hat{v}^{{\rm eff},A} 
-	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
-\end{equation}
-%
-Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
-Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
-%
-\begin{multline} \label{e:ci-oep-master}
- \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
-	+ \sum_{i\in A} \sum_{j\in B} 
-	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
-	 \\ +
-	\sum_{i\in A} \sum_{k\in A}
-	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
-	 +
-        \sum_{j\in B} \sum_{l\in B}
-          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
-\end{multline}
-%
-where the exchange\hyp{}like component is
-%
-\begin{equation}
-	\Delta_{\rm ex}^{AB} P = 
-	s \sum_{ik\in A} \sum_{jl\in B}
-	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
-\end{equation}
-%
-and
-the overlap\hyp{}like and the Coulomb\hyp{}like 
-OEP's are defined as follows
-%
-\begin{subequations}
-\begin{align} \label{e:oep-exch-like}
-	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
-	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \;, \\
-	%
-	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
-        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
-        \\ \label{e:oep-coul-like}
-	& \qquad\qquad+ \frac{1}{2}
-        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
-\end{align}
-\end{subequations}
-%
-and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.
-
-\subsection{\label{s.242}Separability of P-Coefficients}
-
-Until that point, the effective potentials from Eqs.~\eqref{e:oep-exch-like}
-and \eqref{e:oep-coul-like} are not separable in terms of the interacting molecules
-because the $\bf P$ tensors are derived from the CI wavefunction of
-the entire interacting system $AB$. However, if there exist an approximate
-separation as, for example,
-%
-\begin{align}
- P^{AAAB}_{mkij} &\approx X^A_{mk} Y^{AB}_{ij} \;, \\
- P^{AABB}_{ikjl} &\approx Z^A_{ik} Y^{AB}_{} Z^{ B}_{jl} \;,
-\end{align}
-%
-and if the matrices ${\bf X}^{A(B)}$, 
-${\bf Z}^{A(B)}$ and ${\bf P}^{AA(BB)}$ as well as the tensors
-${\bf P}^{AAAA(BBBB)}$ 
-are properties of the unperturbed (isolated) molecules,
-further rearrangements in Eq.~\eqref{e:ci-oep-master} are possible
-leading to computationally more efficient expressions.
-
-Note also that the matrices ${\bf Y}^{AB}$ and ${\bf P}^{AB}$ depend on their instantaneous 
-relative positions.
-
-%then the molecule\hyp{}specific OEP's can be defined.
-
-\section{\label{s.334}Practical OEP-Based Calculations}
-
-There are two general cases of basis function partitioning scheme
+Two general cases of basis function partitioning scheme
+are considered in this work,
 in which it is possible to define an effective one\hyp{}electron
@@
 %
-The above double integral can be approximated by the
+The above integral can be approximated by the
 distributed multipole (DMTP)
@@
 
-\section{\label{s:4}Calculation Details}
-
-All the models that were used to test the theory presented in this work
-were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-
-\section{\label{s:5}Results and Discussion}
-\subsection{\label{ss.5.1}Pauli-Repulsion Interaction Energy}
-
-At short intermolecular interactions, wavefunction overlap increases
-which causes the electrons from either molecules to move apart to minimize the repulsive interactions.
-The resulting interaction energy can be estimated based on perturbation theory approaches.
-In this section, we demonstrate the use of OEP operators to significanlty reduce the computational
-cost of such calculations, that can be used in fragment\hyp{}based modelling of extended molecular
-aggregates in terms of force field methods.
-
-As a starting point we consider perturbation theory model of Murrell et al.,
-in which repulsive energy is originally given by
-%
-\begin{equation}
- \Delta^{AB} E^{\rm Rep} = 
- \Delta^{AB} E^{\rm Rep}(S^{-1}) + 
- \Delta^{AB} E^{\rm Rep}(S^{-2}) \;,
-\end{equation}
-%
-where the first\hyp{}order term with respect to the intermolecular overlap is
-%
-\begin{multline} \label{e:rep.murrell-etal.S1}
-  \Delta^{AB}  E^{\rm Rep}(S^{-1}) = -2\sum_{i\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}
-               S_{ij} \Big\{
-           V^A_{ij} 
-         + V^B_{ji} \\
- + \sum_{k\in A}^{\rm Occ} \left[ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right] 
- + \sum_{l\in B}^{\rm Occ} \left[ 2\BraKet{ij}{ll} - \BraKet{il}{jl} \right]
-                \Big\}
-\end{multline}
-%
-and the second\hyp{}order term is given in Eq.(XXX) in Ref. %cite
-In this work, we focus to simplify only the former term because this is usually
-the bottleneck of evaluating the Pauli repulsive contributions to
-forces in molecular dynamics when using the effective potential method, whereas
-very inexpensive approximation to the latter term has already been worked out previously. %cite
-Note that the Coulomb and exchange integrals can be re\hyp{}cast as follows:
-%
-\begin{subequations}
- \begin{align}
- \BraKet{ij}{kk} &\equiv \sum_{\mu\in A} 
-     C_{\mu i}^A \tBraKet{\mu}{\hat{v}_{kk}}{j} \;, \\
- \BraKet{ik}{jk} &\equiv \sum_{\mu\in A} 
-     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
- \end{align}
-\end{subequations}
-%
-%Similarly, the exchange integrals reduce to
-%
-%\begin{equation}
-% \BraKet{ik}{jk} \equiv \sum_{\mu\in A} 
-%     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
-%\end{equation}
-%
-In the above equations, the auxiliary potential operators are given by
-%
-\begin{equation}
-  \hat{v}_{ik} \equiv \int d{\bf r} \Ket{{\bf r}} 
-        \left[
-        \int d{\bf r}' \frac{\phi_i^{*}({\bf r}') \phi_k({\bf r}')}{\vert {\bf r}' - {\bf r}\vert}
-        \right] \Bra{{\bf r}} \;.
-\end{equation}
-%
-Now, using the prescription in Eq.~\eqref{e:ft-reduction} one can define a joint OEP
-operator from nuclear, Coulomb and exchange parts as
-%
-\begin{equation}
- V^A_{ij} + 
- \sum_{k\in A}^{\rm Occ} 
-  \left\{ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right\}
-\equiv \sum_{\mu\in A} \tBraKet{\mu}{ 
-\hat{v}^{A[\mu i]}_{\rm eff}
- }{j}
-\end{equation}
-%
-with
-%
-\begin{equation}
- \hat{v}^{A[\mu i]}_{\rm eff} \equiv C_{\mu i} \hat{v}^A_{\rm nuc} + 
- \sum_{k\in A}^{\rm Occ} \left[
- 2C_{\mu i}^A \hat{v}^A_{kk} - C_{\mu k}^A \hat{v}^A_{ik}
- \right] \;.
-\end{equation}
-%
-%The nuclear part can also be incorporated within the effective potential
-%resulting in the following reduction
-%
-%
-On the other hand, it immediately follows that
-%
-\begin{equation}
- \hat{v}^{A[\mu i]}_{\rm eff} \Ket{\mu} = 
-  \sum_{k\in A}^{\rm Occ} \left\{
-     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
-  \right\} \;.
-\end{equation}
-%
-For practical calculations, the right hand side of the above equation can be expanded 
-in the auxiliary basis,
-%
-\begin{equation} \label{e:v-oep.rep}
-  \sum_{k\in A}^{\rm Occ} \left\{
-     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
-  \right\} \cong
-  \sum_{\xi\in A}^{\rm DF} 
-  V_{\xi i}^A \Ket{\xi} \;,
-\end{equation}
-%
-where the matrix ${\bf V}^A$ can be considered as effective fragment parameters.
-Doing the same operations on the twin operators associated with the molecule $B$
-original theory of Murrell et. al from Eq.~\eqref{e:rep.murrell-etal.S1}
-reduces to
-%
-\begin{equation}
-  \Delta^{AB}  E^{\rm Rep}(S^{-1}) \cong 
- -2\sum_{i\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}
-               S_{ij} \Big\{
-           \sum_{\xi \in A}^{\rm DF} V_{\xi i}^A S_{\xi j}
-         + \sum_{\eta\in B}^{\rm DF} V_{\eta j}^B S_{\eta i}
-                \Big\}
-\end{equation}
-%
-which is a fully OEP\hyp{}based result. Note, that now only 
-overlap integrals, which are relatively computationally very inexpensive,
-need to be evaluated.
-
-As another benchmark model for accuracy, we compare our results with 
-the exchange and repulsion energy term 
-from the density\hyp{}based scheme of partitioning the interaction energy
-(here abbreviated as DDS),
-that was developed by Mandado and Hermida-Ramon. 
-%In the above equations, the Pauli deformation density 
-%%
-%\begin{equation}
-%   \Delta {\bf D}^{\rm Pauli} \equiv {\bf D}^{oo} - {\bf D}
-%\end{equation}
-%%
-%where \f$ {\bf D}^{oo}\f$ and \f$ {\bf D}\f$ are the density matrix formed from
-%mutually orthogonal sets of molecular orbitals within the entire aggregate (formed
-%by symmetric orthogonalization of MO's) and the density matrix of the unperturbed
-%system (that can be understood as a Hadamard sum \f$ {\bf D} \equiv {\bf D}^A \oplus {\bf D}^B\f$).
-% 
-%At HF level, the Pauli deformation density matrix is given by
-%\f[
-%  \Delta {\bf D}^{\rm Pauli} = {\bf C} \left[ {\bf S}^{-1} - {\bf 1} \right] {\bf C}^\dagger
-%\f]
-%whereas the density matrix constructed from mutually orthogonal orbitals is
-%\f[
-%   {\bf D}^{oo} = {\bf C} {\bf S}^{-1} {\bf C}^\dagger
-%\f]
-%In the above equations, \f$ {\bf S} \f$ is the overlap matrix between doubly occupied molecular orbitals
-% of the entire aggregate.
-It is also mentioned here that the DDS model of exchange-repulsion at HF level of theory
-is equivalent to the perturbative method of Hayes and Stone and the two methods
-yield the same results, although the exchange and repulsive interaction energies
-are slightly different.
-
-\subsubsection{\label{ssdff.sd}Defining OEP's for Repulsive Potential}
-
-\subsection{\label{ss.5.2}Charge-Transfer Interaction Energy}
-
-Charge transfer (CT) between molecules occurs when the net electronic populations 
-of interacting molecules change which leads to an additional stabilization 
-of a molecular aggregate. Since the effect of CT on the interaction energy 
-is known to be often non\hyp{}negligible, it is quite important to model 
-CT energies with reasonable accuracy and efficiency when performing energy 
-and molecular dynamics calculations with \emph{ab initio} force fields, 
-especially in donor\hyp{}acceptor systems such as H\hyp{}bonded species. 
-However, estimating the CT stabilization energy is usually much more expensive 
-than estimating the contributions due to other effects such as electrostatics, 
-exchange\hyp{}repulsion and dispersion.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
-
-As an illustration of application of the OEP technique for efficient calculation 
-of CT energy we consider the second\hyp{}order perturbation theory of Murrell et al.,
-in which the energy associated with the charge transfer from molecule $A$ to $B$ is given by
+\section{\label{s:34r}Charge Transfer Interaction Energy for Fragment Potentials}
+
+\subsection{\label{s:34rg} Murrell et al.'s Model: Starting Point}
+
+After the OEP technique has been formulated, it can be applied 
+to the calculation of interaction\hyp{}induced properties following the
+fragment\hyp{}based philosophy. 
+According to the second\hyp{}order intermolecular perturbation theory of Murrell et al.,
+the energy associated with the charge transfer from molecule $A$ to $B$ is given by
 %
@@
 %
-In the above expression, the following effective potentials,
+In the above expression, the following effective one\hyp{}electron operators,
 %
@@
 were introduced without making any approximation to the original equation
-from Ref. One can immediately notice that the five summation terms
+from Ref. 
+The goal of this work is to apply the technique sketched in Eq.XXX
+to eliminate ERI's from Eq.~\eqref{e:ct-murell-etal.vin} in terms of OEP's.
+
+\subsection{\label{s:34rag} Murrell et al.'s Model: Application of OEP Technique}
+
+One can immediately notice that the five summation terms
 from Eq.~\eqref{e:ct-murell-etal.vin} can be classified into three groups
@@
 %\end{multline}
-%%
+%
+
+\section{\label{s:4}Calculation Details}
+
+All the models that were used to test the theory presented in this work
+were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
 
@@
            \BraKet{\alpha\beta}{\gamma\delta} \;,\\
-   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{i} \;.
-     &=
+   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{n} 
+     &= \sum_{x}^{\rm At} W_{\alpha n}^{(x)} \nonumber   \\
+       + \sum_{\beta\gamma\delta} &
+           \left\{
+             2 C_{\beta n} D_{\gamma\delta} - C_{\gamma n} D_{\beta \delta}
+           \right\}
+           \BraKet{\alpha\beta}{\gamma\delta}
+    \;.
  \end{align}

### 2019-11-05 papers/2_oep_technique/do-r1.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+base=paper-r1
+
+latex  ${base}.tex
+bibtex ${base}.aux
+dvipdf ${base}.dvi

### 2019-11-11 papers/2_oep_technique/paper-r1.tex
Removing derivation errors in the group II and III of OL and OEP-OL expressions

@@
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
-      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
+      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right) 
-        \BraKet{nj}{ik} \;.
+        S_{kj} \BraKet{nj}{ik} \;.
 \end{multline}
@@
  \hat{v}^B_{\rm tot} &= \hat{v}^B_{\rm nuc} + 2\sum_{j\in B}^{\rm Occ} \hat{v}^B_{jj} \quad\text{ and }\\ 
- \hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
+%\hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
 \end{align}
@@
 weighted by the overlap integrals $S_{nk}$. Using the distributed multipole 
-expansion this group can be written as
+expansion based on the charge centroids of the localized molecular orbitals (LMO's),
+${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$ with $\chi_i({\bf r})$ being localized,
+this group can be written as
 %
 \begin{equation} \label{e:ct.group-2}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
+\end{equation}
+%
+Note that, in general, $\rho_{ki}^A$ is complex whereas $\rho^B_{\rm tot}$ is always real.
+Since the interaction energy must also be real, it must hold that
+%
+\begin{equation} \label{e:ct.group-2-d}
+ \rho_{ki}^A \odot \rho^B_{\rm tot} = \delta_{ik} \rho_{ii}^A \odot \rho^B_{\rm tot} \;.
 \end{equation}
@@
 Consider now a situation in which the localized occupied MO's are used.
-If the multipole series expansions are truncated
-at the distributed charges $q_i$ and dipole moments ${\BM{\upmu}_i}$, 
-located at the centroids ${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$, 
-then the interaction energy becomes
+In this case, the distributed multipole expansion
+of the above becomes
 %
 \begin{multline}
- \rho_{ki}^A \odot \rho^B_{\rm tot} 
+ \rho_{ii}^A \odot \rho^B_{\rm tot} 
  \approx 
- q_{ki} 
+ q_{i} 
  \left[
@@
  \right] \\
- + \text{ dipole terms } \;.
+ + \text{ quadrupole terms } \;,
 \end{multline}
 %
-Note however that $q_{ik} = -\delta_{ik}$ and $q_j = -1$. Moreover, 
-if the localized orbitals are perfectly spherical distributed dipole moments vanish.
+because the distributed charges
+$q_i = -1$ whereas the distributed dipole moments %${\BM{\upmu}_i}$ 
+centered at their
+respective LMO charge centroids vanish.
 This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows:
@@
 \begin{equation} \label{e:ct.group-ii.final}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \approx S_{ni}  \left[
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \approx 2S_{ni}  \left[
   %\sum_{i\in A}^{\rm LC} 
@@
 \begin{multline}
-       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right) 
+        S_{kj} 
         \BraKet{nj}{ik} \\ = -
@@
  \left\{ 
-  \delta_{ik} \left( \hat{v}^A_{k} + \hat{v}^A_{ik} \right)
-  - \hat{v}^A_{ik}
+  \delta_{ik} \hat{v}^A_{\rm tot} - \hat{v}^A_{ik}
  \right\} 
@@
  \delta_{ik} 
- \left[
-  v^A_{\rm tot} ({\bf r}) - 2 v^A_{kk} ({\bf r}) + v^A_{ik} ({\bf r})
- \right] 
-  - v^A_{ik} ({\bf r}) \;.
+ {v}^A_{\rm tot} ({\bf r})- {v}^A_{ik} ({\bf r}) \;.
 \end{equation}
@@
 %
-\begin{multline} \label{e:ct.group-iii.final.a}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+\begin{equation} \label{e:ct.group-iii.final.a}
+        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right) 
-        \BraKet{nj}{ik}  \\
- \approx 0 \;.
-\end{multline}
+        S_{kj} 
+        \BraKet{nj}{ik} \approx 0 \;.
+\end{equation}
 %
@@
 In this work, the effective charges were computed via the CAMM or ESP methods as
-discussed in Section~XXX. Applying the SOA 
+discussed in Section~XXX. 
+Note now that both effective densities in 
+$\rho^B_{jn} \odot \rho^{A,{\rm eff}}_{ik}$ are generally complex and one cannot apply the SOA
+for $\rho^{A,{\rm eff}}_{ik}$. In the LMO picture however, it can be shown that
+$\rho_{ik} ({\bf r}) \ll \rho_{ii} ({\bf r})$ 
+for $i\ne k$ in most locations. Therefore,
+the SOA is still applicable in the LMO picture
 and the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
+simplifies to
+%
+\begin{equation} \label{e:ct.group-iii.oep.simplified}
+ {v}^{A,{\rm eff}}_{ik}({\bf r}) \approx
+ \delta_{ik} \left\{
+ {v}^A_{\rm tot} ({\bf r})- {v}^A_{ii} ({\bf r}) \right\} \;,
+\end{equation}
+%
+which
 leads to 
@@
 \begin{multline} \label{e:ct.group-iii.final.b}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right) 
+        S_{kj}
         \BraKet{nj}{ik}  \\
@@
    \frac{Z_x}{r_{xy}}
-  + \frac{2}{r_{iy}}
+  + \frac{1}{r_{iy}}
   - \sum_{k\in A}^{\rm Occ}
@@
  +
-S_{ni} u_i^{BA} \;, \\
+2 S_{ni} u_i^{BA} \;, \\
 %

### 2019-11-13 papers/2_oep_technique/paper-r2.tex
Second revision of CT paper added

--- 2019-11-11 papers/2_oep_technique/paper-r1.tex
@@
 
-\title{Ab Initio Effective One-Electron Potentials Revisited:
-General Theory and Example Applications for Charge-Transfer Energy Calculations}
+\title{Ab Initio Effective One-Electron Potential Operators:
+Applications for Charge-Transfer Energy in Effective Fragment Potentials}
 
@@
 molecular aggregates such as interacting molecules in condensed phases.
-%In this Work, \emph{ab initio} effective potentials are studied for a number of theories
-%of the intemolecular interaction energy.
 \end{abstract}
@@
 \tableofcontents
-
-%\section{\label{s:1}Introduction}
 
@@
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
-      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
+      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \BraKet{nj}{ik} \;.
+        S_{kj} \left(1+\delta_{ik}\right) \BraKet{nj}{ik} \;.
 \end{multline}
@@
  \hat{v}^B_{\rm tot} &= \hat{v}^B_{\rm nuc} + 2\sum_{j\in B}^{\rm Occ} \hat{v}^B_{jj} \quad\text{ and }\\ 
-%\hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
+ \hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
 \end{align}
@@
 \begin{equation} \label{e:ct.group-2}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
+      - \sum_{k\in A}^{\rm Occ} S_{nk}  \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
 \end{equation}
@@
 \begin{equation} \label{e:ct.group-ii.final}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \left(1+\delta_{ik}\right) \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \approx 2S_{ni}  \left[
+      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \approx S_{ni}  \left[
   %\sum_{i\in A}^{\rm LC} 
@@
 \begin{multline}
-       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
+       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} 
+        S_{kj} \left( 1 - \delta_{ik} \right)
         \BraKet{nj}{ik} \\ = -
@@
  \left\{ 
-  \delta_{ik} \hat{v}^A_{\rm tot} - \hat{v}^A_{ik}
+  \delta_{ik} \left( \hat{v}^A_{k} + \hat{v}^A_{ik} \right)   - \hat{v}^A_{ik}
  \right\} 
@@
  \delta_{ik} 
- {v}^A_{\rm tot} ({\bf r})- {v}^A_{ik} ({\bf r}) \;.
+ \left[
+  v^A_{\rm tot} ({\bf r}) - 2 v^A_{kk} ({\bf r}) + v^A_{ik} ({\bf r})
+ \right] 
+  - v^A_{ik} ({\bf r}) \;.
 \end{equation}
@@
 %
-\begin{equation} \label{e:ct.group-iii.final.a}
+\begin{multline} \label{e:ct.group-iii.final.a}
         - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} 
-        \BraKet{nj}{ik} \approx 0 \;.
-\end{equation}
+        S_{kj} \left( 1 - \delta_{ik} \right)
+        \BraKet{nj}{ik} \\ 
+    \approx 0 \;.
+\end{multline}
 %
@@
  \delta_{ik} \left\{
- {v}^A_{\rm tot} ({\bf r})- {v}^A_{ii} ({\bf r}) \right\} \;,
+ {v}^A_{\rm tot} ({\bf r})- 2{v}^A_{kk} ({\bf r}) \right\} \;,
 \end{equation}
@@
    \frac{Z_x}{r_{xy}}
-  + \frac{1}{r_{iy}}
+  + \frac{2}{r_{iy}}
   - \sum_{k\in A}^{\rm Occ}
@@
  +
-2 S_{ni} u_i^{BA} \;, \\
+S_{ni} u_i^{BA} \;, \\
 %
@@
 were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
+
+\section{\label{s:fwefsdv}Results and Discussion}
+
+%
+\begin{figure}[t]
+\includegraphics[width=0.5\textwidth]{fig-2.eps}
+\caption{\label{f:fig-1} {\bf Accuracy of the OEP approach to estimate the charge transfer energy
+in the PT/HF formulation.} 
+Asymptotic dependencies of the charge transfer energy is plotted for 
+(a) water dimer, 
+(b) water\hyp{}ammonium complex, and 
+(c) ammonium\hyp{}nitrate complex,
+where one molecule has been translated by $R$ from the starting geometry
+along the vector specified in Fig. 1.
+The counterpoise\hyp{}corrected total interaction energy
+is also shown for comparison in light grey color in this figure.
+All data were obtained at HF/6-31+G** level of theory.
+} 
+\end{figure}
+%
 

### 2019-11-13 papers/2_oep_technique/paper-r2.tex
Figure 2 added to the manuscript. Computational details elaborated.

@@
 %
-In this work, the effective charges were computed via the CAMM or ESP methods as
+In this work, the effective charges were computed via the Mulliken method as
 discussed in Section~XXX. 
@@
 %
+\begin{equation} \label{e:ct-murell-etal.vin.oep}
+V_{in}^{A\rightarrow B} 
+       \approx 
+ G_{1;in}^{A\rightarrow B} 
++G_{2;in}^{A\rightarrow B} 
++G_{3;in}^{A\rightarrow B} \;,
+\end{equation}
+%
+where
+%
 \begin{subequations}
-\begin{align} \label{e:ct-murell-etal.vin.oep}
-\Approx{ V_{in}^{A\rightarrow B} }{1}
-       &= \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} 
- +
-S_{ni} u_i^{BA} \;, \\
-%
-\Approx{ V_{in}^{A\rightarrow B} }{2} &=
-\Approx{ V_{in}^{A\rightarrow B} }{1} -
-  \sum_{j   \in B}^{\rm Occ} S_{ij}
+\begin{align}
+ G_{1;in}^{A\rightarrow B} &\equiv \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;, \\
+ G_{2;in}^{A\rightarrow B} &\equiv S_{ni} u_i^{BA} \;, \\
+ G_{3;in}^{A\rightarrow B} &\equiv
+ -\sum_{j   \in B}^{\rm Occ} S_{ij}
   \sum_{y   \in B}^{\rm At} q_y^{B,(nj)} w_{yi}^{BA}
-\;,
+\;.
 \end{align}
@@
 %
-where the auxiliary variables are
+In the above, the auxiliary variables are
 %
@@
    \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
-   \Approx{V_{i'n}^{A\rightarrow B}}{} 
+   V_{i'n}^{A\rightarrow B}
  \rvert^2 }{\varepsilon_i - \varepsilon_n} \;,
@@
  \Bigg(
-   \Approx{V_{in}^{A\rightarrow B}}{\rm G1} \\
+   G_{1,in}^{A\rightarrow B} \\
   +
@@
    \left\{
-   \Approx{V_{i'n}^{A\rightarrow B}}{\rm G2} 
-  +\Approx{V_{i'n}^{A\rightarrow B}}{\rm G3}
+   G_{2,i'n}^{A\rightarrow B}
+  +G_{3,i'n}^{A\rightarrow B}
   \right\}
@@
 %
-where the subscripts G$n$ ($n$ = 1, 2, 3) denote a particular group of 
+where the subscripts $G_n$ ($n$ = 1, 2, 3) denote a particular group of 
 terms from Eqs.~\eqref{e:ct.group-i.final}, \eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.a}
@@
 The total CT energy is given by the sum of the above contribution and the twin contribution
-due to CT from molecule $B$ to $A$.
+due to CT from molecule $B$ to $A$,
+%
+\begin{equation}
+ \Delta E^{\rm CT} = E^{A\rightarrow B} + E^{B\rightarrow A} \;.
+\end{equation}
+%
 
@@
 
+Four selected molecular pairs composed of H$_2$O, CH$_3$OH, NH$_4^+$ and NO$_3^-$
+(Figure 1) were chosen to validate the OEP\hyp{}based model of the CT energy
+at the PT/HF formulation level. Structures of the molecular dimers were first 
+energy\hyp{}optimized at the HF/6-31+G(d,p) level by using
+the Gaussian 16 quantum chemistry program package.\cite{Gaussian16}
+To obtain the PT/HF benchmark CT energy, it was estimated
+from the following approximation:
+%
+\begin{equation}\label{e:ct-pthf-ref}
+ \Delta E^{\rm CT}_{\rm PT/HF} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
+ + \frac{1}{2} {\bf F}^{\rm T} \cdot {\bf D}^{-1} \cdot {\bf F} \;,
+\end{equation}
+%
+where $\Delta E^{\rm Pol}_{\rm DDS/HF}$ is the Hartree\hyp{}Fock polarization energy
+defined
+according to the density decomposition scheme (here referred as to the DDS) 
+proposed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+whereas the remainder term is the 
+approximate pure induction energy due to the polarizability effects.\cite{Li.Netzloff.Gordon.JCP.2006}
+The DDS polarization energy encompasses all the charge delocalization
+effects that are not associated with the Pauli exclusion principle
+in the first order, and is computed by
+%
+\begin{multline}\label{e:de-pol-dds}
+\Delta E^{\rm Pol}_{\rm DDS/HF} \equiv \Delta E^{\rm HF} \\
+ - \left(
+   \Delta E^{\rm Coul}_{\rm DDS/HF}
+ + \Delta E^{\rm Exch}_{\rm DDS/HF}
+ + \Delta E^{\rm Rep}_{\rm DDS/HF} 
+   \right)\;.
+\end{multline}
+%
+In the above, $\Delta E^{\rm Coul}_{\rm DDS/HF}$ is the pure Coulombic interaction energy
+between the unperturbed (non\hyp{}interacting) charge densities of the monomers in the complex,
+and
+$\Delta E^{\rm Exch}_{\rm DDS/HF}$ and $\Delta E^{\rm Rep}_{\rm DDS/HF}$
+are the exchange and Pauli\hyp{}repulsion interaction energies, respectively. 
+At the HF level of theory, their sum is in fact
+equivalent to the first\hyp{}order exchange\hyp{}repulsion energy 
+in the intermolecular perturbation theory with exchange of Hayes and Stone.\cite{Hayes.Stone.MolPhys.1984} 
+It was also shown by 
+Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011}
+that the Coulombic and exchange\hyp{}repulsion term
+in the DDS method reproduce the first\hyp{}order Coulombic and exchange\hyp{}repulsion
+interaction energy in the symmetry adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994}
+(SAPT)
+with accuracy below 0.5 kcal/mole.
+Therefore, Eq.~\eqref{e:ct-pthf-ref} is believed to be sufficient
+in order to compute the PT/HF CT energy, because there is no dispersion
+contribution to the DDS polarization energy at the HF level.
+
+The pure induction energy was estimated by using the distributed
+dipole\hyp{}dipole polarizability approximation by using the computational
+methodology by Li et al.,\cite{Li.Netzloff.Gordon.JCP.2006} 
+in which ${\bf F}$ constitutes the electric fields aligned in a super\hyp{}vector 
+according to each polarizable center, i.e. ${\bf F} = \{ {\bf F}_1, {\bf F}_2, \ldots\}$,
+whereas
+the ${\bf D}$ matrix is defined as
+%
+\begin{equation}
+ {\bf D}_{ab} = 
+\left\{\begin{matrix}
+{\BM\alpha}_a^{-1} &\text{if $a=b$} \\ 
+0                  &\text{if $a\ne b \land a,b \in$ same molecule}\\ 
+-{\bf T}_{ab}      &\text{if $a\ne b \land a,b \notin$ same molecule}
+\end{matrix}\right.
+\end{equation}
+%
+and ${\bf T}_{ab}$ is the dipole\hyp{}dipole interaction tensor
+given by
+%
+\begin{equation}
+ {\bf T}_{ab} \equiv \frac{3}{r_{ab}^5} {\bf r}_{ab} \otimes {\bf r}_{ab}
+  - \frac{1}{r_{ab}^3} {\bf 1} \;,
+\end{equation}
+%
+with ${\bf r}_{ab} \equiv {\bf r}_{a} - {\bf r}_{b}$ being the
+relative distance between the LMO centroids.
+The distributed polarizabilities {\bf\textalpha} ${\boldsymbol{\mathrm{\upalpha}}}_a$ centered at the LMO centroids
+were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
+method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
+Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
+
 All the models that were used to test the theory presented in this work
@@
 
+\subsection{\label{s:gsdzdv}Accuracy of OEP-based model}
 %
@@
 
-Until this moment in this work, two models have been tested with application of the OEP technique.
+The utmost goal of this work is to reduce the computational cost of
+the CT PT/HF energy evaluation in the calculations involving effective fragment
+potentials. The EFP2 CT expression has the computational cost
+roughly proportional to $(10on^2 + 2n^3 + o^2n^2)v$, where the
+$o$ and $n$ denote the number of occupied orbitals and the number of atomic basis functions,
+respectively, whereas $v$ is the relative cost. On the other hand, OEP\hyp{}based expression
+from Eq.~\eqref{} has the cost of approximately $on(Q+N+o)$, where 
+$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
+respectively.
+
 In this section, computational cost is estimated from the formulas.

### 2019-11-15 papers/2_oep_technique/paper-r2.tex
CT paper revision added new sections

@@
 Four selected molecular pairs composed of H$_2$O, CH$_3$OH, NH$_4^+$ and NO$_3^-$
-(Figure 1) were chosen to validate the OEP\hyp{}based model of the CT energy
+(Figure 1) molecules or ions were chosen to validate the OEP\hyp{}based model of the CT energy
 at the PT/HF formulation level. Structures of the molecular dimers were first 
@@
  \Delta E^{\rm CT}_{\rm PT/HF} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
- + \frac{1}{2} {\bf F}^{\rm T} \cdot {\bf D}^{-1} \cdot {\bf F} \;,
+ - \Delta E^{\rm Ind} \;,
 \end{equation}
@@
 proposed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
-whereas the remainder term is the 
-approximate pure induction energy due to the polarizability effects.\cite{Li.Netzloff.Gordon.JCP.2006}
+whereas $\Delta E^{\rm Ind}$ term is the 
+approximate pure induction energy due to the polarizability effects.\cite{Stone.TheTheoryOfIntermolecularForces.1996}
 The DDS polarization energy encompasses all the charge delocalization
@@
 $\Delta E^{\rm Exch}_{\rm DDS/HF}$ and $\Delta E^{\rm Rep}_{\rm DDS/HF}$
-are the exchange and Pauli\hyp{}repulsion interaction energies, respectively. 
-At the HF level of theory, their sum is in fact
+are the exchange and the Pauli\hyp{}repulsion interaction energies, respectively. 
+At the HF level of theory, the sum of the latter two terms is in fact
 equivalent to the first\hyp{}order exchange\hyp{}repulsion energy 
@@
 (SAPT)
-with accuracy below 0.5 kcal/mole.
+with accuracy below 0.3 kcal/mole.
 Therefore, Eq.~\eqref{e:ct-pthf-ref} is believed to be sufficient
@@
 The pure induction energy was estimated by using the distributed
-dipole\hyp{}dipole polarizability approximation by using the computational
-methodology by Li et al.,\cite{Li.Netzloff.Gordon.JCP.2006} 
-in which ${\bf F}$ constitutes the electric fields aligned in a super\hyp{}vector 
-according to each polarizable center, i.e. ${\bf F} = \{ {\bf F}_1, {\bf F}_2, \ldots\}$,
-whereas
-the ${\bf D}$ matrix is defined as
+dipole\hyp{}dipole polarizability approximation according to the computational
+methodology of Li et al.,\cite{Li.Netzloff.Gordon.JCP.2006}
+in which
+%
+\begin{equation}\label{e:ind-ref}
+ \Delta E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\bf D}^{-1} \cdot {\bf F} \;.
+\end{equation}
+%
+In the above equation, ${\bf F} = \{ \ldots, {\bf F}({\bf r}_a), \ldots\}$
+is a super\hyp{}vector of electric fields at the $a$th distributed site
+due to the surrounding unperturbed molecular charge densities in the complex
+(i.e., the charge\hyp{}density of the molecule that contains $a$th center
+is not included in ${\bf F}({\bf r}_a)$).
+%The electric field components are
+%aligned in a super\hyp{}vector 
+%according to each polarizable center, i.e. ${\bf F} = \{ \ldots, {\bf F}({\bf r}_a), \ldots\}$,
+%where $a$ labels the LMO centroids.
+The ${\bf D}$ matrix is defined as
 %
@@
 \left\{\begin{matrix}
-{\BM\alpha}_a^{-1} &\text{if $a=b$} \\ 
+{\BM\upalpha}_a^{-1} &\text{if $a=b$} \\ 
 0                    &\text{if $a\ne b \land a,b \in$ same molecule}\\ 
@@
 relative distance between the LMO centroids.
-The distributed polarizabilities {\bf\textalpha} ${\boldsymbol{\mathrm{\upalpha}}}_a$ centered at the LMO centroids
+The electric fields at the LMO centroids were computed from
+the unperturbed HF charge\hyp{}density distributions
+%
+\begin{equation}
+ {\bf F}({\bf r}_a) = \sum_x^{\rm At} \frac{Z_x}{r_{ax}^3} {\bf r}_{ax} 
+  + 2 \sum_{\alpha\gamma} D_{\alpha\gamma} 
+%    \tBraKet{\gamma}{\frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3} }{\alpha}_{\bf r} \;.
+  \left< \gamma \left| 
+    \frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3}
+               \right| \alpha \right>_{\bf r} \;.
+\end{equation}
+%
+The distributed anisotropic dipole\hyp{}dipole
+polarizabilities ${\BM{\upalpha}}_a$ centered at the LMO centroids
 were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock

### 2019-11-16 papers/2_oep_technique/paper-r2.tex
New fogures added and EFP2 equation initiated

@@
  U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
+\end{equation}
+%
+Xu and Gordon used the expansion of the overlap density 
+in Taylor series and found four different approximate
+theories for the CT energy. The optimal theory, which was shown
+by them to best reproduce the reduced variational scheme (RVS)
+reads
+%
+\begin{equation}
+ E^{A\rightarrow B} \approx 
+ 2\sum_{i\in A}^{\rm Occ}
+  \sum_{n\in B}^{\rm Vir}
+   \frac{ \left| V_{in}^{A\rightarrow B} \right|^2}{F_{ii} - T_{nn}} \;,
+\end{equation}
+%
+where
+%
+\begin{multline}
+ \left| V_{in}^{A\rightarrow B} \right|^2 \approx
+ \frac{U_{in}}{1-\sum_{m\in A}^{\rm All} S_{mn}^2}
+ \Bigg\{
+   U_{in} \\ + \sum_{j\in B}^{\rm Occ} S_{ij} 
+   \left( 
+     T_{nj} - \sum_{m\in A}^{\rm All} S_{nm} T_{mj}
+   \right)
+ \Bigg\} \;,
+\end{multline}
+%
+and 
+%
+\begin{equation}
+ U_{in} \equiv V_{in}^{{\rm EF},B} - \sum_{m\in A}^{\rm All} V_{im}^{{\rm EF},B} S_{mn} \;.
 \end{equation}
@@
 centered at their
-respective LMO charge centroids vanish.
+respective LMO charge centroids vanish.\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
 This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows:
@@
 \subsection{\label{s:gsdzdv}Accuracy of OEP-based model}
-%
-\begin{figure}[t]
-\includegraphics[width=0.5\textwidth]{fig-2.eps}
-\caption{\label{f:fig-1} {\bf Accuracy of the OEP approach to estimate the charge transfer energy
-in the PT/HF formulation.} 
-Asymptotic dependencies of the charge transfer energy is plotted for 
+
+First, the asymptotic dependence of the CT energy
+was investigated for four chosen molecular pairs and the results are
+shown in Figure~\ref{f:fig-1} and Figure~S1 for the methanol-water system. 
+Overall, OEP technique of effective
+elimination of ERI's sketched in Eq.~\eqref{e:ft-reduction} 
+works
+quantitatively well for water dimer and ammonium nitrate (Figures~\ref{f:fig-1}(a) and (c)),
+and methanol-water (Figure~S1)
+reproducing the reference Otto\hyp{}Ladik CT energy,
+whereas in the case of water\hyp{}ammonium cation complex the OEP
+method overestimates the CT energy by around 50\%, even for large
+intermonomer separations. Careful inspection
+(see SI)
+revealed that the reason is the SOA in $G_2$ and $G_3$ and not
+the generalized density fitting in $G_1$ terms, in the case of which
+the aug-cc-pVDZ-jkfit basis set was sufficient as auxiliary space of OEP transformation.
+On the other hand, analysis of the statistical sets of dimeric complexes
+shown in Figure~XXX
+confirmed that in the great majority of cases the 
+OEP technique gives reasonably accurate results with respect
+to the parent model with relative errors less than 10\%,
+unless systems are charged.
+
+In the case of water dimer and methanol-water complex, EFP2 and PT/HF estimations
+coincide with the OEP model. However, in the presence of ions, 
+the deviation of the OEP model from the PT/HF energy 
+become significant, in contrast to the EFP2 model, which gives very good estimates
+of the PT/HF energy in mostly all cases.
+Naturally, the accuracy of the Otto\hyp{}Ladik model limits the absolute accuracy
+of the OEP model, which is manifested in the ammonium nitrate system
+in Figure~\ref{f:fig-1}(c) or the problems with SOA. 
+In order to improve the performance for ionic systems,
+further studies are necessary, that will modify the original formulation
+of Otto and Ladik and SOA treatment. Nevertheless, except from the ionic systems,
+it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable
+(Figure~XX) across a wide range of various interacting complexes including 
+systems with H-boding, dominated by dipole-dipole or dispersion interactions or 
+exhibiting substantial CT.
+
+%
+\begin{figure}[h]
+\includegraphics[width=0.46\textwidth]{fig-2.eps}
+\caption{\label{f:fig-1} {\bf Asymptotic dependence of the charge transfer energy
+in the PT/HF formulation for selected bi\hyp{}molecular complexes.} 
+The charge transfer energy estimates are plotted for 
 (a) water dimer, 
@@
 \end{figure}
+%
+
+%
+\begin{figure*}[h]
+\includegraphics[width=0.8\textwidth]{fig-3.eps}
+\caption{\label{f:fig-1} {\bf Performance of the OEP-based vs EFP2 method for the PT/HF charge transfer interaction energy.}
+(a) accuracy of the ERI elimination technique; 
+(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
+PT/HF estimate. 
+The structures of bi\hyp{}molecular complexes
+were taken from the databases developed by the Truhlar's group:
+the HB6/04 hydrogen bonding database 
+(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the DI6/04 dipoleinteraction database 
+(green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the CT7/04 charge-transfer complex database 
+(blue),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the WI7/05 weak interaction database 
+(yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+and PPS5/05 the $\pi$-$\pi$ stacking database 
+(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+6-311++G** basis set was used for the CT energy calculations.
+} 
+\end{figure*}
 %

### 2019-11-16 oepdev/libsolver/solver_energy_ct.cc

@@ -1300,8 +1300,10 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
             v_ba_v2->set(i, n, v); 
        }
   }
-  v_ab_v2->gemm(false, false, 1.0, oep_1->localizer()->U(), v_ab_v2->clone(), 0.0);
-  v_ba_v2->gemm(false, false, 1.0, oep_2->localizer()->U(), v_ba_v2->clone(), 0.0);
+  psi::SharedMatrix v_ab_v2_copy = v_ab_v2->clone(); v_ab_v2->zero();
+  psi::SharedMatrix v_ba_v2_copy = v_ba_v2->clone(); v_ba_v2->zero();
+  v_ab_v2->gemm(false, false, 1.0, oep_1->localizer()->U(), v_ab_v2_copy, 0.0);
+  v_ba_v2->gemm(false, false, 1.0, oep_2->localizer()->U(), v_ba_v2_copy, 0.0);
 
   // ===> Compute V3 term <=== //
   std::shared_ptr<psi::Matrix> v_ab_v3 = std::make_shared<psi::Matrix>("", nocc_1, nvir_2);
@@ -1329,8 +1331,10 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
             v_ba_v3->set(i, n, v);
        }
   }
-  v_ab_v3->gemm(false, false, 1.0, oep_1->localizer()->U(), v_ab_v3->clone(), 0.0);
-  v_ba_v3->gemm(false, false, 1.0, oep_2->localizer()->U(), v_ba_v3->clone(), 0.0);
+  psi::SharedMatrix v_ab_v3_copy = v_ab_v3->clone(); v_ab_v3->zero();
+  psi::SharedMatrix v_ba_v3_copy = v_ba_v3->clone(); v_ba_v3->zero();
+  v_ab_v3->gemm(false, false, 1.0, oep_1->localizer()->U(), v_ab_v3_copy, 0.0);
+  v_ba_v3->gemm(false, false, 1.0, oep_2->localizer()->U(), v_ba_v3_copy, 0.0);
 
 
   // ---> Add coupling constant contributions <--- //

### 2019-11-16 papers/2_oep_technique/do-r2.sh

@@ -1,6 +1,10 @@
 #!/bin/bash
 base=paper-r2
+si=supplementary-r2
 
-latex  ${base}.tex
-bibtex ${base}.aux
-dvipdf ${base}.dvi
+for i in $base $si
+do
+  latex  ${i}.tex
+  bibtex ${i}.aux
+  dvipdf ${i}.dvi
+done

### 2019-11-20 papers/2_oep_technique/paper-r2.tex
Added multiple figures and new tables

--- 2019-11-15 papers/2_oep_technique/paper-r2.tex
@@
 
-\section{\label{ss.5.2}Introduction}
+\section{\label{s:1.introduction}Introduction}
 
@@
 of interacting molecules change which leads to an additional stabilization 
-of a molecular aggregate. The effect of CT on the interaction energy 
-is known to be often non\hyp{}negligible, especially
-in donor\hyp{}acceptor systems such as H\hyp{}bonded species
-and charged complexes.
-\emph{Ab initio} fragmentation methods, such as
-the second generation of the effective fragment potential method (EFP2),
-are often used to describe the homogeneous extended molecular systems 
-such as solutions
-and recently even biomolecules. EFP2 relies on the intermolecular 
-perturbation theory, and the total interaction energy is approximated as
-a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
-charge densities of the monomers, exchange\hyp{}repulsion energy,
-induction energy due to mutual polarization, intermolecular dispersion energy,
-as well as the CT energy,
-%
-\begin{equation}
+of a molecular aggregate.\cite{Otto.Ladik.IJQC.1980,Stone.TheTheoryOfIntermolecularForces.1996}
+Although CT processes can be relatively easily scrutinized based
+on the total amount of charge transferred between interacting species,\cite{Otto.Ladik.IJQC.1980}
+evaluation of its contribution to the intermolecular interaction
+potential\cite{Stone.TheTheoryOfIntermolecularForces.1996} is far from trivial due to 
+its very complex quantum mechanical (QM) origins even at the Hartree\hyp{}Fock approximation level,\cite{Otto.Ladik.ChemPhys.1975}
+and cannot be realized in terms of any classical nor semi\hyp{}classical approach.
+In fact, CT does not naturally emerge in the symmetry 
+adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT)
+due to certain technical aspects related with the use of finite basis set expansions
+in modern quantum chemistry calculations and minimizing the basis set superimposition error (BSSE). 
+Recently, Stone and Misquitta showed that CT energy
+can be extracted from SAPT calculations by comparing the induction energies
+of fictional systems in which basis functions are centered either on the monomers only, or the entire 
+interacting complex.\cite{Stone.Misquitta.CPL.2009}
+CT energy was formulated by Murrell et al.~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
+in their perturbation theory in the region of small wavefunction overlap up to second order.
+However, their theory is also computationally too expensive to apply for efficient
+calculation of intermolecular forces in molecular dynamics because
+it involces instantaneous calculations of the electron repulsion integrals (ERI's)
+in molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}
+
+The apparent difficulty in theoretically characterising the CT energy
+in terms of the interacting molecular fragments
+is indeed a challenge in developing modern force fields 
+or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
+for modeling structure
+and dynamics in condensed phases.
+This needs to be contrasted with the
+Coulombic electrostatics,
+non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
+dispersion and induction for instance, 
+which to a certain extent
+can be well described by, respectively, 
+the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
+Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
+the van der Waals repulsive and attractive potentials, 
+as well as the polarizability models, i.e., all being relatively simple and computationally
+inexpensive to evaluate mathematical models. 
+Due to this reason, the
+CT effects 
+%while usually of significant and non\hyp{}negligible magnitude as compared 
+%to the total interaction energy, 
+are not included in any molecular mechanics force field developed up to date
+such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
+The only way to
+incorporate the CT effects in the condensed\hyp{}phase simulations,
+apart from performing full QM molecular dynamics simulations,
+is through the second generation of the effective fragment potential 
+(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
+   Nguyen.Pachter.Day.JCP.2014,
+   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
+developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
+EFP2 is an \emph{ab initio} force field
+derived on the grounds of the first\hyp{}principles perturbation theory
+at the HF level and including the intermolecular dispersion effects
+by the response theory based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
+That is to say, the total intermolecular 
+interaction potential is approximated as 
+%
+\begin{equation}\label{e:u-efp2}
  U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
@@
 %
-However, calculating the CT stabilization energy is usually much more expensive 
-than evaluating the contributions due to other effects.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
+where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
+distributions of the monomers, approximated by the distributed multipole approximation,
+$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
+principle, that could be roughly understood as the Heitler\hyp{}London interaction
+energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
+are the induction and dispersion energies obtained from the distributed polarizability
+approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.
+
+Despite the considerable success of the EFP2 theory
+in accurately modeling the extended molecular systems 
+like solutions and recently even biomolecules
+with the level of accuracy
+roughly arguably comparable to the M{\o}ller\hyp{}Plesset perturbation theory,
+evaluation of the CT energy in EFP2 model is still too costly for
+typical uses in the
+molecular dynamics simulations. It has been reported that 
+the implementation of the EFP2 CT energy component in GAMESS
+is on average 3-10 times more demanding than the exchange\hyp{}repulsion
+component, becoming a bottleneck of evaluation of total interaction 
+energy and gradients.\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
+Li.Gordon.Jensen.JCP.2006}
+In effect, the CT energy component is usually ignored in most of the applications. In the 
+recent LIBEFP library, CT term were not implemented in the official release 
+of the code.\cite{Kaliman.Slipchenko.JCC.2015}
+
+%However, calculating the CT stabilization energy is usually much more expensive 
+%than evaluating the contributions due to other effects.
 %However, the computational cost of the CT term is usually at least few times larger
@@
 %to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$.
-In effect,
-this part of the EFP2 energy is usually ignored in most of the applications. In the 
-recent LIBEFP library, CT term were even not implemented due to those reasons.
-
-In this work, we propose an alternative formulation of the CT energy compatible with the
-EFP2 method. We start from the perturbation theory of Murrell et al.
-and
-introduce the effective one\hyp{}electron potential (OEP) operators
-to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
-Later we validate the model against the EFP2 method in Section XXX. Finally,
-we conclude this work and provide few remarks on future development in Section XXX.
-
-\section{\label{s:2}Effective One-Electron Potential Operators}
+
+The effect of CT on the interaction energy 
+is known to be often non\hyp{}negligible, especially
+in donor\hyp{}acceptor systems such as H\hyp{}bonded species
+and charged complexes, although highly depends on the theoretical
+approach being used. 
+The approach adopted by the founders of EFP2 model
+is based on
+%The most common methodology,
+%adopted also by the founders of EFP2 model,
+%is 
+the approximation to SAPT at HF level of theory 
+which is based on the isolated monomeric wave functons
+as references (here abbreviated to CT/HF0).
+In this formulation, the CT energy can be understood
+as a part of the polarization energy that cannot be
+linked to dispersion and pure induction interactions. 
+In this definition of CT, mixing of virtual and occupied MO's
+technically belonging to different unperturbed wavefunction occurs,
+unlike the pure induction and dispersion, in which such mixings are contained
+solely within one monomer.
+%This level of definition of the CT energy will be referred here as to
+%the IPTO-CT energy. 
+%For defining the CT/HF0, 
+%most common reference is the HF approximation
+%most certainly because the HF energy functional is known
+%and its properties are well understood.
+%In the EFP2 model, the total interaction energy is approximated as
+%a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
+%charge densities of the monomers, exchange\hyp{}repulsion energy,
+%induction energy due to mutual polarization, intermolecular dispersion energy,
+%as well as the CT energy.
+%For developing the efficient 
+%\emph{ab initio} fragmentation method 
+%
+% the second generation of the effective fragment potential method (EFP2),
+%the Hartree\hyp{}Fock (HF) model or the density functional theory (DFT) were
+%employed as starting points 
+%Modelling the structure and dynamics of extended molecular aggregates, such as
+%liquids and solutions require
+%EFP2 relies on the intermolecular 
+%perturbation theory, and the total interaction energy is approximated as ...
+%
+%\begin{equation}
+% U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
+%\end{equation}
+%%
+
+One of the main goals of this work, apart from developing
+more efficient model of the CT/HF0 energy, is to
+extend the effective one\hyp{}electron potentials (here referred to as the OEP) 
+technique
+used previously\cite{
+Otto.Ladik.ChemPhys.1975,
+Holas.March.PhysRevA.1991,
+Weber.Thiel.TCA.2000,
+Neese.JCP.2005,
+Cisneros.Andres.Piquemal.Darden.JCP.2005,
+Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
+Li.Gordon.Jensen.JCP.2006,
+Blasiak.Lee.Cho.JCP.2013,
+Blasiak.Maj.Cho.Gora.JCTC.2015}
+to simplify the rigorous and costly quantum chemistry models of extended systems
+with a particular emphasis on solvation phenomena and molecular dynamics.
+The presented OEP technique of removing ERI's from the working
+equations
+follows the notion of the importance of 
+one\hyp{}electron densities in chemistry\cite{Holas.March.PhysRevA.1991},
+thus reducing the complicated summations involving ERI's to much shorter expressions involving
+only one\hyp{}electron integrals.
+Therefore, the OEP computational method is first presented in Section~II.
+Next, in Section III,
+the new technique is used to derive an alternative formulation of the CT/HF0 energy
+compatible with the EFP2 method. Subsequently, after details of computations
+are discussed in Section IV, in Section V the validation of the 
+OEP\hyp{}based CT/HF0 model is presented and its performance in terms 
+of accuracy and computational speed is compared against the EFP2
+model. Finally, few concluding remarks and comments for future work
+are given in Section VI.
+%ERI's, which are oftentimes the main bottleneck
+%in simulating large systems, will be projected out from the working equations
+%to ultimately form the effective fragments in a form of small matrices that are much easier to handle.
+%The proposed methodology aims in removing ERI's from the working equations
+%to ultimately form the effective fragments that are much easier to handle.
+%
+%It is believed that the proposed concept can be helpful in designing future methods tuned 
+%for studying electronically excited states in solutions, better describing molecular
+%dynamics as well as enabling the vibrational analysis of extended systems
+%In this work, we propose an alternative formulation of the CT energy compatible with the
+%EFP2 method. We start from the perturbation theory of Murrell et al.
+%and
+%introduce the effective one\hyp{}electron potential (OEP) operators
+%to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
+%Later we validate the model against the EFP2 method in Section XXX. Finally,
+%we conclude this work and provide few remarks on future development in Section XXX.
+
+\section{\label{s:2.oep}Effective One-Electron Potential Operators}
 
@@
 Formally, the electronic part can be expanded in terms of an effective
-one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
+bond order matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
 %
 \begin{equation} \label{e:d-spectral}
-	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
+	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
 	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
@@
 %\begin{align}
-\begin{equation}
+\begin{equation} \label{e:oep-operator-nuc}
         \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
@@
 %
-\begin{equation}
+\begin{equation} \label{e:oep-matrix-element}
 	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
 	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
-        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;,
+        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
 \end{equation}
@@
 %
-\begin{equation}
+\begin{equation} \label{e:oep-matrix-element-nuc}
  W_{\alpha\beta}^{(x)} = 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:1el-operator}
 \tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:eri}
 	\BraKet{\alpha\beta}{\gamma\delta} \equiv
@@
 
-%\subsection{\label{s.334}Practical OEP-Based Calculations}
-
-\subsection{\label{ss:2.3}Incorporating Electron Repulsion Integrals into Effective Potentials}
+\subsection{\label{ss:2.1.oep-technique}Incorporating Electron Repulsion Integrals into Effective Potentials}
 
@@
 The above design has the advantage that it opens the possibility to define first\hyp{}principles
-effective fragments as long as the $P_{ij}$ and $P_{ijkl}$ 
-from Eq.~\eqref{e:exp-val-series} are computable and can be approximately
+effective fragments as long as 
+the functionals ${\mathcal{F}}_t$ 
+%the $P_{ij}$ and $P_{ijkl}$ 
+%from Eq.~\eqref{e:exp-val-series} 
+are computable and can be approximately
 partitioned in between the interacting fragments.
@@
 
-\subsection{\label{s.333}Distributed Multipole Expansion of OEP's}
+\subsection{\label{ss:2.2.oep-DMTP}Distributed Multipole Expansion of OEP's}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:coulomb-energy:general}
  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
@@
 %
-\begin{multline}
+\begin{multline} \label{e:coulomb-energy:dmtp}
   \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
@@
 There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
-of Stone and Anderton, %cite
-the cumulative atomic multipole moments of Sokalski and Poirier, %cite
-the localised distributed multipole expansion (LMTP) of Etchtebest et al. %cite
-or schemes based on fitting to electrostatic potential.
-
-\subsection{\label{s.3344}Density Fitting of OEP's}
+of Stone\cite{Stone.JCTC.2005}
+the cumulative atomic multipole moments of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983}
+the localised distributed multipole expansion (LMTP) of Etchtebest, Lavery and Pullman\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
+or schemes based on fitting to electrostatic potential such as ChelpG\cite{Breneman.Wiberg.JCC.1990} for instance.
+
+\subsection{\label{ss:2.3.oep-GDF}Density Fitting of OEP's}
 
@@
 
-\subsubsection{Density Fitting in Nearly-Complete Space}
+\subsubsection{\label{sss:2.3.1.GDF-1}Density Fitting in Nearly-Complete Space}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-compl}
  \hat{v}^{A}\Ket{i} = \sum_{\xi}^{\rm RI} V^A_{i\xi} \Ket{\xi} \;.
@@
 %
-\begin{equation}
+\begin{equation} \label{e:error-compl}
  \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm RI} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
@@
 %
-\begin{equation} \label{e:gdf-coml.v}
+\begin{equation} \label{e:gdf-compl.v}
 % {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
@@
 %
-\begin{equation}
+\begin{equation} \label{e:a-coeff-compl}
  a^{(i)}_\eta = \lambda \sum_{x} W_{\eta i}^{(x)} + 
@@
 
-\subsubsection{Density Fitting in Incomplete Space}
+\subsubsection{\label{sss:2.3.1.GDF-2}Density Fitting in Incomplete Space}
 
@@
 set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
-matrix elements of the OEP operator in a incomplete auxiliary space. This can be achieved by minimizing 
+matrix elements of the OEP operator in an incomplete auxiliary space. This can be achieved by minimizing 
 the following objective function
@@
 %
-\begin{equation}
+\begin{equation} \label{e:error-incompl}
  \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm DF} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
@@
 %
-\begin{equation} \label{e:v-noncompl}
+\begin{equation} \label{e:v-incompl}
 % {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-compl-b-coeff}
  b^{(i)}_\eta = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} R_{\varepsilon\eta} \;,
@@
 which can be easily evaluated by noting that
-$H_{i\varepsilon}$ are given by Eq.~\eqref{e:gdf-coml.v}.
+$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}.
 Thus, the matrix elements of the OEP operator can be found
-by using Eq.~\eqref{e:v-noncompl}.
+by using Eq.~\eqref{e:v-incompl}.
 %Therefore, in order to use this generalized density fitting scheme
@@
 
-\section{\label{s:34r}Charge Transfer Interaction Energy for Fragment Potentials}
-
-\subsection{\label{s:34rg} Murrell et al.'s Model: Starting Point}
-
-After the OEP technique has been formulated, it can be applied 
-to the calculation of interaction\hyp{}induced properties following the
-fragment\hyp{}based philosophy. 
-According to the second\hyp{}order intermolecular perturbation theory of Murrell et al.,
-the energy associated with the charge transfer from molecule $A$ to $B$ is given by
+\section{\label{s:3.ct}Charge Transfer Interaction Energy for Fragment Potentials}
+
+In the CT/HF0 treatments of bi\hyp{}molecular complexes, the CT energy
+can be expressed as a sum of the energy stabilization due to 
+CT from molecule $A$ to $B$ and vice versa, i.e.,
+%
+\begin{equation} \label{e:ct-hf0}
+ U^{\rm CT} =  E^{A\rightarrow B} + E^{B\rightarrow A}\;.
+\end{equation}
+%
+
+\subsection{\label{ss:3.1.EFP2} EFP2 Model}
+
+Li, Gordon and Jensen used the expansion of the overlap density 
+in Taylor series and found four different approximate
+theories for the CT energy.\cite{Li.Gordon.Jensen.JCP.2006} 
+The optimal theory, which was shown
+to well reproduce the CT/HF0 energies obtained by using 
+the reduced variational space (RVS)
+analysis of Stevens and Fink\cite{Stevens.Fink.CPL.1987},
+reads
+%
+\begin{equation} \label{e:u-efp2-ct}
+ E^{A\rightarrow B} \approx 
+ 2\sum_{i\in A}^{\rm Occ}
+  \sum_{n\in B}^{\rm Vir}
+   \frac{ \left| V_{in}^{A\rightarrow B} \right|^2}{\varepsilon_{i} - T_{nn}} \;,
+\end{equation}
+%
+where
+%
+\begin{multline} \label{e:u-efp2-ct-v2}
+ \left| V_{in}^{A\rightarrow B} \right|^2 \approx
+ \frac{U_{in}}{1-\sum_{m\in A}^{\rm All} S_{mn}^2}
+ \Bigg\{
+   U_{in} \\ + \sum_{j\in B}^{\rm Occ} S_{ij} 
+   \left( 
+     T_{nj} - \sum_{m\in A}^{\rm All} S_{nm} T_{mj}
+   \right)
+ \Bigg\} \;,
+\end{multline}
+%
+and 
+%
+\begin{equation} \label{e:uin-efp2}
+ U_{in} \equiv V_{in}^{{\rm EF},B} - \sum_{m\in A}^{\rm All} V_{im}^{{\rm EF},B} S_{mn} \;,
+\end{equation}
+%
+in which the summations extend over occupied (denoted by `Occ'), virtual (denoted by `Vir') 
+or both (denoted by `All') MO's. The effective potential 
+matrices are defined by
+%
+\begin{equation} \label{e:v-efp2-eff}
+ V_{in}^{{\rm EF},B} \equiv 
+ \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} \;,
+\end{equation}
+%
+and are evaluated by expanding the $\hat{v}^B_{\rm tot}$ operator
+in distributed multipoles. The resulting calculation requires computation
+of potential, electric field and electric field gradient one\hyp{}electron integrals
+and their contractions with the DMTP distributions.
+The apparent success of the EFP2 scheme is rooted in the dramatic simplifications of the
+\emph{ab initio} expressions for interaction energy, in which
+the relatively very costly ERI's
+have been effectively removed from the working models while maintaining the required accuracy. 
+That is, the canonical orbital energies $\varepsilon_i$ are
+constant parameters, whereas the overlap $S_{nm}$, kinetic energy $T_{mn}$ and effective
+one\hyp{}electron electrostatic potential $V_{in}^{{\rm EF}}$ matrix elements
+are all certain types of the one\hyp{}electron integrals, orders of magnitude cheaper to
+evaluate than ERI's. Unfortunately, due to extensive summations over virtual orbitals
+evaluating Eq.~\eqref{e:u-efp2-ct} is still considerable in cost because typically large basis sets
+need to be used for generating the EFP2 parameters. Moreover, calculation of $V_{in}^{{\rm EF}}$
+is much more expensive as compared to other types of one\hyp{}electron integrals.
+
+In the following subsections, the alternative model of the CT/HF0 energy is proposed
+by introducing OEP's. Although application of the OEP method to the CT/HF0 formulation in EFP2 method
+is probably possible, it would be relatively difficult to discuss the resulting OEP\hyp{}based
+EFP2 models because there are four distinct versions of this theory
+with a set of different approximations, selected based on performance assessment rather than 
+a rigorous derivation manner.
+Instead, perturbation theory of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
+with the
+explicit formulation for closed shell systems by Otto and Ladik\cite{Otto.Ladik.ChemPhys.1975},
+which is somewhat more rigorous than the EFP2 CT model,
+is chosen as a starting point in this work. 
+It is believed that this choice will allow a clear demonstration 
+of the OEP technique in fragment\hyp{}based modelling.
+
+\subsection{\label{ss:3.2.OL} Murrell et al.'s Model: Starting Point}
+
+%According to the second\hyp{}order intermolecular perturbation theory 
+%with wavefunction overlap of Murrell et al.,
+%the energy associated with the charge transfer from molecule $A$ to $B$ is given by
+The CT/HF0 energy can be expressed
+by\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975}
 %
@@
 where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
-of the occupied (denoted by `Occ') and virtual (denoted by `Vir') 
+of the occupied and virtual
 HF molecular orbitals, respectively,
@@
 
-\subsection{\label{s:34rag} Murrell et al.'s Model: Application of OEP Technique}
+\subsection{\label{ss:3.3.OEP} Murrell et al.'s Model: Application of OEP Technique}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-oep-group-1-prefinal}
  V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf R}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
@@
 leads to a much simpler formula
-that requires only the one- and two\hyp{}electron integrals. 
+that requires only the one\hyp{} and two\hyp{}electron integrals. 
 Therefore, by combining
-equations %which equations?
+equations~\eqref{e:v-oep-group-1-prefinal}, \eqref{e:v-compl-b-coeff} and \eqref{e:gdf-compl.v}
 one arrives to
 %
-\begin{equation}
+\begin{equation} \label{e:v-oep-group-1-final}
  V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:v-oep-group-1-final:H-matrix}
  H_{n\varepsilon}^B = \sum_{\zeta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\zeta}
@@
 %%
-\begin{align}
+\begin{align} \label{e:v-oep-group-1-final:a-vector}
  a^{(n)}_\zeta &= \tBraKet{\zeta}{\hat{v}^B_{\rm tot}}{n}
@@
 can be combined into a single summation term, i.e.,
-\begin{multline}
+\begin{multline} \label{e:ct.group-iii}
        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
@@
 %
-\begin{equation}
+\begin{equation} \label{e:oed-group-iii}
  D^{B,(nj)}_{\beta\delta} = C_{\beta n} C_{\delta j} \;.
@@
 In this work, the effective charges were computed via the Mulliken method as
-discussed in Section~XXX. 
+discussed in Appendix~\ref{e:a1-v.oep-camm} with $\lambda=0$. 
 Note now that both effective densities in 
@@
 %
-\begin{subequations}
+\begin{subequations} \label{e:ct-murrell-etal.vin.oep-gterms}
 \begin{align}
@@
 The total CT energy is given by the sum of the above contribution and the twin contribution
-due to CT from molecule $B$ to $A$,
-%
-\begin{equation}
- \Delta E^{\rm CT} = E^{A\rightarrow B} + E^{B\rightarrow A} \;.
-\end{equation}
-%
+due to CT from molecule $B$ to $A$ according to Eq.~\eqref{e:ct-hf0}.
 
@@
 
-\section{\label{s:4}Calculation Details}
+\section{\label{s:4.calculations}Calculation Details}
 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ind-ref.d-matrix}
  {\bf D}_{ab} = 
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ind-ref.t-matrix}
  {\bf T}_{ab} \equiv \frac{3}{r_{ab}^5} {\bf r}_{ab} \otimes {\bf r}_{ab}
@@
 %
-\begin{equation}
+\begin{equation} \label{e:ind-ref.efield}
  {\bf F}({\bf r}_a) = \sum_x^{\rm At} \frac{Z_x}{r_{ax}^3} {\bf r}_{ax} 
@@
 
-\section{\label{s:fwefsdv}Results and Discussion}
-
-\subsection{\label{s:gsdzdv}Accuracy of OEP-based model}
-%
-\begin{figure}[t]
-\includegraphics[width=0.5\textwidth]{fig-2.eps}
-\caption{\label{f:fig-1} {\bf Accuracy of the OEP approach to estimate the charge transfer energy
-in the PT/HF formulation.} 
-Asymptotic dependencies of the charge transfer energy is plotted for 
+\section{\label{s:5.results}Results and Discussion}
+
+\subsection{\label{ss:5.1.accuracy}Accuracy of OEP-based model}
+
+First, the asymptotic dependence of the CT energy
+was investigated for four chosen molecular pairs and the results are
+shown in Figure~\ref{f:fig-2} and Figure~S1 for the methanol-water system. 
+Overall, OEP technique of effective
+elimination of ERI's sketched in Eq.~\eqref{e:ft-reduction} 
+works
+quantitatively well for water dimer and ammonium nitrate (Figures~\ref{f:fig-2}(a) and (c)),
+and methanol-water (Figure~S1)
+reproducing the reference Otto\hyp{}Ladik CT energy,
+whereas in the case of water\hyp{}ammonium cation complex the OEP
+method overestimates the CT energy by around 50\%, even for large
+intermonomer separations. Careful inspection
+(see SI)
+revealed that the reason is the SOA in $G_2$ and $G_3$ and not
+the generalized density fitting in $G_1$ terms, in the case of which
+the aug-cc-pVDZ-jkfit basis set was sufficient as auxiliary space of OEP transformation.
+On the other hand, analysis of the statistical sets of dimeric complexes
+shown in Figure~XXX
+confirmed that in the great majority of cases the 
+OEP technique gives reasonably accurate results with respect
+to the parent model with relative errors less than 10\%,
+unless systems are charged.
+%
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[h]
+\caption[Accuracy of approximate CT/HF0 energy methods across wide range of bi\hyp{}molecular complexes]
+{{\bf Accuracy of approximate CT/HF0 energy methods across wide range of bi\hyp{}molecular complexes\footnotemark[1]}
+}
+\label{t:2}
+\begin{ruledtabular}
+\begin{tabular}{lclcccccccc}
+         &&       && \multicolumn{7}{c}{RMSE (kcal/mol) } \\
+\cline{4-11}
+\multicolumn{3}{c}{Database}
+   && OL\footnotemark[2]  
+                  && \multicolumn{3}{c}{OEP\footnotemark[3]}  
+                  && EFP2\footnotemark[4] \\
+\cline{1-3}
+\cline{4-5}
+\cline{6-9}
+\cline{10-11}
+NCB31\footnotemark[5] 
+      &&  HB6/04  && 1.38 && 1.36 && 0.37 && 2.45 \\ % 
+      &&  DI6/04  && 0.35 && 0.66 && 0.22 && 0.44 \\ % 
+      &&  CT7/04  && 0.98 && 2.82 && 0.71 && 4.06 \\ % 
+      &&  WI7/04  && 0.01 && 0.09 && 0.05 && 0.02 \\ % 
+      &&  PPS5/05 && 0.07 && 1.31 && 0.79 && 0.60 \\ % 
+      &&  Total   && 0.83 && 1.69 && 0.53 && 2.39 \\ % 
+\cline{1-3}
+\cline{4-5}
+\cline{6-9}
+\cline{10-11}
+
+BBI\footnotemark[6]   
+      &&  Subset\footnotemark[7]
+                  && 0.26 && 0.56 && 0.13* && 0.12 \\ %
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{Validated against Eq.~\eqref{e:ct-pthf-ref}.}
+\footnotetext[2]{Otto and Ladik's expression from Eq.~(4) in Ref.~\cite{Otto.Ladik.ChemPhys.1975}}
+\footnotetext[3]{This work, Eq.~\eqref{e:ect-oep.final-scaled} 
+with $C=1.00$ (left column) and $C=1.56$ (right columm).
+GDF(1)/aug-cc-pVDZ-jkfit scheme was used for the extended density fitting of group (i) OEP's.}
+\footnotetext[4]{Ref.~\cite{Li.Gordon.Jensen.JCP.2006}, Eq.~\eqref{e:u-efp2-ct}.}
+\footnotetext[5]{Ref.~\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Schultz.Truhlar.JCTC.2006}}
+\footnotetext[6]{Ref.~\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}}
+\footnotetext[7]{100 backbone\hyp{}sidechains complexes were randomly selected.}
+%
+\end{table}
+}
+%
+
+In the case of water dimer and methanol-water complex, EFP2 and PT/HF estimations
+coincide with the OEP model. However, in the presence of ions, 
+the deviation of the OEP model from the PT/HF energy 
+become significant, in contrast to the EFP2 model, which gives very good estimates
+of the PT/HF energy in mostly all cases.
+Naturally, the accuracy of the Otto\hyp{}Ladik model limits the absolute accuracy
+of the OEP model, which is manifested in the ammonium nitrate system
+in Figure~\ref{f:fig-2}(c) or the problems with SOA. 
+In order to improve the performance for ionic systems,
+further studies are necessary, that will modify the original formulation
+of Otto and Ladik and SOA treatment. Nevertheless, except from the ionic systems,
+it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable
+(Figure~XX) across a wide range of various interacting complexes including 
+systems with H-boding, dominated by dipole-dipole or dispersion interactions or 
+exhibiting substantial CT.
+%
+\begin{figure}[h]
+\includegraphics[width=0.46\textwidth]{fig-2.pdf}
+\caption{\label{f:fig-2} {\bf Asymptotic dependence of the charge transfer energy
+in the CT/HF0 formulation for selected bi\hyp{}molecular complexes.} 
+The charge transfer energy estimates are plotted for 
 (a) water dimer, 
@@
 where one molecule has been translated by $R$ from the starting geometry
-along the vector specified in Fig. 1.
+optimized at HF/6-31+G(d,p) level,
+along the vector specified in the insets.
 The counterpoise\hyp{}corrected total interaction energy
 is also shown for comparison in light grey color in this figure.
-All data were obtained at HF/6-31+G** level of theory.
+Interaction energies were obtained at the HF/6-311++G(2df,2pd) level of theory
+with the single GDF scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
 } 
@@
 
-\subsection{\label{s:413s6}Reduction of computational costs}
+%
+\begin{figure*}[h]
+\includegraphics[width=0.9\textwidth]{fig-3.pdf}
+\caption{\label{f:fig-3} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
+in small bi\hyp{}molecular neutral complexes.}
+(a) accuracy of the ERI elimination technique; 
+(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
+CT/HF0 estimate. 
+The structures of bi\hyp{}molecular complexes
+were taken from the NCB31 database developed by the Truhlar's group:
+the HB6/04 hydrogen bonding database 
+(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the DI6/04 dipoleinteraction database 
+(green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the CT7/04 charge-transfer complex database 
+(blue),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the WI7/05 weak interaction database 
+(yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+and PPS5/05 the $\pi$-$\pi$ stacking database 
+(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+6-311++G** basis set was used for the CT energy calculations
+and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
+for the OEP calculations.
+} 
+\end{figure*}
+%
+
+%
+\begin{figure*}[h]
+\includegraphics[width=0.9\textwidth]{fig-4.pdf}
+\caption{\label{f:fig-4} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
+in aminoacid pairs.}
+(a) accuracy of the ERI elimination technique; 
+(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
+CT/HF0 estimate. 
+The structures of 100 aminoacid pairs
+were randomly selected from the BBI Bioo database
+for backbone\hyp{}sidechain interactions.\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}
+6-31+G(d,p) basis set was used for the CT energy calculations
+and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
+for the OEP calculations.
+} 
+\end{figure*}
+%
+
+\subsection{\label{ss:5.2.cost}Reduction of computational costs}
 
@@
 respectively, whereas $v$ is the relative cost. On the other hand, OEP\hyp{}based expression
-from Eq.~\eqref{} has the cost of approximately $on(Q+N+o)$, where 
+from Eq.~\eqref{e:ect-oep.final} has the cost of approximately $on(Q+N+o)$, where 
 $Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
 respectively. 
-
-In this section, computational cost is estimated from the formulas.
-In
-Table~\ref{t:oep-costs} 
 %
@@
 \renewcommand{\arraystretch}{1.4}
-\begin{table}[b]
-\caption[Computational cost of the benchmark and OEP-based methods: calculation of coupling constant]
-{{\bf Computational cost of the benchmark and OEP-based methods\footnotemark[1]}
+\begin{table}[h]
+\caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
+{{\bf}Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.\footnotemark[1]
 }
-\label{t:oep-costs}
+\label{t:3}
 \begin{ruledtabular}
-\begin{tabular}{lcccccc}
-Method             && Murrell et al. &&    EFP2        &&      OEP     \\ 
-	\cline{1-7}
-Group (i)         &&  &&        &&   $obQ$     \\
-Group (ii)        &&  &&        &&   $ob$      \\
-Group (iii)       &&  &&        &&   $ob(n+o)$     \\
-Total Cost        &&  && $10ob^2 + 2b^3 + o^2b^2$       &&   $ob(Q+n+o)$     \\
+\begin{tabular}{lldcd}
+ AO's &  & \multicolumn{3}{c}{Exponents (a.u.)}\\
+      &  & \text{Methanol}         && \text{Water}           \\
+\cline{1-2}
+\cline{3-5}
+ H    &1s&   7.7369294534   &&   29.5837988322 \\
+ C    &1s& 880.3654511045   &&                 \\
+      &2s& 123.8276940826   &&                 \\
+      &2p&  11.6748155072   &&                 \\
+ O    &1s& 902.9798919021   && 1030.5721050297 \\
+      &2s& 203.7584408153   &&  142.1022267153 \\
+      &2p&  20.2442429692   &&   40.9175702474 \\
 \end{tabular}
@@
 %
-\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
-$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
+\footnotetext[1]{Fitting performed according to Appendix.~\eqref{a:auxiliary-basis} assuming 6-311++G(d,p) 
+and aug-cc-pVDZ-jkfit primary and test basis sets, respectively.}
 %
@@
 %
+
+In this section, computational cost is estimated from the formulas.
+%In
+%Table~\ref{t:oep-costs} 
+%
+%{
+%\renewcommand{\arraystretch}{1.4}
+%\begin{table}[b]
+%\caption[Computational cost of the benchmark and OEP-based methods: calculation of coupling constant]
+%{{\bf Computational cost of the benchmark and OEP-based methods\footnotemark[1]}
+%}
+%\label{t:oep-costs}
+%\begin{ruledtabular}
+%\begin{tabular}{lcccccc}
+%Method             && Murrell et al. &&    EFP2        &&      OEP     \\ 
+%	\cline{1-7}
+%Group (i)         &&  &&        &&   $obQ$     \\
+%Group (ii)        &&  &&        &&   $ob$      \\
+%Group (iii)       &&  &&        &&   $ob(n+o)$     \\
+%Total Cost        &&  && $10ob^2 + 2b^3 + o^2b^2$       &&   $ob(Q+n+o)$     \\
+%\end{tabular}
+%\end{ruledtabular}
+%%
+%\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
+%$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
+%%
+%\end{table}
+%}
+%
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[h]
+\caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
+{{\bf}CPU timings in miliseconds of CT/HF0 single point energy calculations
+for water-methanol complex at 6-311++G(d,p) primary basis set\footnotemark[1]
+}
+\label{t:4.oep-costs}
+\begin{ruledtabular}
+\begin{tabular}{llccc}
+                      && Water-Methanol && Water dimer \\
+\cline{3-5}
+OL                    && 774    &&\\
+EFP2                  &&  42.3  &&\\
+OEP/aug-cc-pVQZ-jkfit &&   4.7  &&\\
+OEP/aug-cc-pVTZ-jkfit &&   4.0  &&\\
+OEP/aug-cc-pVDZ-jkfit &&   3.8  &&\\
+OEP/6-311++G(d,p)     &&   3.0  &&\\
+OEP/mini              &&   2.5  &&\\
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{CPU: XXX. See the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
+%
+\end{table}
+}
+%
+
+%
 It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
 and many orders of magnitude lesser than the original theory of Murrell et al.
-
-\section{\label{s:6}Summary and a few concluding remarks}
-
-Bla.
+%
+\begin{figure}[h]
+\includegraphics[width=0.46\textwidth]{fig-5.pdf}
+\caption{\label{f:fig-5} {\bf Asymptotic dependence of the charge transfer energy
+in the CT/HF0 formulation for methanol-warer complex with minimal OEP auxiliary basis.} 
+Water molecule has been translated by $R$ from the starting geometry
+optimized at HF/6-31+G(d,p) level,
+along the vector specified in the insets.
+The counterpoise\hyp{}corrected total interaction energy
+is also shown for comparison in light grey color in this figure.
+Interaction energies were obtained at the HF/6-311++G(d,p) level of theory.
+In this figure, `OEP' denotes single GDF scheme with aug-cc-pVQZ-jkfit auxiliary basis set
+whereas `OEP/mini` denotes double GDF scheme with the minimal auxiliary basis
+developed here for water and methanol.
+} 
+\end{figure}
+%
+
+\section{\label{s:6.conclusions}Summary and a few concluding remarks}
+
+In this work, the one\hyp{}electron effective potential operator technique
+for eliminating electron repulsion integrals in the fragment\hyp{}based
+theories of intermolecular interactions
+was derived and applied to calculating the charge transfer energy
+in condensed phases simulations. It was shown that in general case two types of OEP's
+can be defined and worked out either through the density fitting or the distributed
+multipole expansion. For the first group of OEP's, the density fitting was generalized
+and two computational treatments of OEP's were developed: the single GDF scheme
+for RI\hyp{}type auxiliary basis sets, as well as the double GDF scheme, for
+essentially all other types of basis sets. The presented validation of the OEP technique
+against the Otto\hyp{}Ladik CT/HF0 model as parent theory showed that in most cases
+elimination of ERI's is quantitatively accurate, with few exceptions
+where only qualitative accuracy was found. 
+Finally, it was found that, while the proposed OEP\hyp{}based Otto\hyp{}Ladik
+model is of comparable accuracy as the CT/HF0 formulation within the
+state\hyp{}of\hyp{}the\hyp{}art EFP2 model, it significantly outperforms the latter
+in terms of computational efficiency reaching speedups up to 30--40 times.
+%Since EFP2 method is based on the PT/HF0 foundations
+%with fragments defined in gas phase state,
+%interaction energies without the CT contribution
+%in this model
+%are
+%conceptually incomplete.
+Therefore, it is believed that the OEP\hyp{}based
+model can be incorporated within the EFP2 method, enabling
+the full (CT\hyp{}including) EFP2 energy calculations and 
+molecular dynamics simulations in large systems. However, further 
+study is necessary and the following issues need to be carefully addressed:
 
@@
 This project is carried out under POLONEZ programme which has received funding from the European Union's
-Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
+Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
 No.~665778. This project is funded by National Science Centre, Poland 
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
+We thank Professor Marcos Mandado from University of Vigo in Spain
+for providing us with his benchmark calculations to validate our implementation
+of the DDS method.
 \end{acknowledgments}
@@
 is the `test' basis set used to probe the accuracy of the density fitting.
-The forms of the Pauli and charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
-can be directly derived from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
+The forms of the %Pauli and 
+charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
+can be directly derived 
+from Eq.~\eqref{e:v-oep.ct}
+%from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
 by expanding the MO's in terms of atomic basis functions. The working formulae
@@
 
+\section{Coulomb-like OEP via Cumulative Atomic Multipole Moments\label{a:oep-camm}}
+
+Overlap\hyp{}like OEP's are always associated with the effective one\hyp{}particle density
+(or bond\hyp{}order) 
+matrices and the $\lambda$ parameter from Eq.~\eqref{e:oep-operator} being either 1 or 0. 
+Therefore, one can use this fact to compute the effective distributed multipole
+moments. In this work, the cumulative atomic multipole moments (CAMM)
+of Sokalski and Poirier with Mulliken partitioning of the AO space
+were chosen because of their simplicity.
+One of examples
+of such OEP\hyp{}based models reported already in the literature is
+the transition cumulative atomic multipole moments (TrCAMM) model
+for estimation of the excitation energy transfer (EET) couplings in the F{\"o}rster
+limit\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}, 
+or the vibrational solvatochromic distributed multipole moments
+(SolCAMM) for the determination of the vibrational frequency shifts
+of spatially localized IR probes.\cite{Blasiak.Lee.Cho.JCP.2013}
+In general, having the OED and $\lambda$, the effective cumulative atomic charges
+and dipole moments are given by
+%
+\begin{subequations} \label{e:a1-v.oep-camm}
+ \begin{align}
+  q_x^{\rm eff} &= \lambda Z_x - 
+       \sum_{\alpha\in x}\sum_{\beta} 
+        P_{\alpha\beta}^{\rm eff} \BraKet{\alpha}{\beta} \;, \\
+  {\BM\upmu}_x^{\rm eff} &= 
+       \sum_{\alpha\in x}\sum_{\beta} 
+        P_{\alpha\beta}^{\rm eff}
+   \left\{ 
+    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\bf r}{\beta} 
+   \right\}
+   \;,
+ \end{align}
+\end{subequations}
+%
+where ${\bf R}_x$ is the position vector of the $x$-th nucleus.
+Note that $P_{\alpha\beta}^{\rm eff}$ can refer to the particular electron
+spin, or their sum as a bond order matrix as well.
+Equations for the effective distributed quadrupoles, octupoles and hexadecapoles
+can be easily obtained by analogy from Ref.~\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}
+
 %

### 2019-11-20 papers/2_oep_technique/do-r2.sh

@@ -2,9 +2,9 @@
 base=paper-r2
 si=supplementary-r2
 
-for i in $base $si
+for i in $base #$si
 do
-  latex  ${i}.tex
+  pdflatex  ${i}.tex
   bibtex ${i}.aux
-  dvipdf ${i}.dvi
+  #dvipdf ${i}.dvi
 done

### 2019-11-22 papers/2_oep_technique/paper-r2.tex
Paper revision before sending out for coreading

@@
 
+\author{Joanna D. Bednarska}
 \author{Marta Cho{\l}uj} 
-\author{Joanna D. Bednarska}
 \author{Wojciech Bartkowiak}
@@
 \begin{abstract}
-The concept of the effective one-electron potentials has been useful for many decades
+The concept of the effective one\hyp{}electron potentials (OEP) has been useful for many decades
 in efficient description of electronic structure of chemical systems, especially extended
 molecular aggregates such as interacting molecules in condensed phases.
+Here, the general method for effective elimination of the electron repulsion integrals (ERI)
+based on OEP's is presented, that is tuned towards the fragment\hyp{}based calculation methodologies
+such as the second generation of the effective fragment potentials (EFP2) method. 
+Two general types of the OEP operator matrix elements
+are distinguished and treated either via the distributed multipole expansion
+or the extended density fitting schemes developed in this work. 
+The OEP technique was then applied
+to address the problem of using usually incomplete EFP2 settings in most of
+applications in interaction energy and 
+molecular dynamics simulations 
+due to prohibitively high computational cost of evaluating the
+charge transfer (CT) effects as compared to other effects.
+The alternative OEP\hyp{}based CT energy model is proposed
+in the context of the
+intermolecular perturbation theory with Hartree\hyp{}Fock uninteracting gas\hyp{}phase 
+reference wavefunctions, compatible with the EFP2 formulation.
+It is found that the computational cost can be reduced
+up to 20 times as compared to the CT energy method within the EFP2 scheme
+without compromising the accuracy for a wide range of weakly interacting 
+neutral molecular complexes. 
+Therefore, it is believed that
+the proposed model can be used in calculations within 
+the EFP2 framework, making the CT energy term no longer the bottleneck
+in EFP2\hyp{}based simulations of complex systems.
 \end{abstract}
@@
 Although CT processes can be relatively easily scrutinized based
-on the total amount of charge transferred between interacting species,\cite{Otto.Ladik.IJQC.1980}
+on the total amount of electronic charge transferred between interacting species,\cite{Otto.Ladik.IJQC.1980}
 evaluation of its contribution to the intermolecular interaction
 potential\cite{Stone.TheTheoryOfIntermolecularForces.1996} is far from trivial due to 
-its very complex quantum mechanical (QM) origins even at the Hartree\hyp{}Fock approximation level,\cite{Otto.Ladik.ChemPhys.1975}
+its notably complex quantum mechanical (QM) origins 
+even at the Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
+approximation level,\cite{Otto.Ladik.ChemPhys.1975}
 and cannot be realized in terms of any classical nor semi\hyp{}classical approach.
@@
 adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT)
-due to certain technical aspects related with the use of finite basis set expansions
-in modern quantum chemistry calculations and minimizing the basis set superimposition error (BSSE). 
-Recently, Stone and Misquitta showed that CT energy
+due to technical aspects related with the use of finite basis set expansions
+in modern quantum chemistry calculations and avoiding the basis set superimposition error (BSSE). 
+Stone and Misquitta showed that CT energy
 can be extracted from SAPT calculations by comparing the induction energies
@@
 in their perturbation theory in the region of small wavefunction overlap up to second order.
-However, their theory is also computationally too expensive to apply for efficient
+However, all these theories are computationally too expensive to apply for efficient
 calculation of intermolecular forces in molecular dynamics because
-it involces instantaneous calculations of the electron repulsion integrals (ERI's)
-in molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}
+they involces instantaneous calculations of the electron repulsion integrals (ERI's)
+and their four\hyp{}index transformation to molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}
 
@@
 and dynamics in condensed phases.
-This needs to be contrasted with the
-Coulombic electrostatics,
+This needs to be contrasted with the Coulombic electrostatics,
 non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
@@
 inexpensive to evaluate mathematical models. 
-Due to this reason, the
-CT effects 
+Due to this reason, the CT effects 
 %while usually of significant and non\hyp{}negligible magnitude as compared 
 %to the total interaction energy, 
-are not included in any molecular mechanics force field developed up to date
-such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
+are not explicitly included in any molecular mechanics force field developed up to date.
+%such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
 The only way to
 incorporate the CT effects in the condensed\hyp{}phase simulations,
-apart from performing full QM molecular dynamics simulations,
+apart from performing full QM electronic structure simulations,
 is through the second generation of the effective fragment potential 
@@
 at the HF level and including the intermolecular dispersion effects
-by the response theory based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
+by the response theory. 
+%based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
 That is to say, the total intermolecular 
-interaction potential is approximated as 
+interaction potential, which can also be perceived as an approximation
+to SAPT at HF level with added intermolecular dispersion correction, is approximated as 
 %
@@
 Li.Gordon.Jensen.JCP.2006}
-In effect, the CT energy component is usually ignored in most of the applications. In the 
-recent LIBEFP library, CT term were not implemented in the official release 
-of the code.\cite{Kaliman.Slipchenko.JCC.2015}
+In effect, the CT energy component is usually ignored in most of the applications. 
+As a noteworthy example, 
+CT term were not implemented in the official release 
+of the recent LIBEFP library for linking quantum chemistry packages 
+with the EFP2 functionalities.\cite{Kaliman.Slipchenko.JCC.2015}
 
@@
 and charged complexes, although highly depends on the theoretical
-approach being used. 
+approach being used, i.e., the choice of the reference
+wavefunctions.
 The approach adopted by the founders of EFP2 model
-is based on
+is based on gas\hyp{}phase reference (unperturbed) HF wavefunctions,
 %The most common methodology,
@@
 %is 
-the approximation to SAPT at HF level of theory 
-which is based on the isolated monomeric wave functons
-as references (here abbreviated to CT/HF0).
+%which is based on the isolated monomeric wave functons
+%as references 
+here broadly referred to as the CT/HF0 level of theory.
 In this formulation, the CT energy can be understood
@@
 linked to dispersion and pure induction interactions. 
-In this definition of CT, mixing of virtual and occupied MO's
-technically belonging to different unperturbed wavefunction occurs,
+In technical words, mixing of virtual and occupied MO's,
+that are assigned according to certain criteria 
+to different unperturbed wavefunctions, occurs
 unlike the pure induction and dispersion, in which such mixings are contained
@@
 One of the main goals of this work, apart from developing
-more efficient model of the CT/HF0 energy, is to
+more efficient model of the CT/HF0 energy that
+is compatible with the EFP2 energetics, is to
 extend the effective one\hyp{}electron potentials (here referred to as the OEP) 
 technique
-used previously\cite{
+widely
+used previously\cite{Roothaan.RevModPhys.1951,
+Hohenberg.Kohn.PhysRev.1964,
+Kohn.Sham.PhysRev.1965,
 Otto.Ladik.ChemPhys.1975,
@@
 follows the notion of the importance of 
-one\hyp{}electron densities in chemistry\cite{Holas.March.PhysRevA.1991},
+one\hyp{}electron densities in chemistry\cite{Kohn.Sham.PhysRev.1965,Holas.March.PhysRevA.1991},
 thus reducing the complicated summations involving ERI's to much shorter expressions involving
-only one\hyp{}electron integrals.
-Therefore, the OEP computational method is first presented in Section~II.
-Next, in Section III,
+only one\hyp{}electron integrals (OEI's).
+Therefore, the OEP computational method is first presented in Section~\ref{s:2.oep}.
+Next, in Section~\ref{s:3.ct},
 the new technique is used to derive an alternative formulation of the CT/HF0 energy
 compatible with the EFP2 method. Subsequently, after details of computations
-are discussed in Section IV, in Section V the validation of the 
+are discussed in Section~\ref{s:4.calculations}, in Section~\ref{s:5.results} the validation of the 
 OEP\hyp{}based CT/HF0 model is presented and its performance in terms 
 of accuracy and computational speed is compared against the EFP2
-model. Finally, few concluding remarks and comments for future work
-are given in Section VI.
+model. Finally, few concluding remarks and outlook of future work
+are given in Section~\ref{s:6.conclusions}.
 %ERI's, which are oftentimes the main bottleneck
@@
 To establish the notation within the present work, 
-the formalism of one\hyp{}electron potential (OEP) operators is
+the formalism of one\hyp{}electron potential operators is
 given first.
@@
 %
-where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
+where $\lambda$ is a certain parameter and is assumed to be either 1 or 0 in this work.
 Formally, the electronic part can be expanded in terms of an effective
-bond order matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
+bond order matrix, or one\hyp{}particle density (OED), represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
 %
@@
         \int d{\bf r} \Ket{{\bf r}} 
-        v^{\rm eff}({\bf r})
+        v^{\rm eff}_{\rm el}({\bf r})
 	%\left[
@@
 %for any one\hyp{}electron function $f({\bf r})$. 
-and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
+and $\BraKet{\bf r}{\alpha} \equiv \varphi_\alpha({\bf r})$.
 The matrix element of the effective potential operator
@@
 for any one\hyp{}electron operator $\mathscr{O}(1)$ 
-and the electron repulsion integral (ERI)
+and the ERI
 is defined according to
@@
 %or basis function.
-$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
+$\hat{v}_{kl}^A$ is the effective OEP operator given by Eq.~\eqref{e:oep-operator} 
 with the
 effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
-The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
+The summations over $k$ and $l$ can be incorporated into the total effective OEP operator
 $\hat{v}_{\text{eff}}^A$
@@
 %
-Thus, the total computational effort is, in principle, reduced from the fourth-fold
-sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
+Thus, the total computational effort is, in principle, reduced from the fourth\hyp{}fold
+sum involving evaluation of ERI's to the two\hyp{}fold sums of cheaper OEI's.
 It is also possible to generalize the above expression even further by
@@
 %from Eq.~\eqref{e:exp-val-series} 
-are computable and can be approximately
+are well defined, computable and can be approximately
 partitioned in between the interacting fragments.
@@
 
-Two general cases of basis function partitioning scheme
-are considered in this work,
-in which it is possible to define an effective one\hyp{}electron
-potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
-The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
-The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
+Three classes of ERI's can be recognized 
+based on the basis function partitioning scheme
+within the system composed of two molecules (shall be $A$ and $B$
+throughout the course of this work),
+in which it is possible to define an OEP. 
+They are as follows:
+%
+\begin{enumerate}
+\item the Coulomb\hyp{}like ERI's of the type 
+$\BraKet{AA}{BB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in B}\phi_{l\in B}}$,
+\item the overlap\hyp{}like ERI's of the type 
+$\BraKet{AA}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in A}\phi_{l\in B}}$, and
+\item the exchange\hyp{}like ERI's of the type 
+$\BraKet{AB}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{j\in B}}{\phi_{k\in A}\phi_{l\in B}}$.
+\end{enumerate}
+%
+Only the first two classes of ERI's can be incorporated into OEP's,
+while exchange\hyp{}like ERI's cannot.
+The two groups, which are listed in Table~\ref{t:oep-matrix-element-types},
+are usually approximated via
+expanding the OEP operator in distributed multipole (DMTP) expansion series, and integrating over
+one remaining electron coordinate,\cite{Li.Gordon.Jensen.JCP.2006} i.e.,
+%
+\begin{multline} \label{e:coulomb-energy:dmtp-oei}
+ \tBraKet{j}{\hat{v}_{\rm eff}^A}{l} \cong \\
+ \left< j \left| 
+                   \sum_{w\in A} \left\{
+                                 T^{(0)}_{wu} \hat{q}_w 
+                          - {\bf T}^{(1)}_{wu} \cdot \hat{\BM\upmu}_w 
+            + \frac{1}{3}   {\bf T}^{(2)}_{wu} :     \hat{\BM\Theta}_w 
+                    - \ldots  \right\}
+              \right| l \right> \;.
+\end{multline}
+%
+In the above equation, $\hat{q}_{\rm eff}^{(w)}$,
+$\hat{\BM\upmu}^{(w)}_{\rm eff}$ and $\hat{\BM\Theta}^{(w)}_{\rm eff}$
+are the quantum operators of effective distributed monopole (charge),
+dipole moment and quadrupole moment, respectively, centered on the $w$th site, 
+whereas ${{\bf T}^{(r)}_{wu}}$ are the so called interaction tensors of rank $r$ 
+that can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
+Note that $\Ket{j}$ and $\Ket{l}$ can belong to either molecule.
+In this way, ERI's are no longer needed and the computational cost reduces appreciably.
+In this work, this method is however considered already too expensive
+for application in the CT/HF0 energy because evaluation of
+Eq.~\eqref{e:coulomb-energy:dmtp-oei} requires calculation
+of electrostatic potential and electrostatic potential gradient(s) OEI's. These kind of
+integrals are typically the most expensive as compared to other standard OEI's
+such as overlap or kinetic energy integrals.
+Therefore, Coulomb and overlap\hyp{}like OEP matrix elements
+%The first type is a Coulomb\hyp{}like contribution, for which 
+will be treated via more approximate and less expensive approaches
+%the multipole expansion or density fitting
+which are discussed next.
+% can be used.
+%The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
+%Note that the exchange\hyp{}like integrals cannot be treated with the OEP technique.
+
 %
@@
 Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
-Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
+ERI class           &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
 DF\footnotemark[2]/RI\footnotemark[3] Form    
@@
 
-\subsection{\label{ss:2.2.oep-DMTP}Distributed Multipole Expansion of OEP's}
+\subsection{\label{ss:2.2.oep-DMTP}Semi-classical multipole expansion}
 
@@
 because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
-and $\rho^A_{\rm eff}$. 
+and $\rho^A_{\rm eff}$. These matrix elements require ERI's of type
+$\BraKet{AA}{BB}$ only.
 In general, given certain two effective one\hyp{}electron density distributions,
@@
 %
-The above integral can be approximated by the
-distributed multipole (DMTP)
-expansions of the respective densities as follows
+The above integral can be approximated by 
+applying the DMTP expansion 
+to both effective potential operators
+and dropping the `$\hat{}$' symbol in the multipole operators
+which leads to\cite{Stone.TheTheoryOfIntermolecularForces.1996}
 %
@@
 %
-In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
-are the effective distributed monopole (charge)
-and dipole moment, respectively, centered on the $u$th site, 
-whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
-that can be found elsewhere.
 The symbol `$\odot$' denotes the sum of all the tensor contractions
@@
 There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
-of Stone\cite{Stone.JCTC.2005}
-the cumulative atomic multipole moments of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983}
+of Stone\cite{Stone.JCTC.2005},
+the cumulative atomic multipole moments (CAMM) of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983},
 the localised distributed multipole expansion (LMTP) of Etchtebest, Lavery and Pullman\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
@@
 
-\subsection{\label{ss:2.3.oep-GDF}Density Fitting of OEP's}
-
-This scheme is most useful in the case of matrix elements of the type
+\subsection{\label{ss:2.3.oep-GDF}Extended Density Fitting of OEP's}
+
+This scheme, which will be referred to as the EDF scheme, 
+is applicable in case of matrix elements of 
 $
  \tBraKet{i^A}{\hat{v}_{\rm eff}^A}{j^B}
-$.
+$ type. 
+ These matrix elements require ERI's of type
+$\BraKet{AA}{AB}$ only.
 To get the \emph{ab initio} representation of such an overlap\hyp{}like matrix element,
@@
 the typical density fitting or resolution of identity, both of which are nowadays widely used 
-to compute electron\hyp{}repulsion integrals (ERI's) more efficiently.
+to compute electron\hyp{}repulsion integrals (ERI's) more efficiently,
+and reduce computational cost of post\hyp{}Hartree\hyp{}Fock methods. Density fitting was also
+applied to design \emph{ab initio} force fields.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,Cisneros.Andres.Piquemal.Darden.JCP.2005}
 
@@
 %
+Since only one step is required to obtain the OEP matrix, the scheme will
+be abbreviated as EDF-1.
 
@@
 Note that, while $R_{\eta\xi}$ is a typical 2\hyp{}center ERI 
-that can be evaluated by standard means,
+that can be evaluated by standard means,\cite{McMurchie.Davidson.JComputPhys.1978}
 $b^{(i)}_\eta$ matrix elements are not at all trivial to calculate
@@
 by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 
-and impractical even for medium sized molecules. 
+%and impractical 
+even for medium sized molecules.\cite{Barca.Loos.JCP.2017}
 However, one can introduce the effective potential in order to eliminate one integration. 
-This can be achieved by performing additional density fitting in a nearly complete intermediate basis 
+This can be achieved by performing additional density fitting 
+in a nearly complete intermediate basis\cite{Barca.Loos.JCP.2017}  % add more citations here
 %
@@
 %The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.
-
 It is emphasized here that, as long as the state vector $\Ket{i}$, OEP operator $\hat{v}^A$ 
@@
 $V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.
+
+Since, in contrast to the EDF-1 scheme, now two density fitting steps are required to
+obtain OEP matrix, it will be referred to as the EDF-2 scheme.
+In the limitting case, when the auxiliary basis is the same as the intermediate basis,
+EDF-2 is equivalent to EDF-1.
+As the theory required to define OEP's and their computational
+evaluation in the fragment\hyp{}based formulation has been finally given in this Section,
+the application of the OEP technique of ERI elimination is
+further discussed for the evaluation of the CT/HF0 energy in the next sections.
 
@@
 in which the summations extend over occupied (denoted by `Occ'), virtual (denoted by `Vir') 
-or both (denoted by `All') MO's. The effective potential 
+or both (denoted by `All') canonical HF MO's. The effective potential 
 matrices are defined by
@@
 and are evaluated by expanding the $\hat{v}^B_{\rm tot}$ operator
-in distributed multipoles. The resulting calculation requires computation
-of potential, electric field and electric field gradient one\hyp{}electron integrals
-and their contractions with the DMTP distributions.
+in distributed multipoles according to Eq.~\eqref{e:coulomb-energy:dmtp-oei}.
+%The resulting calculation requires computation
+%of potential, electric field and electric field gradient one\hyp{}electron integrals
+%and their contractions with the DMTP distributions.
 The apparent success of the EFP2 scheme is rooted in the dramatic simplifications of the
@@
 have been effectively removed from the working models while maintaining the required accuracy. 
-That is, the canonical orbital energies $\varepsilon_i$ are
+That is, in the case of the CT/EFP2 energy component, 
+the canonical orbital energies $\varepsilon_i$ are
 constant parameters, whereas the overlap $S_{nm}$, kinetic energy $T_{mn}$ and effective
@@
 evaluating Eq.~\eqref{e:u-efp2-ct} is still considerable in cost because typically large basis sets
-need to be used for generating the EFP2 parameters. Moreover, calculation of $V_{in}^{{\rm EF}}$
-is much more expensive as compared to other types of one\hyp{}electron integrals.
+need to be used for generating the EFP2 parameters. In effect, calculation of $V_{in}^{{\rm EF}}$
+is much more expensive as compared to other types of one\hyp{}electron integrals
+and is the bottleneck of CT/EFP2 energy calculation. Recently, Xu and Gordon
+extended the previous formulation by reducing the size of the virtual orbital space
+via the quasiatomic minimal\hyp{}basis orbitals (QUAMBO's).\cite{Xu.Gordon.JCP.2013} 
+Unfortunately, this was still insufficient to overcome
+the high cost of evaluation of the potential integrals, making the CT term still the most time\hyp{}consuming
+to evaluate.
 
@@
 with a set of different approximations, selected based on performance assessment rather than 
-a rigorous derivation manner.
+a rigorous derivation manner.\cite{Li.Gordon.Jensen.JCP.2006}
 Instead, perturbation theory of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
@@
 
-\subsection{\label{ss:3.2.OL} Murrell et al.'s Model: Starting Point}
+\subsection{\label{ss:3.2.OL} Otto-Ladik's Model: Starting Point}
 
@@
  E^{A\rightarrow B} = 2 \sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir} 
-  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n}
+  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n} \;,
 \end{equation}
@@
 %
-where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
-of the occupied and virtual
-HF molecular orbitals, respectively,
-whereas the coupling constant is given by
+%where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
+%of the occupied and virtual
+%HF molecular orbitals, respectively,
+%whereas the coupling constant is given by
+where the coupling constant is given by Otto and Ladik,\cite{Otto.Ladik.ChemPhys.1975}
+here referred to as the OL method,
 %
@@
 from Ref. 
-The goal of this work is to apply the technique sketched in Eq.XXX
+The goal of this work is to apply the technique sketched in Eq.~\eqref{e:ft-reduction}
 to eliminate ERI's from Eq.~\eqref{e:ct-murell-etal.vin} in terms of OEP's.
 
-\subsection{\label{ss:3.3.OEP} Murrell et al.'s Model: Application of OEP Technique}
+\subsection{\label{ss:3.3.OEP} Otto-Ladik's Model: Application of OEP Technique}
 
 One can immediately notice that the five summation terms
-from Eq.~\eqref{e:ct-murell-etal.vin} can be classified into three groups
+from Eq.~\eqref{e:ct-murell-etal.vin} can be classified 
+based on Table~\ref{t:oep-matrix-element-types} 
+into three groups
 regarding the type of ERI's that are required:
@@
 which, by combining with the first term from Eq.~\eqref{e:ct-murell-etal.vin}, 
-allows to apply the density fitting in an
-incomplete basis set as follows
+allows to apply the two\hyp{}step extended density fitting (EDF-2) 
+developed in Section~\ref{sss:2.3.1.GDF-2} as follows
 %
@@
 %
-with the error density defined by
+with the error density given by
 %
@@
 leads to a much simpler formula
-that requires only the one\hyp{} and two\hyp{}electron integrals. 
+that requires only OEI's and ERI's.
+%the one\hyp{} and two\hyp{}electron integrals. 
 Therefore, by combining
@@
 because only the overlap integrals between the $i$th MO on molecule $A$
-and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated on the fly.
+and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated. % on the fly.
 Note that the only approximation made so far was the application of density fitting
@@
 basis sets are sufficiently large, the errors
-due to this approximation can be minimal and negligible.
+due to this approximation can be minimal and negligible in principle.
 
@@
 We shall refer to this approximation as to
-the spherical overlap approximation (SOA). Note that in order for this approximation to be valid,
-occupied molecular orbitals need to be localized.
+the localized overlap approximation (LOA). Note that in order for this approximation to be valid,
+occupied molecular orbitals need to be spatially localized.
 
@@
 Note now that both effective densities in 
-$\rho^B_{jn} \odot \rho^{A,{\rm eff}}_{ik}$ are generally complex and one cannot apply the SOA
-for $\rho^{A,{\rm eff}}_{ik}$. In the LMO picture however, it can be shown that
-$\rho_{ik} ({\bf r}) \ll \rho_{ii} ({\bf r})$ 
+$\rho^B_{jn} \odot \rho^{A,{\rm eff}}_{ik}$ are generally complex and one cannot apply the LOA
+for $\rho^{A,{\rm eff}}_{ik}$. 
+In the LMO picture however, it can be shown that
+$\lvert\rho_{ik} ({\bf r}) \rvert \ll \rho_{ii} ({\bf r}) $ 
 for $i\ne k$ in most locations. Therefore,
-the SOA is still applicable in the LMO picture
+the LOA is still applicable in the LMO picture
 and the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
@@
 \paragraph{Final OEP-based forms of the coupling constant.}
-Gathering the results from Eqs. the coupling constant
+Gathering the results from Eqs.~\eqref{e:ct.group-i.final},
+\eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.b} the coupling constant
 can be given as follows
@@
 %
-\begin{subequations}
-\begin{align} \label{e:ct-murell-etal.vin.oep.aux-var}
+\begin{subequations} \label{e:ct-murell-etal.vin.oep.aux-var}
+\begin{align} \label{e:ct-murell-etal.vin.oep.aux-var.u} 
  u_i^{BA} &\equiv    
@@
                 \;.
+  \label{e:ct-murell-etal.vin.oep.aux-var.w}
 \end{align}
@@
 %
-To compute the interaction energy due to CT from $A$ to $B$ by using the above SOA\hyp{}based
+To compute the interaction energy due to CT from $A$ to $B$ by using the above LOA\hyp{}based
 approximations, coupling elements need to be transformed from localized to canonical MO basis, i.e.,
@@
 
-Four selected molecular pairs composed of H$_2$O, CH$_3$OH, NH$_4^+$ and NO$_3^-$
-(Figure 1) molecules or ions were chosen to validate the OEP\hyp{}based model of the CT energy
-at the PT/HF formulation level. Structures of the molecular dimers were first 
-energy\hyp{}optimized at the HF/6-31+G(d,p) level by using
-the Gaussian 16 quantum chemistry program package.\cite{Gaussian16}
-To obtain the PT/HF benchmark CT energy, it was estimated
-from the following approximation:
+Four complexes: 
+(i) (H$_2$O)$_2$, 
+(ii) H$_2$O--CH$_3$OH, 
+(iii) H$_2$O--NH$_4^+$ and 
+(iv) NO$_3^-$--NH$_4^+$,
+were chosen as model systems to analyse the asymptotic dependence 
+of CT/HF0 energy. 
+The reference (zero\hyp{}displacement) geometries
+were obtained by
+energy\hyp{}optimizations at the HF/6-31+G(d,p) level,
+as implemented in 
+the {\sc Gaussian~16} quantum chemistry program package.\cite{Gaussian16}
+Subsequently, 30 displaced geometries for each model complex
+were obtained by translating one of the monomers along the vector 
+co\hyp{}linear with the H-bond or N--N distance in the case of ammonium nitrate.
+The reference structures as well as the translation vectors are
+indicated in the insets of Figure~\ref{f:fig-1} and S1.
+To perform statistical error analysis, structural databases
+of bi\hyp{}molecular complexes in the non\hyp{}covalent
+interactions database NCB31 developed by the Truhlar's 
+group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
+interactions in proteins,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+as implemented in the {\sc Psi4}
+program,\cite{Psi4.JCTC.2017}
+were utilized.
+In particular, the subsets from the NCB31 database were separately considered:
+the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+
+The benchmark CT/HF0 energy was estimated
+from the following assumption:
 %
 \begin{equation}\label{e:ct-pthf-ref}
- \Delta E^{\rm CT}_{\rm PT/HF} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
+ \Delta E^{\rm CT}_{\rm HF0} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
  - \Delta E^{\rm Ind} \;,
@@
 in the DDS method reproduce the first\hyp{}order Coulombic and exchange\hyp{}repulsion
-interaction energy in the symmetry adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994}
-(SAPT)
+interaction energy in SAPT\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994}
 with accuracy below 0.3 kcal/mole.
 Therefore, Eq.~\eqref{e:ct-pthf-ref} is believed to be sufficient
-in order to compute the PT/HF CT energy, because there is no dispersion
+for evaluating the CT/HF0 energy, because there is no dispersion
 contribution to the DDS polarization energy at the HF level.
@@
 \begin{equation}\label{e:ind-ref}
- \Delta E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\bf D}^{-1} \cdot {\bf F} \;.
+ \Delta E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\BM \Delta}^{-1} \cdot {\bf F} \;.
 \end{equation}
@@
 %where $a$ labels the LMO centroids.
-The ${\bf D}$ matrix is defined as
+The matrix elements of ${\BM \Delta}$ are defined as
 %
 \begin{equation} \label{e:ind-ref.d-matrix}
- {\bf D}_{ab} = 
+ {\BM \Delta}_{ab} = 
 \left\{\begin{matrix}
@@
     \frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3}
-               \right| \alpha \right>_{\bf r} \;.
-\end{equation}
-%
+               \right| \alpha \right>_{\bf r} \;,
+\end{equation}
+%
+in which ${\bf D}$ is the one\hyp{}particle density matrix.
 The distributed anisotropic dipole\hyp{}dipole
@@
 method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
-Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
-
-All the models that were used to test the theory presented in this work
+It was found here that accuracy of the LOA 
+from Eqs.~\eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.b}
+is usually better 
+when the Boys method\cite{Boys.RevModPhys.1960} is used, as compared
+to the Pipek\hyp{}Mezey method\cite{Pipek.Mezey.JCP.1989}. Henceforth, the former
+method was used for molecular orbital localization throughout all the production calculations.
+
+All the models that were used to test the theory presented in this work,
+i.e., the OL, EFP2 and OEP methods, as well as the benchmark EDF, CT/HF0 and DDS methods,
 were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
+DDS method was implemented in the dimer\hyp{}centered basis set
+in order to eliminate the BSSE and properly inlcude the CT processes.\cite{Stone.Misquitta.CPL.2009}
 
@@
 
-First, the asymptotic dependence of the CT energy
-was investigated for four chosen molecular pairs and the results are
-shown in Figure~\ref{f:fig-2} and Figure~S1 for the methanol-water system. 
+%
+\begin{figure}[h]
+\includegraphics[width=0.44\textwidth]{fig-1.pdf}
+\caption{\label{f:fig-1} {\bf Asymptotic dependence of the charge transfer energy
+in the CT/HF0 formulation for selected bi\hyp{}molecular complexes.} 
+%The charge transfer energy estimates are plotted for 
+(a) water dimer, 
+(b) water\hyp{}ammonium complex, and 
+(c) ammonium\hyp{}nitrate complex,
+were one molecule has been translated
+along the vector specified in the insets relative to initial geometry,
+optimized at HF/6-31+G(d,p) level.
+The 
+%counterpoise\hyp{}corrected 
+total interaction energy
+is also shown for comparison in light grey color in this figure.
+Interaction energies were obtained at the HF/6-311++G(2df,2pd) level of theory
+with the single GDF scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
+} 
+\end{figure}
+%
+The asymptotic dependence of the CT energy
+is
+shown in Figure~\ref{f:fig-1} and Figure~S1 for the methanol\hyp{}water system. 
 Overall, OEP technique of effective
 elimination of ERI's sketched in Eq.~\eqref{e:ft-reduction} 
-works
-quantitatively well for water dimer and ammonium nitrate (Figures~\ref{f:fig-2}(a) and (c)),
-and methanol-water (Figure~S1)
-reproducing the reference Otto\hyp{}Ladik CT energy,
-whereas in the case of water\hyp{}ammonium cation complex the OEP
+is sufficiently accurate
+for water dimer and ammonium nitrate (Figures~\ref{f:fig-1}(a) and (c)),
+and methanol\hyp{}water system (Figure~S1) in all distances studied,
+reproducing the reference Otto\hyp{}Ladik CT energy, with errors not
+exceeding 10\% relative to OL estimates (compare open and filled green circles).
+However, in the case of water\hyp{}ammonium cation complex the OEP
 method overestimates the CT energy by around 50\%, even for large
-intermonomer separations. Careful inspection
+intermonomer separations, where the profile exhibits
+unphysical long\hyp{}range tail, visible in Figure~\ref{f:fig-1}(b). 
+Careful inspection
 (see SI)
-revealed that the reason is the SOA in $G_2$ and $G_3$ and not
-the generalized density fitting in $G_1$ terms, in the case of which
-the aug-cc-pVDZ-jkfit basis set was sufficient as auxiliary space of OEP transformation.
-On the other hand, analysis of the statistical sets of dimeric complexes
-shown in Figure~XXX
-confirmed that in the great majority of cases the 
-OEP technique gives reasonably accurate results with respect
-to the parent model with relative errors less than 10\%,
-unless systems are charged.
+revealed that increasing the size of the 
+auxiliary basis set from aug-cc-pVDZ-jkfit to aug-cc-pVQZ-jkfit
+slightly improves the performance of the OEP method, however the long\hyp{}range
+tail is still appreciable in magnitude even in the aug-cc-pVQZ-jkfit auxiliary basis.
+
+At the moment, it is unclear what causes these errors at large separations. 
+The LOA in $G_2$ and $G_3$ may be too drastic approximations
+when one of the monomers is charged, and more accurate
+DMTP expansion for the effective densities could be required,
+also including dipole and higher multipole moments. % and as well as finer distribution of centres.
+Interestingly in the case of ammonium nitrate (Figure~\ref{f:fig-1}(c)),
+the relative errors between OEP and OL estimates
+are not grater than 9\% even though both of the monomers are charged.
+On the other hand, OL method itself strongly overestimates the CT/HF0
+energies in this system. Therefore, among the four
+model complexes, EFP2 and OEP energies
+coincide with the CT/HF0 estimates 
+only in the case of water dimer and methanol-water complex,
+in contrast to the EFP2 model, which gives very good estimates
+of the CT/HF0 energy in mostly all cases at all separations.
+
+Analysis of the statistical sets of dimeric complexes
+shown in Figures~\ref{f:fig-2}(a) and Figures~\ref{f:fig-3}(a)
+suggests that, at least in the case of neutral systems,
+ERI elimination technique (compare OL and OEP)
+is of acceptable accuracy with the $R^2$ coefficients
+being 74\% for the NCB31, and 62\% for the BBI dataset. 
+The absolute accuracy of the OEP model, measured with respect to the benchmark
+CH/HF0 estimates, is shown in Figures~\ref{f:fig-2}(b) and Figures~\ref{f:fig-3}(b)
+and the root mean square errors (RMSE)
+are listed in Table~\ref{t:2}. In NCB31 set OEP performs better than EFP2
+with
+RMSE of 1.69 and 2.39 kcal/mol, respectively. This is especially
+the case for charge\hyp{}transfer\hyp{}dominated systems (subset CT7/04),
+in which EFP2 erroneously predict vanishing or even small
+positive CT energiers (RMSE 4.06), in contrast to OEP method that gives correct
+signs in all cases and good correlation with the benchmark.
+On the other hand, EFP2 method very accurately reproduces the CT/HF0 energies
+in the BBI set (R$^2=95\%$) whereas OEP model tends to significantly overestimate
+the values by a factor of 1.55 with RMSE of 0.56 kcal/mol, 
+unlike the OL model which is more accurate
+(RMSE of 0.22 and R$^2=78\%$). 
 %
@@
 \renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
+\begin{table}[t]
 \caption[Accuracy of approximate CT/HF0 energy methods across wide range of bi\hyp{}molecular complexes]
@@
 
-In the case of water dimer and methanol-water complex, EFP2 and PT/HF estimations
-coincide with the OEP model. However, in the presence of ions, 
-the deviation of the OEP model from the PT/HF energy 
-become significant, in contrast to the EFP2 model, which gives very good estimates
-of the PT/HF energy in mostly all cases.
-Naturally, the accuracy of the Otto\hyp{}Ladik model limits the absolute accuracy
+Since the nature of the EFP2 and OEP models is highly approximate,
+and the most important goal is to construct efficient and reasonably
+accurate \emph{ab initio} force field, one can use the fact that
+OEP model almost always overestimates the CT/HF0 energies
+with rather good linear correlation,
+unlike EFP2 model, which can also strongly underestimate it.
+Therefore, the simple semi\hyp{}empirical scaling model is proposed here
+for the OEP\hyp{}based estimate, i.e.,
+%
+\begin{equation} \label{e:ect-oep.final-scaled}
+ \Delta E^{\rm CT} \approx C \times \Delta E^{\rm OEP}
+\end{equation}
+%
+where $C$ is an empircial constant and $\Delta E^{\rm OEP}$ is given by
+Eqs.~\eqref{e:ct-hf0} and Eq.~\eqref{e:ect-oep.final}. Based on
+OEP performance from NCB31 and BBI sets, it was estimated
+that $C\approx 1.55$. Application of the semi\hyp{}empirical model
+from Eq.~\eqref{e:ect-oep.final-scaled} significantly improved
+absolute accuracy of the model in all datasets. 
+For instance, RMSE was reduced down to 0.13 kcal/mol in the BBI set,
+comparable to EFP2 model.
+%but still with somewhat large RMSE of 2.82 kcal/mol.
+%confirmed that in the great majority of cases the 
+%OEP technique gives reasonably accurate results with respect
+%to the parent model with relative errors less than 10\%,
+%unless systems are charged.
+Naturally, the accuracy of the OL model limits the absolute accuracy
 of the OEP model, which is manifested in the ammonium nitrate system
-in Figure~\ref{f:fig-2}(c) or the problems with SOA. 
+in Figure~\ref{f:fig-1}(c) or the problems with LOA. 
 In order to improve the performance for ionic systems,
 further studies are necessary, that will modify the original formulation
-of Otto and Ladik and SOA treatment. Nevertheless, except from the ionic systems,
-it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable
-(Figure~XX) across a wide range of various interacting complexes including 
+of OL and LOA treatment. Nevertheless, except from the ionic systems,
+it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable,
+especially after applying the empirical scaling,
+across a wide range of various interacting complexes including 
 systems with H-boding, dominated by dipole-dipole or dispersion interactions or 
 exhibiting substantial CT.
-%
-\begin{figure}[h]
-\includegraphics[width=0.46\textwidth]{fig-2.pdf}
-\caption{\label{f:fig-2} {\bf Asymptotic dependence of the charge transfer energy
-in the CT/HF0 formulation for selected bi\hyp{}molecular complexes.} 
-The charge transfer energy estimates are plotted for 
-(a) water dimer, 
-(b) water\hyp{}ammonium complex, and 
-(c) ammonium\hyp{}nitrate complex,
-where one molecule has been translated by $R$ from the starting geometry
-optimized at HF/6-31+G(d,p) level,
-along the vector specified in the insets.
-The counterpoise\hyp{}corrected total interaction energy
-is also shown for comparison in light grey color in this figure.
-Interaction energies were obtained at the HF/6-311++G(2df,2pd) level of theory
-with the single GDF scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
-} 
-\end{figure}
-%
+%This
+%indicates that further investigation is required for charged systems
+%which is out of the scope of the present work.
+
+%However, in the presence of ions, 
+%the deviation of the OEP model from the CT/HF0 energy 
+%become significant,
 
@@
 \begin{figure*}[h]
-\includegraphics[width=0.9\textwidth]{fig-3.pdf}
-\caption{\label{f:fig-3} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
+\includegraphics[width=0.9\textwidth]{fig-2.pdf}
+\caption{\label{f:fig-2} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
 in small bi\hyp{}molecular neutral complexes.}
@@
 (black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-6-311++G** basis set was used for the CT energy calculations
+6-311++G(d,p) basis set was used for the CT energy calculations
 and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
@@
 \begin{figure*}[h]
-\includegraphics[width=0.9\textwidth]{fig-4.pdf}
-\caption{\label{f:fig-4} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
+\includegraphics[width=0.9\textwidth]{fig-3.pdf}
+\caption{\label{f:fig-3} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
 in aminoacid pairs.}
@@
 The structures of 100 aminoacid pairs
-were randomly selected from the BBI Bioo database
-for backbone\hyp{}sidechain interactions.\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}
+were taken from the BBI subset of the BioFragment Database
+for backbone\hyp{}backbone interactions 
+in proteins.\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}
 6-31+G(d,p) basis set was used for the CT energy calculations
-and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
+and the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
 for the OEP calculations.
@@
 
+%
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[t]
+\caption[Estimated computational cost of the EFP2 and OEP methods for calculation of CT/HF0 energy]
+{{\bf Estimated computational cost of the EFP2 and OEP methods for calculation of CT/HF0 energy\footnotemark[1]}
+}
+\label{t:oep-costs}
+\begin{ruledtabular}
+\begin{tabular}{lccc}
+Method     && EFP2 & OEP \\
+\hline
+%Constants  && $\varepsilon_i$ 
+%            & $\varepsilon_i$, $\varepsilon_n$ \\
+Constant   && \multirow{2}{*}{$\varepsilon_i$} 
+            & \multirow{2}{*}{$\varepsilon_i$, $\varepsilon_n$} \\
+parameters && & \\
+\hline
+%\multirow{2}{*}{Rotatable}  
+Superimposable  && \multirow{2}{*}{$C_{\alpha i}$, $C_{\alpha n}$, $\{\alpha\}$} 
+            & $C_{\alpha i}$, $C_{\alpha n}$, $\{\alpha\}$, \\
+parameters && & $\{\eta\}$, $V_{n\eta}^{{\rm eff},B}$ \\
+\hline
+\multirow{3}{*}{Calculables} 
+           && $S_{ij}$, $S_{nk}$, $S_{nw}$,            &   $S_{\eta i}$, $S_{ni}$, $S_{ij}$, \\
+           && $T_{nn}$, $T_{kj}$, $T_{wj}$, $T_{nj}$   &  $u_i^{BA}$, $w_{yi}^{BA}$ \\
+           && $V_{in}^{{\rm EF},B}$ , $V_{ik}^{{\rm EF},B}$, $V_{iw}^{{\rm EF},B}$
+                                                       & \\
+\hline
+Cost       && $(10on^2 + 2n^3 + o^2n^2)v$ & $on(Q+N+o)$ \\ 
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
+$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
+%
+\end{table}
+}
+%
+
 \subsection{\label{ss:5.2.cost}Reduction of computational costs}
@@
 The utmost goal of this work is to reduce the computational cost of
-the CT PT/HF energy evaluation in the calculations involving effective fragment
-potentials. The EFP2 CT expression has the computational cost
+the CT/HF0 energy evaluation in the calculations involving effective fragment
+potentials, no longer making it the bottleneck of the EFP2\hyp{}based interaction
+energy calculations. In Table~\ref{t:oep-costs} estimation of the
+computational cost of the EFP2 and OEP models is shown.
+It is apparent that EFP2 has much more quantities to compute on the fly
+as compared to the OEP method (`Calculables' in the table).
+Clearly,
+evaluation of the EFP2 CT expression from Eq.~\eqref{e:u-efp2-ct} 
+requires quite a number of different types of OEI's. According to our estimations
+that assume sequential (two\hyp{}step) two\hyp{}index AO\hyp{}MO transformations of OEI matrices
+and large AO basis sets,
+the computational cost is
 roughly proportional to $(10on^2 + 2n^3 + o^2n^2)v$, where the
 $o$ and $n$ denote the number of occupied orbitals and the number of atomic basis functions,
-respectively, whereas $v$ is the relative cost. On the other hand, OEP\hyp{}based expression
-from Eq.~\eqref{e:ect-oep.final} has the cost of approximately $on(Q+N+o)$, where 
-$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
-respectively. 
+respectively. $v$ is the relative cost of evaluation of the
+most expensive type of OEI's, i.e., the potential and potential gradient(s)
+integrals contracted with the DMTP's, that are necessary to compute ${\bf V}^{\rm EF}$ matrices
+from Eq.~\eqref{e:v-efp2-eff}. 
 %
@@
 \renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
-\caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
-{{\bf}Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.\footnotemark[1]
+\begin{table}[t]
+\caption[Minimal uncontracted auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations]
+{{\bf Minimal uncontracted auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations\footnotemark[1]}
 }
@@
 %
-\footnotetext[1]{Fitting performed according to Appendix.~\eqref{a:auxiliary-basis} assuming 6-311++G(d,p) 
+\footnotetext[1]{Fitting performed according to Appendix~\eqref{a:auxiliary-basis} assuming 6-311++G(d,p) 
 and aug-cc-pVDZ-jkfit primary and test basis sets, respectively.}
@@
 %
-
-In this section, computational cost is estimated from the formulas.
-%In
-%Table~\ref{t:oep-costs} 
-%
-%{
-%\renewcommand{\arraystretch}{1.4}
-%\begin{table}[b]
-%\caption[Computational cost of the benchmark and OEP-based methods: calculation of coupling constant]
-%{{\bf Computational cost of the benchmark and OEP-based methods\footnotemark[1]}
-%}
-%\label{t:oep-costs}
-%\begin{ruledtabular}
-%\begin{tabular}{lcccccc}
-%Method             && Murrell et al. &&    EFP2        &&      OEP     \\ 
-%	\cline{1-7}
-%Group (i)         &&  &&        &&   $obQ$     \\
-%Group (ii)        &&  &&        &&   $ob$      \\
-%Group (iii)       &&  &&        &&   $ob(n+o)$     \\
-%Total Cost        &&  && $10ob^2 + 2b^3 + o^2b^2$       &&   $ob(Q+n+o)$     \\
-%\end{tabular}
-%\end{ruledtabular}
-%%
-%\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
-%$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
-%%
-%\end{table}
-%}
-%
+On the contrary, 
+OEP\hyp{}based expression
+from Eq.~\eqref{e:ect-oep.final} requires only overlap OEI's that are the least expensive, 
+and has the cost of approximately $on(Q+N+o)$, where 
+$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
+respectively. Note also that, among the calculables that are needed on the fly in each
+OEP\hyp{}based CT energy evaluation, are the auxiliary vectors and matrices 
+from Eqs.~\eqref{e:ct-murell-etal.vin.oep.aux-var.u}
+and \eqref{e:ct-murell-etal.vin.oep.aux-var.w},
+the cost of which is negligible. The amount of effective fragment parameters,
+that need to be superimposed during the calculations by applying rotation
+of orbitals and basis functions (`Superimposable parameters')
+is rather the same in EFP2 and OEP models. 
+This includes the LCAO\hyp{}MO coefficients in canonical basis,
+and the primary basis set, with an addition of the auxiliary basis set
+for the OEP model. Therefore, the cost of parameter superimposition should
+not be significantly larger as in the EFP2 formulation, provided sufficiently 
+small auxiliary basis set is used.
 {
@@
 \caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
-{{\bf}CPU timings in miliseconds of CT/HF0 single point energy calculations
-for water-methanol complex at 6-311++G(d,p) primary basis set\footnotemark[1]
+{{\bf CPU timings in miliseconds of CT/HF0 single point energy calculations
+for water-methanol complex at 6-311++G(d,p) primary basis set\footnotemark[1]}
 }
-\label{t:4.oep-costs}
+\label{t:5}
 \begin{ruledtabular}
@@
 \cline{3-5}
-OL                    && 774    &&\\
-EFP2                  &&  42.3  &&\\
-OEP/aug-cc-pVQZ-jkfit &&   4.7  &&\\
-OEP/aug-cc-pVTZ-jkfit &&   4.0  &&\\
-OEP/aug-cc-pVDZ-jkfit &&   3.8  &&\\
-OEP/6-311++G(d,p)     &&   3.0  &&\\
-OEP/mini              &&   2.5  &&\\
+OL                    &&14749.1 &&  2909.9\\     %14716.1 &&  2876.9 added 33 sec and 17 sec for ERI/AO calculation from timer.dat
+EFP2                  &&  42.3  &&    15.4\\
+OEP/aug-cc-pVQZ-jkfit &&   4.7  &&     1.8\\
+OEP/aug-cc-pVTZ-jkfit &&   4.0  &&     1.5\\
+OEP/aug-cc-pVDZ-jkfit &&   3.8  &&     1.4\\
+OEP/6-311++G(d,p)     &&   3.0  &&     1.0\\
+OEP/mini              &&   2.5  &&     0.75\\
 \end{tabular}
@@
 %
-\footnotetext[1]{CPU: XXX. See the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
+\footnotetext[1]{1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor, calculations performed on 1 core. 
+See also the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
 %
@@
 
-%
-It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
-and many orders of magnitude lesser than the original theory of Murrell et al.
-%
-\begin{figure}[h]
-\includegraphics[width=0.46\textwidth]{fig-5.pdf}
-\caption{\label{f:fig-5} {\bf Asymptotic dependence of the charge transfer energy
+In this work, the minimal uncontracted basis sets were optimized for water and methanol
+that consist of only $s$ and $p$\hyp{}type functions,
+and the orbital exponents are shown in Table~\ref{t:3},
+and the procedure is outlined in Appendix~\ref{a:auxiliary-basis}.
+Unfortunately, we were not able to successfully fit the auxiliary basis sets
+for NH$_4^+$ and NO$_3^-$ using current approach
+and the basis set optimization falls out of scope of this work.
+Nevertheless, we found that in the case of neutral systems minimal basis
+sets are already sufficient to reach required accuracy.
+In Figure~\ref{f:fig-4} the asymptotic dependence of the CT/HF0 energy
+for water\hyp{}methanol system is correctly reproduced by the OEP/mini variant
+which
+performs comparably well as the OEP/aug-cc-pVQZ-jkfit. 
+The total time required for evaluation of the CT/HF0 energies by using EFP2 and OEP
+models was measured for this system and also for the water-water system in their reference geometries,
+and the results are shown in Table~\ref{t:5}. 
+Profiling of the code for OL, EFP2 and OEP methods was performed 
+for all the computational operations required for a
+single point calculation in a hypothetical sequential run on multiple
+geometries. Therefore, calculations of ERI's in AO basis 
+along with their four\hyp{}index transformation to MO basis were included in
+time profiling of the OL routine in our computer code. For the EFP2 
+and OEP methods, all the calculables were taken into account in the profiling,
+i.e., calculations of OEI's in AO basis and their two\hyp{}index transformations
+to MO basis. Calculations of canonical orbital 
+energies and LCAO\hyp{}MO coefficients through the self\hyp{}consistent field
+HF procedure, as well as calculations of the
+OEP matrices from Eq.~\eqref{e:v-oep.ct} 
+via EDF-2 scheme and atomic effective charges from Eq.~\eqref{e:a1-v.oep-camm.q}
+using the OED in Eq.~\eqref{e:oed-group-iii}
+was not included because these quantities constitute the effective fragments
+by definition.
+The above profiling settings allows one to compare the performances
+of OL, EFP2 and OEP methods as effective fragment parameter methods.
+
+Even for the rather large auxiliary quadruple\hyp{}$\zeta$ basis set,
+the OEP\hyp{}based model is roughly 9 times more efficient than evaluating the EFP2 CT energy.
+Reducing the size of auxiliary basis set up to minimal basis from Table~\ref{t:3}
+scales down the cost further by a factor of $1.9$ in water-methanol system,
+and $2.4$ in water dimer system, reaching in total roughly 20\hyp{}fold
+reduction in CPU time as compared to the EFP2 model, and more than three orders of magnitude
+as compared to the OL model. Since the size of such a minimal basis
+is very small as compared to the primary basis sets recommended for constructing
+the EFP2 parameters, the additional superimposition costs of the auxiliary basis
+should be negligible.
+It is also worth mentioning here that, while using the usual 6-311++G(d,p) basis
+in a role of auxiliary basis set 
+gave already accurate results, the STO-3G basis as auxiliary space resulted
+in CT energies of around $-8$ kcal/mol (data not shown). Therefore, the basis set optimization
+described in Appendix~\ref{a:auxiliary-basis} was necessary for the minimal basis to be applicable
+in the CT calculations using OEP technique, with OEP's generated by the EDF-2 scheme
+developed in Section~\ref{sss:2.3.1.GDF-2}. 
+%
+%It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
+%and many orders of magnitude lesser than the original theory of Murrell et al.
+%
+\begin{figure}[t]
+\includegraphics[width=0.46\textwidth]{fig-4.pdf}
+\caption{\label{f:fig-4} {\bf Asymptotic dependence of the charge transfer energy
 in the CT/HF0 formulation for methanol-warer complex with minimal OEP auxiliary basis.} 
@@
 along the vector specified in the insets.
-The counterpoise\hyp{}corrected total interaction energy
+The total interaction energy
 is also shown for comparison in light grey color in this figure.
@@
 whereas `OEP/mini` denotes double GDF scheme with the minimal auxiliary basis
-developed here for water and methanol.
+given in Table~\ref{t:3}.
 } 
@@
 theories of intermolecular interactions
-was derived and applied to calculating the charge transfer energy
-in condensed phases simulations. It was shown that in general case two types of OEP's
+was presented.
+It was shown that in general case two types of OEP's
 can be defined and worked out either through the density fitting or the distributed
-multipole expansion. For the first group of OEP's, the density fitting was generalized
-and two computational treatments of OEP's were developed: the single GDF scheme
-for RI\hyp{}type auxiliary basis sets, as well as the double GDF scheme, for
-essentially all other types of basis sets. The presented validation of the OEP technique
+multipole expansion. For the first group of OEP's, the density fitting was extended
+and two computational treatments of OEP's were developed: the EDF-1 scheme
+for RI\hyp{}type auxiliary basis sets, as well as the EDF-2 scheme, for
+essentially all other types of basis sets. 
+The OEP technique was then applied to calculating the charge transfer energy
+in condensed phases simulations. 
+The presented validation of the OEP technique
 against the Otto\hyp{}Ladik CT/HF0 model as parent theory showed that in most cases
@@
 where only qualitative accuracy was found. 
-Finally, it was found that, while the proposed OEP\hyp{}based Otto\hyp{}Ladik
+Finally, it was concluded that, while the proposed OEP\hyp{}based Otto\hyp{}Ladik
 model is of comparable accuracy as the CT/HF0 formulation within the
 state\hyp{}of\hyp{}the\hyp{}art EFP2 model, it significantly outperforms the latter
-in terms of computational efficiency reaching speedups up to 30--40 times.
+in terms of computational efficiency reaching speedups up to 20 times.
+It should be emphasized that there is still more room for optimization of the
+OEP technique. First of all, not only group (i) terms, but also
+all the summations over the virtual orbitals can be optimized,
+e.g., by using the QUANBO's, following Xu and Gordon.\cite{Xu.Gordon.JCP.2013}
+In addition, effective atomic charges are most of the time very small
+and could be neglected, leading to further cost reductions.
 %Since EFP2 method is based on the PT/HF0 foundations
@@
 the full (CT\hyp{}including) EFP2 energy calculations and 
-molecular dynamics simulations in large systems. However, further 
-study is necessary and the following issues need to be carefully addressed:
+molecular dynamics simulations in large systems. However, the current model
+needs to be improved by further investigating the ionic
+systems in detail and possibly revising the LOA and EDF approximations.
+Nevertheless,
+it is anticipated here that the OEP method of ERI elimination can be used in 
+virtually any other 
+\emph{ab initio} fragment\hyp{}based
+approaches for condensed\hyp{}phase simulations, where ERI pose the
+computational challenge when confronted with the size of the system.
+
+%However, further 
+%study is necessary and the following issues need to be carefully addressed:
 
@@
 \begin{subequations} \label{e:a1-v.oep-camm}
- \begin{align}
+ \begin{align} \label{e:a1-v.oep-camm.q}
   q_x^{\rm eff} &= \lambda Z_x - 

### 2019-11-29 papers/2_oep_technique/paper-r2.tex
Paper 2 added plotting sources

@@
 \begin{equation}\label{e:u-efp2}
- U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
+ U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;,
 \end{equation}
@@
  b^{(i)}_\eta &= \iint 
-                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 
+                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v}^A \phi_i({\bf r}_2) } 
                             {\vert {\bf r}_1 - {\bf r}_2\vert}  
@@
 the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+and 
+the PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
 
@@
 (yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-and PPS5/05 the $\pi$-$\pi$ stacking database 
+and the PPS5/05 the $\pi$-$\pi$ stacking database 
 (black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
@@
 
-Even for the rather large auxiliary quadruple\hyp{}$\zeta$ basis set,
+Even for the rather large quadruple\hyp{}$\zeta$ auxiliary basis set,
 the OEP\hyp{}based model is roughly 9 times more efficient than evaluating the EFP2 CT energy.
@@
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
+Wroclaw Centre for Networking and Supercomputing (WCSS) is acknowledged for
+computational resources.
 We thank Professor Marcos Mandado from University of Vigo in Spain
@@
 %from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
-by expanding the MO's in terms of atomic basis functions. The working formulae
-are given below:
-%
-\begin{subequations} \label{e:a1-v.oep}
- \begin{align}
-   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm Rep}}{i}
-     &= \sum_{x}^{\rm At} W_{\alpha i}^{(x)} \nonumber   \\ 
-        + \sum_{\beta\gamma\delta} &
-           \left\{ 
-             2 C_{\beta i} D_{\gamma\delta} - C_{\gamma i} D_{\beta \delta}
-           \right\}
-           \BraKet{\alpha\beta}{\gamma\delta} \;,\\
-   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{n} 
-     &= \sum_{x}^{\rm At} W_{\alpha n}^{(x)} \nonumber   \\
-       + \sum_{\beta\gamma\delta} &
+by expanding the MO's in terms of atomic basis functions. The working formula
+is %given below:
+%
+\begin{multline} \label{e:a1-v.oep}
+   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{{\rm G}_1}}{n} 
+     = \sum_{x}^{\rm At} W_{\alpha n}^{(x)}   \\
+       + \sum_{\beta\gamma\delta} 
            \left\{
@@
            \BraKet{\alpha\beta}{\gamma\delta}
-    \;.
- \end{align}
-\end{subequations}
-%
-In the above equations, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
+%    \;.
+\end{multline}
+%
+for the OEP operator defined by 
+%
+\begin{equation}
+ \sum_{\eta\in B}^{\rm DF} \hat{v}_{\rm eff}^{{\rm G}_1} \Ket{\eta} \equiv 
+ \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta} \;.
+\end{equation}
+%
+In Eq.~\eqref{e:a1-v.oep}, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
 is the one\hyp{}particle density matrix in AO basis.
@@
    \left\{ 
-    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\bf r}{\beta} 
+    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\hat{\bf r}}{\beta} 
    \right\}

### 2019-11-29 papers/2_oep_technique/do-r2.sh

@@ -2,7 +2,7 @@
 base=paper-r2
 si=supplementary-r2
 
-for i in $base #$si
+for i in $base $si
 do
   pdflatex  ${i}.tex
   bibtex ${i}.aux

### 2019-11-29 papers/2_oep_technique/figures/make.fig-1.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+figure_dir="./"
+
+gnuplot plot.fig-1.gpl
+
+for plot in fig-1 ; do
+    epstopdf ${plot}-inc.eps
+    sed -f ${plot}.sed ${plot}.tex > ${plot}-s.tex
+    pdflatex ${plot}-s.tex
+    pdftops -eps -r 600 ${plot}-s.pdf
+    #mv -v ${plot}.pdf ${plot}.eps ${figure_dir}
+    mv -v ${plot}-s.pdf ${plot}.pdf
+    rm -v ${plot}-s.aux ${plot}-s.tex ${plot}-s.log ${plot}-inc*
+    #rm -v ${plot}.aux ${plot}.tex ${plot}.log 
+    mv -v ${plot}-s.eps ${plot}.eps
+done

### 2019-11-29 papers/2_oep_technique/figures/water_methanol.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+pymol -p <<EOF
+load ../xyz/water-methanol.xyz
+set_view (\
+     0.020379238,    0.412595391,   -0.910679519,\
+    -0.174068779,   -0.895494044,   -0.409611315,\
+    -0.984519064,    0.166866705,    0.053572617,\
+     0.000005444,    0.000003511,  -14.221484184,\
+    -0.747403860,    0.374430537,    1.037632346,\
+    11.221332550,   17.239648819,  -20.000000000 )
+set stick_radius, 0.06
+set sphere_scale, 0.15
+show sticks
+show spheres
+set transparency, 0.9
+set opaque_background, off
+set ray_opaque_background, off
+ray 4000, 4000
+png water_methanol, dpi=1200
+EOF
+

### 2019-11-29 main.cc
Maintenance of the code - adding more options for EDF-2 auxiliary basis optimization

@@ -128,6 +128,7 @@ SharedWavefunction oepdev(SharedWavefunction ref_wfn, Options& options)
         else 
              throw psi::PSIEXCEPTION("OEPDEV: Invalid OEP build category selected!");
 
+      //oep->localize(); // add later option for it
         oep->compute();
 	oep->print_header();
 	oep->print();

### 2019-11-29 oepdev/libsolver/solver_energy_ct.cc

@@ -1273,6 +1273,10 @@ double ChargeTransferEnergySolver::compute_oep_based_murrell_etal()
   std::shared_ptr<psi::Vector> u_2 = this->compute_u_vector(rmo_2, rmo_1, mol_1);
   std::shared_ptr<psi::Matrix> w_1 = this->compute_w_matrix(mol_1, mol_2, rmo_1);
   std::shared_ptr<psi::Matrix> w_2 = this->compute_w_matrix(mol_2, mol_1, rmo_2);
+  u_1->set_name("Auxiliary calculables: Vector u_1"); u_1->print();
+  w_1->set_name("Auxiliary calculables: Matrix w_1"); w_1->print();
+  u_2->set_name("Auxiliary calculables: Vector u_2"); u_2->print();
+  w_2->set_name("Auxiliary calculables: Matrix w_2"); w_2->print();
 
   // ---> Get distributed effective charges <--- //
   std::vector<std::shared_ptr<psi::Matrix>> q_1 = oep_1->oep("Otto-Ladik.V3.CAMM-nj").dmtp->charges();

### 2019-11-29 include/oepdev_options.h
Davidson-Liu run will now raise error upon lack of convergence unless prevented by setting option

@@ -114,6 +114,7 @@ int read_options(std::string name, Options& options)
         options.add_double ("DAVIDSON_LIU_THRESH_SMALL"   , 1.0E-06              );
         options.add_int    ("DAVIDSON_LIU_SPACE_MAX"      , 200                  );
         options.add_int    ("DAVIDSON_LIU_SPACE_START"    ,-1                    );
+        options.add_bool   ("DAVIDSON_LIU_STOP_WHEN_UNCONVERGED", true           );
 
 
         // ---> CPHF Global Options  <--- //

### 2019-11-29 oepdev/libutil/davidson_liu.cc

@@ -44,6 +44,8 @@ void oepdev::DavidsonLiu::run_davidson_liu() {
   this->davidson_liu_compute_diagonal_hamiltonian();
 
   psi::outfile->Printf("\n @Davidson-Liu: Starting iteration process.\n");
+  bool success = false;
+
   while (conv > eps) {
 
    //this->guess_vectors_davidson_liu_->orthonormalize(); -> not needed since guess vectors are already orthonormal (but maybe useful in the future)
@@ -57,15 +59,17 @@ void oepdev::DavidsonLiu::run_davidson_liu() {
      iter++;
      if (iter > maxit) {
          psi::outfile->Printf(" @Davidson-Liu: Maximum iterations %d exceeded!\n", maxit);
+         success = false;
          break;}
 
      if (conv < eps  ) {
          psi::outfile->Printf(" @Davidson-Liu: Maximum iterations converged successfully!\n");
+         success = true;
          break;}
 
   }
 
-  this->davidson_liu_finalize();
+  this->davidson_liu_finalize(success);
   psi::outfile->Printf("\n @Davidson-Liu: Done.\n");
 }
 
@@ -223,9 +227,13 @@ double oepdev::DavidsonLiu::davidson_liu_compute_convergence()
   return E_old_davidson_liu_->rms();
 }
 
-void oepdev::DavidsonLiu::davidson_liu_finalize()
+void oepdev::DavidsonLiu::davidson_liu_finalize(bool success)
 {
   this->sigma_vectors_davidson_liu_.clear();
   this->guess_vectors_davidson_liu_->reset();
   this->davidson_liu_finalized_ = true;
+
+  if (!success && options_.get_bool("DAVIDSON_LIU_STOP_WHEN_UNCONVERGED")) {
+       throw psi::PSIEXCEPTION(" Davidson-Liu did not converge! Restart with different options.");
+  }
 }

### 2019-11-29 oepdev/libutil/davidson_liu.h

@@ -30,6 +30,7 @@ namespace oepdev{
  *  * `DAVIDSON_LIU_SPACE_MAX`    - Maximum number of guess vectors. Default: 200.     
  *  * `DAVIDSON_LIU_SPACE_START`  - Starting amount of guess vectors. Must be larger or equal to number of roots. 
  *                                  Default: -1, which means that number of roots is taken.
+ *  * `DAVIDSON_LIU_STOP_WHEN_UNCONVERGED` - Raise error when iterations do not converge. Default: True.
  *
  *
  * # Usage in C++ programming
@@ -175,7 +176,7 @@ class DavidsonLiu {
     virtual void davidson_liu_compute_sigma() = 0;
     virtual void davidson_liu_add_guess_vectors();
     virtual double davidson_liu_compute_convergence();
-    virtual void davidson_liu_finalize();
+    virtual void davidson_liu_finalize(bool);
 
 
 };

### 2019-11-29 papers/2_oep_technique.CT/paper-r0.tex
Adding pauli-repulsion paper

--- 2019-07-11 papers/2_oep_technique/paper-r0.tex
@@
 \begin{align}
- P^{AAAB}_{mkij} &\approx X^A_{mk} Y^{AB}_{ij} \;, \\
+ P^{AB}_{ij} &\approx \sum_{k\in A} \sum_{l\in B} X^A_{ik} S_{kl}^{AB} X^B_{lj} \;,\\
+ P^{AAAB}_{mkij} &\approx \sum_{k\in A} \sum_{l\in B} 
+                 X^A_{mkik'} S^{AB}_{k'l'} Y^{B}_{l'j} \;, \\
  P^{AABB}_{ikjl} &\approx Z^A_{ik} Y^{AB}_{} Z^{ B}_{jl} \;,

### 2019-11-29 papers/2_oep_technique.CT/supplementary-r2.tex
SI added to CT paper

--- 2019-11-20 papers/2_oep_technique/paper-r2.tex
@@
 
-\title{Ab Initio Effective One-Electron Potential Operators:
+\title{Supplementary Information\\Ab Initio Effective One-Electron Potential Operators:
 Applications for Charge-Transfer Energy in Effective Fragment Potentials}
@@
 
-\begin{abstract}
-The concept of the effective one-electron potentials has been useful for many decades
-in efficient description of electronic structure of chemical systems, especially extended
-molecular aggregates such as interacting molecules in condensed phases.
-\end{abstract}
-
 \pacs{}
@@
 
-\section{\label{s:1.introduction}Introduction}
-
-Charge transfer (CT) between molecules occurs when the net electronic populations 
-of interacting molecules change which leads to an additional stabilization 
-of a molecular aggregate.\cite{Otto.Ladik.IJQC.1980,Stone.TheTheoryOfIntermolecularForces.1996}
-Although CT processes can be relatively easily scrutinized based
-on the total amount of charge transferred between interacting species,\cite{Otto.Ladik.IJQC.1980}
-evaluation of its contribution to the intermolecular interaction
-potential\cite{Stone.TheTheoryOfIntermolecularForces.1996} is far from trivial due to 
-its very complex quantum mechanical (QM) origins even at the Hartree\hyp{}Fock approximation level,\cite{Otto.Ladik.ChemPhys.1975}
-and cannot be realized in terms of any classical nor semi\hyp{}classical approach.
-In fact, CT does not naturally emerge in the symmetry 
-adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT)
-due to certain technical aspects related with the use of finite basis set expansions
-in modern quantum chemistry calculations and minimizing the basis set superimposition error (BSSE). 
-Recently, Stone and Misquitta showed that CT energy
-can be extracted from SAPT calculations by comparing the induction energies
-of fictional systems in which basis functions are centered either on the monomers only, or the entire 
-interacting complex.\cite{Stone.Misquitta.CPL.2009}
-CT energy was formulated by Murrell et al.~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
-in their perturbation theory in the region of small wavefunction overlap up to second order.
-However, their theory is also computationally too expensive to apply for efficient
-calculation of intermolecular forces in molecular dynamics because
-it involces instantaneous calculations of the electron repulsion integrals (ERI's)
-in molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}
-
-The apparent difficulty in theoretically characterising the CT energy
-in terms of the interacting molecular fragments
-is indeed a challenge in developing modern force fields 
-or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
-for modeling structure
-and dynamics in condensed phases.
-This needs to be contrasted with the
-Coulombic electrostatics,
-non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
-dispersion and induction for instance, 
-which to a certain extent
-can be well described by, respectively, 
-the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
-Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
-the van der Waals repulsive and attractive potentials, 
-as well as the polarizability models, i.e., all being relatively simple and computationally
-inexpensive to evaluate mathematical models. 
-Due to this reason, the
-CT effects 
-%while usually of significant and non\hyp{}negligible magnitude as compared 
-%to the total interaction energy, 
-are not included in any molecular mechanics force field developed up to date
-such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
-The only way to
-incorporate the CT effects in the condensed\hyp{}phase simulations,
-apart from performing full QM molecular dynamics simulations,
-is through the second generation of the effective fragment potential 
-(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
-   Nguyen.Pachter.Day.JCP.2014,
-   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
-developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
-EFP2 is an \emph{ab initio} force field
-derived on the grounds of the first\hyp{}principles perturbation theory
-at the HF level and including the intermolecular dispersion effects
-by the response theory based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
-That is to say, the total intermolecular 
-interaction potential is approximated as 
-%
-\begin{equation}\label{e:u-efp2}
- U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
-\end{equation}
-%
-where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
-distributions of the monomers, approximated by the distributed multipole approximation,
-$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
-principle, that could be roughly understood as the Heitler\hyp{}London interaction
-energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
-are the induction and dispersion energies obtained from the distributed polarizability
-approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.
-
-Despite the considerable success of the EFP2 theory
-in accurately modeling the extended molecular systems 
-like solutions and recently even biomolecules
-with the level of accuracy
-roughly arguably comparable to the M{\o}ller\hyp{}Plesset perturbation theory,
-evaluation of the CT energy in EFP2 model is still too costly for
-typical uses in the
-molecular dynamics simulations. It has been reported that 
-the implementation of the EFP2 CT energy component in GAMESS
-is on average 3-10 times more demanding than the exchange\hyp{}repulsion
-component, becoming a bottleneck of evaluation of total interaction 
-energy and gradients.\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
-Li.Gordon.Jensen.JCP.2006}
-In effect, the CT energy component is usually ignored in most of the applications. In the 
-recent LIBEFP library, CT term were not implemented in the official release 
-of the code.\cite{Kaliman.Slipchenko.JCC.2015}
-
-%However, calculating the CT stabilization energy is usually much more expensive 
-%than evaluating the contributions due to other effects.
-%However, the computational cost of the CT term is usually at least few times larger
-%than the next costly contribution $U^{\rm Ex-Rep}$, which is still too costly
-%to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$.
-
-The effect of CT on the interaction energy 
-is known to be often non\hyp{}negligible, especially
-in donor\hyp{}acceptor systems such as H\hyp{}bonded species
-and charged complexes, although highly depends on the theoretical
-approach being used. 
-The approach adopted by the founders of EFP2 model
-is based on
-%The most common methodology,
-%adopted also by the founders of EFP2 model,
-%is 
-the approximation to SAPT at HF level of theory 
-which is based on the isolated monomeric wave functons
-as references (here abbreviated to CT/HF0).
-In this formulation, the CT energy can be understood
-as a part of the polarization energy that cannot be
-linked to dispersion and pure induction interactions. 
-In this definition of CT, mixing of virtual and occupied MO's
-technically belonging to different unperturbed wavefunction occurs,
-unlike the pure induction and dispersion, in which such mixings are contained
-solely within one monomer.
-%This level of definition of the CT energy will be referred here as to
-%the IPTO-CT energy. 
-%For defining the CT/HF0, 
-%most common reference is the HF approximation
-%most certainly because the HF energy functional is known
-%and its properties are well understood.
-%In the EFP2 model, the total interaction energy is approximated as
-%a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
-%charge densities of the monomers, exchange\hyp{}repulsion energy,
-%induction energy due to mutual polarization, intermolecular dispersion energy,
-%as well as the CT energy.
-%For developing the efficient 
-%\emph{ab initio} fragmentation method 
-%
-% the second generation of the effective fragment potential method (EFP2),
-%the Hartree\hyp{}Fock (HF) model or the density functional theory (DFT) were
-%employed as starting points 
-%Modelling the structure and dynamics of extended molecular aggregates, such as
-%liquids and solutions require
-%EFP2 relies on the intermolecular 
-%perturbation theory, and the total interaction energy is approximated as ...
-%
-%\begin{equation}
-% U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
-%\end{equation}
-%%
-
-One of the main goals of this work, apart from developing
-more efficient model of the CT/HF0 energy, is to
-extend the effective one\hyp{}electron potentials (here referred to as the OEP) 
-technique
-used previously\cite{
-Otto.Ladik.ChemPhys.1975,
-Holas.March.PhysRevA.1991,
-Weber.Thiel.TCA.2000,
-Neese.JCP.2005,
-Cisneros.Andres.Piquemal.Darden.JCP.2005,
-Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
-Li.Gordon.Jensen.JCP.2006,
-Blasiak.Lee.Cho.JCP.2013,
-Blasiak.Maj.Cho.Gora.JCTC.2015}
-to simplify the rigorous and costly quantum chemistry models of extended systems
-with a particular emphasis on solvation phenomena and molecular dynamics.
-The presented OEP technique of removing ERI's from the working
-equations
-follows the notion of the importance of 
-one\hyp{}electron densities in chemistry\cite{Holas.March.PhysRevA.1991},
-thus reducing the complicated summations involving ERI's to much shorter expressions involving
-only one\hyp{}electron integrals.
-Therefore, the OEP computational method is first presented in Section~II.
-Next, in Section III,
-the new technique is used to derive an alternative formulation of the CT/HF0 energy
-compatible with the EFP2 method. Subsequently, after details of computations
-are discussed in Section IV, in Section V the validation of the 
-OEP\hyp{}based CT/HF0 model is presented and its performance in terms 
-of accuracy and computational speed is compared against the EFP2
-model. Finally, few concluding remarks and comments for future work
-are given in Section VI.
-%ERI's, which are oftentimes the main bottleneck
-%in simulating large systems, will be projected out from the working equations
-%to ultimately form the effective fragments in a form of small matrices that are much easier to handle.
-%The proposed methodology aims in removing ERI's from the working equations
-%to ultimately form the effective fragments that are much easier to handle.
-%
-%It is believed that the proposed concept can be helpful in designing future methods tuned 
-%for studying electronically excited states in solutions, better describing molecular
-%dynamics as well as enabling the vibrational analysis of extended systems
-%In this work, we propose an alternative formulation of the CT energy compatible with the
-%EFP2 method. We start from the perturbation theory of Murrell et al.
-%and
-%introduce the effective one\hyp{}electron potential (OEP) operators
-%to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
-%Later we validate the model against the EFP2 method in Section XXX. Finally,
-%we conclude this work and provide few remarks on future development in Section XXX.
-
-\section{\label{s:2.oep}Effective One-Electron Potential Operators}
-
-To establish the notation within the present work, 
-the formalism of one\hyp{}electron potential (OEP) operators is
-given first.
-The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
-produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
-is 
-%
-\begin{equation} \label{e:v-eff}
-	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
-\end{equation}
-%
-where ${\bf r}$ is a spatial coordinate. 
-For convenience, the total effective density can be split into nuclear and electronic contributions,
-%
-\begin{equation} \label{e:rho-eff}
- \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
-\end{equation}
-%
-where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
-Formally, the electronic part can be expanded in terms of an effective
-bond order matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
-%
-\begin{equation} \label{e:d-spectral}
-	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
-	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
-\end{equation}
-%
-Based on that, the operator form of the effective potential 
-can be written as
-%
-\begin{equation} \label{e:oep-operator}
-	\hat{v}^{\rm eff} = 
-        \lambda \hat{v}_{\rm nuc} +
-        %\sum_{\alpha\beta} 
-	%\Ket{\alpha} 
-        \int d{\bf r} \Ket{{\bf r}} 
-        v^{\rm eff}({\bf r})
-	%\left[
-	%\int
-	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
-	%\rho^{\rm eff}({\bf r}')
-	%d{\bf r}' 
-	%\right]
-	%\Bra{\beta}  
-        \Bra{{\bf r}}
-\end{equation}
-%
-with the nuclear
-%and Coulomb operators 
-operator
-defined by
-%
-%\begin{subequations}
-%\begin{align}
-\begin{equation} \label{e:oep-operator-nuc}
-        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
-                     \int d{\bf r} \Ket{{\bf r}} 
-                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
-                     \Bra{{\bf r}} \;.
-	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
-	%f({\bf r}) 
-	%\int
-	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-\end{equation}
-%\end{align}
-%\end{subequations}
-%
-%for any one\hyp{}electron function $f({\bf r})$. 
-and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
-The matrix element of the effective potential operator
-is therefore given by
-%
-\begin{equation} \label{e:oep-matrix-element}
-	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
-	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
-        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
-\end{equation}
-%
-where 
-%
-\begin{equation} \label{e:oep-matrix-element-nuc}
- W_{\alpha\beta}^{(x)} = 
- Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
-\end{equation}
-%
-$Z_x$ is the atomic number of the $x$th atom,
-and the symbol `At' denotes all atoms that contribute to the effective potential.
-In the above equation and throughout the work, 
-%
-\begin{equation} \label{e:1el-operator}
-\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
-\end{equation}
-%
-for any one\hyp{}electron operator $\mathscr{O}(1)$ 
-and the electron repulsion integral (ERI)
-is defined according to
-%
-\begin{equation} \label{e:eri}
-	\BraKet{\alpha\beta}{\gamma\delta} \equiv
-	\iint 
-	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
-	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
-	d{\bf r}_1 d{\bf r}_2  \;.
-\end{equation}
-%
-
-\subsection{\label{ss:2.1.oep-technique}Incorporating Electron Repulsion Integrals into Effective Potentials}
-
-Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
-ERI's. In this work, OEP's are defined by the following transformation
-%
- \begin{equation}
- \mathcal{F}
- \left[ 
-   \BraKet{ij}{k^Al^A}
-	 \right] = \tBraKet{i}{\hat{v}_{kl}^A}{j}  \;,
- \end{equation}
-%
-where 
-% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
-%or basis function.
-$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
-with the
-effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
-The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
-$\hat{v}_{\text{eff}}^A$
-to produce
-%
-\begin{equation}
-	\sum_{ij}\sum_{kl\in A} \mathcal{F}\left[ 
-   \BraKet{ij}{k^Al^A}
- \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j}  \;.
-\end{equation}
-%
-Thus, the total computational effort is, in principle, reduced from the fourth-fold
-sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
-It is also possible to generalize the above expression even further by
-summing over all possible functionals ${\mathcal{F}}_t$
-%
-\begin{equation} \label{e:ft-reduction}
-	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
-   \BraKet{ij}{k^Al^A}
- \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
-\end{equation}
-%
-The above design has the advantage that it opens the possibility to define first\hyp{}principles
-effective fragments as long as 
-the functionals ${\mathcal{F}}_t$ 
-%the $P_{ij}$ and $P_{ijkl}$ 
-%from Eq.~\eqref{e:exp-val-series} 
-are computable and can be approximately
-partitioned in between the interacting fragments.
-%the resulting effective potentials are fully first\hyp{}principles
-%and no extensive case\hyp{}dependent fitting procedures are necessary as long as the $P_{ij}$ and $P_{ijkl}$
-%are computable.
-%Only the effective potentials of \emph{independent} fragment $A$ need to be determined once and for all
-%and stored in a file. 
-%Note also, that, in principle, there is no approximation 
-%made here at that moment.
-
-Two general cases of basis function partitioning scheme
-are considered in this work,
-in which it is possible to define an effective one\hyp{}electron
-potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
-The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
-The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
-%
-{
-\renewcommand{\arraystretch}{1.4}
-\begin{table}[b]
-\caption[Types of matrix elements with OEP operators]
-{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
-}
-\label{t:oep-matrix-element-types}
-\begin{ruledtabular}
-\begin{tabular}{lcccc}
-Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
-                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
-	\cline{1-5}
-Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
-Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
-DF\footnotemark[2]/RI\footnotemark[3] Form    
-&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
-&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
-DMTP\footnotemark[4] Form                     
-&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
-\end{tabular}
-\end{ruledtabular}
-%
-\footnotetext[1]{Test footnotemark.}
-\footnotetext[2]{Density Fitting}
-\footnotetext[3]{Resolution of Identity}
-\footnotetext[4]{Distributed Multipole Expansion}
-%
-\end{table}
-}
-%
-
-\subsection{\label{ss:2.2.oep-DMTP}Distributed Multipole Expansion of OEP's}
-
-This scheme is most applicable in the case of matrix elements of the type
-$
- \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
-$
-because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
-and $\rho^A_{\rm eff}$. 
-In general, given certain two effective one\hyp{}electron density distributions,
-the associated effective Coulombic interaction energy can be estimated from the classical formula
-according to
-%
-\begin{equation} \label{e:coulomb-energy:general}
- \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
- {\vert {\bf r}_1 - {\bf r}_2 \vert} 
-d{\bf r}_1 d{\bf r}_2 \;.
-\end{equation}
-%
-The above integral can be approximated by the
-distributed multipole (DMTP)
-expansions of the respective densities as follows
-%
-\begin{multline} \label{e:coulomb-energy:dmtp}
-  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
- \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
- \sum_{u\in A} \sum_{w\in B} \Big\{ 
- T^{(0)}_{uw}
- q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
- - {\bf T}^{(1)}_{uw} \cdot 
-   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
- - {\bf T}^{(2)}_{uw} : 
-  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
- \ldots
- \Big\}
-\end{multline}
-%
-In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
-are the effective distributed monopole (charge)
-and dipole moment, respectively, centered on the $u$th site, 
-whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
-that can be found elsewhere.
-The symbol `$\odot$' denotes the sum of all the tensor contractions
-performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
-The choice of the distribution centres as well as the truncation order of the multipole expansion
-is crucial in compromising the accuracy and computational cost of the resulting expressions.
-There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
-of Stone\cite{Stone.JCTC.2005}
-the cumulative atomic multipole moments of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983}
-the localised distributed multipole expansion (LMTP) of Etchtebest, Lavery and Pullman\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
-or schemes based on fitting to electrostatic potential such as ChelpG\cite{Breneman.Wiberg.JCC.1990} for instance.
-
-\subsection{\label{ss:2.3.oep-GDF}Density Fitting of OEP's}
-
-This scheme is most useful in the case of matrix elements of the type
-$
- \tBraKet{i^A}{\hat{v}_{\rm eff}^A}{j^B}
-$.
-To get the \emph{ab initio} representation of such an overlap\hyp{}like matrix element,
-one can use a procedure similar to
-the typical density fitting or resolution of identity, both of which are nowadays widely used 
-to compute electron\hyp{}repulsion integrals (ERI's) more efficiently.
-
-\subsubsection{\label{sss:2.3.1.GDF-1}Density Fitting in Nearly-Complete Space}
-
-An arbitrary one\hyp{}electron potential of molecule $A$ acting on any state vector 
-associated with molecule $A$ can be expanded in an auxiliary space centered 
-on $A$ as
-%
-\begin{equation}
-   \hat{v}^{A}\Ket{i} = \sum_{\xi\eta}^{\rm RI} \hat{v}^{A}\Ket{\xi} [{\bf S}^{-1}]_{\xi\eta} \BraKet{\eta}{i}
-\end{equation}
-%
-under the necessary assumption that the auxiliary basis set is nearly complete,
-i.e., 
-$\sum_{\xi\eta}^{\rm RI} \Ket{\xi}[{\bf S}^{-1}]_{\xi\eta} \Bra{\eta} \cong 1$.
-In the equations above, 
-the `RI' symbol denotes a certain auxiliary basis set that fulfills the resolution of identity. 
-%In a special case when the basis set is orthogonal (e.g., molecular orbitals)
-%the above relation simplifies to
-%\begin{equation}
-%   v\vert i) = \sum_{\xi} v\vert \xi) ( \xi \vert i)
-%\end{equation}
-Such a general expansion can be obtained by 
-utilizing the density fitting
-in the nearly\hyp{}complete space,
-%
-\begin{equation} \label{e:v-compl}
- \hat{v}^{A}\Ket{i} = \sum_{\xi}^{\rm RI} V^A_{i\xi} \Ket{\xi} \;.
-\end{equation}
-%
-In the above equation,
-the matrix ${\bf V}^A$
-is the projection of the state vector $\hat{v}^{A} \Ket{i}$
-onto the complete basis $\{ \varphi_\xi \}$.
-Let $Z_i[{\bf V}^A]$ be the least\hyp{}squares objective function 
-%
-\begin{equation} \label{e:z-compl}
- Z_i[{\bf V}^A] = \int \vert \Xi_i({\bf r}) \vert^2 d{\bf r}
-\end{equation}
-%
-with the error density defined by
-%
-\begin{equation} \label{e:error-compl}
- \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm RI} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
-\end{equation}
-%
-By requiring that
-%
-\begin{equation} \label{e:z-necessary-requirement}
- \frac{\partial Z_i[{\bf V}^A]}{\partial V^A_{i\mu}} = 0 \text{ for all $\mu$}
-\end{equation}
-%
-one finds the coefficients of the $i$th row of ${\bf V}^A$ to be
-%
-\begin{equation} \label{e:gdf-compl.v}
-% {\bf v}_i = {\bf a}^{(i)} \cdot {\bf S}^{-1}
-  V^A_{i\xi} = \sum_\eta^{\rm RI} [{\bf S}^{-1}]_{\xi\eta} a^{(i)}_\eta \;,
-\end{equation}
-%
-where the auxiliary matrices are given by
-%
-\begin{subequations}
-\begin{align}
- a^{(i)}_\eta &= \int \varphi^*_\eta({\bf r}) \hat{v}^A \phi_i({\bf r}) d{\bf r} \;,\\  
- S_{\eta\xi}  &= \int \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) d{\bf r} \;.
-\end{align}
-\end{subequations}
-%
-%Since matrix elements of an OEP operator in auxiliary space can be computed 
-%in the same way as the matrix elements with any other basis function, 
-%one can formally write the following identity
-%\begin{equation}
-% (X \vert v\vert i) = \sum_{\xi\eta} S_{X \xi} [{\bf S}^{-1}]_{\xi\eta} (\eta\vert v\vert i)
-%\end{equation}
-%where $ X $ is an arbitrary orbital.
-%When the other orbital
-%does not belong to molecule $A$ but to the (changing) environment, it is 
-%straightforward to compute the resulting matrix element, which is simply given as  
-%\begin{equation}
-%   (j_{\in B} \vert v^A \vert i_{\in A}) = \sum_\xi {S_{j\xi}} {G_{i\xi}}
-%\end{equation}
-%where $j$ denotes the other (environmental) basis function.
-%
-The working formula for $a^{(i)}_\eta$ can be found by applying 
-potential form from Eq.~\eqref{e:v-eff}
-along with the spectral representation of the effective density from Eq.~\eqref{e:d-spectral} 
-which finally gives
-%
-\begin{equation} \label{e:a-coeff-compl}
- a^{(i)}_\eta = \lambda \sum_{x} W_{\eta i}^{(x)} + 
- \sum_{\alpha\beta} D^A_{\alpha\beta} 
-  \BraKet{\alpha\beta}{\eta i} \;.
-%  \iint \frac{ \phi_\alpha^*({\bf r}_1)  \phi_\beta({\bf r}_1) \varphi^*_\eta({\bf r}_2) \phi_i({\bf r}_2)}
-% {\vert {\bf r}_1 - {\bf r}_2 \vert }
-% d{\bf r}_1 d{\bf r}_2
-\end{equation}
-%
-
-\subsubsection{\label{sss:2.3.1.GDF-2}Density Fitting in Incomplete Space}
-
-Density fitting scheme from previous section has practical disadvantage of a nearly\hyp{}complete basis set
-being usually very large (spanned by large amount of basis set vectors). 
-Since most of basis sets used in quantum chemistry do not form a nealy complete
-set, it is beneficial to design a modified scheme in which it is possible to obtain the effective 
-matrix elements of the OEP operator in an incomplete auxiliary space. This can be achieved by minimizing 
-the following objective function
-%
-\begin{equation} \label{e:z-incompl}
-	Z_i[{\bf V}^A] = \iint 
-        \frac{ \Xi_i^*({\bf r}_1) \Xi_i({\bf r}_2) }{\vert {\bf r}_1 - {\bf r}_2 \vert}  
-         d{\bf r}_1 d{\bf r}_2  \;.
-\end{equation}
-%
-with the error density defined by
-%
-\begin{equation} \label{e:error-incompl}
- \Xi_i({\bf r}) = v^A({\bf r}) \phi_i({\bf r}) - \sum_\xi^{\rm DF} V^A_{i\xi} \varphi_\xi({\bf r}) \;.
-\end{equation}
-%
-where the symbol `DF' denotes the generally incomplete auxiliary basis set.
-From the requirement given in Eq.~\eqref{e:z-necessary-requirement}
-one obtains
-%
-\begin{equation} \label{e:v-incompl}
-% {\bf G}^{(i)} = {\bf b}^{(i)} \cdot {\bf A}^{-1}
-  V^A_{i\xi} = \sum_\eta^{\rm DF} [{\bf R}^{-1}]_{\xi\eta} b^{(i)}_\eta \;,
-\end{equation}
-%
-where 
-%
-\begin{subequations}
-\begin{align}
- b^{(i)}_\eta &= \iint 
-                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 
-                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
-                 d{\bf r}_1 d{\bf r}_2 \;, \\
- R_{\eta\xi}  &= \iint 
-                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
-                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
-                 d{\bf r}_1 d{\bf r}_2 \;.
-\end{align}
-\end{subequations}
-%
-%The symbol $ \vert\vert $ is to denote the operator $ r_{12}^{-1}$ and double integration over $ {\bf r}_1 $
-%and $ {\bf r}_2 $. 
-Note that, while $R_{\eta\xi}$ is a typical 2\hyp{}center ERI 
-that can be evaluated by standard means,
-$b^{(i)}_\eta$ matrix elements are not at all trivial to calculate
-because the OEP operator, which contains integration over an electron coordinate,
-is present inside the double integral. Therefore, the following triple integral,
-%
-\begin{equation} \label{e:triple-integral}
- b^{(i)}_\eta = \iiint 
-           \frac{ \varphi^*_\eta({\bf r}_1) \phi_i({\bf r}_2)  \rho^{\rm eff}({\bf r}_3) }
-            {\vert {\bf r}_1 - {\bf r}_2 \vert \vert {\bf r}_3 - {\bf r}_2 \vert}
-           d{\bf r}_1 d{\bf r}_2 d{\bf r}_3 \;,
-\end{equation}
-%
-has to be computed.
-Obtaining all the necessery integrals of this kind directly 
-by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 
-and impractical even for medium sized molecules. 
-However, one can introduce the effective potential in order to eliminate one integration. 
-This can be achieved by performing additional density fitting in a nearly complete intermediate basis 
-%
-\begin{equation}
- \hat{v}^{\rm eff} \Ket{i} = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} \Ket{\varepsilon} \;,
-\end{equation}
-%
-%where the symbol `RI' denotes the intermediate basis from a chosen resolution of identity.
-The working equation is therefore given by
-%
-\begin{equation} \label{e:v-compl-b-coeff}
- b^{(i)}_\eta = \sum_\varepsilon^{\rm RI} H_{i\varepsilon} R_{\varepsilon\eta} \;,
-\end{equation}
-%
-%with
-%%
-%\begin{equation}
-% R_{\varepsilon\eta} \equiv \iint 
-%                       \frac{ \varphi^*_\varepsilon({\bf r}_1) \varphi_\eta({\bf r}_2) } 
-%                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
-%                 d{\bf r}_1 d{\bf r}_2 \;,
-%\end{equation}
-%%
-which can be easily evaluated by noting that
-$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}.
-Thus, the matrix elements of the OEP operator can be found
-by using Eq.~\eqref{e:v-incompl}.
-%Therefore, in order to use this generalized density fitting scheme
-%one must to compute two\hyp{}centre electron repulsion integrals
-%as well as four\hyp{}centre asymmetric electron repulsion integrals of the type $ (\alpha\beta\gamma||\eta) $.
-%The evaluation of such integrals is discussed in Appendix~\ref{a:mcmurchie-davidson}.
-
-It is emphasized here that, as long as the state vector $\Ket{i}$, OEP operator $\hat{v}^A$ 
-and the auxiliary and intermediate basis sets depend solely on one unperturbed molecule $A$, the matrix elements
-$V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.
-
-\section{\label{s:3.ct}Charge Transfer Interaction Energy for Fragment Potentials}
-
-In the CT/HF0 treatments of bi\hyp{}molecular complexes, the CT energy
-can be expressed as a sum of the energy stabilization due to 
-CT from molecule $A$ to $B$ and vice versa, i.e.,
-%
-\begin{equation} \label{e:ct-hf0}
- U^{\rm CT} =  E^{A\rightarrow B} + E^{B\rightarrow A}\;.
-\end{equation}
-%
-
-\subsection{\label{ss:3.1.EFP2} EFP2 Model}
-
-Li, Gordon and Jensen used the expansion of the overlap density 
-in Taylor series and found four different approximate
-theories for the CT energy.\cite{Li.Gordon.Jensen.JCP.2006} 
-The optimal theory, which was shown
-to well reproduce the CT/HF0 energies obtained by using 
-the reduced variational space (RVS)
-analysis of Stevens and Fink\cite{Stevens.Fink.CPL.1987},
-reads
-%
-\begin{equation} \label{e:u-efp2-ct}
- E^{A\rightarrow B} \approx 
- 2\sum_{i\in A}^{\rm Occ}
-  \sum_{n\in B}^{\rm Vir}
-   \frac{ \left| V_{in}^{A\rightarrow B} \right|^2}{\varepsilon_{i} - T_{nn}} \;,
-\end{equation}
-%
-where
-%
-\begin{multline} \label{e:u-efp2-ct-v2}
- \left| V_{in}^{A\rightarrow B} \right|^2 \approx
- \frac{U_{in}}{1-\sum_{m\in A}^{\rm All} S_{mn}^2}
- \Bigg\{
-   U_{in} \\ + \sum_{j\in B}^{\rm Occ} S_{ij} 
-   \left( 
-     T_{nj} - \sum_{m\in A}^{\rm All} S_{nm} T_{mj}
-   \right)
- \Bigg\} \;,
-\end{multline}
-%
-and 
-%
-\begin{equation} \label{e:uin-efp2}
- U_{in} \equiv V_{in}^{{\rm EF},B} - \sum_{m\in A}^{\rm All} V_{im}^{{\rm EF},B} S_{mn} \;,
-\end{equation}
-%
-in which the summations extend over occupied (denoted by `Occ'), virtual (denoted by `Vir') 
-or both (denoted by `All') MO's. The effective potential 
-matrices are defined by
-%
-\begin{equation} \label{e:v-efp2-eff}
- V_{in}^{{\rm EF},B} \equiv 
- \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} \;,
-\end{equation}
-%
-and are evaluated by expanding the $\hat{v}^B_{\rm tot}$ operator
-in distributed multipoles. The resulting calculation requires computation
-of potential, electric field and electric field gradient one\hyp{}electron integrals
-and their contractions with the DMTP distributions.
-The apparent success of the EFP2 scheme is rooted in the dramatic simplifications of the
-\emph{ab initio} expressions for interaction energy, in which
-the relatively very costly ERI's
-have been effectively removed from the working models while maintaining the required accuracy. 
-That is, the canonical orbital energies $\varepsilon_i$ are
-constant parameters, whereas the overlap $S_{nm}$, kinetic energy $T_{mn}$ and effective
-one\hyp{}electron electrostatic potential $V_{in}^{{\rm EF}}$ matrix elements
-are all certain types of the one\hyp{}electron integrals, orders of magnitude cheaper to
-evaluate than ERI's. Unfortunately, due to extensive summations over virtual orbitals
-evaluating Eq.~\eqref{e:u-efp2-ct} is still considerable in cost because typically large basis sets
-need to be used for generating the EFP2 parameters. Moreover, calculation of $V_{in}^{{\rm EF}}$
-is much more expensive as compared to other types of one\hyp{}electron integrals.
-
-In the following subsections, the alternative model of the CT/HF0 energy is proposed
-by introducing OEP's. Although application of the OEP method to the CT/HF0 formulation in EFP2 method
-is probably possible, it would be relatively difficult to discuss the resulting OEP\hyp{}based
-EFP2 models because there are four distinct versions of this theory
-with a set of different approximations, selected based on performance assessment rather than 
-a rigorous derivation manner.
-Instead, perturbation theory of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
-with the
-explicit formulation for closed shell systems by Otto and Ladik\cite{Otto.Ladik.ChemPhys.1975},
-which is somewhat more rigorous than the EFP2 CT model,
-is chosen as a starting point in this work. 
-It is believed that this choice will allow a clear demonstration 
-of the OEP technique in fragment\hyp{}based modelling.
-
-\subsection{\label{ss:3.2.OL} Murrell et al.'s Model: Starting Point}
-
-%According to the second\hyp{}order intermolecular perturbation theory 
-%with wavefunction overlap of Murrell et al.,
-%the energy associated with the charge transfer from molecule $A$ to $B$ is given by
-The CT/HF0 energy can be expressed
-by\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975}
-%
-\begin{equation} \label{e:ct-murell-etal.e}
- E^{A\rightarrow B} = 2 \sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir} 
-  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n}
-\end{equation}
-%
-%coupling constants between occupied and virtual molecular orbitals 
-%are re-expressed by the effective one-electron potentials. The resulting formula provides new possibilities 
-%of optimization and applications in condensed phase simulations. [3]
-%
-where $\varepsilon_i$ and $\varepsilon_n$ are the canonical energies
-of the occupied and virtual
-HF molecular orbitals, respectively,
-whereas the coupling constant is given by
-%
-\begin{multline} \label{e:ct-murell-etal.vin}
- V^{A\rightarrow B}_{in} = 
-        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
-      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
-      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
-     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left(1+\delta_{ik}\right) \BraKet{nj}{ik} \;.
-\end{multline}
-%
-In the above expression, the following effective one\hyp{}electron operators,
-%
-\begin{subequations} 
-\begin{align} \label{e:ct-murell-etal.vtot-vi}
- \hat{v}^B_{\rm tot} &= \hat{v}^B_{\rm nuc} + 2\sum_{j\in B}^{\rm Occ} \hat{v}^B_{jj} \quad\text{ and }\\ 
- \hat{v}^A_{i      } &= \hat{v}^A_{\rm tot} - 2\hat{v}^A_{ii} \;,
-\end{align}
-\end{subequations}
-%
-were introduced without making any approximation to the original equation
-from Ref. 
-The goal of this work is to apply the technique sketched in Eq.XXX
-to eliminate ERI's from Eq.~\eqref{e:ct-murell-etal.vin} in terms of OEP's.
-
-\subsection{\label{ss:3.3.OEP} Murrell et al.'s Model: Application of OEP Technique}
-
-One can immediately notice that the five summation terms
-from Eq.~\eqref{e:ct-murell-etal.vin} can be classified into three groups
-regarding the type of ERI's that are required:
-(i) $\BraKet{AB}{BB}$ -- the first two terms;
-(ii) $\BraKet{AA}{BB}$ -- the third term and
-(iii) $\BraKet{BB}{AA}$ -- the two last terms. 
-Note also that there are no exchange\hyp{}like terms needed in this case.
-Therefore, all the contributions can be re\hyp{}cast in terms of the OEP's.
-%Note that groups (ii) and (iii) correspond to the Coulomb interactions between
-%The group (i) corresponds to the overlap\hyp{}like integrals, whereas the groups (ii) and (iii)
-%correspond to the Coulomb\hyp{}like integrals
-
-\paragraph{Group (i).}
-Group (i) can be rewritten by noticing that
-%
-\begin{equation} \label{e:ct-murell-etal.group-i.notice}
- \sum_{j\in B}^{\rm Occ} 
- \BraKet{nj}{ij} =
- \Bra{i} \left[ \hat{v}^B_{nj} \Ket{j} \right]  \;,
-\end{equation}
-%
-which, by combining with the first term from Eq.~\eqref{e:ct-murell-etal.vin}, 
-allows to apply the density fitting in an
-incomplete basis set as follows
-%
-\begin{equation} \label{e:v-oep.ct}
-\Bra{i} \left[ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
-\cong \Bra{i} \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta}
-\end{equation}
-%
-with the error density defined by
-%
-\begin{multline}
- \Xi_n({\bf r}) = 
-  \left\{
-      v_{nuc}^B({\bf r}) + 2 \sum_{j\in B}^{\rm Occ}
-   \int \frac{ \vert \phi_j({\bf r}') \vert^2 }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-  \right\}
-  \phi_n({{\bf r}})  \\
- - \sum_{j\in B}^{\rm Occ} \phi_j({{\bf r}})
-   \int \frac{ \phi^*_n({\bf r}') \phi_j({{\bf r}'}) }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
- - \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \varphi_\eta({\bf r}) \;.
-\end{multline}
-%
-Substituting the above equation into Eq.~\eqref{e:z-incompl}
-%and then
-%using resolution of identity to simplify the 
-leads to
-%
-\begin{equation} \label{e:v-oep-group-1-prefinal}
- V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} \left[ {\bf R}^{-1} \right]_{\xi\eta} b_\eta^{(n)} \;,
-\end{equation}
-%
-where
-%
-\begin{multline}
- b^{(n)}_\eta = \iint 
-           d{\bf r}_1 d{\bf r}_2  
-           \frac{ \varphi^*_\eta({\bf r}_1) }
-            {\vert {\bf r}_1 - {\bf r}_2 \vert } \\ \times
-          \left[ 
-           \hat{v}^B_{\rm tot}  \phi_n({\bf r}_2) 
-         - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \phi_j({{\bf r}_2})
-           \right]  \;.
-\end{multline}
-%
-The above result is given by a sum of triple integrals from Eq.~\eqref{e:triple-integral}.
-However, the following application of the resolution of identity,
-%
-\begin{equation}
- \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
- \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon} \;,
-\end{equation}
-%
-leads to a much simpler formula
-that requires only the one\hyp{} and two\hyp{}electron integrals. 
-Therefore, by combining
-equations~\eqref{e:v-oep-group-1-prefinal}, \eqref{e:v-compl-b-coeff} and \eqref{e:gdf-compl.v}
-one arrives to
-%
-\begin{equation} \label{e:v-oep-group-1-final}
- V^B_{n\xi} = \sum_{\eta\in B}^{\rm DF} 
-          \sum_{\varepsilon\in B}^{\rm RI}
-         \left[ {\bf R}^{-1} \right]_{\xi\eta}
-         R_{\eta \varepsilon} 
-         H_{n\varepsilon}^B \;,
-\end{equation}
-%
-where
-%
-\begin{equation} \label{e:v-oep-group-1-final:H-matrix}
- H_{n\varepsilon}^B = \sum_{\zeta\in B}^{\rm RI} \left[ {\bf S}^{-1} \right]_{\varepsilon\zeta}
-   a_\zeta^{(n)}
-\end{equation}
-%
-and
-%%
-\begin{align} \label{e:v-oep-group-1-final:a-vector}
- a^{(n)}_\zeta &= \tBraKet{\zeta}{\hat{v}^B_{\rm tot}}{n}
-      - \sum_{j\in B}^{\rm Occ} \tBraKet{\zeta}{\hat{v}^B_{nj}}{j} \nonumber \\
- &= \sum_{y\in B} W^{(y)}_{\zeta n} 
-  + \sum_{j\in B}^{\rm Occ} 
-  \left\{
-   2\BraKet{\zeta n}{jj} - \BraKet{\zeta j}{nj} 
-  \right\} \;.
-\end{align}
-%
-Note that all the calculations that are required to obtain $V^B_{n\xi}$ are performed
-solely on the densities and basis sets associated with the unperturbed molecule $B$.
-Therefore, $V^B_{n\xi}$ can be considered as effective fragment parameters
-used to compute the first two terms of Eq.~\eqref{e:ct-murell-etal.vin} by
-%
-\begin{equation} \label{e:ct.group-i.final}
-        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
-      - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
-       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;,
-\end{equation}
-%
-which is a great simplification over the original form of group (i)
-because only the overlap integrals between the $i$th MO on molecule $A$
-and $\eta$th auxiliary orbital on molecule $B$ need to be evaluated on the fly.
-Note that the only approximation made so far was the application of density fitting
-and resolution of identity. If the auxiliary and intermediate
-basis sets are sufficiently large, the errors
-due to this approximation can be minimal and negligible.
-
-\paragraph{Group (ii).}
-The term belonging to this group can be considered as a sum of interaction
-energies between the total charge density distribution of molecule $B$
-and the partial density $\rho_{ik}({\bf r})$ of molecule $A$,
-weighted by the overlap integrals $S_{nk}$. Using the distributed multipole 
-expansion based on the charge centroids of the localized molecular orbitals (LMO's),
-${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$ with $\chi_i({\bf r})$ being localized,
-this group can be written as
-%
-\begin{equation} \label{e:ct.group-2}
-      - \sum_{k\in A}^{\rm Occ} S_{nk}  \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
-\end{equation}
-%
-Note that, in general, $\rho_{ki}^A$ is complex whereas $\rho^B_{\rm tot}$ is always real.
-Since the interaction energy must also be real, it must hold that
-%
-\begin{equation} \label{e:ct.group-2-d}
- \rho_{ki}^A \odot \rho^B_{\rm tot} = \delta_{ik} \rho_{ii}^A \odot \rho^B_{\rm tot} \;.
-\end{equation}
-%
-Consider now a situation in which the localized occupied MO's are used.
-In this case, the distributed multipole expansion
-of the above becomes
-%
-\begin{multline}
- \rho_{ii}^A \odot \rho^B_{\rm tot} 
- \approx 
- q_{i} 
- \left[
-  %\sum_{i\in A}^{\rm LC} 
- \sum_{y\in B}^{\rm At}
-  \frac{Z_y}{\vert {\bf r}_y - {\bf r}_i \vert } 
- +
- 2\sum_{j\in B}^{\rm LC}
-  \frac{q_j}{\vert {\bf r}_j - {\bf r}_i \vert } 
- \right] \\
- + \text{ quadrupole terms } \;,
-\end{multline}
-%
-because the distributed charges
-$q_i = -1$ whereas the distributed dipole moments %${\BM{\upmu}_i}$ 
-centered at their
-respective LMO charge centroids vanish.\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
-This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows:
-%
-\begin{equation} \label{e:ct.group-ii.final}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \approx S_{ni}  \left[
-  %\sum_{i\in A}^{\rm LC} 
- \sum_{y\in B}^{\rm At}
-  \frac{Z_y}{ r_{yi} } 
- -
- \sum_{j\in B}^{\rm LC}
-  \frac{2}{r_{ji}} 
- \right] \;.
-\end{equation}
-%
-Therefore, only overlap integrals and relative distances between
-atomic and LC's are needed, which leads to a great reduction of the
-calculation cost,
-as compared either to the original expression or to the multipole expansion (left\hyp{} and right\hyp{}hand sides
-of Eq.~\eqref{e:ct.group-2}, respectively).
-We shall refer to this approximation as to
-the spherical overlap approximation (SOA). Note that in order for this approximation to be valid,
-occupied molecular orbitals need to be localized.
-
-\paragraph{Group (iii).}
-The terms with the overlap integrals involving the occupied MO on $A$
-can be combined into a single summation term, i.e.,
-\begin{multline} \label{e:ct.group-iii}
-       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
-     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right)
-        \BraKet{nj}{ik} \\ = -
- \sum_{k\in A}^{\rm Occ} 
- \sum_{j\in B}^{\rm Occ}
- S_{kj} 
- \tBraKet{j}{
- \underbrace{
- \left\{ 
-  \delta_{ik} \left( \hat{v}^A_{k} + \hat{v}^A_{ik} \right)   - \hat{v}^A_{ik}
- \right\} 
-  }_{\hat{v}^{A,{\rm eff}}_{ik}}
- }{n}  \\ \cong -
-%
-  \sum_{k\in A}^{\rm Occ} 
- \sum_{j\in B}^{\rm Occ}
- S_{kj} 
- \rho_{nj}^B \odot \rho^{A,{\rm eff}}_{ik} \;,
-\end{multline}
-%
-where the effective potential ${v}^{A,{\rm eff}}_{ik}$ (with the associated 
-effective density $\rho^{A,{\rm eff}}_{ik}$) 
-is defined by
-%
-\begin{equation} \label{e:ct.group-iii.oep}
- {v}^{A,{\rm eff}}_{ik}({\bf r}) \equiv
- \delta_{ik} 
- \left[
-  v^A_{\rm tot} ({\bf r}) - 2 v^A_{kk} ({\bf r}) + v^A_{ik} ({\bf r})
- \right] 
-  - v^A_{ik} ({\bf r}) \;.
-\end{equation}
-%
-Now, it is first assumed here that $\rho^{B}_{nj} ({\bf r}) \approx 0$ if occupied MO's on $B$ are localized.
-This means that
-%
-\begin{multline} \label{e:ct.group-iii.final.a}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
-     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right)
-        \BraKet{nj}{ik} \\ 
-    \approx 0 \;.
-\end{multline}
-%
-The above result is a crude approximation, since the channels $j\rightarrow n$
-are quite important contributions in most of perturbation theory treatments.
-%virtual MO's can penetrate the occupied MO's to some extent.
-In order to include the $\rho^{B}_{nj}$ density, it is approximately represented here by a set of effective 
-cumulative
-atomic charges $\{ q^{B,(nj)}_{y} \}$ associated with the effective one\hyp{}particle density matrix
-%
-\begin{equation} \label{e:oed-group-iii}
- D^{B,(nj)}_{\beta\delta} = C_{\beta n} C_{\delta j} \;.
-\end{equation}
-%
-In this work, the effective charges were computed via the Mulliken method as
-discussed in Appendix~\ref{e:a1-v.oep-camm} with $\lambda=0$. 
-Note now that both effective densities in 
-$\rho^B_{jn} \odot \rho^{A,{\rm eff}}_{ik}$ are generally complex and one cannot apply the SOA
-for $\rho^{A,{\rm eff}}_{ik}$. In the LMO picture however, it can be shown that
-$\rho_{ik} ({\bf r}) \ll \rho_{ii} ({\bf r})$ 
-for $i\ne k$ in most locations. Therefore,
-the SOA is still applicable in the LMO picture
-and the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
-simplifies to
-%
-\begin{equation} \label{e:ct.group-iii.oep.simplified}
- {v}^{A,{\rm eff}}_{ik}({\bf r}) \approx
- \delta_{ik} \left\{
- {v}^A_{\rm tot} ({\bf r})- 2{v}^A_{kk} ({\bf r}) \right\} \;,
-\end{equation}
-%
-which
-leads to 
-%
-\begin{multline} \label{e:ct.group-iii.final.b}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
-     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj}
-        \BraKet{nj}{ik}  \\
- \approx  -
- \sum_{j\in B}^{\rm Occ} S_{ij}
- \sum_{y\in B}^{\rm At} 
- q^{B,(nj)}_{y} 
- \left[ 
-   \sum_{x\in A}^{\rm At}
-   \frac{Z_x}{r_{xy}}
-  + \frac{2}{r_{iy}}
-  - \sum_{k\in A}^{\rm Occ}
-    \frac{2}{r_{ky}} 
- \right]
- \;.
-\end{multline}
-%
-
-\paragraph{Final OEP-based forms of the coupling constant.}
-Gathering the results from Eqs. the coupling constant
-can be given as follows
-%
-\begin{equation} \label{e:ct-murell-etal.vin.oep}
-V_{in}^{A\rightarrow B} 
-       \approx 
- G_{1;in}^{A\rightarrow B} 
-+G_{2;in}^{A\rightarrow B} 
-+G_{3;in}^{A\rightarrow B} \;,
-\end{equation}
-%
-where
-%
-\begin{subequations} \label{e:ct-murrell-etal.vin.oep-gterms}
-\begin{align}
- G_{1;in}^{A\rightarrow B} &\equiv \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;, \\
- G_{2;in}^{A\rightarrow B} &\equiv S_{ni} u_i^{BA} \;, \\
- G_{3;in}^{A\rightarrow B} &\equiv
- -\sum_{j   \in B}^{\rm Occ} S_{ij}
-  \sum_{y   \in B}^{\rm At} q_y^{B,(nj)} w_{yi}^{BA}
-\;.
-\end{align}
-\end{subequations}
-%
-In the above, the auxiliary variables are
-%
-\begin{subequations}
-\begin{align} \label{e:ct-murell-etal.vin.oep.aux-var}
- u_i^{BA} &\equiv    
- \sum_{y\in B}^{\rm At}
-  \frac{Z_y}{ r_{yi} } 
- -
- \sum_{j\in B}^{\rm LC}
-  \frac{2}{r_{ji}} 
-                \;, \\
- w_{yi}^{BA} &\equiv 
-   \sum_{x\in A}^{\rm At}
-   \frac{Z_x}{r_{xy}}
-  + \frac{2}{r_{iy}}
-  - \sum_{k\in A}^{\rm Occ}
-    \frac{2}{r_{ky}} 
-                \;.
-\end{align}
-\end{subequations}
-%
-To compute the interaction energy due to CT from $A$ to $B$ by using the above SOA\hyp{}based
-approximations, coupling elements need to be transformed from localized to canonical MO basis, i.e.,
-%
-\begin{equation} \label{e:ect-oep.prefinal}
- E^{A\rightarrow B} \approx 
- 2 
- \sum_{i\in A}^{\rm CMO}
- \sum_{n\in B}^{\rm Vir}
- \frac{
- \lvert
-   \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
-   V_{i'n}^{A\rightarrow B}
- \rvert^2 }{\varepsilon_i - \varepsilon_n} \;,
-\end{equation}
-%
-where ${\bf L}^A$ is the CMO\hyp{}LMO transformation matrix for occupied orbitals of $A$. 
-Note however, that the effective potential parameters from the density fitted term of group (i) 
-do not involve any occupied orbitals.
-Therefore, to save computational costs, only groups (ii) and (iii) need to be transformed
-whereas for group (i) the overlap integrals from Eq.~\eqref{e:ct.group-i.final} 
-can be already evaluated in CMO basis.
-Thus, the final working formula for the interaction energy due to CT from $A$ to $B$ reads
-%
-\begin{multline} \label{e:ect-oep.final}
- E^{A\rightarrow B} \approx
- 2 
- \sum_{i\in A}^{\rm CMO}
- \sum_{n\in B}^{\rm Vir}
- \frac{1}{\varepsilon_i - \varepsilon_n} \times 
- \Bigg(
-   G_{1,in}^{A\rightarrow B} \\
-  +
-   \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
-   \left\{
-   G_{2,i'n}^{A\rightarrow B}
-  +G_{3,i'n}^{A\rightarrow B}
-  \right\}
- \Bigg)^2 \;,
-\end{multline}
-%
-where the subscripts $G_n$ ($n$ = 1, 2, 3) denote a particular group of 
-terms from Eqs.~\eqref{e:ct.group-i.final}, \eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.a}
-or \eqref{e:ct.group-iii.final.b}. 
-The total CT energy is given by the sum of the above contribution and the twin contribution
-due to CT from molecule $B$ to $A$ according to Eq.~\eqref{e:ct-hf0}.
+\section{Asymptotic dependence of CT interactions for methanol-water complex}
 
 %
-%he above formula does not explicitly define an OEP due to the summation over different
-%state vectors. 
-%However, it can be achieved
-%by transforming from MO to AO basis, i.e.,
-%%
-%\begin{equation}
-% \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
-% = \sum_{\beta\in B}^{\rm AO} 
-% \left\{ 
-%  C_{\beta n} \hat{v}^B_{\rm tot} - \sum_{j\in B}^{\rm Occ} C_{\beta j} \hat{v}^B_{nj} 
-% \right\} \Ket{\beta} \;,
-%\end{equation}
-%%
-%which leads to the OEP in spatial representation
-%%
-%\begin{multline}
-% v_{\beta n}({\bf r}) \equiv C_{\beta n} v_{nuc}^B({\bf r}) \\ + 
-%  \sum_{j\in B}^{\rm Occ} \Bigg\{ 
-%   2 C_{\beta n} \int \frac{ \vert \phi_j({\bf r}') \vert^2 }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-% %\\
-% - C_{\beta j} \int \frac{ \phi^*_n({\bf r}') \phi_j({\bf r}') }{\vert {\bf r}' - {\bf r} \vert} d{\bf r}'
-%  \Bigg\}  
-%%
-%\end{multline}
-%
-
-\section{\label{s:4.calculations}Calculation Details}
-
-Four selected molecular pairs composed of H$_2$O, CH$_3$OH, NH$_4^+$ and NO$_3^-$
-(Figure 1) molecules or ions were chosen to validate the OEP\hyp{}based model of the CT energy
-at the PT/HF formulation level. Structures of the molecular dimers were first 
-energy\hyp{}optimized at the HF/6-31+G(d,p) level by using
-the Gaussian 16 quantum chemistry program package.\cite{Gaussian16}
-To obtain the PT/HF benchmark CT energy, it was estimated
-from the following approximation:
-%
-\begin{equation}\label{e:ct-pthf-ref}
- \Delta E^{\rm CT}_{\rm PT/HF} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
- - \Delta E^{\rm Ind} \;,
-\end{equation}
-%
-where $\Delta E^{\rm Pol}_{\rm DDS/HF}$ is the Hartree\hyp{}Fock polarization energy
-defined
-according to the density decomposition scheme (here referred as to the DDS) 
-proposed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
-whereas $\Delta E^{\rm Ind}$ term is the 
-approximate pure induction energy due to the polarizability effects.\cite{Stone.TheTheoryOfIntermolecularForces.1996}
-The DDS polarization energy encompasses all the charge delocalization
-effects that are not associated with the Pauli exclusion principle
-in the first order, and is computed by
-%
-\begin{multline}\label{e:de-pol-dds}
-\Delta E^{\rm Pol}_{\rm DDS/HF} \equiv \Delta E^{\rm HF} \\
- - \left(
-   \Delta E^{\rm Coul}_{\rm DDS/HF}
- + \Delta E^{\rm Exch}_{\rm DDS/HF}
- + \Delta E^{\rm Rep}_{\rm DDS/HF} 
-   \right)\;.
-\end{multline}
-%
-In the above, $\Delta E^{\rm Coul}_{\rm DDS/HF}$ is the pure Coulombic interaction energy
-between the unperturbed (non\hyp{}interacting) charge densities of the monomers in the complex,
-and
-$\Delta E^{\rm Exch}_{\rm DDS/HF}$ and $\Delta E^{\rm Rep}_{\rm DDS/HF}$
-are the exchange and the Pauli\hyp{}repulsion interaction energies, respectively. 
-At the HF level of theory, the sum of the latter two terms is in fact
-equivalent to the first\hyp{}order exchange\hyp{}repulsion energy 
-in the intermolecular perturbation theory with exchange of Hayes and Stone.\cite{Hayes.Stone.MolPhys.1984} 
-It was also shown by 
-Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011}
-that the Coulombic and exchange\hyp{}repulsion term
-in the DDS method reproduce the first\hyp{}order Coulombic and exchange\hyp{}repulsion
-interaction energy in the symmetry adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994}
-(SAPT)
-with accuracy below 0.3 kcal/mole.
-Therefore, Eq.~\eqref{e:ct-pthf-ref} is believed to be sufficient
-in order to compute the PT/HF CT energy, because there is no dispersion
-contribution to the DDS polarization energy at the HF level.
-
-The pure induction energy was estimated by using the distributed
-dipole\hyp{}dipole polarizability approximation according to the computational
-methodology of Li et al.,\cite{Li.Netzloff.Gordon.JCP.2006}
-in which
-%
-\begin{equation}\label{e:ind-ref}
- \Delta E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\bf D}^{-1} \cdot {\bf F} \;.
-\end{equation}
-%
-In the above equation, ${\bf F} = \{ \ldots, {\bf F}({\bf r}_a), \ldots\}$
-is a super\hyp{}vector of electric fields at the $a$th distributed site
-due to the surrounding unperturbed molecular charge densities in the complex
-(i.e., the charge\hyp{}density of the molecule that contains $a$th center
-is not included in ${\bf F}({\bf r}_a)$).
-%The electric field components are
-%aligned in a super\hyp{}vector 
-%according to each polarizable center, i.e. ${\bf F} = \{ \ldots, {\bf F}({\bf r}_a), \ldots\}$,
-%where $a$ labels the LMO centroids.
-The ${\bf D}$ matrix is defined as
-%
-\begin{equation} \label{e:ind-ref.d-matrix}
- {\bf D}_{ab} = 
-\left\{\begin{matrix}
-{\BM\upalpha}_a^{-1} &\text{if $a=b$} \\ 
-0                    &\text{if $a\ne b \land a,b \in$ same molecule}\\ 
--{\bf T}_{ab}        &\text{if $a\ne b \land a,b \notin$ same molecule}
-\end{matrix}\right.
-\end{equation}
-%
-and ${\bf T}_{ab}$ is the dipole\hyp{}dipole interaction tensor
-given by
-%
-\begin{equation} \label{e:ind-ref.t-matrix}
- {\bf T}_{ab} \equiv \frac{3}{r_{ab}^5} {\bf r}_{ab} \otimes {\bf r}_{ab}
-  - \frac{1}{r_{ab}^3} {\bf 1} \;,
-\end{equation}
-%
-with ${\bf r}_{ab} \equiv {\bf r}_{a} - {\bf r}_{b}$ being the
-relative distance between the LMO centroids.
-The electric fields at the LMO centroids were computed from
-the unperturbed HF charge\hyp{}density distributions
-%
-\begin{equation} \label{e:ind-ref.efield}
- {\bf F}({\bf r}_a) = \sum_x^{\rm At} \frac{Z_x}{r_{ax}^3} {\bf r}_{ax} 
-  + 2 \sum_{\alpha\gamma} D_{\alpha\gamma} 
-%    \tBraKet{\gamma}{\frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3} }{\alpha}_{\bf r} \;.
-  \left< \gamma \left| 
-    \frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3}
-               \right| \alpha \right>_{\bf r} \;.
-\end{equation}
-%
-The distributed anisotropic dipole\hyp{}dipole
-polarizabilities ${\BM{\upalpha}}_a$ centered at the LMO centroids
-were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
-method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
-Orbital localization was performed by using the Boys method.\cite{Boys.RevModPhys.1960}
-
-All the models that were used to test the theory presented in this work
-were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-
-\section{\label{s:5.results}Results and Discussion}
-
-\subsection{\label{ss:5.1.accuracy}Accuracy of OEP-based model}
-
-First, the asymptotic dependence of the CT energy
-was investigated for four chosen molecular pairs and the results are
-shown in Figure~\ref{f:fig-2} and Figure~S1 for the methanol-water system. 
-Overall, OEP technique of effective
-elimination of ERI's sketched in Eq.~\eqref{e:ft-reduction} 
-works
-quantitatively well for water dimer and ammonium nitrate (Figures~\ref{f:fig-2}(a) and (c)),
-and methanol-water (Figure~S1)
-reproducing the reference Otto\hyp{}Ladik CT energy,
-whereas in the case of water\hyp{}ammonium cation complex the OEP
-method overestimates the CT energy by around 50\%, even for large
-intermonomer separations. Careful inspection
-(see SI)
-revealed that the reason is the SOA in $G_2$ and $G_3$ and not
-the generalized density fitting in $G_1$ terms, in the case of which
-the aug-cc-pVDZ-jkfit basis set was sufficient as auxiliary space of OEP transformation.
-On the other hand, analysis of the statistical sets of dimeric complexes
-shown in Figure~XXX
-confirmed that in the great majority of cases the 
-OEP technique gives reasonably accurate results with respect
-to the parent model with relative errors less than 10\%,
-unless systems are charged.
-%
-{
-\renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
-\caption[Accuracy of approximate CT/HF0 energy methods across wide range of bi\hyp{}molecular complexes]
-{{\bf Accuracy of approximate CT/HF0 energy methods across wide range of bi\hyp{}molecular complexes\footnotemark[1]}
-}
-\label{t:2}
-\begin{ruledtabular}
-\begin{tabular}{lclcccccccc}
-         &&       && \multicolumn{7}{c}{RMSE (kcal/mol) } \\
-\cline{4-11}
-\multicolumn{3}{c}{Database}
-   && OL\footnotemark[2]  
-                  && \multicolumn{3}{c}{OEP\footnotemark[3]}  
-                  && EFP2\footnotemark[4] \\
-\cline{1-3}
-\cline{4-5}
-\cline{6-9}
-\cline{10-11}
-NCB31\footnotemark[5] 
-      &&  HB6/04  && 1.38 && 1.36 && 0.37 && 2.45 \\ % 
-      &&  DI6/04  && 0.35 && 0.66 && 0.22 && 0.44 \\ % 
-      &&  CT7/04  && 0.98 && 2.82 && 0.71 && 4.06 \\ % 
-      &&  WI7/04  && 0.01 && 0.09 && 0.05 && 0.02 \\ % 
-      &&  PPS5/05 && 0.07 && 1.31 && 0.79 && 0.60 \\ % 
-      &&  Total   && 0.83 && 1.69 && 0.53 && 2.39 \\ % 
-\cline{1-3}
-\cline{4-5}
-\cline{6-9}
-\cline{10-11}
-
-BBI\footnotemark[6]   
-      &&  Subset\footnotemark[7]
-                  && 0.26 && 0.56 && 0.13* && 0.12 \\ %
-\end{tabular}
-\end{ruledtabular}
-%
-\footnotetext[1]{Validated against Eq.~\eqref{e:ct-pthf-ref}.}
-\footnotetext[2]{Otto and Ladik's expression from Eq.~(4) in Ref.~\cite{Otto.Ladik.ChemPhys.1975}}
-\footnotetext[3]{This work, Eq.~\eqref{e:ect-oep.final-scaled} 
-with $C=1.00$ (left column) and $C=1.56$ (right columm).
-GDF(1)/aug-cc-pVDZ-jkfit scheme was used for the extended density fitting of group (i) OEP's.}
-\footnotetext[4]{Ref.~\cite{Li.Gordon.Jensen.JCP.2006}, Eq.~\eqref{e:u-efp2-ct}.}
-\footnotetext[5]{Ref.~\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Schultz.Truhlar.JCTC.2006}}
-\footnotetext[6]{Ref.~\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}}
-\footnotetext[7]{100 backbone\hyp{}sidechains complexes were randomly selected.}
-%
-\end{table}
-}
-%
-
-In the case of water dimer and methanol-water complex, EFP2 and PT/HF estimations
-coincide with the OEP model. However, in the presence of ions, 
-the deviation of the OEP model from the PT/HF energy 
-become significant, in contrast to the EFP2 model, which gives very good estimates
-of the PT/HF energy in mostly all cases.
-Naturally, the accuracy of the Otto\hyp{}Ladik model limits the absolute accuracy
-of the OEP model, which is manifested in the ammonium nitrate system
-in Figure~\ref{f:fig-2}(c) or the problems with SOA. 
-In order to improve the performance for ionic systems,
-further studies are necessary, that will modify the original formulation
-of Otto and Ladik and SOA treatment. Nevertheless, except from the ionic systems,
-it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable
-(Figure~XX) across a wide range of various interacting complexes including 
-systems with H-boding, dominated by dipole-dipole or dispersion interactions or 
-exhibiting substantial CT.
-%
 \begin{figure}[h]
-\includegraphics[width=0.46\textwidth]{fig-2.pdf}
-\caption{\label{f:fig-2} {\bf Asymptotic dependence of the charge transfer energy
-in the CT/HF0 formulation for selected bi\hyp{}molecular complexes.} 
-The charge transfer energy estimates are plotted for 
-(a) water dimer, 
-(b) water\hyp{}ammonium complex, and 
-(c) ammonium\hyp{}nitrate complex,
-where one molecule has been translated by $R$ from the starting geometry
-optimized at HF/6-31+G(d,p) level,
-along the vector specified in the insets.
+\includegraphics[width=0.5\textwidth]{fig-s1.pdf}
+\caption{\label{f:fig-s1} {\bf Asymptotic dependence of the charge transfer energy
+in the CT/HF0 formulation for water-methanol complex.} 
+Water molecule has been translated by $R$ from the starting geometry
+along the vector specified in the inset picture.
 The counterpoise\hyp{}corrected total interaction energy
 is also shown for comparison in light grey color in this figure.
-Interaction energies were obtained at the HF/6-311++G(2df,2pd) level of theory
-with the single GDF scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
+All data were obtained at HF/6-311++G(2df,2pd) level of theory
+with the single GDF scheme and aug-cc-pVTZ-jkfit auxiliary basis set.
 } 
@@
 
-%
-\begin{figure*}[h]
-\includegraphics[width=0.9\textwidth]{fig-3.pdf}
-\caption{\label{f:fig-3} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
-in small bi\hyp{}molecular neutral complexes.}
-(a) accuracy of the ERI elimination technique; 
-(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
-CT/HF0 estimate. 
-The structures of bi\hyp{}molecular complexes
-were taken from the NCB31 database developed by the Truhlar's group:
-the HB6/04 hydrogen bonding database 
-(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-the DI6/04 dipoleinteraction database 
-(green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-the CT7/04 charge-transfer complex database 
-(blue),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-the WI7/05 weak interaction database 
-(yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-and PPS5/05 the $\pi$-$\pi$ stacking database 
-(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-6-311++G** basis set was used for the CT energy calculations
-and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
-for the OEP calculations.
-} 
-\end{figure*}
-%
-
-%
-\begin{figure*}[h]
-\includegraphics[width=0.9\textwidth]{fig-4.pdf}
-\caption{\label{f:fig-4} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
-in aminoacid pairs.}
-(a) accuracy of the ERI elimination technique; 
-(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
-CT/HF0 estimate. 
-The structures of 100 aminoacid pairs
-were randomly selected from the BBI Bioo database
-for backbone\hyp{}sidechain interactions.\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}
-6-31+G(d,p) basis set was used for the CT energy calculations
-and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
-for the OEP calculations.
-} 
-\end{figure*}
-%
-
-\subsection{\label{ss:5.2.cost}Reduction of computational costs}
-
-The utmost goal of this work is to reduce the computational cost of
-the CT PT/HF energy evaluation in the calculations involving effective fragment
-potentials. The EFP2 CT expression has the computational cost
-roughly proportional to $(10on^2 + 2n^3 + o^2n^2)v$, where the
-$o$ and $n$ denote the number of occupied orbitals and the number of atomic basis functions,
-respectively, whereas $v$ is the relative cost. On the other hand, OEP\hyp{}based expression
-from Eq.~\eqref{e:ect-oep.final} has the cost of approximately $on(Q+N+o)$, where 
-$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
-respectively. 
-%
-{
-\renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
-\caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
-{{\bf}Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.\footnotemark[1]
-}
-\label{t:3}
-\begin{ruledtabular}
-\begin{tabular}{lldcd}
- AO's &  & \multicolumn{3}{c}{Exponents (a.u.)}\\
-      &  & \text{Methanol}         && \text{Water}           \\
-\cline{1-2}
-\cline{3-5}
- H    &1s&   7.7369294534   &&   29.5837988322 \\
- C    &1s& 880.3654511045   &&                 \\
-      &2s& 123.8276940826   &&                 \\
-      &2p&  11.6748155072   &&                 \\
- O    &1s& 902.9798919021   && 1030.5721050297 \\
-      &2s& 203.7584408153   &&  142.1022267153 \\
-      &2p&  20.2442429692   &&   40.9175702474 \\
-\end{tabular}
-\end{ruledtabular}
-%
-\footnotetext[1]{Fitting performed according to Appendix.~\eqref{a:auxiliary-basis} assuming 6-311++G(d,p) 
-and aug-cc-pVDZ-jkfit primary and test basis sets, respectively.}
-%
-\end{table}
-}
-%
-
-In this section, computational cost is estimated from the formulas.
-%In
-%Table~\ref{t:oep-costs} 
-%
-%{
-%\renewcommand{\arraystretch}{1.4}
-%\begin{table}[b]
-%\caption[Computational cost of the benchmark and OEP-based methods: calculation of coupling constant]
-%{{\bf Computational cost of the benchmark and OEP-based methods\footnotemark[1]}
-%}
-%\label{t:oep-costs}
-%\begin{ruledtabular}
-%\begin{tabular}{lcccccc}
-%Method             && Murrell et al. &&    EFP2        &&      OEP     \\ 
-%	\cline{1-7}
-%Group (i)         &&  &&        &&   $obQ$     \\
-%Group (ii)        &&  &&        &&   $ob$      \\
-%Group (iii)       &&  &&        &&   $ob(n+o)$     \\
-%Total Cost        &&  && $10ob^2 + 2b^3 + o^2b^2$       &&   $ob(Q+n+o)$     \\
-%\end{tabular}
-%\end{ruledtabular}
-%%
-%\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
-%$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
-%%
-%\end{table}
-%}
-%
-{
-\renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
-\caption[Minimal auxiliary basis set optimized for OEP\hyp{}based CT/HF0 calculations.]
-{{\bf}CPU timings in miliseconds of CT/HF0 single point energy calculations
-for water-methanol complex at 6-311++G(d,p) primary basis set\footnotemark[1]
-}
-\label{t:4.oep-costs}
-\begin{ruledtabular}
-\begin{tabular}{llccc}
-                      && Water-Methanol && Water dimer \\
-\cline{3-5}
-OL                    && 774    &&\\
-EFP2                  &&  42.3  &&\\
-OEP/aug-cc-pVQZ-jkfit &&   4.7  &&\\
-OEP/aug-cc-pVTZ-jkfit &&   4.0  &&\\
-OEP/aug-cc-pVDZ-jkfit &&   3.8  &&\\
-OEP/6-311++G(d,p)     &&   3.0  &&\\
-OEP/mini              &&   2.5  &&\\
-\end{tabular}
-\end{ruledtabular}
-%
-\footnotetext[1]{CPU: XXX. See the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
-%
-\end{table}
-}
-%
-
-%
-It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
-and many orders of magnitude lesser than the original theory of Murrell et al.
-%
-\begin{figure}[h]
-\includegraphics[width=0.46\textwidth]{fig-5.pdf}
-\caption{\label{f:fig-5} {\bf Asymptotic dependence of the charge transfer energy
-in the CT/HF0 formulation for methanol-warer complex with minimal OEP auxiliary basis.} 
-Water molecule has been translated by $R$ from the starting geometry
-optimized at HF/6-31+G(d,p) level,
-along the vector specified in the insets.
-The counterpoise\hyp{}corrected total interaction energy
-is also shown for comparison in light grey color in this figure.
-Interaction energies were obtained at the HF/6-311++G(d,p) level of theory.
-In this figure, `OEP' denotes single GDF scheme with aug-cc-pVQZ-jkfit auxiliary basis set
-whereas `OEP/mini` denotes double GDF scheme with the minimal auxiliary basis
-developed here for water and methanol.
-} 
-\end{figure}
-%
-
-\section{\label{s:6.conclusions}Summary and a few concluding remarks}
-
-In this work, the one\hyp{}electron effective potential operator technique
-for eliminating electron repulsion integrals in the fragment\hyp{}based
-theories of intermolecular interactions
-was derived and applied to calculating the charge transfer energy
-in condensed phases simulations. It was shown that in general case two types of OEP's
-can be defined and worked out either through the density fitting or the distributed
-multipole expansion. For the first group of OEP's, the density fitting was generalized
-and two computational treatments of OEP's were developed: the single GDF scheme
-for RI\hyp{}type auxiliary basis sets, as well as the double GDF scheme, for
-essentially all other types of basis sets. The presented validation of the OEP technique
-against the Otto\hyp{}Ladik CT/HF0 model as parent theory showed that in most cases
-elimination of ERI's is quantitatively accurate, with few exceptions
-where only qualitative accuracy was found. 
-Finally, it was found that, while the proposed OEP\hyp{}based Otto\hyp{}Ladik
-model is of comparable accuracy as the CT/HF0 formulation within the
-state\hyp{}of\hyp{}the\hyp{}art EFP2 model, it significantly outperforms the latter
-in terms of computational efficiency reaching speedups up to 30--40 times.
-%Since EFP2 method is based on the PT/HF0 foundations
-%with fragments defined in gas phase state,
-%interaction energies without the CT contribution
-%in this model
-%are
-%conceptually incomplete.
-Therefore, it is believed that the OEP\hyp{}based
-model can be incorporated within the EFP2 method, enabling
-the full (CT\hyp{}including) EFP2 energy calculations and 
-molecular dynamics simulations in large systems. However, further 
-study is necessary and the following issues need to be carefully addressed:
-
-\begin{acknowledgments}
-This project is carried out under POLONEZ programme which has received funding from the European Union's
-Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
-No.~665778. This project is funded by National Science Centre, Poland 
-(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
-We thank Professor Marcos Mandado from University of Vigo in Spain
-for providing us with his benchmark calculations to validate our implementation
-of the DDS method.
-\end{acknowledgments}
-
-%%
-\appendix
-
-\section{Optimized Auxiliary Basis Sets for OEP Applications\label{a:auxiliary-basis}}
-
-To fit the auxiliary DF basis for the treatment of the overlap\hyp{}like
-OEP matrix elements with the operator $\hat{v}_{\rm eff}$, 
-the following objective function is minimized
-%
-\begin{equation} \label{e:a1-obj}
- Z[\{\xi\}] = \sum_{\alpha i} \left[ 
-     \tBraKet{\alpha}{\hat{v}_{\rm eff}}{i} - 
-     \sum_{\xi}^{\rm DF} V_{\xi i} S_{\alpha \xi} 
-    \right]^2 \;,
-\end{equation}
-%
-where $\{\xi\}$ is the auxiliary basis set to optimize, whereas $\{\alpha\}$
-is the `test' basis set used to probe the accuracy of the density fitting.
-The forms of the %Pauli and 
-charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
-can be directly derived 
-from Eq.~\eqref{e:v-oep.ct}
-%from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
-by expanding the MO's in terms of atomic basis functions. The working formulae
-are given below:
-%
-\begin{subequations} \label{e:a1-v.oep}
- \begin{align}
-   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm Rep}}{i}
-     &= \sum_{x}^{\rm At} W_{\alpha i}^{(x)} \nonumber   \\ 
-        + \sum_{\beta\gamma\delta} &
-           \left\{ 
-             2 C_{\beta i} D_{\gamma\delta} - C_{\gamma i} D_{\beta \delta}
-           \right\}
-           \BraKet{\alpha\beta}{\gamma\delta} \;,\\
-   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{n} 
-     &= \sum_{x}^{\rm At} W_{\alpha n}^{(x)} \nonumber   \\
-       + \sum_{\beta\gamma\delta} &
-           \left\{
-             2 C_{\beta n} D_{\gamma\delta} - C_{\gamma n} D_{\beta \delta}
-           \right\}
-           \BraKet{\alpha\beta}{\gamma\delta}
-    \;.
- \end{align}
-\end{subequations}
-%
-In the above equations, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
-is the one\hyp{}particle density matrix in AO basis.
-
-\section{Coulomb-like OEP via Cumulative Atomic Multipole Moments\label{a:oep-camm}}
-
-Overlap\hyp{}like OEP's are always associated with the effective one\hyp{}particle density
-(or bond\hyp{}order) 
-matrices and the $\lambda$ parameter from Eq.~\eqref{e:oep-operator} being either 1 or 0. 
-Therefore, one can use this fact to compute the effective distributed multipole
-moments. In this work, the cumulative atomic multipole moments (CAMM)
-of Sokalski and Poirier with Mulliken partitioning of the AO space
-were chosen because of their simplicity.
-One of examples
-of such OEP\hyp{}based models reported already in the literature is
-the transition cumulative atomic multipole moments (TrCAMM) model
-for estimation of the excitation energy transfer (EET) couplings in the F{\"o}rster
-limit\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}, 
-or the vibrational solvatochromic distributed multipole moments
-(SolCAMM) for the determination of the vibrational frequency shifts
-of spatially localized IR probes.\cite{Blasiak.Lee.Cho.JCP.2013}
-In general, having the OED and $\lambda$, the effective cumulative atomic charges
-and dipole moments are given by
-%
-\begin{subequations} \label{e:a1-v.oep-camm}
- \begin{align}
-  q_x^{\rm eff} &= \lambda Z_x - 
-       \sum_{\alpha\in x}\sum_{\beta} 
-        P_{\alpha\beta}^{\rm eff} \BraKet{\alpha}{\beta} \;, \\
-  {\BM\upmu}_x^{\rm eff} &= 
-       \sum_{\alpha\in x}\sum_{\beta} 
-        P_{\alpha\beta}^{\rm eff}
-   \left\{ 
-    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\bf r}{\beta} 
-   \right\}
-   \;,
- \end{align}
-\end{subequations}
-%
-where ${\bf R}_x$ is the position vector of the $x$-th nucleus.
-Note that $P_{\alpha\beta}^{\rm eff}$ can refer to the particular electron
-spin, or their sum as a bond order matrix as well.
-Equations for the effective distributed quadrupoles, octupoles and hexadecapoles
-can be easily obtained by analogy from Ref.~\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}
-
-%
-%\section{\label{a:mcmurchie-davidson} McMurchie-Davidson Method for Asymmetric Electron Repulsion Integrals}
-%
-%Evaluation of integrals $(ijk\vert l)$,
-%that are necessary for the generalized density fitting of OEP's,
-%can be easily carried out by re\hyp{}expressing the
-%unnormalized product of three primitive Gaussian\hyp{}type (GTO) functions,
-%%
-%\begin{equation} \label{e:1}
-%[ijk] \equiv \phi_i({\bf r}) \phi_j({\bf r}) \phi_k({\bf r}) \;,
-%\end{equation}
-%%
-%in terms of Hermite functions. It can be shown that
-%%
-%\begin{multline} \label{e:2}
-%   [ijk] = E_{ijk} \sum_{N=0}^{n_1+n_2+n_3} \sum_{L=0}^{l_1+l_2+l+3} \sum_{M=0}^{m_1+m_2+m_3} 
-%	\\
-%          d_N^{n_1n_2n_3} d_L^{l_1l_2l_3} d_M^{m_1m_2m_3}
-%          \Lambda_N(x_R)\Lambda_L(y_R)\Lambda_M(z_R)e^{-\alpha_Rr_R^2}
-%\end{multline}
-%%
-%in which the McMurchie-Davidson $d3$ coefficients are given by the following recurrence relationships
-%%
-% \begin{align} \label{e:3}
-%  d_N^{n_1+1,n_2,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-%                             + \vert {\bf R} - {\bf A}\vert_x d_N^{n_1n_2n_3} \nonumber \\ 
-%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
-%  d_N^{n_1,n_2+1,n_3} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-%                             + \vert {\bf R} - {\bf B}\vert_x d_N^{n_1n_2n_3} \nonumber \\
-%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} \\
-%  d_N^{n_1,n_2,n_3+1} &= \frac{1}{2\alpha_R} d_{N-1}^{n_1n_2n_3} 
-%                             + \vert {\bf R} - {\bf C}\vert_x d_N^{n_1n_2n_3} \nonumber \\
-%			    & \qquad\qquad + (N+1) d_{N+1}^{n_1n_2n_3} 
-%\end{align}
-%%
-%with $d_0^{000} = 1$.
-%In Eq.~\eqref{e:1},
-%$\phi_i({\bf r})$ is given by
-%%
-%\begin{equation}
-%\phi_i({\bf r}) \equiv x_A^{n_1} y_A^{l_1} z_A^{m_1} e^{-\alpha_1r_A^2}
-%\end{equation}
-%%
-%where 
-%${\bf r}_A \equiv {\bf r} - {\bf A}$,
-%${\bf A}$ is the centre of the GTO, $\alpha_1$ its exponent, whereas $n_1,l_1,m_1$
-%the Cartesian angular momenta, with the total angular momentum $\theta_1 = n_1+l_1+m_1$.
-%
-%It can be easily shown that the multiplicative constant $E_{ijk}$ is given by
-%%
-%\begin{multline}
-%  E_{ijk}(\alpha_1,\alpha_2,\alpha_3)  = \exp{\left[-\frac{\alpha_1\alpha_2}
-%                                        {\alpha_1+\alpha_2}\vert {\bf A}-{\bf B}\vert^2\right]} \\ \times
-%                                         \exp{\left[-\frac{(\alpha_1+\alpha_2)\alpha_3}
-%                                        {\alpha_1+\alpha_2+\alpha_3}
-%                                         \vert {\bf P}-{\bf C}\vert^2\right]} 
-%\end{multline}
-%%
+\section{Statistical validation of OEP method}
 

### 2019-12-08 papers/2_oep_technique.CT/paper-r3.tex
Revised paper uploaded

--- 2019-11-29 papers/2_oep_technique/paper-r2.tex
@@
+\documentclass[aip,jcp,amsmath,amssymb,reprint,floatfix]{revtex4-1}
 \input{header.tex}
@@
 molecular aggregates such as interacting molecules in condensed phases.
-Here, the general method for effective elimination of the electron repulsion integrals (ERI)
-based on OEP's is presented, that is tuned towards the fragment\hyp{}based calculation methodologies
-such as the second generation of the effective fragment potentials (EFP2) method. 
+Here, a general method for effective OEP\hyp{}based elimination of electron repulsion integrals (ERI),
+that is tuned
+towards the fragment\hyp{}based calculation methodologies
+such as the second generation of the effective fragment potentials (EFP2) method,
+is presented.
 Two general types of the OEP operator matrix elements
@@
 or the extended density fitting schemes developed in this work. 
-The OEP technique was then applied
-to address the problem of using usually incomplete EFP2 settings in most of
+The OEP technique is then applied
+to address the problem of using incomplete EFP2 settings in many
 applications in interaction energy and 
@@
 Therefore, it is believed that
-the proposed model can be used in calculations within 
+the proposed model can be used within 
 the EFP2 framework, making the CT energy term no longer the bottleneck
@@
 its notably complex quantum mechanical (QM) origins 
-even at the Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
-approximation level,\cite{Otto.Ladik.ChemPhys.1975}
+even at the Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} (HF)
+approximation level,\cite{Jensen.JCP.2001,Otto.Ladik.ChemPhys.1975}
 and cannot be realized in terms of any classical nor semi\hyp{}classical approach.
@@
 calculation of intermolecular forces in molecular dynamics because
-they involces instantaneous calculations of the electron repulsion integrals (ERI's)
+they involve calculation of the electron repulsion integrals (ERI's)
 and their four\hyp{}index transformation to molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}
@@
 for modeling structure
-and dynamics in condensed phases.
+and dynamics in condensed phases.\cite{Demerdash.Mao.Liu.Head-Gordon.Head-Gordon.JCP.2017}
 This needs to be contrasted with the Coulombic electrostatics,
@@
 can be well described by, respectively, 
-the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
+the distributed multipole moments of charge densities,\cite{Sokalski.Poirier.CPL.1983,
 Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
@@
 %to the total interaction energy, 
-are not explicitly included in any molecular mechanics force field developed up to date.
+are not explicitly included in any molecular mechanics force field 
+developed up to date.\cite{Demerdash.Yap.Head-Gordon.AnnuRevPhysChem.2014}
 %such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
 The only way to
-incorporate the CT effects in the condensed\hyp{}phase simulations,
-apart from performing full QM electronic structure simulations,
+incorporate the CT effects in the condensed\hyp{}phase simulations
+in \emph{ab initio} manner,
+apart from performing full QM electronic structure simulations
+or non\hyp{}force\hyp{}field\hyp{}based 
+fragmentation techniques\cite{Leverentz.Maerzke.Keasler.Siepmann.Truhlar.PCCP.2012,
+Dahlke.Truhlar.JCTC.2008},
 is through the second generation of the effective fragment potential 
@@
    Nguyen.Pachter.Day.JCP.2014,
-   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
-developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
+   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996,
+   Sattasathuchana.Xu.Gordon.JPCA.2019,
+   Blasiak.Londergan.Webb.Cho.ACR.2017,
+   Kuroki.Mori.ChemLett.2016,
+   Ghosh.Cho.Choi.JPCB.2014}
+developed by Jensen, Gordon and co\hyp{}workers.\cite{Jensen.JCP.2001,Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
 EFP2 is an \emph{ab initio} force field
-derived on the grounds of the first\hyp{}principles perturbation theory
-at the HF level and including the intermolecular dispersion effects
-by the response theory. 
+derived on the grounds of the first\hyp{}principles %perturbation theory
+at the HF level\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998,Li.Netzloff.Gordon.JCP.2006} 
+and including the intermolecular dispersion effects
+by the response theory.\cite{Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014}
 %based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
@@
 where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
-distributions of the monomers, approximated by the distributed multipole approximation,
+distributions of the monomers, treated by the distributed multipole approximation
+with damping to account for the charge\hyp{}penetration effects,\cite{Slipchenko.Gordon.JCC.2007}
 $U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
-principle, that could be roughly understood as the Heitler\hyp{}London interaction
-energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
+principle,\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998} 
+that could be roughly understood as the Heitler\hyp{}London interaction
+energy\cite{Chalasinski.Gutowski.MolPhys.1985} with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
 are the induction and dispersion energies obtained from the distributed polarizability
-approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.
+approximation,\cite{Li.Netzloff.Gordon.JCP.2006,Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014} 
+and finally $U^{\rm CT}$ is the CT energy,\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013} 
+the focus of this work.
 
@@
 in accurately modeling the extended molecular systems 
-like solutions and recently even biomolecules
+like solutions\cite{Sattasathuchana.Xu.Gordon.JPCA.2019,
+   Blasiak.Londergan.Webb.Cho.ACR.2017,
+   Kuroki.Mori.ChemLett.2016,
+   Ghosh.Cho.Choi.JPCB.2014}
+and recently even biomolecules\cite{Blasiak.Ritchie.Webb.Cho.PCCP.2016,
+Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018,
+Gurunathan.Acharya.Ghosh.Kosenkov.Kaliman.Shao.Krylov.Slipchenko.JPCB.2016,
+Ghosh.Kosenkov.Vanovschi.Williams.Herbert.Gordon.Schmidt.Slipchenko.Krylov.JPCA.2010}
 with the level of accuracy
-roughly arguably comparable to the M{\o}ller\hyp{}Plesset perturbation theory,
-evaluation of the CT energy in EFP2 model is still too costly for
+reaching in many cases\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
+the M{\o}ller\hyp{}Plesset perturbation theory\cite{Moller.Plesset.PhysRev.1934},
+evaluation of the CT energy in EFP2 model is still relatively costly for
 typical uses in the
 molecular dynamics simulations. It has been reported that 
-the implementation of the EFP2 CT energy component in GAMESS
-is on average 3-10 times more demanding than the exchange\hyp{}repulsion
+the implementation of the EFP2 CT energy and gradient in GAMESS US computer program\cite{GAMESS.JCC.1993}
+is on average 30--40 times more demanding than the exchange\hyp{}repulsion
 component, becoming a bottleneck of evaluation of total interaction 
@@
 Li.Gordon.Jensen.JCP.2006}
-In effect, the CT energy component is usually ignored in most of the applications. 
-As a noteworthy example, 
-CT term were not implemented in the official release 
+In effect, the CT energy component is often ignored 
+in many applications.\cite{Kuroki.Mori.JPCB.2019,
+Blasiak.Londergan.Webb.Cho.ACR.2017,
+Tommaso.Lafiosca.Cappelli.JCTC.2017,
+Gurunathan.Acharya.Ghosh.Kosenkov.Kaliman.Shao.Krylov.Slipchenko.JPCB.2016,
+Budzak.Laurent.Laurence.Medved.Jacquemin.JCTC.2016,
+Smith.Gordon.Slipchenko.JPCA.2011,
+Smith.Gordon.Slipchenko.JPCA.2011.2,
+Ghosh.Kosenkov.Vanovschi.Williams.Herbert.Gordon.Schmidt.Slipchenko.Krylov.JPCA.2010,
+Smith.Gordon.Slipchenko.JPCA.2008}
+In fact,
+EFP2 CT term is available only in the GAMESS US quantum chemistry program,\cite{GAMESS.JCC.1993}
+whereas it is neither supported in the official release 
 of the recent LIBEFP library for linking quantum chemistry packages 
-with the EFP2 functionalities.\cite{Kaliman.Slipchenko.JCC.2015}
+with the EFP2 functionalities,\cite{Kaliman.Slipchenko.JCC.2015}
+nor in the Q\hyp{}CHEM quantum chemistry 
+program,\cite{Ghosh.Kosenkov.Vanovschi.Flick.Kaliman.Shao.Gilbert.Krylov.Slipchenko.JCC.2013}
+contrary to electrostatic, exchange\hyp{}repulsion, induction and dispersion EFP2 terms.
 
@@
 in donor\hyp{}acceptor systems such as H\hyp{}bonded species
-and charged complexes, although highly depends on the theoretical
+and charged complexes,\cite{Devarajan.Gaenko.Gordon.Windus.Fragmentation.2017,
+Devarajan.Windus.Gordon.JPCA.2011} 
+although highly depends on the theoretical
 approach being used, i.e., the choice of the reference
@@
 %as references 
-here broadly referred to as the CT/HF0 level of theory.
+which will be here broadly referred to as the CT/HF0 level of theory
+and the symbol `0' denotes unperturbed wavefunctions.
 In this formulation, the CT energy can be understood
@@
 unlike the pure induction and dispersion, in which such mixings are contained
-solely within one monomer.
+solely within one monomer.\cite{Stone.TheTheoryOfIntermolecularForces.1996}
 %This level of definition of the CT energy will be referred here as to
@@
 is compatible with the EFP2 energetics, is to
-extend the effective one\hyp{}electron potentials (here referred to as the OEP) 
+extend the one\hyp{}electron effective potentials (here referred to as the OEP) 
 technique
@@
 to both effective potential operators
-and dropping the `$\hat{}$' symbol in the multipole operators
+and dropping the $\hat{}$ symbol in the multipole operators
 which leads to\cite{Stone.TheTheoryOfIntermolecularForces.1996}
@@
 There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
-of Stone\cite{Stone.JCTC.2005},
+of Stone and Alderton\cite{Stone.Alderton.MolPhys.1985,Stone.JCTC.2005},
 the cumulative atomic multipole moments (CAMM) of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983},
@@
 %
-Now, it is first assumed here that $\rho^{B}_{nj} ({\bf r}) \approx 0$ if occupied MO's on $B$ are localized.
-This means that
-%
-\begin{multline} \label{e:ct.group-iii.final.a}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
-     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
-        S_{kj} \left( 1 - \delta_{ik} \right)
-        \BraKet{nj}{ik} \\ 
-    \approx 0 \;.
-\end{multline}
-%
-The above result is a crude approximation, since the channels $j\rightarrow n$
-are quite important contributions in most of perturbation theory treatments.
+%Now, it is first assumed here that $\rho^{B}_{nj} ({\bf r}) \approx 0$ if occupied MO's on $B$ are localized.
+%This means that
+%%
+%\begin{multline} \label{e:ct.group-iii.final.a}
+%        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
+%     + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
+%        S_{kj} \left( 1 - \delta_{ik} \right)
+%        \BraKet{nj}{ik} \\ 
+%    \approx 0 \;.
+%\end{multline}
+%%
+%The above result is a crude approximation, since the channels $j\rightarrow n$
+%are quite important contributions in most of perturbation theory treatments.
 %virtual MO's can penetrate the occupied MO's to some extent.
@@
 can be already evaluated in CMO basis.
+This leads to the coupling constant of the following form,
+%
+\begin{equation} \label{e:ect-oep.final-vin}
+   V_{in}^{A\rightarrow B} \approx 
+%
+   G_{1,in}^{A\rightarrow B} \\
+  +
+   \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
+   \left\{
+   G_{2,i'n}^{A\rightarrow B}
+  +G_{3,i'n}^{A\rightarrow B}
+  \right\} \;,
+\end{equation}
+%
+where the subscripts $G_n$ ($n$ = 1, 2, 3) denote a particular group of 
+terms from Eqs.~\eqref{e:ct.group-i.final}, \eqref{e:ct.group-ii.final} 
+and \eqref{e:ct.group-iii.final.b}.
+%or \eqref{e:ct.group-iii.final.b}. 
 Thus, the final working formula for the interaction energy due to CT from $A$ to $B$ reads
@@
   \right\}
- \Bigg)^2 \;,
+ \Bigg)^2 \;.
 \end{multline}
 %
-where the subscripts $G_n$ ($n$ = 1, 2, 3) denote a particular group of 
-terms from Eqs.~\eqref{e:ct.group-i.final}, \eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.a}
-or \eqref{e:ct.group-iii.final.b}. 
 The total CT energy is given by the sum of the above contribution and the twin contribution
@@
 were obtained by
-energy\hyp{}optimizations at the HF/6-31+G(d,p) level,
+performing energy\hyp{}optimizations at the HF/6-31+G(d,p) level,
 as implemented in 
@@
 the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the DI6/04 dipole interaction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
 the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
@@
 and 
-the PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+the PPS5/05 $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
 
@@
 according to the density decomposition scheme (here referred as to the DDS) 
-proposed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
 whereas $\Delta E^{\rm Ind}$ term is the 
@@
 are the exchange and the Pauli\hyp{}repulsion interaction energies, respectively.
-At the HF level of theory, the sum of the latter two terms is in fact
-equivalent to the first\hyp{}order exchange\hyp{}repulsion energy 
+At the HF level of theory, the sum of the latter two terms,
+i.e., the exchange\hyp{}repulsion energy, $\Delta E^{\rm Ex-Rep}_{\rm DDS/HF}$,
+is in fact
+equivalent to the exchange\hyp{}repulsion energy 
 in the intermolecular perturbation theory with exchange of Hayes and Stone.\cite{Hayes.Stone.MolPhys.1984} 
@@
 All the models that were used to test the theory presented in this work,
-i.e., the OL, EFP2 and OEP methods, as well as the benchmark EDF, CT/HF0 and DDS methods,
+i.e., the OL, EFP2 and OEP methods, as well as the EDF and the benchmark CT/HF0 and DDS methods,
 were implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
-DDS method was implemented in the dimer\hyp{}centered basis set
+DDS method was implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985}
 in order to eliminate the BSSE and properly inlcude the CT processes.\cite{Stone.Misquitta.CPL.2009}
+Estimations of the $\Delta E^{\rm Coul}_{\rm DDS/HF}$, $\Delta E^{\rm Ex-Rep}_{\rm DDS/HF}$,
+$\Delta E^{\rm Ind}$ and $\Delta E^{\rm CT}_{\rm EFP2}$ interaction energy components 
+obtained by using the computer 
+code developed for this work were
+found to be equivalent with the EFP2 interaction energy components
+obtained by using the {\sc GAMESS US} package\cite{GAMESS.JCC.1993} 
+(see Supplementary Information).
+For the CT EFP2 component,
+%Due to convenience and simplicity of implementation, 
+potential energy integrals
+from Eq.~\eqref{e:coulomb-energy:dmtp-oei} 
+were calculated with the
+CAMM up to quadrupoles
+(distributed centers are atoms), 
+instead of the DMA 
+(distributed centers are atoms and mid\hyp{}bond points) 
+as implemented in most
+of quantum chemistry programs. The choice of CAMM versus DMA
+was due to convenience of implementation and, because
+the quantitative accuracy of our EFP2 CT energy code is comparable to
+the EFP2 code of {\sc GAMESS US}, using only atomic distribution centers
+does not affect the interpretation
+of results in Section~\ref{s:5.results}.
 
@@
 BBI\footnotemark[6]   
-      &&  Subset\footnotemark[7]
-                  && 0.26 && 0.56 && 0.13* && 0.12 \\ %
+      &&  Total
+                  && 0.26 && 0.56 && 0.13 && 0.12 \\ %
 \end{tabular}
@@
 \footnotetext[6]{Ref.~\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}}
-\footnotetext[7]{100 backbone\hyp{}sidechains complexes were randomly selected.}
 %
@@
 OEP performance from NCB31 and BBI sets, it was estimated
-that $C\approx 1.55$. Application of the semi\hyp{}empirical model
+that $C\approx 1.56$. Application of the semi\hyp{}empirical model
 from Eq.~\eqref{e:ect-oep.final-scaled} significantly improved
@@
 \includegraphics[width=0.9\textwidth]{fig-2.pdf}
-\caption{\label{f:fig-2} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
-in small bi\hyp{}molecular neutral complexes.}
-(a) accuracy of the ERI elimination technique; 
-(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
-CT/HF0 estimate. 
-The structures of bi\hyp{}molecular complexes
-were taken from the NCB31 database developed by the Truhlar's group:
+\caption{\label{f:fig-2} {\bf Accuracy of the ERI elimination technique.}
+(a) NCB31 
+database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
+and
+(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+from the BioFragment Database.
+The EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+for the OEP calculations.
+} 
+\end{figure*}
+%
+
+%
+\begin{figure*}[h]
+\includegraphics[width=0.9\textwidth]{fig-3.pdf}
+\caption{\label{f:fig-3} {\bf 
+Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
+in bi\hyp{}molecular neutral complexes.}
+(a) NCB31 set from database
+developed by the Truhlar's group\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}:
 the HB6/04 hydrogen bonding database 
-(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-the DI6/04 dipoleinteraction database 
+(black),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
+the DI6/04 dipole interaction database 
 (green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
@@
 and the PPS5/05 the $\pi$-$\pi$ stacking database 
-(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-6-311++G(d,p) basis set was used for the CT energy calculations
-and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
-for the OEP calculations.
-} 
-\end{figure*}
-%
-
-%
-\begin{figure*}[h]
-\includegraphics[width=0.9\textwidth]{fig-3.pdf}
-\caption{\label{f:fig-3} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
-in aminoacid pairs.}
-(a) accuracy of the ERI elimination technique; 
-(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
-CT/HF0 estimate. 
-The structures of 100 aminoacid pairs
-were taken from the BBI subset of the BioFragment Database
-for backbone\hyp{}backbone interactions 
-in proteins.\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017}
-6-31+G(d,p) basis set was used for the CT energy calculations
-and the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+(red);\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
+(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+from the BioFragment Database.
+The EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
 for the OEP calculations.
@@
             & \multirow{2}{*}{$\varepsilon_i$, $\varepsilon_n$} \\
-parameters && & \\
+parameters\footnotemark[2] 
+           && & \\
 \hline
 %\multirow{2}{*}{Rotatable}  
-Superimposable  && \multirow{2}{*}{$C_{\alpha i}$, $C_{\alpha n}$, $\{\alpha\}$} 
-            & $C_{\alpha i}$, $C_{\alpha n}$, $\{\alpha\}$, \\
-parameters && & $\{\eta\}$, $V_{n\eta}^{{\rm eff},B}$ \\
+Superimposable  && \multirow{2}{*}{$C_{\alpha i}^A$, $C_{\beta n}^B$, $\{\alpha\}$, $\{\beta\}$  } 
+            & $C_{\alpha i}^A$, $C_{\alpha n}^B$, $\{\alpha\}$, $\{\beta\}$, \\
+parameters\footnotemark[2] 
+                && & $\{\eta\}$, $V_{n\eta}^{{\rm eff},B}$, $L_{ii'}^A$ \\
 \hline
-\multirow{3}{*}{Calculables} 
-           && $S_{ij}$, $S_{nk}$, $S_{nw}$,            &   $S_{\eta i}$, $S_{ni}$, $S_{ij}$, \\
-           && $T_{nn}$, $T_{kj}$, $T_{wj}$, $T_{nj}$   &  $u_i^{BA}$, $w_{yi}^{BA}$ \\
+\multirow{3}{*}{Calculables\footnotemark[2]} 
+           && $S_{ij}$, $S_{nk}$, $S_{nw}$,            &   $S_{ij}$, $S_{\eta i}$, $S_{ni}$, \\
+           && $T_{nn}$, $T_{kj}$, $T_{wj}$, $T_{nj}$,  &  $u_i^{BA}$, $w_{yi}^{BA}$\\
            && $V_{in}^{{\rm EF},B}$ , $V_{ik}^{{\rm EF},B}$, $V_{iw}^{{\rm EF},B}$
@@
 \hline
-Cost       && $(10on^2 + 2n^3 + o^2n^2)v$ & $on(Q+N+o)$ \\ 
+\multirow{4}{*}{Cost\footnotemark[3]}    
+                         && $sp\left( 2p^2+2op+o^2\right)$ & \\
+                         &&$+tp\left( 2p^2+2op+o^2\right)$ & $sop\left( 2p+o+a\right)$ \\ 
+                         &&$+vop\left( 3p+o\right)$        & $+op(a+oN+2o)$\\             
+                         && $+o^2p$  & \\
 \end{tabular}
@@
 %
-\footnotetext[1]{Numbers of: $o$ - occupied molecular orbitals; $b$ - atomic basis functions; $n$ - atoms; 
-$Q$ - auxiliary basis functions. It was assumed that the number of virtual orbitals is equal to $b$.}
+\footnotetext[1]{Based on coupling constant
+expressions from Eq.~\eqref{e:u-efp2-ct-v2} and Eq.~\eqref{e:ect-oep.final-vin} 
+for EFP2 and OEP
+method, respectively.}
+\footnotetext[2]{The subscript meaning is as follows: 
+primary basis set functions of $A$: $\alpha$;
+primary basis set functions of $B$: $\beta$;
+auxiliary basis set functions of $B$: $\eta$,
+occupied MO's of $A$: $i$, $i'$, $k$;
+occupied MO's of $B$: $j$;
+virtual MO's of $A$: $w$; 
+virtual MO's of $B$: $n$; 
+atoms of $B$: $y$.}
+\footnotetext[3]{
+Numbers of: 
+primary basis set functions - $p$;
+auxiliary basis set functions - $a$;
+occupied MO's - $o$;
+atoms - $N$. 
+Relative costs: $v$ - multipole potential, $t$ - kinetic energy and $s$ - overlap OEI's.
+It was assumed that the number of virtual orbitals is equal to $n$.}
 %
@@
 and large AO basis sets,
-the computational cost is
-roughly proportional to $(10on^2 + 2n^3 + o^2n^2)v$, where the
-$o$ and $n$ denote the number of occupied orbitals and the number of atomic basis functions,
-respectively. $v$ is the relative cost of evaluation of the
-most expensive type of OEI's, i.e., the potential and potential gradient(s)
-integrals contracted with the DMTP's, that are necessary to compute ${\bf V}^{\rm EF}$ matrices
+the computational cost
+has an order of $2p^3(s+t) + 3vop^2$, where the
+$o$ and $p$ denote the number of occupied orbitals and the number of atomic basis functions,
+respectively. 
+$s$, $t$ and $v$ are the relative costs of evaluation of the
+overlap, kinetic energy, and multipole potential OEI's, respectively, with the latter being
+most expensive but
+%type of OEI's, i.e., the potential and potential gradient(s)
+%integrals contracted with the DMTP's, 
+%that are 
+necessary to compute ${\bf V}^{\rm EF}$ matrices
 from Eq.~\eqref{e:v-efp2-eff}. 
@@
 from Eq.~\eqref{e:ect-oep.final} requires only overlap OEI's that are the least expensive, 
-and has the cost of approximately $on(Q+N+o)$, where 
-$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
-respectively. Note also that, among the calculables that are needed on the fly in each
+and has the cost of approximate magnitude of $2sop^2$ for relatively small auxiliary basis sets.
+%where 
+%$Q$ and $N$ denote the number of auxiliary basis functions and the number of atoms,
+%respectively. 
+Note also that, among the calculables that are needed on the fly in each
 OEP\hyp{}based CT energy evaluation, are the auxiliary vectors and matrices 
@@
 small auxiliary basis set is used.
+For example,
+assuming a water dimer system and 6-311++G(d,p) primary and minimal auxiliary basis set
+with $s=t=v\approx1$,
+the OEP method is predicted to be roughly 12--16 times faster than EFP2 method. In practice,
+the parameters $s$, $t$ and $v$ will have larger values, especially $v$.
+
 {
@@
 OEP/6-311++G(d,p)     &&   3.0  &&     1.0\\
-OEP/mini              &&   2.5  &&     0.75\\
+OEP/mini\footnotemark[2]&& 2.5  &&     0.75\\
 \end{tabular}
@@
 See also the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
+\footnotetext[2]{This work, Table~\ref{t:3}.}
 %
@@
 
-In this work, the minimal uncontracted basis sets were optimized for water and methanol
+In this work, the minimal uncontracted auxiliary basis sets were optimized for water and methanol
 that consist of only $s$ and $p$\hyp{}type functions,
 and the orbital exponents are shown in Table~\ref{t:3},
-and the procedure is outlined in Appendix~\ref{a:auxiliary-basis}.
+whereas the procedure is outlined in Appendix~\ref{a:auxiliary-basis}.
 Unfortunately, we were not able to successfully fit the auxiliary basis sets
@@
 In this work, the one\hyp{}electron effective potential operator technique
-for eliminating electron repulsion integrals in the fragment\hyp{}based
+of eliminating electron repulsion integrals in the fragment\hyp{}based
 theories of intermolecular interactions
@@
 all the summations over the virtual orbitals can be optimized,
-e.g., by using the QUANBO's, following Xu and Gordon.\cite{Xu.Gordon.JCP.2013}
+e.g., by using the QUAMBO's, following Xu and Gordon.\cite{Xu.Gordon.JCP.2013}
 In addition, effective atomic charges are most of the time very small
 and could be neglected, leading to further cost reductions.
+In addition, the possibility of the OEP\hyp{}based formulation of the direct EFP2 CT formulation
+should be studied in the future as well.
 %Since EFP2 method is based on the PT/HF0 foundations

### 2019-12-09 papers/2_oep_technique.CT/paper-r3.tex
Revisions in II paper added. III paper initialized.

@@
 as implemented in 
-the {\sc Gaussian~16} quantum chemistry program package.\cite{Gaussian16}
+the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
 Subsequently, 30 displaced geometries for each model complex
@@
 \begin{ruledtabular}
-\begin{tabular}{llccc}
-                      && Water-Methanol && Water dimer \\
-\cline{3-5}
-OL                    &&14749.1 &&  2909.9\\     %14716.1 &&  2876.9 added 33 sec and 17 sec for ERI/AO calculation from timer.dat
-EFP2                  &&  42.3  &&    15.4\\
-OEP/aug-cc-pVQZ-jkfit &&   4.7  &&     1.8\\
-OEP/aug-cc-pVTZ-jkfit &&   4.0  &&     1.5\\
-OEP/aug-cc-pVDZ-jkfit &&   3.8  &&     1.4\\
-OEP/6-311++G(d,p)     &&   3.0  &&     1.0\\
-OEP/mini\footnotemark[2]&& 2.5  &&     0.75\\
+\begin{tabular}{llccccc}
+                      && \multicolumn{2}{c}{Water-Methanol} && \multicolumn{2}{c}{Water dimer} \\
+\cline{3-4}
+\cline{6-7}
+OL                    &&1.48$\times 10^{4}$ 
+                                &(-0.81)&& 2.91$\times 10^{3}$     
+                                                    &(-0.85) \\  %14716.1 &&  2876.9 added 33 sec and 17 sec for ERI/AO calculation from timer.dat
+EFP2                  &&  42.7  &(-1.27)&&   15.6   &(-1.06) \\
+OEP/aug-cc-pVQZ-jkfit &&   4.36 &(-1.03)&&    1.64  &(-1.05) \\
+OEP/aug-cc-pVTZ-jkfit &&   3.63 &(-1.07)&&    1.35  &(-1.09) \\
+OEP/aug-cc-pVDZ-jkfit &&   3.60 &(-1.05)&&    1.26  &(-1.03) \\
+OEP/6-311++G(d,p)     &&   2.74 &(-1.37)&&    0.933 &(-1.31) \\
+OEP/mini\footnotemark[2]&& 2.25 &(-1.34)&&    0.623 &(-1.13) \\
 \end{tabular}
@@
 \footnotetext[1]{1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor, calculations performed on 1 core. 
+CT/HF0 energies are given in parentheses for reference (kcal/mol).
 See also the implementation details in Section~\ref{s:4.calculations} and Section~\ref{ss:5.2.cost}.}
@@
 that consist of only $s$ and $p$\hyp{}type functions,
-and the orbital exponents are shown in Table~\ref{t:3},
+and the orbital exponents are shown in Table~\ref{t:3}
 whereas the procedure is outlined in Appendix~\ref{a:auxiliary-basis}.
@@
 The total time required for evaluation of the CT/HF0 energies by using EFP2 and OEP
-models was measured for this system and also for the water-water system in their reference geometries,
+models was measured for this system and also for the water\hyp{}water system in their reference geometries,
 and the results are shown in Table~\ref{t:5}. 
@@
 using the OED in Eq.~\eqref{e:oed-group-iii}
-was not included because these quantities constitute the effective fragments
+were not included because these quantities constitute the effective fragments
 by definition.
 The above profiling settings allows one to compare the performances
-of OL, EFP2 and OEP methods as effective fragment parameter methods.
+of the OL, EFP2 and OEP methods as effective fragment parameter methods.
 
 Even for the rather large quadruple\hyp{}$\zeta$ auxiliary basis set,
-the OEP\hyp{}based model is roughly 9 times more efficient than evaluating the EFP2 CT energy.
+the OEP\hyp{}based model is roughly 10 times more efficient than evaluating the EFP2 CT energy.
 Reducing the size of auxiliary basis set up to minimal basis from Table~\ref{t:3}
-scales down the cost further by a factor of $1.9$ in water-methanol system,
-and $2.4$ in water dimer system, reaching in total roughly 20\hyp{}fold
+scales down the cost further by a factor of $2.0$ in water\hyp{}methanol system,
+and $2.6$ in water dimer system, reaching in total roughly 20\hyp{}fold
 reduction in CPU time as compared to the EFP2 model, and more than three orders of magnitude
@@
 in a role of auxiliary basis set 
-gave already accurate results, the STO-3G basis as auxiliary space resulted
-in CT energies of around $-8$ kcal/mol (data not shown). Therefore, the basis set optimization
+gave already accurate results, the minimal STO-3G basis as auxiliary space resulted
+in CT energies of $-8.4$ and $-8.8$ kcal/mol in water\hyp{}methanol and water dimer systems (data not shown),
+much lower than the reference values of $-0.81$ and $-0.85$ kcal/mol, respecively. 
+Therefore, the basis set optimization
 described in Appendix~\ref{a:auxiliary-basis} was necessary for the minimal basis to be applicable
@@
 e.g., by using the QUAMBO's, following Xu and Gordon.\cite{Xu.Gordon.JCP.2013}
-In addition, effective atomic charges are most of the time very small
-and could be neglected, leading to further cost reductions.
+Secondly, effective atomic charges $q_y^{B,(nj)}$ from Eq.~\eqref{e:ct.group-iii.final.b}
+are most of the time very small
+and could be neglected, leading to further cost reductions (not shown here).
 In addition, the possibility of the OEP\hyp{}based formulation of the direct EFP2 CT formulation

### 2019-12-09 papers/3_oep_technique.Pauli/paper-r0.tex
Revisions in II paper added. III paper initialized.

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

\title{Ab Initio Effective One-Electron Potential Operators. II.
Applications for Exchange-Repulsion Energy in Effective Fragment Potentials}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}
Recently, it was shown that
the effective one\hyp{}electron potentials (OEP) method
of effective elimination of the electron repulsion integrals
can be used to lower the computational cost
of evaluation of the charge transfer interaction energy
within the effective fragment potential method (EFP2)
by the factor of 20, making it no longer a bottleneck
in EFP2\hyp{}based simulations of complex systems.
Here, OEP technique 
is used to enhance computational efficiency in evaluating the
exchange\hyp{}repulsion EFP2 interaction energy
by incorporating the kinetic energy operator in the OEP
via the extended density fitting in incomplete auxiliary basis.
It is shown that the computational cost can be reduced by a factor
of 1.5 and kinetic energy integrals are no longer necessary to evaluate.
\end{abstract}

\pacs{}

\maketitle

\tableofcontents

\section{\label{s:1.introduction}Introduction}

The exchange\hyp{}repulsion (EXR) interaction energy in the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
describes the non\hyp{}electrostatic repulsion
between two Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
isolated and unperturbed wavefunctions in an interacting
bi\hyp{}molecular complex.
The EXR EFP2 term was derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to second order with respect to the wavefunction overlap,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the benchmark EXR energy according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost of
the exchange and second\hyp{}order terms
is much less than the first\hyp{}order term,
which requires evaluaton of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

In this work, an alternative formulation for the first\hyp{}order
term is proposed based on the effective one\hyp{}electron potential (OEP)
operator technique developed in the preceeding contribution.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
It is shown that calculation of kinetic energy integrals can be avoided
by introducing OEP's.

\section{Results}

A theoretical foundation of the
first\hyp{}order repulsive term in the EFP2 model
is the perturbation theory of
Murrell et~al.~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
in which
%
\begin{multline} \label{e:rep.murrell-etal.S1}
    E^{\rm Rep}(S^{-1}) = -2\sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc}
               S_{ij} \Big\{
           W^A_{ij} 
         + W^B_{ji} \\
 + \sum_{k\in A}^{\rm Occ} \left[ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right] 
 + \sum_{l\in B}^{\rm Occ} \left[ 2\BraKet{ij}{ll} - \BraKet{il}{jl} \right]
                \Big\}
\end{multline}
%
where $W^A_{ij} \equiv \tBraKet{i}{\hat{v}^A_{\rm nuc}}{j}$,
$\hat{v}^A_{\rm nuc}$
is the nucleus\hyp{}electron potential operator
and
ERI is defined by
%
\begin{equation} \label{e:eri}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%
Now, note that the Coulomb and exchange integrals can be re\hyp{}cast as follows:
%
\begin{subequations}
 \begin{align}
 \BraKet{ij}{kk} &\equiv \sum_{\mu\in A} 
     C_{\mu i}^A \tBraKet{\mu}{\hat{v}_{kk}}{j} \;, \\
 \BraKet{ik}{jk} &\equiv \sum_{\mu\in A} 
     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
 \end{align}
\end{subequations}
%
%Similarly, the exchange integrals reduce to
%
%\begin{equation}
% \BraKet{ik}{jk} \equiv \sum_{\mu\in A} 
%     C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
%\end{equation}
%
In the above equations, the auxiliary potential operators are given by
%
\begin{equation}
  \hat{v}_{ik} \equiv \int d{\bf r} \Ket{{\bf r}} 
        \left[
        \int d{\bf r}' \frac{\phi_i^{*}({\bf r}') \phi_k({\bf r}')}{\vert {\bf r}' - {\bf r}\vert}
        \right] \Bra{{\bf r}} \;.
\end{equation}
%
ERI's can be effectively eliminated from Eq.~\eqref{e:rep.murrell-etal.S1} 
by using the following prescription~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
%
\begin{equation} \label{e:ft-reduction}
	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
\end{equation}
%
where ${\mathcal{F}}_t$ is a certain and well defined functional of ERI's.
Application of Eq.~\eqref{e:ft-reduction} enables to define
a joint OEP
operator constructed from nuclear, Coulomb and exchange parts 
as
%
\begin{equation}
 V^A_{ij} + 
 \sum_{k\in A}^{\rm Occ} 
  \left\{ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right\}
\equiv \sum_{\mu\in A} \tBraKet{\mu}{ 
\hat{v}^{A[\mu i]}_{\rm eff}
 }{j}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{A[\mu i]}_{\rm eff} \equiv C_{\mu i} \hat{v}^A_{\rm nuc} + 
 \sum_{k\in A}^{\rm Occ} \left[
 2C_{\mu i}^A \hat{v}^A_{kk} - C_{\mu k}^A \hat{v}^A_{ik}
 \right] \;.
\end{equation}
%
%The nuclear part can also be incorporated within the effective potential
%resulting in the following reduction
%
%
On the other hand, it immediately follows that
%
\begin{equation}
 \hat{v}^{A[\mu i]}_{\rm eff} \Ket{\mu} = 
  \sum_{k\in A}^{\rm Occ} \left\{
     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
  \right\} \;.
\end{equation}
%
For practical calculations, the right hand side of the above equation can be expanded 
in the auxiliary basis,
%
\begin{equation} \label{e:v-oep.rep}
  \sum_{k\in A}^{\rm Occ} \left\{
     2\hat{v}^A_{kk} \Ket{i} - \hat{v}^A_{ik} \Ket{k}
  \right\} \cong
  \sum_{\xi\in A}^{\rm DF} 
  V_{\xi i}^A \Ket{\xi} \;,
\end{equation}
%
where the matrix ${\bf V}^A$ can be considered as effective fragment parameters.
Doing the same operations on the twin operators associated with the molecule $B$
original theory of Murrell et. al from Eq.~\eqref{e:rep.murrell-etal.S1}
reduces to
%
\begin{equation} \label{e:rep.murrell-etal.S1.oep}
    E^{\rm Rep}(S^{-1}) \cong 
 -2\sum_{i\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}
               S_{ij} \Big\{
           \sum_{\xi \in A}^{\rm DF} V_{\xi i}^A S_{\xi j}
         + \sum_{\eta\in B}^{\rm DF} V_{\eta j}^B S_{\eta i}
                \Big\} \;,
\end{equation}
%
where 
effective fragment parameters
%the matrix elements $V_{\xi i}^A$ 
can be obtained 
from the extended density fitting\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} (EDF)
by
%
\begin{equation} \label{e:rep.murrell-etal.S1.oep-edf}
            V_{\xi i}^X = \sum_{\eta\in X}^{\rm DF} 
                          \sum_{\varepsilon,\zeta\in X}^{\rm RI}
%                         \sum_{\zeta\in A}^{\rm RI}
                          \left[ {\bf R}^{-1} \right]_{\xi\eta} R_{\eta\varepsilon} 
                          \left[ {\bf S}^{-1} \right]_{\varepsilon\zeta} 
                          \tBraKet{\zeta}{\hat{v}_{\rm eff}^{X[\zeta i]}}{i}
                \;.
\end{equation}
%
In the procedure, the required OEP matrix elements can be calculated from
%
\begin{equation} \label{e:rep.murrell-etal.S1.oep-v}
   \tBraKet{\zeta}{\hat{v}_{\rm eff}^{X[\zeta i]}}{i}
     = \sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} 
        + \sum_{\beta\gamma\delta\in X}^{\rm AO}
           \left\{ 
             2 C_{\beta i}^X D_{\gamma\delta}^X - C_{\gamma i}^X D_{\beta \delta}^X
           \right\}
           \BraKet{\zeta\beta}{\gamma\delta} \;,
\end{equation}
%
where ${\bf D}^X$ and ${\bf C}^X$ is the one\hyp{}particle AO density and
the LCAO\hyp{}MO matrices of isolated molecule $X$, respectively.
%where $\{\phi_\alpha\}$ is the test basis set, assumed here to be identical
%to the intermediate basis set necessary to perform the extended density fitting (EDF).
%
Eq.~\eqref{e:rep.murrell-etal.S1.oep} and Eq.~\eqref{e:rep-efp2.s1} have almost
the same form, with two exceptions in the new OEP formulation: 
(i) only overlap integrals, which are relatively computationally inexpensive,
need to be evaluated, instead of additional kinetic energy integrals in the EFP2
formulation that are slightly more expensive;
(ii) overlap integrals need to be evaluated also between
auxiliary basis set. The smaller the size of the auxiliary basis, the less expensive
evaluation of the $E^{\rm Rep}(S^{-1})$ becomes. 
%\section{\label{s:5.results}Results and Discussion}

Three complexes: 
(i) (H$_2$O)$_2$, 
(ii) H$_2$O--HOCH$_3$, 
(iii) (DMSO)$_2$,
were chosen as model systems to analyze the asymptotic dependence 
of EXR energy, which is shown in Figure~XXX. 
The reference (zero\hyp{}displacement) geometries
were obtained as described in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} 
and the structures  
along with the translation direction are depicted
in the insets in Figure~XXX.
Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
Small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO were developed
to minimize the computational cost of evaluating the summation terms
from Eq.~\eqref{e:rep.murrell-etal.S1.oep}. For this purpose, the basis set optimization 
as described in the Appendix~A in Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
in conjunction with the basin hopping global optimization algorithm,  %cite
available in the SciPy Python module. %cite

To investigate the accuracy of the OEP\hyp{}based repulsion
term across
%perform statistical error analysis for 
a variety of
interacting systems, databases
of bi\hyp{}molecular complexes in the non\hyp{}covalent
interactions database NCB31 developed by the Truhlar's 
group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
interactions in proteins,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
as implemented in the {\sc Psi4}
program,\cite{Psi4.JCTC.2017}
were utilized as sources of equilibrium structures. 
In Figure~YYY we show statistical error analysis...

%In particular, the subsets from the NCB31 database were separately considered:
%the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
%and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}

\section{\label{s:6.conclusions}Summary and a few concluding remarks}

Therefore, the following model is proposed for the
total exchange\hyp{}repulsion energy in the EFP2 model: 
(i) $E^{\rm Ex}$ and $E^{\rm Rep}(S^{-2})$ are evaluated
from the original EFP2 formulae of Jensen and Gordon, i.e.,
from Eqs.~\eqref{e:exc-efp2} and \eqref{e:rep-efp2.s2}, respectively;
(ii) $E^{\rm Rep}(S^{-1})$ term is evaluated from Eq.~\eqref{e:rep.murrell-etal.S1.oep}
and assuming a small auxiliary basis sets, optimized for each
EFP2 fragment separately.

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

%%

% -----------------------
\bibliography{references}
% -----------------------

\end{document}

### 2019-12-10 papers/3_oep_technique.Pauli/paper-r0.tex
Figures added to paper II

@@
 evaluation of the $E^{\rm Rep}(S^{-1})$ becomes. 
+The alternative EFP2 formulation %with OEP\hyp{}based first\hyp{}order term
+is therefore still given by Eq.~\eqref{e:exr-efp2}
+%
+%\begin{equation} \label{e:exrep-new}
+%E^{\rm Ex-Rep} \approx E^{\rm Ex} + E^{\rm Rep}_{\rm OEP} + E^{\rm Rep}(S^{-2}) \;,
+%\end{equation}
+%
+but with $E^{\rm Rep}(S^{-1})$ replaced by OEP\hyp{}based
+first\hyp{}order term from Eq.~\eqref{e:rep.murrell-etal.S1.oep}.
+%with $E^{\rm Ex}$ given by Eq.~\eqref{e:exc-efp2}, 
+%$E^{\rm Rep}(S^{-1})$ given by Eq.~\eqref{e:rep.murrell-etal.S1.oep} and
+%$E^{\rm Rep}(S^{-2})$ given by Eq.~\eqref{e:rep-efp2.s2}.
 %\section{\label{s:5.results}Results and Discussion}
+
+We implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
+(EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as the repulsion term of Murrell et~al.
+from Eq.~\eqref{e:rep.murrell-etal.S1} (here abbreviated as S1-Ref) for validation of the EDF technique.
+Benchmark EXR energy were estimated from the density decomposition scheme 
+at the Hartree\hyp{}Fock level of theory (here referred as to the DDS/HF) 
+developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+which is equivalent to the exact EXR energy 
+in the intermolecular perturbation theory with exchange of Hayes and Stone
+at this level of theory.\cite{Hayes.Stone.MolPhys.1984} 
+To localize molecular orbitals, procedure that is necessary for the EFP2 calculations,
+the Boys method\cite{Boys.RevModPhys.1960} was used. Note here that, since
+Murrell et~al.'s theory is invariant with respect to unitary transformation
+of molecular orbitals, Eq.~\eqref{e:rep.murrell-etal.S1.oep} is also invariant and does not require
+orbital localization. 
+%in which
+%
+%\begin{equation}\label{e:dds-exr}
+% \Delta E^{\rm Ex-Rep}_{\rm DDS/HF} = 
+%\end{equation}
+%
 
@@
 from Eq.~\eqref{e:rep.murrell-etal.S1.oep}. For this purpose, the basis set optimization 
-as described in the Appendix~A in Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
+as described in Appendix~A to Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
 in conjunction with the basin hopping global optimization algorithm,  %cite
 available in the SciPy Python module. %cite
+
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[h]
+\caption[]
+{{\bf CPU timings in miliseconds of exchange\hyp{}repulsion single point 
+energy calculations via various methods\footnotemark[1]}
+}
+\label{t:5}
+\begin{ruledtabular}
+\begin{tabular}{llccccccccc}
+                      && \multicolumn{2}{c}{(H$_2$O)$_2$\footnotemark[2]} 
+                      && \multicolumn{2}{c}{H$_2$O--HOCH$_3$\footnotemark[2]} 
+                      && \multicolumn{2}{c}{(DMSO)$_2$\footnotemark[3]} \\
+\cline{3-4}
+\cline{6-7}
+\cline{9-10}
+Murrell et al.\footnotemark[4] && 3$\times 10^4$  & (10.0)  
+                               && 3$\times 10^4$  & (10.0)
+                               && 3$\times 10^4$  & (10.0) \\
+Hayes-Stone\footnotemark[5]    && 3$\times 10^4$  & (10.0)  
+                               && 3$\times 10^4$  & (10.0)
+                               && 3$\times 10^4$  & (10.0) \\
+EFP2                           && 3$\times 10^4$  & (10.0)  
+                               && 3$\times 10^4$  & (10.0)
+                               && 3$\times 10^4$  & (10.0) \\
+OEP/large\footnotemark[6]      && 3$\times 10^4$  & (10.0)  
+                               && 3$\times 10^4$  & (10.0)
+                               && 3$\times 10^4$  & (10.0) \\
+OEP/mini\footnotemark[6]       && 3$\times 10^4$  & (10.0)  
+                               && 3$\times 10^4$  & (10.0)
+                               && 3$\times 10^4$  & (10.0) \\
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor, calculations performed on 1 core. 
+Exchange\hyp{}repulsion energies are given in parentheses for reference (kcal/mol).
+See also the implementation details in the main text.}
+\footnotetext[2]{Primary basis set: 6-311++G(d,p).}
+\footnotetext[3]{Primary basis set: 6-31++G(d,p).}
+\footnotetext[4]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
+\footnotetext[5]{Reference~\cite{Hayes.Stone.MolPhys.1984}.}
+\footnotetext[6]{Eq.~\eqref{}. Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
+from Eq.~\eqref{e:rep.murrell-etal.S1.oep}
+are: `large' - aug-cc-pVDZ-jkfit for (H$_2$O)$_2$ and H$_2$O--HOCH$_3$ systems as well as
+aug-cc-pVDZ-ri for (DMSO)$_2$ system;
+`mini' - optimized small basis sets (see Supporting Information).}
+%
+\end{table}
+}
+%
 

### 2019-12-10 papers/3_oep_technique.Pauli/figures/water_water.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+pymol -p <<EOF
+load ../xyz/water-water.xyz
+set_view (\
+     0.323731661,    0.940398216,   -0.104131773,\
+     0.006585651,    0.107815512,    0.994144022,\
+     0.946119666,   -0.322521687,    0.028708948,\
+     0.000000000,    0.000000000,  -14.178204536,\
+    -0.009005010,   -0.003893316,    1.248221874,\
+    11.178204536,   17.178203583,  -20.000000000 )
+
+set stick_radius, 0.06
+set sphere_scale, 0.15
+show sticks
+show spheres
+set transparency, 0.9
+set opaque_background, off
+set ray_opaque_background, off
+ray 4000, 4000
+png water_water, dpi=1200
+EOF
+

### 2019-12-11 papers/3_oep_technique.Pauli/paper-r0.tex
Timing table added

@@
 
+\author{Robert W. G{\'o}ra}
 \author{Wojciech Bartkowiak}
@@
 \renewcommand{\arraystretch}{1.4}
-\begin{table}[h]
+\begin{table*}[h]
 \caption[]
@@
 \cline{9-10}
-Murrell et al.\footnotemark[4] && 3$\times 10^4$  & (10.0)  
-                               && 3$\times 10^4$  & (10.0)
-                               && 3$\times 10^4$  & (10.0) \\
-Hayes-Stone\footnotemark[5]    && 3$\times 10^4$  & (10.0)  
-                               && 3$\times 10^4$  & (10.0)
-                               && 3$\times 10^4$  & (10.0) \\
-EFP2                           && 3$\times 10^4$  & (10.0)  
-                               && 3$\times 10^4$  & (10.0)
-                               && 3$\times 10^4$  & (10.0) \\
-OEP/large\footnotemark[6]      && 3$\times 10^4$  & (10.0)  
-                               && 3$\times 10^4$  & (10.0)
-                               && 3$\times 10^4$  & (10.0) \\
-OEP/mini\footnotemark[6]       && 3$\times 10^4$  & (10.0)  
-                               && 3$\times 10^4$  & (10.0)
-                               && 3$\times 10^4$  & (10.0) \\
+Hayes \& Stone\footnotemark[4] && 1.16$\times 10^3$  & (4.81)  
+                               && 1.12$\times 10^4$  & (4.72)
+                               && 1.31$\times 10^5$  & (7.34) \\
+Murrell et al.\footnotemark[5] && 2.33$\times 10^2$  & (4.53)  
+                               && 3.87$\times 10^3$  & (4.44)
+                               && 2.79$\times 10^4$  & (6.84) \\
+EFP2                           && 0.496              & (4.45)  
+                               && 0.84               & (4.44)
+                               && 4.65   & (6.56) \\
+OEP/large\footnotemark[6]      && 0.884  & (4.83)  
+                               && 1.836  & (4.83)
+                               && 8.95   & (6.77) \\
+OEP/mini\footnotemark[6]       && 0.324  & (5.13)  
+                               && 0.566  & (4.81)
+                               && 3.00   & (7.41) \\
 \end{tabular}
@@
 \footnotetext[3]{Primary basis set: 6-31++G(d,p).}
-\footnotetext[4]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
-\footnotetext[5]{Reference~\cite{Hayes.Stone.MolPhys.1984}.}
+\footnotetext[4]{Reference~\cite{Hayes.Stone.MolPhys.1984}.}
+\footnotetext[5]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
 \footnotetext[6]{Eq.~\eqref{}. Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
@@
 %
-\end{table}
+\end{table*}
 }

### 2019-12-12 papers/3_oep_technique.Pauli/paper-r0.tex
Data figures and tables added. Discussion completed

--- 2019-12-09 papers/3_oep_technique.Pauli/paper-r0.tex
@@
 
+%
+\begin{figure}[t]
+\includegraphics[width=0.44\textwidth]{fig-1.pdf}
+\caption{\label{f:fig-1} {\bf Asymptotic dependence of the exchange\hyp{}repulsion interaction energy
+for selected bi\hyp{}molecular complexes.} 
+%The charge transfer energy estimates are plotted for 
+(a) water dimer, 
+(b) water\hyp{}methanol complex, and 
+(c) DMSO dimer in $S_i$ symmetry,
+were one molecule has been translated
+along the vector specified in the insets relative to initial geometry,
+optimized at HF/6-31+G(d,p) level of theory.
+The 
+total interaction energy
+is also shown for comparison in light grey color in this figure.
+Interaction energies were obtained by using the 6-311++G(d,p)
+primary basis set for (a) and (b), and 6-31++G(d,p) for (c).
+In OEP calculations, EDF-1 scheme combined with small optimized auxiliary basis sets
+(Supplementary Information) were used.
+} 
+\end{figure}
+%
+
 A theoretical foundation of the
@@
 %
-On the other hand, it immediately follows that
+{
+\renewcommand{\arraystretch}{1.4}
+\begin{table*}[t]
+\caption[]
+{{\bf CPU timings in miliseconds of exchange\hyp{}repulsion single point 
+energy calculations via various methods\footnotemark[1]}
+}
+\label{t:tab-1}
+\begin{ruledtabular}
+\begin{tabular}{llccccccccc}
+                      && \multicolumn{2}{c}{(H$_2$O)$_2$\footnotemark[2]} 
+                      && \multicolumn{2}{c}{H$_2$O--HOCH$_3$\footnotemark[2]} 
+                      && \multicolumn{2}{c}{(DMSO)$_2$\footnotemark[3]} \\
+\cline{3-4}
+\cline{6-7}
+\cline{9-10}
+Hayes \& Stone\footnotemark[4] && 1.16$\times 10^3$  & (4.81)  
+                               && 1.12$\times 10^4$  & (4.72)
+                               && 1.31$\times 10^5$  & (7.34) \\
+Murrell et al.\footnotemark[5] && 2.33$\times 10^2$  & (4.53)  
+                               && 3.87$\times 10^3$  & (4.44)
+                               && 2.79$\times 10^4$  & (6.84) \\
+EFP2                           && 0.496              & (4.45)  
+                               && 0.84               & (4.44)
+                               && 4.65   & (6.56) \\
+OEP/large\footnotemark[6]      && 0.884  & (4.83)  
+                               && 1.836  & (4.83)
+                               && 8.95   & (6.77) \\
+OEP/mini\footnotemark[6]       && 0.324  & (5.13)  
+                               && 0.566  & (4.81)
+                               && 3.00   & (7.41) \\
+\end{tabular}
+\end{ruledtabular}
+%
+\footnotetext[1]{1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor, calculations performed on 1 core. 
+Exchange\hyp{}repulsion energies are given in parentheses for reference (kcal/mol).
+Time profiling of the code performance was carried out as in Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}.
+See also the implementation details in the main text.}
+\footnotetext[2]{Primary basis set: 6-311++G(d,p).}
+\footnotetext[3]{Primary basis set: 6-31++G(d,p).}
+\footnotetext[4]{Reference~\cite{Hayes.Stone.MolPhys.1984}.}
+\footnotetext[5]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
+\footnotetext[6]{Eq.~\eqref{}. Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
+from Eq.~\eqref{e:rep.murrell-etal.S1.oep}
+are: `large' - aug-cc-pVDZ-jkfit for (H$_2$O)$_2$ and H$_2$O--HOCH$_3$ systems as well as
+aug-cc-pVDZ-ri for (DMSO)$_2$ system;
+`mini' - optimized small basis sets (see Supporting Information).}
+%
+\end{table*}
+}
+%
+\\ On the other hand, it immediately follows that
 %
@@
 %
-\begin{equation} \label{e:rep.murrell-etal.S1.oep-v}
+\begin{multline} \label{e:rep.murrell-etal.S1.oep-v}
    \tBraKet{\zeta}{\hat{v}_{\rm eff}^{X[\zeta i]}}{i}
-     = \sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} 
+     = \sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\
         + \sum_{\beta\gamma\delta\in X}^{\rm AO}
@@
            \BraKet{\zeta\beta}{\gamma\delta} \;,
-\end{equation}
+\end{multline}
 %
@@
 evaluation of the $E^{\rm Rep}(S^{-1})$ becomes. 
+The alternative EFP2 formulation %with OEP\hyp{}based first\hyp{}order term
+is therefore still given by Eq.~\eqref{e:exr-efp2}
+%
+%\begin{equation} \label{e:exrep-new}
+%E^{\rm Ex-Rep} \approx E^{\rm Ex} + E^{\rm Rep}_{\rm OEP} + E^{\rm Rep}(S^{-2}) \;,
+%\end{equation}
+%
+but with $E^{\rm Rep}(S^{-1})$ replaced by OEP\hyp{}based
+first\hyp{}order term from Eq.~\eqref{e:rep.murrell-etal.S1.oep}.
+%with $E^{\rm Ex}$ given by Eq.~\eqref{e:exc-efp2}, 
+%$E^{\rm Rep}(S^{-1})$ given by Eq.~\eqref{e:rep.murrell-etal.S1.oep} and
+%$E^{\rm Rep}(S^{-2})$ given by Eq.~\eqref{e:rep-efp2.s2}.
 %\section{\label{s:5.results}Results and Discussion}
 
+We implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
+(EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as 
+the approximate intermolecular perturbation theory with exchange of 
+Murrell et~al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965},
+%from Eq.~\eqref{e:rep.murrell-etal.S1} 
+the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984}
+and the density decomposition scheme 
+at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
+developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+for validation of the EDF technique.
+When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
+the DDS/HF model is
+equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
+which is the case in this work. Using DCBS also allows one to eliminate the basis set
+superimposition error (DCBS).
+To localize molecular orbitals, procedure that is necessary for the EFP2 calculations,
+the Boys method\cite{Boys.RevModPhys.1960} was used. Note here that, since
+Murrell et~al.'s theory is invariant with respect to unitary transformation
+of molecular orbitals, the OEP\hyp{}based expression 
+in Eq.~\eqref{e:rep.murrell-etal.S1.oep} is also invariant and does not require
+orbital localization. 
+%in which
+%
+%\begin{equation}\label{e:dds-exr}
+% \Delta E^{\rm Ex-Rep}_{\rm DDS/HF} = 
+%\end{equation}
+%
+%
+\begin{figure*}[t]
+\includegraphics[width=0.8\textwidth]{fig-2.pdf}
+\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of exchange\hyp{}repulsion energy
+across various bi\hyp{}molecular systems.} 
+(a) NCB31 
+database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
+of non\hyp{}covalent interactions
+and
+(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
+For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+was used.
+} 
+\end{figure*}
+%
+
 Three complexes: 
-(i) (H$_2$O)$_2$, 
-(ii) H$_2$O--HOCH$_3$, 
-(iii) (DMSO)$_2$,
+(H$_2$O)$_2$, 
+H$_2$O--HOCH$_3$, and
+(DMSO)$_2$,
 were chosen as model systems to analyze the asymptotic dependence 
-of EXR energy, which is shown in Figure~XXX. 
+of EXR energy, which is shown in Figure~\ref{f:fig-1}. 
 The reference (zero\hyp{}displacement) geometries
@@
 along with the translation direction are depicted
-in the insets in Figure~XXX.
+in the insets in Figure~\ref{f:fig-1}.
 Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
@@
 the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
-Small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO were developed
-to minimize the computational cost of evaluating the summation terms
-from Eq.~\eqref{e:rep.murrell-etal.S1.oep}. For this purpose, the basis set optimization 
-as described in the Appendix~A in Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
-in conjunction with the basin hopping global optimization algorithm,  %cite
-available in the SciPy Python module. %cite
+OEP and EFP2 models correctly describe the EXR energy at all separations
+for the studied model systems with very similar accuracy. 
+However,
+contrary to our previous application of the OEP technique\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019},
+where approximately 20\hyp{}fold speedups as compared to the EFP2 model were achieved
+for the evaluation of the charge\hyp{}transfer energy,
+the computational cost of the OEP EXR method is roughly 2 times higher than the EFP2 model
+when the usual density fitting auxiliary basis sets such as aug-cc-pVDZ-jkfit or aug-cc-pVDZ-ri are used
+(Table~\ref{t:tab-1}).
+However, application of our small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
+%developed
+developed via the basis set optimization~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
+in conjunction with the basin hopping global optimization algorithm  %cite
+as implemented in the SciPy Python library
+(see Supplementary Information),
+%to minimize the computational cost of evaluating the summation terms
+%from Eq.~\eqref{e:rep.murrell-etal.S1.oep},
+reduces costs appreciably and makes the OEP method on average 1.5 times faster than EFP2.
+For instance, calculation of the EXR energy in the DMSO dimer by using 6-31++G(d,p) primary
+basis set and 
+1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
+requires about 4.6~ms and 3.0~ms when EFP2 and OEP model is used, respectively.
+%Note that the benchmark Hayes \& Stone's calculations predict slightly larger repulsion energies
 
@@
 a variety of
-interacting systems, databases
-of bi\hyp{}molecular complexes in the non\hyp{}covalent
+interacting systems, range of
+bi\hyp{}molecular complexes from the non\hyp{}covalent
 interactions database NCB31 developed by the Truhlar's 
@@
 as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
-interactions in proteins,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+interactions in proteins of Sherrill group,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
 as implemented in the {\sc Psi4}
 program,\cite{Psi4.JCTC.2017}
-were utilized as sources of equilibrium structures. 
-In Figure~YYY we show statistical error analysis...
+were utilized, and the
+EXR energies were compared with the reference DDS/HF results (Figure~\ref{f:fig-2}).
+On average, OEP and EXR tend to underestimate the EXR energy by 0.7\% and 7.6\%
+in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
+respectively.
+This is mostly due to the lack of BSSE, that is accounted for
+only in the reference calculations with the DDS/HF model, and the effect is usually larger
+when smaller primary basis sets are used.
+The OEP model is of slightly better accuracy as the EFP2 model
+in systems sudied in this work, with 
+root mean square errors of OEP and EFP2 model being
+2.0 kcal/mol and 2.6 kcal/mol for NCB31/6-311++G(d,p) data set, 
+as well as
+0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set, respectively.
+It is an accidental fruitful cancellation of errors because
+the first\hyp{}order repulsive terms have much less differences with respect to each other
+(RMSE is 0.7\% only in all data sets).
+Nevertheless, together with the performance data of both models
+from Table~\ref{t:tab-1}, OEP model performs better than EFP2 as long as the minimal
+auxiliary basis sets are used.
 
@@
 
+In this work, the OEP technique of effective ERI elimination
+proposed by us previously was utilized to reduce the computational cost
+of evaluation of the EXR energy in the EFP2 model.
+Starting from the Murrell et~al.'s model, which is a foundation of the EFP2 EXR term,
+OEP\hyp{}based expression was derived which requires evaluation of only overlap
+OEI's, apart from additional kinetic energy OEI's required for EFP2. 
 Therefore, the following model is proposed for the

### 2019-12-12 papers/3_oep_technique.Pauli/do-r1.sh
SI added

@@ -1,0 +1,10 @@
+#!/bin/bash
+base=paper-r1
+si=supplementary-r1
+
+for i in $base $si
+do
+  pdflatex  ${i}.tex
+  bibtex ${i}.aux
+  #dvipdf ${i}.dvi
+done

### 2019-12-13 papers/3_oep_technique.Pauli/paper-r2.tex
RWG suggestions to Paper III

--- 2019-12-12 papers/3_oep_technique.Pauli/paper-r0.tex
@@
 \begin{abstract}
-Recently, it was shown that
-the effective one\hyp{}electron potentials (OEP) method
-of effective elimination of the electron repulsion integrals
-can be used to lower the computational cost
-of evaluation of the charge transfer interaction energy
+In Paper~I, the effective one\hyp{}electron
+potentials (OEP) method was introduced and demonstrated as an
+efficient approach
+to reduce the computational cost
+of evaluation of the charge-transfer interaction energy
 within the effective fragment potential method (EFP2)
-by the factor of 20, making it no longer a bottleneck
+by an average factor of 20, making it no longer a bottleneck
 in EFP2\hyp{}based simulations of complex systems.
-Here, OEP technique 
+Here, the OEP technique 
 is used to enhance computational efficiency in evaluating the
 exchange\hyp{}repulsion EFP2 interaction energy
-by incorporating the kinetic energy operator in the OEP
-via the extended density fitting in incomplete auxiliary basis.
-It is shown that the computational cost can be reduced by a factor
-of 1.5 and kinetic energy integrals are no longer necessary to evaluate.
+by redefining the first\hyp{}order repulsive term of 
+Murrell~et al. [Murrell et al., Proc. R. Soc. Lond. A 284, 566 (1965)] through
+the extended density fitting in incomplete auxiliary basis.
+In the proposed approach the evaluation kinetic energy integrals is no
+longer required and the computational cost can be reduced roughly by a factor
+of 1.5 as compared to the original EFP2 formulation.
 \end{abstract}
@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 describes the non\hyp{}electrostatic repulsion
-between two Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
-isolated and unperturbed wavefunctions in an interacting
+between two isolated and unperturbed Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
+wavefunctions in an interacting
 bi\hyp{}molecular complex.
@@
 of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
-up to second order with respect to the wavefunction overlap,
+up to the second order with respect to the wavefunction overlap,
 and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
@@
 is remarkably efficient and gives usually very accurate
-estimates of the benchmark EXR energy according to the symmetry 
+estimates of the reference EXR energies according to the symmetry 
 adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
-The computational cost of
+The computational cost to evaluate
 the exchange and second\hyp{}order terms
-is much less than the first\hyp{}order term,
-which requires evaluaton of not only the one\hyp{}electron overlap,
+is much smaller than that of the first\hyp{}order term,
+which requires evaluation of not only the one\hyp{}electron overlap,
 but also the kinetic
@@
 term is proposed based on the effective one\hyp{}electron potential (OEP)
-operator technique developed in the preceeding contribution.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
-It is shown that calculation of kinetic energy integrals can be avoided
-by introducing OEP's.
+operator technique developed in the preceding contribution.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
+It is shown that using this approach calculation of kinetic energy integrals can be avoided
+what leads to increased computational efficiency.
 
@@
 In OEP calculations, EDF-1 scheme combined with small optimized auxiliary basis sets
-(Supplementary Information) were used.
+(cf. Supplementary Information) were used.
 } 
@@
 and
-ERI is defined by
+ERIs are defined by
 %
@@
 %
-ERI's can be effectively eliminated from Eq.~\eqref{e:rep.murrell-etal.S1} 
+ERIs can be effectively eliminated from Eq.~\eqref{e:rep.murrell-etal.S1} 
 by using the following prescription~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
@@
 %
-where ${\mathcal{F}}_t$ is a certain and well defined functional of ERI's.
+where ${\mathcal{F}}_t$ is a certain and well defined functional of ERIs.
 Application of Eq.~\eqref{e:ft-reduction} enables to define
 a joint OEP
-operator constructed from nuclear, Coulomb and exchange parts 
-as
+operator constructed from nuclear, Coulomb and exchange
+contributions as
 %
@@
 \footnotetext[5]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
-\footnotetext[6]{Eq.~\eqref{}. Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
+\footnotetext[6]{Same as EFP2 but with $E^{\rm Rep}(S^{-1})$ replaced by formula in Eq.~\eqref{e:rep.murrell-etal.S1.oep}. 
+Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
 from Eq.~\eqref{e:rep.murrell-etal.S1.oep}
@@
 %
-where the matrix ${\bf V}^A$ can be considered as effective fragment parameters.
+where the matrix ${\bf V}^A$ can be considered as a set of effective fragment parameters.
 Doing the same operations on the twin operators associated with the molecule $B$
@@
 %
-where 
+where the
 effective fragment parameters
@@
 %
-In the procedure, the required OEP matrix elements can be calculated from
+In this procedure, the required OEP matrix elements can be calculated from
 %
@@
 %
-where ${\bf D}^X$ and ${\bf C}^X$ is the one\hyp{}particle AO density and
+where ${\bf D}^X$ and ${\bf C}^X$ are the one\hyp{}particle AO density and
 the LCAO\hyp{}MO matrices of isolated molecule $X$, respectively.
@@
 the same form, with two exceptions in the new OEP formulation: 
-(i) only overlap integrals, which are relatively computationally inexpensive,
-need to be evaluated, instead of additional kinetic energy integrals in the EFP2
-formulation that are slightly more expensive;
+(i) only overlap integrals are needed, which are relatively easy to
+compute. Furthermore, there is no need to evaluate the kinetic
+energy integrals, as in the original EFP2
+formulation, which are computationally slightly more demanding;
 (ii) overlap integrals need to be evaluated also between
@@
 
-We implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
+In our in\hyp{}house plug-in to the {\sc Psi4}
+quantum chemistry
+program,\cite{Psi4.JCTC.2017}
+we implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
 (EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as 
@@
 %from Eq.~\eqref{e:rep.murrell-etal.S1} 
-the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984}
-and the density decomposition scheme 
-at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
-developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
-for validation of the EDF technique.
-When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
-the DDS/HF model is
-equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
-which is the case in this work. Using DCBS also allows one to eliminate the basis set
-superimposition error (DCBS).
-To localize molecular orbitals, procedure that is necessary for the EFP2 calculations,
-the Boys method\cite{Boys.RevModPhys.1960} was used. Note here that, since
-Murrell et~al.'s theory is invariant with respect to unitary transformation
+the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
+here referred to as the HS model.
+The latter was
+implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
+to eliminate the basis set
+superposition error (DCBS) in the benchmark calculations.
+%and the density decomposition scheme 
+%at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
+%developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+%for validation of the EDF technique.
+%When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
+%the DDS/HF model is
+%equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
+%which is the case in this work. 
+%Using DCBS also allows one to eliminate the basis set
+%superimposition error (DCBS).
+The Boys method\cite{Boys.RevModPhys.1960} was used to localize
+molecular orbitals---the same as in the original EFP2 formulation,
+Note here that, since
+Murrell et~al. theory is invariant with respect to unitary transformation
 of molecular orbitals, the OEP\hyp{}based expression 
@@
 and the structures  
-along with the translation direction are depicted
+along the translation direction are depicted
 in the insets in Figure~\ref{f:fig-1}.
@@
 for the evaluation of the charge\hyp{}transfer energy,
-the computational cost of the OEP EXR method is roughly 2 times higher than the EFP2 model
+the CPU timings of EXR calculations in the OEP method are comparable but roughly 2
+times higher than that of the EFP2 model
 when the usual density fitting auxiliary basis sets such as aug-cc-pVDZ-jkfit or aug-cc-pVDZ-ri are used
 (Table~\ref{t:tab-1}).
-However, application of our small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
+Computational cost of $\Delta E^{\rm Rep}(S^{-1})$ in EFP2 model
+is approximately $s(on^2 + o^2n) + t(on^2 + o^2n) + 2o^3$, whereas in OEP model
+is $s(on^2 + o^2n) + 2ao^2$, both of which have comparable magnitudes. 
+Here, $s$ and $t$ denote the relative costs of evaluation of the overlap
+and kinetic-energy one electron integrals (OEIs), whereas $o$, $n$ and $a$ is
+the number of LMOs, primary AOs and auxiliary AOs, respectively.
+It is clear that the critical parameter is the number of auxiliary basis functions $a$,
+which should be comparable with the number of occupied orbitals $o$ 
+as $t$ is usually only 2--3 times larger than $s$.
+To investigate this issue more thoroughly we developed small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
 %developed
-developed via the basis set optimization~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
-in conjunction with the basin hopping global optimization algorithm  %cite
-as implemented in the SciPy Python library
-(see Supplementary Information),
+via the basis set optimization method described in Appendix~A to Paper I
+in conjunction with the basin hopping global optimization algorithm\cite{Wales.EnergyLandscapes.2003,
+Wales.Scheraga.Science.1999,
+Wales.Doye.JPCA.1997,Li.Scheraga.PNAS.1987}
+as implemented in the SciPy Python library\cite{SciPy.2019}
+(see Supplementary Information).
 %to minimize the computational cost of evaluating the summation terms
 %from Eq.~\eqref{e:rep.murrell-etal.S1.oep},
-reduces costs appreciably and makes the OEP method on average 1.5 times faster than EFP2.
+Applying such basis sets
+reduces computational effort appreciably and makes the OEP method on average 1.5 times faster than EFP2.
 For instance, calculation of the EXR energy in the DMSO dimer by using 6-31++G(d,p) primary
 basis set and 
-1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
-requires about 4.6~ms and 3.0~ms when EFP2 and OEP model is used, respectively.
+1.2~GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
+requires about 4.6~ms and 3.0~ms when EFP2 and OEP models are used, respectively.
 %Note that the benchmark Hayes \& Stone's calculations predict slightly larger repulsion energies
@@
 a variety of
-interacting systems, range of
+interacting systems, a selection of
 bi\hyp{}molecular complexes from the non\hyp{}covalent
@@
 program,\cite{Psi4.JCTC.2017}
-were utilized, and the
-EXR energies were compared with the reference DDS/HF results (Figure~\ref{f:fig-2}).
-On average, OEP and EXR tend to underestimate the EXR energy by 0.7\% and 7.6\%
+were utilized. The computed
+EXR energies were compared with the reference HS results (Figure~\ref{f:fig-2}).
+On average, OEP and EFP2 methods tend to underestimate the EXR energy by 0.7\% and 7.6\%
 in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
 respectively.
-This is mostly due to the lack of BSSE, that is accounted for
-only in the reference calculations with the DDS/HF model, and the effect is usually larger
+This is mostly due to BSSE, that is corrected for
+only in the reference calculations with the HS model, and the effect is usually larger
 when smaller primary basis sets are used.
-The OEP model is of slightly better accuracy as the EFP2 model
-in systems sudied in this work, with 
-root mean square errors of OEP and EFP2 model being
-2.0 kcal/mol and 2.6 kcal/mol for NCB31/6-311++G(d,p) data set, 
-as well as
-0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set, respectively.
-It is an accidental fruitful cancellation of errors because
-the first\hyp{}order repulsive terms have much less differences with respect to each other
-(RMSE is 0.7\% only in all data sets).
+The OEP model performs slightly better when compared to the EFP2
+method for systems studied in this work, with 
+root-mean-square errors of OEP and EFP2 models of
+2.0 kcal/mol and 2.6 kcal/mol, respectively, for NCB31/6-311++G(d,p) data set, 
+or
+0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set.
+The slightly better performance of OEP is most likely due to
+fruitful cancellation of errors, since
+the mutual differences between EFP2 and OEP estimations of 
+$\Delta E^{\rm Rep}(S^{-1})$ are on the average much smaller
+(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
+%
+% RWG: Aren't they increasing with the magnitude of interaction
+% energy? For instance what is the difference in magnitudes for the
+% last complex in NCB31 of 48kcal/mol and the last three complexes
+% in BBI above 15 kcal/mol. 
+% 
+% From what I see the OEP is consistently closer to the reference
+% values - thus I think it is a bit more than fruitful cancellation.
+% Can you find one example when EFP2 is closer to the reference?
+% Have you tried to compute this in other (larger) basis sets? If
+% this trend would be consistent in larger basis set then it is
+% pretty obvious that it is not just fruitful cancellation.
+%
 Nevertheless, together with the performance data of both models
-from Table~\ref{t:tab-1}, OEP model performs better than EFP2 as long as the minimal
-auxiliary basis sets are used.
+from Table~\ref{t:tab-1}, OEP model consistently outperforms the
+EFP2 approach in terms of accuracy and also performance, provided
+that the minimal auxiliary basis set is used.
 
@@
 
-In this work, the OEP technique of effective ERI elimination
-proposed by us previously was utilized to reduce the computational cost
-of evaluation of the EXR energy in the EFP2 model.
-Starting from the Murrell et~al.'s model, which is a foundation of the EFP2 EXR term,
-OEP\hyp{}based expression was derived which requires evaluation of only overlap
-OEI's, apart from additional kinetic energy OEI's required for EFP2. 
-Therefore, the following model is proposed for the
+In this work, the effective one-electron potential (OEP) technique,
+proposed in the preceding paper for the
+effective elimination of electron
+repulsion integrals in ab initio calculations,
+was utilized to reduce the computational cost
+of evaluation of the exchange repulsion (EXR) energy in the EFP2 model.
+Starting from the first-order formula for EXR in Murrell et~al. perturbation theory, 
+being a foundation of the EFP2 EXR term,
+OEP\hyp{}based expression was derived which requires evaluation of
+only the overlap one-electron integrals, in contrast to EFP2
+approach which requires also the kinetic-energy one-electron
+integrals. 
+The reported results indicate that a following model should be more
+efficient for calculations of the 
 total exchange\hyp{}repulsion energy in the EFP2 model: 

### 2019-12-21 papers/2_oep_technique.CT/paper-r4.tex
Revision 4 of Paper 2 added

--- 2019-12-09 papers/2_oep_technique.CT/paper-r3.tex
@@
 \begin{document}
-\preprint{AIP/123-OEP}
-
-\title{Ab Initio Effective One-Electron Potential Operators:
+%\preprint{AIP/123-OEP}
+
+\title{Ab Initio Effective One-Electron Potential Operators. I.
 Applications for Charge-Transfer Energy in Effective Fragment Potentials}
@@
 molecular dynamics simulations 
-due to prohibitively high computational cost of evaluating the
+due to relatively high computational cost of evaluating the
 charge transfer (CT) effects as compared to other effects.
@@
 \begin{equation}\label{e:u-efp2}
- U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;,
-\end{equation}
-%
-where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
+ E^{\rm EFP2} \approx E^{\rm Coul} + E^{\rm Ex-Rep} + E^{\rm Ind} + E^{\rm Disp} + E^{\rm CT} \;,
+\end{equation}
+%
+where $E^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
 distributions of the monomers, treated by the distributed multipole approximation
 with damping to account for the charge\hyp{}penetration effects,\cite{Slipchenko.Gordon.JCC.2007}
-$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
+$E^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
 principle,\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998} 
 that could be roughly understood as the Heitler\hyp{}London interaction
-energy\cite{Chalasinski.Gutowski.MolPhys.1985} with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
+energy\cite{Chalasinski.Gutowski.MolPhys.1985} with $E^{\rm Coul}$ subtracted, $E^{\rm Ind}$ and $E^{\rm Disp}$
 are the induction and dispersion energies obtained from the distributed polarizability
 approximation,\cite{Li.Netzloff.Gordon.JCP.2006,Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014} 
-and finally $U^{\rm CT}$ is the CT energy,\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013} 
+and finally $E^{\rm CT}$ is the CT energy,\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013} 
 the focus of this work.
@@
 \begin{equation} \label{e:d-spectral}
-	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
+	\rho^{\rm eff}_{\rm el}({\bf r}) = -\sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
 	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
@@
 	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
-	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
+	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} -
         \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
@@
 
-Three classes of ERI's can be recognized 
+Three unique classes of ERI's can be recognized 
 based on the basis function partitioning scheme
 within the system composed of two molecules (shall be $A$ and $B$
-throughout the course of this work),
-in which it is possible to define an OEP. 
+throughout the course of this work). 
 They are as follows:
@@
                    \sum_{w\in A} \left\{
-                                 T^{(0)}_{wu} \hat{q}_w 
-                          - {\bf T}^{(1)}_{wu} \cdot \hat{\BM\upmu}_w 
-            + \frac{1}{3}   {\bf T}^{(2)}_{wu} :     \hat{\BM\Theta}_w 
+                                 T^{(0)}_{we} \hat{q}_w 
+                          - {\bf T}^{(1)}_{we} \cdot \hat{\BM\upmu}_w 
+            + \frac{1}{3}   {\bf T}^{(2)}_{we} :     \hat{\BM\Theta}_w 
                     - \ldots  \right\}
-              \right| l \right> \;.
+              \right| l \right>_{{\bf r}_e} \;.
 \end{multline}
@@
 are the quantum operators of effective distributed monopole (charge),
-dipole moment and quadrupole moment, respectively, centered on the $w$th site, 
-whereas ${{\bf T}^{(r)}_{wu}}$ are the so called interaction tensors of rank $r$ 
-that can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
+dipole moment and quadrupole moment, respectively, centered on the $w$th site at ${\bf r}_w$, 
+whereas ${{\bf T}^{(d)}_{we}}$ are the so called interaction tensors of rank $d$ between ${\bf r}_w$ and ${\bf r}_e$, with the latter being the electronic coordinate.
+Explicit forms of interaction tensors can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
 Note that $\Ket{j}$ and $\Ket{l}$ can belong to either molecule.
@@
 \caption[Types of matrix elements with OEP operators]
-{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
+{{\bf Types of matrix elements with OEP operators}
 }
@@
 ERI class           &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
-DF\footnotemark[2]/RI\footnotemark[3] Form    
-&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
-&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
-DMTP\footnotemark[4] Form                     
+DF\footnotemark[1]/RI\footnotemark[2] Form    
+&& $\sum_{\xi\in A}^{\rm } v^A_{i\xi} S^{AB}_{\xi j} $  
+&& $\sum_{\xi\zeta\in A}^{\rm } S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
+DMTP\footnotemark[3] Form                     
 && --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
@@
 %
-\footnotetext[1]{Test footnotemark.}
-\footnotetext[2]{Density Fitting}
-\footnotetext[3]{Resolution of Identity}
-\footnotetext[4]{Distributed Multipole Expansion}
+\footnotetext[1]{Density Fitting}
+\footnotetext[2]{Resolution of Identity}
+\footnotetext[3]{Distributed Multipole Expansion}
 %
@@
 \begin{equation} \label{e:coulomb-energy:general}
- \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
+  E_{\rm eff} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
  {\vert {\bf r}_1 - {\bf r}_2 \vert} 
@@
 \begin{multline} \label{e:coulomb-energy:dmtp}
-  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
+  E_{\rm eff} \approx
  \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
@@
 one can use a procedure similar to
-the typical density fitting or resolution of identity, both of which are nowadays widely used 
+the typical density fitting (DF) or resolution of identity (RI), both of which are nowadays widely used 
 to compute electron\hyp{}repulsion integrals (ERI's) more efficiently,
-and reduce computational cost of post\hyp{}Hartree\hyp{}Fock methods. Density fitting was also
+and reduce computational cost of post\hyp{}Hartree\hyp{}Fock methods.\cite{Hesselmann.Jansen.Schutz.JCP.2005} Density fitting was also
 applied to design \emph{ab initio} force fields.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,Cisneros.Andres.Piquemal.Darden.JCP.2005}
@@
 In the equations above, 
-the `RI' symbol denotes a certain auxiliary basis set that fulfills the resolution of identity. 
+the `RI' symbol denotes a certain auxiliary basis set that fulfills such resolution of identity. 
 %In a special case when the basis set is orthogonal (e.g., molecular orbitals)
@@
 %\end{equation}
-Such a general expansion can be obtained by 
+The above general expansion can be also obtained by 
 utilizing the density fitting
@@
 is the projection of the state vector $\hat{v}^{A} \Ket{i}$
-onto the complete basis $\{ \varphi_\xi \}$.
+onto the nearly\hyp{}complete basis $\{ \varphi_\xi \}$.
 Let $Z_i[{\bf V}^A]$ be the least\hyp{}squares objective function 
@@
 \begin{equation} \label{e:a-coeff-compl}
- a^{(i)}_\eta = \lambda \sum_{x} W_{\eta i}^{(x)} + 
- \sum_{\alpha\beta} D^A_{\alpha\beta} 
+ a^{(i)}_\eta = \lambda \sum_{x} W_{\eta i}^{(x)} - 
+ \sum_{\alpha\beta} P^A_{\alpha\beta} 
   \BraKet{\alpha\beta}{\eta i} \;.
@@
 matrix elements of the OEP operator in an incomplete auxiliary space. This can be achieved by minimizing 
-the following objective function
+the following objective function\cite{Cisneros.Andres.Piquemal.Darden.JCP.2005,Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006}
 %
@@
 Obtaining all the necessery integrals of this kind directly 
-by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 
+by performing integrations of Eq.~\eqref{e:triple-integral} is very costly 
 %and impractical 
@@
 which can be easily evaluated by noting that
-$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}.
-Thus, the matrix elements of the OEP operator can be found
+$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}
+--- thus, the matrix elements of the OEP operator can be found
 by using Eq.~\eqref{e:v-incompl}.
@@
 and the auxiliary and intermediate basis sets depend solely on one unperturbed molecule $A$, the matrix elements
-$V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.
+$V^A_{i\xi}$ can be calculated just once and stored as effective fragment parameters.
 
@@
 the application of the OEP technique of ERI elimination is
-further discussed for the evaluation of the CT/HF0 energy in the next sections.
+further discussed for the evaluation of the CT/HF0 energy in the following sections.
 
@@
 \begin{equation} \label{e:ct-hf0}
- U^{\rm CT} =  E^{A\rightarrow B} + E^{B\rightarrow A}\;.
+ E^{\rm CT} =  E^{A\rightarrow B} + E^{B\rightarrow A}\;.
 \end{equation}
@@
   \sum_{n\in B}^{\rm Vir}
-   \frac{ \left| V_{in}^{A\rightarrow B} \right|^2}{\varepsilon_{i} - T_{nn}} \;,
+   \frac{ \left| U_{in}^{A\rightarrow B} \right|^2}{\varepsilon_{i} - T_{nn}} \;,
 \end{equation}
@@
 \begin{multline} \label{e:u-efp2-ct-v2}
- \left| V_{in}^{A\rightarrow B} \right|^2 \approx
- \frac{U_{in}}{1-\sum_{m\in A}^{\rm All} S_{mn}^2}
+ \left| U_{in}^{A\rightarrow B} \right|^2 \approx
+ \frac{u_{in}}{1-\sum_{m\in A}^{\rm All} S_{mn}^2}
  \Bigg\{
-   U_{in} \\ + \sum_{j\in B}^{\rm Occ} S_{ij} 
+   u_{in} \\ + \sum_{j\in B}^{\rm Occ} S_{ij} 
    \left( 
@@
 \begin{equation} \label{e:uin-efp2}
- U_{in} \equiv V_{in}^{{\rm EF},B} - \sum_{m\in A}^{\rm All} V_{im}^{{\rm EF},B} S_{mn} \;,
+ u_{in} \equiv U_{in}^{{\rm EF},B} - \sum_{m\in A}^{\rm All} U_{im}^{{\rm EF},B} S_{mn} \;,
 \end{equation}
@@
 in which the summations extend over occupied (denoted by `Occ'), virtual (denoted by `Vir') 
-or both (denoted by `All') canonical HF MO's. The effective potential 
-matrices are defined by
+or both (denoted by `All') canonical HF MO's. The effective potential energy
+matrix elements are defined by
 %
 \begin{equation} \label{e:v-efp2-eff}
- V_{in}^{{\rm EF},B} \equiv 
- \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} \;,
+ U_{in}^{{\rm EF},B} \equiv 
+ -\tBraKet{i}{\hat{v}^B_{\rm tot} }{n} \;,
 \end{equation}
@@
 constant parameters, whereas the overlap $S_{nm}$, kinetic energy $T_{mn}$ and effective
-one\hyp{}electron electrostatic potential $V_{in}^{{\rm EF}}$ matrix elements
+one\hyp{}electron electrostatic potential $U_{in}^{{\rm EF}}$ matrix elements
 are all certain types of the one\hyp{}electron integrals, orders of magnitude cheaper to
@@
 evaluating Eq.~\eqref{e:u-efp2-ct} is still considerable in cost because typically large basis sets
-need to be used for generating the EFP2 parameters. In effect, calculation of $V_{in}^{{\rm EF}}$
+need to be used for generating the EFP2 parameters. In effect, calculation of $U_{in}^{{\rm EF}}$
 is much more expensive as compared to other types of one\hyp{}electron integrals
@@
 Unfortunately, this was still insufficient to overcome
-the high cost of evaluation of the potential integrals, making the CT term still the most time\hyp{}consuming
+the high cost of evaluation of the potential energy integrals, making the CT term still the most time\hyp{}consuming
 to evaluate.
@@
  E^{A\rightarrow B} = 2 \sum_{i\in A}^{\rm Occ} \sum_{n\in B}^{\rm Vir} 
-  \frac{\vert V^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n} \;,
+  \frac{\vert U^{A\rightarrow B}_{in} \vert^2 }{\varepsilon_i - \varepsilon_n} \;,
 \end{equation}
@@
 where the coupling constant is given by Otto and Ladik,\cite{Otto.Ladik.ChemPhys.1975}
-here referred to as the OL method,
+here referred to as the OL method, as
 %
 \begin{multline} \label{e:ct-murell-etal.vin}
- V^{A\rightarrow B}_{in} = 
-        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
+ U^{A\rightarrow B}_{in} = 
+      - \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
-      - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+      + \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} \\
+      + \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
@@
 %
-In the above expression, the following effective one\hyp{}electron operators,
+In the above expression, the following effective one\hyp{}electron potential operators,
 %
@@
 were introduced without making any approximation to the original equation
-from Ref. 
-The goal of this work is to apply the technique sketched in Eq.~\eqref{e:ft-reduction}
-to eliminate ERI's from Eq.~\eqref{e:ct-murell-etal.vin} in terms of OEP's.
+from Ref.\cite{Otto.Ladik.ChemPhys.1975} 
+Note that ERI's in MO basis are necessary to evaluate all terms
+in Eq.~\eqref{e:ct-murell-etal.vin}, e.g., 
+$\tBraKet{i}{\hat{v}^B_{\rm tot} }{n} \equiv W^B_{nj} - 
+2\sum_{j\in B}\BraKet{in}{jj}$.
+Therefore, the goal of this work is to apply the technique sketched in Eq.~\eqref{e:ft-reduction}
+to effectively eliminate ERI's from Eq.~\eqref{e:ct-murell-etal.vin} in terms of OEP's.
 
@@
 regarding the type of ERI's that are required:
-(i) $\BraKet{AB}{BB}$ -- the first two terms;
-(ii) $\BraKet{AA}{BB}$ -- the third term and
-(iii) $\BraKet{BB}{AA}$ -- the two last terms. 
+(i) overlap\hyp{}like $\BraKet{AB}{BB}$ -- the first two terms;
+(ii) Coulomb\hyp{}like $\BraKet{AA}{BB}$ -- the third term and
+(iii)  Coulomb\hyp{}like $\BraKet{BB}{AA}$ -- the two last terms. 
 Note also that there are no exchange\hyp{}like terms needed in this case.
@@
  \BraKet{nj}{ij} =
- \Bra{i} \left[ \hat{v}^B_{nj} \Ket{j} \right]  \;,
+ \Bra{i} \left[ -\hat{v}^B_{nj} \Ket{j} \right]  \;,
 \end{equation}
@@
 \begin{equation} \label{e:v-oep.ct}
-\Bra{i} \left[ \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
+\Bra{i} \left[ -\hat{v}^B_{\rm tot} \Ket{n} + \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j} \right]
 \cong \Bra{i} \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta}
@@
 \begin{equation}
- \hat{v}^B_{\rm tot} \Ket{n} - \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
+-\hat{v}^B_{\rm tot} \Ket{n} + \sum_{j\in B}^{\rm Occ} \hat{v}^B_{nj} \Ket{j}
  \cong \sum_{\varepsilon\in B}^{\rm RI} H_{n\varepsilon}^B \Ket{\varepsilon} \;,
@@
 \begin{align} \label{e:v-oep-group-1-final:a-vector}
- a^{(n)}_\zeta &= \tBraKet{\zeta}{\hat{v}^B_{\rm tot}}{n}
-      - \sum_{j\in B}^{\rm Occ} \tBraKet{\zeta}{\hat{v}^B_{nj}}{j} \nonumber \\
- &= \sum_{y\in B} W^{(y)}_{\zeta n} 
+ a^{(n)}_\zeta &= -\tBraKet{\zeta}{\hat{v}^B_{\rm tot}}{n}
+      + \sum_{j\in B}^{\rm Occ} \tBraKet{\zeta}{\hat{v}^B_{nj}}{j} \nonumber \\
+ &=-\sum_{y\in B} W^{(y)}_{\zeta n} 
   + \sum_{j\in B}^{\rm Occ} 
@@
 \begin{equation} \label{e:ct.group-i.final}
-        \tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
+        -\tBraKet{i}{\hat{v}^B_{\rm tot} }{n} 
       - \sum_{j\in B}^{\rm Occ} \BraKet{nj}{ij} 
@@
 ${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$ with $\chi_i({\bf r})$ being localized,
-this group can be written as
+this group can be approximated by
 %
 \begin{equation} \label{e:ct.group-2}
-      - \sum_{k\in A}^{\rm Occ} S_{nk}  \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
-\end{equation}
-%
-Note that, in general, $\rho_{ki}^A$ is complex whereas $\rho^B_{\rm tot}$ is always real.
-Since the interaction energy must also be real, it must hold that
+       \sum_{k\in A}^{\rm Occ} S_{nk}  \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \approx  S_{ni} \rho_{ii}^A \odot \rho^B_{\rm tot} \;.
+\end{equation}
+%
+Here, it was assumed that
+$\lvert\rho_{ik} ({\bf r}) \rvert \ll \rho_{ii} ({\bf r}) $ 
+for $i\ne k$ in most locations in the case of LMO's,
+which allows one to conjecture that
 %
 \begin{equation} \label{e:ct.group-2-d}
- \rho_{ki}^A \odot \rho^B_{\rm tot} = \delta_{ik} \rho_{ii}^A \odot \rho^B_{\rm tot} \;.
-\end{equation}
-%
-Consider now a situation in which the localized occupied MO's are used.
-In this case, the distributed multipole expansion
-of the above becomes
+ \rho_{ki}^A \odot \rho^B_{\rm tot} \approx \delta_{ik} \rho_{ii}^A \odot \rho^B_{\rm tot} \;.
+\end{equation}
+%
+Now, the DMTP expansion 
+of the interaction energy in the right hand 
+side of Eq.~\eqref{e:ct.group-2-d} can be given by
+%Consider now a situation in which the occupied LMO's are used.
+%In this case, the distributed multipole expansion
+%of the above becomes
 %
@@
  +
- 2\sum_{j\in B}^{\rm LC}
+ 2\sum_{j\in B}^{\rm Occ}
   \frac{q_j}{\vert {\bf r}_j - {\bf r}_i \vert } 
@@
 respective LMO charge centroids vanish.\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
-This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows:
+This means that Eq.~\eqref{e:ct.group-2} can be finally given as follows:
 %
 \begin{equation} \label{e:ct.group-ii.final}
-      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
- \approx S_{ni}  \left[
+       \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
+ \approx - S_{ni}  \left[
   %\sum_{i\in A}^{\rm LC} 
@@
  -
- \sum_{j\in B}^{\rm LC}
+ \sum_{j\in B}^{\rm Occ}
   \frac{2}{r_{ji}} 
@@
 Therefore, only overlap integrals and relative distances between
-atomic and LC's are needed, which leads to a great reduction of the
+atomic and LMO centroid positions are needed, which leads to a great reduction of the
 calculation cost,
@@
 We shall refer to this approximation as to
-the localized overlap approximation (LOA). Note that in order for this approximation to be valid,
+the localized overlap approximation (LOA) resulting
+in similar expressions to the ones obtained by Jensen and Gordon
+in their exchange\hyp{}repulsion interaction energy EFP2 model
+(see Eq.~(39) in Ref.\cite{Jensen.Gordon.MolPhys.1996}). Note that in order for this approximation to be valid,
 occupied molecular orbitals need to be spatially localized.
@@
 \begin{multline} \label{e:ct.group-iii}
-       - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
+        \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{i} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
         S_{kj} \left( 1 - \delta_{ik} \right)
-        \BraKet{nj}{ik} \\ = -
+        \BraKet{nj}{ik} \\ = 
  \sum_{k\in A}^{\rm Occ} 
@@
   }_{\hat{v}^{A,{\rm eff}}_{ik}}
- }{n}  \\ \cong -
+ }{n}  \\ \cong 
 %
@@
 \begin{equation} \label{e:oed-group-iii}
- D^{B,(nj)}_{\beta\delta} = C_{\beta n} C_{\delta j} \;.
-\end{equation}
-%
-In this work, the effective charges were computed via the Mulliken method as
+ P^{B,(nj)}_{\beta\delta} = C_{\beta n} C_{\delta j} \;.
+\end{equation}
+%
+In this work, the effective charges were defined via the Mulliken method as
 discussed in Appendix~\ref{e:a1-v.oep-camm} with $\lambda=0$. 
-Note now that both effective densities in 
-$\rho^B_{jn} \odot \rho^{A,{\rm eff}}_{ik}$ are generally complex and one cannot apply the LOA
-for $\rho^{A,{\rm eff}}_{ik}$. 
-In the LMO picture however, it can be shown that
-$\lvert\rho_{ik} ({\bf r}) \rvert \ll \rho_{ii} ({\bf r}) $ 
-for $i\ne k$ in most locations. Therefore,
-the LOA is still applicable in the LMO picture
-and the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
+By applying the LOA in LMO picture for the $\rho_{ik}^A$ density
+the effective potential from Eq.~\eqref{e:ct.group-iii.oep} 
 simplifies to
@@
 \begin{multline} \label{e:ct.group-iii.final.b}
-        - \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
+         \sum_{j\in B}^{\rm Occ} S_{ij} \tBraKet{j}{\hat{v}^A_{\rm tot} }{n}  
      + \sum_{k\in A}^{\rm Occ} \sum_{j\in B}^{\rm Occ}  
@@
         \BraKet{nj}{ik}  \\
- \approx  -
+ \approx  
  \sum_{j\in B}^{\rm Occ} S_{ij}
@@
 \begin{equation} \label{e:ct-murell-etal.vin.oep}
-V_{in}^{A\rightarrow B} 
+U_{in}^{A\rightarrow B} 
        \approx 
@@
  G_{1;in}^{A\rightarrow B} &\equiv \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;, \\
- G_{2;in}^{A\rightarrow B} &\equiv S_{ni} u_i^{BA} \;, \\
+ G_{2;in}^{A\rightarrow B} &\equiv -S_{ni} u_i^{BA} \;, \\
  G_{3;in}^{A\rightarrow B} &\equiv
- -\sum_{j   \in B}^{\rm Occ} S_{ij}
+  \sum_{j   \in B}^{\rm Occ} S_{ij}
   \sum_{y   \in B}^{\rm At} q_y^{B,(nj)} w_{yi}^{BA}
@@
    \sum_{i'\in A}^{\rm LMO} L_{ii'}^A
-   V_{i'n}^{A\rightarrow B}
+   U_{i'n}^{A\rightarrow B}
  \rvert^2 }{\varepsilon_i - \varepsilon_n} \;,
@@
 \begin{equation}\label{e:ct-pthf-ref}
- \Delta E^{\rm CT}_{\rm HF0} \cong \Delta E^{\rm Pol}_{\rm DDS/HF} 
- - \Delta E^{\rm Ind} \;,
-\end{equation}
-%
-where $\Delta E^{\rm Pol}_{\rm DDS/HF}$ is the Hartree\hyp{}Fock polarization energy
+ E^{\rm CT}_{\rm HF0} \cong  E^{\rm Pol}_{\rm DDS/HF} 
+ -  E^{\rm Ind} \;,
+\end{equation}
+%
+where $E^{\rm Pol}_{\rm DDS/HF}$ is the Hartree\hyp{}Fock polarization energy
 defined
@@
 developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
-whereas $\Delta E^{\rm Ind}$ term is the 
+whereas $E^{\rm Ind}$ term is the 
 approximate pure induction energy due to the polarizability effects.\cite{Stone.TheTheoryOfIntermolecularForces.1996}
@@
 \begin{multline}\label{e:de-pol-dds}
-\Delta E^{\rm Pol}_{\rm DDS/HF} \equiv \Delta E^{\rm HF} \\
+ E^{\rm Pol}_{\rm DDS/HF} \equiv \Delta E^{\rm HF} \\
  - \left(
-   \Delta E^{\rm Coul}_{\rm DDS/HF}
- + \Delta E^{\rm Exch}_{\rm DDS/HF}
- + \Delta E^{\rm Rep}_{\rm DDS/HF} 
+   E^{\rm Coul}_{\rm DDS/HF}
+ + E^{\rm Exch}_{\rm DDS/HF}
+ + E^{\rm Rep}_{\rm DDS/HF} 
    \right)\;.
@@
 %
-In the above, $\Delta E^{\rm Coul}_{\rm DDS/HF}$ is the pure Coulombic interaction energy
+In the above, $\Delta E^{\rm HF}$ is the total HF interaction
+energy,
+$E^{\rm Coul}_{\rm DDS/HF}$ is the pure Coulombic interaction energy
 between the unperturbed (non\hyp{}interacting) charge densities of the monomers in the complex,
 and
-$\Delta E^{\rm Exch}_{\rm DDS/HF}$ and $\Delta E^{\rm Rep}_{\rm DDS/HF}$
+$E^{\rm Exch}_{\rm DDS/HF}$ and $E^{\rm Rep}_{\rm DDS/HF}$
 are the exchange and the Pauli\hyp{}repulsion interaction energies, respectively.
 At the HF level of theory, the sum of the latter two terms,
-i.e., the exchange\hyp{}repulsion energy, $\Delta E^{\rm Ex-Rep}_{\rm DDS/HF}$,
+i.e., the exchange\hyp{}repulsion energy, $E^{\rm Ex-Rep}_{\rm DDS/HF}$,
 is in fact
@@
 \begin{equation}\label{e:ind-ref}
- \Delta E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\BM \Delta}^{-1} \cdot {\bf F} \;.
+ E^{\rm Ind} \approx - \frac{1}{2} {\bf F}^{\rm T} \cdot {\BM \Delta}^{-1} \cdot {\bf F} \;.
 \end{equation}
@@
 \left\{\begin{matrix}
-{\BM\upalpha}_a^{-1} &\text{if $a=b$} \\ 
+{\BM\upalpha}_a^{-1} &\text{if $a=b$}\qquad\qquad\qquad\qquad\qquad\qquad \\ 
 0                    &\text{if $a\ne b \land a,b \in$ same molecule}\\ 
--{\bf T}_{ab}        &\text{if $a\ne b \land a,b \notin$ same molecule}
+-{\bf T}_{ab}        &\text{if $a\ne b \land a,b \notin$ same molecule}&
 \end{matrix}\right.
@@
  {\bf F}({\bf r}_a) = \sum_x^{\rm At} \frac{Z_x}{r_{ax}^3} {\bf r}_{ax} 
-  + 2 \sum_{\alpha\gamma} D_{\alpha\gamma} 
+  + 2 \sum_{\alpha\gamma}^{\rm AO} D_{\alpha\gamma} 
 %    \tBraKet{\gamma}{\frac{{\bf r}_a - {\bf r}}{\vert {\bf r}_a - {\bf r} \vert^3} }{\alpha}_{\bf r} \;.
@@
 from Eqs.~\eqref{e:ct.group-ii.final} and \eqref{e:ct.group-iii.final.b}
-is usually better 
+is usually slightly better 
 when the Boys method\cite{Boys.RevModPhys.1960} is used, as compared
@@
 in order to eliminate the BSSE and properly inlcude the CT processes.\cite{Stone.Misquitta.CPL.2009}
-Estimations of the $\Delta E^{\rm Coul}_{\rm DDS/HF}$, $\Delta E^{\rm Ex-Rep}_{\rm DDS/HF}$,
-$\Delta E^{\rm Ind}$ and $\Delta E^{\rm CT}_{\rm EFP2}$ interaction energy components 
+Estimations of the $E^{\rm Coul}_{\rm DDS/HF}$, $E^{\rm Ex-Rep}_{\rm DDS/HF}$,
+$E^{\rm Ind}$ and $E^{\rm CT}_{\rm EFP2}$ interaction energy components 
 obtained by using the computer 
@@
 \begin{equation} \label{e:ect-oep.final-scaled}
- \Delta E^{\rm CT} \approx C \times \Delta E^{\rm OEP}
-\end{equation}
-%
-where $C$ is an empircial constant and $\Delta E^{\rm OEP}$ is given by
+ E^{\rm CT}_{\rm scaled} \approx C \times E^{\rm CT}_{\rm unscaled}
+\end{equation}
+%
+where $C$ is an empircial constant and $E^{\rm CT}_{\rm unscaled}$ is given by
 Eqs.~\eqref{e:ct-hf0} and Eq.~\eqref{e:ect-oep.final}. Based on
@@
 further studies are necessary, that will modify the original formulation
-of OL and LOA treatment. Nevertheless, except from the ionic systems,
+of OL and LOA treatment, 
+or use different sets of virtual orbitals\cite{Xu.Gordon.JCP.2013}. Nevertheless, except from the ionic systems,
 it is clear that the level of accuracy of the EFP2 and OEP models are rather comparable,
@@
 from the BioFragment Database.
-The EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
-for the OEP calculations.
+For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+was used.
 } 
@@
 Constant   && \multirow{2}{*}{$\varepsilon_i$} 
-            & \multirow{2}{*}{$\varepsilon_i$, $\varepsilon_n$} \\
+            & \multirow{2}{*}{$\varepsilon_i$, $\varepsilon_n$, $L_{ii'}^A$} \\
 parameters\footnotemark[2] 
@@
 parameters\footnotemark[2] 
-                && & $\{\eta\}$, $V_{n\eta}^{{\rm eff},B}$, $L_{ii'}^A$ \\
+                && & $\{\eta\}$, $V_{n\eta}^{{\rm eff},B}$ \\
 \hline
@@
            && $T_{nn}$, $T_{kj}$, $T_{wj}$, $T_{nj}$,  &  $u_i^{BA}$, $w_{yi}^{BA}$\\
-           && $V_{in}^{{\rm EF},B}$ , $V_{ik}^{{\rm EF},B}$, $V_{iw}^{{\rm EF},B}$
+           && $U_{in}^{{\rm EF},B}$ , $U_{ik}^{{\rm EF},B}$, $U_{iw}^{{\rm EF},B}$
                                                        &  \\
@@
 virtual MO's of $B$: $n$; 
-atoms of $B$: $y$.}
+atoms of $B$: $y$. Analysis is based on $E^{A\rightarrow B}$ term.}
 \footnotetext[3]{
@@
 computational cost of the EFP2 and OEP models is shown.
-It is apparent that EFP2 has much more quantities to compute on the fly
+It is apparent that EFP2 requires 
+much more quantities to be computed 
 as compared to the OEP method (`Calculables' in the table).
@@
 evaluation of the EFP2 CT expression from Eq.~\eqref{e:u-efp2-ct} 
-requires quite a number of different types of OEI's. According to our estimations
+involves quite a number of different types of OEI's. According to our estimations
 that assume sequential (two\hyp{}step) two\hyp{}index AO\hyp{}MO transformations of OEI matrices
@@
 the computational cost
-has an order of $2p^3(s+t) + 3vop^2$, where the
+is of an order of $2p^3(s+t) + 3vop^2$, where the
 $o$ and $p$ denote the number of occupied orbitals and the number of atomic basis functions,
 respectively. 
-$s$, $t$ and $v$ are the relative costs of evaluation of the
+Here, $s$, $t$ and $v$ are the relative costs of evaluation of the
 overlap, kinetic energy, and multipole potential OEI's, respectively, with the latter being
@@
 %that are 
-necessary to compute ${\bf V}^{\rm EF}$ matrices
+necessary to compute ${\bf U}^{\rm EF}$ matrices
 from Eq.~\eqref{e:v-efp2-eff}. 
@@
 %
+\noindent \hspace{-20pt}
 On the contrary, 
@@
 %respectively. 
-Note also that, among the calculables that are needed on the fly in each
+Note also that, among the calculables that are needed in each
 OEP\hyp{}based CT energy evaluation, are the auxiliary vectors and matrices 
@@
 the cost of which is negligible. The amount of effective fragment parameters,
-that need to be superimposed during the calculations by applying rotation
+that needs to be superimposed during the calculations by applying rotation
 of orbitals and basis functions (`Superimposable parameters')
@@
 and the results are shown in Table~\ref{t:5}. 
-Profiling of the code for OL, EFP2 and OEP methods was performed 
+Time profiling of the code for OL, EFP2 and OEP methods was performed 
 for all the computational operations required for a
 single point calculation in a hypothetical sequential run on multiple
-geometries. Therefore, calculations of ERI's in AO basis 
+geometries. Therefore, calculations of intermonomer 
+ERI's in AO basis 
 along with their four\hyp{}index transformation to MO basis were included in
@@
 by definition.
-The above profiling settings allows one to compare the performances
+The above time profiling setting allows one to compare the performances
 of the OL, EFP2 and OEP methods as effective fragment parameter methods.
@@
 theories of intermolecular interactions
-was presented.
+was proposed.
 It was shown that in general case two types of OEP's
@@
 Therefore, it is believed that the OEP\hyp{}based
-model can be incorporated within the EFP2 method, enabling
+model can be incorporated within the EFP2 method, strongly
+facilitating
 the full (CT\hyp{}including) EFP2 energy calculations and 
@@
 \emph{ab initio} fragment\hyp{}based
-approaches for condensed\hyp{}phase simulations, where ERI pose the
+approaches for condensed\hyp{}phase simulations, where ERI's pose the
 computational challenge when confronted with the size of the system.
@@
 computational resources.
-We thank Professor Marcos Mandado from University of Vigo in Spain
+We cordially thank Professor Marcos Mandado from University of Vigo in Spain
 for providing us with his benchmark calculations to validate our implementation
@@
 %from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
-by expanding the MO's in terms of atomic basis functions. The working formula
+by expanding the MO's in terms of the AO's. The working formula
 is %given below:
@@
    \tBraKet{\alpha}{\hat{v}_{\rm eff}^{{\rm G}_1}}{n} 
-     = \sum_{x}^{\rm At} W_{\alpha n}^{(x)}   \\
+     = -\sum_{x}^{\rm At} W_{\alpha n}^{(x)}   \\
        + \sum_{\beta\gamma\delta} 

### 2019-12-21 papers/3_oep_technique.Pauli/paper-r3.tex
Figures in Paper 3 updated

--- 2019-12-13 papers/3_oep_technique.Pauli/paper-r2.tex
@@
 \begin{document}
-\preprint{AIP/123-OEP}
 
@@
 the extended density fitting in incomplete auxiliary basis.
-In the proposed approach the evaluation kinetic energy integrals is no
+In the proposed approach the evaluation of the kinetic energy integrals is no
 longer required and the computational cost can be reduced roughly by a factor
@@
 defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
+$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
+integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
+of unperturbed monomer $X$, all defined in LMO basis.
 It is known that evaluating Eq.~\eqref{e:exr-efp2}
@@
                S_{ij} \Big\{
-           W^A_{ij} 
-         + W^B_{ji} \\
+         - W^A_{ij} 
+         - W^B_{ji} \\
  + \sum_{k\in A}^{\rm Occ} \left[ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right] 
@@
  \begin{align}
- \BraKet{ij}{kk} &\equiv \sum_{\mu\in A} 
+ \BraKet{ij}{kk} &\equiv - \sum_{\mu\in A} 
      C_{\mu i}^A \tBraKet{\mu}{\hat{v}_{kk}}{j} \;, \\
- \BraKet{ik}{jk} &\equiv \sum_{\mu\in A} 
+ \BraKet{ik}{jk} &\equiv - \sum_{\mu\in A} 
      C_{\mu k}^A \tBraKet{\mu}{\hat{v}_{ik}}{j} \;.
@@
 \begin{equation}
-  \hat{v}_{ik} \equiv \int d{\bf r} \Ket{{\bf r}} 
+  \hat{v}_{ik} \equiv - \int d{\bf r} \Ket{{\bf r}} 
         \left[
@@
 \begin{equation}
- V^A_{ij} + 
+ -W^A_{ij} + 
  \sum_{k\in A}^{\rm Occ} 
@@
 %
-In this procedure, the required OEP matrix elements can be calculated from
+In this procedure, $R_{\alpha\gamma}= \iint d{\bf r}_1 d{\bf r}_2 \frac{\phi_{\alpha}^*({\bf r}_1) \phi_{\beta}({\bf r}_2) }{\lvert {\bf r}_1 - {\bf r}_2 \rvert}$, ${\bf S}$ is the matrix of overlap integrals in AO basis
+and the required OEP matrix elements can be calculated from
 %
@@
    \tBraKet{\zeta}{\hat{v}_{\rm eff}^{X[\zeta i]}}{i}
-     = \sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\
+     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\
         + \sum_{\beta\gamma\delta\in X}^{\rm AO}

### 2019-12-21 papers/3_oep_technique.Pauli/supplementary-r3.tex
Figures in Paper 3 updated

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

% numbering for Figures in SI
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\title{Supplementary Information\\Ab Initio Effective One-Electron Potential Operators. II.
Applications for Exchange-Repulsion Energy in Effective Fragment Potentials}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Robert W. G{\'o}ra}
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\pacs{}

\maketitle

\tableofcontents

\section{Minimalistic auxiliary basis sets for water, methanol and DMSO molecules}

The basis sets are given below. Note that in case of DMSO, the basis set size is smaller than the
minimal basis size (3p orbital on S atoms were not necessary to obtain good fitting in the test basis set).

\begin{verbatim}
[ water ]
cartesian
# primary basis set: 6-311++G** 6D
****
H     0
S   1   1.00
        0.500             1.000000
****
O     0
S   2   1.00
      937.182             0.986162 
        6.892             0.014190 
S   1   1.00
       45.156             1.000000
P   2   1.00
       24.823             0.277579 
        2.783             0.301334 
****

[ methanol ]
cartesian
# primary basis set: 6-311++G** 6D
****
H     0
S   1   1.00
        0.539             1.000000
****
C     0
S   1   1.00
      939.683             1.000000
S   1   1.00
       30.185             1.000000
P   1   1.00
        5.832             1.000000
P   1   1.00
        0.326             1.000000
****
O     0
S   1   1.00
      731.638             1.000000
S   1   1.00
       37.387             1.000000
P   1   1.00
        9.214             1.000000
P   1   1.00
        0.368             1.000000
****

[ dmso ]
cartesian
# primary basis set: 6-31++G** 6D
****
H     0
S   1   1.00
        0.365             1.000000
****
C     0
S   1   1.00
      207.472             1.000000
S   1   1.00
       15.052             1.000000
P   1   1.00
        7.228             1.000000
****
O     0
S   1   1.00
      302.339             1.000000
S   1   1.00
       15.248             1.000000
P   1   1.00
        7.612             1.000000
****
S     0
S   1   1.00
      689.986             1.000000
S   1   1.00
      145.419             1.000000
S   1   1.00
        9.046             1.000000
P   1   1.00
       28.979             1.000000
****
\end{verbatim}

% -----------------------
\bibliography{references}
% -----------------------

\end{document}

### 2019-12-30 papers/3_oep_technique.Pauli/supplementary-r3.tex
Figures in Paper 2 modified

@@
 
+\section{Accuracy of OEP and EFP2 models for exchange-repulsion energy estimation}
+
+%
+\begin{figure*}[t]
+\includegraphics[width=0.8\textwidth]{fig-2.pdf}
+\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of exchange\hyp{}repulsion energy
+across various bi\hyp{}molecular systems.} 
+(a) NCB31 
+database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
+of non\hyp{}covalent interactions
+and
+(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
+For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+was used.
+} 
+\end{figure*}
+%
+
+%
+\begin{figure*}[t]
+\includegraphics[width=0.8\textwidth]{fig-s2.pdf}
+\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of first\hyp{}order repulsion energy
+across various bi\hyp{}molecular systems.} 
+(a) NCB31 
+database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
+of non\hyp{}covalent interactions
+and
+(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
+of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
+For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
+was used.
+} 
+\end{figure*}
+%
+
+\clearpage
+\section{References}
+
 % -----------------------

### 2019-12-30 papers/3_oep_technique.Pauli/figures/make.fig-s1.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+figure_dir="./"
+
+gnuplot plot.fig-s1.gpl
+
+for plot in fig-s1 ; do
+    epstopdf ${plot}-inc.eps
+    pdflatex ${plot}.tex
+    pdftops -eps ${plot}.pdf
+    #mv -v ${plot}.pdf ${plot}.eps ${figure_dir}
+    #rm -v ${plot}.pdf
+    rm -v ${plot}.aux ${plot}.tex ${plot}.log ${plot}-inc*
+done

### 2019-12-30 papers/3_oep_technique.Pauli/paper-r3.tex
Revision 3 updated

@@
 \begin{document}
+%\preprint{AIP/123-OEP}
 
@@
 
-The exchange\hyp{}repulsion (EXR) interaction energy in the
+The exchange\hyp{}repulsion (EXR) interaction energy 
+usually describes the non\hyp{}electrostatic repulsion
+between two isolated and unperturbed 
+wavefunctions in an interacting
+molecular 
+complex.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975,Hayes.Stone.MolPhys.1984,Jeziorski.Moszynski.Szalewicz.ChemRev.1994,Jensen.Gordon.MolPhys.1996,Hesselmann.Jansen.Schutz.JCP.2005,Mandado.Hermida-Ramon.JCTC.2011}
+In the
 second generation of the
@@
    Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
-describes the non\hyp{}electrostatic repulsion
-between two isolated and unperturbed Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
-wavefunctions in an interacting
-bi\hyp{}molecular complex.
-The EXR EFP2 term was derived from the intermolecular perturbation theory
+the EXR energy is an important component of the total interaction energy.
+It is
+derived from the intermolecular perturbation theory
 of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
-up to the second order with respect to the wavefunction overlap,
+up to the second order with respect to the wavefunction overlap
+assuming Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} (HF) wavefunctions,
 and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
@@
          - W^A_{ij} 
-         - W^B_{ji} \\
+         - W^B_{ij} \\
  + \sum_{k\in A}^{\rm Occ} \left[ 2\BraKet{ij}{kk} - \BraKet{ik}{jk} \right] 
@@
 $\hat{v}^A_{\rm nuc}$
-is the nucleus\hyp{}electron potential operator
+is the electrostatic potential operator due to nuclei
 and
@@
 %from Eq.~\eqref{e:rep.murrell-etal.S1} 
-the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
+and the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
 here referred to as the HS model.
@@
 The Boys method\cite{Boys.RevModPhys.1960} was used to localize
-molecular orbitals---the same as in the original EFP2 formulation,
+molecular orbitals---the same as in the original EFP2 formulation.
 Note here that, since
@@
 (a) NCB31 
-database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
+database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
 Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
@@
 (Table~\ref{t:tab-1}).
-Computational cost of $\Delta E^{\rm Rep}(S^{-1})$ in EFP2 model
+Computational cost of $E^{\rm Rep}(S^{-1})$ in EFP2 model
 is approximately $s(on^2 + o^2n) + t(on^2 + o^2n) + 2o^3$, whereas in OEP model
@@
 program,\cite{Psi4.JCTC.2017}
-were utilized. The computed
-EXR energies were compared with the reference HS results (Figure~\ref{f:fig-2}).
-On average, OEP and EFP2 methods tend to underestimate the EXR energy by 0.7\% and 7.6\%
-in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
-respectively.
+was utilized. The computed
+EXR energies were compared with the reference Murrell~et~al. results (Figure~\ref{f:fig-2})
+as well as with the reference HS results (Figure~S1).
+On average, OEP and EFP2 methods are in good agreement
+with the reference models and correlation $R^2$ coefficients
+are between 90--99\% in all data sets. 
+The mutual differences between EFP2 and OEP estimations of 
+$E^{\rm Rep}(S^{-1})$ are on the average 0.7 kcal/mol  (Figure~S2) which shows that the ERI elimination proposed here is accurate.
+Root mean square errors (RMSE) of EXR energy estimation
+via OEP and EFP2 models
+%against both %the Murrell~et~al. and the HS reference
+are around 2.0~kcal/mol and 2.6~kcal/mol in the NCB31/6-311++G(d,p) data set, and 0.6~kcal/mol 
+and 0.3--1.2~kcal/mol
+in the BBI/6-31+G(d,p) data set, respectively. However,
+while the OEP and EFP2 models tend to overestimate
+the EXR energy as compared to the Murrell~et~al. reference
+by 7--12\%
+(except for the EFP2 model in the BBI/6-31+G(d,p) data set that shows very good agreement), 
+they consistently underestimate the EXR energy
+as compared to the HS reference by 0.7--7\% (OEP)
+and 8--15\% (EFP2).
 This is mostly due to BSSE, that is corrected for
-only in the reference calculations with the HS model, and the effect is usually larger
-when smaller primary basis sets are used.
-The OEP model performs slightly better when compared to the EFP2
-method for systems studied in this work, with 
-root-mean-square errors of OEP and EFP2 models of
-2.0 kcal/mol and 2.6 kcal/mol, respectively, for NCB31/6-311++G(d,p) data set, 
-or
-0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set.
-The slightly better performance of OEP is most likely due to
-fruitful cancellation of errors, since
-the mutual differences between EFP2 and OEP estimations of 
-$\Delta E^{\rm Rep}(S^{-1})$ are on the average much smaller
-(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
+only in the reference calculations with the HS model.
+%, and the effect is usually larger
+%when smaller primary basis sets are used.
+%
+% though the.
+%
+%much smaller
+%(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
+%
+%with root mean square errors (RMSE)
+%not larger than 2.3~kcal/mol and 2.9~kcal/mol
+%in NCB31/6-311++G(d,p) data set,
+%as well as
+%0.6~kcal/mol and 1.2~kcal/mol 
+%for BBI/6-31+G(d,p) data set. 
+%On the one hand, when compared to the HS model as reference,
+%OEP and EFP2 tend to underestimate the EXR energy
+%in all studied cases. On the other hand, EFP2 well
+%reproduces EXR energies obtained by using the Murrell~et~al. %model,
+%while OEP slightly overestimates the 
+%
+%However, due to the nature of
+%approximations made, the origins of errors are 
+%
+%tend to underestimate the EXR energy by 0.7\% and 7.6\%
+%in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
+%respectively.
+%The OEP model performs slightly better when compared to the EFP2
+%method for systems studied in this work, with 
+%root-mean-square errors of OEP and EFP2 models of
+%2.0 kcal/mol and 2.6 kcal/mol, respectively, for NCB31/6-311++G(d,p) data set, 
+%or
+%0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set.
+%On the other hand, when compared to 
+%the Murrell~et~al. theory as reference (see Supplementary Information),
+%EFP2 model performs 
+%slightly better than OEP model. Note that in this case BSSE
+%is not included in the reference calculations.
+%Therefore
+%The slightly better performance of OEP vs HS reference 
+%is most likely due to
+%fruitful cancellation of errors, 
+%
+%Careful inspection (Figure~S1) revealed that when compared to 
+%the Murrell~et~al. theory as reference, EFP2 model performs 
+%slightly better than OEP model.
 %
@@
 %
-Nevertheless, together with the performance data of both models
+Nevertheless,
+OEP model is rather of roughly comparable accuracy as EFP2 model in all the systems studied.
+Together with the performance data of both models
 from Table~\ref{t:tab-1}, OEP model consistently outperforms the
-EFP2 approach in terms of accuracy and also performance, provided
+EFP2 approach provided
 that the minimal auxiliary basis set is used.
-
-%In particular, the subsets from the NCB31 database were separately considered:
-%the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-%the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-%the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
-%the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
-%and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
 
@@
 (grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
+Wroclaw Centre for Networking and Supercomputing (WCSS) is acknowledged for
+computational resources.
 \end{acknowledgments}

### 2019-12-30 papers/3_oep_technique.Pauli/submit.tex
Submitted versions of papers added

--- 2019-12-30 papers/3_oep_technique.Pauli/paper-r3.tex
@@
+\documentclass[aip,jcp,amsmath,amssymb,preprint,floatfix]{revtex4-1}
 \input{header.tex}
 \begin{document}
-%\preprint{AIP/123-OEP}
 
@@
 %
-\begin{multline} \label{e:rep-efp2.s1}
- E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
- \times
+\begin{equation} \label{e:rep-efp2.s1}
+ E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} 
+% \times
  \left\{
@@
  \right\}
-\end{multline}
+\end{equation}
 %
@@
 %
-\begin{multline} \label{e:rep-efp2.s2}
+\begin{equation} \label{e:rep-efp2.s2}
  E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
- \\ \times 
+ %\\ \times 
  \Bigg\{
@@
  \Bigg\} \;.
-\end{multline}
+\end{equation}
 %
@@
 %
-\begin{multline} \label{e:rep.murrell-etal.S1.oep-v}
+\begin{equation} \label{e:rep.murrell-etal.S1.oep-v}
    \tBraKet{\zeta}{\hat{v}_{\rm eff}^{X[\zeta i]}}{i}
-     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\
+     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} %\\
         + \sum_{\beta\gamma\delta\in X}^{\rm AO}
@@
            \BraKet{\zeta\beta}{\gamma\delta} \;,
-\end{multline}
+\end{equation}
 %

### 2020-02-21 papers/2_oep_technique.CT/paper-r3.tex

@@
 %
-On the contrary, 
+\\ On the contrary, 
 OEP\hyp{}based expression

### 2020-02-21 papers/2_oep_technique.CT/supplementary-r2.tex

@@
+\documentclass[aip,jcp,amsmath,amssymb,preprint,floatfix]{revtex4-1}
 \input{header.tex}
@@
 \preprint{AIP/123-OEP}
+
+% numbering for Figures in SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
 
@@
 is also shown for comparison in light grey color in this figure.
-All data were obtained at HF/6-311++G(2df,2pd) level of theory
-with the single GDF scheme and aug-cc-pVTZ-jkfit auxiliary basis set.
+All data were obtained at HF/6-311++G(d,p) level of theory
+with the EDF-1 scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
 } 
@@
 
+In Figure~\ref{f:fig-s2} interaction energy components
+obtained by using the code developed in this work are
+compared to the ones obtained by using the EFP2 calculation routines
+in GAMESS US quantum chemistry package\cite{GAMESS.JCC.1993} (version: Sept 30, 2017 R2 Public Release). 
+For this, EFP2 parameters
+were generated by running the \texttt{MAKEFP} routine
+with \texttt{CTVVO} option switched to \texttt{.FALSE.} which corresponds
+to using all the canonical molecular orbitals from HF calculations
+instead of valence virtual orbitals.
+Generally, Coulombic energies (Figure~S2(a)) 
+are in very good agreement with a few exeptions in the PPS5 set,
+for which the differences are due to the inaccuracies in the DMTP
+expansion used in the EFP2 model -- note that the Coulombic component in the DDS/HF method
+is free from multipole approximation and is equivalent to the first\hyp{}order SAPT Coulombic energy.
+Exchange\hyp{}repulsion and induction energies (Figures~S2(b) and S2(c), respectively)
+are in very good agreement. CT energies (Figure~S2(d)) 
+evaluated with CAMM up to quadrupoles for the potential energy matrix elements (black filled circles)
+are generally also in decent quantitative agreement,
+except for NH$_3$--FCl, (HCOOH)$_2$ and (HCONH$_2$)$_2$ 
+%and parallel\hyp{}displaced benzene dimer
+where the differences
+are between 2--4 kcal/mol. Only in the case of the latter three systems,
+adding distributed octupoles (blue crosses in Figure~S2(d)) 
+lowers the CT energy values by around 1 kcal/mol,
+with negligible changes for the rest.
+
+%
+\begin{figure}[h]
+\includegraphics[width=0.8\textwidth]{fig-s2.pdf}
+\caption{\label{f:fig-s2} {\bf Validation of implementation of the interaction energy components
+with respect to the EFP2 GAMESS US code.} 
+Results obtained for the NCB31 database set assuming 6-311++G(d,p) primary basis set. 
+(a) - Coulombic, (b) - exchange-repulsion, (c) - induction, and (d) - charge-transfer
+interaction energy. In (d), black circles and blue crosses correspond to the
+CT EFP2 energies with CAMM up to distributed quadrupoles and octupoles, respectively.
+For more details see the main text (Section III, \emph{Calculation Details}).
+} 
+\end{figure}
+%
+
 % -----------------------

### 2020-02-21 papers/3_oep_technique.Pauli/paper-r1.tex

--- 2019-12-12 papers/3_oep_technique.Pauli/paper-r0.tex
@@
 \begin{abstract}
-Recently, it was shown that
+In Paper~I, it was shown that
 the effective one\hyp{}electron potentials (OEP) method
@@
 exchange\hyp{}repulsion EFP2 interaction energy
-by incorporating the kinetic energy operator in the OEP
-via the extended density fitting in incomplete auxiliary basis.
+by re\hyp{}defining the first\hyp{}order repulsive term of 
+Murrell~et al. through
+the extended density fitting in incomplete auxiliary basis.
 It is shown that the computational cost can be reduced by a factor
-of 1.5 and kinetic energy integrals are no longer necessary to evaluate.
+of 1.5 as compared to the original EFP2 formulation
+and kinetic energy integrals are no longer necessary to evaluate.
 \end{abstract}
@@
 It is shown that calculation of kinetic energy integrals can be avoided
-by introducing OEP's.
+by introducing OEP's leading to enhancepent of computational efficiency.
 
@@
 
-We implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
+In our in\hyp{}house plugin to the {\sc Psi4}
+quantum chemistry
+program,\cite{Psi4.JCTC.2017}
+we implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
 (EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as 
@@
 %from Eq.~\eqref{e:rep.murrell-etal.S1} 
-the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984}
-and the density decomposition scheme 
-at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
-developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
-for validation of the EDF technique.
-When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
-the DDS/HF model is
-equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
-which is the case in this work. Using DCBS also allows one to eliminate the basis set
-superimposition error (DCBS).
+the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
+here referred to as the HS model.
+The latter was
+implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
+to eliminate the basis set
+superimposition error (DCBS) in the benchmark calculations.
+%and the density decomposition scheme 
+%at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
+%developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
+%for validation of the EDF technique.
+%When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
+%the DDS/HF model is
+%equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
+%which is the case in this work. 
+%Using DCBS also allows one to eliminate the basis set
+%superimposition error (DCBS).
 To localize molecular orbitals, procedure that is necessary for the EFP2 calculations,
@@
 (Table~\ref{t:tab-1}).
-However, application of our small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
+Computational cost of $\Delta E^{\rm Rep}(S^{-1})$ in EFP2 model
+is approximately $s(on^2 + o^2n) + t(on^2 + o^2n) + 2o^3$, whereas in OEP model
+is $s(on^2 + o^2n) + 2ao^2$, both of which have comparable magnitudes. 
+Here, $s$ and $t$ denote the relative cost of evaluation of the overlap
+and kinetic energy one electron integrals (OEI's), whereas $o$, $n$ and $a$ is
+the number of LMO's, primary AO's and auxiliary AO's, respectively.
+It is clear that the critical parameter is the number of auxiliary basis functions $a$,
+which should be comparable with the number of occupied orbitals $o$ 
+as $t$ is usually only 2--3 times larger than $s$.
+We developed small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
 %developed
-developed via the basis set optimization~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019}
-in conjunction with the basin hopping global optimization algorithm  %cite
-as implemented in the SciPy Python library
-(see Supplementary Information),
+via the basis set optimization method described in Appendix~A to Paper I
+in conjunction with the basin hopping global optimization algorithm\cite{Wales.EnergyLandscapes.2003,
+Wales.Scheraga.Science.1999,
+Wales.Doye.JPCA.1997,Li.Scheraga.PNAS.1987}
+as implemented in the SciPy Python library\cite{SciPy.2019}
+(see Supplementary Information).
 %to minimize the computational cost of evaluating the summation terms
 %from Eq.~\eqref{e:rep.murrell-etal.S1.oep},
+Applying such basis sets
 reduces costs appreciably and makes the OEP method on average 1.5 times faster than EFP2.
@@
 basis set and 
-1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
+1.2~GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
 requires about 4.6~ms and 3.0~ms when EFP2 and OEP model is used, respectively.
@@
 were utilized, and the
-EXR energies were compared with the reference DDS/HF results (Figure~\ref{f:fig-2}).
+EXR energies were compared with the reference HS results (Figure~\ref{f:fig-2}).
 On average, OEP and EXR tend to underestimate the EXR energy by 0.7\% and 7.6\%
@@
 This is mostly due to the lack of BSSE, that is accounted for
-only in the reference calculations with the DDS/HF model, and the effect is usually larger
+only in the reference calculations with the HS model, and the effect is usually larger
 when smaller primary basis sets are used.
@@
 It is an accidental fruitful cancellation of errors because
-the first\hyp{}order repulsive terms have much less differences with respect to each other
-(RMSE is 0.7\% only in all data sets).
+the EFP2 and OEP estimations of 
+$\Delta E^{\rm Rep}(S^{-1})$ have much less differences with respect to each other
+(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
 Nevertheless, together with the performance data of both models

### 2020-02-21 papers/3_oep_technique.Pauli/paper-r2.tex

@@
 \begin{document}
-\preprint{AIP/123-OEP}
+%\preprint{AIP/123-OEP}
 

### 2020-02-21 papers/4_eet.I.Theory/paper-r0.tex

--- 2019-08-28 papers/3_eet.I.Theory/paper-r0.tex
@@
 \begin{equation}
- \mathscr{F} \approx \mathscr{F}^{A}_0 + \mathscr{F}^{B}_0 \;,
+% \mathscr{F} \approx \mathscr{F}^{A}_0 + \mathscr{F}^{B}_0 \;,
+ \mathscr{F} \approx \hat{T} + 
+   \sum_X^{A,B} \left\{ 
+      \hat{V}^{X}_{\rm nuc} + \hat{J}({\bf P}^{X}_g) - \frac{1}{2} \hat{K}({\bf P}^{X}_g)
+ \right\} \;,
 \end{equation}
 %
-where
+where $\hat{T}$ is the kinetic energy operator, 
+$\hat{V}^{X}_{\rm nuc}$ is the operator associated with the electrostatic potential energy
+between nuclei and electrons in molecule $X$, and the Coulomb and exchange operators
+are defined by
+%
+\begin{subequations}
+\begin{align}
+ \hat{J}({\bf P}^{X}_g) &\equiv \sum_{\sigma\lambda} P^{X(g)}_{\lambda\sigma}
+           \iint \frac{\varphi_\lambda({\bf r}_1) \varphi^*_\sigma({\bf r}_1) }{\vert {\bf r}_1 - {\bf r}_2 \vert}
+           \square \; d{\bf r}_1 d{\bf r}_2 \;,\\
+ \hat{K}({\bf P}^{X}_g) &\equiv \sum_{\sigma\lambda} P^{X(g)}_{\lambda\sigma}
+           \iint \frac{\varphi_\lambda({\bf r}_1) \varphi^*_\sigma({\bf r}_2) }{\vert {\bf r}_1 - {\bf r}_2 \vert}
+           \square \; d{\bf r}_1 d{\bf r}_2 \;.
+\end{align} 
+\end{subequations}
+%
+In the above definitions, the `$\square$' symbol reminds that $\hat{J}$ and $\hat{K}$ are the integral operators
+and is given by
 %
 \begin{equation}
- \mathscr{F}^{X}_0 = \hat{T}^{X} + \hat{V}^{X}_{\rm nuc} + \hat{J}({\bf P}^{X}_g) - \frac{1}{2} \hat{K}({\bf P}^{X}_g) \;.
+ \square \equiv \iint \Ket{{\bf r}} \Bra{{\bf r}'} \; d{\bf r} d{\bf r}' \;.
 \end{equation}
 %
-Note that, without loss of generality, one can rewrite Eq.~\eqref{e:v0-et1}
+Note that the kinetic energy operator cannot be decomposed into particular molecules. However, 
+consider the following partitioning of the total Fock operator:
+%
+\begin{equation}
+ \mathscr{F} \approx \mathscr{G}^{A}_0 + \mathscr{G}^{B}_0 \;,
+\end{equation}
+%
+in which the auxiliary operator is defined as
+%
+\begin{equation}
+ \mathscr{G}^{X}_0 = \mathscr{F}^{X}_0 - \frac{1}{2}\hat{T}^{X} \;.
+\end{equation}
+%
+Here, the kinetic energy operator was symmetrically partitioned in between two molecules.
+%
+Now,
+note that, without loss of generality, one can rewrite Eq.~\eqref{e:v0-et1}
 as
@@
  V^{{\rm ET1},(0)} = t_{H\rightarrow L}^A \times \Big\{ 
- \tbraket{\phi_L^A}{\mathscr{F}}{\phi_L^B} \\
+ \tbraket{\phi_L^A}{\left( \mathscr{G}^A_0 + \mathscr{G}^B_0 \right)}{\phi_L^B} \\
   + 2 \tbraket{\phi_H^A}{\hat{v}_{LH}^A}{\phi_L^B}
@@
 \begin{equation}
- \bra{\phi_L^A} \left\{ \mathscr{F}_0^A - \hat{v}^A_{HH} \right\}
+ \bra{\phi_L^A} \left\{ \mathscr{G}_0^A - \hat{v}^A_{HH} \right\}
  + \bra{\phi_H^A} 2 \hat{v}^A_{LH} 
@@
 \begin{equation}
- \mathscr{F}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B;{\rm ET}}_{\eta;L} \ket{\eta} 
+ \mathscr{G}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B;{\rm ET}}_{\eta;L} \ket{\eta} 
 \end{equation}
@@
 \begin{align}
- \bra{\phi_H^A} \left\{ -\mathscr{F}_0^A - \hat{v}^A_{LL} \right\}
+ \bra{\phi_H^A} \left\{ -\mathscr{G}_0^A - \hat{v}^A_{LL} \right\}
  + \bra{\phi_L^A} 2 \hat{v}^A_{HL} 
  &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT}}_{\xi,HL} \\
- -\mathscr{F}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm HT}}_{\eta,H} \ket{\eta} 
+ -\mathscr{G}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm HT}}_{\eta,H} \ket{\eta} 
 \end{align}
@@
 %
-a_{\alpha;L}^{X;{\rm ET}} &= \sum_\beta C_{\beta L}^X F_{\alpha\beta}^X \;,\\
+a_{\alpha;L}^{X;{\rm ET}} &= \sum_\beta C_{\beta L}^X G_{\alpha\beta}^X \;,\\
 %
@@
 %
-a_{\alpha;H}^{X;{\rm HT}} &=-\sum_\beta C_{\beta H}^X F_{\alpha\beta}^X \;,\\
+a_{\alpha;H}^{X;{\rm HT}} &=-\sum_\beta C_{\beta H}^X G_{\alpha\beta}^X \;,\\
 %
@@
 %
-Here, the Fock matrix in space spanned by the intermediate and primary AO basis sets
+Here, the auxiliary one\hyp{}electron ${\mathscr{G}}^X$ operator represented
+in space spanned by the intermediate and primary AO basis sets
 needs to be computed. The respective contributions to ${\bf a}$ vectors can be
@@
 \begin{multline}
- \sum_\beta C_{\beta U}^X F_{\alpha\beta}^X = 
- \sum_\beta C_{\beta U}^X h_{\alpha\beta}^X \\
+ \sum_\beta C_{\beta U}^X G_{\alpha\beta}^X = 
+ \sum_\beta C_{\beta U}^X \left( \frac{1}{2} T_{\alpha\beta}  + V_{{\rm nuc};\alpha\beta}^X \right)\\
    + \sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
@@
 %
-where ${\bf h}^X$ is the one\hyp{}electron core Hamiltonian of molecule $X$.
-
+where
+%
+\begin{subequations}
+\begin{align}
+ T_{\alpha\beta} &= -\frac{1}{2} \int \varphi^*_\alpha({\bf r}) \nabla^2_{{\bf r}} \varphi_\beta({\bf r})\; d{\bf r} 
+ \text{ and}\\ 
+ V_{{\rm nuc};\alpha\beta}^X &= -\sum_{x\in X} Z_x \int 
+           \frac{\varphi^*_\alpha({\bf r})  \varphi_\beta({\bf r}) }{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;.
+\end{align}
+\end{subequations}
+%
+In the above, $Z_x$ is the atomic number of the $x$th atom and $\nabla^2_{{\bf r}}\equiv 
+\frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2}$.
 % -----------------------

### 2020-02-21 gefp/gefp/density/dms.py
DMS routines revised and first qualitative model is working

@@ -12,128 +12,110 @@ import psi4, oepdev
 import scipy.optimize
 import scipy.spatial.transform
 from abc import ABC, abstractmethod
-from .partitioning import DensityDecomposition
-from ..math.matrix import move_atom_rotate_molecule, rotate_ao_matrix, matrix_power
-
-__all__ = ["DMSFit", "DMS", "Computer"]
-
-class Computer: #TODO
-  def __init__(self, *dms):
-      self._set_work(*dms)
-
-  def _set_work(self, *dms):
-      self._data = {}
-      self._data["dms"] = dms
-      #self._data["nbf"] = numpy.array([x._bfs.nbf() for x in dms])
-      #self._data["d"] = numpy.array([x._D for x in dms])
-
-      self._dim = sum(self._data["nbf"])
-      self._work = numpy.zeros((self._dim, self._dim))
-
-  def deformation_density(self, return_all=False):
-      dD_pau = self._deformation_density_pauli()
-      dD_pol = self._deformation_density_polar()
-      dD = dD_pau + dD_pol
-      if return_all: return dD_pau, dD_pol, dD
-      return dD
-
-  # protected
-  @abstractmethod
-  def _deformation_density_pauli(self): pass
-
-  @abstractmethod
-  def _deformation_density_polar(self): pass
-#      for i in range(self._nfrag):
-#          for j in range(self._nfrag):
-#              computer_ij = Computer_Dimer(*self._dms_set[i,j])
-
-class UnperturbedField_Computer(Computer): pass
-
-#class DMS_Computer(Computer):
-#
-#  def __init__(self, *dms):
-#      Computer.__init__(self, *dms)
-#
-#  def _deformation_density_pauli(self):
-#      raise NotImplementedError
-#      S_mo_t    = numpy.zeros((self._nmo_t, self._nmo_t), numpy.float64)
-#      W_mo_t    = numpy.zeros((self._nmo_t             ), numpy.float64)
-#      C_ao_mo_t = numpy.zeros((self._nbf_t, self._nmo_t), numpy.float64)
-#
-#      for i in range(self._nfrag):
-#          bfs_i = self._nbf_set[i]
-#          non_i = self._data["non"][i]
-#          noc_i = self._data["noc"][i]
-#          nmo_i = self._data["nmo"][i]
-#          nbf_i = self._data["nbf"][i]
-#          ofm_i = self._data["ofm"][i]
-#          ofb_i = self._data["ofb"][i]
[…]

### 2020-11-17 include/doxygen/oepdev_manual.h
Manual of EOPDev updated

@@ -36,7 +36,8 @@ re-expresses the physics of intermolecular interactions in terms
 of effective one-electron functions that are easier to handle 
 in practice.
 
-This Project will focus on finding a unified way to simplify various equations 
+This Project focuses on finding a unified way to simplify various
+fragment-based approaches
 of Quantum Chemistry of extended molecular systems, i.e., 
 molecular aggregates such as interacting chromophores
 and molecules solvated by water and other solvents. 
@@ -52,9 +53,9 @@ while introducing no or minor approximations to the original theories.
 \section resprojmeth Research Project Methodology
 
 In this Project the new theoretical protocol based on the 
-one-electron effective potentials (OEP's) is developed. 
+effective one-electron potentials (EOP's) is developed. 
 The main principle is to rewrite arbitrary sum of functions \f$ f \f$ 
-of electron repulsion integrals (ERI's) by defining \ref poepdesign "OEP's"
+of electron repulsion integrals (ERI's) by defining \ref poepdesign "EOP's"
 according to the following general prescription:
 \f{align*}{
 \sum_f f\left[ 
@@ -69,7 +70,7 @@ according to the following general prescription:
 where \f$ A \f$ and \f$ B \f$ denote different molecules 
 and \f$ \phi_i \f$ is the \f$ i \f$-th molecular orbital
 or basis function. Here,
-\f$ v_{kl}^B \f$ denotes the \ref poeptypes *ab initio* "OEP matrix element". The technique described above 
+\f$ v_{kl}^B \f$ denotes the \ref poeptypes *ab initio* "EOP matrix element". The technique described above 
 will be applied to simplify expressions for 
  - short-range excitation 
    energy transfer couplings between chlorophyll subunits of reaction centres 
@@ -83,47 +84,55 @@ in fragment-based *ab initio* molecular dynamics protocols of new generation.
 
 \section impact Expected Impact on the Development of Science, Civilization and Society
 
-The proposed OEP's are expected to significantly develop 
+The proposed EOP's are expected to significantly develop 
 the fragment-based methods that are widely used 
 in physical chemistry and modelling of biologically 
-important systems. Owing to universality of OEP's, 
+important systems. Owing to universality of EOP's, 
 they could find applications in many branches of chemical science: 
 *non-empirical* molecular dynamics, short-range
 resonance energy transfer in photosynthesis,
 electronic and vibrational solvatochromism, 
 multidimensional spectroscopy and so on. In particular:
 
[…]

### 2020-11-17 main.cc

@@ -4,8 +4,6 @@
  * oepdev by Bartosz Błasiak (email: blasiak.bartosz@gmail.com)
  *
  * Copyright (c) 2017 Bartosz Błasiak.
- * NOTE: oepdev is not open-source and cannot be redistributed, copied and downloaded
- * without a written consent of the Project Administrator.
  * 
  * oepdev is a plugin to:
  *

### 2020-11-17 pymodule.py

@@ -3,8 +3,6 @@
 # @BEGIN LICENSE
 #
 # oepdev by Bartosz Błasiak (email: blasiak.bartosz@gmail.com).
-# oepdev copy, use and redistribution not allowed without written 
-# consent of the Project Administrator.
 #
 # oepdev: A plugin to:
 #

### 2020-11-17 include/doxygen/oepdev_manual.h
Test of licence

@@ -562,7 +562,7 @@ helps in producing self-maintaining code and is much easier to use. Use:
 \subsection spadvpreppsi Preparing Psi4
 
 EOPDev is a Psi4 plugin. It requires 
-  - Psi4, version 1.2.1 (git commit `9d4a61c`). Has to be modified (see below).
+  - Psi4, version 1.2.1 (git commit `7a58e8c`). Has to be modified (see below).
   - Eigen3, any version.
 
 \note

### 2020-11-18 include/doxygen/oepdev_manual.h
Patch to Psi4 1.2.1 updated

@@ -96,14 +96,15 @@ multidimensional spectroscopy and so on. In particular:
 
  - the EOP-based models of Pauli repulsion energy and charge-transfer (CT) energy
    could be used to improve the computational performance of the second generation effective fragment
-   potential method (EFP2). At present, the CT term is very time consuming and due to this reasons
-   it is not used in many of applications of EFP2 to perform molecular dynamics 
+   potential method (EFP2). Particularly, the EFP2 CT term is relatively time consuming and due to this reason
+   it is sometimes omitted in the applications of EFP2 to perform molecular dynamics 
    simulations.\cite OEP-1.2020
  - the EOP-based model of EET couplings could significantly improve modelling of energy transfer
    in the light harvesting complexes. At present, short-range phenomena (Dexter mechanisms of EET)
-   are very difficult to efficiently and quantitatively asses when performing statistical averaging
+   are very difficult to efficiently and quantitatively evaluate when performing statistical averaging
    and applying to large molecular aggregates. Such Dexter effects could be computed by using EOP's
-   in much more efficient manner without loosing high accuracy of parent TDFI-TI method.\cite OEP-2.2020
+   in much more efficient manner without loosing high accuracy of state-of-the-art methods such as
+   TDFI-TI method.\cite OEP-2.2020
  - the density matrix polarization (DMS) tensors could be used in new generation fragment-based *ab initio*
    molecular dynamics protocols that rigorously take into consideration electron correlation effects.\cite Blasiak.JCP.2018
 
@@ -119,13 +120,13 @@ the EOPDev platform is developed.
 Accurate and efficient *ab initio* \ref pimplementedmodels "models"
 based on EOP's are implemented in the EOPDev code, along with
 the state-of-the-art benchmark and competiting methods. 
-Written entirely in C++, 
+Written in C++ with an extensive Python interface, 
 EOPDev is a plugin to Psi4 quantum chemistry package. Therefore,
 compilation and running the EOPDev code is straightforward and follows
 the API interface similar to the one used in Psi4 with 
 just a few \ref pprogramming "specific programing conventions".
 The detailed discussion about using the EOPDev code
-can be found in \ref advanced "advanced usage section". 
+can be found in \ref advanced "usage section". 
 \note The `OEP' abbreviation, rather than the `EOP',
       is used throughout the code. It is because the earlier versions
       of EOPDev utilized the former abbreviation consecutively for the shared libraries,
[…]

### 2020-11-19 include/doxygen/oepdev_manual.h
Revision - added TOC

@@ -18,18 +18,18 @@ energetic and dynamical properties of **extended molecular systems** efficiently
 accurately, 
 \cite Warshel.Levitt.JMolBiol.1976 
 \cite Senn.Thiel.Angew.2009 
-\cite Demerdash.Yap.T-Head-Gordon.AnnuRevPhysChem.2014 
+\cite Demerdash.Yap.T-Head-Gordon.ARPC.2014 
 \cite Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012
 there exist particularly difficult cases in which 
 modelling is still challenging: 
 
   - describing electronic transitions in solution or 
     when coupled with other electronic transition 
-    via resonance energy transfer, \cite Barbati.JACS.2014 Szabla.Sponer.Jiri.Gora.JPCL.2015 Bednarska.Zalesny.Tian.Murugan.Agren.Bartkowiak.Molecules.2017 Jedrzejewska.Grabarz.Bartkowiak.Osmialowski.SpectChimActA.2018
+    via resonance energy transfer, \cite Barbati.JACS.2014 \cite Szabla.Sponer.Jiri.Gora.JPCL.2015 \cite Bednarska.Zalesny.etal.Molecules.2017 \cite Jedrzejewska.Grabarz.etal.SCAA.2018
   - performing molecular dynamics 
     at very high level of theory including dynamic electron correlation, \cite Curchod.Martinez.ChemRev.2018
   - vibrational frequency calculations of particular normal mode in 
-    condensed phases \cite Blasiak.Londergan.Webb.Cho.ACR.2017 Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018 Lewis.Gruenke.Oliver.Ballottari.Bassi.Fleming.JPCL.2016
+    condensed phases \cite Blasiak.Londergan.Webb.Cho.ACR.2017 \cite Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018 \cite Lewis.Gruenke.etal.JPCL.2016
  
 and so on. The reason behind (sometimes prohibitively) 
 high costs of fully *ab initio* calculations in the above areas 
@@ -38,12 +38,26 @@ rather then (conceptually more straightforward) electronic densities and potenti
 On the other hand, 
 it has been pointed out before
 that the one-electron density 
-distributions are of particular importance in chemistry. \cite Kohn.Sham.PhysRev.1965 Holas.March.PhysRevA.1991
[…]

### 2020-11-19 include/doxygen/oepdev_modules.h

@@ -1,5 +1,5 @@
 /** 
- *  @defgroup OEPDEV_EOPS The Generalized One-Electron Potentials Library
+ *  @defgroup OEPDEV_OEPS The Generalized One-Electron Potentials Library
  *  @brief
  *  Implements the goal of this project: The Generalized One-Electron Potentials
  *  (EOPs). You will find here EOPs for computation of Pauli repulsion energy,

### 2020-11-19 include/doxygen/oepdev_manual.h
Final manual added

@@ -455,7 +455,7 @@ Therefore it should follow the programming etiquette of Psi4. Also,
 EOPDev has additional programming tips to make the code more versatile and easy to develop further.
 Here, I emphasise on most important aspects regarding the proposed **programming rules**.
 
-\section scontrmain Main routine and libraries
+\section scontrmain Main Routine and Libraries
 
 EOPDev has only *one* source file in the plugin base directory, i.e., `main.cc`. This is the main
 driver routine that handles the functionality of the whole EOP testing platform: specifies options for 
@@ -480,7 +480,7 @@ Things to remember:
      If you want to add graphics, put it in the `images` directory.
 
 
-\section sheader Header files in libraries
+\section sheader Header Files in Libraries
 
 Header files are handy in obtaining a quick glimpse of the functionality within certain library. Each library
 directory should contain at least one header file in EOPDev. However, header files can be problematic if not managed properly. 
@@ -511,7 +511,7 @@ Things to remember:
       For instance, all classes and functions in `oepdev` module are implemented within the namespace of the same label.
       Considering addition of other local namespaces within a module can also be useful in certain cases.
 
-\section senviron Environmental variables
+\section senviron Environmental Variables
 
 Defining the set of intrinsic environmental variables can help in code management 
 and conditional compilation. The EOPDev environmental variables are defined in
@@ -523,7 +523,7 @@ OEPDEV_XXXX
 \endcode
 where `XXXX` is the descriptive name of variable.
 
-\section sdocumentation Documenting the code
+\section sdocumentation Documenting the Code
 
 Code has to be documented (at best at a time it is being created). The place for documentation 
 is always in header files. Additional documentation can be also placed in source files. Leaving a chunk of code
@@ -544,7 +544,7 @@ Things to remember:
       a one-liner documentation (starting from `///`).
    4. **Do not be afraid of long names in the code**. Self-documenting code is a bless!
 
-\section snaming Naming conventions
+\section snaming Naming Conventions
 
 Naming is important because it helps to create more readable and clear self-documented code. 
 
@@ -570,7 +570,7 @@ Some loose suggestions:
       * good name: `DIISManager`, bad name: `DIIS`.
       * good name: `EETCouplingSolver`, bad name: `EETSolver`, very bad: `EET`.
      
-\section stime Track timing when evaluating the code
+\section stime Track Timing When Evaluating the Code
 
[…]

### 2020-11-19 oepdev/libtest/unitary_optimizer_2.cc
Missing test file added

@@ -1,0 +1,62 @@
+#include <iostream>
+#include "test.h"
+#include "../libutil/unitary_optimizer.h"
+
+using namespace std;
+
+
+double oepdev::test::Test::test_unitaryOptimizer_2()
+{
+  double result = 0.0;
+
+  double P[27] = {  -5.7237E-02,
+    -4.8214E-01,  -8.0553E-01,  -1.5206E-01,  -7.4836E-01,  -2.9927E-01,  -4.5974E-01,  -5.1164E-02,
+    -3.7566E-01,  -1.6202E-01,  -9.9207E-01,  -1.0660E-02,  -9.2229E-01,  -6.7787E-01,  -5.3848E-02,
+    -9.9106E-01,  -1.7727E-01,  -1.3879E-01,  -5.6017E-01,  -7.4425E-01,  -1.9731E-01,  -5.2214E-01,
+    -8.6566E-01,  -7.2151E-02,  -1.0403E-01,  -5.0845E-01,  -1.4330E-01 };
+
+  // the reference Xmin and Xmax values were set by checking the output of the UnitaryOptimizer_2
+  // for the given P tensor. To fully ensure the correct result, revise this test.
+  
+
+  const double Xmin_ref[9] = {  -0.44555350,   0.58229464,   0.680011039,
+                                 0.42752510,   0.80575731,  -0.409850497,
+                                -0.78657762,   0.10811146,  -0.607953579};
+  
+  const double Xmax_ref[9] = {   0.73412257,   0.55674902,  -0.388708844,
+                                 0.65062607,  -0.74056654,   0.168068170,
+                                -0.19429297,  -0.37628674,  -0.905902049};
+
+  oepdev::UnitaryOptimizer_2 optimizer(P, 3, 1.0e-8, 100, true);
+
+  psi::outfile->Printf(" ==> Unitary Minimization 2 in 3 Dimensions <==\n");
+  bool success_min = optimizer.minimize();
+  std::shared_ptr<psi::Matrix> X_min = optimizer.X();
+
+  psi::outfile->Printf(" ==> Unitary Maximization 2 in 3 Dimensions <==\n");
+  bool success_max = optimizer.maximize();
+  std::shared_ptr<psi::Matrix> X_max = optimizer.X();
+
+  // Accumulate errors
+  double** xmin = X_min->pointer();
+  double** xmax = X_max->pointer();
+  const double*  xmin_ref = Xmin_ref;
[…]

### 2020-11-19 oepdev/libutil/util.cc
Git revision in preambule removed

@@ -22,8 +22,6 @@ void preambule(void) {
       outfile->Printf("                From Fragment-Based Methods of Quantum Chemistry             \n");
       outfile->Printf("                                  Version 1.0.0                              \n");
       outfile->Printf("                                                                             \n");
-      outfile->Printf("                                Git: Rev {203ae6c}                           \n");
-      outfile->Printf("                                                                             \n");
       outfile->Printf("    -------------------------------------------------------------------------\n");
       outfile->Printf("                                                                             \n");
       outfile->Printf("    Bartosz Błasiak                                                          \n");

### 2020-11-19 include/doxygen/oepdev_manual.h
Final commint before first public release!

@@ -740,7 +740,12 @@ To summarize, follow these steps:
      methods or handling with molecule and wavefunction objects. In particular, GEFP
      contains efficient and fully automatized implementation 
      of the \ref sdensfitincomplalt "extended density fitting of EOPs".
-     
+  5. **Test**. Run the tests:
+     \code{.sh}
+     cd ${EOPDEV_PATH}
+     ctest
+     \endcode
+     If any of the tests do not work, resolve possible compilation issues (including Psi4). 
 
 
 \section padvcodestr EOPDev Code Structure
